### 2023-06-16 SDSS-abstract/SDSS2023-extended-abstract.tex
abstract for SDSS 2023

% File SDSS2020_SampleExtendedAbstract.tex
\documentclass[10pt]{article}\usepackage[]{graphicx}\usepackage[table]{xcolor}
% maxwidth is the original width if it is less than linewidth
% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\usepackage{sdss2020} % Uses Times Roman font (either newtx or times package)
\usepackage{url}
\usepackage{latexsym}
\usepackage{amsmath, amsthm, amsfonts}
\usepackage{algorithm, algorithmic}  
\usepackage{graphicx}
\usepackage[table]{xcolor}
\usepackage{caption}

\usepackage{titlesec}
\setlength{\belowcaptionskip}{-15pt}
\usepackage{parskip}
\setlength{\parskip}{0.2\baselineskip plus 1pt}

\title{Testing statistical charts for accuracy and interpretation}

\author{Anonymous}
% \author{
%  Kiegan Rice \\
%   Affiliation / Address line 1  \\
%   Affiliation / Address line 2  \\
%   {\tt email@domain} \\\And
%   Heike Hofmann \\
%   Department of Statistics  \\
%   Iowa State University  \\
%   {\tt hofmann@iastate.edu} \\\And
%   Nola du Toit \\
%   Affiliation / Address line 1  \\
%   Affiliation / Address line 2  \\
%   {\tt email@domain} \\\And
%   Edward Mulrow \\
%   Affiliation / Address line 1  \\
%   Affiliation / Address line 2  \\
%   {\tt email@domain} \\}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\titlespacing{\subsection}{0pt}{1ex}{0ex}
\titlespacing{\section}{0pt}{1ex}{0ex}

%\titlespacing*{\section} {0pt}{3.5ex plus 1ex minus .2ex}{2.3ex plus .2ex}
%\titlespacing*{\subsection} {0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}

\maketitle
% \begin{abstract}
% The use of visuals is a key component in scientific communication, and decisions about the design of a data visualization should be informed by what design elements best support the audience's ability to perceive and understand the components of the data visualization. We build on the foundations of Cleveland and McGill's work in graphical perception \cite{cleveland}, employing a large, nationally-representative, probability-based panel of survey respondents to test perception in statistical charts. Our findings provide actionable guidance for data visualization practitioners to employ in their work. 
% \end{abstract}

{\bf Keywords:} Graphical Perception, Data Visualization, Scientific Communication

\section{Introduction}

%Why is it important that graphics are good? 
Communication of scientific results to the general public is essential. The use of visuals is a key component in scientific communication; visuals can help provide context, explain scientific concepts, highlight findings, and display patterns in data.

Decisions about the design of a data visualization are made by the author, and may be driven by subject matter conventions, branding choices, or personal style preferences. When used to communicate with the general public, choices about data visualizations should be informed by knowledge on what best supports the audience in understanding the data and conclusions correctly. %We want  choices about data visualizations to be informed by knowledge on what best supports the statement in the data.

%We distinguish between two components of audience understanding: perception and cognition. Being able to accurately read information from a chart (perception) is a crucial first step and the basis to enable the second state -- connecting the individual pieces of information from the chart to a connected whole of insight about the data (cognition).

Cleveland and McGill~\shortcite{cleveland} proposed a list of basic graphical perception tasks and experimentally determined an accuracy-based ranking of those tasks. This ranking has been reproduced by others~\cite{heer}. The tasks proposed by Cleveland and McGill involve assessments of elementary chart elements which focus solely on the \textit{structural} components of a chart: the mapping of a statistical quantity to a particular shape, size, and position in an image. These studies do not address the impact of \textit{aesthetics} on perception: orientation of elements in a chart, context provided, color mappings, and supporting visual elements such as borders, labels, and reference or grid lines.

Cleveland and McGill recruited colleagues and their spouses for their evaluation, while the Heer study used respondents to Amazon Mechanical Turk\footnote{https://www.mturk.com/}.

We extended the tasks proposed by Cleveland and McGill to complete a series of tests on graphical perception in modern data visualization, incorporating and varying both structural and aesthetic elements within the charts presented to viewers. In addition, we utilized a nationally-representative sample of U.S. adults to better understand how the general public as a whole perceives graphics, in contrast to the convenience samples used by prior studies.  
%("We analyzed a total of 2,880 responses from the two experimental runs" HH: not sure how many participants that corresponds to).

%KR: here are the citations
%\cite{cleveland}, \cite{heer}, \cite{lu2021jnd}, 
%\cite{skau}, \cite{eells}, \cite{vonHuhn}, \cite{spence}

%Overview by  Tableau researchers: https://research.tableau.com/sites/default/files/bar.pdf

\section{Methods}

We conducted a series of tests focusing on the public's ability to perceive differences between two values displayed in a chart. Each test asked the respondents to identify which of two elements displayed in a chart was larger. We varied the \textit{structure} of the visual elements as well as the \textit{aesthetics} used in the chart.

\subsection{Test population}

We employ AmeriSpeak's Omnibus\footnote{https://amerispeak.norc.org/us/en/amerispeak/our-capabilities/amerispeak-omnibus.html} survey, which utilizes a probability-based panel and surveys a sample of 1,000 nationally representative adults 18 and older. The advantage of using probability-based panel approach is two-fold. First, we have access to a large sample of survey participants and thus have greater power in making inference about graphical perception abilities. Second, the sample is representative of the general adult public in the U.S., which is an important target audience for scientific communication.

In some rounds, respondents were split into two groups and each group received a different stimulus; each stimulus in every round was seen and responded to by at least 465 respondents. %\textbf{OVER 400 RESPONDENTS -- DO WE HAVE THE ACTUAL MIN HANDY?}. HH: round 1, colors were evaluated by 465.

\subsection{Test stimulus}

\begin{figure*}
\centering
\begin{tabular}{lll}
(a) vertical & (b) horizontal & (c) horizontal wide \\
\includegraphics[width=0.25\textwidth]{images/vertical.PNG} &
\includegraphics[width=0.25\textwidth]{images/horizontal.PNG} &
\includegraphics[width=0.25\textwidth]{images/horizontal-wide.PNG} 
\end{tabular}
\caption{\label{stimulus} From left to right three stacked barcharts shown to panelists. The difference in value between A and B is the same throughout all charts (B is larger). The area of the representation is kept constant; the difference in heights/widths between bars A and B changes from 7 pixels in (a) and (b) to 11 pixels in (c). }
\end{figure*}

The tasks focused on determining which of two very similar values were larger. The two values were chosen close to the \textit{just noticeable difference} (the theoretical difference for which about half of the population is able to determine the difference correctly), making this a visually hard task at the boundary of our perception \cite{lu2021jnd}. Our assumption is that even small changes to this task can have a large effect on a viewer's perception resulting in measurable changes of the overall accuracy.  %Respondents could also select the option that the values were the same. \textbf{KR: WE SHOULD NOTE IN RESULTS SECTION RESULTS ARE CONSISTENT EITHER WAY, OR CUT THIS SENTENCE IN THE ABSTRACT.}

Values were depicted as pieces of a stacked bar chart, shown in Figure~\ref{stimulus}. Responses were collected from a series of variations to structure and aesthetics. Differences in structure included whether the marked pieces were aligned along a common baseline or not, as well as the orientation of the chart as a horizontal or vertical stacked bar chart (see Figure~\ref{stimulus}a,b). Differences in aesthetics included variations in the color scheme used, use of grid lines, and removal of all relevant context (not shown). 
%
\section{Results}
Each one of the stacked barcharts in Figure~\ref{stimulus} was shown to panelists in two versions: with bars A and B aligned along a common axis and the stacked version shown in the image.  The results of these evaluations are shown in Figure~\ref{fig:aligned}. Like Cleveland and McGill, we find that assessing the difference between aligned bars is easier than between unaligned bars (McNemar test statistic 372.15, $p$-value = $2.2 \times 10^{-16}$).

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{images/fig-aligned-round4-1.png}
\caption{\label{fig:aligned} A large number of respondents chose the answer 'the bars are the same size'. While correct for the purposes of interpretation, technically this answer is not the most accurate. Differences between aligned bars are correctly identified at about twice the rate of stacked bars. }
\end{figure}
The design choices are clearly having an impact on accuracy. The change from a horizontal to a vertical design leads to a significant loss in accuracy for both aligned and stacked bars. 
The re-scaled design of the wide horizontal bars reclaims some of the loss for stacked bars and outperforms the vertical design by a similar margin in aligned bars. 
This improves performance of the wide horizontal design over the horizontal design is expected: bars are overall wider and closer together, both factors positively affect accuracy \cite{lu2021jnd}.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{images/zoomers3-1.png}
\caption{\label{zooming}Devices with small screens lead to about twice the zooming rate as tablets or desktops. Panelists zoom into the vertical design, but less so into horizontal designs.}
\end{figure}

Another factor contributing to the difference in accuracy between the vertical and the horizontal is how participants interact with the different designs: generally, about half of all participants make use of the option to zoom into charts (which improves accuracy by about 5 percentage points on average).  Figure~\ref{zooming} shows that zooming behavior of panelists changes depending on the design of the chart.

\section{Conclusions/Further Work}
We have shown that the format of the AmeriSpeak panel allows us to reproduce findings established in the literature.  This provides us with a unique opportunity to create an experimentally validated portfolio of graphical insights:
Grid lines improve accuracy in comparisons, isolating the visual task from its context in the chart seems oddly detrimental to an accurate assessment, forcing panelists to rank objects by size (rather than offering the choice of 'same') does not change the relative rate between the other choices, but increases time spent on an evaluation and reduces certainty in one's response. 
%
\vspace{-10pt}
\bibliographystyle{sdss2020} % Please do not change the bibliography style
\bibliography{references}
\end{document}

### 2023-06-16 SDSS-2023/testing-charts-accuracy.qmd
setup with template from 2022

---
title: "Testing Charts: viewer's perceptual accuracy in surveys"
bibliography: references.bib
abstract: " "
keywords: 'list, keywords, here, 3-6, are, fine'
format: 
  pdf:
    include-in-header:
     # - edits.tex
      - authors.tex
      # - refs.tex
    keep-tex: true
    # template: template.tex
documentclass: jds
---

<!-- Article Types -->
<!-- https://jds-online.org/journal/JDS/information/Article%20Types -->

<!-- Statistical Data Science -->

<!-- This section is the home base of the reformed journal covering statistical methods that are motivated by real-world applications. It is not for papers with technical proofs that push the frontiers of theoretical developments. In addition to classic topics in Statistics, cutting-edge works on big data, visualization, machine learning, and artificial intelligence are also welcome. -->

<!-- Computing in Data Science -->

<!-- Computing is an indispensable component of all data science and big data applications. As more journals on data science and big data emerge, it is of great interest for the data science community to have a highly regarded outlet with a specialization in computing, covering a wide spectrum from methods, algorithms, software implementations, to case studies. Existing journals on statistical computing have their own traditions and may not meet the increasing demands. Some research works on cutting-edge problems may not fit well in any existing journal. -->

<!-- This section covers the following types of articles. -->
<!-- 1. Software: Articles here are similar to those in the Journal of Statistical Software. They are not referencing manuals but vignettes that introduce the methods being implemented as well as the usage of the software with reproducible code chunks. The software implementation can be in any computer language with a sufficiently large user base. -->
<!-- 2. Algorithms: Articles here focus on the performance side of the computing needs arising from domain applications. For example, one can propose algorithms that make infeasible tasks feasible or speed up existing algorithms. -->
<!-- 3. Methods: Articles here are similar to those in Journal of Computational and Graphic Statistics or Statistics and Computing. The computing methods need to be motivated by a domain application with the properties carefully studied. -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, 
                      eval = F, 
                      fig.path="./images/", 
                      fig.env="figure*", 
                      fig.align = "center",
                      dpi = 72) 
library(readr)
library(tidyverse)
library(knitr)
library(bibtex)
theme_set(theme_bw())

library(readxl)
library(lubridate)
library(patchwork)
```

```{r eyefitting-lmer-residualplots, warning = F, message = F}
#| fig-cap: "quarto figure caption."
#| out-width: \columnwidth
#| label: fig-label-we-decide-on
#| fig-res: 300
#| fig-width: 9 # set only if different from the default
#| fig-height: 9
#| fig-env: figure
#|

data.frame(x = runif(100), y = runif(100)) %>%
  ggplot(aes(x = x, y = y)) +
  geom_point()
```

# Introduction

# Development

# Application

## Validation Study

## Data Analysis and Results

# Conclusion

# Supplementary Material {-}

+ **Participant Data (Linear):** Link to csv file with the data. 
+ **Data Analysis Code:** Link to an html document with annotated code chunks.

# References

### 2023-06-16 SDSS-2023/testing-charts-accuracy.qmd
set up for edits

@@
     include-in-header:
-     # - edits.tex
+      - edits.tex
       - authors.tex
@@
 
+\hh{This is Heike's color for making changes}
 # Introduction

### 2023-06-16 SDSS-2023/testing-charts-accuracy.qmd
hopefully got all of the files we want to ignore in git

@@
                       fig.align = "center",
-                      dpi = 72) 
+                      dpi = 72,
+                      cache = TRUE) 
 library(readr)

### 2023-06-16 SDSS-2023/testing-charts-accuracy.qmd
settings for figures and images

--- 2023-06-16 SDSS-2023/testing-charts-accuracy.qmd
@@
 knitr::opts_chunk$set(echo = F, 
-                      eval = F, 
-                      fig.path="./images/", 
+                      eval = T, 
+                      fig.path="./figures/", # figures are created on the fly, images are pictures like the stimuli
                       fig.env="figure*", 
                       fig.align = "center",
-                      dpi = 72) 
+                      dpi = 72, # we can set the resolution up to 300 later.
+                      cache = TRUE) 
 library(readr)
@@
 
-```{r eyefitting-lmer-residualplots, warning = F, message = F}
+```{r showplot, warning = F, message = F}
 #| fig-cap: "quarto figure caption."

### 2023-06-17 SDSS-2023/testing-charts-accuracy.qmd
authors

@@
 title: "Testing Charts: viewer's perceptual accuracy in surveys"
-bibliography: references.bib
-abstract: " "
+bibliography: "`r rbbt::bbt_write_bib('references.bib', overwrite = TRUE, library_id=rbbt::bbt_library_id('Graphics Research'))`"
+abstract: "The use of visuals is a key component in scientific communication, and decisions about the design of a data visualization should be informed by what design elements best support the audience's ability to perceive and understand the components of the data visualization.
+We build on the foundations of Cleveland and McGill's work in graphical perception, employing a large, nationally-representative, probability-based panel of survey respondents to test perception in statistical charts.
+Our findings provide actionable guidance for data visualization practitioners to employ in their work."
 keywords: 'list, keywords, here, 3-6, are, fine'
@@
 
+@clevelandGraphicalPerceptionTheory1984
+
 ```{r setup, include = FALSE}
@@
                       fig.align = "center",
+                      out.width = "50%",
                       dpi = 72, # we can set the resolution up to 300 later.
@@
 #| fig-res: 300
-#| fig-width: 9 # set only if different from the default
-#| fig-height: 9
+#| fig-width: 5 # set only if different from the default
+#| fig-height: 5
 #| fig-env: figure
@@
 \hh{This is Heike's color for making changes}
-# Introduction
+## Introduction
 
-# Development
+How do structural design choices in a data visualization impact viewers’ ability to identify the larger of two elements?
 
-# Application
+How do aesthetic design choices in a data visualization impact viewers’ ability to identify the larger of two elements?
 
-## Validation Study
+How is viewer behavior (zooming, time spent on question, certainty of response) impacted by structural and aesthetic design choices in a data visualization?
+
+Structural design choices:
+
+Structural design choices
+
+Mapping, Stacked bar, Vertical, Horizontal, Horizontal wide
+
+Facetted bar
+
+Only have a split sample for this
+
+Pie, Alignment
+
+We have this for all above mappings, but the setup is a little different for facetted bar
+
+Aesthetic design choices (structural choices seems stronger/there could be a lot to talk about there… should we skip aesthetic on this one?)
+
+Colors
+
+Use of gridlines
+
+Outcomes/responses for modeling:
+
+Binary accuracy (correct/incorrect – ‘they are the same’ is incorrect here)
+
+Ordinal response (a/b/they are the same)
+
+Zooming behavior (zoomed/did not zoom)
+
+Time spent on question (continuous, in seconds)
+
+Certainty
+
+I’ve noted this below, but: how do model? Ordinal response? Binary (certain or very certain vs everybody else)?
+
+## Survey setup - Stimulus description
+
+We decided to show respondents a stacked bar chart (because there is little research on stacked bars) and ask to compare the sizes of two tiles (labelled A vs B or C vs D) to each other.
+
+DATA THAT MAY BE INCLUDED IN ANALYSIS:
+
+ROUNDS 1-2: Color variations on vertical stacked bar, aligned and unaligned
+
+ROUND 3: Horizontal and horizontal wide, aligned and unaligned
+
+ROUND 5: Horizontal wide gridlines (only dark grid split sample)
+
+ROUND 6: Facetted bar (split sample w/o forcing choice)
+
+ROUND 7: Aligned vs unaligned pie (full sample)
 
 ## Data Analysis and Results
+
+COMBINING SAMPLES AND WEIGHTING NOTES:
+
+We can combine responses across samples into one combined dataset, but we need to adjust weights accordingly so that each sample is weighted equally in the model
+
+Question for Ed: If we compare a full sample to a split sample, do we still want to weight these ‘equally’?
+
+Analysis should be done using the ‘survey’ package and weights should be taken into account.
+
+ANALYSIS PLAN:
+
+Models below structured as:
+
+Response
+
+Covariates to use in each model
+
+STRUCTURAL VARIATION – START HERE
+
+Binary accuracy across structural choices
+
+Model 1:
+
+Alignment only, just vertical stacked bar
+
+Model 2:
+
+Alignment
+
+Bar vs pie (comparable question for pie is A vs B)
+
+Model 3:
+
+Alignment
+
+Vertical x horizontal x horizontal wide
+
+Model 4:
+
+Alignment
+
+Every structure (vertical bar, horizontal bar, horizontal wide bar, facet bar, pie)
+
+Visuals:
+
+% yes across each different structural condition
+
+Facet by aligned/unaligned?
+
+Model estimates + CIs
+
+Ordinal response
+
+Model 1:
+
+Alignment only, just vertical stacked bar
+
+Model 2:
+
+Alignment
+
+Bar vs pie (comparable question for pie is A vs B)
+
+Model 3:
+
+Alignment
+
+Vertical x horizontal x horizontal wide
+
+Model 4:
+
+Alignment
+
+Every structure (vertical bar, horizontal bar, horizontal wide bar, facet bar, pie)
+
+Visuals:
+
+All responses across each different structural condition
+
+Facet by aligned/unaligned?
+
+Model estimates + CIs
+
+Zooming behavior (zoomed/did not zoom)
+
+Model 1:
+
+Device type
+
+Alignment
+
+Vertical x horizontal x horizontal wide
+
+Visuals:
+
+% zoomed by device + alignment (already have this chart)
+
+Model estimates + CIs
+
+Time spent on question (in seconds)
+
+Model 1:
+
+Device type
+
+Zoom
+
+Alignment
+
+Vertical x horizontal x horizontal wide
+
+Model 2 (this may not be feasible for comparison depending on what level the ‘TOTALTIME’ is captured at):
+
+Device type
+
+Zoom
+
+Alignment
+
+Every structure (vertical bar, horizontal bar, horizontal wide bar, facet bar, pie)
+
+Visuals:
+
+Distribution of time spent variable
+
+Facet by device type, zoom, structural condition, alignment? Play around with it
+
+Average time spent by each of the conditions
+
+Certainty?
+
+Same models as above, but I’m not sure how we want to do the response. Ordinal response? Binary (certain or very certain vs everybody else)?
+
+AESTHETIC VARIATION – ONLY IF TIME
+
+Binary accuracy (correct/incorrect – ‘they are the same’ is incorrect here) across structural choices
+
+Model 1:
+
+Dark grid vs no grid (only have for horizontal wide)
+
+Response choice (ordinal response)
+
+Model 1:
+
+Dark grid vs no grid (only have for horizontal wide)
+
+Zooming behavior (zoomed/did not zoom)
+
+Model 1:
+
+Device type
+
+Dark grid vs no grid
+
+Time spent on question (in seconds)
+
+Model 1:
+
+Device type
+
+Zoom
+
+Dark grid vs no grid
+
+Certainty?
+
+Same models as above, but I’m not sure how we want to do the response. Ordinal response? Binary (certain or very certain vs everybody else)?
 

### 2023-06-18 SDSS-2023/testing-charts-accuracy.qmd
rounds 1 and 2 description, and motivation of JND

@@
 
-I’ve noted this below, but: how do model? Ordinal response? Binary (certain or very certain vs everybody else)?
+I’ve noted this below, but: how to model? Ordinal response? Binary (certain or very certain vs everybody else)?
 
@@
 
-We decided to show respondents a stacked bar chart (because there is little research on stacked bars) and ask to compare the sizes of two tiles (labelled A vs B or C vs D) to each other.
+Asking perceptual questions in a survey is different from the controlled environment of a cognitive lab, where these kind of questions would usually be addressed.
+This means, that instructions to participants have to be delivered in a very short and easily understandable, because questions arising from the task can not be answered.
+Similarly, rather than asking the same (or similar) type of question multiple hundred times, in a survey we can ask only a few questions.
+In order to observe any effect, we need to ask questions that are perceptually hard, which means that we need to ask questions about stimuli that are close to our perceptual threshold.
+The **Just-Noticeable Difference** (JND) is defined as the smallest difference that will be detected 50% of the time.
+For the height difference of barcharts, @luModelingJustNoticeable2022 determined that the distance between bars is the main contributor to differentiating sizes.
+Based on their models, a tile difference of 155 pixels leads to a just noticeable difference of 3.5 pixels.
+\autoref{tbl-tasks-12} shows the two stacked barcharts shown to participants in the survey.
+The marked tiles in each plot are 155 pixels apart.
+Their heights are 205 (left) and 213 pixels (right), respectively, corresponding to about twice the JND.
+This difference should lead to a relatively high accuracy rate for participants and simultaneously limit the amount of frustration resulting from a task that is perceived as 'too hard'.
+
+<!-- log_{10}(JND) = -.4653 + .0065 distance -->
+
+Both charts in \autoref{tbl-tasks-12} show the same data with slight modifications to the order of the levels – the first and second level in each of the bars are reversed between the left and the right chart.
+Participants were asked to compare the relative sizes of the tiles marked A and B (C and D, respectively) and select the correct response out of the possible choices:
+
+1. A is bigger
+2. B is bigger
+3. They are the same
+
+Answer 2 is the correct answer for both charts.
+Both charts are shown at the same size, i.e.
+in both cases the difference in size between the bars is exactly the same, the vertical distance between the bars is the exact same amount.
+This leaves the vertical positioning of the bars as the only difference between the charts. Any differences in observed accuracy can therefore be attributed to this difference in presentation.
+
+|                                   |                                   |
+|-----------------------------------|-----------------------------------|
+| **Unaligned positions**           | **Aligned positions**             |
+| ![Unaligned positions marked A and B](../SDSS-2023/images/vertical-ab.PNG) | ![Aligned positions marked C and D](../SDSS-2023/images/vertical-cd.PNG) |
+
+: The two stacked barcharts every participant got to see.
+In each barchart, the two marked tiles are to be compared for their size.
+in both instances, the tile on the right is (very slightly) larger.
+{#tbl-tasks-12 tbl-colwidths="\[50,50\]"}
 
@@
 
-We can combine responses across samples into one combined dataset, but we need to adjust weights accordingly so that each sample is weighted equally in the model
+\hh{We can combine responses across samples into one combined dataset, but we need to adjust weights accordingly so that each sample is weighted equally in the model [@omuircheartaighCombiningSamplesVs2002].}
 
@@
 Alignment only, just vertical stacked bar
+
+\hh{Using the AmeriSpeak survey tool, a total of 1902 participants were exposed to two barcharts each, as shown in \autoref{tbl-tasks-12}.}
 

### 2023-06-18 SDSS-2023/testing-charts-accuracy.qmd
stage edits

@@
 
+<!-- motivate the use of just noticeable differences -->
 Asking perceptual questions in a survey is different from the controlled environment of a cognitive lab, where these kind of questions would usually be addressed.
@@
 
-<!-- log_{10}(JND) = -.4653 + .0065 distance -->
+<!-- Lu et al calculation: log_{10}(JND) = -.4653 + .0065 distance -->
 
+<!-- describe stimulus -->
 Both charts in \autoref{tbl-tasks-12} show the same data with slight modifications to the order of the levels – the first and second level in each of the bars are reversed between the left and the right chart.

### 2023-06-19 SDSS-2023/testing-charts-accuracy.qmd
references

@@
 title: "Testing Charts: viewer's perceptual accuracy in surveys"
-bibliography: "`r rbbt::bbt_write_bib('references.bib', overwrite = TRUE, library_id=rbbt::bbt_library_id('Graphics Research'))`"
+bibliography: "`r rbbt::bbt_write_bib('references.bib', overwrite = TRUE, library_id=rbbt::bbt_library_id('Graphics Research'), translator='bibtex')`"
 abstract: "The use of visuals is a key component in scientific communication, and decisions about the design of a data visualization should be informed by what design elements best support the audience's ability to perceive and understand the components of the data visualization.
@@
 library(patchwork)
+library(survey)
+```
+
+```{r data}
+# other github repo. 
+round1 <- readRDS("~/Documents/projects/NORC/visual-studies/data/round1.rds")
+round2 <- readRDS("~/Documents/projects/NORC/visual-studies/data/round2.rds")
+
+survey1 <- survey::svydesign(
+  data = round1 %>% 
+            filter(ab != "SKIPPED ON WEB", cd != "SKIPPED ON WEB"), 
+  ids = round1$CaseId, weights = round1$weight)
+
+survey2 <- survey::svydesign(
+  data = round2 %>% 
+            filter(ab != "SKIPPED ON WEB", cd != "SKIPPED ON WEB"), 
+  ids = round2$CaseId, weights = round2$weight)
+
 ```
@@
 Analysis should be done using the ‘survey’ package and weights should be taken into account.
+
+```{r}
+round_n1 <- nrow(round1)
+round_n2 <- nrow(round2)
+# effective sample size n/D, where D is the design effect
+d1 <- deff(survey1)
+```
+All calculations in this paper are done in R [@R2022] using the `survey` package [@lumleyAnalysisComplexSurvey2004] version 4.0 [@survey] based on @lumleyComplexSurveysGuide2010.
 

### 2023-06-19 analysis/analysis.qmd
analysis

---
title: "Supplement: Analysis and code for 'Testing Perceptual Accuracy in Charts using Surveys'"
format:
  jasa-html: default
  jasa-pdf:
    keep-tex: true  
    journal:
      blinded: false
date: last-modified
bibliography: ../SDSS\-2023/references.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, 
                      eval = T, 
                      fig.path="./figures/", # figures are created on the fly, images are pictures like the stimuli
                      fig.env="figure*", 
                      fig.align = "center",
                      out.width = "50%",
                      dpi = 72, # we can set the resolution up to 300 later.
                      cache = TRUE) 
library(readr)
library(tidyverse)
library(knitr)
library(bibtex)
theme_set(theme_bw())

library(readxl)
library(lubridate)
library(patchwork)
library(survey)
#library(ggmosaic)
library(ggpcp)

```

## Survey rounds

```{r data}
# other github repo. 
round1 <- readRDS("~/Documents/projects/NORC/visual-studies/data/round1.rds")
round2 <- readRDS("~/Documents/projects/NORC/visual-studies/data/round2.rds")

round1_adj <- round1 %>% 
            filter(ab != "SKIPPED ON WEB", cd != "SKIPPED ON WEB") %>%
            mutate(CorrectAB = ab=="B is bigger", CorrectCD = cd=="D is bigger",
                   CorrectABCD = paste(CorrectAB, CorrectCD),
                   one = 1)

round2_adj <- round2 %>% 
            filter(ab != "SKIPPED ON WEB", cd != "SKIPPED ON WEB") %>%
            mutate(CorrectAB = ab=="B is bigger", CorrectCD = cd=="D is bigger",
                   CorrectABCD = paste(CorrectAB, CorrectCD),
                   one = 1)

survey1 <- survey::svydesign(
  data = round1_adj, 
  ids = ~CaseId, weights = ~weight)

survey2 <- survey::svydesign(
  data = round2_adj,
  ids = ~CaseId, weights = ~weight)
```

The data for this paper were collected in several rounds as part of the NORC Omnibus.

| Name    | Date  | # Participants | Weights $\sum_i w_i$ |
|---------|-------|--------:|----------:|
| Round 1 | April 2022 |  942    |    `r round(sum(round1_adj$weight),1)`       |
| Round 2 | May 2022 |  960   |    `r round(sum(round2_adj$weight),1)`       |
|         |       |         |           |

: Survey rounds: dates, number of participants (nominal sample size), and sum of weights. {#tbl-rounds}

@omuircheartaighCombiningSamplesVs2002 suggest combining surveys $S_1$ and $S_2$ by 
multiplying weights in $S_1$ and $S_2$ by $\lambda$ and $1 - \lambda$, respectively.
$$
\lambda = \frac{n_1/d_1}{n_1/d_1 + n_2/d_2},
$$
where $n_1$ and $n_2$ are the nominal sample sizes and $d_1$ and $d_2$ are the design effects for the estimators. 
Instead of using design effects itself, $d_1$ and $d_2$ are estimated as 
$$
d_1 = 1 + CV(w_i \in S_1)^2 \ \ \ \text{ and } \ \ \ d_2 = 1 + CV(w_i \in S_2)^2
$$
$CV$ is the coefficient of variation of the weights within each sample.

@omuircheartaighCombiningSamplesVs2002 calculated
$\lambda$ separately for any combination of race/ethnicity by sex.

We will calculate $\lambda$ separately whenever we include demographic variables in the analysis, otherwise we will use a single adjustment for the weights.

All calculations are done in R [@R2022] using the `survey` package [@lumleyAnalysisComplexSurvey2004] version 4.0 [@survey] based on @lumleyComplexSurveysGuide2010.

## Model 1: Aligned versus Unaligned {#sec-model1}

```{r lambdas}
n1 <- nrow(round1_adj)
n2 <- nrow(round2_adj)
neff1 <- svytotal(~one, survey1, deff=TRUE)
neff2 <- svytotal(~one, survey2, deff=TRUE)
#abcd1 <- svytotal(~CorrectABCD, survey1, deff=TRUE)
#abcd2 <- svytotal(~CorrectABCD, survey2, deff=TRUE)

d1 <- 1 + cv(neff1)^2
d2 <- 1 + cv(neff2)^2

denom <- n1/d1 + n2/d2

lambda <- n1/(d1*denom) 
```

```{r combine_rounds_1_and_2}
round_12 <- round1_adj %>% 
            mutate(weight = lambda[1]*weight) %>% 
            select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight) %>%
    rbind(round2_adj %>% 
            mutate(weight = (1-lambda[1])*weight) %>% 
            select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight))
round_12_long <- round_12 %>% 
  pivot_longer(CorrectAB:CorrectCD, names_to = "Question", values_to = "Response")

survey_12 <- 
  survey::svydesign(
  data = round_12_long, 
  ids = ~CaseId, weights = ~weight)
```

```{r ttest}
paired <- svyttest(Response~Question, survey_12)
abcd <- svyby(~Response, by=~Question, survey_12, svytotal)

marginal_results <- data.frame(abcd)
```

Accuracy comparing aligned versus unaligned bars. The data used for this is a combination of rounds 1 and 2, with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(sum(round_12_long$weight/2),1)`. \autoref{fig-alignment}(a) shows that more than twice the number of responses is accurate, when the tiles are aligned along the same axis. 
Because each particpant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts. The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: < 2.2e-16).

```{r, fig.height = 5, fig.width = 10, warning=FALSE}
#| fig-cap: "On the left (a), a stacked barchart shows the number of respondents with correct (green) and incorrect (grey) responses to the two comparison questions. When tiles are aligned along the same axis, more than twice the number of responses is accurate. The shaded area along the top of the green tiles corresponds to  95\\% confidence intervals around (marginal) correct responses. On the right (b), a parallel coordinate plot shows all combinations of responses. There's a huge asymmetry in the number of responses, where participants answered only one of the questions correctly. A lot more responses are correct when comparing aligned tiles than unaligned tiles."
#| label: fig-alignment
#| out-width: 80%
#| fig-env: figure

gg1 <- round_12_long %>%
  mutate(
    Response = ifelse(Response, "Correct", "Not correct"),
    Response = factor(Response, levels=c("Not correct", "Correct")),
    Question = ifelse(Question=="CorrectAB",
                      "Unaligned: A or B?", "Aligned: C or D?")
  ) %>%
#  filter(Response == "Correct") %>%
  ggplot(aes(x = Question, fill = Response, weight = weight)) +
  geom_bar() +
  ylab("Number of (effective) Respondents") +
  xlab("") +
  scale_fill_manual(values=c("grey", "forestgreen")) +
  scale_y_continuous(
      sec.axis = sec_axis(~./sum(round_12$weight)*100,
                          name="Percent responses")
  ) +
  theme(legend.position = "bottom") +
  ggtitle("(a) Accuracy of survey responses")  +
  geom_rect(
    aes(
      x = rev(c("Aligned: C or D?", "Unaligned: A or B?")),
      xmin = 2:1 - 0.45,
      ymin = ResponseTRUE - 1.96* se.ResponseTRUE,
      xmax = 2:1 + 0.45,
      ymax = ResponseTRUE + 1.96* se.ResponseTRUE,
      weight = 1),
    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
    data = marginal_results
  )

gg2 <- round_12 %>%
  mutate(
    `Unaligned: A or B?` = ifelse(CorrectAB, "Correct", "Not correct"), 
    `Aligned: C or D?` = ifelse(CorrectCD, "Correct", "Not correct")
    ) %>%
  pcp_select(`Aligned: C or D?`, `Unaligned: A or B?`) %>%
  pcp_scale() %>%
  pcp_arrange() %>%
  ggplot(aes_pcp()) + 
    geom_pcp_axes() + 
    geom_pcp(aes(colour = CorrectABCD), alpha = 0.2)+
    geom_pcp_boxes(fill=NA, colour = "black", linewidth=0.25) + 
    geom_pcp_labels(colour = "black") +
  scale_y_continuous("Number responses", breaks=seq(0,1,by=0.25), 
                     labels = round(sum(round_12$weight)*seq(0,1,by=0.25),0),
                     sec.axis = sec_axis(~.*100,
                          name="Percent responses")) +
  xlab("") + 
  scale_colour_manual(values = c("grey", "#84AC84", "#70A570", "forestgreen")) +
  ggtitle("(b) Combinations of responses") +
  theme(legend.position = "none")

gg1 + gg2
```

## References{-}

### 2023-06-20 SDSS-2023/testing-charts-accuracy.qmd
analysis step .0001

@@
 library(survey)
+library(ggpcp)
 ```
@@
 
+round1_adj <- round1 %>% 
+            filter(ab != "SKIPPED ON WEB", cd != "SKIPPED ON WEB") %>%
+            mutate(CorrectAB = ab=="B is bigger", CorrectCD = cd=="D is bigger",
+                   CorrectABCD = paste(CorrectAB, CorrectCD),
+                   one = 1)
+
+round2_adj <- round2 %>% 
+            filter(ab != "SKIPPED ON WEB", cd != "SKIPPED ON WEB") %>%
+            mutate(CorrectAB = ab=="B is bigger", CorrectCD = cd=="D is bigger",
+                   CorrectABCD = paste(CorrectAB, CorrectCD),
+                   one = 1)
+
 survey1 <- survey::svydesign(
-  data = round1 %>% 
-            filter(ab != "SKIPPED ON WEB", cd != "SKIPPED ON WEB"), 
-  ids = round1$CaseId, weights = round1$weight)
+  data = round1_adj, 
+  ids = ~CaseId, weights = ~weight)
 
 survey2 <- survey::svydesign(
-  data = round2 %>% 
-            filter(ab != "SKIPPED ON WEB", cd != "SKIPPED ON WEB"), 
-  ids = round2$CaseId, weights = round2$weight)
-
-```
-
-```{r showplot, warning = F, message = F}
-#| fig-cap: "quarto figure caption."
-#| out-width: \columnwidth
-#| label: fig-label-we-decide-on
-#| fig-res: 300
-#| fig-width: 5 # set only if different from the default
-#| fig-height: 5
-#| fig-env: figure
-#|
-
-data.frame(x = runif(100), y = runif(100)) %>%
-  ggplot(aes(x = x, y = y)) +
-  geom_point()
+  data = round2_adj,
+  ids = ~CaseId, weights = ~weight)
 ```
@@
 
-```{r}
-round_n1 <- nrow(round1)
-round_n2 <- nrow(round2)
-# effective sample size n/D, where D is the design effect
-d1 <- deff(survey1)
+All calculations in this paper are done in R [@R2022] using the `survey` package [@lumleyAnalysisComplexSurvey2004] version 4.0 [@survey] based on @lumleyComplexSurveysGuide2010.
+
+```{r lambdas}
+n1 <- nrow(round1_adj)
+n2 <- nrow(round2_adj)
+neff1 <- svytotal(~one, survey1, deff=TRUE)
+neff2 <- svytotal(~one, survey2, deff=TRUE)
+#abcd1 <- svytotal(~CorrectABCD, survey1, deff=TRUE)
+#abcd2 <- svytotal(~CorrectABCD, survey2, deff=TRUE)
+
+d1 <- 1 + cv(neff1)^2
+d2 <- 1 + cv(neff2)^2
+
+denom <- n1/d1 + n2/d2
+
+lambda <- n1/(d1*denom) 
 ```
-All calculations in this paper are done in R [@R2022] using the `survey` package [@lumleyAnalysisComplexSurvey2004] version 4.0 [@survey] based on @lumleyComplexSurveysGuide2010.
+
+```{r combine_rounds_1_and_2}
+round_12 <- round1_adj %>% 
+            mutate(weight = lambda[1]*weight) %>% 
+            select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight) %>%
+    rbind(round2_adj %>% 
+            mutate(weight = (1-lambda[1])*weight) %>% 
+            select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight))
+round_12_long <- round_12 %>% 
+  pivot_longer(CorrectAB:CorrectCD, names_to = "Question", values_to = "Response")
+
+survey_12 <- 
+  survey::svydesign(
+  data = round_12_long, 
+  ids = ~CaseId, weights = ~weight)
+```
+
+```{r ttest}
+paired <- svyttest(Response~Question, survey_12)
+abcd <- svyby(~Response, by=~Question, survey_12, svytotal)
+
+marginal_results <- data.frame(abcd)
+```
+
+The data used for assessing the accuracy of comparisons in \autoref{tbl-tasks-12} is collected in two rounds of the NORC Omnibus survey.
+Rounds 1 and 2 are combined by adjusting the weights  with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(sum(round_12_long$weight/2),1)`.
+Figure \ref{fig-alignment}(a) shows that more than twice the number of responses is accurate, when the tiles are aligned along the same axis.
+Because each particpant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts.
+The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: < 2.2e-16).
+
+```{r accuracy, fig.height = 5, fig.width = 10, warning=FALSE}
+#| fig-cap: "On the left (a), a stacked barchart shows the number of respondents with correct (green) and incorrect (grey) responses to the two comparison questions.
+When tiles are aligned along the same axis, more than twice the number of responses is accurate.
+The shaded area along the top of the green tiles corresponds to  95\\% confidence intervals around (marginal) correct responses.
+On the right (b), a parallel coordinate plot shows all combinations of responses.
+There's a huge asymmetry in the number of responses, where participants answered only one of the questions correctly.
+A lot more responses are correct when comparing aligned tiles than unaligned tiles."
+#| label: fig-alignment
+#| out-width: 80%
+#| fig-env: figure
+
+gg1 <- round_12_long %>%
+  mutate(
+    Response = ifelse(Response, "Correct", "Not correct"),
+    Response = factor(Response, levels=c("Not correct", "Correct")),
+    Question = ifelse(Question=="CorrectAB",
+                      "Unaligned: A or B?", "Aligned: C or D?")
+  ) %>%
+#  filter(Response == "Correct") %>%
+  ggplot(aes(x = Question, fill = Response, weight = weight)) +
+  geom_bar() +
+  ylab("Number of (effective) Respondents") +
+  xlab("") +
+  scale_fill_manual(values=c("grey", "forestgreen")) +
+  scale_y_continuous(
+      sec.axis = sec_axis(~./sum(round_12$weight)*100,
+                          name="Percent responses")
+  ) +
+  theme(legend.position = "bottom") +
+  ggtitle("(a) Accuracy of survey responses")  +
+  geom_rect(
+    aes(
+      x = rev(c("Aligned: C or D?", "Unaligned: A or B?")),
+      xmin = 2:1 - 0.45,
+      ymin = ResponseTRUE - 1.96* se.ResponseTRUE,
+      xmax = 2:1 + 0.45,
+      ymax = ResponseTRUE + 1.96* se.ResponseTRUE,
+      weight = 1),
+    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+    data = marginal_results
+  )
+
+gg2 <- round_12 %>%
+  mutate(
+    `Unaligned: A or B?` = ifelse(CorrectAB, "Correct", "Not correct"), 
+    `Aligned: C or D?` = ifelse(CorrectCD, "Correct", "Not correct")
+    ) %>%
+  pcp_select(`Aligned: C or D?`, `Unaligned: A or B?`) %>%
+  pcp_scale() %>%
+  pcp_arrange() %>%
+  ggplot(aes_pcp()) + 
+    geom_pcp_axes() + 
+    geom_pcp(aes(colour = CorrectABCD), alpha = 0.2)+
+    geom_pcp_boxes(fill=NA, colour = "black", linewidth=0.25) + 
+    geom_pcp_labels(colour = "black") +
+  scale_y_continuous("Number responses", breaks=seq(0,1,by=0.25), 
+                     labels = round(sum(round_12$weight)*seq(0,1,by=0.25),0),
+                     sec.axis = sec_axis(~.*100,
+                          name="Percent responses")) +
+  xlab("") + 
+  scale_colour_manual(values = c("grey", "#84AC84", "#70A570", "forestgreen")) +
+  ggtitle("(b) Combinations of responses") +
+  theme(legend.position = "none")
+
+gg1 + gg2
+```
 

### 2023-06-20 analysis/analysis.qmd
analysis step .0001

@@
 |---------|-------|--------:|----------:|
-| Round 1 | April 2022 |  942    |    `r round(sum(round1_adj$weight),1)`       |
-| Round 2 | May 2022 |  960   |    `r round(sum(round2_adj$weight),1)`       |
+| Round 1 | April 2022 |  `r nrow(round1_adj)`    |    `r round(sum(round1_adj$weight),1)`       |
+| Round 2 | May 2022 |  `r nrow(round2_adj)`   |    `r round(sum(round2_adj$weight),1)`       |
 |         |       |         |           |
@@
 The data used for this is a combination of rounds 1 and 2, with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(sum(round_12_long$weight/2),1)`.
-\autoref{fig-alignment}(a) shows that more than twice the number of responses is accurate, when the tiles are aligned along the same axis.
+Figure \ref{fig-alignment}(a) shows that more than twice the number of responses is accurate, when the tiles are aligned along the same axis.
 Because each particpant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts.
@@
 
-```{r, fig.height = 5, fig.width = 10, warning=FALSE}
+```{r accuracy, fig.height = 5, fig.width = 10, warning=FALSE}
 #| fig-cap: "On the left (a), a stacked barchart shows the number of respondents with correct (green) and incorrect (grey) responses to the two comparison questions.

### 2023-06-20 SDSS-2023/testing-charts-accuracy.qmd
ordinal response is just an additional logistic regression

@@
 library(ggpcp)
+
+# Kish's Effective Sample Size
+neff <- function(weight) {
+  n <- length(weight)
+  L <- var(weight)/mean(weight)^2
+  n/(1+L)
+}
 ```
@@
 The data used for assessing the accuracy of comparisons in \autoref{tbl-tasks-12} is collected in two rounds of the NORC Omnibus survey.
-Rounds 1 and 2 are combined by adjusting the weights  with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(sum(round_12_long$weight/2),1)`.
+Rounds 1 and 2 are combined by adjusting the weights  with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`.
 Figure \ref{fig-alignment}(a) shows that more than twice the number of responses is accurate, when the tiles are aligned along the same axis.

### 2023-06-20 analysis/analysis.qmd
ordinal response is just an additional logistic regression

@@
 
+# Kish's Effective Sample Size
+neff <- function(weight) {
+  n <- length(weight)
+  L <- var(weight)/mean(weight)^2
+  n/(1+L)
+}
 ```
@@
             mutate(CorrectAB = ab=="B is bigger", CorrectCD = cd=="D is bigger",
-                   CorrectABCD = paste(CorrectAB, CorrectCD),
-                   one = 1)
+                   CorrectABCD = paste(CorrectAB, CorrectCD))
 
@@
 
-| Name    | Date  | # Participants | Weights $\sum_i w_i$ |
-|---------|-------|--------:|----------:|
-| Round 1 | April 2022 |  `r nrow(round1_adj)`    |    `r round(sum(round1_adj$weight),1)`       |
-| Round 2 | May 2022 |  `r nrow(round2_adj)`   |    `r round(sum(round2_adj$weight),1)`       |
-|         |       |         |           |
+| Name    | Date  | # Participants | effective sample size | Sum of weights $\sum_i w_i$ |
+|---------|-------|--------:|----------:|----------:|
+| Round 1 | April 2022 |  `r nrow(round1_adj)`    |    `r round(neff(round1_adj$weight),1)`       | `r round(sum(round1_adj$weight),1)`       |
+| Round 2 | May 2022 |  `r nrow(round2_adj)`   |    `r round(neff(round2_adj$weight),1)`       | `r round(sum(round2_adj$weight),1)`       |
+|         |       |         |           |           |
 
-: Survey rounds: dates, number of participants (nominal sample size), and sum of weights. {#tbl-rounds}
+: Survey rounds: dates, number of participants (nominal sample size), effective sample size, and sum of weights. {#tbl-rounds}
 
-@omuircheartaighCombiningSamplesVs2002 suggest combining surveys $S_1$ and $S_2$ by 
+We are using a strategy of *combining* (rather than cumulating) surveys $S_1$ and $S_2$, as described in @omuircheartaighCombiningSamplesVs2002,  by 
 multiplying weights in $S_1$ and $S_2$ by $\lambda$ and $1 - \lambda$, respectively.
@@
 $$
-$CV$ is the coefficient of variation of the weights within each sample.
+$CV$ is the coefficient of variation of the weights within each sample, and can be estimated as
+
+$$ 
+CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
+$$
 
@@
 
-## Model 1: Aligned versus Unaligned {#sec-model1}
+## Model 1: Comparing Aligned and Unaligned Tiles {#sec-model1}
 
 ```{r lambdas}
+
+d1 <- 1+var(round1_adj$weight)/mean(round1_adj$weight)^2
+d2 <- 1+var(round2_adj$weight)/mean(round2_adj$weight)^2
+
 n1 <- nrow(round1_adj)
 n2 <- nrow(round2_adj)
-neff1 <- svytotal(~one, survey1, deff=TRUE)
-neff2 <- svytotal(~one, survey2, deff=TRUE)
-#abcd1 <- svytotal(~CorrectABCD, survey1, deff=TRUE)
-#abcd2 <- svytotal(~CorrectABCD, survey2, deff=TRUE)
 
-d1 <- 1 + cv(neff1)^2
-d2 <- 1 + cv(neff2)^2
+neff1 <- n1/d1
+neff2 <- n2/d2
 
@@
 round_12 <- round1_adj %>% 
-            mutate(weight = lambda[1]*weight) %>% 
+            mutate(weight = lambda*weight) %>% 
             select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight) %>%
     rbind(round2_adj %>% 
-            mutate(weight = (1-lambda[1])*weight) %>% 
+            mutate(weight = (1-lambda)*weight) %>% 
             select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight))
+
+wt_scalar <- neff(round_12$weight)/sum(round_12$weight)
+round_12$weight <- wt_scalar*round_12$weight 
+# should not affect the further analysis, but helps with the interpretation of results
+
 round_12_long <- round_12 %>% 
@@
 
-Accuracy comparing aligned versus unaligned bars.
-The data used for this is a combination of rounds 1 and 2, with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(sum(round_12_long$weight/2),1)`.
-Figure \ref{fig-alignment}(a) shows that more than twice the number of responses is accurate, when the tiles are aligned along the same axis.
+The data used for this is a combination of rounds 1 and 2, with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`.
+
+### Binary response
+
+Defining an accurate response as "B is bigger" in the chart of unaligned tiles, and "D is bigger" in the aligned case, while calling the other two responses as incorrect, leads to the data shown in figure \ref{fig-alignment}(a): we see
+ that more than twice the number of responses is accurate, when the tiles are aligned along the same axis. 
 Because each particpant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts.
@@
   scale_y_continuous("Number responses", breaks=seq(0,1,by=0.25), 
-                     labels = round(sum(round_12$weight)*seq(0,1,by=0.25),0),
+                     labels = round(sum(round_12$weight*wt_scalar)*seq(0,1,by=0.25),0),
                      sec.axis = sec_axis(~.*100,
@@
 
+### Ordinal response
+
+```{r}
+round_12_ord_long <- round_12 %>% 
+  pivot_longer(ab:cd, names_to = "Question", values_to = "Response")
+
+round_12_ord_long <- round_12_ord_long %>% 
+  mutate(
+    Response = ifelse(Response %in% c("B is bigger", "D is bigger"), "Y is bigger (Correct)", 
+                      ifelse(Response == "They are the same", "They are the same", "X is bigger (Wrong)"))
+  )
+
+survey_12_ord <- 
+  survey::svydesign(
+  data = round_12_ord_long, 
+  ids = ~CaseId, weights = ~weight)
+
+model1_ord <- svyloglin(~Response*Question, survey_12_ord)
+
+# two logisting regressions
+model1_logist1 <- svyglm(I(Response == "Y is bigger (Correct)")~Question, survey_12_ord)
+# parameter Questioncd is effect of increase in accuracy due to alignment
+
+model1_logist2 <- svyglm(I(Response %in% c("Y is bigger (Correct)", "They are the same"))~Question, survey_12_ord)
+# For aligned tiles, 4.55 percentage points more respondents correctly pick D as the correct response or said the two tiles were the same size than picked B (or said they were the same size) in the unaligned tiles.
+
+summary(model1_logist2)
+```
+
 ## References{-}

### 2023-06-21 analysis/analysis.qmd
add figures for html

@@
 
-wt_scalar <- neff(round_12$weight)/sum(round_12$weight)
-round_12$weight <- wt_scalar*round_12$weight 
-# should not affect the further analysis, but helps with the interpretation of results
+round_12$weight <- neff(round_12$weight)/sum(round_12$weight)*round_12$weight 
+# does not affect the further analysis, but helps with the interpretation of results
 
@@
 Because each particpant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts.
-The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: < 2.2e-16).
+The difference in mean accuracy is `r round(paired$estimate,1)`.
+This difference is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: < 2.2e-16).
 
 ```{r accuracy, fig.height = 5, fig.width = 10, warning=FALSE}
-#| fig-cap: "On the left (a), a stacked barchart shows the number of respondents with correct (green) and incorrect (grey) responses to the two comparison questions.
+#| fig-cap: "On the left (a), a stacked barchart shows the number of respondents with correct (blue) and incorrect (orange) responses to the two comparison questions.
 When tiles are aligned along the same axis, more than twice the number of responses is accurate.
-The shaded area along the top of the green tiles corresponds to  95\\% confidence intervals around (marginal) correct responses.
+The shaded area along the top of the blue tiles corresponds to  95\\% confidence intervals around (marginal) correct responses.
 On the right (b), a parallel coordinate plot shows all combinations of responses.
@@
   xlab("") +
-  scale_fill_manual(values=c("grey", "forestgreen")) +
+  scale_fill_manual(values=c("orange", "steelblue")) +
   scale_y_continuous(
@@
     `Unaligned: A or B?` = ifelse(CorrectAB, "Correct", "Not correct"), 
-    `Aligned: C or D?` = ifelse(CorrectCD, "Correct", "Not correct")
+    `Aligned: C or D?` = ifelse(CorrectCD, "Correct", "Not correct"),
+    changed = `Unaligned: A or B?` != `Aligned: C or D?`,
+    changed = ifelse(changed, "changed", "unchanged")
     ) %>%
@@
     geom_pcp_axes() + 
-    geom_pcp(aes(colour = CorrectABCD), alpha = 0.2)+
+    geom_pcp(aes(colour = changed), alpha = 0.2)+
     geom_pcp_boxes(fill=NA, colour = "black", linewidth=0.25) + 
@@
   scale_y_continuous("Number responses", breaks=seq(0,1,by=0.25), 
-                     labels = round(sum(round_12$weight*wt_scalar)*seq(0,1,by=0.25),0),
+                     labels = round(sum(round_12$weight)*seq(0,1,by=0.25),0),
                      sec.axis = sec_axis(~.*100,
@@
   xlab("") + 
-  scale_colour_manual(values = c("grey", "#84AC84", "#70A570", "forestgreen")) +
+  scale_colour_manual("Response", values = c("purple", "grey")) +
   ggtitle("(b) Combinations of responses") +
-  theme(legend.position = "none")
+  theme(legend.position = "bottom") +
+  guides(colour=guide_legend(override.aes = list(alpha = 1)))
 
@@
 
-```{r}
+```{r data-ordinal}
 round_12_ord_long <- round_12 %>% 
@@
   mutate(
-    Response = ifelse(Response %in% c("B is bigger", "D is bigger"), "Y is bigger (Correct)", 
-                      ifelse(Response == "They are the same", "They are the same", "X is bigger (Wrong)"))
+    Response3 = ifelse(Response %in% c("B is bigger", "D is bigger"), "Y is bigger (Correct)", 
+                      ifelse(Response == "They are the same", "They are the same", "X is bigger (Wrong)")),
+    Response3 = factor(Response3, levels = c("X is bigger (Wrong)", "They are the same", "Y is bigger (Correct)")),
+    Question = ifelse(Question=="ab",
+                      "Unaligned: A or B?", "Aligned: C or D?")
   )
@@
   ids = ~CaseId, weights = ~weight)
+```
 
-model1_ord <- svyloglin(~Response*Question, survey_12_ord)
+```{r}
+paired2 <- svyttest(I(Response3=="X is bigger (Wrong)")~Question, survey_12_ord)
+```
+
+When moving beyond the binary accuracy measure of the response and using all three levels of the response as dependent variable, we see that by far the largest change in responses comes from a change from 'they are the same' to the (perceptually) correct response of 'D is bigger' when the tiles are aligned along the same axis for a comparison (see \ref{fig-ordinal}).
+Even though the number of wrong responses (orange) is only slightly different (`r round(paired2$estimate,3)`) between aligned and unaligned comparisons, it is still highly significant ($t$ statistic: `r round(paired2$statistic,1)`, df: `r paired2$parameter`, $p$-value: < 2.2e-16).
+
+```{r fig-ordinal, fig.height = 5, fig.width = 10, warning=FALSE}
+#| fig-cap: "On the left (a), a stacked barchart shows the number of respondents for each level of response: by aligning tiles, the number of responses of 'they are the same' is cut in half.
+On the right (b), a parallel coordinate plot shows all combinations of responses.
+The largest nuber of changes are from 'D is bigger' to 'they are the same'.
+"
+#| label: fig-ordinal
+#| out-width: 80%
+#| fig-env: figure
+
+gg1 <- round_12_ord_long %>%
+#  filter(Response == "Correct") %>%
+  ggplot(aes(x = Question, fill = Response3, weight = weight)) +
+  geom_bar() +
+  ylab("Number of (effective) Respondents") +
+  xlab("") +
+  scale_fill_manual(values=c("darkorange", "grey", "steelblue")) +
+  scale_y_continuous(
+      sec.axis = sec_axis(~./sum(round_12$weight)*100,
+                          name="Percent responses")
+  ) +
+  theme(legend.position = "none")
+
+gg1_build <- gg1 %>% ggplot2::ggplot_build()
+gg1_labels <- gg1_build$data[[1]] %>% mutate(
+  x = (xmin+xmax)/2,
+  y = (ymin+ymax)/2,
+  label = c("C is bigger (Wrong)", "They are the same", "D is bigger (Correct)",
+            "A is bigger (Wrong)", "They are the same", "B is bigger (Correct)"),
+  Response3 = rep(c("X is bigger (Wrong)", "They are the same", "Y is bigger (Correct)"),2)
+)
+
+gg1 <- gg1 +
+  ggtitle("(a) Three levels of survey responses")  +
+  geom_label(aes(x = x, y = y, label=label, weight=NULL), 
+             fill="white", alpha = 0.5, colour="black", data = gg1_labels)
+
+gg2 <- round_12 %>%
+  mutate(
+    `Unaligned: A or B?` = factor(ab, levels=c("B is bigger", "They are the same", "A is bigger")), 
+    `Aligned: C or D?` = factor(cd, levels=c("D is bigger", "They are the same", "C is bigger")), 
+    changed = (as.numeric(ab) != as.numeric(cd)),
+    changed = ifelse(changed, "changed", "unchanged")
+    ) %>%
+  pcp_select(`Aligned: C or D?`, `Unaligned: A or B?`) %>%
+  pcp_scale() %>%
+  pcp_arrange() %>%
+  ggplot(aes_pcp()) + 
+    geom_pcp_axes() + 
+    geom_pcp(aes(colour = changed), alpha = 0.2)+
+    geom_pcp_boxes(fill=NA, colour = "black", linewidth=0.25) + 
+    geom_pcp_labels(colour = "black") +
+  scale_y_continuous("Number responses", breaks=seq(0,1,by=0.25), 
+                     labels = round(sum(round_12$weight)*seq(0,1,by=0.25),0),
+                     sec.axis = sec_axis(~.*100,
+                          name="Percent responses")) +
+  xlab("") + 
+  scale_colour_manual("Response", values = c("purple", "grey")) +
+  ggtitle("(b) Combinations of responses") +
+  theme(legend.position = "bottom") + 
+  guides(colour=guide_legend(override.aes = list(alpha = 1)))
+
+gg1 + gg2
+```
+
+```{r model-ord}
+
+#model1_ord <- svyloglin(~Response3*Question, survey_12_ord)
 
 # two logisting regressions
-model1_logist1 <- svyglm(I(Response == "Y is bigger (Correct)")~Question, survey_12_ord)
+model1_logist1 <- svyglm(I(Response3 == "Y is bigger (Correct)")~Question, survey_12_ord, family = quasibinomial())
 # parameter Questioncd is effect of increase in accuracy due to alignment
 
-model1_logist2 <- svyglm(I(Response %in% c("Y is bigger (Correct)", "They are the same"))~Question, survey_12_ord)
+model1_logist2 <- svyglm(I(Response3 %in% c("Y is bigger (Correct)", "They are the same"))~Question, survey_12_ord, family = quasibinomial())
 # For aligned tiles, 4.55 percentage points more respondents correctly pick D as the correct response or said the two tiles were the same size than picked B (or said they were the same size) in the unaligned tiles.
+# exp(1.93092+0.49588)/(1 + exp(1.93092+0.49588)) - exp(1.93092)/(1 + exp(1.93092))
 
-summary(model1_logist2)
+antilog <- function(x) {exp(x)/(1+ exp(x))}
+
+contrasts1 <- as.data.frame(summary(model1_logist1)$coefficients)
+contrasts2 <- as.data.frame(summary(model1_logist2)$coefficients)
+contrasts <- data.frame(c("correct | same or wrong", "correct or same | wrong"), rbind(contrasts1[2,], contrasts2[2,]), 
+percent = 100*c(antilog(sum(model1_logist1$coefficients)) - antilog(model1_logist1$coefficients[1]),
+antilog(sum(model1_logist2$coefficients)) - antilog(model1_logist2$coefficients[1]))
+)
+names(contrasts) <- c("level","log odds", "SE", "t-statistic", "p-value", "percent change")
+knitr::kable(contrasts, digits = c(NA, 1, 2, 1, 4, 1), row.names = FALSE, caption = "Table 2: A change from aligned to unaligned tiles is detrimental to the level of correct (or correct and same) responses.", label = "tab:logodds")
 ```

### 2023-06-23 analysis/analysis.qmd
expand analysis to facetting

@@
 date: last-modified
-bibliography: ../SDSS\-2023/references.bib
+bibliography: "`r rbbt::bbt_write_bib('references.bib', overwrite = TRUE, library_id=rbbt::bbt_library_id('Graphics Research'), translator='bibtex')`"
 ---
@@
 library(ggpcp)
+library(broom)
+library(latex2exp)
+library(gt) # grammar of tables
 
@@
 }
+
+# inverse of the logistic function
+inv_logit <- function(x) {exp(x)/(1+ exp(x))}
 ```
@@
 
-```{r data}
+```{r data-1}
 # other github repo. 
@@
 
+```{r data-rnd7}
+# other github repo. 
+round7 <- readRDS("~/Documents/projects/NORC/visual-studies/data/round7.rds")
+
+# filter out all of the 'skipped on web':
+round7_adj <- round7 %>% mutate(
+  weight = WEIGHT_NOLA
+)
+```
+
 The data for this paper were collected in several rounds as part of the NORC Omnibus.
@@
 | Round 2 | May 2022 |  `r nrow(round2_adj)`   |    `r round(neff(round2_adj$weight),1)`       | `r round(sum(round2_adj$weight),1)`       |
+| Round 7 | Oct 22 |  `r nrow(round7_adj)`   |    `r round(neff(round7_adj$weight),1)`       | `r round(sum(round7_adj$weight),1)`       |
 |         |       |         |           |           |
@@
 $$
-$CV$ is the coefficient of variation of the weights within each sample, and can be estimated as
+$CV$ is the coefficient of variation of the weights within each sample, and is estimated as @kishSurveySampling1965
 
@@
 
-@omuircheartaighCombiningSamplesVs2002 calculated
+@omuircheartaighCombiningSamplesVs2002 estimate
 $\lambda$ separately for any combination of race/ethnicity by sex. 
+We will use that strategy whenever we include demographic variables in the analysis, otherwise we will use a single adjustment for the weights.
 
-We will calculate $\lambda$ separately whenever we include demographic variables in the analysis, otherwise we will use a single adjustment for the weights.
+All calculations are done in R [@RLanguage] using the `survey` package [@lumleyAnalysisComplexSurvey2004] version 4.0 [@survey] based on @lumleyComplexSurveysGuide2010.
 
-All calculations are done in R [@R2022] using the `survey` package [@lumleyAnalysisComplexSurvey2004] version 4.0 [@survey] based on @lumleyComplexSurveysGuide2010.
-
-## Model 1: Comparing Aligned and Unaligned Tiles {#sec-model1}
+## Model 1: Comparing Aligned and Unaligned Tiles in Vertical Stacked Barcharts {#sec-model1}
 
@@
 
+The data used for this is a combination of rounds 1 and 2, with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`.
+Figure \ref{fig-tasks-12} shows the two stacked barcharts shown to all panelists.
+
+```{r}
+#| label: fig-tasks-12
+#| fig-cap: "The two stacked barcharts every participant got to see.
+In each barchart, the two marked tiles are to be compared for their size.
+in both instances, the tile on the right is (very slightly) larger."
+#| fig-height: 3
+#| fig-width: 4
+#| out-width: 50%
+#| layout-nrow: 1
+#| fig-subcap:
+#| - "**Unaligned tiles:** Is A bigger than B?"
+#| - "**Aligned tiles:** Is C bigger than D?"
+knitr::include_graphics("images/vertical-ab.PNG")
+knitr::include_graphics("images/vertical-cd.PNG")
+```
+
 ```{r ttest}
 paired <- svyttest(Response~Question, survey_12)
+logit_12 <- svyglm(Response ~ Question, survey_12, family = quasibinomial())
+coefs_12 <- broom::tidy(logit_12)
 abcd <- svyby(~Response, by=~Question, survey_12, svytotal)
@@
 ```
-
-The data used for this is a combination of rounds 1 and 2, with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`.
 
@@
 Because each particpant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts.
-The difference in mean accuracy is `r round(paired$estimate,1)`.
-This difference is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: < 2.2e-16).
+The difference in mean accuracy is `r round(paired$estimate,2)`.
+This difference is highly significant ($t$ statistic: `r round(paired$statistic,2)`, df: `r paired$parameter`, $p$-value: < 2.2e-16).
 
@@
 gg1 + gg2
+```
+
+This test is equivalent to a logistic regression with response $Y_{ij}$, the response of participant $i$  to question $Q_j$ $j = 1/2$ is correct (=1) or incorrect (=0):
+$$
+\text{logit } P \left(Y = 1 \mid Q_j \right) = \alpha + \beta_j, 
+$$
+We assume that $Y \mid Q_j$ has a Bernoulli distribution with success probability $p_j$, $j = 1, 2$. 
+For the purpose of estimability, we will assume that $\beta_1 = 0$, i.e.
+$\text{logit} (p_1) = \alpha$ and $\beta_2 = \text{logit}(p_2) - \text{logit}(p_1)$.
+The odds of an accurate answer with aligned bars is therefore  exp (`r round(coefs_12$estimate[2],1)`) =  `r round(exp(coefs_12$estimate[2]),1)` times higher than for the unaligned bars of Question 1.
+
+```{r}
+coefs_12$term <- c("$$\\widehat{\\alpha}$$", "$$\\widehat{\\beta}_2$$")
+
+coefs_12 %>% 
+  gt() %>%
+  tab_header("Estimates for Model 1",
+             " accuracy of responses for unaligned (Question 1) and aligned bars (Question 2)") %>%
+  cols_label(
+    std.error = "SE",
+    statistic = "t-statistic",
+    p.value = "p-value"
+  ) %>%
+  fmt_number(
+    columns = c(estimate, std.error),
+    decimals = 2
+  ) %>%
+  fmt_number(
+    columns = statistic,
+    decimals = 1
+  ) %>%
+  fmt_scientific(
+    columns = p.value,
+    decimals = 1
+  )
 ```
@@
 
-antilog <- function(x) {exp(x)/(1+ exp(x))}
-
 contrasts1 <- as.data.frame(summary(model1_logist1)$coefficients)
@@
 contrasts <- data.frame(c("correct | same or wrong", "correct or same | wrong"), rbind(contrasts1[2,], contrasts2[2,]), 
-percent = 100*c(antilog(sum(model1_logist1$coefficients)) - antilog(model1_logist1$coefficients[1]),
-antilog(sum(model1_logist2$coefficients)) - antilog(model1_logist2$coefficients[1]))
+percent = 100*c(inv_logit(sum(model1_logist1$coefficients)) - inv_logit(model1_logist1$coefficients[1]),
+inv_logit(sum(model1_logist2$coefficients)) - inv_logit(model1_logist2$coefficients[1]))
 )
@@
 
+## Model 2: Comparing Aligned and Unaligned Elements Facets {#sec-model2}
+
+The survey setup and all stimuli are shown in \ref{fig-tasks-7}.
+A total of `r nrow(round7_adj)` panelists  (Effective sample size: `r round(neff(round7_adj$weight), 1)`) were shown four sets of facetted charts each.
+For questions 3 and 4, the panel was split into two (roughly) halves.
+One set of panelists was shown the floating wedges, while the other half was shown the framed wedges on the right.
+
+For each chart, panelists were asked to compare the sizes of the marked tiles against each other.
+
+|                              |                                |                                |
+|------------------------------|--------------------------------|--------------------------------|
+|   **Everybody:**       |  **Question 1:** Facetted Pie, unaligned | **Question 2:** Facetted Pie, aligned |
+|   | ![Facetted pie charts, unaligned](images/round7/unaligned pie - all.png)   |   ![Facetted pie charts, aligned](images/round7/aligned pie - all.png)  |
+|   **Split Sample**      | Effective sample size:  `r round(neff(filter(round7_adj, RND_05 == 0)$weight), 1)`                   |   Effective sample size: `r round(neff(filter(round7_adj, RND_05 == 1)$weight), 1)`          |
+|   **Question 3:**  unaligned  |    Floating pie wedges | Framed pie wedges |
+|            | ![Question 3](images/round7/unaligned pie floating - sample 1.png)                    |            ![Question 3](images/round7/unaligned pie framed - sample 2.png)          |
+|  **Question 4:**  aligned     |     ![Question 4](images/round7/aligned pie floating - sample 1.png)                     |               ![Question 4](images/round7/aligned pie framed - sample 2.png)                    |
+
+: Setup of questions and stimuli for Round 7. {#tbl-tasks-7 tbl-colwidths="\[20,40,40\]"}
+
+For a comparison of wedges marked X and Y, there were three possible options for the response: (1) X is bigger, (2) Y is bigger or (3) They are the same.
+
+The four labelled pieces A, B, C, and D correspond to values 46, 47.6, 45.2, and 44.4, respectively, i.e underlying each evaluation is the order D < C < A < B.
+Nominally, the differences between the size of the marked elements can be expressed as multiples in $\delta = 0.8$, with $B = A + 2\delta = C + 3 \delta = D + 4 \delta$.
+A difference of $\delta = 0.8$ corresponds to about a $3^\degree$ difference for the angle of a wedge.
+We would expect that an increase in the difference in the values increases the accuracy in the responses.
+This is mitigated by the distance between wedges: more distance (should) lead to a decrease in accuracy.
+Similarly, if pieces have an axis or a line of reference in common, we would expect an increase in accuracy.
+
+There are different possible strategies for evaluating the size of a wedge in a pie chart.
+Panelists can use the **angle** of the wedge (or a suitable derivative of it, such as the complementary angle or the remainder to 180 degrees), its **area** (as part of the circle or some derivative) or the **arc length** of the wedge.
+Studies by  @kosaraJudgmentErrorPie2016 and @kosaraEvidenceAreaPrimary2019 show that the area of a piece in a pie chart is the primary evaluation method.
+
+```{r eval=TRUE}
+widths <- c(28, 16, 31, 38)
+heights <- c(206, 209, 206, 203)
+
+tan_est <- (pi-atan(widths/(heights-105)))/pi*180
+cos_est <- (pi/2+acos(widths/105))/pi*180
+alpha_est <- round((tan_est+cos_est)/2,1)
+
+values <- c(46, 47.6, 45.2,  44.4)
+names(values) <- LETTERS[1:4]
+```
+
+<!--
+In the (aligned wedges), those values are mapped to the angles `r alpha_est`, for an approximate ratio of angle to value of `r round(alpha_est/values,2)` degrees to 1.
+A circle with a radius of 105 pixels has a circumference of 660 pixels.
+The ratio of the wedges' arc length to the values are `r round((alpha_est/180) *105/values,2)` pixels to 1.
+
+A circle with a radius of 105 pixels has an area of 34636 square pixels. The ratio of the wedges' areas to the values are `r round((alpha_est/360) *105^2/values,1)` square pixels to 1.
+
+We only evaluated the aligned wedges for these ratios, most likely there are small differences in the exact pixel values for the unaligned wedges due to rounding issues.
+-->
+
+```{r survey-round7}
+response_type <- round7_adj %>%
+  pivot_longer(
+    matches("NOLA[1-6]_[1-6]"), 
+    names_to = "taskquestion", values_to = "responses")
+
+response_type <- response_type %>% filter(!is.na(responses), responses!="SKIPPED ON WEB")
+
+response_type <- response_type %>% separate(col="taskquestion", into=c("task", "question"), sep="_")
+
+response_type <- response_type %>% mutate(
+  question = parse_number(question),
+  labelled = c("AB", "AC", "AD", "BD", "BC", "CD")[question]
+)
+
+# D < C < A < B.
+response_type <- response_type %>% mutate(
+  correct = ifelse(labelled %in% c("AB", "BC", "BD"), responses == "B is bigger", 
+              ifelse(labelled %in% c("AC", "AD"), responses == "A is bigger",
+                     responses == "C is bigger")),
+  correct3 = correct | responses == "They are the same"
+)
+
+response_type <- response_type %>% 
+  mutate(
+    responses = factor(responses, levels = c("B is bigger", "A is bigger", "C is bigger", "D is bigger", "They are the same")),
+    taskname = c("Pie", "Pie", "Float", "Float", "Frame", "Frame")[parse_number(gsub("NOLA_","",task))],
+    taskname = factor(taskname, levels=c("Frame", "Float", "Pie")),
+    concept = c("Un-", "A-", "Un-", "A-", "Un-", "A-")[parse_number(gsub("NOLA_","",task))],
+    signal = ifelse(labelled %in% "BD", 4, 
+              ifelse(labelled %in% "BC", 3, 
+                ifelse(labelled %in% c("AB", "AD"), 2, 1)))
+  )
+
+survey_7 <- svydesign(~CaseId, data = response_type, weights = ~weight)
+```
+
+```{r fig-pies, eval=TRUE, fig.width = 10, fig.height = 5, fig.cap="Panelist responses by task and comparison.
+On average, the accuracy for aligned pieces is higher than for unaligned pieces.
+Similarly, the accuracy for framed pieces is higher than for floating pieces, which in turn has a higher accuracy than the full pie chart.", out.width="100%", warning =FALSE}
+# D < C < A < B.
+#   1   1.  2
+
+correct_ci <- as.data.frame(svyby(~correct, by=~interaction(labelled, concept,taskname), svymean, design = survey_7))
+correct_ci <- correct_ci %>% 
+  mutate(
+    labels = row.names(correct_ci),
+  ) 
+correct_ci <- correct_ci %>% separate_wider_delim(
+    labels, delim = ".", names =c("labelled", "concept", "taskname")
+  ) %>% mutate(
+    taskname = factor(taskname, levels=c("Frame", "Float", "Pie")),
+    labelled = factor(labelled, levels =c("BD", "BC", "AB", "AD", "AC", "CD")),
+    ct = interaction(concept, taskname)
+  )
+
+correct3_ci <- as.data.frame(svyby(~correct3, by=~interaction(labelled, concept,taskname), svymean, design = survey_7))
+correct3_ci <- correct3_ci %>% 
+  mutate(
+    labels = row.names(correct3_ci),
+  ) 
+correct3_ci <- correct3_ci %>% separate_wider_delim(
+    labels, delim = ".", names =c("labelled", "concept", "taskname")
+  ) %>% mutate(
+    taskname = factor(taskname, levels=c("Frame", "Float", "Pie")),
+    labelled = factor(labelled, levels =c("BD", "BC", "AB", "AD", "AC", "CD")),
+    ct = interaction(concept, taskname)
+  )
+
+gg2b <- response_type %>% 
+  mutate(
+    labelled = factor(labelled, levels =c("BD", "BC", "AB", "AD", "AC", "CD")),
+    ct = interaction(concept, taskname)
+  ) %>% 
+  ggplot(aes(x = ct, fill=factor(as.numeric(correct)+as.numeric(correct3)), weight = weight)) +
+  geom_bar(position="fill") +
+  facet_grid(~labelled, space = "free") +
+  theme_bw() +
+  ggtitle("Which piece is bigger?") +
+  scale_fill_manual(values=rev(c("steelblue", "grey", "orange"))) + 
+  scale_x_discrete(labels=c("Frame\nA", "\nN", "Float\nA", "\nU", "Pie\nA", "\nU")) +
+  ylab("") +
+  geom_rect(
+    aes(
+      xmin = as.numeric(ct) - 0.45,
+      ymin = (correctTRUE - 1.96* se.correctTRUE),
+      xmax = as.numeric(ct) + 0.45,
+      ymax = (correctTRUE + 1.96* se.correctTRUE),
+      weight = 1), 
+    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+    data = correct_ci
+  ) +
+  geom_rect(
+    aes(
+      xmin = as.numeric(ct) - 0.45,
+      ymin = (correct3TRUE - 1.96* se.correct3TRUE),
+      xmax = as.numeric(ct) + 0.45,
+      ymax = (correct3TRUE + 1.96* se.correct3TRUE),
+      weight = 1), 
+    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+    data = correct3_ci
+  ) +
+  geom_hline(yintercept=c(.25,.50,.75), colour = "grey50", linewidth=0.35) +
+  geom_hline(yintercept=c(.25,.50,.75), colour = "grey90", linewidth=0.35,
+             linetype="dashed") +
+  xlab("") +
+  theme(legend.position = "none")
+
+gg2b
+```
+
+Figure \ref{fig-pies} shows a summary of panelists' accuracy of their responses.
+The panels are sorted according to the level of difficulty of a comparison as hypothesized above from lowest difficulty (BD) to highest (CD).
+We see that the percentage of correct responses follows (on average) this hypothesized order.
+Using the comparison of A versus B as the baseline, we see from Table \ref{tab-model7}, that comparisons BC and BD have a significantly higher accuracy (in that order), while the accuracy of comparisons AD, AC, and CD are significantly worse (HH: should we do pairwise comparisons for that?).
+Comparisons involving aligned pieces have (with two exceptions) a higher rate of accuracy than a  comparison of the same-size, unaliged pieces.  
+Generally, framed pieces have the highest rate of accuracy, while the comparisons within the full context of the pie chart fared the worst, both in the percent of 'the same' responses as well as in identifying the smaller piece as the bigger.
+
+Using a frame as an additional visual cue was shown by @clevelandGraphicalPerceptionTheory1984 to significantly improve the accuracy of perceiving the height of bars (cf.
+Fig.
+12, p.
+538; difference between 'length' and 'position along an unaligned identical axis').
+Here, we find that framed pieces show the same trend in accuracy, however, this effect fails to become significant in the model (tasknameFloat).
+
+The fact that the full pie chart sees the lowest accuracy rates is similar to the finding in @talbotFourExperimentsPerception2014 for bars.
+@talbotFourExperimentsPerception2014 report that other bars serve as distractors when comparing two other bars.
+
+```{r}
+logit_7 <- svyglm(correct ~ labelled + concept + taskname, family = quasibinomial(), design = survey_7)
+coefs_7 <- broom::tidy(logit_7)
+
+#logit_7b <- svyglm(correct ~ factor(signal) + concept + taskname, family = quasibinomial(), design = survey_7)
+#logit_7c <- svyglm(correct ~ signal + concept + taskname, family = quasibinomial(), design = survey_7)
+
+coefs_7 %>% 
+  mutate(
+    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value))) %>%
+  gt() %>%
+  tab_header("Estimates for Model 2",
+             " accuracy of responses across conditions of Round 7") %>%
+  cols_label(
+    std.error = "SE",
+    statistic = "t-statistic",
+    p.value = "p-value"
+  ) %>%
+  fmt_number(
+    columns = c(estimate, std.error),
+    decimals = 2
+  ) %>%
+  fmt_number(
+    columns = statistic,
+    decimals = 1
+  ) %>%
+  cols_align(
+    columns = p.value,
+    align = "right"
+  )
+```
+
 ## References{-}

### 2023-06-24 analysis/analysis.qmd
model 2 - halfway?

@@
 inv_logit <- function(x) {exp(x)/(1+ exp(x))}
+
+# calculate lambda for combining weights
+calculate_lambda <- function(weights1, weights2) {
+  d1 <- 1+var(weights1)/mean(weights1)^2
+  d2 <- 1+var(weights2)/mean(weights2)^2
+
+  n1 <- length(weights1)
+  n2 <- length(weights2)
+
+  denom <- n1/d1 + n2/d2
+
+  lambda <- n1/(d1*denom) 
+  lambda
+}
+
+vs_repo_local <- "~/Documents/projects/NORC" # link to local version of the visual_studies repo
 ```
@@
 # other github repo. 
-round1 <- readRDS("~/Documents/projects/NORC/visual-studies/data/round1.rds")
-round2 <- readRDS("~/Documents/projects/NORC/visual-studies/data/round2.rds")
+round1 <- readRDS(file.path(vs_repo_local, "visual-studies/data/round1.rds"))
+round2 <- readRDS(file.path(vs_repo_local, "visual-studies/data/round2.rds"))
 
@@
 
+```{r data-rnd6}
+# other github repo. 
+round6 <- readRDS(file.path(vs_repo_local, "visual-studies/data/round6.rds"))
+
+round6_adj <- round6 %>% mutate(
+  weight = WEIGHT_NOLA
+) %>% filter(RND_01 == 0) # same type of questions as round 7
+```
+
 ```{r data-rnd7}
 # other github repo. 
-round7 <- readRDS("~/Documents/projects/NORC/visual-studies/data/round7.rds")
+round7 <- readRDS(file.path(vs_repo_local, "visual-studies/data/round7.rds"))
 
@@
 | Round 2 | May 2022 |  `r nrow(round2_adj)`   |    `r round(neff(round2_adj$weight),1)`       | `r round(sum(round2_adj$weight),1)`       |
-| Round 7 | Oct 22 |  `r nrow(round7_adj)`   |    `r round(neff(round7_adj$weight),1)`       | `r round(sum(round7_adj$weight),1)`       |
+| Round 6 | Sep 2022 |  `r nrow(round6_adj)` [Split]  |    `r round(neff(round6_adj$weight),1)`       | `r round(sum(round6_adj$weight),1)`       |
+| Round 7 | Oct 2022 |  `r nrow(round7_adj)`   |    `r round(neff(round7_adj$weight),1)`       | `r round(sum(round7_adj$weight),1)`       |
 |         |       |         |           |           |
@@
 ```{r combine_rounds_1_and_2}
+# all weights are scaled to add to 1 first, then lambda is used to 
+# get the convex weight assignment, 
+# then the weights are scaled to the effective sample size
 round_12 <- round1_adj %>% 
-            mutate(weight = lambda*weight) %>% 
+            mutate(weight = lambda*weight/sum(weight)) %>% 
             select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight) %>%
     rbind(round2_adj %>% 
-            mutate(weight = (1-lambda)*weight) %>% 
+            mutate(weight = (1-lambda)*weight/sum(weight)) %>% 
             select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight))
@@
 coefs_12 %>% 
+  mutate(
+    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value))
+  ) %>%
   gt() %>%
@@
   ) %>%
-  fmt_scientific(
+  cols_align(
     columns = p.value,
-    decimals = 1
+    align = "right"
   )
@@
 
-## Model 2: Comparing Aligned and Unaligned Elements Facets {#sec-model2}
+## Model 2: Aligned and unaligned comparisons between facetted barcharts and facetted piecharts {#sec-model2}
+
+A total of three different charts were shown to (a part) of the panelists from Round 6 and all panelists of Round 7. 
+The four labelled pieces A, B, C, and D correspond to values 46, 47.6, 45.2, and 44.4, respectively, i.e underlying each evaluation is the order D < C < A < B.
+
+```{r}
+#| label: fig-tasks-67
+#| fig-cap: "Stimuli shown to panelists. For each of the charts, all marked pieces were evaluated against each other. "
+#| fig-height: 3
+#| fig-width: 4
+#| layout-nrow: 1
+#| fig-subcap:
+#| - "**Round 6**: Facetted barcharts"
+#| - "**Round 7**: Facetted piecharts with unaligned wedges"
+#| - "**Round 7**: Facetted piecharts with aligned wedges"
+knitr::include_graphics("images/Round 6 - Sept 2022.png")
+knitr::include_graphics("images/round7/unaligned pie - all.png")
+knitr::include_graphics("images/round7/aligned pie - all.png")
+```
+
+```{r combine_rnds67}
+lambda6 <- calculate_lambda(round6_adj$weight, round7_adj$weight)
+round6_adj <- round6_adj %>% 
+  mutate(
+    weight = lambda6*weight/sum(weight)
+  )
+round7_adj <- round7_adj %>% 
+  mutate(
+    weight = (1-lambda6)*weight/sum(weight)
+  )
+
+response_6 <- round6_adj %>%
+  pivot_longer(
+    ab1:cd1, # barcharts 
+    names_to = "taskquestion", values_to = "responses") %>%
+  filter(!is.na(responses), responses!="SKIPPED ON WEB") %>%
+  mutate(
+    labelled = toupper(gsub("1", "", taskquestion)),
+    task = "Bar"
+  )
+
+response_7 <- round7_adj %>%
+  pivot_longer(
+    matches("NOLA[1-2]_[1-6]"), # only pie charts
+    names_to = "taskquestion", values_to = "responses") %>%
+  filter(!is.na(responses), responses!="SKIPPED ON WEB") %>%
+  separate(col="taskquestion", into=c("task", "question"), sep="_")  %>%
+  mutate(
+    question = parse_number(question),
+    labelled = c("AB", "AC", "AD", "BD", "BC", "CD")[question],
+    task = c("Pie\nUn-", "Pie\nA-")[parse_number(task)]
+  )
+
+# D < C < A < B.
+response_67 <- rbind(
+    response_6 %>% select(CaseId, labelled, task, responses, weight),
+    response_7 %>% select(CaseId, labelled, task, responses, weight)
+  ) %>% mutate(
+  correct = ifelse(labelled %in% c("AB", "BC", "BD"), responses == "B is bigger", 
+              ifelse(labelled %in% c("AC", "AD"), responses == "A is bigger",
+                     responses == "C is bigger")),
+  correct3 = correct | responses == "They are the same",
+  response3 = factor(correct+correct3, labels = c("Wrong", "Same", "Correct"))
+)
+
+survey_67 <- svydesign(~CaseId, data = response_67, weights = ~weight)
+```
+
+Figure \ref{fig-response-67} gives an overview of all responses to size assessments of the marked tiles.
+The panels are ordered according to the difference in signal and hypothesized difficulty from easiest (left) to hardest (right) comparison.
+We see, that in general the percentage of correct responses decreases from left to right.
+Design does not seem to matter for easy comparisons (BD, BC) or the hard comparison (CD). 
+The three comparisons in the middle show interesting patterns, highlighted in the reordered figure \ref{fig-response-67b}.
+
+```{r}
+#| label: fig-response-67
+#| fig-cap: "Responses for comparisons AB through CD on the three different facetted designs.
+Responses to the same comparison are shown side-by-side for the three designs.
+Comparisons are ordered from least difficult (BD) to most difficult (CD).
+The increasing difficulty of comparisons is reflected in the overall decreasing percentage of correct responses (height of blue tiles).
+The overlaid rectangles represent 95% confidence intervals."
+#| fig-height: 4.25
+#| fig-width: 7
+#| out-width: 80%
+
+props_ci <- as.data.frame(svyby(~response3, by=~interaction(labelled, task), svymean, design = survey_67))
+props_ci <- props_ci %>% 
+  mutate(
+    labels = row.names(props_ci),
+  ) 
+props_ci <- props_ci %>% separate_wider_delim(
+    labels, delim = ".", names =c("labelled", "task")
+  ) %>% mutate(
+    task = factor(task, levels=c("Bar", "Pie\nA-", "Pie\nUn-")),
+    labelled = factor(labelled, levels =c("BD", "BC", "AB", "AD", "AC", "CD"))
+  )
+
+response_67 %>% 
+  mutate(
+    labelled = factor(labelled, levels=c("BD", "BC", "AB", "AD", "AC", "CD"))
+  ) %>% 
+  ggplot(aes(x = task)) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  facet_grid(.~labelled) + 
+  theme(legend.position="bottom") + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") + 
+  geom_rect(
+    aes(
+      xmin = as.numeric(task) - 0.45,
+      ymin = (response3Correct - 1.96* se.response3Correct),
+      xmax = as.numeric(task) + 0.45,
+      ymax = (response3Correct + 1.96* se.response3Correct),
+    ), 
+    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+    data = props_ci
+  ) +
+  geom_segment(
+    aes(
+      x = as.numeric(task) - 0.45,
+      y = (response3Correct ),
+      xend = as.numeric(task) + 0.45,
+      yend = (response3Correct )
+    ), 
+    alpha = 1, colour = "black", linewidth = 0.25,
+    data = props_ci
+  ) +
+  geom_rect(
+    aes(
+      xmin = as.numeric(task) - 0.45,
+      ymin = (1- response3Wrong - 1.96* se.response3Wrong),
+      xmax = as.numeric(task) + 0.45,
+      ymax = (1-response3Wrong + 1.96* se.response3Wrong),
+    ), 
+    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+    data = props_ci
+  ) +
+  geom_segment(
+    aes(
+      x = as.numeric(task) - 0.45,
+      y = (1- response3Wrong ),
+      xend = as.numeric(task) + 0.45,
+      yend = (1-response3Wrong )
+    ), 
+    alpha = 1, colour = "black", linewidth = 0.25,
+    data = props_ci
+  ) +
+  scale_y_continuous("Percent responses", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25))
+```
+
+For the facetteed barcharts, there is an abrupt drop in accuracy of responses between 'easy' and 'hard' comparisons.
+(HH: All of these comparisons could be affected by the context of their light blue neighbors.
+'B' is bigger than its light blue neighbor, and also the overall biggest.
+'A' is only a bit smaller than 'B', but quite a bit smaller than the light blue bar next to it.
+This seems to make comparisons with A harder.)  For pie charts the accuracy in responses is reacting a lot more gentle to the increasing difficulty.
+There is no apparent effect in pie charts due to alignment.
+However, all comparisons are very close to 50% - the intrinsic reference lines [@lipkusVisualCommunicationRisk1999] in pie charts (at 180 degrees, and to a lesser degree at 90 degree angles) might be helping with this particular set of comparisons.
+
+```{r}
+#| label: fig-response-67b
+#| fig-cap: "Responses for comparisons AB through CD on the three different facetted designs, reordered.
+Now, responses to the same design are shown side-by-side across comparisons.
+Comparisons are ordered from least difficult (BD) to most difficult (CD).
+The three different designs result in different response patterns for increasing comparison difficulty.
+"
+#| fig-height: 4.25
+#| fig-width: 7
+#| out-width: 80%
+
+props_ci <- as.data.frame(svyby(~response3, by=~interaction(labelled, task), svymean, design = survey_67))
+props_ci <- props_ci %>% 
+  mutate(
+    labels = row.names(props_ci),
+  ) 
+props_ci <- props_ci %>% separate_wider_delim(
+  labels, delim = ".", names =c("labelled", "task")
+) %>% mutate(
+  task = factor(task, levels=c("Bar", "Pie\nA-", "Pie\nUn-")),
+  labelled = factor(labelled, levels =c("BD", "BC", "AB", "AD", "AC", "CD"))
+)
+
+response_67 %>% 
+  mutate(
+    labelled = factor(labelled, levels=c("BD", "BC", "AB", "AD", "AC", "CD"))
+  ) %>% 
+  ggplot(aes(x = labelled)) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  facet_grid(.~task) + 
+  theme(legend.position="bottom") + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") + 
+  geom_rect(
+    aes(
+      xmin = as.numeric(labelled) - 0.45,
+      ymin = (response3Correct - 1.96* se.response3Correct),
+      xmax = as.numeric(labelled) + 0.45,
+      ymax = (response3Correct + 1.96* se.response3Correct),
+    ), 
+    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+    data = props_ci
+  ) +
+  geom_segment(
+    aes(
+      x = as.numeric(labelled) - 0.45,
+      y = (response3Correct ),
+      xend = as.numeric(labelled) + 0.45,
+      yend = (response3Correct )
+    ), 
+    alpha = 1, colour = "black", linewidth = 0.25,
+    data = props_ci
+  ) +
+  geom_rect(
+    aes(
+      xmin = as.numeric(labelled) - 0.45,
+      ymin = (1- response3Wrong - 1.96* se.response3Wrong),
+      xmax = as.numeric(labelled) + 0.45,
+      ymax = (1-response3Wrong + 1.96* se.response3Wrong),
+    ), 
+    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+    data = props_ci
+  ) +
+  geom_segment(
+    aes(
+      x = as.numeric(labelled) - 0.45,
+      y = (1- response3Wrong ),
+      xend = as.numeric(labelled) + 0.45,
+      yend = (1-response3Wrong )
+    ), 
+    alpha = 1, colour = "black", linewidth = 0.25,
+    data = props_ci
+  ) +
+  scale_y_continuous("Percent responses", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25))
+```
+
+## Model X: Comparing Aligned and Unaligned Elements Facets {#sec-modelX}
 
@@
 
-For each chart, panelists were asked to compare the sizes of the marked tiles against each other.
+For each chart, panelists were asked to compare all marked tiles against each other.
 
@@
 ```
-
-<!--
-In the (aligned wedges), those values are mapped to the angles `r alpha_est`, for an approximate ratio of angle to value of `r round(alpha_est/values,2)` degrees to 1.
-A circle with a radius of 105 pixels has a circumference of 660 pixels.
-The ratio of the wedges' arc length to the values are `r round((alpha_est/180) *105/values,2)` pixels to 1.
-
-A circle with a radius of 105 pixels has an area of 34636 square pixels. The ratio of the wedges' areas to the values are `r round((alpha_est/360) *105^2/values,1)` square pixels to 1.
-
-We only evaluated the aligned wedges for these ratios, most likely there are small differences in the exact pixel values for the unaligned wedges due to rounding issues.
--->
 

### 2023-06-26 SDSS-2023/testing-charts-accuracy.qmd
a bit of the intro

@@
 
-@clevelandGraphicalPerceptionTheory1984
-
 ```{r setup, include = FALSE}
@@
 library(survey)
+#library(ggmosaic)
 library(ggpcp)
+library(broom)
+library(latex2exp)
+library(gt) # grammar of tables
 
@@
 }
+
+# inverse of the logistic function
+inv_logit <- function(x) {exp(x)/(1+ exp(x))}
+
+# calculate lambda for combining weights
+calculate_lambda <- function(weights1, weights2) {
+  d1 <- 1+var(weights1)/mean(weights1)^2
+  d2 <- 1+var(weights2)/mean(weights2)^2
+
+  n1 <- length(weights1)
+  n2 <- length(weights2)
+
+  denom <- n1/d1 + n2/d2
+
+  lambda <- n1/(d1*denom) 
+  lambda
+}
+
+vs_repo_local <- "~/Documents/projects/NORC" # link to local version of the visual_studies repo
 ```
@@
 
-How do structural design choices in a data visualization impact viewers’ ability to identify the larger of two elements?
+<!-- What is our main objective? -->
+What do viewers see, when we show them a data chart?
+A crucial step in the process of 'understanding' a chart, is that a visual allows us to make comparisons between its parts.
+@clevelandGraphicalPerceptionTheory1984 defined for the purpose of their seminal study the better visual as the one that allows viewers to make more accurate comparisons.
+What we always need to remember, is that these study results are based on only a few individuals recruited from the authors' coworkers and their spouses.
+@heerCrowdsourcingGraphicalPerception2010 have reproduced Cleveland and McGill's rankings using a crowd sourcing platform.
 
-How do aesthetic design choices in a data visualization impact viewers’ ability to identify the larger of two elements?
+<!-- Bias from crowd-sourcing? sample size?  -->
+crowd workers are known to be biased towards male, young and relatively higher education: @borgoCrowdsourcingInformationVisualization2017
 
-How is viewer behavior (zooming, time spent on question, certainty of response) impacted by structural and aesthetic design choices in a data visualization?
+Here, we seek to -- first -- answer the question whether it is possible to use a survey to reproduce (some of) the rankings.
 
-Structural design choices:
+```{r}
+#| label: fig-aligned-unaligned
+#| fig-cap: "The only difference between the two pairs of rectangles A, B and C, D is their alignment, i.e.
+A and C are of identical size, as are B and D.
+While the predominant response for the unaligned pair on the left is 'they are of the same size', more than half of the viewers respond with 'D is bigger' to the aligned pair of bars on the right."
+#| fig-subcap:
+#| - "**Unaligned bars**"
+#| - "**Aligned bars**"
+#| fig-cap-location: "bottom"
+#| layout: [[45,45], [100]]
+#| fig-align: "center"
+#| out-width: 100%
+knitr::include_graphics("images/round5-modified.PNG")
+knitr::include_graphics("images/round3-modified.PNG")
+```
+
+<!-- motivate the use of just noticeable differences -->
+Asking perceptual questions in a survey is different from the controlled environment of a cognitive lab, where these kind of questions would usually be addressed.
+This means, that instructions to participants have to be delivered in a very short and easily understandable, because questions arising from the task can not be answered.
+Similarly, rather than asking the same (or similar) type of question with varied signals multiple hundred times, in a survey we can ask only a few questions.
+In order to observe any effect, we need to ask questions that are perceptually hard, which means that we need to ask questions about stimuli that are close to our perceptual threshold.
+The **Just-Noticeable Difference** (JND) is defined as the smallest difference that will be detected 50% of the time.
+We are using results from studies on barcharts and pie charts [@luModelingJustNoticeable2022] to inform the differences in charts shown to survey panelists.
+
+- How do structural design choices in a data visualization impact viewers’ ability to identify the larger of two elements?
+
+- How do aesthetic design choices in a data visualization impact viewers’ ability to identify the larger of two elements?
+
+- How is viewer behavior (zooming, time spent on question, certainty of response) impacted by structural and aesthetic design choices in a data visualization?
 
@@
 
-<!-- motivate the use of just noticeable differences -->
-Asking perceptual questions in a survey is different from the controlled environment of a cognitive lab, where these kind of questions would usually be addressed.
-This means, that instructions to participants have to be delivered in a very short and easily understandable, because questions arising from the task can not be answered.
-Similarly, rather than asking the same (or similar) type of question multiple hundred times, in a survey we can ask only a few questions.
-In order to observe any effect, we need to ask questions that are perceptually hard, which means that we need to ask questions about stimuli that are close to our perceptual threshold.
-The **Just-Noticeable Difference** (JND) is defined as the smallest difference that will be detected 50% of the time.
-For the height difference of barcharts, @luModelingJustNoticeable2022 determined that the distance between bars is the main contributor to differentiating sizes.
-Based on their models, a tile difference of 155 pixels leads to a just noticeable difference of 3.5 pixels.
 \autoref{tbl-tasks-12} shows the two stacked barcharts shown to participants in the survey.
 The marked tiles in each plot are 155 pixels apart.
-Their heights are 205 (left) and 213 pixels (right), respectively, corresponding to about twice the JND.
+Based on @luModelingJustNoticeable2022's model, a difference of 155 pixels leads to a just noticeable difference of 3.5 pixels.
+The heights of the bars are 205 (left) and 213 pixels (right), respectively, corresponding to about twice the JND.
 This difference should lead to a relatively high accuracy rate for participants and simultaneously limit the amount of frustration resulting from a task that is perceived as 'too hard'.
@@
 
-All calculations in this paper are done in R [@R2022] using the `survey` package [@lumleyAnalysisComplexSurvey2004] version 4.0 [@survey] based on @lumleyComplexSurveysGuide2010.
+All calculations in this paper are done in R [@RLanguage] using the `survey` package [@lumleyAnalysisComplexSurvey2004] version 4.0 [@survey] based on @lumleyComplexSurveysGuide2010.
 

### 2023-06-26 analysis/analysis.qmd
Two more dependent variables
a bit of the intro

@@
 
+### How certain are participants?
+
+### Certainty and Accuracy
+
+HH: For rounds 1 and 2 we do not have individual timing of tasks.
+
 ## Model 2: Aligned and unaligned comparisons between facetted barcharts and facetted piecharts {#sec-model2}
@@
   correct3 = correct | responses == "They are the same",
-  response3 = factor(correct+correct3, labels = c("Wrong", "Same", "Correct"))
+  response3 = factor(correct+correct3, labels = c("Wrong", "Same", "Correct")),
+  labelled = factor(labelled, levels=c("BD", "BC", "AB", "AD", "AC", "CD")),
+  signal = ifelse(labelled=="BD", 4, 
+             ifelse(labelled=="BC", 3,
+               ifelse(labelled %in% c("AB", "AD"), 2, 1)))
 )
@@
   scale_y_continuous("Percent responses", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25))
+```
+
+```{r model-67}
+# still putzing around with it
+logit_67_correct <- svyglm(correct~task+labelled + task, family=quasibinomial(), design=survey_67)
+
+logit_67_cont <- svyglm(correct~task*signal, family=quasibinomial(), design=survey_67)
+
 ```

### 2023-07-05 analysis/analysis.qmd
still working on the aligned vs unaligned model

@@
             mutate(weight = lambda*weight/sum(weight)) %>% 
-            select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight) %>%
+            select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
+                   ab_certain, cd_certain, weight) %>%
     rbind(round2_adj %>% 
             mutate(weight = (1-lambda)*weight/sum(weight)) %>% 
-            select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight))
+            select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, 
+                   ab_certain, cd_certain, weight))
 
@@
 #| label: fig-alignment
-#| out-width: 80%
+#| out-width: 90%
 #| fig-env: figure
@@
 
-### Certainty and Accuracy
+```{r certainty-data}
+round_12_certainty_long <- round_12 %>% 
+  pivot_longer(ab_certain:cd_certain, names_to = "Question", values_to = "Response") %>%
+  filter(Response != "SKIPPED ON WEB") %>%  # 13 participants
+  mutate(
+    Correct = ifelse(Question == "ab_certain", CorrectAB, CorrectCD),
+    Question = ifelse(Question=="ab_certain", "A vs B (Unaligned)", "C vs D (Aligned)"),
+# ordering of levels in Response is from extremely=1 to not at all=5 - not very intuitive
+    Response = factor(Response, levels = rev(levels(Response)))
+  )
 
-HH: For rounds 1 and 2 we do not have individual timing of tasks.
+survey_12_crt <- 
+  survey::svydesign(
+  data = round_12_certainty_long, 
+  ids = ~CaseId, weights = ~weight)
+```
+
+```{r}
+#| label: fig-certainty
+#| out-width: 45%
+#| fig-height: 5
+#| fig-width: 5
+#| fig-cap: "Two views of certainty: The barcharts in (a) show that aligning bars along the same access is associated with more participants who are very certain of their response compared to the unaligned comparison of A vs B, where the majority of participants resports 'moderate' certainty.
+In the barcharts in (b) we see that participants are not quite as certain of their responses when their answers are correct as when their answers are not correct.
+Both of these are significant (see model 2.3)."
+#| layout-nrow: 1
+#| fig-subcap: 
+#|   - Certainty by Task
+#|   - Certainty by Accuracy of Response
+round_12_certainty_long %>% 
+  ggplot(aes(x = Response, weight=weight)) + geom_bar() + 
+  ylab("Number of (weighted) responses") + 
+  facet_grid(Question~.) +
+  ggtitle("Certainty by Task")
+
+round_12_certainty_long %>% 
+  ggplot(aes(x = Response, weight=weight)) + geom_bar() + 
+  ylab("Number of (weighted) responses") + 
+  facet_grid(Correct~.) +
+  ggtitle("Certainty by Correctness")
+
+```
+
+Using linear scores for the response of `Certainty`, with `not certain at all` assigned a score of 1 and `extremely certain` assigned a score of 5, we can estimate the effects of task and correctness on certainty  by using  a general normal linear model of the form:
+$$
+Y_{k} = \mu + \tau_{i(k)} + \gamma_{j(k)} + \epsilon_{k},
+$$
+where $k = 1, ..., N$, $\mu$ is average certainty (measured on a scale from 1 to 5), $\tau_{i(k)}$ is the task in the $k$th response (with $i(k) = 1$ for unaligned and $i(k)=2$ for aligned comparisons) and $\gamma_{j(k)}$ is the correctness of the $k$th response (with $j(k) = 1$ for a correct and $j(k)=2$ for a wrong answer).
+We also assume that errors are normally distributed, i.e.
+$\epsilon_k \stackrel{i.i.d}{\sim} N(0, \sigma^2)$ for all $k = 1, ..., N$.
+We use a constraint of first effects as zero to make effects estimable, i.e.
+$\widehat{\mu}$ has to be interpreted as the average certainty of participants who are wrong on the unaligned comparison.
+
+```{r model-certainty}
+model_certain <- svyglm(as.numeric(Response)~Question+Correct, design = survey_12_crt)
+
+coefs_certainty <- broom::tidy(model_certain)
+
+coefs_certainty$term <- c("$$\\widehat{\\mu}$$", "$$\\widehat{\\tau}_2$$", "$$\\widehat{\\gamma}_2$$")
+
+coefs_certainty %>% 
+  mutate(
+    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value))
+  ) %>%
+  gt() %>%
+  tab_header("Estimates for Model 2.3",
+             " certainty of responses by task and correctness of response") %>%
+  cols_label(
+    std.error = "SE",
+    statistic = "t-statistic",
+    p.value = "p-value"
+  ) %>%
+  fmt_number(
+    columns = c(estimate, std.error),
+    decimals = 2
+  ) %>%
+  fmt_number(
+    columns = statistic,
+    decimals = 1
+  ) %>%
+  cols_align(
+    columns = p.value,
+    align = "right"
+  )
+```
 

### 2023-07-05 analysis/analysis.qmd
certainty

@@
     Question = ifelse(Question=="ab_certain", "A vs B (Unaligned)", "C vs D (Aligned)"),
-# ordering of levels in Response is from extremely=1 to not at all=5 - not very intuitive
-    Response = factor(Response, levels = rev(levels(Response)))
+# ordering of levels in Response is from extremely=6 to not at all=2 - not very intuitive
+    Response = factor(Response, levels = c("not at all", "slightly", "moderately", "very", "extremely")) # gets rid of the 'SKIPPED ON WEB' and makes the ordering go from 1 to 5
   )
@@
 
-Using linear scores for the response of `Certainty`, with `not certain at all` assigned a score of 1 and `extremely certain` assigned a score of 5, we can estimate the effects of task and correctness on certainty  by using  a general normal linear model of the form:
+Using linear scores for the response of `Certainty`, with `not certain at all` assigned a score of 1 and `extremely certain` assigned a score of 5, we can estimate the effects of task and correctness on certainty  by using  a cell-means model of the form:
 $$
-Y_{k} = \mu + \tau_{i(k)} + \gamma_{j(k)} + \epsilon_{k},
+Y_{k} = \mu_{ij(k)} + \epsilon_{k},
 $$
-where $k = 1, ..., N$, $\mu$ is average certainty (measured on a scale from 1 to 5), $\tau_{i(k)}$ is the task in the $k$th response (with $i(k) = 1$ for unaligned and $i(k)=2$ for aligned comparisons) and $\gamma_{j(k)}$ is the correctness of the $k$th response (with $j(k) = 1$ for a correct and $j(k)=2$ for a wrong answer).
+where $k = 1, ..., N$, $\mu_{ij(k)}$ is average certainty (measured on a scale from 1 to 5) of the four combinations of task and correctness, where $i = 1, 2$ encodes unaligned/aligned, and $j=1,2$ encodes wrong, correct, respectively.
 We also assume that errors are normally distributed, i.e.
 $\epsilon_k \stackrel{i.i.d}{\sim} N(0, \sigma^2)$ for all $k = 1, ..., N$.
-We use a constraint of first effects as zero to make effects estimable, i.e.
-$\widehat{\mu}$ has to be interpreted as the average certainty of participants who are wrong on the unaligned comparison.
+What we find, is that we see the lowest score of certainty for correct responses to the unaligned comparison task.
+For aligned comparisons, the correctness of the response does not matter for the associated certainty.
 
 ```{r model-certainty}
-model_certain <- svyglm(as.numeric(Response)~Question+Correct, design = survey_12_crt)
+model_certain <- svyglm(as.numeric(Response)~Question:Correct-1, design = survey_12_crt)
 
 coefs_certainty <- broom::tidy(model_certain)
+terms <- coefs_certainty$term
 
-coefs_certainty$term <- c("$$\\widehat{\\mu}$$", "$$\\widehat{\\tau}_2$$", "$$\\widehat{\\gamma}_2$$")
+coefs_certainty$term <- c("$$\\widehat{\\mu}_{11}$$", "$$\\widehat{\\mu}_{21}$$", "$$\\widehat{\\mu}_{12}$$", "$$\\widehat{\\mu}_{22}$$")
 
@@
     align = "right"
+  )
+```
+
+```{r}
+#| out-width: 80%
+#| fig-cap: "Certainty by task and correctness.
+Correct answers in the unaligned task have the least certainty associated with them.
+An incoorect response on the unaligned task shows a significant oost in certainty.
+"
+round_12_certainty_long %>% 
+  ggplot(aes(x = Response, weight=weight)) + geom_bar() + 
+  ylab("Number of (weighted) responses") + 
+  facet_grid(Correct~Question) +
+  geom_vline(
+    aes(xintercept = estimate),
+    colour = "darkorange",
+    data = data.frame(
+      coefs_certainty, 
+      Correct= gsub(".*(FALSE|TRUE)", "\\1", terms),
+      Question = gsub("Question(.*):Correct.*", "\\1", terms))
   )

### 2023-07-05 analysis/analysis.qmd
references to figures and tables work now

@@
 date: last-modified
-bibliography: "`r rbbt::bbt_write_bib('references.bib', overwrite = TRUE, library_id=rbbt::bbt_library_id('Graphics Research'), translator='bibtex')`"
+bibliography: references.bib 
 ---
 
+```{r run-to-update-bibfile, eval=TRUE, include=FALSE}
+keys <- rbbt::bbt_detect_citations("analysis.qmd")
+citations_only <- grep(pattern="[fig|tbl]-.*", keys, value=TRUE, invert=TRUE)
+rbbt::bbt_write_bib(keys=citations_only, 'references.bib', overwrite = TRUE,
+                    library_id=rbbt::bbt_library_id('Graphics Research'),
+                    translator='bibtex')
+```
 ```{r setup, include = FALSE}
@@
 The data used for this is a combination of rounds 1 and 2, with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`.
-Figure \ref{fig-tasks-12} shows the two stacked barcharts shown to all panelists.
+@fig-tasks-12 shows the two stacked barcharts shown to all panelists.
 
@@
 
-Defining an accurate response as "B is bigger" in the chart of unaligned tiles, and "D is bigger" in the aligned case, while calling the other two responses as incorrect, leads to the data shown in figure \ref{fig-alignment}(a): we see
+Defining an accurate response as "B is bigger" in the chart of unaligned tiles, and "D is bigger" in the aligned case, while calling the other two responses as incorrect, leads to the data shown in @fig-alignment(a): we see
  that more than twice the number of responses is accurate, when the tiles are aligned along the same axis. 
@@
 
-When moving beyond the binary accuracy measure of the response and using all three levels of the response as dependent variable, we see that by far the largest change in responses comes from a change from 'they are the same' to the (perceptually) correct response of 'D is bigger' when the tiles are aligned along the same axis for a comparison (see \ref{fig-ordinal}).
+When moving beyond the binary accuracy measure of the response and using all three levels of the response as dependent variable, we see that by far the largest change in responses comes from a change from 'they are the same' to the (perceptually) correct response of 'D is bigger' when the tiles are aligned along the same axis for a comparison (see @fig-ordinal).
 Even though the number of wrong responses (orange) is only slightly different (`r round(paired2$estimate,3)`) between aligned and unaligned comparisons, it is still highly significant ($t$ statistic: `r round(paired2$statistic,1)`, df: `r paired2$parameter`, $p$-value: < 2.2e-16).
@@
 #| label: fig-tasks-67
-#| fig-cap: "Stimuli shown to panelists. For each of the charts, all marked pieces were evaluated against each other. "
+#| fig-cap: "Stimuli shown to panelists. For each of the charts, all marked pieces were evaluated pairwise for their size difference. "
 #| fig-height: 3
@@
 
-Figure \ref{fig-response-67} gives an overview of all responses to size assessments of the marked tiles.
+### Binary and Ordinal Correctness as Response
+
+@fig-response-67 gives an overview of all responses to size assessments of the marked tiles.
 The panels are ordered according to the difference in signal and hypothesized difficulty from easiest (left) to hardest (right) comparison.
@@
 Design does not seem to matter for easy comparisons (BD, BC) or the hard comparison (CD). 
-The three comparisons in the middle show interesting patterns, highlighted in the reordered figure \ref{fig-response-67b}.
+The three comparisons in the middle show interesting patterns, highlighted in the reordered @fig-response-67b.
 
@@
 
+### How certain are participants?
+
+```{r data-certainty-pie}
+
+```
+
 ## Model X: Comparing Aligned and Unaligned Elements Facets {#sec-modelX}
 
-The survey setup and all stimuli are shown in \ref{fig-tasks-7}.
+The survey setup and all stimuli are shown in @tbl-tasks-7.
 A total of `r nrow(round7_adj)` panelists  (Effective sample size: `r round(neff(round7_adj$weight), 1)`) were shown four sets of facetted charts each.
@@
 
-Figure \ref{fig-pies} shows a summary of panelists' accuracy of their responses.
+@fig-pies shows a summary of panelists' accuracy of their responses.
 The panels are sorted according to the level of difficulty of a comparison as hypothesized above from lowest difficulty (BD) to highest (CD).
 We see that the percentage of correct responses follows (on average) this hypothesized order.
-Using the comparison of A versus B as the baseline, we see from Table \ref{tab-model7}, that comparisons BC and BD have a significantly higher accuracy (in that order), while the accuracy of comparisons AD, AC, and CD are significantly worse (HH: should we do pairwise comparisons for that?).
+Using the comparison of A versus B as the baseline, we see from @tbl-model7, that comparisons BC and BD have a significantly higher accuracy (in that order), while the accuracy of comparisons AD, AC, and CD are significantly worse (HH: should we do pairwise comparisons for that?).
 Comparisons involving aligned pieces have (with two exceptions) a higher rate of accuracy than a  comparison of the same-size, unaliged pieces.  

### 2023-07-05 analysis/analysis.qmd
no need

@@
   jasa-pdf:
+    include-in-header:
+      - edits.tex
+      - authors.tex
     keep-tex: true  

### 2023-07-07 analysis/analysis.qmd
update references

@@
 
-```{r run-to-update-bibfile, eval=TRUE, include=FALSE}
+```{r run-to-update-bibfile, eval=FALSE, include=FALSE}
 keys <- rbbt::bbt_detect_citations("analysis.qmd")
@@
 ```
+
 ```{r setup, include = FALSE}
@@
 response_67 <- rbind(
-    response_6 %>% select(CaseId, labelled, task, responses, weight),
-    response_7 %>% select(CaseId, labelled, task, responses, weight)
+    response_6 %>% dplyr::select(CaseId, labelled, task, responses, weight),
+    response_7 %>% dplyr::select(CaseId, labelled, task, responses, weight)
   ) %>% mutate(
@@
              ifelse(labelled=="BC", 3,
-               ifelse(labelled %in% c("AB", "AD"), 2, 1)))
+               ifelse(labelled %in% c("AB", "AD"), 2, 1))),
+  task_labelled = factor(paste(task, labelled, sep="_"))
 )
@@
 The three comparisons in the middle show interesting patterns, highlighted in the reordered @fig-response-67b.
+Colored letters below and above the bars encode significances between pairs of bars of the same color: two proportions are significantly different at a 5% significance level, if they do not have a letter in common [@piephoAlgorithmLetterBasedRepresentation2004].
+Significances are adjusted for multiple comparisons as implemented in the `multcomp` package [@hothornSimultaneousInferenceGeneral2008].
+
+```{r model-67, warning = FALSE}
+# same model as for plot 
+logit_67_correct <- svyglm(correct~task_labelled-1, family=quasibinomial(), design=survey_67)
+
+logit_67_correct3 <- svyglm(correct3~task_labelled-1, family=quasibinomial(), design=survey_67)
+
+ci_confint <- svyby(~correct, by=~task_labelled, design=survey_67, svyciprop)
+ci_confint3 <- svyby(~correct3, by=~task_labelled, design=survey_67, svyciprop)
+
+coefs_67 <- broom::tidy(logit_67_correct)
+coefs_67_relaxed <- broom::tidy(logit_67_correct3)
+
+# create all pairwise comparisons for tasks and questions.
+cross_compare <- glht(logit_67_correct, mcp(task_labelled="Tukey"))
+cross_compare3 <- glht(logit_67_correct3, mcp(task_labelled="Tukey"))
+
+# get letters for each confidence 
+ci_confint$letters <- cld(cross_compare)$mcletters$Letters
+ci_confint3$letters <- cld(cross_compare3)$mcletters$Letters
+
+ci_confint <- ci_confint %>% 
+separate_wider_delim(
+  task_labelled, delim = "_", names =c("task", "labelled")
+)  %>% mutate(
+  task = factor(task, levels=c("Bar", "Pie\nA-", "Pie\nUn-")),
+  labelled = factor(labelled, levels =c("BD", "BC", "AB", "AD", "AC", "CD"))
+)
+
+ci_confint3 <- ci_confint3 %>% 
+separate_wider_delim(
+  task_labelled, delim = "_", names =c("task", "labelled")
+)  %>% mutate(
+  task = factor(task, levels=c("Bar", "Pie\nA-", "Pie\nUn-")),
+  labelled = factor(labelled, levels =c("BD", "BC", "AB", "AD", "AC", "CD"))
+)
+```
 
@@
 The increasing difficulty of comparisons is reflected in the overall decreasing percentage of correct responses (height of blue tiles).
-The overlaid rectangles represent 95% confidence intervals."
+The overlaid rectangles represent 95% confidence intervals.
+The letters in blue and orange encode significances between pairwise proportions: two bars have a significantly different proportion (at a 5% significance level) if they do not share a letter."
 #| fig-height: 4.25
@@
   ) +
-  scale_y_continuous("Percent responses", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25))
+  scale_y_continuous("Percent responses", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25), limits = c(-0.1, 1.1)) + 
+  geom_text(aes(label=letters), y = -0.05, data = ci_confint, colour = "steelblue") +
+  geom_text(aes(label=letters), y = 1.05, data = ci_confint3, colour = "darkorange") 
 ```
@@
   ) +
-  scale_y_continuous("Percent responses", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25))
-```
-
-```{r model-67}
-# still putzing around with it
-logit_67_correct <- svyglm(correct~task+labelled + task, family=quasibinomial(), design=survey_67)
-
-logit_67_cont <- svyglm(correct~task*signal, family=quasibinomial(), design=survey_67)
-
+  scale_y_continuous("Percent responses", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25), limits = c(-0.1, 1.1)) + 
+  geom_text(aes(label=letters), y = -0.05, data = ci_confint, colour = "steelblue") +
+  geom_text(aes(label=letters), y = 1.05, data = ci_confint3, colour = "darkorange") 
+```
+
+```{r estimates, eval=FALSE}
+coefs_67 %>% 
+  mutate(
+    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value))) %>%
+  gt() %>%
+  tab_header("Estimates for Model 3.2",
+             " accuracy of responses across pairwise evalulations") %>%
+  cols_label(
+    estimate = "log odds",
+    std.error = "SE",
+    statistic = "t-statistic",
+    p.value = "p-value"
+  ) %>%
+  fmt_number(
+    columns = c(estimate, std.error),
+    decimals = 2
+  ) %>%
+  fmt_number(
+    columns = statistic,
+    decimals = 1
+  ) %>%
+  cols_align(
+    columns = p.value,
+    align = "right"
+  )
 ```

### 2023-07-07 analysis/analysis.qmd
timing

@@
     labelled = c("AB", "AC", "AD", "BD", "BC", "CD")[question],
+    time_spent = ifelse(parse_number(task) == 1, NOLA1_TOTALTIME, NOLA2_TOTALTIME),
     task = c("Pie\nUn-", "Pie\nA-")[parse_number(task)]
@@
 response_67 <- rbind(
-    response_6 %>% dplyr::select(CaseId, labelled, task, responses, weight),
-    response_7 %>% dplyr::select(CaseId, labelled, task, responses, weight)
+    response_6 %>% dplyr::select(CaseId, labelled, task, responses, 
+                                 time_spent = NOLA1_TOTALTIME, weight),
+    response_7 %>% dplyr::select(CaseId, labelled, task, responses, 
+                                 time_spent, weight)
   ) %>% mutate(
@@
   correct3 = correct | responses == "They are the same",
+  time_spent = parse_number(time_spent),
   response3 = factor(correct+correct3, labels = c("Wrong", "Same", "Correct")),
@@
 
-### How certain are participants?
-
-```{r data-certainty-pie}
-
+### How much time did participants spend on each task?
+
+Note: we do not have certainty for round 7. We also do not have timing by individual comparison.
+
+```{r vis-time-pie}
+#| label: fig-time-67
+#| fig-cap: "Comparisons for aligned pie slices were on average the fastest.
+Comparisons in the facetted barcharts took the longest.
+The difference in average times are significant for aligned pies versus either of the other designs, but not different between barcharts and unaligned pie charts."
+response_67 %>%
+  mutate(task = reorder(task, time_spent)) %>%
+  ggplot(aes(y = task, x = time_spent)) + 
+  geom_boxplot() +
+  scale_x_log10(breaks=c(1, 5, 10, 20, 30, 60, 120, 300, 600), labels = c("1 sec", "5 secs", "10", "20", "30", "1 min", "2", "5", "10 mins")) + 
+  ylab("") + xlab("") + 
+  ggtitle("Time spent on comparing all six pairs by design") 
+```
+
+```{r model-time}
+time_mod <- svyglm(time_spent~task, design = survey_67)
+summary(time_mod)
 ```

### 2023-07-07 analysis/analysis.qmd
(1) area of wedges vs tiles (2) effective number of responses for rounds 6 and 7

@@
 
-neff1 <- n1/d1
+neff1 <- n1/d1 # same as neff(round1_adj$weight)
 neff2 <- n2/d2
@@
 
-lambda <- n1/(d1*denom) 
+lambda <- n1/(d1*denom) # same as calculate_lambda(round1_adj$weight, round2_adj$weight)
 ```
@@
 Correct answers in the unaligned task have the least certainty associated with them.
-An incoorect response on the unaligned task shows a significant oost in certainty.
+An incoorect response on the unaligned task shows a significant boost in certainty.
 "
@@
 
-A total of three different charts were shown to (a part) of the panelists from Round 6 and all panelists of Round 7. 
-The four labelled pieces A, B, C, and D correspond to values 46, 47.6, 45.2, and 44.4, respectively, i.e underlying each evaluation is the order D < C < A < B.
-
-```{r}
-#| label: fig-tasks-67
-#| fig-cap: "Stimuli shown to panelists. For each of the charts, all marked pieces were evaluated pairwise for their size difference. "
-#| fig-height: 3
-#| fig-width: 4
-#| layout-nrow: 1
-#| fig-subcap:
-#| - "**Round 6**: Facetted barcharts"
-#| - "**Round 7**: Facetted piecharts with unaligned wedges"
-#| - "**Round 7**: Facetted piecharts with aligned wedges"
-knitr::include_graphics("images/Round 6 - Sept 2022.png")
-knitr::include_graphics("images/round7/unaligned pie - all.png")
-knitr::include_graphics("images/round7/aligned pie - all.png")
-```
+The data this section is based on comes from (a part) of Round 6 and all responses from Round 7.
 
@@
 ```
+
+A total of three different charts were shown to (a part) of the panelists from Round 6 and all panelists of Round 7. 
+Each of the charts has four labelled pieces A, B, C, and D.
+
+```{r}
+#| label: fig-tasks-67
+#| fig-cap: "Stimuli shown to panelists. For each of the charts, all marked pieces were evaluated pairwise for their size difference. "
+#| fig-height: 3
+#| fig-width: 4
+#| layout-nrow: 1
+#| fig-subcap:
+#| - "**Round 6**: Facetted barcharts"
+#| - "**Round 7**: Facetted piecharts with unaligned wedges"
+#| - "**Round 7**: Facetted piecharts with aligned wedges"
+knitr::include_graphics("images/Round 6 - Sept 2022.png")
+knitr::include_graphics("images/round7/unaligned pie - all.png")
+knitr::include_graphics("images/round7/aligned pie - all.png")
+```
+
+The questions asked for each of these charts were of the form: "which of the marked pieces is bigger? Just your best guess is fine".
+
+A.	A or B?
+
+RESPONSE OPTIONS:
+
+1.	A is bigger
+2.	B is bigger
+3.	They are the same
+
+Each participant was asked a total of six comparisons of this type: A vs B, A vs C, A vs D, B vs C, B vs C, and C vs D.
+The number of valid responses and the resulting effective sample sizes are shown in @tbl-combine-67
+
+```{r tbl-combine-67}
+#| label: tbl-combine-67
+#| tbl-cap: Table of the number of effective responses by task 
+dframe <- data.frame(
+  round = c("Round 6", "Round 7", "Round 7"),
+  task = c("Bar", "Pie-Aligned", "Pie-Unaligned"),
+  responses = c(nrow(response_6), nrow(response_7 %>% filter(task=="Pie\nA-")), nrow(response_7 %>% filter(task=="Pie\nUn-"))),
+  effective = c(neff(response_6$weight), neff((response_7 %>% filter(task=="Pie\nA-"))$weight),
+                neff((response_7 %>% filter(task=="Pie\nUn-"))$weight)),
+  participants = c(length(unique(response_6$CaseId)),
+                   nrow(unique(response_7 %>% filter(task=="Pie\nA-") %>% dplyr::select(CaseId))),
+                   nrow(unique(response_7 %>% filter(task=="Pie\nUn-") %>% dplyr::select(CaseId)))),
+  lambda = c(calculate_lambda(response_6$weight, response_7$weight), 
+             calculate_lambda((response_7 %>% filter(task=="Pie\nA-"))$weight, 
+                               c(response_6$weight,(response_7 %>% filter(task=="Pie\nUn-"))$weight)),
+             calculate_lambda((response_7 %>% filter(task=="Pie\nUn-"))$weight, 
+                               c(response_6$weight,(response_7 %>% filter(task=="Pie\nA-"))$weight)))
+)
+
+dframe %>% 
+  group_by(round) %>%
+  gt() %>%
+  tab_header("Number of (effective) responses by task") %>%
+  cols_label(
+    responses = md("number of responses:<br>nominal"),
+    effective = md("number of responses:<br>effective"),
+    participants = md("number of participants"),
+    lambda = md("λ"),
+  ) %>%
+  fmt_number(
+    everything(), 
+    decimals = 0,
+    use_seps=TRUE
+  ) %>%
+  fmt_number(
+    columns = effective,
+    decimals = 1
+  ) %>%
+  fmt_number(
+    columns = c(lambda),
+    decimals = 4
+  ) %>%
+  tab_options(
+    column_labels.font.weight = "bold",
+    row_group.font.weight = "bold"
+  )
+```
+
+The four labelled pieces A, B, C, and D correspond to values 46, 47.6, 45.2, and 44.4, respectively, i.e underlying each evaluation is the order D < C < A < B.
+Nominally, the differences between the size of the marked elements can be expressed as multiples in $\delta = 0.8$, with $B = A + 2\delta = C + 3 \delta = D + 4 \delta$.
+A difference of $\delta = 0.8$ corresponds to about a $3^\degree$ difference for the angle of a wedge.
+We would expect that an increase in the difference in the values increases the accuracy in the responses.
+This is mitigated by the distance between wedges: more distance (should) lead to a decrease in accuracy.
+Similarly, if pieces have an axis or a line of reference in common, we would expect an increase in accuracy.
 
@@
 
+Besides intrinsic reference lines, the overall size of the stimulus might be another factor contributing to the difference in accuracy between the tiles and the wedges.
+In the shown example, the marked tiles have an  area between $133 \times 55 = 7315$  pixels (for D) to  $144 \times 55 = 7920$ pixels (for B), while the pie wedges  are based on circles with a diameter of 105 pixels, and therefore cover visually more than twice the area of the tiles in the barchart.
+The area of D in form of a pie-wedge is $\frac{44.4}{100} \times 105^2 \pi  \approx$ 15,378 pixels and B has an area of $\frac{47.6}{100} \times 105^2 \pi  \approx$ 16,487 pixels.
+
+There are different possible strategies for evaluating the size of a wedge in a pie chart.
+Panelists can use the **angle** of the wedge (or a suitable derivative of it, such as the complementary angle or the remainder to 180 degrees), its **area** (as part of the circle or some derivative) or the **arc length** of the wedge.
+Studies by  @kosaraJudgmentErrorPie2016 and @kosaraEvidenceAreaPrimary2019 show that the area of a piece in a pie chart is the primary evaluation method.
+
 ### How much time did participants spend on each task?
 
-Note: we do not have certainty for round 7. We also do not have timing by individual comparison.
+Note: we do not have certainty for round 7. We also do not have timing by individual comparisons.
 
@@
 For a comparison of wedges marked X and Y, there were three possible options for the response: (1) X is bigger, (2) Y is bigger or (3) They are the same.
-
-The four labelled pieces A, B, C, and D correspond to values 46, 47.6, 45.2, and 44.4, respectively, i.e underlying each evaluation is the order D < C < A < B.
-Nominally, the differences between the size of the marked elements can be expressed as multiples in $\delta = 0.8$, with $B = A + 2\delta = C + 3 \delta = D + 4 \delta$.
-A difference of $\delta = 0.8$ corresponds to about a $3^\degree$ difference for the angle of a wedge.
-We would expect that an increase in the difference in the values increases the accuracy in the responses.
-This is mitigated by the distance between wedges: more distance (should) lead to a decrease in accuracy.
-Similarly, if pieces have an axis or a line of reference in common, we would expect an increase in accuracy.
-
-There are different possible strategies for evaluating the size of a wedge in a pie chart.
-Panelists can use the **angle** of the wedge (or a suitable derivative of it, such as the complementary angle or the remainder to 180 degrees), its **area** (as part of the circle or some derivative) or the **arc length** of the wedge.
-Studies by  @kosaraJudgmentErrorPie2016 and @kosaraEvidenceAreaPrimary2019 show that the area of a piece in a pie chart is the primary evaluation method.
 

### 2023-07-10 SDSS-2023/testing-charts-accuracy.qmd
(1) area of wedges vs tiles (2) effective number of responses for rounds 6 and 7
changed template for JDS ... not quite there yet but getting closer

@@
 title: "Testing Charts: viewer's perceptual accuracy in surveys"
-bibliography: "`r rbbt::bbt_write_bib('references.bib', overwrite = TRUE, library_id=rbbt::bbt_library_id('Graphics Research'), translator='bibtex')`"
+short-title: Testing charts in surveys # running title
+author:
+  - first: Kiegan                 # first name
+    first-init: K.              # initial of first name
+    last: Rice                  # last name
+    marker: '1,2'               # identifier for the affiliation
+    footnote-id: 1              # optional identifier for the footnote
+    email: foo@bar.com          # optional email address
+  - first: Brett
+    first-init: B.
+    last: Berger
+    marker: 2
+    footnote-id: 2
+  - first: Carl
+    first-init: C.
+    last: Camp
+    marker: 3
+  - first: David
+    first-init: D.
+    last: Dodge
+    marker: 3
+footnote:
+  - content: Corresponding author
+    type: corresp               # set type for the corresponding author
+    id: 1
+  - content: Another footnote.
+    id: 2
+affiliation:
+  - institution: University of Achievement
+    prefix: Department A        # optional
+    country: Country A
+    marker: 1
+  - institution: Institution B
+    country: Country B
+    marker: 2
+  - institution: Institution C
+    country: Country C
+    marker: 3
+bibliography: references.bib
 abstract: "The use of visuals is a key component in scientific communication, and decisions about the design of a data visualization should be informed by what design elements best support the audience's ability to perceive and understand the components of the data visualization.
@@
   pdf:
+    template: template-jdsart.cls.tex
     include-in-header: 
-      - edits.tex
-      - authors.tex
-      # - refs.tex
-    keep-tex: true
-    # template: template.tex
-documentclass: jds
+      - text: |
+          \usepackage{amsfonts,amsmath,amssymb,amsthm}
+          \usepackage{booktabs}
+          \usepackage{graphicx}
+          \usepackage{longtable}
+          \usepackage{hyperref}
+      - file: edits.tex
+    cite-method: natbib
 ---
+
+<!--
+\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
+\setlength{\parindent}{0pt}
+\usepackage{amsfonts,amsmath,amssymb,amsthm}
+  \usepackage{booktabs}
+  \usepackage{lipsum}
+  \usepackage{subfig}
+  \usepackage{graphicx}
+  \usepackage{longtable}
+\hh{hello world!}
+-->
+
+```{r run-to-update-bibfile, eval=FALSE, include=FALSE}
+keys <- rbbt::bbt_detect_citations("testing-charts-accuracy.Rmd")
+citations_only <- grep(pattern="[fig|tbl|eq][-|:].*", keys, value=TRUE, invert=TRUE)
+rbbt::bbt_write_bib(keys=citations_only, 'references.bib', overwrite = TRUE,
+                    library_id=rbbt::bbt_library_id('Graphics Research'),
+                    translator='bibtex')
+```
 
@@
 
-\hh{This is Heike's color for making changes}
 ## Introduction
@@
 
-|                                   |                                   |
-|-----------------------------------|-----------------------------------|
-| **Unaligned positions**           | **Aligned positions**             |
-| ![Unaligned positions marked A and B](../SDSS-2023/images/vertical-ab.PNG) | ![Aligned positions marked C and D](../SDSS-2023/images/vertical-cd.PNG) |
+```{r}
+#| label: tbl-tasks-12
+#| fig-cap: "The only difference between the two pairs of rectangles A, B and C, D is their alignment, i.e.
+A and C are of identical size, as are B and D.
+While the predominant response for the unaligned pair on the left is 'they are of the same size', more than half of the viewers respond with 'D is bigger' to the aligned pair of bars on the right."
+#| tbl-subcap:
+#| - "**Unaligned bars**"
+#| - "**Aligned bars**"
+#| layout: [[45,45], [100]]
+#| fig-align: "center"
+#| out-width: 100%
+knitr::include_graphics("../SDSS-2023/images/vertical-ab.PNG")
+knitr::include_graphics("../SDSS-2023/images/vertical-cd.PNG")
+```
 
-: The two stacked barcharts every participant got to see.
+<!-- |                                   |                                   | -->
+<!-- |-----------------------------------|-----------------------------------| -->
+<!-- | **Unaligned positions**           | **Aligned positions**             | -->
+<!-- | ![Unaligned positions marked A and B](../SDSS-2023/images/vertical-ab.PNG) | ![Aligned positions marked C and D](../SDSS-2023/images/vertical-cd.PNG) | -->
+
+<!-- : The two stacked barcharts every participant got to see.
 In each barchart, the two marked tiles are to be compared for their size.
 in both instances, the tile on the right is (very slightly) larger.
-{#tbl-tasks-12 tbl-colwidths="\[50,50\]"}
+{#tbl-tasks-12 tbl-colwidths="\[50,50\]"} -->
 
@@
 Figure \ref{fig-alignment}(a) shows that more than twice the number of responses is accurate, when the tiles are aligned along the same axis.
-Because each particpant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts.
+Because each participant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts.
 The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: < 2.2e-16).

### 2023-07-12 SDSS-2023/testing-charts-accuracy.qmd
still hammering on the format

@@
     cite-method: natbib
+    keep-tex: true
 ---
@@
 
-<!-- Article Types -->
-<!-- https://jds-online.org/journal/JDS/information/Article%20Types -->
-
-<!-- Statistical Data Science -->
-
-<!-- This section is the home base of the reformed journal covering statistical methods that are motivated by real-world applications.
-It is not for papers with technical proofs that push the frontiers of theoretical developments.
-In addition to classic topics in Statistics, cutting-edge works on big data, visualization, machine learning, and artificial intelligence are also welcome.
--->
-
-<!-- Computing in Data Science -->
-
-<!-- Computing is an indispensable component of all data science and big data applications.
-As more journals on data science and big data emerge, it is of great interest for the data science community to have a highly regarded outlet with a specialization in computing, covering a wide spectrum from methods, algorithms, software implementations, to case studies.
-Existing journals on statistical computing have their own traditions and may not meet the increasing demands.
-Some research works on cutting-edge problems may not fit well in any existing journal.
--->
-
-<!-- This section covers the following types of articles. -->
-<!-- 1.
-Software: Articles here are similar to those in the Journal of Statistical Software.
-They are not referencing manuals but vignettes that introduce the methods being implemented as well as the usage of the software with reproducible code chunks.
-The software implementation can be in any computer language with a sufficiently large user base.
--->
-<!-- 2.
-Algorithms: Articles here focus on the performance side of the computing needs arising from domain applications.
-For example, one can propose algorithms that make infeasible tasks feasible or speed up existing algorithms.
--->
-<!-- 3.
-Methods: Articles here are similar to those in Journal of Computational and Graphic Statistics or Statistics and Computing.
-The computing methods need to be motivated by a domain application with the properties carefully studied.
--->
-
 ```{r setup, include = FALSE}
@@
                       out.width = "50%",
-                      dpi = 72, # we can set the resolution up to 300 later.
-                      cache = TRUE) 
+                      fig.height = 5,
+                      fig.width = 5,
+                      dpi = 72#, # we can set the resolution up to 300 later.
+                      #cache = TRUE) 
+)
 library(readr)
@@
 
-```{r}
+```{r out.width='\\textwidth'}
 #| label: fig-aligned-unaligned
@@
 #| - "**Aligned bars**"
-#| fig-cap-location: "bottom"
-#| layout: [[45,45], [100]]
+#| layout-nrow: 1
 #| fig-align: "center"
-#| out-width: 100%
-knitr::include_graphics("images/round5-modified.PNG")
-knitr::include_graphics("images/round3-modified.PNG")
+knitr::include_graphics(c("images/round5-modified.PNG", "images/round3-modified.PNG"))
 ```
@@
 
-\autoref{tbl-tasks-12} shows the two stacked barcharts shown to participants in the survey.
+\autoref{fig-tasks-12} shows the two stacked barcharts shown to participants in the survey.
 The marked tiles in each plot are 155 pixels apart.
@@
 <!-- describe stimulus -->
-Both charts in \autoref{tbl-tasks-12} show the same data with slight modifications to the order of the levels – the first and second level in each of the bars are reversed between the left and the right chart.
+Both charts in \autoref{fig-tasks-12} show the same data with slight modifications to the order of the levels – the first and second level in each of the bars are reversed between the left and the right chart.
 Participants were asked to compare the relative sizes of the tiles marked A and B (C and D, respectively) and select the correct response out of the possible choices:
@@
 ```{r}
-#| label: tbl-tasks-12
-#| fig-cap: "The only difference between the two pairs of rectangles A, B and C, D is their alignment, i.e.
-A and C are of identical size, as are B and D.
-While the predominant response for the unaligned pair on the left is 'they are of the same size', more than half of the viewers respond with 'D is bigger' to the aligned pair of bars on the right."
-#| tbl-subcap:
+#| label: fig-tasks-12
+#| fig-cap: "The two stacked barcharts every participant got to see.
+In each barchart, the two marked tiles are to be compared for their size.
+In both instances, the tile on the right is (very slightly) larger."
+#| fig-subcap:
 #| - "**Unaligned bars**"
 #| - "**Aligned bars**"
-#| layout: [[45,45], [100]]
+#| layout-nrow: 1
 #| fig-align: "center"
 #| out-width: 100%
+#| fig-asp: 1
 knitr::include_graphics("../SDSS-2023/images/vertical-ab.PNG")
@@
 
-```{r accuracy, fig.height = 5, fig.width = 10, warning=FALSE}
+```{r accuracy,   warning=FALSE}
 #| fig-cap: "On the left (a), a stacked barchart shows the number of respondents with correct (green) and incorrect (grey) responses to the two comparison questions.
@@
 #| out-width: 80%
-#| fig-env: figure
+#| fig-height: 5 
+#| fig-width: 10
 

### 2023-07-12 SDSS-2023/testing-charts-accuracy.qmd
default to html

@@
     last: Rice                  # last name
-    marker: '1,2'               # identifier for the affiliation
+    marker: 1               # identifier for the affiliation
     footnote-id: 1              # optional identifier for the footnote
     email: foo@bar.com          # optional email address
-  - first: Brett
-    first-init: B.
-    last: Berger
+  - first: Heike
+    first-init: H.
+    last: Hofmann
     marker: 2
-    footnote-id: 2
-  - first: Carl
-    first-init: C.
-    last: Camp
-    marker: 3
-  - first: David
-    first-init: D.
-    last: Dodge
-    marker: 3
+  - first: Nola
+    first-init: N.
+    last: duToit
+    marker: 1
+  - first: Edward
+    first-init: E.
+    last: Mulrow
+    marker: 1
 footnote:
@@
     id: 1
-  - content: Another footnote.
-    id: 2
 affiliation:
-  - institution: University of Achievement
-    prefix: Department A        # optional
-    country: Country A
+  - institution: National Opinion Research Center (NORC)
+    country: United States
     marker: 1
-  - institution: Institution B
-    country: Country B
+  - institution: Iowa State University
+    department: Statistics
+    country: United States
     marker: 2
-  - institution: Institution C
-    country: Country C
-    marker: 3
 bibliography: references.bib
@@
 format:
+  html: default
   pdf:
@@
     keep-tex: true
+    keep-md: true
 ---
+
+\hh{hello?}
+
+\kr{Kiegan, this is your editing color :) - you can choose a different color in edits.tex}
 
@@
                       fig.align = "center",
-                      out.width = "50%",
+                      out.width = "100%",
                       fig.height = 5,
                       fig.width = 5,
-                      dpi = 72#, # we can set the resolution up to 300 later.
+                      dev = "png", dev.args = list(type = "cairo-png")
+#                      dpi = 72#, # we can set the resolution up to 300 later.
                       #cache = TRUE) 
 )
+
 library(readr)
@@
 #| fig-align: "center"
-knitr::include_graphics(c("images/round5-modified.PNG", "images/round3-modified.PNG"))
+knitr::include_graphics("images/round5-modified.PNG")
+knitr::include_graphics("images/round3-modified.PNG")
 ```
@@
 #| fig-asp: 1
-knitr::include_graphics("../SDSS-2023/images/vertical-ab.PNG")
-knitr::include_graphics("../SDSS-2023/images/vertical-cd.PNG")
+knitr::include_graphics("images/vertical-ab.PNG")
+knitr::include_graphics("images/vertical-cd.PNG")
 ```

### 2023-07-12 analysis/analysis.qmd
default to html

@@
   jasa-pdf:
-    include-in-header:
-      - edits.tex
-      - authors.tex
     keep-tex: true  
+    keep-md: true
     journal:
@@
                       dpi = 72, # we can set the resolution up to 300 later.
-                      cache = TRUE) 
+                      cache = FALSE) 
 library(readr)
@@
 library(gt) # grammar of tables
+library(multcomp)
 
@@
             mutate(weight = lambda*weight/sum(weight)) %>% 
-            select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
+            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
                    ab_certain, cd_certain, weight) %>%
@@
             mutate(weight = (1-lambda)*weight/sum(weight)) %>% 
-            select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, 
+            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, 
                    ab_certain, cd_certain, weight))
@@
 Nominally, the differences between the size of the marked elements can be expressed as multiples in $\delta = 0.8$, with $B = A + 2\delta = C + 3 \delta = D + 4 \delta$.
-A difference of $\delta = 0.8$ corresponds to about a $3^\degree$ difference for the angle of a wedge.
+A difference of $\delta = 0.8$ corresponds to about a  3 degree difference for the angle of a wedge.
 We would expect that an increase in the difference in the values increases the accuracy in the responses.
@@
 
-```{r model-67, warning = FALSE}
+```{r model-67, warning = FALSE, cache=TRUE}
 # same model as for plot 

### 2023-07-13 SDSS-2023/testing-charts-accuracy.qmd
jds format for now

@@
 format:
-  html: default
   pdf:
-    template: template-jdsart.cls.tex
+    documentclass: jds
+    fig-width: 6
+    fig-asp: 0.618
+    fig-align: center
     include-in-header: 
@@
                       fig.env="figure*", 
-                      fig.align = "center",
-                      out.width = "100%",
-                      fig.height = 5,
-                      fig.width = 5,
+                 #     fig.align = "center",
+                      out.width = "\\textwidth",
+                #      fig.height = 5,
+                 #     fig.width = 5,
                       dev = "png", dev.args = list(type = "cairo-png")
@@
 
-```{r out.width='\\textwidth'}
+```{r}
 #| label: fig-aligned-unaligned
@@
 #| fig-align: "center"
+#| out-width: 50%
 knitr::include_graphics("images/round5-modified.PNG")
@@
 #| layout-nrow: 1
-#| fig-align: "center"
 #| out-width: 100%
-#| fig-asp: 1
 knitr::include_graphics("images/vertical-ab.PNG")

### 2023-07-13 analysis/analysis.qmd
jds format for now

@@
   mutate(
-    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value))
+    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value)),
+        term = ifelse(term=="(Intercept)", "Intercept", term)
   ) %>%
@@
   mutate(
+    term = ifelse(term=="(Intercept)", "Intercept", term),
     p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value))) %>%
@@
 time_mod <- svyglm(time_spent~task, design = survey_67)
-summary(time_mod)
+coef_time <- broom::tidy(time_mod)
+
+coef_time %>% mutate(
+    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value)),
+    term = ifelse(term=="(Intercept)", "Intercept", term)
+  ) %>%
+  gt() %>%
+  tab_header("Estimates for Model 3.2",
+             " time spent (in seconds) on assessing the six pairs") %>%
+  cols_label(
+    std.error = "SE",
+    statistic = "t-statistic",
+    p.value = "p-value"
+  ) %>%
+  fmt_number(
+    columns = c(std.error),
+    decimals = 2
+  ) %>%
+  fmt_number(
+    columns = c(estimate, statistic),
+    decimals = 1
+  ) %>%
+  cols_align(
+    columns = p.value,
+    align = "right"
+  )
+
 ```

### 2023-07-13 analysis/analysis.qmd
labels for tables

@@
   jasa-pdf:
-    keep-tex: true  
-    keep-md: true
     journal:
@@
 ```{r model-ord}
+#| label: tbl-logodds
+#| tbl-cap: A change from aligned to unaligned tiles is detrimental to the level of correct (or correct and same) responses.
 
@@
 names(contrasts) <- c("level","log odds", "SE", "t-statistic", "p-value", "percent change")
-knitr::kable(contrasts, digits = c(NA, 1, 2, 1, 4, 1), row.names = FALSE, caption = "Table 2: A change from aligned to unaligned tiles is detrimental to the level of correct (or correct and same) responses.", label = "tab:logodds")
+knitr::kable(contrasts, digits = c(NA, 1, 2, 1, 4, 1), row.names = FALSE)
 ```
@@
       coefs_certainty, 
-      Correct= gsub(".*(FALSE|TRUE)", "\\1", terms),
+      Correct= as.logical(gsub(".*(FALSE|TRUE)", "\\1", terms)),
       Question = gsub("Question(.*):Correct.*", "\\1", terms))
@@
 ```{r}
+#| label: tbl-model7
+#| tbl-cap: Estimates for the cell means model
 logit_7 <- svyglm(correct ~ labelled + concept + taskname, family = quasibinomial(), design = survey_7)

### 2023-07-13 analysis/analysis.qmd
link to data

@@
 library(multcomp)
+library(here)
 
@@
 
-vs_repo_local <- "~/Documents/projects/NORC" # link to local version of the visual_studies repo
+vs_repo_local <- here::here("data") # link to the data folder (of the visual_studies repo)
 ```
@@
 # other github repo. 
-round1 <- readRDS(file.path(vs_repo_local, "visual-studies/data/round1.rds"))
-round2 <- readRDS(file.path(vs_repo_local, "visual-studies/data/round2.rds"))
+round1 <- readRDS(file.path(vs_repo_local, "round1.rds"))
+round2 <- readRDS(file.path(vs_repo_local, "round2.rds"))
 
@@
 # other github repo. 
-round6 <- readRDS(file.path(vs_repo_local, "visual-studies/data/round6.rds"))
+round6 <- readRDS(file.path(vs_repo_local, "round6.rds"))
 
@@
 # other github repo. 
-round7 <- readRDS(file.path(vs_repo_local, "visual-studies/data/round7.rds"))
+round7 <- readRDS(file.path(vs_repo_local, "round7.rds"))
 
@@
 The odds of an accurate answer with aligned bars is therefore  exp (`r round(coefs_12$estimate[2],1)`) =  `r round(exp(coefs_12$estimate[2]),1)` times higher than for the unaligned bars of Question 1.
+Estimates for this model fit are shown in @tbl-coefs-12.
 
 ```{r}
+#| label: tbl-coefs-12
+#| tbl-cap: Estimates for the logit model of accuracy by task
 coefs_12$term <- c("$$\\widehat{\\alpha}$$", "$$\\widehat{\\beta}_2$$")

### 2023-07-13 SDSS-2023/testing-charts-accuracy.qmd
link to data folder

@@
 library(gt) # grammar of tables
+library(here)
 
@@
 
-vs_repo_local <- "~/Documents/projects/NORC" # link to local version of the visual_studies repo
+vs_repo_local <- here::here("data") # link to data folder
 ```
@@
 # other github repo. 
-round1 <- readRDS("~/Documents/projects/NORC/visual-studies/data/round1.rds")
-round2 <- readRDS("~/Documents/projects/NORC/visual-studies/data/round2.rds")
+round1 <- readRDS(file.path(vs_repo_local, "round1.rds"))
+round2 <- readRDS(file.path(vs_repo_local, "round2.rds"))
 

### 2023-07-13 SDSS-2023/testing-charts-accuracy.qmd
move packages to preamble

@@
 keywords: 'list, keywords, here, 3-6, are, fine'
-format:
-  pdf:
-    documentclass: jds
-    fig-width: 6
-    fig-asp: 0.618
-    fig-align: center
-    include-in-header: 
-      - text: |
+preamble: >
   \usepackage{amsfonts,amsmath,amssymb,amsthm}
@@
   \usepackage{graphicx}
-          \usepackage{longtable}
-          \usepackage{hyperref}
-      - file: edits.tex
-    cite-method: natbib
-    keep-tex: true
-    keep-md: true
+  \newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
+  \newcommand{\kr}[1]{{\textcolor{teal}{#1}}}
+  \setlength{\parindent}{0pt}
+format:
+  pdf: default
+documentclass: jds
 ---
 
-\hh{hello?}
+<!-- \hh{hello?} -->
+<!-- fig-width: 6.5 -->
+<!-- fig-height: 4.5 -->
+<!-- template: template-jdsart.cls.tex -->
 
-\kr{Kiegan, this is your editing color :) - you can choose a different color in edits.tex}
+<!-- \kr{Kiegan, this is your editing color :) - you can choose a different color above} -->
 
@@
                       #cache = TRUE) 
+)
+
+knitr::opts_chunk$set(
+  jds.rmd::pdf_article()$knitr
 )
@@
 
-\hh{We can combine responses across samples into one combined dataset, but we need to adjust weights accordingly so that each sample is weighted equally in the model [@omuircheartaighCombiningSamplesVs2002].}
+We can combine responses across samples into one combined dataset, but we need to adjust weights accordingly so that each sample is weighted equally in the model [@omuircheartaighCombiningSamplesVs2002].
 
@@
 
-\hh{Using the AmeriSpeak survey tool, a total of 1902 participants were exposed to two barcharts each, as shown in \autoref{tbl-tasks-12}.}
+Using the AmeriSpeak survey tool, a total of 1902 participants were exposed to two barcharts each, as shown in \autoref{tbl-tasks-12}.
 

### 2023-07-13 SDSS-2023/testing-charts-accuracy.qmd
editing color works ...

@@
   \usepackage{graphicx}
-  \newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
-  \newcommand{\kr}[1]{{\textcolor{teal}{#1}}}
-  \setlength{\parindent}{0pt}
 format:
@@
 
-<!-- \hh{hello?} -->
+```{=latex}
+  \newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
+  \newcommand{\kr}[1]{{\textcolor{teal}{#1}}}
+  \setlength{\parindent}{0pt}
+```
+
+\hh{hello?}
+
+\kr{Kiegan, this is your editing color :) - you can choose a different color above}
+
 <!-- fig-width: 6.5 -->
@@
 <!-- template: template-jdsart.cls.tex -->
-
-<!-- \kr{Kiegan, this is your editing color :) - you can choose a different color above} -->
-
-<!--
-\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
-\setlength{\parindent}{0pt}
-\usepackage{amsfonts,amsmath,amssymb,amsthm}
-  \usepackage{booktabs}
-  \usepackage{lipsum}
-  \usepackage{subfig}
-  \usepackage{graphicx}
-  \usepackage{longtable}
-\hh{hello world!}
--->
 

### 2023-07-13 SDSS-2023/testing-charts-accuracy.qmd
expand on cleveland and mcgill, add mackinlay

@@
 @clevelandGraphicalPerceptionTheory1984 defined for the purpose of their seminal study the better visual as the one that allows viewers to make more accurate comparisons.
-What we always need to remember, is that these study results are based on only a few individuals recruited from the authors' coworkers and their spouses.
-@heerCrowdsourcingGraphicalPerception2010 have reproduced Cleveland and McGill's rankings using a crowd sourcing platform.
+Based on mappings of quantitative variables to different graphical elements,
+Cleveland and McGill's study resulted in a ranking of perceptual tasks from most accurate to least accurate. 
+@mackinlayAutomatingDesignGraphical1986 took this approach and expanded the tasks and their order to  ordinal and nominal scales, as shown in @fig-perceptual-tasks.
+
+```{r}
+#| label: fig-perceptual-tasks
+#| fig-cap: Ranking of perceptual tasks, as given by @mackinlayAutomatingDesignGraphical1986.
+The ranking of tasks on the quantitative scale are empirically verified by @clevelandGraphicalPerceptionTheory1984.
+knitr::include_graphics("images/mackinlay-tasks.png")
+```
+
+What we need to remember, is that Mackinlay's work is purely theoretical.
+Cleveland and McGill's rankings based on a study.
+However, this study is a small convenience sample, consisting of only a few individuals recruited from among the authors' coworkers and their spouses.
+@heerCrowdsourcingGraphicalPerception2010 have reproduced Cleveland and McGill's rankings using a crowd sourcing platform. A total of XXX amazon turkers were involved.
 
 <!-- Bias from crowd-sourcing? sample size?  -->
-crowd workers are known to be biased towards male, young and relatively higher education: @borgoCrowdsourcingInformationVisualization2017
+Crowd workers are known to be biased towards male, young and relatively higher education: @borgoCrowdsourcingInformationVisualization2017
 

### 2023-07-13 SDSS-2023/testing-charts-accuracy.qmd
mackinlay reference

@@
 ```{r run-to-update-bibfile, eval=FALSE, include=FALSE}
-keys <- rbbt::bbt_detect_citations("testing-charts-accuracy.Rmd")
+keys <- rbbt::bbt_detect_citations("testing-charts-accuracy.qmd")
 citations_only <- grep(pattern="[fig|tbl|eq][-|:].*", keys, value=TRUE, invert=TRUE)

### 2023-07-14 analysis/analysis.qmd
images and round 3

@@
 
+```{r data-rnd3}
+round3 <- readRDS(file.path(vs_repo_local, "round3.rds"))
+
+# filter out all of the 'skipped on web':
+round3_adj <- round3 %>% mutate(
+  weight = WEIGHT_NOLA
+)
+```
+
 The data for this paper were collected in several rounds as part of the NORC Omnibus.
@@
 | Round 2 | May 2022 |  `r nrow(round2_adj)`   |    `r round(neff(round2_adj$weight),1)`       | `r round(sum(round2_adj$weight),1)`       |
+| Round 3 | Jun 2022 |  `r nrow(round3_adj)`  |    `r round(neff(round3_adj$weight),1)`       | `r round(sum(round3_adj$weight),1)`       |
 | Round 6 | Sep 2022 |  `r nrow(round6_adj)` [Split]  |    `r round(neff(round6_adj$weight),1)`       | `r round(sum(round6_adj$weight),1)`       |
@@
 ```
+## Model 3: Aligned and unaligned comparisons between horizontal and vertical barcharts {#sec-model3}
+
+@tbl-designs shows the three designs that were considered here: We are using the **vertical** design as a baseline and have established that we get similar responses in different rounds of the survey.
+This design also produces stable results under slight modifications, such as the exact color scheme.
+
+```{r}
+#| label: tbl-designs
+#| tbl-cap: Design and stimulus of round 3
+#| tbl-subcap:
+#|   - Horizontal stacked bars
+#|   - Horizontal wide
+#| layout-nrow: 2
+
+knitr::include_graphics("images/round3/horizontal-ab.PNG")
+knitr::include_graphics("images/round3/hwide-ab.PNG")
+knitr::include_graphics("images/round3/horizontal-cd.PNG")
+knitr::include_graphics("images/round3/hwide-cd.PNG")
+```
+
+We are mostly interested in a comparison between this vertical design and the **wide horizontal** design favored by many NORC designers (XXX can I say that this way?).
+This design *should* help two-fold with decision making, because a.
+the bars are overall wider, increasing the physical representation of the stimulus, b.
+the bars are closer together.
+Both of these factors have been shown [@lu2022] to be beneficial for an accurate assessment of small differences between bars.
+
+In order to separate these two effects from the effect of horizontal vs vertical assessment of bars, we introduced the **horizontal** design as a transition, in which the bars are kept to the exact same size as in the vertical design.
+Similarly, the spacing of the bars is identical between these two designs.
+
+Interestingly, the resulting effects on the accuracy of the responses are not completely straightforward: the shift from a horizontal to a vertical design seems to be detrimental to an accurate perception in both situations.
+The re-scaled design of the wide horizontal bars reclaims some of the loss for unaligned bars and outperforms the vertical design by a similar margin in aligned bars.
+A contributing factor to this outcome might be the way that participants interact with the different designs.
+Generally, about half of all participants make use of the option to zoom into charts - and it can be shown that zooming into charts helps with the overall accuracy (which is in agreement with the findings by @lu2022 about the physical size of stimuli).
+The rate of zooming into charts is about twice as high for participants filling out the survey on their smart phones than on other, usually bigger devices such as tablets or desktops.
+However, different designs lead to different rates of zooming: when dealing with the vertical design, the rate of zooming is higher than for the two horizontal designs.
 

### 2023-07-14 SDSS-2023/testing-charts-accuracy.qmd
starting to fill in outline for intro, etc.

@@
 format:
-  pdf: default
-documentclass: jds
+  html: default
+editor: 
+  markdown: 
+    wrap: sentence
 ---
@@
 ```
-
 \hh{hello?}
@@
 <!-- fig-width: 6.5 -->
+
 <!-- fig-height: 4.5 -->
+
 <!-- template: template-jdsart.cls.tex -->
@@
 <!-- What is our main objective? -->
+
+\kr{Should the abstract match -- or be close to -- our SDSS short abstract?}
+
+Outline for introduction:
+
+-   What do viewers see when we show them a data chart?
+
+-   Visuals allow us to make comparisons between values
+
+-   Design of a data visualization impacts viewer ability to make those comparisons
+
+    -   And more broadly, impacts how viewers interact with a chart
+
+-   There is prior work studying this topic
+
+    -   Cleveland and McGill laid the foundations
+
+    -   Mackinlay expanded to theoretical scales
+
+    -   Heer/Bostock expanded to larger study population
+
+-   However, prior work has utilized specific populations of study respondents.
+    These populations are not representative of the general population, a common target audience for data visualization and scientific communication work
+
+-   Our work focuses on the following:
+
+    -   Utilizing a survey panel as a mechanism to test perception in data visualizations
+
+    -   Testing whether prior results from convenience samples hold with a nationally representative sample
+
+    -   Studying variations in data visualization design and their impact on respondent behavior and accuracy of participants' responses.
+
+-   This paper presents a series of tests completed and the resulting findings.
+
+-   The remainder of the paper is organized as follows:
+
+    -   we first describe the stimulus design
+
+    -   then we describe the study population (and survey design?)
+
+    -   subsequently, we share analyses of resulting survey responses across a series of stimuli
+
+    -   finally, we discuss implications of this study and next steps.
+
 What do viewers see, when we show them a data chart?
@@
 @clevelandGraphicalPerceptionTheory1984 defined for the purpose of their seminal study the better visual as the one that allows viewers to make more accurate comparisons.
-Based on mappings of quantitative variables to different graphical elements,
-Cleveland and McGill's study resulted in a ranking of perceptual tasks from most accurate to least accurate. 
+Based on mappings of quantitative variables to different graphical elements, Cleveland and McGill's study resulted in a ranking of perceptual tasks from most accurate to least accurate.
 @mackinlayAutomatingDesignGraphical1986 took this approach and expanded the tasks and their order to ordinal and nominal scales, as shown in @fig-perceptual-tasks.
@@
 However, this study is a small convenience sample, consisting of only a few individuals recruited from among the authors' coworkers and their spouses.
-@heerCrowdsourcingGraphicalPerception2010 have reproduced Cleveland and McGill's rankings using a crowd sourcing platform. A total of XXX amazon turkers were involved.
+@heerCrowdsourcingGraphicalPerception2010 have reproduced Cleveland and McGill's rankings using a crowd sourcing platform.
+A total of XXX amazon turkers were involved.
 
 <!-- Bias from crowd-sourcing? sample size?  -->
+
 Crowd workers are known to be biased towards male, young and relatively higher education: @borgoCrowdsourcingInformationVisualization2017
@@
 knitr::include_graphics("images/round3-modified.PNG")
+
 ```
 
+## Study design -- stimulus
+
 <!-- motivate the use of just noticeable differences -->
+
 Asking perceptual questions in a survey is different from the controlled environment of a cognitive lab, where these kind of questions would usually be addressed.
@@
 
-- How do structural design choices in a data visualization impact viewers’ ability to identify the larger of two elements?
+-   How do structural design choices in a data visualization impact viewers' ability to identify the larger of two elements?
 
-- How do aesthetic design choices in a data visualization impact viewers’ ability to identify the larger of two elements?
+-   How do aesthetic design choices in a data visualization impact viewers' ability to identify the larger of two elements?
 
@@
 
-Aesthetic design choices (structural choices seems stronger/there could be a lot to talk about there… should we skip aesthetic on this one?)
+Aesthetic design choices (structural choices seems stronger/there could be a lot to talk about there... should we skip aesthetic on this one?)
 
@@
 
-Binary accuracy (correct/incorrect – ‘they are the same’ is incorrect here)
+Binary accuracy (correct/incorrect -- 'they are the same' is incorrect here)
 
@@
 
-I’ve noted this below, but: how to model? Ordinal response? Binary (certain or very certain vs everybody else)?
+I've noted this below, but: how to model?
+Ordinal response?
+Binary (certain or very certain vs everybody else)?
+
+## Study design - survey population
+
+We utilize NORC's AmeriSpeak Omnibus survey, which utilizes a probability-based panel and surveys a sample of 1,000 nationally representative adults 18 and older.
+The advantage of using a probability-based approach is two-fold.
+First, we have access to a large sample of survey participants and thus have greater power in making inference about graphical perceptional abilities.
+Second, the sample is representative of the general adult public in the U.S., which is an important target audience for scientific communication.
 
@@
 <!-- describe stimulus -->
-Both charts in \autoref{fig-tasks-12} show the same data with slight modifications to the order of the levels – the first and second level in each of the bars are reversed between the left and the right chart.
+
+Both charts in \autoref{fig-tasks-12} show the same data with slight modifications to the order of the levels -- the first and second level in each of the bars are reversed between the left and the right chart.
 Participants were asked to compare the relative sizes of the tiles marked A and B (C and D, respectively) and select the correct response out of the possible choices:
@@
 Answer 2 is the correct answer for both charts.
-Both charts are shown at the same size, i.e.
-in both cases the difference in size between the bars is exactly the same, the vertical distance between the bars is the exact same amount.
-This leaves the vertical positioning of the bars as the only difference between the charts. Any differences in observed accuracy can therefore be attributed to this difference in presentation.
+Both charts are shown at the same size, i.e. in both cases the difference in size between the bars is exactly the same, the vertical distance between the bars is the exact same amount.
+This leaves the vertical positioning of the bars as the only difference between the charts.
+Any differences in observed accuracy can therefore be attributed to this difference in presentation.
 
@@
 <!-- |                                   |                                   | -->
+
 <!-- |-----------------------------------|-----------------------------------| -->
+
 <!-- | **Unaligned positions**           | **Aligned positions**             | -->
+
 <!-- | ![Unaligned positions marked A and B](../SDSS-2023/images/vertical-ab.PNG) | ![Aligned positions marked C and D](../SDSS-2023/images/vertical-cd.PNG) | -->
@@
 
-## Data Analysis and Results
+# Results
+
+How should we group results?
+
+-   By stimulus/structural variation?
+
+-   By response (binary, ordinal, certainty, time spent, etc.)
+
+Results to discuss:
+
+-   Differences in accuracy between aligned and unaligned bar
+
+-   Differences in accuracy between framed, floating, and general pie
+
+-   Differences in timing and zooming behavior across all designs
+
+-
 
@@
 
-Question for Ed: If we compare a full sample to a split sample, do we still want to weight these ‘equally’?
+Question for Ed: If we compare a full sample to a split sample, do we still want to weight these 'equally'?
 
-Analysis should be done using the ‘survey’ package and weights should be taken into account.
+Analysis should be done using the 'survey' package and weights should be taken into account.
 
@@
 Because each participant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts.
-The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: < 2.2e-16).
+The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: \< 2.2e-16).
 
@@
 
-STRUCTURAL VARIATION – START HERE
+STRUCTURAL VARIATION -- START HERE
 
@@
 
-% yes across each different structural condition
+\% yes across each different structural condition
 
@@
 
-% zoomed by device + alignment (already have this chart)
+\% zoomed by device + alignment (already have this chart)
 
@@
 
-Model 2 (this may not be feasible for comparison depending on what level the ‘TOTALTIME’ is captured at):
+Model 2 (this may not be feasible for comparison depending on what level the 'TOTALTIME' is captured at):
 
@@
 
-Facet by device type, zoom, structural condition, alignment? Play around with it
+Facet by device type, zoom, structural condition, alignment?
+Play around with it
 
@@
 
-Same models as above, but I’m not sure how we want to do the response. Ordinal response? Binary (certain or very certain vs everybody else)?
+Same models as above, but I'm not sure how we want to do the response.
+Ordinal response?
+Binary (certain or very certain vs everybody else)?
 
-AESTHETIC VARIATION – ONLY IF TIME
+AESTHETIC VARIATION -- ONLY IF TIME
 
-Binary accuracy (correct/incorrect – ‘they are the same’ is incorrect here) across structural choices
+Binary accuracy (correct/incorrect -- 'they are the same' is incorrect here) across structural choices
 
@@
 
-Same models as above, but I’m not sure how we want to do the response. Ordinal response? Binary (certain or very certain vs everybody else)?
+Same models as above, but I'm not sure how we want to do the response.
+Ordinal response?
+Binary (certain or very certain vs everybody else)?
 
@@
 
-# Supplementary Material {-}
+Key findings to discuss:
 
-+ **Participant Data (Linear):** Link to csv file with the data. 
-+ **Data Analysis Code:** Link to an html document with annotated code chunks.
+-   Design choices impact viewer behavior
+
+-   Design choices impact viewer accuracy
+
+-   Reproduced prior convenience sample results using a large, nationally-representative survey population
+
+-   Surveys can be used to ask these types of questions -- in the midst of other topics and with a limited number of questions, we are able to ask perception questions and produce results consistent with prior studies
+
+-
+
+# Supplementary Material {.unnumbered}
+
+-   **Participant Data (Linear):** Link to csv file with the data.
+-   **Data Analysis Code:** Link to an html document with annotated code chunks.
 

### 2023-07-14 SDSS-2023/testing-charts-accuracy.qmd
adding some more outlining to the study design and study description

@@
 
+    -   We consider structural (and aesthetic?) variations on bar charts and pie charts
+
+    -   Our tests are centered around the following central research questions:
+
+        -   How do structural design choices in a data visualization impact viewers' ability to identify the larger of two elements?
+
+        -   How is viewer behavior (zooming, time spent on question, certainty of response) impacted by structural and aesthetic design choices in a data visualization?
+
 -   The remainder of the paper is organized as follows:
@@
 
+Outline for study design - stimulus
+
+-   Use of Just-Noticeable Difference
+
+    -   Differences between cognitive lab and survey
+
+    -   Understanding how design impacts the most difficult comparisons
+
+-   Stimuli within context of data viz
+
+    -   Title, legend, axes, etc.
+
+-   Structural variations
+
+    -   Alignment -- across all tests
+
+    -   Stacked bar chart -- vertical, horizontal, and horizontal wide
+
+        -   Vertical and horizontal bars are the same exact visual stimuli but rotated/presented differently
+
+    -   Facetted bar chart
+
+    -   Facetted pie chart -- 4 colors, one color framed, one color floating
+
+-   Results are organized as follows:
+
+    -   ....
+        Haven't decided yet.
+
 <!-- motivate the use of just noticeable differences -->
@@
 
-We utilize NORC's AmeriSpeak Omnibus survey, which utilizes a probability-based panel and surveys a sample of 1,000 nationally representative adults 18 and older.
+Participants were recruited as part of NORC's AmeriSpeak panel, which utilizes a probability-based sampling methodology and samples U.S.
+households from NORC's National Sample Frame that provides coverage of over 97% of U.S.
+households.
+The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households **CITE DENNIS 2022**.
+Each test was conducted using the AmeriSpeak OMnibus survey, which runs biweekly and pulls a sample of around 1,000 U.S. adults and asks participants to answer questions on a variety of topics.
 The advantage of using a probability-based approach is two-fold.
@@
 Second, the sample is representative of the general adult public in the U.S., which is an important target audience for scientific communication.
+
+Given the nature of pulling a sample from the panel for each Omnibus round, there is a possibility that some participants may have been included in multiple rounds; however, our collected data is not a longitudinal or panel study, so we do not have repeated responses from the same participants across each of our tests.
+Each of the tests was run at least a month apart, so if respondents participated in multiple rounds, there was at least a one month gap between viewing each visual stimulus.
 

### 2023-07-14 analysis/analysis.qmd
round 3 - overview

@@
   weight = WEIGHT_NOLA
+) %>% filter(
+  ab != "SKIPPED ON WEB",
+  cd != "SKIPPED ON WEB"
 )
@@
 
-@tbl-designs shows the three designs that were considered here: We are using the **vertical** design as a baseline and have established that we get similar responses in different rounds of the survey.
-This design also produces stable results under slight modifications, such as the exact color scheme.
+We are mostly interested in a comparison between this vertical design and the **wide horizontal** design favored by many NORC designers (XXX can I say that this way?).
+This design *should* help two-fold with decision making, because
+
+    a. the bars are overall wider, increasing the physical representation of the stimulus, 
+    b. the bars are closer together.
+
+Both of these factors have been shown [@hughesJustNoticeableDifferences2001; @luModelingJustNoticeable2022] to be beneficial for an accurate assessment of small differences between bars.
+
+In order to separate these two effects from the effect of horizontal vs vertical assessment of bars, we introduce the **horizontal** design as a transition, in which the bars are kept to the exact same size as in the vertical design.
+Similarly, the spacing of the bars is identical between these two designs.
+
+```{r round3-participants}
+# P_EXP is 1 or 2:
+# panelists in 1 see HZ1-AB.png and HZ1-CD.png (horizontal tall)
+# panelists in 2 see HZ2-AB.png and HZ2-CD.png (horizontal wide)
+n_tall <- round3_adj %>% filter(P_EXP == 1) %>% nrow()
+n_tall_eff <- round3_adj %>% filter(P_EXP == 1) %>% pull(weight) %>% neff()
+
+n_wide <- round3_adj %>% filter(P_EXP == 2) %>% nrow()
+n_wide_eff <- round3_adj %>% filter(P_EXP == 2) %>% pull(weight) %>% neff()
+```
+
+We use the results from rounds 1 and 2 to assess perceptual accuracy in the vertical stacked barchart (as shown in @fig-tasks-12). 
+@fig-designs shows the two designs shown to participants of Round 3.
+The survey was administered using a split design on the type of visualization, but all participants were asked to compare the size of two unaligned tiles (A and B) and two aligned tiles (C and D).
+Roughly half ($n_1 = $ `r n_tall` for an efficient sample size of `r n_tall_eff`) of the round 6 panelists were shown the horizontal tall barchart on the left, the other participants ($n_2 = $ `r n_wide`; efficient $n_2^{(eff)} = $ `r n_wide_eff`) saw the horizontal wide barchart.
 
 ```{r}
-#| label: tbl-designs
-#| tbl-cap: Design and stimulus of round 3
-#| tbl-subcap:
-#|   - Horizontal stacked bars
-#|   - Horizontal wide
+#| label: fig-designs
+#| fig-cap: Design and stimulus of round 3
+#| fig-subcap:
+#|   - Horizontal tall, stacked bars, unaligned
+#|   - Horizontal wide, stacked bars, unaligned
+#|   - Horizontal tall, stacked bars, aligned
+#|   - Horizontal wide, stacked bars, aligned
 #| layout-nrow: 2
@@
 
-We are mostly interested in a comparison between this vertical design and the **wide horizontal** design favored by many NORC designers (XXX can I say that this way?).
-This design *should* help two-fold with decision making, because a.
-the bars are overall wider, increasing the physical representation of the stimulus, b.
-the bars are closer together.
-Both of these factors have been shown [@lu2022] to be beneficial for an accurate assessment of small differences between bars.
-
-In order to separate these two effects from the effect of horizontal vs vertical assessment of bars, we introduced the **horizontal** design as a transition, in which the bars are kept to the exact same size as in the vertical design.
-Similarly, the spacing of the bars is identical between these two designs.
-
 Interestingly, the resulting effects on the accuracy of the responses are not completely straightforward: the shift from a horizontal to a vertical design seems to be detrimental to an accurate perception in both situations.
@@
 A contributing factor to this outcome might be the way that participants interact with the different designs.
-Generally, about half of all participants make use of the option to zoom into charts - and it can be shown that zooming into charts helps with the overall accuracy (which is in agreement with the findings by @lu2022 about the physical size of stimuli).
+Generally, about half of all participants make use of the option to zoom into charts - and it can be shown that zooming into charts helps with the overall accuracy (which is in agreement with the findings by @luModelingJustNoticeable2022 about the physical size of stimuli).
 The rate of zooming into charts is about twice as high for participants filling out the survey on their smart phones than on other, usually bigger devices such as tablets or desktops.

### 2023-07-14 SDSS-2023/testing-charts-accuracy.qmd

@@
 
+## Respondents
+
+Description of respondents -- \# of respondents in each round, broad overview of demo characteristics(?), raw sample size and effective sample size
+
 How should we group results?
@@
 -   By response (binary, ordinal, certainty, time spent, etc.)
+
+-   By accuracy vs by behavior (zooming, time spent, etc.)
 
@@
 
--
-
 COMBINING SAMPLES AND WEIGHTING NOTES:
@@
 We can combine responses across samples into one combined dataset, but we need to adjust weights accordingly so that each sample is weighted equally in the model [@omuircheartaighCombiningSamplesVs2002].
-
-Question for Ed: If we compare a full sample to a split sample, do we still want to weight these 'equally'?
-
-Analysis should be done using the 'survey' package and weights should be taken into account.
 
@@
 
--
+-   Studying how viewers interact with charts can be done in a survey, and more expansive work on this topic should be completed to understand how the general public interacts with and understands charts.
 

### 2023-07-14 analysis/analysis.qmd
number of participants and responses

@@
 
-Both of these factors have been shown [@hughesJustNoticeableDifferences2001; @luModelingJustNoticeable2022] to be beneficial for an accurate assessment of small differences between bars.
+Both of these factors have been shown [@luModelingJustNoticeable2022] to be beneficial for an accurate assessment of small differences between bars.
 
@@
 The survey was administered using a split design on the type of visualization, but all participants were asked to compare the size of two unaligned tiles (A and B) and two aligned tiles (C and D).
-Roughly half ($n_1 = $ `r n_tall` for an efficient sample size of `r n_tall_eff`) of the round 6 panelists were shown the horizontal tall barchart on the left, the other participants ($n_2 = $ `r n_wide`; efficient $n_2^{(eff)} = $ `r n_wide_eff`) saw the horizontal wide barchart.
+Roughly half ($n_1 = $ `r n_tall` for an efficient sample size of `r round(n_tall_eff,1)`) of the round 6 panelists were shown the horizontal tall barchart on the left, the other participants ($n_2 =$   `r n_wide`; efficient $n_2^{(eff)} =$    `r round(n_wide_eff,1)`) saw the horizontal wide barchart.
 
@@
 knitr::include_graphics("images/round3/hwide-cd.PNG")
+```
+
+```{r combine-123}
+lambda_1_2 <- calculate_lambda(round1_adj$weight, round2_adj$weight)
+
+round_12 <- round1_adj %>% 
+            mutate(weight = lambda_1_2*weight/sum(weight),
+                   round = "Round 1") %>% 
+            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
+                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, weight) %>%
+    rbind(round2_adj %>% 
+            mutate(weight = (1-lambda_1_2)*weight/sum(weight),
+                   round = "Round 2") %>% 
+            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, 
+                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, weight))
+
+lambda_12_3 <- calculate_lambda(round_12$weight, round3_adj$weight)
+
+round_123 <- round_12 %>% 
+              mutate(weight = lambda_12_3*weight/sum(weight),
+                     task = "Vertical") %>% 
+  rbind(
+    round3_adj %>% 
+      mutate(
+        weight = (1-lambda_12_3)*weight/sum(weight),
+        round = "Round 3",
+        CorrectAB = ab == "B is bigger",
+        CorrectCD = cd == "D is bigger",
+        CorrectABCD = paste(CorrectAB, CorrectCD),
+        task = ifelse(P_EXP==1, "Horizontal", "Horizontal wide")
+       ) %>% 
+        dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
+                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, task, weight) 
+  )
+
+round_123$weight <- neff(round_123$weight)/sum(round_123$weight)*round_123$weight 
+# does not affect the further analysis, but helps with the interpretation of results
+
+round_123_long <- round_123 %>% 
+  pivot_longer(CorrectAB:CorrectCD, names_to = "Question", values_to = "Response")
+
+survey_123 <- 
+  survey::svydesign(
+  data = round_123_long, 
+  ids = ~CaseId, weights = ~weight)
+```
+
+We combine rounds 1, 2 and 3 as shown in @tbl-combine-123 to incorporate all relevant responses to evaluate the three different designs of charts in terms of their accuracy in assessing small differences in tile sizes.
+
+```{r tbl-combine-123}
+#| label: tbl-combine-123
+#| tbl-cap: Table of the number of effective responses by type of visualization and round
+
+lambdas <- round_123_long %>% group_by(round, task) %>% 
+  summarize(
+    lambda = calculate_lambda(
+      round_123_long$weight[round_123_long$task==task[1] & round_123_long$round==round[1]],
+      round_123_long$weight[!(round_123_long$task==task[1] & round_123_long$round==round[1])])
+  )
+
+dframe <- round_123_long %>% group_by(round, task) %>% 
+  summarize(
+    nominal = n(),
+    effective = round(neff(weight),1),
+    participants = length(unique(CaseId))
+  ) %>% left_join(
+    lambdas, by = c("round", "task")
+  )
+
+dframe %>% 
+  group_by(round) %>%
+  gt() %>%
+  tab_header("Number of (effective) responses by task") %>%
+  cols_label(
+    nominal = md("number of responses:<br>nominal"),
+    effective = md("number of responses:<br>effective"),
+    participants = md("number of participants"),
+    lambda = md("λ"),
+  ) %>%
+  fmt_number(
+    everything(), 
+    decimals = 0,
+    use_seps=TRUE
+  ) %>%
+  fmt_number(
+    columns = effective,
+    decimals = 1
+  ) %>%
+  fmt_number(
+    columns = c(lambda),
+    decimals = 4
+  ) %>%
+  tab_options(
+    column_labels.font.weight = "bold",
+    row_group.font.weight = "bold"
+  )
 ```

### 2023-07-15 analysis/analysis.qmd
raccuracy for task & chart

@@
               mutate(weight = lambda_12_3*weight/sum(weight),
-                     task = "Vertical") %>% 
+                     chart = "Vertical") %>% 
   rbind(
@@
         CorrectABCD = paste(CorrectAB, CorrectCD),
-        task = ifelse(P_EXP==1, "Horizontal", "Horizontal wide")
+        chart = ifelse(P_EXP==1, "Horizontal", "Horizontal wide")
        ) %>% 
         dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
-                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, task, weight) 
+                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, chart, weight) 
   )
@@
 round_123_long <- round_123 %>% 
-  pivot_longer(CorrectAB:CorrectCD, names_to = "Question", values_to = "Response")
+  pivot_longer(ab:cd, names_to = "task", values_to = "Response")
+
+round_123_long <- round_123_long %>% 
+  mutate(
+    correct = Response %in% c("B is bigger", "D is bigger"),
+    correct3 = correct | (Response == "They are the same"),
+    task_chart = factor(paste(task, chart, sep="-"))
+  )
 
@@
 
-```{r tbl-combine-123}
+```{r tbl-combine-123, warning = FALSE, message = FALSE}
 #| label: tbl-combine-123
@@
 
-lambdas <- round_123_long %>% group_by(round, task) %>% 
+lambdas <- round_123_long %>% group_by(round, chart) %>% 
   summarize(
     lambda = calculate_lambda(
-      round_123_long$weight[round_123_long$task==task[1] & round_123_long$round==round[1]],
-      round_123_long$weight[!(round_123_long$task==task[1] & round_123_long$round==round[1])])
-  )
-
-dframe <- round_123_long %>% group_by(round, task) %>% 
+      round_123_long$weight[round_123_long$chart==chart[1] & round_123_long$round==round[1]],
+      round_123_long$weight[!(round_123_long$chart==chart[1] & round_123_long$round==round[1])])
+  )
+
+dframe <- round_123_long %>% group_by(round, chart) %>% 
   summarize(
@@
   ) %>% left_join(
-    lambdas, by = c("round", "task")
+    lambdas, by = c("round", "chart")
   )
@@
 ```
-
-Interestingly, the resulting effects on the accuracy of the responses are not completely straightforward: the shift from a horizontal to a vertical design seems to be detrimental to an accurate perception in both situations.
-The re-scaled design of the wide horizontal bars reclaims some of the loss for unaligned bars and outperforms the vertical design by a similar margin in aligned bars.
+```{r model-123, warning = FALSE, message = FALSE}
+
+logit_123_correct <- svyglm(correct~task_chart-1, family=quasibinomial(), design=survey_123)
+logit_123_correct3 <- svyglm(correct3~task_chart-1, family=quasibinomial(), design=survey_123)
+
+ci_confint <- svyby(~correct, by=~task_chart, design=survey_123, svyciprop)
+ci_confint3 <- svyby(~correct3, by=~task_chart, design=survey_123, svyciprop)
+
+coefs_123 <- broom::tidy(logit_123_correct)
+coefs_123_relaxed <- broom::tidy(logit_123_correct3)
+
+# create all pairwise comparisons for tasks and questions.
+cross_compare <- glht(logit_123_correct, mcp(task_chart="Tukey"))
+cross_compare3 <- glht(logit_123_correct3, mcp(task_chart="Tukey"))
+
+# get letters for each confidence 
+ci_confint$letters <- cld(cross_compare)$mcletters$Letters
+ci_confint3$letters <- cld(cross_compare3)$mcletters$Letters
+
+ci_confint <- ci_confint %>% 
+  separate_wider_delim(
+    task_chart, delim = "-", names =c("task", "chart")
+    )  %>% mutate(
+  task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+  chart = factor(chart, levels =c("Vertical", "Horizontal", "Horizontal wide"))
+)
+
+ci_confint3 <- ci_confint3 %>% 
+  separate_wider_delim(
+    task_chart, delim = "-", names =c("task", "chart")
+    ) %>% mutate(
+  task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+  chart = factor(chart, levels =c("Vertical", "Horizontal", "Horizontal wide"))
+) 
+```
+
+```{r}
+#| label: fig-response-123
+#| fig-cap: "Responses for accuracy in the three designs.
+Responses to the same task are shown side-by-side for the three designs.
+The overlaid rectangles represent 95% confidence intervals.
+The letters in blue and orange encode significances between pairwise proportions: two bars have a significantly different proportion (at a 5% significance level) if they do not share a letter.
+There is no significant difference between the three designs for wrong responses.
+When tiles are unaligned, the horizontal wide barchart is showing the highest accuracy.
+For aligned tile, the horizontal wide design and the vertical design do not show a significant difference in accuracy."
+#| fig-height: 4.25
+#| fig-width: 7
+#| out-width: 80%
+
+round_123_long %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = chart)) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  facet_grid(.~task) + 
+  theme(legend.position="bottom") + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") + 
+  geom_rect(
+    aes(
+      xmin = as.numeric(chart) - 0.45,
+      ymin = correct - 1.96* `se.as.numeric(correct)`,
+      xmax = as.numeric(chart) + 0.45,
+      ymax = correct + 1.96* `se.as.numeric(correct)`,
+    ), 
+    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+    data = ci_confint
+  ) +
+  geom_segment(
+    aes(
+      x = as.numeric(chart) - 0.45,
+      y = correct,
+      xend = as.numeric(chart) + 0.45,
+      yend = correct
+    ), 
+    alpha = 1, colour = "black", linewidth = 0.25,
+    data = ci_confint
+  ) +
+  geom_rect(
+    aes(
+      xmin = as.numeric(chart) - 0.45,
+      ymin = correct3 - 1.96* `se.as.numeric(correct3)`,
+      xmax = as.numeric(chart) + 0.45,
+      ymax = correct3 + 1.96* `se.as.numeric(correct3)`,
+    ), 
+    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+    data = ci_confint3
+  ) +
+  geom_segment(
+    aes(
+      x = as.numeric(chart) - 0.45,
+      y = correct3,
+      xend = as.numeric(chart) + 0.45,
+      yend = correct3
+    ), 
+    alpha = 1, colour = "black", linewidth = 0.25,
+    data = ci_confint3
+  ) +
+  scale_y_continuous("Percent responses", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25), limits = c(-0.1, 1.1)) + 
+  geom_text(aes(label=letters), y = -0.05, data = ci_confint, colour = "steelblue") +
+  geom_text(aes(label=letters), y = 1.05, data = ci_confint3, colour = "darkorange") 
+```
+
+### Accuracy in different barchart designs
+
+@fig-response-123 shows the results of a cell-means model with ordinal response $Y$,
+
+$$
+\text{logit }P(Y_k \le y) = \mu_{ij(k)}
+$$
+where $Y_k$ is the $k$th participant's response with $Y_k \in \{1, 2, 3\}$, where 'correct' is encoded as 1, 'they are the same' is encoded as 2, and 'wrong' is encoded as 3; $i \in \{1, 2\}$ is the comparison type (1 = Aligned, 2 = Unaligned),  and $j \in \{1, 2, 3\}$ is the chart design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide.
+
+HH: the definition of this model isn't quite right yet - will fix it in the morning.
+
+Interestingly, while we should, theoretically, see an improvement in accuracy when shifting from the vertical to the horizontal wide design, the resulting effects on the accuracy of the responses are not completely straightforward:
+The shift from a vertical to the (tall) horizontal design is detrimental to an accurate perception for both aligned and unaligned comparisons.  
+The re-scaled design of the wide horizontal bars reclaims some of the loss for unaligned bars and outperforms the vertical design by a similar margin in aligned bars, but does not out-perform the vertical design when comparing unaligned tiles.
+
+### Zooming
+
 A contributing factor to this outcome might be the way that participants interact with the different designs.
@@
 However, different designs lead to different rates of zooming: when dealing with the vertical design, the rate of zooming is higher than for the two horizontal designs.
+
+### Certainty
 

### 2023-07-15 analysis/analysis.qmd
model formula fixed

@@
 
-@fig-response-123 shows the results of a cell-means model with ordinal response $Y$,
+@fig-response-123 shows the results of a cell-means model with ordinal response $Y_k$, where $Y_k$ is the $k$th participant's response, $Y_k \in \{1, 2, 3\}$, where 'correct' is encoded as 1, 'they are the same' is encoded as 2, and 'wrong' is encoded as 3:
 
 $$
-\text{logit }P(Y_k \le y) = \mu_{ij(k)}
+\text{logit }P(Y_k \le \ell) = \mu_{ij\ell(k)},
 $$
-where $Y_k$ is the $k$th participant's response with $Y_k \in \{1, 2, 3\}$, where 'correct' is encoded as 1, 'they are the same' is encoded as 2, and 'wrong' is encoded as 3; $i \in \{1, 2\}$ is the comparison type (1 = Aligned, 2 = Unaligned),  and $j \in \{1, 2, 3\}$ is the chart design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide.
-
-HH: the definition of this model isn't quite right yet - will fix it in the morning.
+where  $\ell \in \{1, 2\}$; $i \in \{1, 2\}$ is the comparison type (1 = Aligned, 2 = Unaligned),  and $j \in \{1, 2, 3\}$ is the chart design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide.
 

### 2023-07-15 analysis/analysis.qmd
table of log odds

@@
   mutate(task = reorder(task, time_spent)) %>%
-  ggplot(aes(y = task, x = time_spent)) + 
+  ggplot(aes(y = task, x = time_spent, fill=task)) + 
   geom_boxplot() +
@@
   ylab("") + xlab("") + 
-  ggtitle("Time spent on comparing all six pairs by design") 
+  ggtitle("Time spent on comparing all six pairs by design") +
+  theme(legend.position = "none")
 ```
@@
 ci_confint3$letters <- cld(cross_compare3)$mcletters$Letters
+
+coefs_123 <- coefs_123 %>% left_join(
+  ci_confint %>%
+    mutate(
+      task_chart = paste0("task_chart",task_chart, sep="")
+    ) %>% dplyr::select(task_chart, letters), by=c("term"= "task_chart"))
+
+coefs_123_relaxed <- coefs_123_relaxed %>% left_join(
+  ci_confint3 %>% 
+    mutate(
+      task_chart = paste0("task_chart",task_chart, sep="")
+    ) %>% dplyr::select(task_chart, letters), by=c("term"= "task_chart"))
 
@@
 where  $\ell \in \{1, 2\}$; $i \in \{1, 2\}$ is the comparison type (1 = Aligned, 2 = Unaligned),  and $j \in \{1, 2, 3\}$ is the chart design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide.
+The estimated values and 95% confidence intervals are shown in @tbl-rounds-123
+
+```{r}
+#| label: tbl-rounds-123
+#| tbl-cap: Log-odds for the cell-means model, letters behind numbers indicate pairwise significances.
+Within the same **column** values are significantly different (at 5%) if they do not share the same letter.
+
+dframe <- rbind(
+  coefs_123 %>% mutate(level = c("correct | same or wrong")),
+  coefs_123_relaxed %>% mutate(level = c("correct or same | wrong"))
+) %>% mutate(
+  term = gsub("task_chart", "", term),
+) %>% separate(term, into=c("task", "chart"), sep="-")
+
+dframe %>% 
+  dplyr::select(level, task, chart, estimate, std.error, letters) %>%
+  mutate(
+    task = ifelse(task=="ab", "Unaligned", "Aligned")
+  ) %>%
+  group_by(task) %>%
+  mutate(
+    estimate = exp(estimate),
+    std.error = exp(std.error)
+  ) %>%
+  tidyr::pivot_wider(
+        id_cols = c(task, chart),
+        names_from = level,
+        values_from = c(estimate, std.error, letters)
+      ) %>%
+  gt() %>%
+  tab_header(title = "Log odds (SD) of accuracy by task and chart type") %>%
+  fmt_number(
+    everything(),
+    decimals=2
+  ) %>%
+  cols_merge(
+    columns = c("estimate_correct | same or wrong", "std.error_correct | same or wrong",
+                "letters_correct | same or wrong"),
+    pattern = "<<{1}>><< ({2})>><< {3}>>"
+  ) %>%
+  cols_merge(
+    columns = c("estimate_correct or same | wrong", 
+                "std.error_correct or same | wrong",
+                "letters_correct or same | wrong"),
+    pattern = "<<{1}>><< ({2})>><< {3}>>"
+  ) %>%
+  tab_options(
+    column_labels.font.weight = "bold",
+    row_group.font.weight = "bold"
+  ) %>%
+  cols_label(
+    .list = c(
+      "chart" = "",
+      "estimate_correct | same or wrong" = "correct | same or wrong",
+      "estimate_correct or same | wrong" = "correct or same | wrong"
+    )
+  )
+```
 

### 2023-07-15 analysis/analysis.qmd
confidence intervals for log odds in table 6

@@
   mutate(
-    estimate = exp(estimate),
-    std.error = exp(std.error)
+    ci_low = exp(estimate-1.96*std.error),
+    ci_high = exp(estimate+1.96*std.error),
+    estimate = exp(estimate)
   ) %>%
@@
         names_from = level,
-        values_from = c(estimate, std.error, letters)
+        values_from = c(estimate, ci_low, ci_high, letters)
       ) %>%
   gt() %>%
-  tab_header(title = "Log odds (SD) of accuracy by task and chart type") %>%
+  tab_header(title = "Log odds of accuracy by task and chart type") %>%
   fmt_number(
@@
   cols_merge(
-    columns = c("estimate_correct | same or wrong", "std.error_correct | same or wrong",
-                "letters_correct | same or wrong"),
-    pattern = "<<{1}>><< ({2})>><< {3}>>"
+    columns = c("estimate_correct | same or wrong", "ci_low_correct | same or wrong",
+                "ci_high_correct | same or wrong", "letters_correct | same or wrong"),
+    pattern = "<<{1}>><< [{2}, {3}]>><< {4}>>"
   ) %>%
   cols_merge(
-    columns = c("estimate_correct or same | wrong", 
-                "std.error_correct or same | wrong",
-                "letters_correct or same | wrong"),
-    pattern = "<<{1}>><< ({2})>><< {3}>>"
+    columns = c("estimate_correct or same | wrong", "ci_low_correct or same | wrong",
+                "ci_high_correct or same | wrong", "letters_correct or same | wrong"),
+    pattern = "<<{1}>><< [{2}, {3}]>><< {4}>>"
   ) %>%

### 2023-07-17 analysis/analysis.qmd
logistic regression for zooming

@@
     correct3 = correct | (Response == "They are the same"),
-    task_chart = factor(paste(task, chart, sep="-"))
+    task_chart = factor(paste(task, chart, sep="-")),
+    zoom = ifelse(task=="ab", ab_zoom, cd_zoom),
+    response3 = ifelse(correct, "Correct", ifelse(correct3, "They are the same", "Wrong"))
   )
@@
 ```
+
 ```{r model-123, warning = FALSE, message = FALSE}
@@
 A contributing factor to this outcome might be the way that participants interact with the different designs.
-Generally, about half of all participants make use of the option to zoom into charts - and it can be shown that zooming into charts helps with the overall accuracy (which is in agreement with the findings by @luModelingJustNoticeable2022 about the physical size of stimuli).
-The rate of zooming into charts is about twice as high for participants filling out the survey on their smart phones than on other, usually bigger devices such as tablets or desktops.
-However, different designs lead to different rates of zooming: when dealing with the vertical design, the rate of zooming is higher than for the two horizontal designs.
+Generally, about half of all participants make use of the option to zoom into charts - while zooming does help with the overall accuracy (which is in agreement with the findings by @luModelingJustNoticeable2022 about the physical size of stimuli), the increase is not significant.
+However, different designs lead to different rates of zooming: when dealing with the vertical design, the rate of zooming is significantly higher than for the two horizontal designs.
+
+<!-- The rate of zooming into charts is about twice as high for participants filling out the survey on their smart phones than on other, usually bigger devices such as tablets or desktops. -->
+
+```{r zoom-123}
+logit_zoom <- svyglm(I(zoom=="yes")~response3+task+chart, family=quasibinomial(), design=survey_123)
+
+coefs_zoom <- broom::tidy(logit_zoom)
+```
+
+```{r}
+#| label: fig-zoom-123
+#| fig-cap: "Zooming - not significant for accuracy or task, but changes by the type of chart."
+#| fig-height: 4.25
+#| fig-width: 7
+#| out-width: 80%
+
+zt <- round_123_long %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide")),
+    zoom = factor(zoom, levels=c("no", "yes"))
+  ) %>% 
+  ggplot(aes(x = response3)) + 
+  geom_bar(aes(fill = zoom, weight = weight), position="fill") +
+  facet_grid(.~task) + 
+  theme(axis.text.x = element_text(angle=30, hjust = 1)) + 
+  scale_fill_manual("Zoom", values=c("grey", "#1274F3")) +
+  xlab("") + 
+  ggtitle("Zooming behavior by task")
+
+zd <- round_123_long %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide")),
+    zoom = factor(zoom, levels=c("no", "yes"))
+  ) %>% 
+  ggplot(aes(x = response3)) + 
+  geom_bar(aes(fill = zoom, weight = weight), position="fill") +
+  facet_grid(.~chart) + 
+  scale_fill_manual("Zoom", values=c("grey", "#1274F3")) +
+  xlab("") + 
+  ggtitle("Zooming behavior by chart") +
+  theme(axis.text.x = element_text(angle=30, hjust = 1))
+
+zt + zd + plot_layout(widths=c(5, 7.5), guides='collect')
+
+```
+
+Let $Y_{jk}$ describe the zooming behavior of panelist $k$ on task $j$.
+We model zooming behavior (no = 0, yes = 1) as a logistic regression by correctness of response ($\rho$),  task ($\tau$), and design ($\delta$) of the chart:
+$$
+\text{logit } P(Y_k \le 1) = \mu + \rho_{i(k)} + \tau_{j(k)} + \delta_{\ell(k)},
+$$
+
+```{r}
+#| label: tbl-zoom-123
+#| tbl-cap: Coefficients for logistic regression of zooming by task
+
+coefs_zoom %>%
+  mutate(
+    term = ifelse(term=="(Intercept)", "Intercept", term),
+    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value))
+  ) %>%
+  gt() %>%
+  tab_header("Estimates for logistic regression on zooming behavior") %>%
+  cols_label(
+    std.error = "SE",
+    statistic = "t-statistic",
+    p.value = "p-value"
+  ) %>%
+  fmt_number(
+    columns = c(estimate, std.error),
+    decimals = 2
+  ) %>%
+  fmt_number(
+    columns = statistic,
+    decimals = 1
+  ) %>%
+  cols_align(
+    columns = p.value,
+    align = "right"
+  )
+
+```
 

### 2023-07-17 SDSS-2023/testing-charts-accuracy.qmd
a bit more structure thinking in the results section

@@
 
-Description of respondents -- \# of respondents in each round, broad overview of demo characteristics(?), raw sample size and effective sample size
+Description of respondents
+
+-   \# of respondents in each round
+
+-   broad overview of demo characteristics(?)
+
+-   raw sample size and effective sample size
+
+Paragraph on combining samples across rounds
+
+-   Pull from analysis doc
+
+-   When combining responses from multiple rounds, we make weighting adjustments to ensure that each weights in each sample are properly calibrated to the population total across both rounds.
+    **NEED ED REVIEW OF THIS LANGUAGE.** This adjustment is completed as follows: \[insert Heike text from analysis doc here\].
+
+Test 1
+
+-   Alignment in stacked bars
+
+    -   Accuracy and responses
+
+    -
+
+Test 2
+
+-   Structure of stacked bars
+
+Test 3
+
+-   Facetted bars and pies
 

### 2023-07-17 analysis/analysis.qmd

@@
             dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
-                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, weight) %>%
+                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, weight,
+                   age4, gender, income4, racethnicity, educ5) %>%
     rbind(round2_adj %>% 
@@
             dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, 
-                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, weight))
+                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, weight,
+                   age4, gender, income4, racethnicity, educ5))
 
@@
         dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
-                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, chart, weight) 
+                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, chart, weight,
+                   age4, gender, income4, racethnicity, educ5) 
   )
@@
 
+### Accuracy and Demographics
+
+Let $Y_k$ be the response of participant $k$, on a scale from 1 = 'wrong', 2 = 'they are the same' to 3 = 'correct'. 
+$$
+\text{ logit } P(y_k \le i) = \mu_i + \alpha
+$$
+
+```{r}
+#| label: fig-demographics
+#| fig-cap: "Accuracy across demographics.
+Gender and age are not significant factors.
+However, educational level and income do have -- a small -- effect.
+When income levels increase, the percentage of wrong answers decreases.
+#| With increasing levels in education, we see a small drop in correct answers, as well as a small drop in wrong answers, for a significant increase in 'they are the same'."
+#| fig-height: 4.25
+#| fig-width: 7
+#| out-width: 100%
+
+educ <- round_123_long %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = educ5)) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="bottom",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") 
+gender <- round_123_long %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = gender)) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="bottom",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") 
+income <- round_123_long %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = income4)) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="bottom",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("")
+
+race <- round_123_long %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = racethnicity)) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="bottom",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("")
+
+age <- round_123_long %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = age4)) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="bottom",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("")
+
+gender + age +income + educ + plot_layout(nrow=1)
+```
+
+```{r demographics}
+logit_demo <- svyglm(correct~gender+educ5+age4+income4, design=survey_123, family=quasibinomial())
+logit_demo3 <- svyglm(correct3~gender+educ5+age4+income4, design=survey_123, family=quasibinomial())
+
+summary(logit_demo)
+summary(logit_demo3)
+```
+
 ## Model X: Comparing Aligned and Unaligned Elements Facets {#sec-modelX}

### 2023-07-17 SDSS-2023/testing-charts-accuracy.qmd
move html to default position

@@
   \usepackage{graphicx}
+  \usepackage{hyperref}
 format:
   html: default
+  pdf: 
+    documentclass: jds
+    include-in-header: authors.tex
+    keep-tex: true
 editor: 
@@
 ```
-\hh{hello?}
-
-\kr{Kiegan, this is your editing color :) - you can choose a different color above}
 
 <!-- fig-width: 6.5 -->
-
 <!-- fig-height: 4.5 -->
-
 <!-- template: template-jdsart.cls.tex -->
@@
 The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households **CITE DENNIS 2022**.
-Each test was conducted using the AmeriSpeak OMnibus survey, which runs biweekly and pulls a sample of around 1,000 U.S. adults and asks participants to answer questions on a variety of topics.
+Each test was conducted using the AmeriSpeak Omnibus survey, which runs biweekly and samples around 1,000 U.S.adults to answer questions on a variety of topics.
 The advantage of using a probability-based approach is two-fold.
@@
 Second, the sample is representative of the general adult public in the U.S., which is an important target audience for scientific communication.
+
+\hh{XXX I think what you are trying to say is that we do not know, if participants have seen the stimuli multiple times, but if they did, there was at least a month in between?}
 
@@
 #| label: fig-tasks-12
-#| fig-cap: "The two stacked barcharts every participant got to see.
-In each barchart, the two marked tiles are to be compared for their size.
-In both instances, the tile on the right is (very slightly) larger."
+#| fig-cap: "Two stacked (vertical) barcharts. In each barchart, two tiles are marked. In both instances, the tile on the right is (very slightly) larger. "
 #| fig-subcap:
@@
 knitr::include_graphics("images/vertical-cd.PNG")
+```
+
+```{r}
+#| label: fig-tasks-3
+#| fig-cap: "Changes to to the vertical design are made in two steps.
+First, the original chart is rotated.
+The areas of all tiles are kept the same.
+In a second step, the aspect ratio of tiles is changed while keeping the areas of tiles the same."
+#| fig-subcap:
+#| - "**Tall horizontal bars**, unaligned"
+#| - "**Wide horizontal bars**, unaligned"
+#| layout-nrow: 1
+#| out-width: 100%
+knitr::include_graphics("images/round3/horizontal-ab.PNG")
+knitr::include_graphics("images/round3/hwide-ab.PNG")
 ```
@@
 
--   When combining responses from multiple rounds, we make weighting adjustments to ensure that each weights in each sample are properly calibrated to the population total across both rounds.
+-   When combining responses from multiple rounds, we make weighting adjustments to ensure that weights in each sample are properly calibrated,  to the population total across both rounds.
     **NEED ED REVIEW OF THIS LANGUAGE.** This adjustment is completed as follows: \[insert Heike text from analysis doc here\].
@@
 
--   Design choices impact viewer behavior
-
--   Design choices impact viewer accuracy
-
--   Reproduced prior convenience sample results using a large, nationally-representative survey population
-
 -   Surveys can be used to ask these types of questions -- in the midst of other topics and with a limited number of questions, we are able to ask perception questions and produce results consistent with prior studies
 
+-   Reproduced (some) prior convenience sample results using a large, nationally-representative survey population:
+
+    - Cleveland: position (aligned) versus unaligned
+    - unaligned tiles within stacked bar are a hybrid between framed bar and floats
+    - Talbot: pie charts - individual wedges vs piecharts
+
+-   Design choices impact viewer accuracy:
+
+    - new finding: pie charts accuracy in work better when framed (wedges)
+
+- Other important measures beyond accuracy: time to completion (automatic on web), certainty in response (asked).
+
 -   Studying how viewers interact with charts can be done in a survey, and more expansive work on this topic should be completed to understand how the general public interacts with and understands charts.
+
+    - Design choices impact viewer behavior: zooming
+
+- (some) Demographics matter: age and gender does not seem to matter for perception
+  economic background and education does matter for perception: we need to know, we do not want to create a hurdle in communication
 

### 2023-07-17 analysis/analysis.qmd
demographics model
demographics by task

--- 2023-07-17 analysis/analysis.qmd
@@
 
+### Accuracy and Demographics
+
+Let $Y_k$ be the response of participant $k$, on a scale from 1 = 'wrong', 2 = 'they are the same' to 3 = 'correct'.
+We use a generalized cumulative logistic regression, where $\mu_i$ are intercepts $1 \le i < 3$, $X_k$ are demographics of the $k$th participant (in form of the model matrix), and $\beta_{i}$ are the coefficients.
+$$
+\text{ logit } P(Y_k \le i \mid X_k) = \mu_i + X_{k}'\beta_{i}
+$$
+
+```{r}
+#| label: fig-demographics
+#| fig-cap: "Accuracy across demographics.
+Gender and age are not significant factors.
+However, educational level and income do have -- a small -- effect.
+When income levels increase, the percentage of wrong answers decreases.
+#| With increasing levels in education, we see a small drop in correct answers, as well as a small drop in wrong answers, for a significant increase in 'they are the same'."
+#| fig-height: 4.25
+#| fig-width: 7
+#| out-width: 100%
+
+educ <- round_123_long %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = educ5)) + 
+  facet_grid(.~task) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="bottom",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") 
+gender <- round_123_long %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = gender)) + 
+  facet_grid(.~task) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="bottom",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") 
+income <- round_123_long %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = income4)) + 
+  facet_grid(.~task) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="bottom",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("")
+
+race <- round_123_long %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = racethnicity)) + 
+  facet_grid(.~task) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="bottom",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("")
+
+age <- round_123_long %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = age4)) + 
+   facet_grid(.~task) + 
+ geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="bottom",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("")
+
+gender + age +income + educ + plot_layout(nrow=1)
+```
+
+```{r demographics}
+logit_demo <- svyglm(correct~task:(gender+educ5+age4+income4), design=survey_123, family=quasibinomial())
+logit_demo3 <- svyglm(correct3~task:(gender+educ5+age4+income4), design=survey_123, family=quasibinomial())
+
+summary(logit_demo)
+summary(logit_demo3)
+```
+
 ## Model X: Comparing Aligned and Unaligned Elements Facets {#sec-modelX}

### 2023-07-17 SDSS-2023/testing-charts-accuracy.qmd
pdf updated

@@
 A and C are of identical size, as are B and D.
-While the predominant response for the unaligned pair on the left is 'they are of the same size', more than half of the viewers respond with 'D is bigger' to the aligned pair of bars on the right."
+When participants are asked to compare the size of these tiles in barcharts, the predominant response for the unaligned pair on the left is 'they are of the same size'.
+In contrast, more than half of the viewers respond with 'D is bigger' to the aligned pair of bars on the right."
 #| fig-subcap:
@@
 knitr::include_graphics("images/round3-modified.PNG")
-
 ```

### 2023-07-17 SDSS-2023/testing-charts-accuracy.qmd
filling in study design - stimulus section

@@
 ```
+<!-- fig-width: 6.5 -->
 
-<!-- fig-width: 6.5 -->
 <!-- fig-height: 4.5 -->
+
 <!-- template: template-jdsart.cls.tex -->
@@
 
+    -   \kr{Include somewhere in here:} The advantage of using a probability-based approach is two-fold.
+        First, we have access to a large sample of survey participants and thus have greater power in making inference about graphical perceptional abilities.
+        Second, the sample is representative of the general adult public in the U.S., which is an important target audience for scientific communication.
+
 -   This paper presents a series of tests completed and the resulting findings.
@@
 
-Outline for study design - stimulus
+Each task is made up of two elements: a visual stimulus and a question about that image that the viewer is asked to respond to.
+In our study, each visual stimulus is an image of a data visualization, while each question prompts viewers to identify which of two marked pieces in the data visualization is larger.
 
--   Use of Just-Noticeable Difference
+**What specifically is each task?** Each comparison between marked pairs is designed to be a difficult task, with the difference between the values represented in the two marked pieces being a just-noticeable difference.
+The **Just-Noticeable Difference** (JND) is defined as the smallest difference that will be detected 50% of the time.
+Prior results from studies on bar charts and pie charts [@luModelingJustNoticeable2022] inform the differences in charts shown to our survey panelists.
 
-    -   Differences between cognitive lab and survey
+**Why do we utilize just-noticeable differences?** We employ comparisons at the JND in our tasks in order to maximize our ability to identify the impact of design changes on viewer accuracy and behavior.
+Asking perception tasks in a survey differs from the controlled environment of a cognitive lab, where these kind of questions may usually be assessed.
+Rather than asking the same (or similar) type of question with varied signal strength multiple hundred times, we are limited to only a few questions at a time.
+With a small set of tasks, we need to present tasks that are perceptually hard, and thus ask questions about stimuli that are close to our perceptual threshold.
+Therefore, we focus on questions which vary the presented image, but ask viewers to compare the same values across those varied images.
 
-    -   Understanding how design impacts the most difficult comparisons
+**How do we vary the task?** We ask participants to determine which of two just-noticeably different marked pieces is larger within the data visualization image, and we vary the structural design of that data visualization image.
+We focus on three main sets of structural variation in the design.
+First, we vary the alignment of the pieces in question.
+Viewers are presented with two marked pieces in a chart that do not share a common baseline, then two pieces that do share a common baseline.
+Second, we vary the orientation of the chart -- a vertically oriented stacked bar chart, a horizontally oriented version of the same chart, and then a wider version of the horizontally oriented chart.
+\kr{We need to finalize what we end up showing in the results and then revisit this section...} Third, we consider a facetted bar chart and facetted pie chart.
+
+**How do we present the task?** The format of a survey guides the format and design of the questions asked and how they are presented to respondents.
+First, participant instructions must be delivered in a very short and easily understandable format, because participants cannot ask clarifying questions about the task as they might be able to in a cognitive lab setting.
+\kr{I'm not sure how important this second point is...} Second, participants should be given some context for the tasks they are being asked to complete; in a given round, our set of tasks appear as a group in the midst of other survey questions and topics, and without providing respondents some transition we risk a jarring shift and low respondent engagement in the task.
+Finally, to prevent viewers from being exposed to slight variations of the same stimulus in a row (and risk unforeseen order effects or respondents using prior questions to inform their responses), we either split a survey sample in two and show each subsample a distinct version of the chart or test variations of a chart across distinct rounds of the survey.
+
+**We vary the task by presenting viewers with different structural representations of the same data.**
 
@@
 
--   Structural variations
-
-    -   Alignment -- across all tests
-
-    -   Stacked bar chart -- vertical, horizontal, and horizontal wide
-
-        -   Vertical and horizontal bars are the same exact visual stimuli but rotated/presented differently
-
-    -   Facetted bar chart
-
-    -   Facetted pie chart -- 4 colors, one color framed, one color floating
-
 -   Results are organized as follows:
 
-    -   ....
-        Haven't decided yet.
-
-<!-- motivate the use of just noticeable differences -->
-
-Asking perceptual questions in a survey is different from the controlled environment of a cognitive lab, where these kind of questions would usually be addressed.
-This means, that instructions to participants have to be delivered in a very short and easily understandable, because questions arising from the task can not be answered.
-Similarly, rather than asking the same (or similar) type of question with varied signals multiple hundred times, in a survey we can ask only a few questions.
-In order to observe any effect, we need to ask questions that are perceptually hard, which means that we need to ask questions about stimuli that are close to our perceptual threshold.
-The **Just-Noticeable Difference** (JND) is defined as the smallest difference that will be detected 50% of the time.
-We are using results from studies on barcharts and pie charts [@luModelingJustNoticeable2022] to inform the differences in charts shown to survey panelists.
+    -   .... Haven't decided yet.
 
@@
 The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households **CITE DENNIS 2022**.
-Each test was conducted using the AmeriSpeak Omnibus survey, which runs biweekly and samples around 1,000 U.S.adults to answer questions on a variety of topics.
-The advantage of using a probability-based approach is two-fold.
-First, we have access to a large sample of survey participants and thus have greater power in making inference about graphical perceptional abilities.
-Second, the sample is representative of the general adult public in the U.S., which is an important target audience for scientific communication.
+Each test was conducted using the AmeriSpeak Omnibus survey, which runs biweekly and samples around 1,000 U.S. adults to answer questions on a variety of topics.
 
-\hh{XXX I think what you are trying to say is that we do not know, if participants have seen the stimuli multiple times, but if they did, there was at least a month in between?}
+\hh{XXX I think what you are trying to say is that we do not know, if participants have seen the stimuli multiple times, but if they did, there was at least a month in between?} \kr{Yes that's what I was trying to say here!!}
 
@@
 
-ROUND 5: Horizontal wide gridlines (only dark grid split sample)
-
 ROUND 6: Facetted bar (split sample w/o forcing choice)
@@
 ROUND 7: Aligned vs unaligned pie (full sample)
+
+# Notes on Analysis
+
+I need a better name for this section...
+
+Paragraph on applying survey weights
+
+-   All calculations in this paper are done in R [@RLanguage] using the `survey` package [@lumleyAnalysisComplexSurvey2004] version 4.0 [@survey] based on @lumleyComplexSurveysGuide2010.
+
+Paragraph on combining samples across rounds
+
+-   Pull from analysis doc
+
+-   When combining responses from multiple rounds, we make weighting adjustments to ensure that weights in each sample are properly calibrated to the population total across both rounds so each round is correctly represented in the resulting model [@omuircheartaighCombiningSamplesVs2002].
+    **NEED ED REVIEW OF THIS LANGUAGE.** This adjustment is completed as follows: \[insert Heike text from analysis doc here\].
 
@@
 
-Paragraph on combining samples across rounds
-
--   Pull from analysis doc
-
--   When combining responses from multiple rounds, we make weighting adjustments to ensure that weights in each sample are properly calibrated,  to the population total across both rounds.
-    **NEED ED REVIEW OF THIS LANGUAGE.** This adjustment is completed as follows: \[insert Heike text from analysis doc here\].
-
 Test 1
@@
 
-    -
+    -   Differences in accuracy across demographic groups
 
@@
 -   Structure of stacked bars
+    -   Accuracy and responses
+
+    -
 
@@
 -   Facetted bars and pies
+    -   Accuracy and responses
+
+    -   Timing of responses
 
@@
 -   Differences in timing and zooming behavior across all designs
-
-COMBINING SAMPLES AND WEIGHTING NOTES:
-
-We can combine responses across samples into one combined dataset, but we need to adjust weights accordingly so that each sample is weighted equally in the model [@omuircheartaighCombiningSamplesVs2002].
-
-All calculations in this paper are done in R [@RLanguage] using the `survey` package [@lumleyAnalysisComplexSurvey2004] version 4.0 [@survey] based on @lumleyComplexSurveysGuide2010.
 
@@
 
-- (some) Demographics matter: age and gender does not seem to matter for perception
-  economic background and education does matter for perception: we need to know, we do not want to create a hurdle in communication
+-   (some) Demographics matter: age and gender does not seem to matter for perception economic background and education does matter for perception: we need to know, we do not want to create a hurdle in communication
 

### 2023-07-18 SDSS-2023/testing-charts-accuracy.qmd
adding introduction text and doing some restructuring of the middle sections

@@
 
-Outline for introduction:
+What do viewers see when we show them a data chart?
+A data chart -- at its core -- maps quantitative values to graphical elements representing their relative values.
+Modern data visualizations are much more than a simple, objective mapping of values to a plane; they contain contextual and design elements, and are often structured to support the viewer in understanding a particular view of a set of data or specific pattern underlying the values.
+The design of a data visualization impacts a viewer's ability to achieve that understanding; a poorly designed data visualization may leave viewers struggling to understand the content or context, or make it difficult to complete accurate and useful comparison of values across groups or time points.
+More broadly, the design of a data visualization can change how viewers interact with the chart.
 
--   What do viewers see when we show them a data chart?
+A crucial step in the process of interacting with and understanding a chart is the viewer's employment of comparisons of the parts within.
+@clevelandGraphicalPerceptionTheory1984 observed as such, and in their seminal study defined the better visual among a pair as the one that allows viewers to make more accurate comparisons.
+Based on mappings of quantitative variables to different graphical elements, Cleveland and McGill's study resulted in a ranking of perceptual tasks from most accurate to least accurate, which was then extended by @mackinlayAutomatingDesignGraphical1986 to a theoretical framework ranking tasks' order along their ordinal and nominal scales, as shown in @fig-perceptual-tasks.
 
--   Visuals allow us to make comparisons between values
+Cleveland and McGill's work -- while a foundational user study in graphical perception -- utilized a small convenience sample, consisting of only a few individuals recruited from among the authors' coworkers and their spouses.
+@heerCrowdsourcingGraphicalPerception2010 reproduced Cleveland and McGill's rankings using a larger sample from a crowd sourcing platform.
+A total of XXX Amazon Mechanical Turkers were involved.
+Crowd sourced samples were shown by @borgoCrowdsourcingInformationVisualization2017 to be biased towards more male, younger, and relatively higher education relative to the adult U.S.
+population as a whole.
+These study populations are thus not representative of the general population, a common target audience for data visualization and scientific communication work.
+Further, the populations' emphasis on higher education individuals also leads to results which hold for groups of individuals who may be more likely to have prior exposure to data visualization in the context of scientific communication, or more exposure to data topics in higher education, but may not hold across other groups within the population.
+\kr{I don't know if the prior statement is going too far? Maybe we can rephrase.}
 
--   Design of a data visualization impacts viewer ability to make those comparisons
+<!-- Bias from crowd-sourcing? sample size?  -->
 
-    -   And more broadly, impacts how viewers interact with a chart
+\kr{The mackinlay thing feels like maybe we are overemphasizing the ranking done by cleveland/mcgill and then mackinlay - I think we could emphasize less if we want}
 
--   There is prior work studying this topic
+Our work -- as that of Cleveland and McGill and Heer and Bostock -- centers around studying data visualization design choices and their impact on viewer behavior and accuracy of viewers' responses.
+We seek to answer whether it is possible to reproduce some of their findings in the context of a survey with a large, nationally-representative set of respondents, and within that context we focus on the following research questions:
 
-    -   Cleveland and McGill laid the foundations
+1.  How do structural design choices in a data visualization impact viewers' ability to identify the larger of two elements?
 
-    -   Mackinlay expanded to theoretical scales
+2.  How is viewer interaction with the task impacted by structural design choices in a data visualization?
 
-    -   Heer/Bostock expanded to larger study population
+We employ a probability-based survey panel and run a series of perception tests with nationally-representative samples of respondents from that panel.
+The advantage of using a probability-based approach is two-fold.
+First, we have access to a large sample of survey participants and thus have greater power in making inference about graphical perceptional abilities.
+Second, the sample is representative of the general adult public in the U.S., which is an important target audience for scientific communication, and this allows us to test whether prior results from convenience samples hold with a nationally representative sample.
+\kr{Do we want a stronger statement about this here?}
 
--   However, prior work has utilized specific populations of study respondents.
-    These populations are not representative of the general population, a common target audience for data visualization and scientific communication work
-
--   Our work focuses on the following:
-
-    -   Utilizing a survey panel as a mechanism to test perception in data visualizations
-
-    -   Testing whether prior results from convenience samples hold with a nationally representative sample
-
-    -   Studying variations in data visualization design and their impact on respondent behavior and accuracy of participants' responses.
-
-    -   \kr{Include somewhere in here:} The advantage of using a probability-based approach is two-fold.
-        First, we have access to a large sample of survey participants and thus have greater power in making inference about graphical perceptional abilities.
-        Second, the sample is representative of the general adult public in the U.S., which is an important target audience for scientific communication.
-
--   This paper presents a series of tests completed and the resulting findings.
-
-    -   We consider structural (and aesthetic?) variations on bar charts and pie charts
-
-    -   Our tests are centered around the following central research questions:
-
-        -   How do structural design choices in a data visualization impact viewers' ability to identify the larger of two elements?
-
-        -   How is viewer behavior (zooming, time spent on question, certainty of response) impacted by structural and aesthetic design choices in a data visualization?
-
--   The remainder of the paper is organized as follows:
-
-    -   we first describe the stimulus design
-
-    -   then we describe the study population (and survey design?)
-
-    -   subsequently, we share analyses of resulting survey responses across a series of stimuli
-
-    -   finally, we discuss implications of this study and next steps.
-
-What do viewers see, when we show them a data chart?
-A crucial step in the process of 'understanding' a chart, is that a visual allows us to make comparisons between its parts.
-@clevelandGraphicalPerceptionTheory1984 defined for the purpose of their seminal study the better visual as the one that allows viewers to make more accurate comparisons.
-Based on mappings of quantitative variables to different graphical elements, Cleveland and McGill's study resulted in a ranking of perceptual tasks from most accurate to least accurate.
-@mackinlayAutomatingDesignGraphical1986 took this approach and expanded the tasks and their order to ordinal and nominal scales, as shown in @fig-perceptual-tasks.
+We present viewers with structural variations on bar charts and pie charts and ask them to answer questions comparing the size of elements within those charts.
+In this work, we present the series of tests we completed and the resulting findings.
+The remainder of the paper is organized as follows: first, we describe the design of the visual stimulus used in our perception tests.
+We then describe the population of study respondents and obtained survey sample.
+Subsequently, we share our analyses of the resulting survey responses across each of our tests, including analyses on accuracy of responses and response behavior.
+Finally, we discuss implications of this work and next steps.
 
@@
 
-What we need to remember, is that Mackinlay's work is purely theoretical.
-Cleveland and McGill's rankings based on a study.
-However, this study is a small convenience sample, consisting of only a few individuals recruited from among the authors' coworkers and their spouses.
-@heerCrowdsourcingGraphicalPerception2010 have reproduced Cleveland and McGill's rankings using a crowd sourcing platform.
-A total of XXX amazon turkers were involved.
-
-<!-- Bias from crowd-sourcing? sample size?  -->
-
-Crowd workers are known to be biased towards male, young and relatively higher education: @borgoCrowdsourcingInformationVisualization2017
-
-Here, we seek to -- first -- answer the question whether it is possible to use a survey to reproduce (some of) the rankings.
+\kr{The figure below is still a bit of an orphan, but I think we can revisit the intro ad study setup after we get more of the results in the paper}
 
@@
 Asking perception tasks in a survey differs from the controlled environment of a cognitive lab, where these kind of questions may usually be assessed.
-Rather than asking the same (or similar) type of question with varied signal strength multiple hundred times, we are limited to only a few questions at a time.
+Rather than asking the same (or similar) type of question with varied signal strength dozens or hundreds of times, we are limited to only a few questions at a time.
 With a small set of tasks, we need to present tasks that are perceptually hard, and thus ask questions about stimuli that are close to our perceptual threshold.
@@
 
-**We vary the task by presenting viewers with different structural representations of the same data.**
+\kr{Somewhere here we should also talk about how we measure viewer behavior/interaction with the chart: certainty, time spent, zooming, devices, etc. }
 
--   Stimuli within context of data viz
+### Test 1: Alignment in stacked bar charts
 
-    -   Title, legend, axes, etc.
-
--   Results are organized as follows:
-
-    -   .... Haven't decided yet.
-
--   How do structural design choices in a data visualization impact viewers' ability to identify the larger of two elements?
-
--   How do aesthetic design choices in a data visualization impact viewers' ability to identify the larger of two elements?
-
--   How is viewer behavior (zooming, time spent on question, certainty of response) impacted by structural and aesthetic design choices in a data visualization?
-
-Structural design choices
-
-Mapping, Stacked bar, Vertical, Horizontal, Horizontal wide
-
-Facetted bar
-
-Only have a split sample for this
-
-Pie, Alignment
-
-We have this for all above mappings, but the setup is a little different for facetted bar
-
-Aesthetic design choices (structural choices seems stronger/there could be a lot to talk about there... should we skip aesthetic on this one?)
-
-Colors
-
-Use of gridlines
-
-Outcomes/responses for modeling:
-
-Binary accuracy (correct/incorrect -- 'they are the same' is incorrect here)
-
-Ordinal response (a/b/they are the same)
-
-Zooming behavior (zoomed/did not zoom)
-
-Time spent on question (continuous, in seconds)
-
-Certainty
-
-I've noted this below, but: how to model?
-Ordinal response?
-Binary (certain or very certain vs everybody else)?
-
-## Study design - survey population
-
-Participants were recruited as part of NORC's AmeriSpeak panel, which utilizes a probability-based sampling methodology and samples U.S.
-households from NORC's National Sample Frame that provides coverage of over 97% of U.S.
-households.
-The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households **CITE DENNIS 2022**.
-Each test was conducted using the AmeriSpeak Omnibus survey, which runs biweekly and samples around 1,000 U.S. adults to answer questions on a variety of topics.
-
-\hh{XXX I think what you are trying to say is that we do not know, if participants have seen the stimuli multiple times, but if they did, there was at least a month in between?} \kr{Yes that's what I was trying to say here!!}
-
-Given the nature of pulling a sample from the panel for each Omnibus round, there is a possibility that some participants may have been included in multiple rounds; however, our collected data is not a longitudinal or panel study, so we do not have repeated responses from the same participants across each of our tests.
-Each of the tests was run at least a month apart, so if respondents participated in multiple rounds, there was at least a one month gap between viewing each visual stimulus.
-
-## Survey setup - Stimulus description
-
-\autoref{fig-tasks-12} shows the two stacked barcharts shown to participants in the survey.
+\autoref{fig-tasks-12} shows the two stacked bar charts shown to participants in the first test.
 The marked tiles in each plot are 155 pixels apart.
@@
 
+<!-- |                                   |                                   | -->
+
+<!-- |-----------------------------------|-----------------------------------| -->
+
+<!-- | **Unaligned positions**           | **Aligned positions**             | -->
+
+<!-- | ![Unaligned positions marked A and B](../SDSS-2023/images/vertical-ab.PNG) | ![Aligned positions marked C and D](../SDSS-2023/images/vertical-cd.PNG) | -->
+
+<!-- : The two stacked barcharts every participant got to see.
+In each barchart, the two marked tiles are to be compared for their size.
+in both instances, the tile on the right is (very slightly) larger.
+{#tbl-tasks-12 tbl-colwidths="\[50,50\]"} -->
+
+### Test 2: Orientation of stacked bar charts
+
+Let's talk about the stacked bar flippy setup here
+
 ```{r}
@@
 
-<!-- |                                   |                                   | -->
+### Test 3: Alignment and orientation in pie charts
 
-<!-- |-----------------------------------|-----------------------------------| -->
+Let's talk about the pies here
 
-<!-- | **Unaligned positions**           | **Aligned positions**             | -->
+## Study design - survey population
 
-<!-- | ![Unaligned positions marked A and B](../SDSS-2023/images/vertical-ab.PNG) | ![Aligned positions marked C and D](../SDSS-2023/images/vertical-cd.PNG) | -->
+Participants were recruited as part of NORC's AmeriSpeak panel, which utilizes a probability-based sampling methodology and samples U.S.
+households from NORC's National Sample Frame that provides coverage of over 97% of U.S.
+households.
+The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households **CITE DENNIS 2022**.
+Each test was conducted using the AmeriSpeak Omnibus survey, which runs biweekly and samples around 1,000 U.S. adults to answer questions on a variety of topics.
 
-<!-- : The two stacked barcharts every participant got to see.
-In each barchart, the two marked tiles are to be compared for their size.
-in both instances, the tile on the right is (very slightly) larger.
-{#tbl-tasks-12 tbl-colwidths="\[50,50\]"} -->
+\hh{XXX I think what you are trying to say is that we do not know, if participants have seen the stimuli multiple times, but if they did, there was at least a month in between?} \kr{Yes that's what I was trying to say here!!}
 
-DATA THAT MAY BE INCLUDED IN ANALYSIS:
-
-ROUNDS 1-2: Color variations on vertical stacked bar, aligned and unaligned
-
-ROUND 3: Horizontal and horizontal wide, aligned and unaligned
-
-ROUND 6: Facetted bar (split sample w/o forcing choice)
-
-ROUND 7: Aligned vs unaligned pie (full sample)
+Given the nature of pulling a sample from the panel for each Omnibus round, there is a possibility that some participants may have been included in multiple rounds; however, our collected data is not a longitudinal or panel study, so we do not have repeated responses from the same participants across each of our tests.
+Each of the tests was run at least a month apart, so if respondents participated in multiple rounds, there was at least a one month gap between viewing each visual stimulus.
 
@@
 
-Test 1
+### Test 1
 
@@
 
-Test 2
+### Test 2
 
--   Structure of stacked bars
+-   Structure of stacked bars -- vertical, horizontal, horizontal wide
     -   Accuracy and responses
@@
 
-Test 3
+### Test 3
 
@@
     -   Timing of responses
-
-How should we group results?
-
--   By stimulus/structural variation?
-
--   By response (binary, ordinal, certainty, time spent, etc.)
-
--   By accuracy vs by behavior (zooming, time spent, etc.)
 

### 2023-07-18 analysis/analysis.qmd
table for demographics

--- 2023-07-17 analysis/analysis.qmd
@@
 
+### Accuracy and Demographics
+
+Let $Y_k$ be the response of participant $k$, on a scale from 1 = 'wrong', 2 = 'they are the same' to 3 = 'correct'.
+We use a generalized cumulative logistic regression, where $\mu_i$ are intercepts $1 \le i < 3$, $X_k$ are demographics of the $k$th participant (in form of the model matrix), and $\beta_{i}$ are the coefficients.
+$$
+\text{ logit } P(Y_k \le i \mid X_k) = \mu_i + X_{k}'\beta_{i}
+$$
+
+```{r}
+#| label: fig-demographics
+#| fig-cap: "Panelist's demographics matter, particularly, when the task difficulty increases.
+For aligned tiles, gender, age, and education are not significant factors.
+However, income levels do have - a small - effect.
+When income levels increase, the percentage of wrong answers (orange) decreases, while the percentage of correct answers (blue) increases slightly (significant at below 0.05).
+The more difficult task of comparing unaligned tiles, demographics are more significant.
+XXX With increasing levels in education, we see a small drop in correct answers, as well as a small drop in wrong answers, for a significant increase in 'they are the same'."
+#| fig-subcap: 
+#|   - Comparisons of aligned tiles
+#|   - Comparisons of unaligned tiles
+#| fig-height: 5
+#| fig-width: 7
+#| layout-nrow: 2
+#| out-width: 100%
+
+educ_cd <- round_123_long %>% 
+  filter(task=="cd") %>%
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = educ5)) + 
+  facet_grid(~task) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+    ggtitle("\nby education") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+
+educ_ab <- round_123_long %>% 
+  filter(task=="ab") %>%
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = educ5)) + 
+  facet_grid(.~task) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+    ggtitle("\nby education") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+
+gender_cd <- round_123_long %>% 
+  filter(task=="cd") %>%
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = gender)) + 
+  facet_grid(.~task) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+  ggtitle("Accuracy\nby gender") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+
+gender_ab <- round_123_long %>% 
+  filter(task=="ab") %>%
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = gender)) + 
+  facet_grid(.~task) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+  ggtitle("Accuracy\nby gender") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+
+income_cd <- round_123_long %>% 
+  filter(task=="cd") %>%
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = income4)) + 
+  facet_grid(.~task) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+    ggtitle("\nby income") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+
+income_ab <- round_123_long %>% 
+  filter(task=="ab") %>%
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = income4)) + 
+  facet_grid(.~task) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+    ggtitle("\nby income") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+
+race <- round_123_long %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = racethnicity)) + 
+  facet_grid(.~task) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+    ggtitle("Accuracy by race, ethnicity and task") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+
+age_cd <- round_123_long %>% 
+  filter(task=="cd") %>%
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = age4)) + 
+   facet_grid(.~task) + 
+ geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+    ggtitle("\nby age") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+
+age_ab <- round_123_long %>% 
+  filter(task=="cd") %>%
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = age4)) + 
+   facet_grid(.~task) + 
+ geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+    ggtitle("\nby age") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+gender_cd + age_cd +income_cd + educ_cd + plot_layout(nrow=1, guides = "collect")
+
+gender_ab + age_ab +income_ab + educ_ab + plot_layout(nrow=1, guides = "collect")
+```
+
+```{r demographics-model}
+logit_demo_task_cd_correct <- svyglm(correct~gender+educ5+age4+income4, design=survey_123, family=quasibinomial(), subset = task=="cd")
+logit_demo_task_cd_correct_same <- svyglm(correct3~gender+educ5+age4+income4, design=survey_123, family=quasibinomial(), subset = task=="cd")
+
+logit_demo_task_ab_correct <- svyglm(correct~gender+educ5+age4+income4, design=survey_123, family=quasibinomial(), subset = task=="ab")
+logit_demo_task_ab_correct_same <- svyglm(correct3~gender+educ5+age4+income4, design=survey_123, family=quasibinomial(), subset = task=="ab")
+
+coefs_cd_correct <- broom::tidy(logit_demo_task_cd_correct)
+coefs_cd_correct_same <- broom::tidy(logit_demo_task_cd_correct_same)
+
+coefs_ab_correct <- broom::tidy(logit_demo_task_ab_correct)
+coefs_ab_correct_same <- broom::tidy(logit_demo_task_ab_correct_same)
+```
+
+```{r demographics-table}
+ coefs_clean <- function(dframe) {
+  dframe %>% mutate(
+    variable = c("Intercept", "Gender", rep("Education", 4), rep("Age", 3), rep("Income", 3)),
+    term = ifelse(term=="(Intercept)", "", term),
+    term = gsub("gender", "", term),
+    term = gsub("educ5", "", term),
+    term = gsub("age4", "", term),
+    term = gsub("income4", "", term),
+    ci_low = exp(estimate-1.96*std.error),
+    ci_high = exp(estimate+1.96*std.error),
+    estimate = exp(estimate)
+  )
+}
+
+coefs_clean(coefs_cd_correct) %>% mutate(task = "Aligned", correct="correct") %>%
+ rbind(
+  coefs_clean(coefs_ab_correct) %>% mutate(task = "Unaligned", correct="correct")
+ ) %>%
+ rbind(
+  coefs_clean(coefs_cd_correct_same) %>% mutate(task = "Aligned", correct="not wrong")
+ ) %>%
+ rbind(
+  coefs_clean(coefs_ab_correct_same) %>% mutate(task = "Unaligned", correct="not wrong")
+ ) %>%
+ dplyr::select(-std.error, -statistic) %>%
+ mutate(
+  p.value = ifelse (p.value <  0.001, "***",
+      ifelse (p.value <  0.01, "**",
+             ifelse (p.value < 0.05, "*",
+                    ifelse (p.value < 0.1, ".", ""))))
+ ) %>%
+ group_by(variable) %>%
+ pivot_wider(
+    id_cols = c(term, variable),
+    names_from = c(task, correct),
+    values_from = c(estimate, ci_low, ci_high, p.value),
+  ) %>% 
+ gt() %>%
+  tab_header(title = "Log odds of accuracy by task and demographics of respondents") %>%
+ tab_options(
+   column_labels.font.weight = "bold",
+   row_group.font.weight = "bold"
+ ) %>%
+  fmt_number(
+    decimals=2
+  ) %>%
+  cols_width(
+    contains("p.value") ~ px(20)
+  ) %>%
+  cols_merge(
+    columns = contains("ci") & contains("_Aligned_correct"),
+    pattern = "<< [{1}, {2}]>>"
+  ) %>%
+  cols_merge(
+    columns = contains("ci") & contains("_Aligned_not"),
+    pattern = "<< [{1}, {2}]>>"
+  ) %>%
+  cols_move_to_end(contains("_Aligned_")) %>%
+  tab_spanner(
+    label = "correct | same or wrong",
+    columns = contains("_Aligned_correct")
+  )  %>%
+  tab_spanner(
+    label = "correct or same | wrong",
+    columns = contains("_Aligned_not")
+  ) %>%
+  tab_spanner(
+    label = "Aligned tiles",
+    columns = contains("_Aligned_"),
+    spanners = c("correct | same or wrong", "correct or same | wrong")
+  ) %>%
+  cols_move_to_end(contains("_Unaligned_")) %>%
+  cols_merge(
+    columns = contains("ci") & contains("_Unaligned_correct"),
+    pattern = "<< [{1}, {2}]>>"
+  ) %>%
+  cols_merge(
+    columns = contains("ci") & contains("_Unaligned_not"),
+    pattern = "<< [{1}, {2}]>>"
+  ) %>%
+  tab_spanner(
+    label = "correct | same or wrong ",
+    columns = contains("_Unaligned_correct")
+  )  %>%
+  tab_spanner(
+    label = "correct or same | wrong ",
+    columns = contains("_Unaligned_not")
+  ) %>%
+  tab_spanner(
+    label = "Unaligned tiles",
+    columns = contains("_Unaligned")
+  ) %>%
+   cols_label(
+    .list = c(
+      "estimate_Aligned_correct" = "Est.",
+      "ci_low_Aligned_correct" = "          [95% CI]",
+      "p.value_Aligned_correct" = " ",
+      "estimate_Aligned_not wrong" = "Est.",
+      "ci_low_Aligned_not wrong" = "          [95% CI]",
+      "p.value_Aligned_not wrong" = " ",
+      "estimate_Unaligned_correct" = "Est.",
+      "ci_low_Unaligned_correct" = "          [95% CI]",
+      "p.value_Unaligned_correct" = " ",
+      "estimate_Unaligned_not wrong" = "Est.",
+      "ci_low_Unaligned_not wrong" = "          [95% CI]",
+      "p.value_Unaligned_not wrong" = " "
+    )
+  )
+
+```
+
 ## Model X: Comparing Aligned and Unaligned Elements Facets {#sec-modelX}

### 2023-07-18 analysis/analysis.qmd
order of terms in model

@@
 age_ab <- round_123_long %>% 
-  filter(task=="cd") %>%
+  filter(task=="ab") %>%
   mutate(
@@
   scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
-gender_cd + age_cd +income_cd + educ_cd + plot_layout(nrow=1, guides = "collect")
-
-gender_ab + age_ab +income_ab + educ_ab + plot_layout(nrow=1, guides = "collect")
+gender_cd + age_cd  + educ_cd +income_cd + plot_layout(nrow=1, guides = "collect")
+
+gender_ab + age_ab  + educ_ab +income_ab + plot_layout(nrow=1, guides = "collect")
 ```
@@
 ```{r demographics-model}
-logit_demo_task_cd_correct <- svyglm(correct~gender+educ5+age4+income4, design=survey_123, family=quasibinomial(), subset = task=="cd")
-logit_demo_task_cd_correct_same <- svyglm(correct3~gender+educ5+age4+income4, design=survey_123, family=quasibinomial(), subset = task=="cd")
-
-logit_demo_task_ab_correct <- svyglm(correct~gender+educ5+age4+income4, design=survey_123, family=quasibinomial(), subset = task=="ab")
-logit_demo_task_ab_correct_same <- svyglm(correct3~gender+educ5+age4+income4, design=survey_123, family=quasibinomial(), subset = task=="ab")
+logit_demo_task_cd_correct <- svyglm(correct~gender+age4+educ5+income4, design=survey_123, family=quasibinomial(), subset = task=="cd")
+
+logit_demo_task_cd_correct_same <- svyglm(correct3~gender+age4+educ5+income4, design=survey_123, family=quasibinomial(), subset = task=="cd")
+
+logit_demo_task_ab_correct <- svyglm(correct~gender+age4+educ5+income4, design=survey_123, family=quasibinomial(), subset = task=="ab")
+logit_demo_task_ab_correct_same <- svyglm(correct3~gender+age4+educ5+income4, design=survey_123, family=quasibinomial(), subset = task=="ab")
 

### 2023-07-18 analysis/analysis.qmd
table stubs

@@
   dframe %>% mutate(
-    variable = c("Intercept", "Gender", rep("Education", 4), rep("Age", 3), rep("Income", 3)),
+    variable = c("Intercept", "Gender", rep("Age", 3), rep("Education", 4), rep("Income", 3)),
     term = ifelse(term=="(Intercept)", "", term),

### 2023-07-19 SDSS-2023/testing-charts-accuracy.qmd
re-ran the paper as pdf. The colors help a lot :)

@@
 \kr{We need to finalize what we end up showing in the results and then revisit this section...} Third, we consider a facetted bar chart and facetted pie chart.
+
+\hh{Not sure where should put this, parking it here for right now:}
+\hh{expectations - start}
+What we call 'aligned' and 'unaligned', here, is similar to Cleveland and McGill's set of rankings, but with some modifications:
+both 'aligned' and 'unaligned' bars (in tests 1 and 2) or wedges (in test 3) share the same axis.
+Aligned tiles are additionally anchored in the same position in one dimension, i.e.
+the difference between their sizes can be reduced to a positional assessment.
+Unaligned tiles do not share this anchor, however, the context of the other tiles in the chart provide a frame, which *should* help with an assessment of the tiles' sizes beyond a comparison of (arc) lengths or areas.
+We would expect that comparing unaligned tiles is a harder task (with correspondingly lower levels of accuracy) than a comparison of aligned tiles, with the framing given by the context of the other tiles in the same column (or the same pie) mitigating some of this difficulty.
+@fig-tasks-explain gives an overview of the comparisons of tasks 1 through 3 and the closest corresponding tasks in Cleveland and McGill.
+
+```{r}
+#| label: fig-tasks-explain
+#| fig-cap: Comparisons made in charts in the Cleveland and McGill ranking
+
+cat("still working on the chart")
+```
+
+\hh{expectations - end}
 

### 2023-07-19 SDSS-2023/testing-charts-accuracy.qmd
starting to pull in content from the analysis file.

--- 2023-07-18 SDSS-2023/testing-charts-accuracy.qmd
@@
 Answer 2 is the correct answer for both charts.
-Both charts are shown at the same size, i.e. in both cases the difference in size between the bars is exactly the same, the vertical distance between the bars is the exact same amount.
+Both charts are shown at the same size, i.e. in both cases both the difference in size between the bars and the horizontal distance between the bars is the exact same amount.
 This leaves the vertical positioning of the bars as the only difference between the charts.
-Any differences in observed accuracy can therefore be attributed to this difference in presentation.
+Any differences in observed responses can therefore be attributed to this difference in presentation.
 
@@
 
-Let's talk about the stacked bar flippy setup here
+In \autoref{fig-tasks-3} we see two structural variations on the stacked bar chart seen in Test 1: first, the Task 1 chart sized equally but rotated 90 degrees with bar lengths shown horizontally.
+Second, the horizontal chart with a smaller aspect ratio so the bars appear thinner and longer.
+Note that the wide version of the horizontal bars increases the pixel difference between the two marked pieces relative to the standard versions of the horizontal and vertical bars.
+
+This test utilized a split sample approach, where 50% of participants in the sample were shown the tall horizontal bar chart and 50% were shown the wide horizontal chart.
+Participants were asked to complete the same task as in Test 1; selecting the correct response among:
+
+1.  A is bigger
+2.  B is bigger
+3.  They are the same
+
+Again, answer 2 is the correct answer.
 
@@
 
-## Study design - survey population
+## Study design -- participants
 
@@
 
-# Notes on Analysis
+## Study design -- survey weighting
 
-I need a better name for this section...
+Paragraph on how survey weights are derived?
+-- **Question for Ed -- is this something AmeriSpeak can provide boilerplate language for?**
 
-Paragraph on applying survey weights
+All calculations in this paper are done in R [@RLanguage], and weights are applied in analyses using the `survey` package [@lumleyAnalysisComplexSurvey2004] version 4.0 [@survey] based on @lumleyComplexSurveysGuide2010.
 
--   All calculations in this paper are done in R [@RLanguage] using the `survey` package [@lumleyAnalysisComplexSurvey2004] version 4.0 [@survey] based on @lumleyComplexSurveysGuide2010.
+When combining responses that were gathered during distinct rounds of the Omnibus survey, we make weighting adjustments to ensure that weights in each sample are properly calibrated to the population total across both rounds in the resulting model.
+We *combine* (rather than cumulate) surveys $S_1$ and $S_2$, as described in @omuircheartaighCombiningSamplesVs2002, by
+multiplying weights in $S_1$ and $S_2$ by $\lambda$ and $1 - \lambda$ respectively.
+$$
+\lambda = \frac{n_1/d_1}{n_1/d_1 + n_2/d_2},
+$$
+where $n_1$ and $n_2$ are the nominal sample sizes and $d_1$ and $d_2$ are the design effects for the estimators. 
+Here, $d_1$ and $d_2$ are estimated as 
+$$
+d_1 = 1 + CV(w_i \in S_1)^2 \ \ \ \text{ and } \ \ \ d_2 = 1 + CV(w_i \in S_2)^2
+$$
+where $CV$ is the coefficient of variation of the weights within each sample, and is estimated as in @kishSurveySampling1965 :
 
-Paragraph on combining samples across rounds
+$$ 
+CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
+$$
 
--   Pull from analysis doc
-
--   When combining responses from multiple rounds, we make weighting adjustments to ensure that weights in each sample are properly calibrated to the population total across both rounds so each round is correctly represented in the resulting model [@omuircheartaighCombiningSamplesVs2002].
-    **NEED ED REVIEW OF THIS LANGUAGE.** This adjustment is completed as follows: \[insert Heike text from analysis doc here\].
+Note that @omuircheartaighCombiningSamplesVs2002 estimate $\lambda$ separately for any combination of race/ethnicity by sex.
+We employ that strategy whenever we include demographic variables in the analysis, otherwise we will use a single adjustment for the weights.
 
@@
 
-## Respondents
+### Respondents
 
@@
 
--   raw sample size and effective sample size
+-   raw sample size and effective sample size that we are analyzing **in each test**
 
@@
 
-    -   Accuracy and responses
+    -   Binary accuracy in responses
 
-    -   Differences in accuracy across demographic groups
+        -   Figure: binary correctness and binary ggpcp diagram
+        -   Table:(?) T-test results
+        -   Table: results of cell means model on certainty
 
-### Test 2
+    -   Ordinal accuracy in responses
 
--   Structure of stacked bars -- vertical, horizontal, horizontal wide
-    -   Accuracy and responses
+        -   Figure: ordinal correctness and ordinal ggpcp diagram
 
-    -
+        -   Table: Fitted model for ordinal response and alignment
 
-### Test 3
+    -   Ordinal accuracy in responses by demographics
 
--   Facetted bars and pies
-    -   Accuracy and responses
+        -   Figure: ordinal correctness by demographics, split by aligned and unaligned
 
-    -   Timing of responses
+        -   Table: Fitted model for ordinal response and alignment by demos
 
-Results to discuss:
-
--   Differences in accuracy between aligned and unaligned bar
-
--   Differences in accuracy between framed, floating, and general pie
-
--   Differences in timing and zooming behavior across all designs
+            -   Maybe this table should go in the appendix?
 
@@
 On the right (b), a parallel coordinate plot shows all combinations of responses.
-There's a huge asymmetry in the number of responses, where participants answered only one of the questions correctly.
+There's a huge asymmetry in the number of responses where participants answered only one of the questions correctly.
 A lot more responses are correct when comparing aligned tiles than unaligned tiles."
@@
 
-ANALYSIS PLAN:
+### Test 2
 
-Models below structured as:
+-   Structure of stacked bars -- vertical, horizontal, horizontal wide
+    -   Accuracy and responses
 
-Response
+    -
 
-Covariates to use in each model
+### Test 3
 
-STRUCTURAL VARIATION -- START HERE
+-   Facetted bars and pies
+    -   Accuracy and responses
 
-Binary accuracy across structural choices
+    -   Timing of responses
 
-Model 1:
+Results to discuss:
 
-Alignment only, just vertical stacked bar
+-   Differences in accuracy between aligned and unaligned bar
 
-Using the AmeriSpeak survey tool, a total of 1902 participants were exposed to two barcharts each, as shown in \autoref{tbl-tasks-12}.
+-   Differences in accuracy between framed, floating, and general pie
 
-Model 2:
-
-Alignment
-
-Bar vs pie (comparable question for pie is A vs B)
-
-Model 3:
-
-Alignment
-
-Vertical x horizontal x horizontal wide
-
-Model 4:
-
-Alignment
-
-Every structure (vertical bar, horizontal bar, horizontal wide bar, facet bar, pie)
-
-Visuals:
-
-\% yes across each different structural condition
-
-Facet by aligned/unaligned?
-
-Model estimates + CIs
-
-Ordinal response
-
-Model 1:
-
-Alignment only, just vertical stacked bar
-
-Model 2:
-
-Alignment
-
-Bar vs pie (comparable question for pie is A vs B)
-
-Model 3:
-
-Alignment
-
-Vertical x horizontal x horizontal wide
-
-Model 4:
-
-Alignment
-
-Every structure (vertical bar, horizontal bar, horizontal wide bar, facet bar, pie)
-
-Visuals:
-
-All responses across each different structural condition
-
-Facet by aligned/unaligned?
-
-Model estimates + CIs
-
-Zooming behavior (zoomed/did not zoom)
-
-Model 1:
-
-Device type
-
-Alignment
-
-Vertical x horizontal x horizontal wide
-
-Visuals:
-
-\% zoomed by device + alignment (already have this chart)
-
-Model estimates + CIs
-
-Time spent on question (in seconds)
-
-Model 1:
-
-Device type
-
-Zoom
-
-Alignment
-
-Vertical x horizontal x horizontal wide
-
-Model 2 (this may not be feasible for comparison depending on what level the 'TOTALTIME' is captured at):
-
-Device type
-
-Zoom
-
-Alignment
-
-Every structure (vertical bar, horizontal bar, horizontal wide bar, facet bar, pie)
-
-Visuals:
-
-Distribution of time spent variable
-
-Facet by device type, zoom, structural condition, alignment?
-Play around with it
-
-Average time spent by each of the conditions
-
-Certainty?
-
-Same models as above, but I'm not sure how we want to do the response.
-Ordinal response?
-Binary (certain or very certain vs everybody else)?
-
-AESTHETIC VARIATION -- ONLY IF TIME
-
-Binary accuracy (correct/incorrect -- 'they are the same' is incorrect here) across structural choices
-
-Model 1:
-
-Dark grid vs no grid (only have for horizontal wide)
-
-Response choice (ordinal response)
-
-Model 1:
-
-Dark grid vs no grid (only have for horizontal wide)
-
-Zooming behavior (zoomed/did not zoom)
-
-Model 1:
-
-Device type
-
-Dark grid vs no grid
-
-Time spent on question (in seconds)
-
-Model 1:
-
-Device type
-
-Zoom
-
-Dark grid vs no grid
-
-Certainty?
-
-Same models as above, but I'm not sure how we want to do the response.
-Ordinal response?
-Binary (certain or very certain vs everybody else)?
+-   Differences in timing and zooming behavior across all designs
 

### 2023-07-19 SDSS-2023/testing-charts-accuracy.qmd
dennis reference

--- 2023-07-19 SDSS-2023/testing-charts-accuracy.qmd
@@
 households.
-The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households **CITE DENNIS 2022**.
+The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households **CITE DENNIS 2022** @dennis2019technical. \hh{I don't think I've got the right citation}
 Each test was conducted using the AmeriSpeak Omnibus survey, which runs biweekly and samples around 1,000 U.S. adults to answer questions on a variety of topics.

### 2023-07-19 SDSS-2023/testing-charts-accuracy.qmd
references added, pdf updated

@@
 format:
-  html: default
   pdf: 
@@
     keep-tex: true
+  html: default
 editor: 
@@
 #| label: fig-tasks-explain
-#| fig-cap: Comparisons made in charts in the Cleveland and McGill ranking
+#| fig-cap: Comparisons made in charts across tasks 1 through 3 within the Cleveland and McGill ranking
 
-cat("still working on the chart")
+include_graphics("images/sketch-tasks.png")
 ```

### 2023-07-19 SDSS-2023/testing-charts-accuracy.qmd
a few wording changes on first figure

--- 2023-07-19 SDSS-2023/testing-charts-accuracy.qmd
@@
 households.
-The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households **CITE DENNIS 2022** @dennis2019technical. \hh{I don't think I've got the right citation}
+The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households **CITE DENNIS 2022**.
 Each test was conducted using the AmeriSpeak Omnibus survey, which runs biweekly and samples around 1,000 U.S. adults to answer questions on a variety of topics.
@@
 Rounds 1 and 2 are combined by adjusting the weights with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`.
-Figure \ref{fig-alignment}(a) shows that more than twice the number of responses is accurate, when the tiles are aligned along the same axis.
+Figure \ref{fig-alignment}(a) shows that more than twice the number of responses are correct, when the tiles are aligned along the same axis.
 Because each participant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts. 
@@
 #| fig-cap: "On the left (a), a stacked barchart shows the number of respondents with correct (green) and incorrect (grey) responses to the two comparison questions.
-When tiles are aligned along the same axis, more than twice the number of responses is accurate.
+When tiles are aligned along the same axis, more than twice the number of responses are correct.
 The shaded area along the top of the green tiles corresponds to  95\\% confidence intervals around (marginal) correct responses.

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
move data to the front and clean up a bit

--- 2023-07-19 SDSS-2023/testing-charts-accuracy.qmd
@@
 format:
+  html: default
   pdf: 
@@
     keep-tex: true
-  html: default
 editor: 
@@
 round1 <- readRDS(file.path(vs_repo_local, "round1.rds"))
-round2 <- readRDS(file.path(vs_repo_local, "round2.rds"))
 
@@
 
+round2 <- readRDS(file.path(vs_repo_local, "round2.rds"))
+
 round2_adj <- round2 %>% 
@@
 
-survey1 <- survey::svydesign(
-  data = round1_adj, 
-  ids = ~CaseId, weights = ~weight)
+round3 <- readRDS(file.path(vs_repo_local, "round3.rds"))
 
-survey2 <- survey::svydesign(
-  data = round2_adj,
-  ids = ~CaseId, weights = ~weight)
+# filter out all of the 'skipped on web':
+round3_adj <- round3 %>% mutate(
+  weight = WEIGHT_NOLA
+) %>% filter(
+  ab != "SKIPPED ON WEB",
+  cd != "SKIPPED ON WEB"
+)
+
+round6 <- readRDS(file.path(vs_repo_local, "round6.rds"))
+
+round6_adj <- round6 %>% mutate(
+  weight = WEIGHT_NOLA
+) %>% filter(RND_01 == 0) # same type of questions as round 7
+
+round7 <- readRDS(file.path(vs_repo_local, "round7.rds"))
+
+# filter out all of the 'skipped on web':
+round7_adj <- round7 %>% mutate(
+  weight = WEIGHT_NOLA
+)
+```
+
+```{r data-rnd3}
+round3 <- readRDS(file.path(vs_repo_local, "round3.rds"))
+
+# filter out all of the 'skipped on web':
+round3_adj <- round3 %>% mutate(
+  weight = WEIGHT_NOLA
+) %>% filter(
+  ab != "SKIPPED ON WEB",
+  cd != "SKIPPED ON WEB"
+)
 ```
@@
 
+The data for this paper were collected in several rounds as part of the NORC Omnibus.
+
+| Name    | Date  | # Participants | effective sample size | Sum of weights $\sum_i w_i$ |
+|---------|-------|--------:|----------:|----------:|
+| Round 1 | April 2022 |  `r nrow(round1_adj)`    |    `r round(neff(round1_adj$weight),1)`       | `r round(sum(round1_adj$weight),1)`       |
+| Round 2 | May 2022 |  `r nrow(round2_adj)`   |    `r round(neff(round2_adj$weight),1)`       | `r round(sum(round2_adj$weight),1)`       |
+| Round 3 | Jun 2022 |  `r nrow(round3_adj)`  |    `r round(neff(round3_adj$weight),1)`       | `r round(sum(round3_adj$weight),1)`       |
+| Round 6 | Sep 2022 |  `r nrow(round6_adj)` [Split]  |    `r round(neff(round6_adj$weight),1)`       | `r round(sum(round6_adj$weight),1)`       |
+| Round 7 | Oct 2022 |  `r nrow(round7_adj)`   |    `r round(neff(round7_adj$weight),1)`       | `r round(sum(round7_adj$weight),1)`       |
+|         |       |         |           |           |
+
+: Survey rounds: dates, number of participants (nominal sample size), effective sample size, and sum of weights. {#tbl-rounds}
+
 # Results
@@
 
-```{r lambdas}
-n1 <- nrow(round1_adj)
-n2 <- nrow(round2_adj)
-neff1 <- svytotal(~one, survey1, deff=TRUE)
-neff2 <- svytotal(~one, survey2, deff=TRUE)
-#abcd1 <- svytotal(~CorrectABCD, survey1, deff=TRUE)
-#abcd2 <- svytotal(~CorrectABCD, survey2, deff=TRUE)
+```{r combine_rounds_1_and_2}
 
-d1 <- 1 + cv(neff1)^2
-d2 <- 1 + cv(neff2)^2
-
-denom <- n1/d1 + n2/d2
-
-lambda <- n1/(d1*denom) 
-```
-
-```{r combine_rounds_1_and_2}
+lambda <- calculate_lambda(round1_adj$weight, round2_adj$weight)
 round_12 <- round1_adj %>% 

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
moved relevant code from analysis.qmd over

@@
 library(here)
+library(multcomp)
 
@@
 
+\hh{we might be able to include the lambda values for combining the tables here as well}
+
 | Name    | Date  | # Participants | effective sample size | Sum of weights $\sum_i w_i$ |
@@
             mutate(weight = lambda[1]*weight) %>% 
-            select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight) %>%
+            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight) %>%
     rbind(round2_adj %>% 
             mutate(weight = (1-lambda[1])*weight) %>% 
-            select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight))
+            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight))
 round_12_long <- round_12 %>% 
@@
 
-    -
+```{r combine-123}
+lambda_1_2 <- calculate_lambda(round1_adj$weight, round2_adj$weight)
+
+round_12 <- round1_adj %>% 
+            mutate(weight = lambda_1_2*weight/sum(weight),
+                   round = "Round 1") %>% 
+            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
+                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, weight,
+                   age4, gender, income4, racethnicity, educ5) %>%
+    rbind(round2_adj %>% 
+            mutate(weight = (1-lambda_1_2)*weight/sum(weight),
+                   round = "Round 2") %>% 
+            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, 
+                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, weight,
+                   age4, gender, income4, racethnicity, educ5))
+
+lambda_12_3 <- calculate_lambda(round_12$weight, round3_adj$weight)
+
+round_123 <- round_12 %>% 
+              mutate(weight = lambda_12_3*weight/sum(weight),
+                     chart = "Vertical") %>% 
+  rbind(
+    round3_adj %>% 
+      mutate(
+        weight = (1-lambda_12_3)*weight/sum(weight),
+        round = "Round 3",
+        CorrectAB = ab == "B is bigger",
+        CorrectCD = cd == "D is bigger",
+        CorrectABCD = paste(CorrectAB, CorrectCD),
+        chart = ifelse(P_EXP==1, "Horizontal", "Horizontal wide")
+       ) %>% 
+        dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
+                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, chart, weight,
+                   age4, gender, income4, racethnicity, educ5) 
+  )
+
+round_123$weight <- neff(round_123$weight)/sum(round_123$weight)*round_123$weight 
+# does not affect the further analysis, but helps with the interpretation of results
+
+round_123_long <- round_123 %>% 
+  pivot_longer(ab:cd, names_to = "task", values_to = "Response")
+
+round_123_long <- round_123_long %>% 
+  mutate(
+    correct = Response %in% c("B is bigger", "D is bigger"),
+    correct3 = correct | (Response == "They are the same"),
+    task_chart = factor(paste(task, chart, sep="-")),
+    zoom = ifelse(task=="ab", ab_zoom, cd_zoom),
+    response3 = ifelse(correct, "Correct", ifelse(correct3, "They are the same", "Wrong"))
+  )
+
+survey_123 <- 
+  survey::svydesign(
+  data = round_123_long, 
+  ids = ~CaseId, weights = ~weight)
+```
+
+We combine rounds 1, 2 and 3 as shown in Table above to incorporate all relevant responses to evaluate the three different designs of charts in terms of their accuracy in assessing small differences in tile sizes.
+
+```{r model-123, warning = FALSE, message = FALSE}
+
+logit_123_correct <- svyglm(correct~task_chart-1, family=quasibinomial(), design=survey_123)
+logit_123_correct3 <- svyglm(correct3~task_chart-1, family=quasibinomial(), design=survey_123)
+
+ci_confint <- svyby(~correct, by=~task_chart, design=survey_123, svyciprop)
+ci_confint3 <- svyby(~correct3, by=~task_chart, design=survey_123, svyciprop)
+
+coefs_123 <- broom::tidy(logit_123_correct)
+coefs_123_relaxed <- broom::tidy(logit_123_correct3)
+
+# create all pairwise comparisons for tasks and questions.
+cross_compare <- glht(logit_123_correct, mcp(task_chart="Tukey"))
+cross_compare3 <- glht(logit_123_correct3, mcp(task_chart="Tukey"))
+
+# get letters for each confidence 
+ci_confint$letters <- cld(cross_compare)$mcletters$Letters
+ci_confint3$letters <- cld(cross_compare3)$mcletters$Letters
+
+coefs_123 <- coefs_123 %>% left_join(
+  ci_confint %>%
+    mutate(
+      task_chart = paste0("task_chart",task_chart, sep="")
+    ) %>% dplyr::select(task_chart, letters), by=c("term"= "task_chart"))
+
+coefs_123_relaxed <- coefs_123_relaxed %>% left_join(
+  ci_confint3 %>% 
+    mutate(
+      task_chart = paste0("task_chart",task_chart, sep="")
+    ) %>% dplyr::select(task_chart, letters), by=c("term"= "task_chart"))
+
+ci_confint <- ci_confint %>% 
+  separate_wider_delim(
+    task_chart, delim = "-", names =c("task", "chart")
+    )  %>% mutate(
+  task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+  chart = factor(chart, levels =c("Vertical", "Horizontal", "Horizontal wide"))
+)
+
+ci_confint3 <- ci_confint3 %>% 
+  separate_wider_delim(
+    task_chart, delim = "-", names =c("task", "chart")
+    ) %>% mutate(
+  task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+  chart = factor(chart, levels =c("Vertical", "Horizontal", "Horizontal wide"))
+) 
+```
+
+```{r}
+#| label: fig-response-123
+#| fig-cap: "Responses for accuracy in the three designs.
+Responses to the same task are shown side-by-side for the three designs.
+The overlaid rectangles represent 95% confidence intervals.
+The letters in blue and orange encode significances between pairwise proportions: two bars have a significantly different proportion (at a 5% significance level) if they do not share a letter.
+There is no significant difference between the three designs for wrong responses.
+When tiles are unaligned, the horizontal wide barchart is showing the highest accuracy.
+For aligned tile, the horizontal wide design and the vertical design do not show a significant difference in accuracy."
+#| fig-height: 4.25
+#| fig-width: 7
+#| out-width: 80%
+
+round_123_long %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = chart)) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  facet_grid(.~task) + 
+  theme(legend.position="bottom") + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") + 
+  geom_rect(
+    aes(
+      xmin = as.numeric(chart) - 0.45,
+      ymin = correct - 1.96* `se.as.numeric(correct)`,
+      xmax = as.numeric(chart) + 0.45,
+      ymax = correct + 1.96* `se.as.numeric(correct)`,
+    ), 
+    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+    data = ci_confint
+  ) +
+  geom_segment(
+    aes(
+      x = as.numeric(chart) - 0.45,
+      y = correct,
+      xend = as.numeric(chart) + 0.45,
+      yend = correct
+    ), 
+    alpha = 1, colour = "black", linewidth = 0.25,
+    data = ci_confint
+  ) +
+  geom_rect(
+    aes(
+      xmin = as.numeric(chart) - 0.45,
+      ymin = correct3 - 1.96* `se.as.numeric(correct3)`,
+      xmax = as.numeric(chart) + 0.45,
+      ymax = correct3 + 1.96* `se.as.numeric(correct3)`,
+    ), 
+    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+    data = ci_confint3
+  ) +
+  geom_segment(
+    aes(
+      x = as.numeric(chart) - 0.45,
+      y = correct3,
+      xend = as.numeric(chart) + 0.45,
+      yend = correct3
+    ), 
+    alpha = 1, colour = "black", linewidth = 0.25,
+    data = ci_confint3
+  ) +
+  scale_y_continuous("Percent responses", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25), limits = c(-0.1, 1.1)) + 
+  geom_text(aes(label=letters), y = -0.05, data = ci_confint, colour = "steelblue") +
+  geom_text(aes(label=letters), y = 1.05, data = ci_confint3, colour = "darkorange") 
+```
+
+@fig-response-123 shows the results of a cell-means model with ordinal response $Y_k$, where $Y_k$ is the $k$th participant's response, $Y_k \in \{1, 2, 3\}$, where 'correct' is encoded as 1, 'they are the same' is encoded as 2, and 'wrong' is encoded as 3:
+
+$$
+\text{logit }P(Y_k \le \ell) = \mu_{ij\ell(k)},
+$$
+where  $\ell \in \{1, 2\}$; $i \in \{1, 2\}$ is the comparison type (1 = Aligned, 2 = Unaligned),  and $j \in \{1, 2, 3\}$ is the chart design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide.
+The estimated values and 95% confidence intervals are shown in @tbl-rounds-123
+
+```{r}
+#| label: tbl-rounds-123
+#| tbl-cap: Log-odds for the cell-means model, letters behind numbers indicate pairwise significances.
+Within the same **column** values are significantly different (at 5%) if they do not share the same letter.
+
+dframe <- rbind(
+  coefs_123 %>% mutate(level = c("correct | same or wrong")),
+  coefs_123_relaxed %>% mutate(level = c("correct or same | wrong"))
+) %>% mutate(
+  term = gsub("task_chart", "", term),
+) %>% separate(term, into=c("task", "chart"), sep="-")
+
+dframe %>% 
+  dplyr::select(level, task, chart, estimate, std.error, letters) %>%
+  mutate(
+    task = ifelse(task=="ab", "Unaligned", "Aligned")
+  ) %>%
+  group_by(task) %>%
+  mutate(
+    ci_low = exp(estimate-1.96*std.error),
+    ci_high = exp(estimate+1.96*std.error),
+    estimate = exp(estimate)
+  ) %>%
+  tidyr::pivot_wider(
+        id_cols = c(task, chart),
+        names_from = level,
+        values_from = c(estimate, ci_low, ci_high, letters)
+      ) %>%
+  gt() %>%
+  tab_header(title = "Log odds of accuracy by task and chart type") %>%
+  fmt_number(
+    everything(),
+    decimals=2
+  ) %>%
+  cols_merge(
+    columns = c("estimate_correct | same or wrong", "ci_low_correct | same or wrong",
+                "ci_high_correct | same or wrong", "letters_correct | same or wrong"),
+    pattern = "<<{1}>><< [{2}, {3}]>><< {4}>>"
+  ) %>%
+  cols_merge(
+    columns = c("estimate_correct or same | wrong", "ci_low_correct or same | wrong",
+                "ci_high_correct or same | wrong", "letters_correct or same | wrong"),
+    pattern = "<<{1}>><< [{2}, {3}]>><< {4}>>"
+  ) %>%
+  tab_options(
+    column_labels.font.weight = "bold",
+    row_group.font.weight = "bold"
+  ) %>%
+  cols_label(
+    .list = c(
+      "chart" = "",
+      "estimate_correct | same or wrong" = "correct | same or wrong",
+      "estimate_correct or same | wrong" = "correct or same | wrong"
+    )
+  )
+```
+
+Interestingly, while we should, theoretically, see an improvement in accuracy when shifting from the vertical to the horizontal wide design, the resulting effects on the accuracy of the responses are not completely straightforward:
+The shift from a vertical to the (tall) horizontal design is detrimental to an accurate perception for both aligned and unaligned comparisons.  
+The re-scaled design of the wide horizontal bars reclaims some of the loss for unaligned bars and outperforms the vertical design by a similar margin in aligned bars, but does not out-perform the vertical design when comparing unaligned tiles.
+
+#### Viewers react differently to the charts
+
+A contributing factor to this outcome might be the way that participants interact with the different designs.
+Generally, about half of all participants make use of the option to zoom into charts - while zooming does help with the overall accuracy (which is in agreement with the findings by @luModelingJustNoticeable2022 about the physical size of stimuli), the increase is not significant.
+However, different designs lead to different rates of zooming: when dealing with the vertical design, the rate of zooming is significantly higher than for the two horizontal designs.
+
+<!-- The rate of zooming into charts is about twice as high for participants filling out the survey on their smart phones than on other, usually bigger devices such as tablets or desktops. -->
+
+```{r zoom-123}
+logit_zoom <- svyglm(I(zoom=="yes")~response3+task+chart, family=quasibinomial(), design=survey_123)
+
+coefs_zoom <- broom::tidy(logit_zoom)
+```
+
+```{r}
+#| label: fig-zoom-123
+#| fig-cap: "Zooming - not significant for accuracy or task, but changes by the type of chart."
+#| fig-height: 4.25
+#| fig-width: 7
+#| out-width: 80%
+
+zt <- round_123_long %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide")),
+    zoom = factor(zoom, levels=c("no", "yes"))
+  ) %>% 
+  ggplot(aes(x = response3)) + 
+  geom_bar(aes(fill = zoom, weight = weight), position="fill") +
+  facet_grid(.~task) + 
+  theme(axis.text.x = element_text(angle=30, hjust = 1)) + 
+  scale_fill_manual("Zoom", values=c("grey", "#1274F3")) +
+  xlab("") + 
+  ggtitle("Zooming behavior by task")
+
+zd <- round_123_long %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide")),
+    zoom = factor(zoom, levels=c("no", "yes"))
+  ) %>% 
+  ggplot(aes(x = response3)) + 
+  geom_bar(aes(fill = zoom, weight = weight), position="fill") +
+  facet_grid(.~chart) + 
+  scale_fill_manual("Zoom", values=c("grey", "#1274F3")) +
+  xlab("") + 
+  ggtitle("Zooming behavior by chart") +
+  theme(axis.text.x = element_text(angle=30, hjust = 1))
+
+zt + zd + plot_layout(widths=c(5, 7.5), guides='collect')
+
+```
+
+Let $Y_{jk}$ describe the zooming behavior of panelist $k$ on task $j$.
+We model zooming behavior (no = 0, yes = 1) as a logistic regression by correctness of response ($\rho$),  task ($\tau$), and design ($\delta$) of the chart:
+$$
+\text{logit } P(Y_k \le 1) = \mu + \rho_{i(k)} + \tau_{j(k)} + \delta_{\ell(k)},
+$$
+
+I'll pretty up @tbl-zoom-123 if we are going to keep it.
+
+```{r}
+#| label: tbl-zoom-123
+#| tbl-cap: Coefficients for logistic regression of zooming by task
+
+coefs_zoom %>%
+  mutate(
+    term = ifelse(term=="(Intercept)", "Intercept", term),
+    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value))
+  ) %>%
+  gt() %>%
+  tab_header("Estimates for logistic regression on zooming behavior") %>%
+  cols_label(
+    std.error = "SE",
+    statistic = "t-statistic",
+    p.value = "p-value"
+  ) %>%
+  fmt_number(
+    columns = c(estimate, std.error),
+    decimals = 2
+  ) %>%
+  fmt_number(
+    columns = statistic,
+    decimals = 1
+  ) %>%
+  cols_align(
+    columns = p.value,
+    align = "right"
+  )
+
+```
+
+#### Demographics matter
+
+Let $Y_k$ be the response of participant $k$, on a scale from 1 = 'wrong', 2 = 'they are the same' to 3 = 'correct'.
+We use a generalized cumulative logistic regression, where $\mu_i$ are intercepts $1 \le i < 3$, $X_k$ are demographics of the $k$th participant (in form of the model matrix), and $\beta_{i}$ are the coefficients.
+$$
+\text{ logit } P(Y_k \le i \mid X_k) = \mu_i + X_{k}'\beta_{i}
+$$
+
+```{r}
+#| label: fig-demographics
+#| fig-cap: "Panelist's demographics matter, particularly, when the task difficulty increases.
+For aligned tiles, gender, age, and education are not significant factors.
+However, income levels do have - a small - effect.
+When income levels increase, the percentage of wrong answers (orange) decreases, while the percentage of correct answers (blue) increases slightly (significant at below 0.05).
+The more difficult task of comparing unaligned tiles, demographics are more significant.
+XXX With increasing levels in education, we see a small drop in correct answers, as well as a small drop in wrong answers, for a significant increase in 'they are the same'."
+#| fig-subcap: 
+#|   - Comparisons of aligned tiles
+#|   - Comparisons of unaligned tiles
+#| fig-height: 5
+#| fig-width: 7
+#| layout-nrow: 2
+#| out-width: 100%
+
+educ_cd <- round_123_long %>% 
+  filter(task=="cd") %>%
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = educ5)) + 
+  facet_grid(~task) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+    ggtitle("\nby education") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+
+educ_ab <- round_123_long %>% 
+  filter(task=="ab") %>%
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = educ5)) + 
+  facet_grid(.~task) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+    ggtitle("\nby education") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+
+gender_cd <- round_123_long %>% 
+  filter(task=="cd") %>%
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = gender)) + 
+  facet_grid(.~task) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+  ggtitle("Accuracy\nby gender") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+
+gender_ab <- round_123_long %>% 
+  filter(task=="ab") %>%
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = gender)) + 
+  facet_grid(.~task) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+  ggtitle("Accuracy\nby gender") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+
+income_cd <- round_123_long %>% 
+  filter(task=="cd") %>%
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = income4)) + 
+  facet_grid(.~task) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+    ggtitle("\nby income") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+
+income_ab <- round_123_long %>% 
+  filter(task=="ab") %>%
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = income4)) + 
+  facet_grid(.~task) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+    ggtitle("\nby income") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+
+race <- round_123_long %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = racethnicity)) + 
+  facet_grid(.~task) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+    ggtitle("Accuracy by race, ethnicity and task") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+
+age_cd <- round_123_long %>% 
+  filter(task=="cd") %>%
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = age4)) + 
+   facet_grid(.~task) + 
+ geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+    ggtitle("\nby age") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+
+age_ab <- round_123_long %>% 
+  filter(task=="ab") %>%
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+correct3+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+  ) %>% 
+  ggplot(aes(x = age4)) + 
+   facet_grid(.~task) + 
+ geom_bar(aes(fill = response3, weight = weight), position="fill") +
+  theme(legend.position="none",
+        axis.text.x = element_text(angle=30, hjust=1)) + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+    ggtitle("\nby age") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+gender_cd + age_cd  + educ_cd +income_cd + plot_layout(nrow=1, guides = "collect")
+
+gender_ab + age_ab  + educ_ab +income_ab + plot_layout(nrow=1, guides = "collect")
+```
+
+```{r demographics-model}
+logit_demo_task_cd_correct <- svyglm(correct~gender+age4+educ5+income4, design=survey_123, family=quasibinomial(), subset = task=="cd")
+
+logit_demo_task_cd_correct_same <- svyglm(correct3~gender+age4+educ5+income4, design=survey_123, family=quasibinomial(), subset = task=="cd")
+
+logit_demo_task_ab_correct <- svyglm(correct~gender+age4+educ5+income4, design=survey_123, family=quasibinomial(), subset = task=="ab")
+logit_demo_task_ab_correct_same <- svyglm(correct3~gender+age4+educ5+income4, design=survey_123, family=quasibinomial(), subset = task=="ab")
+
+coefs_cd_correct <- broom::tidy(logit_demo_task_cd_correct)
+coefs_cd_correct_same <- broom::tidy(logit_demo_task_cd_correct_same)
+
+coefs_ab_correct <- broom::tidy(logit_demo_task_ab_correct)
+coefs_ab_correct_same <- broom::tidy(logit_demo_task_ab_correct_same)
+```
+
+@tbl-demographics doesn't survive the formatting to pdf very well - \hh{need to shorten the row names or put into multiple lines}
+
+```{r demographics-table}
+#| label: tbl-demographics
+#| tbl-cap: Demographics matter for perception, particularly when the tasks get harder.
+ coefs_clean <- function(dframe) {
+  dframe %>% mutate(
+    variable = c("Intercept", "Gender", rep("Age", 3), rep("Education", 4), rep("Income", 3)),
+    term = ifelse(term=="(Intercept)", "", term),
+    term = gsub("gender", "", term),
+    term = gsub("educ5", "", term),
+    term = gsub("age4", "", term),
+    term = gsub("income4", "", term),
+    ci_low = exp(estimate-1.96*std.error),
+    ci_high = exp(estimate+1.96*std.error),
+    estimate = exp(estimate)
+  )
+}
+
+coefs_clean(coefs_cd_correct) %>% mutate(task = "Aligned", correct="correct") %>%
+ rbind(
+  coefs_clean(coefs_ab_correct) %>% mutate(task = "Unaligned", correct="correct")
+ ) %>%
+ rbind(
+  coefs_clean(coefs_cd_correct_same) %>% mutate(task = "Aligned", correct="not wrong")
+ ) %>%
+ rbind(
+  coefs_clean(coefs_ab_correct_same) %>% mutate(task = "Unaligned", correct="not wrong")
+ ) %>%
+ dplyr::select(-std.error, -statistic) %>%
+ mutate(
+  p.value = ifelse (p.value <  0.001, "***",
+      ifelse (p.value <  0.01, "**",
+             ifelse (p.value < 0.05, "*",
+                    ifelse (p.value < 0.1, ".", ""))))
+ ) %>%
+ group_by(variable) %>%
+ pivot_wider(
+    id_cols = c(term, variable),
+    names_from = c(task, correct),
+    values_from = c(estimate, ci_low, ci_high, p.value),
+  ) %>% 
+ gt() %>%
+  tab_header(title = "Log odds of accuracy by task and demographics of respondents") %>%
+ tab_options(
+   column_labels.font.weight = "bold",
+   row_group.font.weight = "bold"
+ ) %>%
+  fmt_number(
+    decimals=2
+  ) %>%
+  cols_width(
+    1~px(50),
+    contains("p.value") ~ px(20)
+  ) %>%
+  cols_merge(
+    columns = contains("ci") & contains("_Aligned_correct"),
+    pattern = "<< [{1}, {2}]>>"
+  ) %>%
+  cols_merge(
+    columns = contains("ci") & contains("_Aligned_not"),
+    pattern = "<< [{1}, {2}]>>"
+  ) %>%
+  cols_move_to_end(contains("_Aligned_")) %>%
+  tab_spanner(
+    label = "correct | same or wrong",
+    columns = contains("_Aligned_correct")
+  )  %>%
+  tab_spanner(
+    label = "correct or same | wrong",
+    columns = contains("_Aligned_not")
+  ) %>%
+  tab_spanner(
+    label = "Aligned tiles",
+    columns = contains("_Aligned_"),
+    spanners = c("correct | same or wrong", "correct or same | wrong")
+  ) %>%
+  cols_move_to_end(contains("_Unaligned_")) %>%
+  cols_merge(
+    columns = contains("ci") & contains("_Unaligned_correct"),
+    pattern = "<< [{1}, {2}]>>"
+  ) %>%
+  cols_merge(
+    columns = contains("ci") & contains("_Unaligned_not"),
+    pattern = "<< [{1}, {2}]>>"
+  ) %>%
+  tab_spanner(
+    label = "correct | same or wrong ",
+    columns = contains("_Unaligned_correct")
+  )  %>%
+  tab_spanner(
+    label = "correct or same | wrong ",
+    columns = contains("_Unaligned_not")
+  ) %>%
+  tab_spanner(
+    label = "Unaligned tiles",
+    columns = contains("_Unaligned")
+  ) %>%
+   cols_label(
+    .list = c(
+      "estimate_Aligned_correct" = "Est.",
+      "ci_low_Aligned_correct" = "          [95% CI]",
+      "p.value_Aligned_correct" = " ",
+      "estimate_Aligned_not wrong" = "Est.",
+      "ci_low_Aligned_not wrong" = "          [95% CI]",
+      "p.value_Aligned_not wrong" = " ",
+      "estimate_Unaligned_correct" = "Est.",
+      "ci_low_Unaligned_correct" = "          [95% CI]",
+      "p.value_Unaligned_correct" = " ",
+      "estimate_Unaligned_not wrong" = "Est.",
+      "ci_low_Unaligned_not wrong" = "          [95% CI]",
+      "p.value_Unaligned_not wrong" = " "
+    )
+  )
+
+```
 

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
starting to restructure tasks 1-3

--- 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
@@
 round1 <- readRDS(file.path(vs_repo_local, "round1.rds"))
+round2 <- readRDS(file.path(vs_repo_local, "round2.rds"))
 
@@
 
-round2 <- readRDS(file.path(vs_repo_local, "round2.rds"))
-
 round2_adj <- round2 %>% 
@@
 
-round3 <- readRDS(file.path(vs_repo_local, "round3.rds"))
+survey1 <- survey::svydesign(
+  data = round1_adj, 
+  ids = ~CaseId, weights = ~weight)
 
-# filter out all of the 'skipped on web':
-round3_adj <- round3 %>% mutate(
-  weight = WEIGHT_NOLA
-) %>% filter(
-  ab != "SKIPPED ON WEB",
-  cd != "SKIPPED ON WEB"
-)
-
-round6 <- readRDS(file.path(vs_repo_local, "round6.rds"))
-
-round6_adj <- round6 %>% mutate(
-  weight = WEIGHT_NOLA
-) %>% filter(RND_01 == 0) # same type of questions as round 7
-
-round7 <- readRDS(file.path(vs_repo_local, "round7.rds"))
-
-# filter out all of the 'skipped on web':
-round7_adj <- round7 %>% mutate(
-  weight = WEIGHT_NOLA
-)
-```
-
-```{r data-rnd3}
-round3 <- readRDS(file.path(vs_repo_local, "round3.rds"))
-
-# filter out all of the 'skipped on web':
-round3_adj <- round3 %>% mutate(
-  weight = WEIGHT_NOLA
-) %>% filter(
-  ab != "SKIPPED ON WEB",
-  cd != "SKIPPED ON WEB"
-)
+survey2 <- survey::svydesign(
+  data = round2_adj,
+  ids = ~CaseId, weights = ~weight)
 ```
@@
 In this work, we present the series of tests we completed and the resulting findings.
-The remainder of the paper is organized as follows: first, we describe the design of the visual stimulus used in our perception tests.
+The remainder of the paper is organized as follows: first, we describe the design of the visual stimuli used in our perception tests.
 We then describe the population of study respondents and obtained survey sample.
@@
 Viewers are presented with two marked pieces in a chart that do not share a common baseline, then two pieces that do share a common baseline.
-Second, we vary the orientation of the chart -- a vertically oriented stacked bar chart, a horizontally oriented version of the same chart, and then a wider version of the horizontally oriented chart.
-\kr{We need to finalize what we end up showing in the results and then revisit this section...} Third, we consider a facetted bar chart and facetted pie chart.
+Second, we vary the orientation of the chart -- we rotate the vertical stacked bar chart, and present a horizontally oriented version of the same chart, with identically sized marked pieces.
+Finally, we change the aspect ratio of the chart and present a wider version of the horizontally oriented chart which has longer marked pieces.
 
@@
 
-### Test 1: Alignment in stacked bar charts
+### Task 1: Alignment in stacked bar charts
 
@@
 
-### Test 2: Orientation of stacked bar charts
+### Task 2: Orientation of stacked bar charts
 
-In \autoref{fig-tasks-3} we see two structural variations on the stacked bar chart seen in Test 1: first, the Task 1 chart sized equally but rotated 90 degrees with bar lengths shown horizontally.
+Story here: not a huge difference, but we mainly see a difference in how people react.
+horizontal/vertical/horizontal.
+Even if there's not hugely significant differences by design, we can see differences in behavior.
+
+Combine Tasks 1 and 2 and just present overall, aligned/unaligned by the 3 different orientations? Doesn't matter whether we have a vertical or horizontal position,
+
+We can reproduce
+We can rank
+We can rank better because we have a bigger sample size
+WE can find changes in viewer behavior
+Demographics matter to accuracy and choices
+
+In \autoref{fig-tasks-3} we see two structural variations on the stacked bar chart seen in Task 1: first, the Task 1 chart sized equally but rotated 90 degrees with bar lengths shown horizontally.
 Second, the horizontal chart with a smaller aspect ratio so the bars appear thinner and longer.
@@
 
-### Test 3: Alignment and orientation in pie charts
+### Task 3: Alignment and orientation in pie charts
 
@@
 
-The data for this paper were collected in several rounds as part of the NORC Omnibus.
-
-| Name    | Date  | # Participants | effective sample size | Sum of weights $\sum_i w_i$ |
-|---------|-------|--------:|----------:|----------:|
-| Round 1 | April 2022 |  `r nrow(round1_adj)`    |    `r round(neff(round1_adj$weight),1)`       | `r round(sum(round1_adj$weight),1)`       |
-| Round 2 | May 2022 |  `r nrow(round2_adj)`   |    `r round(neff(round2_adj$weight),1)`       | `r round(sum(round2_adj$weight),1)`       |
-| Round 3 | Jun 2022 |  `r nrow(round3_adj)`  |    `r round(neff(round3_adj$weight),1)`       | `r round(sum(round3_adj$weight),1)`       |
-| Round 6 | Sep 2022 |  `r nrow(round6_adj)` [Split]  |    `r round(neff(round6_adj$weight),1)`       | `r round(sum(round6_adj$weight),1)`       |
-| Round 7 | Oct 2022 |  `r nrow(round7_adj)`   |    `r round(neff(round7_adj$weight),1)`       | `r round(sum(round7_adj$weight),1)`       |
-|         |       |         |           |           |
-
-: Survey rounds: dates, number of participants (nominal sample size), effective sample size, and sum of weights. {#tbl-rounds}
-
 # Results
@@
 
+```{r lambdas}
+n1 <- nrow(round1_adj)
+n2 <- nrow(round2_adj)
+neff1 <- svytotal(~one, survey1, deff=TRUE)
+neff2 <- svytotal(~one, survey2, deff=TRUE)
+#abcd1 <- svytotal(~CorrectABCD, survey1, deff=TRUE)
+#abcd2 <- svytotal(~CorrectABCD, survey2, deff=TRUE)
+
+d1 <- 1 + cv(neff1)^2
+d2 <- 1 + cv(neff2)^2
+
+denom <- n1/d1 + n2/d2
+
+lambda <- n1/(d1*denom) 
+```
+
 ```{r combine_rounds_1_and_2}
-
-lambda <- calculate_lambda(round1_adj$weight, round2_adj$weight)
 round_12 <- round1_adj %>% 

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd

--- 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
@@
 
-### Test 2: Orientation of stacked bar charts
-
-In \autoref{fig-tasks-3} we see two structural variations on the stacked bar chart seen in Test 1: first, the Task 1 chart sized equally but rotated 90 degrees with bar lengths shown horizontally.
+### Task 2: Orientation of stacked bar charts
+
+Story here: not a huge difference, but we mainly see a difference in how people react.
+horizontal/vertical/horizontal.
+Even if there's not hugely significant differences by design, we can see differences in behavior.
+
+Combine Tasks 1 and 2 and just present overall, aligned/unaligned by the 3 different orientations? Doesn't matter whether we have a vertical or horizontal position,
+
+We can reproduce
+We can rank
+We can rank better because we have a bigger sample size
+WE can find changes in viewer behavior
+Demographics matter to accuracy and choices
+
+In \autoref{fig-tasks-3} we see two structural variations on the stacked bar chart seen in Task 1: first, the Task 1 chart sized equally but rotated 90 degrees with bar lengths shown horizontally.
 Second, the horizontal chart with a smaller aspect ratio so the bars appear thinner and longer.

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
restructuring and starting section rewriting

@@
 Second, we vary the orientation of the chart -- we rotate the vertical stacked bar chart, and present a horizontally oriented version of the same chart, with identically sized marked pieces.
-Finally, we change the aspect ratio of the chart and present a wider version of the horizontally oriented chart which has longer marked pieces.
-
+Finally, we change the aspect ratio of the chart and present a wider version of the horizontally oriented chart which has longer, but thinner, marked pieces.
+
+```{=tex}
 \hh{Not sure where should put this, parking it here for right now:}
 \hh{expectations - start}
-What we call 'aligned' and 'unaligned', here, is similar to Cleveland and McGill's set of rankings, but with some modifications:
-both 'aligned' and 'unaligned' bars (in tests 1 and 2) or wedges (in test 3) share the same axis.
-Aligned tiles are additionally anchored in the same position in one dimension, i.e.
-the difference between their sizes can be reduced to a positional assessment.
-Unaligned tiles do not share this anchor, however, the context of the other tiles in the chart provide a frame, which *should* help with an assessment of the tiles' sizes beyond a comparison of (arc) lengths or areas.
+```
+What we call 'aligned' and 'unaligned', here, is similar to Cleveland and McGill's set of rankings, but with some modifications: both 'aligned' and 'unaligned' bars (in tests 1 and 2) or wedges (in test 3) share the same axis.
+Aligned tiles are additionally anchored in the same position in one dimension, i.e. the difference between their sizes can be reduced to a positional assessment.
+Unaligned tiles do not share this anchor, however, the context of the other tiles in the chart provide a frame, which *should* help with an assessment of the tiles' sizes beyond a comparison of (arc) lengths or areas.\
+
 We would expect that comparing unaligned tiles is a harder task (with correspondingly lower levels of accuracy) than a comparison of aligned tiles, with the framing given by the context of the other tiles in the same column (or the same pie) mitigating some of this difficulty.
@@
 First, participant instructions must be delivered in a very short and easily understandable format, because participants cannot ask clarifying questions about the task as they might be able to in a cognitive lab setting.
-\kr{I'm not sure how important this second point is...} Second, participants should be given some context for the tasks they are being asked to complete; in a given round, our set of tasks appear as a group in the midst of other survey questions and topics, and without providing respondents some transition we risk a jarring shift and low respondent engagement in the task.
+Second, we want to utilize content within the data visualization image which is not socially or politically charged for the average U.S.
+adult; this risks participants reacting to the subject matter within the chart rather than focusing on the presented task.
+For this reason, we utilize data on living arrangements of older U.S. adults -- a topic which most U.S. adults will have some familiarity with, but is not inherently polarizing.
 Finally, to prevent viewers from being exposed to slight variations of the same stimulus in a row (and risk unforeseen order effects or respondents using prior questions to inform their responses), we either split a survey sample in two and show each subsample a distinct version of the chart or test variations of a chart across distinct rounds of the survey.
@@
 
-### Task 1: Alignment in stacked bar charts
-
-\autoref{fig-tasks-12} shows the two stacked bar charts shown to participants in the first test.
+When viewing each chart, participants were asked to compare the relative sizes of marked elements within the chart:
+
+"There are many charts used in the news media to portray data visually. Looking at the chart below, which of the marked dark blue pieces is bigger, A or B? Just your best guess is fine.
+
+-   A is bigger
+
+-   B is bigger
+
+-   They are the same"
+
+When presented with the aligned version of each chart, pieces were marked with a C and D and the question text is updated accordingly.
+In all scenarios, the second option (B \[D\] is bigger) is the correct response.
+
+The time in seconds that each respondent spent on each task was recorded, as well as whether the participant zoomed in on each chart while answering the question.
+Respondents were also asked to rate their certainty in their response to each question on a five-point scale.
+
+### Task 1: Vertical stacked bar
+
+\autoref{fig-tasks-12} shows the two stacked bar charts shown to participants in the first task.
 The marked tiles in each plot are 155 pixels apart.
@@
 
-Both charts in \autoref{fig-tasks-12} show the same data with slight modifications to the order of the levels -- the first and second level in each of the bars are reversed between the left and the right chart.
-Participants were asked to compare the relative sizes of the tiles marked A and B (C and D, respectively) and select the correct response out of the possible choices:
-
-1.  A is bigger
-2.  B is bigger
-3.  They are the same
-
-Answer 2 is the correct answer for both charts.
+Both charts in the left column of \autoref{fig-tasks} show the same data with slight modifications to the order of the levels -- the first and second level in each of the bars are reversed between the left and the right chart.
 Both charts are shown at the same size, i.e. in both cases both the difference in size between the bars and the horizontal distance between the bars is the exact same amount.
@@
 
+Task 1 was presented to viewers across two distinct rounds, with four color scheme variations each presented to 50% of respondents in each round.
+There were no significant differences in respondent accuracy or behavior across color schemes, and we combine them here into one set of stimulus responses to compare structural differences.
+\kr{Should we put the color variations and a test for significance in accuracy/a visual in the appendix/supplementary materials? }
+
 ```{r}
-#| label: fig-tasks-12
-#| fig-cap: "Two stacked (vertical) barcharts. In each barchart, two tiles are marked. In both instances, the tile on the right is (very slightly) larger. "
+#| label: fig-tasks
+#| fig-cap: "Each of the stimuli presented to viewers. In each bar chart, two pieces are marked. The larger piece in each chart is piece A or C. "
 #| fig-subcap:
-#| - "**Unaligned bars**"
-#| - "**Aligned bars**"
-#| layout-nrow: 1
+#| - "**Vertical stacked bar, unaligned comparison**"
+#| - "**Horizontal stacked bar, unaligned comparison**"
+#| - "**Wide horizontal stacked bar, unaligned comparison**"
+#| - "**Vertical stacked bar, aligned comparison**"
+#| - "**Horizontal stacked bar, aligned comparison**"
+#| - "**Wide horizontal stacked bar, aligned comparison**"
+#| layout-nrow: 2
 #| out-width: 100%
 knitr::include_graphics("images/vertical-ab.PNG")
+knitr::include_graphics("images/round3/horizontal-ab.PNG")
+knitr::include_graphics("images/round3/hwide-ab.PNG")
 knitr::include_graphics("images/vertical-cd.PNG")
-```
-
-<!-- |                                   |                                   | -->
-
-<!-- |-----------------------------------|-----------------------------------| -->
-
-<!-- | **Unaligned positions**           | **Aligned positions**             | -->
-
-<!-- | ![Unaligned positions marked A and B](../SDSS-2023/images/vertical-ab.PNG) | ![Aligned positions marked C and D](../SDSS-2023/images/vertical-cd.PNG) | -->
-
-<!-- : The two stacked barcharts every participant got to see.
-In each barchart, the two marked tiles are to be compared for their size.
-in both instances, the tile on the right is (very slightly) larger.
-{#tbl-tasks-12 tbl-colwidths="\[50,50\]"} -->
+knitr::include_graphics("images/round3/horizontal-cd.PNG")
+knitr::include_graphics("images/round3/hwide-cd.PNG")
+```
+
+### Task 2: Horizontal stacked bar
+
+The visual mappings in the second task, shown in \autoref{fig-tasks-3a}, are identical to the first task, but the axes of the chart are rotated so that the stacked bar is represented in a horizontal format.
+This represents a structural change in how the data are presented to the viewer while preserving the pixel size of the elements viewers must compare.
+Once again Tasks 2 and 3 were asked within the same survey round and the full sample was randomly split among respondents, with 50% of viewers seeing the Task 2 stimuli and 50% of viewers seeing the Task 3 stimuli.
+
+### Task 3: Wide horizontal stacked bar
+
+The images utilized in the third task again represent the same data, and the overall image has dimensions which are identical to the images in the first two tasks.
+However, the aspect ratio of the plotting area is adjusted to increase the length in pixels of the stacked bar elements, and decrease the relative thickness \kr{wording?} of each bar.
 
@@
 
-Combine Tasks 1 and 2 and just present overall, aligned/unaligned by the 3 different orientations? Doesn't matter whether we have a vertical or horizontal position,
-
-We can reproduce
-We can rank
-We can rank better because we have a bigger sample size
-WE can find changes in viewer behavior
-Demographics matter to accuracy and choices
-
-In \autoref{fig-tasks-3} we see two structural variations on the stacked bar chart seen in Task 1: first, the Task 1 chart sized equally but rotated 90 degrees with bar lengths shown horizontally.
-Second, the horizontal chart with a smaller aspect ratio so the bars appear thinner and longer.
-Note that the wide version of the horizontal bars increases the pixel difference between the two marked pieces relative to the standard versions of the horizontal and vertical bars.
-
-This test utilized a split sample approach, where 50% of participants in the sample were shown the tall horizontal bar chart and 50% were shown the wide horizontal chart.
-Participants were asked to complete the same task as in Test 1; selecting the correct response among:
-
-1.  A is bigger
-2.  B is bigger
-3.  They are the same
-
-Again, answer 2 is the correct answer.
-
-```{r}
-#| label: fig-tasks-3
-#| fig-cap: "Changes to to the vertical design are made in two steps.
-First, the original chart is rotated.
-The areas of all tiles are kept the same.
-In a second step, the aspect ratio of tiles is changed while keeping the areas of tiles the same."
-#| fig-subcap:
-#| - "**Tall horizontal bars**, unaligned"
-#| - "**Wide horizontal bars**, unaligned"
-#| layout-nrow: 1
-#| out-width: 100%
-knitr::include_graphics("images/round3/horizontal-ab.PNG")
-knitr::include_graphics("images/round3/hwide-ab.PNG")
-```
-
-### Task 3: Alignment and orientation in pie charts
-
-Let's talk about the pies here
+Combine Tasks 1 and 2 and just present overall, aligned/unaligned by the 3 different orientations?
+Doesn't matter whether we have a vertical or horizontal position,
+
+We can reproduce We can rank We can rank better because we have a bigger sample size WE can find changes in viewer behavior Demographics matter to accuracy and choices
 
@@
 households.
-The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households **CITE DENNIS 2022** @dennis2019technical. \hh{I don't think I've got the right citation}
-Each test was conducted using the AmeriSpeak Omnibus survey, which runs biweekly and samples around 1,000 U.S. adults to answer questions on a variety of topics.
+The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households @dennis2019technical.
+\hh{I don't think I've got the right citation} Each test was conducted using the AmeriSpeak Omnibus survey, which runs biweekly and samples around 1,000 U.S.
+adults to answer questions on a variety of topics.
 
@@
 
-When combining responses that were gathered during distinct rounds of the Omnibus survey, we make weighting adjustments to ensure that weights in each sample are properly calibrated to the population total across both rounds in the resulting model.
-We *combine* (rather than cumulate) surveys $S_1$ and $S_2$, as described in @omuircheartaighCombiningSamplesVs2002, by
-multiplying weights in $S_1$ and $S_2$ by $\lambda$ and $1 - \lambda$ respectively.
+When combining responses that were gathered during distinct rounds of the Omnibus survey, we make a weighting adjustment to ensure that weights in each sample are properly calibrated to the population total across all rounds in the resulting model.
+
+\kr{Do we need to update this description to just say how we did it for combining across all the rounds represented?}
+
+We *combine* (rather than cumulate) surveys $S_1$ and $S_2$, as described in @omuircheartaighCombiningSamplesVs2002, by multiplying weights in $S_1$ and $S_2$ by $\lambda$ and $1 - \lambda$ respectively.
 $$
 \lambda = \frac{n_1/d_1}{n_1/d_1 + n_2/d_2},
-$$
-where $n_1$ and $n_2$ are the nominal sample sizes and $d_1$ and $d_2$ are the design effects for the estimators. 
-Here, $d_1$ and $d_2$ are estimated as 
-$$
+$$ where $n_1$ and $n_2$ are the nominal sample sizes and $d_1$ and $d_2$ are the design effects for the estimators.
+Here, $d_1$ and $d_2$ are estimated as $$
 d_1 = 1 + CV(w_i \in S_1)^2 \ \ \ \text{ and } \ \ \ d_2 = 1 + CV(w_i \in S_2)^2
-$$
-where $CV$ is the coefficient of variation of the weights within each sample, and is estimated as in @kishSurveySampling1965 :
+$$ where $CV$ is the coefficient of variation of the weights within each sample, and is estimated as in @kishSurveySampling1965 :
 
@@
 
-| Name    | Date  | # Participants | effective sample size | Sum of weights $\sum_i w_i$ |
-|---------|-------|--------:|----------:|----------:|
+<!-- | Name    | Date       |                \# Participants |                effective sample size |         Sum of weights $\sum_i w_i$ | -->
+
+<!-- |---------------|---------------|--------------:|--------------:|--------------:| -->
+
+<!-- | Round 1 | April 2022 |           `r nrow(round1_adj)` | `r round(neff(round1_adj$weight),1)` | `r round(sum(round1_adj$weight),1)` | -->
+
+<!-- | Round 2 | May 2022   |           `r nrow(round2_adj)` | `r round(neff(round2_adj$weight),1)` | `r round(sum(round2_adj$weight),1)` | -->
+
+<!-- | Round 3 | Jun 2022   |           `r nrow(round3_adj)` | `r round(neff(round3_adj$weight),1)` | `r round(sum(round3_adj$weight),1)` | -->
+
+<!-- | Round 6 | Sep 2022   | `r nrow(round6_adj)` \[Split\] | `r round(neff(round6_adj$weight),1)` | `r round(sum(round6_adj$weight),1)` | -->
+
+<!-- | Round 7 | Oct 2022   |           `r nrow(round7_adj)` | `r round(neff(round7_adj$weight),1)` | `r round(sum(round7_adj$weight),1)` | -->
+
+<!-- |         |            |                                |                                      |                                     | -->
+
+<!-- : Survey rounds: dates, number of participants (nominal sample size), effective sample size, and sum of weights. {#tbl-rounds} -->
+
+# Results
+
+### Respondents
+
+| Name    | Date       |      \# Participants |                effective sample size |         Sum of weights $\sum_i w_i$ |
+|---------------|---------------|--------------:|--------------:|--------------:|
 | Round 1 | April 2022 | `r nrow(round1_adj)` | `r round(neff(round1_adj$weight),1)` | `r round(sum(round1_adj$weight),1)` |
@@
 | Round 3 | Jun 2022   | `r nrow(round3_adj)` | `r round(neff(round3_adj$weight),1)` | `r round(sum(round3_adj$weight),1)` |
-| Round 6 | Sep 2022 |  `r nrow(round6_adj)` [Split]  |    `r round(neff(round6_adj$weight),1)`       | `r round(sum(round6_adj$weight),1)`       |
-| Round 7 | Oct 2022 |  `r nrow(round7_adj)`   |    `r round(neff(round7_adj$weight),1)`       | `r round(sum(round7_adj$weight),1)`       |
 |         |            |                      |                                      |                                     |
@@
 
-# Results
-
-### Respondents
-
 Description of respondents
@@
 
-### Test 1
-
--   Alignment in stacked bars
-
-    -   Binary accuracy in responses
-
-        -   Figure: binary correctness and binary ggpcp diagram
-        -   Table:(?) T-test results
-        -   Table: results of cell means model on certainty
-
-    -   Ordinal accuracy in responses
-
-        -   Figure: ordinal correctness and ordinal ggpcp diagram
-
-        -   Table: Fitted model for ordinal response and alignment
-
-    -   Ordinal accuracy in responses by demographics
-
-        -   Figure: ordinal correctness by demographics, split by aligned and unaligned
-
-        -   Table: Fitted model for ordinal response and alignment by demos
-
-            -   Maybe this table should go in the appendix?
+### Accuracy of responses
 
@@
 ```
-
-The data used for assessing the accuracy of comparisons in \autoref{tbl-tasks-12} is collected in two rounds of the NORC Omnibus survey.
-Rounds 1 and 2 are combined by adjusting the weights with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`.
-Figure \ref{fig-alignment}(a) shows that more than twice the number of responses are correct, when the tiles are aligned along the same axis.
-Because each participant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts. 
-The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: \< 2.2e-16).
-
-```{r accuracy,   warning=FALSE}
-#| fig-cap: "On the left (a), a stacked barchart shows the number of respondents with correct (green) and incorrect (grey) responses to the two comparison questions.
-When tiles are aligned along the same axis, more than twice the number of responses are correct.
-The shaded area along the top of the green tiles corresponds to  95\\% confidence intervals around (marginal) correct responses.
-On the right (b), a parallel coordinate plot shows all combinations of responses.
-There's a huge asymmetry in the number of responses where participants answered only one of the questions correctly.
-A lot more responses are correct when comparing aligned tiles than unaligned tiles."
-#| label: fig-alignment
-#| out-width: 80%
-#| fig-height: 5 
-#| fig-width: 10
-
-gg1 <- round_12_long %>%
-  mutate(
-    Response = ifelse(Response, "Correct", "Not correct"),
-    Response = factor(Response, levels=c("Not correct", "Correct")),
-    Question = ifelse(Question=="CorrectAB",
-                      "Unaligned: A or B?", "Aligned: C or D?")
-  ) %>%
-#  filter(Response == "Correct") %>%
-  ggplot(aes(x = Question, fill = Response, weight = weight)) +
-  geom_bar() +
-  ylab("Number of (effective) Respondents") +
-  xlab("") +
-  scale_fill_manual(values=c("grey", "forestgreen")) +
-  scale_y_continuous(
-      sec.axis = sec_axis(~./sum(round_12$weight)*100,
-                          name="Percent responses")
-  ) +
-  theme(legend.position = "bottom") +
-  ggtitle("(a) Accuracy of survey responses")  +
-  geom_rect(
-    aes(
-      x = rev(c("Aligned: C or D?", "Unaligned: A or B?")),
-      xmin = 2:1 - 0.45,
-      ymin = ResponseTRUE - 1.96* se.ResponseTRUE,
-      xmax = 2:1 + 0.45,
-      ymax = ResponseTRUE + 1.96* se.ResponseTRUE,
-      weight = 1),
-    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
-    data = marginal_results
-  )
-
-gg2 <- round_12 %>%
-  mutate(
-    `Unaligned: A or B?` = ifelse(CorrectAB, "Correct", "Not correct"), 
-    `Aligned: C or D?` = ifelse(CorrectCD, "Correct", "Not correct")
-    ) %>%
-  pcp_select(`Aligned: C or D?`, `Unaligned: A or B?`) %>%
-  pcp_scale() %>%
-  pcp_arrange() %>%
-  ggplot(aes_pcp()) + 
-    geom_pcp_axes() + 
-    geom_pcp(aes(colour = CorrectABCD), alpha = 0.2)+
-    geom_pcp_boxes(fill=NA, colour = "black", linewidth=0.25) + 
-    geom_pcp_labels(colour = "black") +
-  scale_y_continuous("Number responses", breaks=seq(0,1,by=0.25), 
-                     labels = round(sum(round_12$weight)*seq(0,1,by=0.25),0),
-                     sec.axis = sec_axis(~.*100,
-                          name="Percent responses")) +
-  xlab("") + 
-  scale_colour_manual(values = c("grey", "#84AC84", "#70A570", "forestgreen")) +
-  ggtitle("(b) Combinations of responses") +
-  theme(legend.position = "none")
-
-gg1 + gg2
-```
-
-### Test 2
-
--   Structure of stacked bars -- vertical, horizontal, horizontal wide
-    -   Accuracy and responses
 
@@
 \text{logit }P(Y_k \le \ell) = \mu_{ij\ell(k)},
-$$
-where  $\ell \in \{1, 2\}$; $i \in \{1, 2\}$ is the comparison type (1 = Aligned, 2 = Unaligned),  and $j \in \{1, 2, 3\}$ is the chart design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide.
+$$ where $\ell \in \{1, 2\}$; $i \in \{1, 2\}$ is the comparison type (1 = Aligned, 2 = Unaligned), and $j \in \{1, 2, 3\}$ is the chart design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide.
 The estimated values and 95% confidence intervals are shown in @tbl-rounds-123
@@
 
-Interestingly, while we should, theoretically, see an improvement in accuracy when shifting from the vertical to the horizontal wide design, the resulting effects on the accuracy of the responses are not completely straightforward:
-The shift from a vertical to the (tall) horizontal design is detrimental to an accurate perception for both aligned and unaligned comparisons.  
+Interestingly, while we should, theoretically, see an improvement in accuracy when shifting from the vertical to the horizontal wide design, the resulting effects on the accuracy of the responses are not completely straightforward: The shift from a vertical to the (tall) horizontal design is detrimental to an accurate perception for both aligned and unaligned comparisons.\
+
 The re-scaled design of the wide horizontal bars reclaims some of the loss for unaligned bars and outperforms the vertical design by a similar margin in aligned bars, but does not out-perform the vertical design when comparing unaligned tiles.
 
-#### Viewers react differently to the charts
+<!-- The data used for assessing the accuracy of comparisons in \autoref{tbl-tasks-12} is collected in two rounds of the NORC Omnibus survey. -->
+<!-- Rounds 1 and 2 are combined by adjusting the weights with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`. -->
+<!-- Figure \ref{fig-alignment}(a) shows that more than twice the number of responses are correct, when the tiles are aligned along the same axis. -->
+<!-- Because each participant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts. -->
+<!-- The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: \< 2.2e-16). -->
+
+<!-- ```{r accuracy,   warning=FALSE} -->
+<!-- #| fig-cap: "On the left (a), a stacked barchart shows the number of respondents with correct (green) and incorrect (grey) responses to the two comparison questions.
+When tiles are aligned along the same axis, more than twice the number of responses are correct.
+The shaded area along the top of the green tiles corresponds to  95\\% confidence intervals around (marginal) correct responses.
+On the right (b), a parallel coordinate plot shows all combinations of responses.
+There's a huge asymmetry in the number of responses where participants answered only one of the questions correctly.
+A lot more responses are correct when comparing aligned tiles than unaligned tiles." -->
+<!-- #| label: fig-alignment -->
+<!-- #| out-width: 80% -->
+<!-- #| fig-height: 5  -->
+<!-- #| fig-width: 10 -->
+
+<!-- gg1 <- round_12_long %>% -->
+<!--   mutate( -->
+<!--     Response = ifelse(Response, "Correct", "Not correct"), -->
+<!--     Response = factor(Response, levels=c("Not correct", "Correct")), -->
+<!--     Question = ifelse(Question=="CorrectAB", -->
+<!--                       "Unaligned: A or B?", "Aligned: C or D?") -->
+<!--   ) %>% -->
+<!-- #  filter(Response == "Correct") %>% -->
+<!--   ggplot(aes(x = Question, fill = Response, weight = weight)) + -->
+<!--   geom_bar() + -->
+<!--   ylab("Number of (effective) Respondents") + -->
+<!--   xlab("") + -->
+<!--   scale_fill_manual(values=c("grey", "forestgreen")) + -->
+<!--   scale_y_continuous( -->
+<!--       sec.axis = sec_axis(~./sum(round_12$weight)*100, -->
+<!--                           name="Percent responses") -->
+<!--   ) + -->
+<!--   theme(legend.position = "bottom") + -->
+<!--   ggtitle("(a) Accuracy of survey responses")  + -->
+<!--   geom_rect( -->
+<!--     aes( -->
+<!--       x = rev(c("Aligned: C or D?", "Unaligned: A or B?")), -->
+<!--       xmin = 2:1 - 0.45, -->
+<!--       ymin = ResponseTRUE - 1.96* se.ResponseTRUE, -->
+<!--       xmax = 2:1 + 0.45, -->
+<!--       ymax = ResponseTRUE + 1.96* se.ResponseTRUE, -->
+<!--       weight = 1), -->
+<!--     fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25, -->
+<!--     data = marginal_results -->
+<!--   ) -->
+
+<!-- gg2 <- round_12 %>% -->
+<!--   mutate( -->
+<!--     `Unaligned: A or B?` = ifelse(CorrectAB, "Correct", "Not correct"),  -->
+<!--     `Aligned: C or D?` = ifelse(CorrectCD, "Correct", "Not correct") -->
+<!--     ) %>% -->
+<!--   pcp_select(`Aligned: C or D?`, `Unaligned: A or B?`) %>% -->
+<!--   pcp_scale() %>% -->
+<!--   pcp_arrange() %>% -->
+<!--   ggplot(aes_pcp()) +  -->
+<!--     geom_pcp_axes() +  -->
+<!--     geom_pcp(aes(colour = CorrectABCD), alpha = 0.2)+ -->
+<!--     geom_pcp_boxes(fill=NA, colour = "black", linewidth=0.25) +  -->
+<!--     geom_pcp_labels(colour = "black") + -->
+<!--   scale_y_continuous("Number responses", breaks=seq(0,1,by=0.25),  -->
+<!--                      labels = round(sum(round_12$weight)*seq(0,1,by=0.25),0), -->
+<!--                      sec.axis = sec_axis(~.*100, -->
+<!--                           name="Percent responses")) + -->
+<!--   xlab("") +  -->
+<!--   scale_colour_manual(values = c("grey", "#84AC84", "#70A570", "forestgreen")) + -->
+<!--   ggtitle("(b) Combinations of responses") + -->
+<!--   theme(legend.position = "none") -->
+
+<!-- gg1 + gg2 -->
+<!-- ``` -->
+
+### Respondent behavior
 
@@
 Let $Y_{jk}$ describe the zooming behavior of panelist $k$ on task $j$.
-We model zooming behavior (no = 0, yes = 1) as a logistic regression by correctness of response ($\rho$),  task ($\tau$), and design ($\delta$) of the chart:
-$$
+We model zooming behavior (no = 0, yes = 1) as a logistic regression by correctness of response ($\rho$), task ($\tau$), and design ($\delta$) of the chart: $$
 \text{logit } P(Y_k \le 1) = \mu + \rho_{i(k)} + \tau_{j(k)} + \delta_{\ell(k)},
@@
 
-#### Demographics matter
+### Differences across demographic groups
 
@@
 
-### Test 3
-
--   Facetted bars and pies
-    -   Accuracy and responses
-
-    -   Timing of responses
-
 Results to discuss:
@@
 
--   Differences in accuracy between framed, floating, and general pie
+-   Differences across structures
 
 -   Differences in timing and zooming behavior across all designs
+
+-   Demographic differences and their interaction with alignment
 

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
minor changes to institution and initial research questions

@@
 affiliation:
-  - institution: National Opinion Research Center (NORC)
+  - institution: NORC at the University of Chicago
     country: United States
@@
   pdf: 
-    documentclass: jds
+    documentclass: jdsart
     include-in-header: authors.tex
@@
 2.  How is viewer interaction with the task impacted by structural design choices in a data visualization?
+
+3.  Are there differences in perception and interaction with the tasks across demographic groups?
 

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
some more moving around of figures

@@
 First, we have access to a large sample of survey participants and thus have greater power in making inference about graphical perceptional abilities.
-Second, the sample is representative of the general adult public in the U.S., which is an important target audience for scientific communication, and this allows us to test whether prior results from convenience samples hold with a nationally representative sample.
-\kr{Do we want a stronger statement about this here?}
-
-We present viewers with structural variations on bar charts and pie charts and ask them to answer questions comparing the size of elements within those charts.
+Second, the sample is representative of the general adult public in the U.S., and this allows us to test whether prior results from convenience samples hold with a nationally representative sample and whether there are differences in those results across demographic subgroups.
+
+We present viewers with structural variations of bar charts and ask them to answer questions comparing the size of elements within those charts.
 In this work, we present the series of tests we completed and the resulting findings.
@@
 
-\kr{The figure below is still a bit of an orphan, but I think we can revisit the intro ad study setup after we get more of the results in the paper}
-
-```{r}
-#| label: fig-aligned-unaligned
-#| fig-cap: "The only difference between the two pairs of rectangles A, B and C, D is their alignment, i.e.
+<!-- \kr{The figure below is still a bit of an orphan, but I think we can revisit the intro ad study setup after we get more of the results in the paper} -->
+
+<!-- ```{r} -->
+<!-- #| label: fig-aligned-unaligned -->
+<!-- #| fig-cap: "The only difference between the two pairs of rectangles A, B and C, D is their alignment, i.e.
 A and C are of identical size, as are B and D.
 When participants are asked to compare the size of these tiles in barcharts, the predominant response for the unaligned pair on the left is 'they are of the same size'.
-In contrast, more than half of the viewers respond with 'D is bigger' to the aligned pair of bars on the right."
-#| fig-subcap:
-#| - "**Unaligned bars**"
-#| - "**Aligned bars**"
-#| layout-nrow: 1
-#| fig-align: "center"
-#| out-width: 50%
-knitr::include_graphics("images/round5-modified.PNG")
-knitr::include_graphics("images/round3-modified.PNG")
-```
+In contrast, more than half of the viewers respond with 'D is bigger' to the aligned pair of bars on the right." -->
+<!-- #| fig-subcap: -->
+<!-- #| - "**Unaligned bars**" -->
+<!-- #| - "**Aligned bars**" -->
+<!-- #| layout-nrow: 1 -->
+<!-- #| fig-align: "center" -->
+<!-- #| out-width: 50% -->
+<!-- knitr::include_graphics("images/round5-modified.PNG") -->
+<!-- knitr::include_graphics("images/round3-modified.PNG") -->
+<!-- ``` -->
 
@@
 
-**How do we present the task?** The format of a survey guides the format and design of the questions asked and how they are presented to respondents.
+**How do we present the task?** Our use of a survey format guides the format and design of the questions asked and how they are presented to respondents.
 First, participant instructions must be delivered in a very short and easily understandable format, because participants cannot ask clarifying questions about the task as they might be able to in a cognitive lab setting.
@@
 
-\kr{Somewhere here we should also talk about how we measure viewer behavior/interaction with the chart: certainty, time spent, zooming, devices, etc. }
-
 When viewing each chart, participants were asked to compare the relative sizes of marked elements within the chart:
@@
 
-### Task 1: Vertical stacked bar
-
-\autoref{fig-tasks-12} shows the two stacked bar charts shown to participants in the first task.
-The marked tiles in each plot are 155 pixels apart.
-Based on @luModelingJustNoticeable2022's model, a difference of 155 pixels leads to a just noticeable difference of 3.5 pixels.
-The heights of the bars are 205 (left) and 213 pixels (right), respectively, corresponding to about twice the JND.
-This difference should lead to a relatively high accuracy rate for participants and simultaneously limit the amount of frustration resulting from a task that is perceived as 'too hard'.
-
-<!-- Lu et al calculation: log_{10}(JND) = -.4653 + .0065 distance -->
-
-<!-- describe stimulus -->
-
-Both charts in the left column of \autoref{fig-tasks} show the same data with slight modifications to the order of the levels -- the first and second level in each of the bars are reversed between the left and the right chart.
-Both charts are shown at the same size, i.e. in both cases both the difference in size between the bars and the horizontal distance between the bars is the exact same amount.
-This leaves the vertical positioning of the bars as the only difference between the charts.
-Any differences in observed responses can therefore be attributed to this difference in presentation.
-
-Task 1 was presented to viewers across two distinct rounds, with four color scheme variations each presented to 50% of respondents in each round.
-There were no significant differences in respondent accuracy or behavior across color schemes, and we combine them here into one set of stimulus responses to compare structural differences.
-\kr{Should we put the color variations and a test for significance in accuracy/a visual in the appendix/supplementary materials? }
-
 ```{r}
 #| label: fig-tasks
-#| fig-cap: "Each of the stimuli presented to viewers. In each bar chart, two pieces are marked. The larger piece in each chart is piece A or C. "
+#| fig-cap: "Each of the visual stimuli presented to viewers. In each bar chart, two pieces are marked. The larger piece in each chart is piece A or C. "
 #| fig-subcap:
-#| - "**Vertical stacked bar, unaligned comparison**"
-#| - "**Horizontal stacked bar, unaligned comparison**"
-#| - "**Wide horizontal stacked bar, unaligned comparison**"
-#| - "**Vertical stacked bar, aligned comparison**"
-#| - "**Horizontal stacked bar, aligned comparison**"
-#| - "**Wide horizontal stacked bar, aligned comparison**"
+#| - "**Vertical stacked bar, unaligned**"
+#| - "**Horizontal stacked bar, unaligned**"
+#| - "**Wide horizontal stacked bar, unaligned**"
+#| - "**Vertical stacked bar, aligned**"
+#| - "**Horizontal stacked bar, aligned**"
+#| - "**Wide horizontal stacked bar, aligned**"
 #| layout-nrow: 2
@@
 
+### Task 1: Vertical stacked bar
+
+\autoref{fig-tasks}(a,d) shows the two stacked bar charts shown to participants in the first task.
+The marked tiles in each plot are 155 pixels apart.
+Based on @luModelingJustNoticeable2022's model, a difference of 155 pixels leads to a just noticeable difference of 3.5 pixels.
+The heights of the bars are 205 (left) and 213 pixels (right), respectively, corresponding to about twice the JND.
+This difference should lead to a relatively high accuracy rate for participants and simultaneously limit the amount of frustration resulting from a task that is perceived as 'too hard'.
+
+<!-- Lu et al calculation: log_{10}(JND) = -.4653 + .0065 distance -->
+
+<!-- describe stimulus -->
+
+Both charts show the same data with slight modifications to the order of the levels -- the first and second level in each of the bars are reversed between the left and the right chart.
+Both charts are shown at the same size, i.e. in both cases both the difference in size between the bars and the horizontal distance between the bars is the exact same amount.
+This leaves the vertical positioning of the bars as the only difference between the charts.
+Any differences in observed responses can therefore be attributed to this difference in presentation.
+
+Task 1 was presented to viewers in one of two rounds (Rounds 1-2), with four color scheme variations each presented to 50% of respondents in each round.
+There were no significant differences in respondent accuracy or behavior across color schemes, and we combine them here into one set of stimulus responses to compare structural differences.
+\kr{Should we put the color variations and a test for significance in accuracy/a visual in the appendix/supplementary materials? }
+
 ### Task 2: Horizontal stacked bar
@@
 This represents a structural change in how the data are presented to the viewer while preserving the pixel size of the elements viewers must compare.
-Once again Tasks 2 and 3 were asked within the same survey round and the full sample was randomly split among respondents, with 50% of viewers seeing the Task 2 stimuli and 50% of viewers seeing the Task 3 stimuli.
+Tasks 2 and 3 were asked within the same survey round (Round 3) and the full sample was randomly split among respondents, with 50% of participants seeing the Task 2 stimuli and 50% of participants seeing the Task 3 stimuli.
 
@@
 However, the aspect ratio of the plotting area is adjusted to increase the length in pixels of the stacked bar elements, and decrease the relative thickness \kr{wording?} of each bar.
-
-### Task 2: Orientation of stacked bar charts
-
-Story here: not a huge difference, but we mainly see a difference in how people react.
-horizontal/vertical/horizontal.
-Even if there's not hugely significant differences by design, we can see differences in behavior.
-
-Combine Tasks 1 and 2 and just present overall, aligned/unaligned by the 3 different orientations?
-Doesn't matter whether we have a vertical or horizontal position,
-
-We can reproduce We can rank We can rank better because we have a bigger sample size WE can find changes in viewer behavior Demographics matter to accuracy and choices
 
@@
 
+There are differences by alignment
+There are not huge differences across structures
+However, we see a difference in how people react across different structures (zooming behavior, certainty? time?)
+
+- We can reproduce prior findings with nat rep sample 
+- We can rank 
+- We can rank better because we have a bigger sample size 
+- We can find changes in viewer behavior 
+- Demographics matter to accuracy and choices
+
 ### Respondents
+
+A total of `r sum(c(nrow(round1_adj), nrow(round2_adj), nrow(round3_adj)))` respondents participated across the three rounds.
+The number of responses and corresponding effective sample sizes in each round are shown in \autoref{tbl-rounds}.
+
+\kr{We need to be careful about not making it seem like the rounds are the same as the tasks, since we have 3 rounds and 3 tasks.
+We describe it, but I feel like this table kind of muddies the waters.
+Could we instead do a table of respondents by task?}
 
@@
 ```
+
+<!-- The data used for assessing the accuracy of comparisons in \autoref{tbl-tasks-12} is collected in two rounds of the NORC Omnibus survey. -->
+<!-- Rounds 1 and 2 are combined by adjusting the weights with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`. -->
+<!-- Figure \ref{fig-alignment}(a) shows that more than twice the number of responses are correct, when the tiles are aligned along the same axis. -->
+<!-- Because each participant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts. -->
+<!-- The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: \< 2.2e-16). -->
 
@@
   ids = ~CaseId, weights = ~weight)
+```
+
+\kr{We need to update this chart to include the data from all 3 rounds instead of just the first 2}
+
+```{r accuracy,   warning=FALSE}
+#| fig-cap: "On the left (a), a stacked barchart shows the number of respondents with correct (green) and incorrect (grey) responses to the two comparison questions.
+When tiles are aligned along the same axis, more than twice the number of responses are correct.
+The shaded area along the top of the green tiles corresponds to  95\\% confidence intervals around (marginal) correct responses.
+On the right (b), a parallel coordinate plot shows all combinations of responses.
+There's a huge asymmetry in the number of responses where participants answered only one of the questions correctly.
+A lot more responses are correct when comparing aligned tiles than unaligned tiles."
+#| label: fig-alignment
+#| out-width: 80%
+#| fig-height: 5
+#| fig-width: 10
+
+gg1 <- round_12_long %>%
+  mutate(
+    Response = ifelse(Response, "Correct", "Not correct"),
+    Response = factor(Response, levels=c("Not correct", "Correct")),
+    Question = ifelse(Question=="CorrectAB",
+                      "Unaligned: A or B?", "Aligned: C or D?")
+  ) %>%
+#  filter(Response == "Correct") %>%
+  ggplot(aes(x = Question, fill = Response, weight = weight)) +
+  geom_bar() +
+  ylab("Number of (effective) Respondents") +
+  xlab("") +
+  scale_fill_manual(values=c("grey", "forestgreen")) +
+  scale_y_continuous(
+      sec.axis = sec_axis(~./sum(round_12$weight)*100,
+                          name="Percent responses")
+  ) +
+  theme(legend.position = "bottom") +
+  ggtitle("(a) Accuracy of survey responses")  +
+  geom_rect(
+    aes(
+      x = rev(c("Aligned: C or D?", "Unaligned: A or B?")),
+      xmin = 2:1 - 0.45,
+      ymin = ResponseTRUE - 1.96* se.ResponseTRUE,
+      xmax = 2:1 + 0.45,
+      ymax = ResponseTRUE + 1.96* se.ResponseTRUE,
+      weight = 1),
+    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+    data = marginal_results
+  )
+
+gg2 <- round_12 %>%
+  mutate(
+    `Unaligned: A or B?` = ifelse(CorrectAB, "Correct", "Not correct"),
+    `Aligned: C or D?` = ifelse(CorrectCD, "Correct", "Not correct")
+    ) %>%
+  pcp_select(`Aligned: C or D?`, `Unaligned: A or B?`) %>%
+  pcp_scale() %>%
+  pcp_arrange() %>%
+  ggplot(aes_pcp()) +
+    geom_pcp_axes() +
+    geom_pcp(aes(colour = CorrectABCD), alpha = 0.2)+
+    geom_pcp_boxes(fill=NA, colour = "black", linewidth=0.25) +
+    geom_pcp_labels(colour = "black") +
+  scale_y_continuous("Number responses", breaks=seq(0,1,by=0.25),
+                     labels = round(sum(round_12$weight)*seq(0,1,by=0.25),0),
+                     sec.axis = sec_axis(~.*100,
+                          name="Percent responses")) +
+  xlab("") +
+  scale_colour_manual(values = c("grey", "#84AC84", "#70A570", "forestgreen")) +
+  ggtitle("(b) Combinations of responses") +
+  theme(legend.position = "none")
+
+gg1 + gg2
 ```
@@
 
-<!-- The data used for assessing the accuracy of comparisons in \autoref{tbl-tasks-12} is collected in two rounds of the NORC Omnibus survey. -->
-<!-- Rounds 1 and 2 are combined by adjusting the weights with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`. -->
-<!-- Figure \ref{fig-alignment}(a) shows that more than twice the number of responses are correct, when the tiles are aligned along the same axis. -->
-<!-- Because each participant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts. -->
-<!-- The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: \< 2.2e-16). -->
-
-<!-- ```{r accuracy,   warning=FALSE} -->
-<!-- #| fig-cap: "On the left (a), a stacked barchart shows the number of respondents with correct (green) and incorrect (grey) responses to the two comparison questions.
-When tiles are aligned along the same axis, more than twice the number of responses are correct.
-The shaded area along the top of the green tiles corresponds to  95\\% confidence intervals around (marginal) correct responses.
-On the right (b), a parallel coordinate plot shows all combinations of responses.
-There's a huge asymmetry in the number of responses where participants answered only one of the questions correctly.
-A lot more responses are correct when comparing aligned tiles than unaligned tiles." -->
-<!-- #| label: fig-alignment -->
-<!-- #| out-width: 80% -->
-<!-- #| fig-height: 5  -->
-<!-- #| fig-width: 10 -->
-
-<!-- gg1 <- round_12_long %>% -->
-<!--   mutate( -->
-<!--     Response = ifelse(Response, "Correct", "Not correct"), -->
-<!--     Response = factor(Response, levels=c("Not correct", "Correct")), -->
-<!--     Question = ifelse(Question=="CorrectAB", -->
-<!--                       "Unaligned: A or B?", "Aligned: C or D?") -->
-<!--   ) %>% -->
-<!-- #  filter(Response == "Correct") %>% -->
-<!--   ggplot(aes(x = Question, fill = Response, weight = weight)) + -->
-<!--   geom_bar() + -->
-<!--   ylab("Number of (effective) Respondents") + -->
-<!--   xlab("") + -->
-<!--   scale_fill_manual(values=c("grey", "forestgreen")) + -->
-<!--   scale_y_continuous( -->
-<!--       sec.axis = sec_axis(~./sum(round_12$weight)*100, -->
-<!--                           name="Percent responses") -->
-<!--   ) + -->
-<!--   theme(legend.position = "bottom") + -->
-<!--   ggtitle("(a) Accuracy of survey responses")  + -->
-<!--   geom_rect( -->
-<!--     aes( -->
-<!--       x = rev(c("Aligned: C or D?", "Unaligned: A or B?")), -->
-<!--       xmin = 2:1 - 0.45, -->
-<!--       ymin = ResponseTRUE - 1.96* se.ResponseTRUE, -->
-<!--       xmax = 2:1 + 0.45, -->
-<!--       ymax = ResponseTRUE + 1.96* se.ResponseTRUE, -->
-<!--       weight = 1), -->
-<!--     fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25, -->
-<!--     data = marginal_results -->
-<!--   ) -->
-
-<!-- gg2 <- round_12 %>% -->
-<!--   mutate( -->
-<!--     `Unaligned: A or B?` = ifelse(CorrectAB, "Correct", "Not correct"),  -->
-<!--     `Aligned: C or D?` = ifelse(CorrectCD, "Correct", "Not correct") -->
-<!--     ) %>% -->
-<!--   pcp_select(`Aligned: C or D?`, `Unaligned: A or B?`) %>% -->
-<!--   pcp_scale() %>% -->
-<!--   pcp_arrange() %>% -->
-<!--   ggplot(aes_pcp()) +  -->
-<!--     geom_pcp_axes() +  -->
-<!--     geom_pcp(aes(colour = CorrectABCD), alpha = 0.2)+ -->
-<!--     geom_pcp_boxes(fill=NA, colour = "black", linewidth=0.25) +  -->
-<!--     geom_pcp_labels(colour = "black") + -->
-<!--   scale_y_continuous("Number responses", breaks=seq(0,1,by=0.25),  -->
-<!--                      labels = round(sum(round_12$weight)*seq(0,1,by=0.25),0), -->
-<!--                      sec.axis = sec_axis(~.*100, -->
-<!--                           name="Percent responses")) + -->
-<!--   xlab("") +  -->
-<!--   scale_colour_manual(values = c("grey", "#84AC84", "#70A570", "forestgreen")) + -->
-<!--   ggtitle("(b) Combinations of responses") + -->
-<!--   theme(legend.position = "none") -->
-
-<!-- gg1 + gg2 -->
-<!-- ``` -->
-
 ### Respondent behavior

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
accuracy narrative pieces

@@
 
-Each task is made up of two elements: a visual stimulus and a question about that image that the viewer is asked to respond to.
+Each task is made up of two elements: a visual stimulus and a question about that stimulus that the viewer is asked to respond to.
 In our study, each visual stimulus is an image of a data visualization, while each question prompts viewers to identify which of two marked pieces in the data visualization is larger.
@@
 ```
-What we call 'aligned' and 'unaligned', here, is similar to Cleveland and McGill's set of rankings, but with some modifications: both 'aligned' and 'unaligned' bars (in tests 1 and 2) or wedges (in test 3) share the same axis.
+What we call 'aligned' and 'unaligned', here, is similar to Cleveland and McGill's set of rankings, but with some modifications: both 'aligned' and 'unaligned' bars share the same axis.
 Aligned tiles are additionally anchored in the same position in one dimension, i.e. the difference between their sizes can be reduced to a positional assessment.
@@
 
-Description of respondents
-
--   \# of respondents in each round
-
--   broad overview of demo characteristics(?)
-
--   raw sample size and effective sample size that we are analyzing **in each test**
+All responses were combined into one set of survey responses, with adjusted combined sample weights and indicators for which task each respondent was exposed to.
+Across the distinct stimuli, we model the resulting accuracy of responses, metrics on viewer interaction with the chart, and differences across demographic groups.
+
+\kr{Insert figure here of demographic breakdown across tasks or rounds?}
 
 ### Accuracy of responses
+
+We first investigate respondent accuracy in selecting the correct response.
+As participants were able to select the option 'They are the same', there are several ways to model accuracy and response.
+The argument could be made that for the purposes of practical interpretation, 'they are the same' is a correct choice, as the options are visually very similar and not substantially different values within the context of the data shown in the chart.
+However, we are interested in understanding whether viewers *can* perceive the difference and correctly identify which piece is larger.
+We consider multiple ways of modeling response to investigate participant response patterns.
+First, we define a measure of binary 'correctness', for which all answers that are not the correct option (B, D is bigger) are 'incorrect', including the selection of 'they are the same'.
+
+\autoref{accuracy}(a) displays all responses along the binary correctness measure, separated by whether the stimulus was an aligned task or unaligned task.
+We can see that levels of accuracy for all responses are significantly higher for the 'easier' (aligned) task, with about twice as many respondents correctly selecting the larger of the two marked elements.
+
+\kr{Can we update the figure and t-test with combined data across all tasks/all rounds 1-3?}
 
@@
 
-<!-- The data used for assessing the accuracy of comparisons in \autoref{tbl-tasks-12} is collected in two rounds of the NORC Omnibus survey. -->
-<!-- Rounds 1 and 2 are combined by adjusting the weights with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`. -->
-<!-- Figure \ref{fig-alignment}(a) shows that more than twice the number of responses are correct, when the tiles are aligned along the same axis. -->
-<!-- Because each participant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts. -->
-<!-- The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: \< 2.2e-16). -->
+\kr{ Can we update this to include the values for ROund 3 as well?}
+Rounds 1 and 2 are combined by adjusting the weights with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`.
+Figure \ref{fig-alignment}(a) shows that more than twice the number of responses are correct, when the tiles are aligned along the same axis.
+Because each participant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts.
+The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: \< 2.2e-16).
 
@@
 
-We combine rounds 1, 2 and 3 as shown in Table above to incorporate all relevant responses to evaluate the three different designs of charts in terms of their accuracy in assessing small differences in tile sizes.
+Next, we consider an ordinal model to investigate response behavior across all three options -- 'A (C) is bigger', 'B (D) is bigger', and 'they are the same', and consider these response patterns across each of the stimuli.
 
@@
 \text{logit }P(Y_k \le \ell) = \mu_{ij\ell(k)},
-$$ where $\ell \in \{1, 2\}$; $i \in \{1, 2\}$ is the comparison type (1 = Aligned, 2 = Unaligned), and $j \in \{1, 2, 3\}$ is the chart design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide.
+$$
+
+where $\ell \in \{1, 2\}$; $i \in \{1, 2\}$ is the comparison type (1 = Aligned, 2 = Unaligned), and $j \in \{1, 2, 3\}$ is the chart design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide.
 The estimated values and 95% confidence intervals are shown in @tbl-rounds-123
@@
 
-Interestingly, while we should, theoretically, see an improvement in accuracy when shifting from the vertical to the horizontal wide design, the resulting effects on the accuracy of the responses are not completely straightforward: The shift from a vertical to the (tall) horizontal design is detrimental to an accurate perception for both aligned and unaligned comparisons.\
-
+The same pattern in accuracy holds across each of the three structural variations; the aligned task has a higher level of accuracy than its unaligned counterpart.
+Interestingly, while we expect an improvement in accuracy when shifting from the horizontal to the horizontal wide design given the larger difference in pixel length between the two pieces, the resulting effects on the accuracy of the responses are not completely straightforward: The shift from a vertical to the (tall) horizontal design is detrimental to an accurate perception for both aligned and unaligned comparisons.
 The re-scaled design of the wide horizontal bars reclaims some of the loss for unaligned bars and outperforms the vertical design by a similar margin in aligned bars, but does not out-perform the vertical design when comparing unaligned tiles.

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
pdf

--- 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
@@
   \usepackage{hyperref}
+documentclass: jds
 format:
-  html: default
   pdf: 
-    documentclass: jdsart
     include-in-header: authors.tex
     keep-tex: true
+  html: default
 editor: 

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
with current pdf

@@
 #| label: fig-tasks
-#| fig-cap: "Each of the visual stimuli presented to viewers. In each bar chart, two pieces are marked. The larger piece in each chart is piece A or C. "
+#| fig-cap: "Each of the visual stimuli presented to viewers. In each bar chart, two pieces are marked. The larger piece in each chart are pieces B and D. "
 #| fig-subcap:
-#| - "**Vertical stacked bar, unaligned**"
-#| - "**Horizontal stacked bar, unaligned**"
-#| - "**Wide horizontal stacked bar, unaligned**"
-#| - "**Vertical stacked bar, aligned**"
-#| - "**Horizontal stacked bar, aligned**"
-#| - "**Wide horizontal stacked bar, aligned**"
+#| - "Vertical stacked bar, unaligned"
+#| - "Horizontal stacked bar, unaligned"
+#| - "Wide horizontal stacked bar, unaligned"
+#| - "Vertical stacked bar, aligned"
+#| - "Horizontal stacked bar, aligned"
+#| - "Wide horizontal stacked bar, aligned"
 #| layout-nrow: 2

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
single spacing 16 pages + references

@@
   \usepackage{hyperref}
+  \usepackage{setspace}
 documentclass: jds
 format:
-  html: default
   pdf: 
@@
   \setlength{\parindent}{0pt}
+  \singlespacing
 ```

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
some more on zooming behavior

--- 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
@@
 
-A contributing factor to this outcome might be the way that participants interact with the different designs.
+A contributing factor to the observed patterns in response accuracy might be the way that participants interact with the different designs.
 Generally, about half of all participants make use of the option to zoom into charts - while zooming does help with the overall accuracy (which is in agreement with the findings by @luModelingJustNoticeable2022 about the physical size of stimuli), the increase is not significant.
@@
 Let $Y_{jk}$ describe the zooming behavior of panelist $k$ on task $j$.
-We model zooming behavior (no = 0, yes = 1) as a logistic regression by correctness of response ($\rho$), task ($\tau$), and design ($\delta$) of the chart: $$
+We model zooming behavior (no = 0, yes = 1) as a logistic regression by correctness of response ($\rho$), task ($\tau$), and design ($\delta$) of the chart:
+
+$$
 \text{logit } P(Y_k \le 1) = \mu + \rho_{i(k)} + \tau_{j(k)} + \delta_{\ell(k)},
@@
 
-I'll pretty up @tbl-zoom-123 if we are going to keep it.
+I'll pretty up @tbl-zoom-123 if we are going to keep it. \kr{I think we should! it's nice to have a model for this}
 
@@
 
+However, zooming behavior *only* differs significantly by structural design -- rates do not differ significantly between the aligned and unaligned tasks, nor do they differ significantly by the chosen response.
+
+\kr{Do we have anything about total time spent on tasks across the 3 structures? We have time on each question for rounds 2-3. And can we pull in some of the certainty stuff?}
+
 ### Differences across demographic groups
@@
 We use a generalized cumulative logistic regression, where $\mu_i$ are intercepts $1 \le i < 3$, $X_k$ are demographics of the $k$th participant (in form of the model matrix), and $\beta_{i}$ are the coefficients.
+
 $$

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
table of rounds with lambda

--- 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
@@
 
-knitr::opts_chunk$set(
-  jds.rmd::pdf_article()$knitr
-)
-
 library(readr)
@@
 
+```{r combine-123}
+lambda_1_2 <- calculate_lambda(round1_adj$weight, round2_adj$weight)
+
+round_12 <- round1_adj %>% 
+            mutate(weight = lambda_1_2*weight/sum(weight),
+                   round = "Round 1") %>% 
+            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
+                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, weight,
+                   age4, gender, income4, racethnicity, educ5) %>%
+    rbind(round2_adj %>% 
+            mutate(weight = (1-lambda_1_2)*weight/sum(weight),
+                   round = "Round 2") %>% 
+            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, 
+                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, weight,
+                   age4, gender, income4, racethnicity, educ5))
+
+lambda_12_3 <- calculate_lambda(round_12$weight, round3_adj$weight)
+
+lambdas <- c(lambda_1_2, 1-lambda_1_2)*lambda_12_3
+lambdas <- c(lambdas, 1-sum(lambdas))
+
+round_123 <- round_12 %>% 
+              mutate(weight = lambda_12_3*weight/sum(weight),
+                     chart = "Vertical") %>% 
+  rbind(
+    round3_adj %>% 
+      mutate(
+        weight = (1-lambda_12_3)*weight/sum(weight),
+        round = "Round 3",
+        CorrectAB = ab == "B is bigger",
+        CorrectCD = cd == "D is bigger",
+        CorrectABCD = paste(CorrectAB, CorrectCD),
+        chart = ifelse(P_EXP==1, "Horizontal", "Horizontal wide")
+       ) %>% 
+        dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
+                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, chart, weight,
+                   age4, gender, income4, racethnicity, educ5) 
+  )
+
+round_123$weight <- neff(round_123$weight)/sum(round_123$weight)*round_123$weight 
+# does not affect the further analysis, but helps with the interpretation of results
+
+round_123_long <- round_123 %>% 
+  pivot_longer(ab:cd, names_to = "task", values_to = "Response")
+
+round_123_long <- round_123_long %>% 
+  mutate(
+    correct = Response %in% c("B is bigger", "D is bigger"),
+    correct3 = correct | (Response == "They are the same"),
+    task_chart = factor(paste(task, chart, sep="-")),
+    zoom = ifelse(task=="ab", ab_zoom, cd_zoom),
+    response3 = ifelse(correct, "Correct", ifelse(correct3, "They are the same", "Wrong"))
+  )
+
+survey_123 <- 
+  survey::svydesign(
+  data = round_123_long, 
+  ids = ~CaseId, weights = ~weight)
+```
+
 ## Introduction
@@
 \kr{Should the abstract match -- or be close to -- our SDSS short abstract?}
+\hh{I don't think it has to}
 
@@
 \kr{I don't know if the prior statement is going too far? Maybe we can rephrase.}
-
+\hh{I think this is going to hold - the demographics are pretty significant}
 <!-- Bias from crowd-sourcing? sample size?  -->
@@
 
+```{=tex}
+\hh{I think we should include the floating bars here to demonstrate that they are worse than the wide horizontal bars. }
+```
+
 ```{r}
 #| label: fig-tasks-explain
-#| fig-cap: Comparisons made in charts across tasks 1 through 3 within the Cleveland and McGill ranking
+#| fig-cap: Comparisons made in charts within the Cleveland and McGill ranking 
+#| out-width: 60%
 
@@
 #| layout-nrow: 2
-#| out-width: 100%
+#| out-width: 90%
 knitr::include_graphics("images/vertical-ab.PNG")
@@
 We describe it, but I feel like this table kind of muddies the waters.
-Could we instead do a table of respondents by task?}
-
-| Name    | Date       |      \# Participants |                effective sample size |         Sum of weights $\sum_i w_i$ |
-|---------------|---------------|--------------:|--------------:|--------------:|
-| Round 1 | April 2022 | `r nrow(round1_adj)` | `r round(neff(round1_adj$weight),1)` | `r round(sum(round1_adj$weight),1)` |
-| Round 2 | May 2022   | `r nrow(round2_adj)` | `r round(neff(round2_adj$weight),1)` | `r round(sum(round2_adj$weight),1)` |
-| Round 3 | Jun 2022   | `r nrow(round3_adj)` | `r round(neff(round3_adj$weight),1)` | `r round(sum(round3_adj$weight),1)` |
-|         |            |                      |                                      |                                     |
-
-: Survey rounds: dates, number of participants (nominal sample size), effective sample size, and sum of weights. {#tbl-rounds}
+Could we instead do a table of respondents by task?} \hh{does this comment still stand after the re-organization?
+...
+I would suggest to move it up to the survey methods and add the lambdas}
+
+| Name    | Date       |      \# Participants |                effective sample size |         Sum of weights $\sum_i w_i$ | $\lambda_i$ |
+|---------------|---------------|--------------:|--------------:|--------------:|--------------:|
+| Round 1 | April 2022 | `r nrow(round1_adj)` | `r round(neff(round1_adj$weight),1)` | `r round(sum(round1_adj$weight),1)` | $\lambda_1$  =  `r round(lambdas[1],3)` |
+| Round 2 | May 2022   | `r nrow(round2_adj)` | `r round(neff(round2_adj$weight),1)` | `r round(sum(round2_adj$weight),1)` | $\lambda_2$  = `r sprintf("%.3f",lambdas[2])` |
+| Round 3 | Jun 2022   | `r nrow(round3_adj)` | `r round(neff(round3_adj$weight),1)` | `r round(sum(round3_adj$weight),1)` | $\lambda_3$  = `r round(lambdas[3],3)` |
+| **Total**        |    ---        |    `r nrow(round_123)`   |  `r round(neff(round_123$weight),1)`   |    |  $\lambda_1w_1 + \lambda_2w_2 + \lambda_3w_3$ |
+
+: Survey rounds: dates, number of participants (nominal sample size), effective sample size, sum of weights, and factors for the convex combination as discussed above. {#tbl-rounds}
 
@@
 \kr{Insert figure here of demographic breakdown across tasks or rounds?}
+
+```{r}
+
+sum(round_123$weight)
+
+```
 
@@
 The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: \< 2.2e-16).
-
-```{r combine-123}
-lambda_1_2 <- calculate_lambda(round1_adj$weight, round2_adj$weight)
-
-round_12 <- round1_adj %>% 
-            mutate(weight = lambda_1_2*weight/sum(weight),
-                   round = "Round 1") %>% 
-            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
-                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, weight,
-                   age4, gender, income4, racethnicity, educ5) %>%
-    rbind(round2_adj %>% 
-            mutate(weight = (1-lambda_1_2)*weight/sum(weight),
-                   round = "Round 2") %>% 
-            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, 
-                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, weight,
-                   age4, gender, income4, racethnicity, educ5))
-
-lambda_12_3 <- calculate_lambda(round_12$weight, round3_adj$weight)
-
-round_123 <- round_12 %>% 
-              mutate(weight = lambda_12_3*weight/sum(weight),
-                     chart = "Vertical") %>% 
-  rbind(
-    round3_adj %>% 
-      mutate(
-        weight = (1-lambda_12_3)*weight/sum(weight),
-        round = "Round 3",
-        CorrectAB = ab == "B is bigger",
-        CorrectCD = cd == "D is bigger",
-        CorrectABCD = paste(CorrectAB, CorrectCD),
-        chart = ifelse(P_EXP==1, "Horizontal", "Horizontal wide")
-       ) %>% 
-        dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
-                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, chart, weight,
-                   age4, gender, income4, racethnicity, educ5) 
-  )
-
-round_123$weight <- neff(round_123$weight)/sum(round_123$weight)*round_123$weight 
-# does not affect the further analysis, but helps with the interpretation of results
-
-round_123_long <- round_123 %>% 
-  pivot_longer(ab:cd, names_to = "task", values_to = "Response")
-
-round_123_long <- round_123_long %>% 
-  mutate(
-    correct = Response %in% c("B is bigger", "D is bigger"),
-    correct3 = correct | (Response == "They are the same"),
-    task_chart = factor(paste(task, chart, sep="-")),
-    zoom = ifelse(task=="ab", ab_zoom, cd_zoom),
-    response3 = ifelse(correct, "Correct", ifelse(correct3, "They are the same", "Wrong"))
-  )
-
-survey_123 <- 
-  survey::svydesign(
-  data = round_123_long, 
-  ids = ~CaseId, weights = ~weight)
-```
 

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
update to figure and t-test

@@
 
-\kr{Can we update the figure and t-test with combined data across all tasks/all rounds 1-3?}
-
-```{r combine_rounds_1_and_2}
-
-lambda <- calculate_lambda(round1_adj$weight, round2_adj$weight)
-round_12 <- round1_adj %>%
-            mutate(weight = lambda[1]*weight) %>%
-            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight) %>%
-    rbind(round2_adj %>%
-            mutate(weight = (1-lambda[1])*weight) %>%
-            dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD, weight))
-round_12_long <- round_12 %>%
-  pivot_longer(CorrectAB:CorrectCD, names_to = "Question", values_to = "Response")
-
-survey_12 <-
-  survey::svydesign(
-  data = round_12_long,
-  ids = ~CaseId, weights = ~weight)
-```
+\kr{Can we update the figure and t-test with combined data across all tasks/all rounds 1-3?}\hh{done}
 
 ```{r ttest}
-paired <- svyttest(Response~Question, survey_12)
-abcd <- svyby(~Response, by=~Question, survey_12, svytotal)
+paired <- svyttest(correct~task, survey_123)
+abcd <- svyby(~correct, by=~task, survey_123, svytotal)
 
@@
 
-\kr{ Can we update this to include the values for ROund 3 as well?}
-Rounds 1 and 2 are combined by adjusting the weights with $\lambda$ = `r round(lambda[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`.
+\kr{ Can we update this to include the values for ROund 3 as well?}\hh{done}
+<!--Rounds 1 and 2 are combined by adjusting the weights with $\lambda$ = `r round(lambdas[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`.-->
+
 Figure \ref{fig-alignment}(a) shows that more than twice the number of responses are correct, when the tiles are aligned along the same axis.
@@
 The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: \< 2.2e-16).
-
-\kr{We need to update this chart to include the data from all 3 rounds instead of just the first 2}
+\hh{how about: $p$-value: $<< 0.0001$ -- the e-16 looks very computery}
+
+\kr{We need to update this chart to include the data from all 3 rounds instead of just the first 2} \hh{done}
 
@@
 
-gg1 <- round_12_long %>%
-  mutate(
-    Response = ifelse(Response, "Correct", "Not correct"),
+gg1 <- round_123_long %>%
+  mutate(
+    Response = ifelse(correct, "Correct", "Not correct"),
     Response = factor(Response, levels=c("Not correct", "Correct")),
-    Question = ifelse(Question=="CorrectAB",
+    Question = ifelse(task=="ab",
                       "Unaligned: A or B?", "Aligned: C or D?")
@@
   scale_y_continuous(
-      sec.axis = sec_axis(~./sum(round_12$weight)*100,
+      sec.axis = sec_axis(~./sum(round_123$weight)*100,
                           name="Percent responses")
@@
       xmin = 2:1 - 0.45,
-      ymin = ResponseTRUE - 1.96* se.ResponseTRUE,
+      ymin = correctTRUE - 1.96* se.correctTRUE,
       xmax = 2:1 + 0.45,
-      ymax = ResponseTRUE + 1.96* se.ResponseTRUE,
+      ymax = correctTRUE + 1.96* se.correctTRUE,
       weight = 1),
@@
 
-gg2 <- round_12 %>%
+gg2 <- round_123 %>%
   mutate(
@@
     geom_pcp_labels(colour = "black") +
-  scale_y_continuous("Number responses", breaks=seq(0,1,by=0.25),
-                     labels = round(sum(round_12$weight)*seq(0,1,by=0.25),0),
+  scale_y_continuous("Number of (effective) Respondents", breaks=seq(0,1,by=0.25),
+                     labels = round(sum(round_123$weight)*seq(0,1,by=0.25),0),
                      sec.axis = sec_axis(~.*100,

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
shorten levels for income and education

@@
 
+levels(round_123$educ5) <- c("Less than HS", "HS or equivalent", "Some college", "Bachelor", "Post graduate")
+
+levels(round_123$income4) <- c("< $30k", "$30k - $60k", "$60k - $100k", "more than $100k")
+
 round_123$weight <- neff(round_123$weight)/sum(round_123$weight)*round_123$weight 
@@
 #|   - Comparisons of unaligned tiles
-#| fig-height: 5
+#| fig-height: 4
 #| fig-width: 7

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
turn the table

@@
   \usepackage{setspace}
+  \usepackage{lscape}
 documentclass: jds
@@
   pdf: 
-    include-in-header: authors.tex
+    includes:
+      in_header: 'gt_packages.sty'
+    include-in-header: 
+      text: |
+        \usepackage{lscape}
+        \author[1]{Kiegan Rice\thanks{Corresponding author. Email: rice-kiegan@norc.org}}
+        \author[2]{Heike Hofmann\footnote{Email: hofmann@iastate.edu}}
+        \author[1]{Nola du Toit\footnote{Email: dutoit-nola@norc.org}} 
+        \author[1]{Edward Mulrow\footnote{Email: mulrow-edward@norc.org}}
+        \affil[1]{NORC at the University of Chicago}
+        \affil[2]{Department of Statistics, Iowa State University}
     keep-tex: true
@@
 ```{=latex}
-  \newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
+  \newcommand{\hh}[1]{{\textcolor{orange}{#1}}} % take out 
   \newcommand{\kr}[1]{{\textcolor{teal}{#1}}}
+
+  \newcommand{\blandscape}{\begin{landscape}}
+  \newcommand{\elandscape}{\end{landscape}}
+
   \setlength{\parindent}{0pt}
@@
 
+\blandscape
+
 ```{r demographics-table}
@@
 #| tbl-cap: Demographics matter for perception, particularly when the tasks get harder.
+
  coefs_clean <- function(dframe) {
@@
   tab_header(title = "Log odds of accuracy by task and demographics of respondents") %>%
+  tab_footnote("Signif. codes:  0 `***' 0.001 `**' 0.01 `*' 0.05 `.' 0.1 ` ' 1") %>%
  tab_options(
    column_labels.font.weight = "bold",
-   row_group.font.weight = "bold"
+   row_group.font.weight = "bold",
+   data_row.padding.horizontal = px(5),
+   page.orientation = "landscape"
  ) %>%
@@
     1~px(50),
-    contains("p.value") ~ px(20)
+    contains("p.value") ~ px(10)
+  ) %>%
+  cols_align(
+    align = "left",
+    columns = contains("p.value")
   ) %>%
@@
 
+\elandscape
+
 Results to discuss:

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
demographics overview across rounds

@@
 ```{r}
-
-sum(round_123$weight)
+#| label: fig-rounds-demographics
+#| fig-cap: "Demographics of respondents"
+#| fig-height: 7
+#| fig-width: 10
+
+round_123 <- round_123 %>% group_by(round) %>% 
+  mutate(
+    weight_pct_round = weight/sum(weight)*100
+  )
+
+gender <- round_123 %>% 
+  ggplot(aes(x = gender, weight = weight_pct_round)) + geom_bar() + 
+  ylab("Percent") + 
+  ggtitle("Gender") +
+  theme(axis.text.x = element_text(angle=30, hjust = 1)) +
+  xlab("") +
+  facet_grid(.~round)
+
+age <- round_123 %>% 
+  ggplot(aes(x = age4, weight = weight_pct_round)) + geom_bar() + 
+  ylab("Percent") + 
+  ggtitle("Age ") +
+  theme(axis.text.x = element_text(angle=30, hjust = 1)) +
+  xlab("") + ylim(c(0,35)) +
+  facet_grid(.~round)
+
+education <- round_123 %>% 
+  ggplot(aes(x = educ5, weight = weight_pct_round)) + geom_bar() + 
+  ylab("Percent") + 
+  ggtitle("Education levels") +
+  theme(axis.text.x = element_text(angle=30, hjust = 1)) +
+  xlab("") + ylim(c(0,35)) +
+  facet_grid(.~round)
+
+income <- round_123 %>% 
+  ggplot(aes(x = income4, weight = weight_pct_round)) + geom_bar() + 
+  ylab("Percent") + 
+  ggtitle("Income levels") +
+  theme(axis.text.x = element_text(angle=30, hjust = 1)) +
+  xlab("") + ylim(c(0,35)) +
+  facet_grid(.~round)
+
+(gender + education + plot_layout(nrow=1, widths=c(2,5))) / 
+  (age + income + plot_layout(nrow=1, widths = c(4,4)))
 
@@
 
-@tbl-demographics doesn't survive the formatting to pdf very well - \hh{need to shorten the row names or put into multiple lines}
+```{=tex}
+\hh{@tbl-demographics doesn't survive the formatting to pdf very well - need to shorten the row names or put into multiple lines} \hh{table rotates in pdf}
+```
 

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
smoothing out across the whole thing, filling in results and conclusions

@@
 ---
-title: "Testing Charts: viewer's perceptual accuracy in surveys"
+title: "Testing Charts: viewers' perceptual accuracy in surveys"
 short-title: Testing charts in surveys # running title
@@
 @clevelandGraphicalPerceptionTheory1984 observed as such, and in their seminal study defined the better visual among a pair as the one that allows viewers to make more accurate comparisons.
-Based on mappings of quantitative variables to different graphical elements, Cleveland and McGill's study resulted in a ranking of perceptual tasks from most accurate to least accurate, which was then extended by @mackinlayAutomatingDesignGraphical1986 to a theoretical framework ranking tasks' order along their ordinal and nominal scales, as shown in @fig-perceptual-tasks.
+Based on mappings of quantitative variables to different graphical elements, Cleveland and McGill's study resulted in a ranking of perceptual tasks from most accurate to least accurate, which was then extended by @mackinlayAutomatingDesignGraphical1986 to a theoretical framework ranking tasks' order along their ordinal and nominal scales.
 
 Cleveland and McGill's work -- while a foundational user study in graphical perception -- utilized a small convenience sample, consisting of only a few individuals recruited from among the authors' coworkers and their spouses.
-@heerCrowdsourcingGraphicalPerception2010 reproduced Cleveland and McGill's rankings using a larger sample from a crowd sourcing platform.
-A total of XXX Amazon Mechanical Turkers were involved.
+@heerCrowdsourcingGraphicalPerception2010 reproduced Cleveland and McGill's rankings using a larger sample from a crowd sourcing platform, employing a total of XXX Amazon Mechanical Turk (mTurk) workers for the study.
 Crowd sourced samples were shown by @borgoCrowdsourcingInformationVisualization2017 to be biased towards more male, younger, and relatively higher education relative to the adult U.S.
-population as a whole.
-These study populations are thus not representative of the general population, a common target audience for data visualization and scientific communication work.
+population as a whole; generally, convenience and opt-in samples are not representative of the general population, a common target audience for data visualization and scientific communication work.
 Further, the populations' emphasis on higher education individuals also leads to results which hold for groups of individuals who may be more likely to have prior exposure to data visualization in the context of scientific communication, or more exposure to data topics in higher education, but may not hold across other groups within the population.
-\kr{I don't know if the prior statement is going too far? Maybe we can rephrase.}
-\hh{I think this is going to hold - the demographics are pretty significant}
 <!-- Bias from crowd-sourcing? sample size?  -->
@@
 
-```{r}
-#| label: fig-perceptual-tasks
-#| fig-cap: Ranking of perceptual tasks, as given by @mackinlayAutomatingDesignGraphical1986.
+<!-- # ```{r} -->
+<!-- # #| label: fig-perceptual-tasks -->
+<!-- # #| fig-cap: Ranking of perceptual tasks, as given by @mackinlayAutomatingDesignGraphical1986.
 The ranking of tasks on the quantitative scale are empirically verified by @clevelandGraphicalPerceptionTheory1984.
-knitr::include_graphics("images/mackinlay-tasks.png")
-```
+-->
+<!-- # knitr::include_graphics("images/mackinlay-tasks.png") -->
+<!-- # ``` -->
 
@@
 
-**What specifically is each task?** Each comparison between marked pairs is designed to be a difficult task, with the difference between the values represented in the two marked pieces being a just-noticeable difference.
+The comparison between sets of marked pairs is designed to be a difficult task, with the difference between the values represented in the two marked pieces being a just-noticeable difference.
 The **Just-Noticeable Difference** (JND) is defined as the smallest difference that will be detected 50% of the time.
@@
 
-**Why do we utilize just-noticeable differences?** We employ comparisons at the JND in our tasks in order to maximize our ability to identify the impact of design changes on viewer accuracy and behavior.
+We employ comparisons at the JND in our tasks in order to maximize our ability to identify the impact of design changes on viewer accuracy and behavior.
 Asking perception tasks in a survey differs from the controlled environment of a cognitive lab, where these kind of questions may usually be assessed.
@@
 With a small set of tasks, we need to present tasks that are perceptually hard, and thus ask questions about stimuli that are close to our perceptual threshold.
-Therefore, we focus on questions which vary the presented image, but ask viewers to compare the same values across those varied images.
-
-**How do we vary the task?** We ask participants to determine which of two just-noticeably different marked pieces is larger within the data visualization image, and we vary the structural design of that data visualization image.
-We focus on three main sets of structural variation in the design.
+Therefore, we focus on questions which vary the structure of the presented image, but ask viewers to compare the same underlying data values across those varied images.
+
+We ask participants to determine which of two just-noticeably different marked pieces is larger within each data visualization image, and focus on three main sets of structural variation in the design of that data visualization image.
 First, we vary the alignment of the pieces in question.
@@
 Aligned tiles are additionally anchored in the same position in one dimension, i.e. the difference between their sizes can be reduced to a positional assessment.
-Unaligned tiles do not share this anchor, however, the context of the other tiles in the chart provide a frame, which *should* help with an assessment of the tiles' sizes beyond a comparison of (arc) lengths or areas.\
-
-We would expect that comparing unaligned tiles is a harder task (with correspondingly lower levels of accuracy) than a comparison of aligned tiles, with the framing given by the context of the other tiles in the same column (or the same pie) mitigating some of this difficulty.
+Unaligned tiles do not share this anchor, however, the context of the other tiles in the chart provide a frame, which *should* help with an assessment of the tiles' sizes beyond a comparison of (arc) lengths or areas.
+
+We would expect that comparing unaligned tiles is a harder task (with correspondingly lower levels of accuracy) than a comparison of aligned tiles, with the framing given by the context of the other tiles in the same column mitigating some of this difficulty.
 @fig-tasks-explain gives an overview of the comparisons of tasks 1 through 3 and the closest corresponding tasks in Cleveland and McGill.
+\kr{The updated sketch is great here!! One note -- can we move the label for the "positions along the same, but unaligned scale", it's not aligned with the others}
 
@@
 
-**How do we present the task?** Our use of a survey format guides the format and design of the questions asked and how they are presented to respondents.
+Our use of a survey format guides the format and design of the questions asked and how they are presented to respondents.
 First, participant instructions must be delivered in a very short and easily understandable format, because participants cannot ask clarifying questions about the task as they might be able to in a cognitive lab setting.
@@
 For this reason, we utilize data on living arrangements of older U.S. adults -- a topic which most U.S. adults will have some familiarity with, but is not inherently polarizing.
-Finally, to prevent viewers from being exposed to slight variations of the same stimulus in a row (and risk unforeseen order effects or respondents using prior questions to inform their responses), we either split a survey sample in two and show each subsample a distinct version of the chart or test variations of a chart across distinct rounds of the survey.
-
+Finally, to prevent viewers from being exposed to slight variations of the same stimulus in a row (and risk unforeseen order effects or respondents using prior questions to inform their responses), we either split a survey sample in two and show each subsample a distinct structural version of the chart or test variations of a chart across distinct rounds of the survey.
+
+#### Question text  
 When viewing each chart, participants were asked to compare the relative sizes of marked elements within the chart:
@@
 
+For a given task, viewers are first presented with the unaligned version of the task, followed by the aligned version of the task.
 The time in seconds that each respondent spent on each task was recorded, as well as whether the participant zoomed in on each chart while answering the question.
@@
 \autoref{fig-tasks}(a,d) shows the two stacked bar charts shown to participants in the first task.
-The marked tiles in each plot are 155 pixels apart.
-Based on @luModelingJustNoticeable2022's model, a difference of 155 pixels leads to a just noticeable difference of 3.5 pixels.
-The heights of the bars are 205 (left) and 213 pixels (right), respectively, corresponding to about twice the JND.
+The marked tiles in each plot are 155 pixels apart, which leads to a JND of 3.5 pixels based on @luModelingJustNoticeable2022's model.
+The heights of the bars are 205 (left) and 213 pixels (right), corresponding to about twice the JND.
 This difference should lead to a relatively high accuracy rate for participants and simultaneously limit the amount of frustration resulting from a task that is perceived as 'too hard'.
@@
 Both charts show the same data with slight modifications to the order of the levels -- the first and second level in each of the bars are reversed between the left and the right chart.
-Both charts are shown at the same size, i.e. in both cases both the difference in size between the bars and the horizontal distance between the bars is the exact same amount.
+Both charts are displayed at the same size, i.e. in both cases both the difference in size between the bars and the horizontal distance between the bars is the exact same amount.
 This leaves the vertical positioning of the bars as the only difference between the charts.
@@
 
-The visual mappings in the second task, shown in \autoref{fig-tasks-3a}, are identical to the first task, but the axes of the chart are rotated so that the stacked bar is represented in a horizontal format.
+The visual mappings in the second task, shown in \autoref{fig-tasks}(b, e), are identical to the first task, but the axes of the chart are rotated so that the stacked bar is represented in a horizontal format.
 This represents a structural change in how the data are presented to the viewer while preserving the pixel size of the elements viewers must compare.
@@
 
-### Task 3: Wide horizontal stacked bar
-
-The images utilized in the third task again represent the same data, and the overall image has dimensions which are identical to the images in the first two tasks.
+### Task 3: Horizontal wide stacked bar
+
+The images utilized in the third task again represent the same data as in the first two tasks, and the overall image has dimensions which are identical to the images in the first two tasks, displaying at the same size during the survey.
 However, the aspect ratio of the plotting area is adjusted to increase the length in pixels of the stacked bar elements, and decrease the relative thickness \kr{wording?} of each bar. 
+\kr{Do we have a measurement of the pixels for this? I think a sentence on that would close out this section nicely.}
 
@@
 adults to answer questions on a variety of topics.
+Our tests require visual inspection of an image; for this reason, our survey questions were only presented to web-based panelists and not panelists who respond via phone interviews.
 
@@
 Given the nature of pulling a sample from the panel for each Omnibus round, there is a possibility that some participants may have been included in multiple rounds; however, our collected data is not a longitudinal or panel study, so we do not have repeated responses from the same participants across each of our tests.
-Each of the tests was run at least a month apart, so if respondents participated in multiple rounds, there was at least a one month gap between viewing each visual stimulus.
+Each of the tests was run at least a month apart, so if respondents did participate in multiple rounds, there was at least a one month gap between viewing each visual stimulus.
 
@@
 Note that @omuircheartaighCombiningSamplesVs2002 estimate $\lambda$ separately for any combination of race/ethnicity by sex.
-We employ that strategy whenever we include demographic variables in the analysis, otherwise we will use a single adjustment for the weights.
-
-The data for this paper were collected in several rounds as part of the NORC Omnibus.
+We employ that strategy whenever we include demographic variables in the analysis, otherwise we use a single adjustment for the weights.
+
+The data for this paper were collected in several rounds as part of the NORC Omnibus, and the resulting number of participants, effective sample sizes, and $\lambda$ values are shown in \autoref{tbl-rounds}.
 
@@
 <!-- : Survey rounds: dates, number of participants (nominal sample size), effective sample size, and sum of weights. {#tbl-rounds} -->
-
-# Results
-
-There are differences by alignment
-There are not huge differences across structures
-However, we see a difference in how people react across different structures (zooming behavior, certainty? time?)
-
-- We can reproduce prior findings with nat rep sample 
-- We can rank 
-- We can rank better because we have a bigger sample size 
-- We can find changes in viewer behavior 
-- Demographics matter to accuracy and choices
-
-### Respondents
-
-A total of `r sum(c(nrow(round1_adj), nrow(round2_adj), nrow(round3_adj)))` respondents participated across the three rounds.
-The number of responses and corresponding effective sample sizes in each round are shown in \autoref{tbl-rounds}.
 
@@
 ...
-I would suggest to move it up to the survey methods and add the lambdas}
+I would suggest to move it up to the survey methods and add the lambdas} \kr{I think moving it here helps, I did also add another table below.
+Let me know what you think.}
 
@@
 
+# Results
+
+### Respondents
+
+A total of `r sum(c(nrow(round1_adj), nrow(round2_adj), nrow(round3_adj)))` respondents participated across the three rounds.
+The number of responses and corresponding effective sample sizes in each round are shown in \autoref{tbl-rounds}.
 All responses were combined into one set of survey responses, with adjusted combined sample weights and indicators for which task each respondent was exposed to.
-Across the distinct stimuli, we model the resulting accuracy of responses, metrics on viewer interaction with the chart, and differences across demographic groups.
-
-\kr{Insert figure here of demographic breakdown across tasks or rounds?}
+The resulting sample sizes corresponding to each task are shown in \autoref{tbl-tasks}.
+\kr{Is this table useful?
+I want to be clear about the difference between tasks and rounds and make sure it's clear what our resulting sizes are..}
+
+| Name    | Description    |      \# Participants |                effective sample size |
+|---------------|---------------|--------------:|--------------:|
+| Task 1 | Vertical| `r nrow(round1_adj) + nrow(round2_adj)` | `r round(neff(round_12$weight),1)` |
+| Task 2 | Horizontal |`r nrow(round3_adj[round3_adj$P_EXP == 1,])` | `r round(neff(round3_adj[round3_adj$P_EXP == 1,]$weight),1)` |
+| Task 3 | Horizontal wide |`r nrow(round3_adj[round3_adj$P_EXP == 2,])` | `r round(neff(round3_adj[round3_adj$P_EXP == 2,]$weight),1)` |
+
+: Survey tasks: number of participants (nominal sample size) and effective sample size for each task. {#tbl-tasks}
+
+Distributions of demographic characteristics across each survey round are shown in \autoref{fig-rounds-demographics}.
+\kr{Split by rounds or by tasks?
+I feel like since we later split by task, maybe we should show the demos by task?
+Or we can just talk about how Round 3 was split randomly along demos for Tasks 2/3?}
 
@@
 The argument could be made that for the purposes of practical interpretation, 'they are the same' is a correct choice, as the options are visually very similar and not substantially different values within the context of the data shown in the chart.
-However, we are interested in understanding whether viewers *can* perceive the difference and correctly identify which piece is larger.
-We consider multiple ways of modeling response to investigate participant response patterns.
-First, we define a measure of binary 'correctness', for which all answers that are not the correct option (B, D is bigger) are 'incorrect', including the selection of 'they are the same'.
-
-\autoref{accuracy}(a) displays all responses along the binary correctness measure, separated by whether the stimulus was an aligned task or unaligned task.
+However, we are interested in understanding whether viewers *can* perceive the difference and correctly identify which piece is larger and thus consider multiple ways of modeling response to investigate participant response patterns.
+We begin by defining a measure of binary 'correctness', for which all answers that are not the correct option (B [D] is bigger) are 'incorrect', including the selection of 'they are the same'.
+
+\autoref{fig-alignment}(a) displays all responses along the binary correctness measure, separated by whether the stimulus was an aligned task or unaligned task.
 We can see that levels of accuracy for all responses are significantly higher for the 'easier' (aligned) task, with about twice as many respondents correctly selecting the larger of the two marked elements.
-
-\kr{Can we update the figure and t-test with combined data across all tasks/all rounds 1-3?}\hh{done}
 
@@
 
-\kr{ Can we update this to include the values for ROund 3 as well?}\hh{done}
 <!--Rounds 1 and 2 are combined by adjusting the weights with $\lambda$ = `r round(lambdas[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`.-->
 
-Figure \ref{fig-alignment}(a) shows that more than twice the number of responses are correct, when the tiles are aligned along the same axis.
-Because each participant was shown both versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts.
-The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: \< 2.2e-16).
-\hh{how about: $p$-value: $<< 0.0001$ -- the e-16 looks very computery}
-
-\kr{We need to update this chart to include the data from all 3 rounds instead of just the first 2} \hh{done}
+Because each participant was shown both the aligned and unaligned versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts.
+The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: $<< 0.0001$).
 
@@
 
-Next, we consider an ordinal model to investigate response behavior across all three options -- 'A (C) is bigger', 'B (D) is bigger', and 'they are the same', and consider these response patterns across each of the stimuli.
+Next, we consider an ordinal model to investigate response behavior across all three options -- 'A [C] is bigger', 'B [D] is bigger', and 'They are the same', and consider these response patterns across each of the stimuli.
 
@@
 where $\ell \in \{1, 2\}$; $i \in \{1, 2\}$ is the comparison type (1 = Aligned, 2 = Unaligned), and $j \in \{1, 2, 3\}$ is the chart design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide.
-The estimated values and 95% confidence intervals are shown in @tbl-rounds-123
+The estimated values and 95% confidence intervals are shown in @tbl-rounds-123.
 
@@
 
-The same pattern in accuracy holds across each of the three structural variations; the aligned task has a higher level of accuracy than its unaligned counterpart.
-Interestingly, while we expect an improvement in accuracy when shifting from the horizontal to the horizontal wide design given the larger difference in pixel length between the two pieces, the resulting effects on the accuracy of the responses are not completely straightforward: The shift from a vertical to the (tall) horizontal design is detrimental to an accurate perception for both aligned and unaligned comparisons.
+One pattern in accuracy holds across each of the three structural variations: the aligned task has a higher level of accuracy than its unaligned counterpart.
+Interestingly, while we expect an improvement in accuracy when shifting from the horizontal to the horizontal wide design given the larger difference in pixel length between the two pieces, the resulting effects on the accuracy of the responses are not completely straightforward: the shift from a vertical to the (tall) horizontal design is detrimental to an accurate perception for both aligned and unaligned comparisons.
 The re-scaled design of the wide horizontal bars reclaims some of the loss for unaligned bars and outperforms the vertical design by a similar margin in aligned bars, but does not out-perform the vertical design when comparing unaligned tiles.
 
+Rates of selecting the incorrect response are low across all three structural variations and the aligned and unaligned tasks; while viewers select the incorrect response at significantly higher rates for all three unaligned tasks relative to the aligned tasks, those rates do not differ significantly across the three structural variations.
+Most of the observed differences in response patterns across structural variations are attributed to respondents' selection between the correct option or the 'they are the same' option.
+
+To better comprehend these observed patterns, we investigate viewer interaction with the tasks more broadly as well as differences in response selection across demographic groups.
+
 ### Respondent behavior
 
-A contributing factor to the observed patterns in response accuracy might be the way that participants interact with the different designs.
-Generally, about half of all participants make use of the option to zoom into charts - while zooming does help with the overall accuracy (which is in agreement with the findings by @luModelingJustNoticeable2022 about the physical size of stimuli), the increase is not significant.
-However, different designs lead to different rates of zooming: when dealing with the vertical design, the rate of zooming is significantly higher than for the two horizontal designs.
+One contributing factor to the observed patterns in response accuracy might be the way that participants interact with the different designs.
+Across all tasks, about half of all participants make use of the option to zoom into charts.
+We observe that while zooming does help with the overall accuracy (which is in agreement with the findings by @luModelingJustNoticeable2022 about the physical size of stimuli), the increase is not significant.
+However, different designs lead to different rates of zooming: we observe in \autoref{fig-zoom-123} that when dealing with the vertical design, the rate of zooming is significantly higher than for the two horizontal designs.
 
@@
 
-Let $Y_{jk}$ describe the zooming behavior of panelist $k$ on task $j$.
+To formalize this pattern in a model, let $Y_{jk}$ describe the zooming behavior of panelist $k$ on task $j$.
 We model zooming behavior (no = 0, yes = 1) as a logistic regression by correctness of response ($\rho$), task ($\tau$), and design ($\delta$) of the chart:
@@
 
-I'll pretty up @tbl-zoom-123 if we are going to keep it. \kr{I think we should! it's nice to have a model for this}
+The resulting model estimates and confidence intervals are displayed in \autoref{tbl-zoom-123}.
+We can observe that zooming behavior *only* differs significantly by structural design; rates do not differ significantly between the aligned and unaligned tasks, nor do they differ significantly by the chosen response.
+This higher rate of zooming does not necessarily lead to higher accuracy; although rates of accuracy are significantly higher for the vertical orientation relative to the horizontal orientation, they are not significantly higher -- and in fact, are significantly lower on the aligned task -- than the accuracy for the horizontal wide orientation.
+\hh{I'll pretty up @tbl-zoom-123 if we are going to keep it.} \kr{I think we should!
+it's nice to have a model for this}
 
@@
 
-However, zooming behavior *only* differs significantly by structural design -- rates do not differ significantly between the aligned and unaligned tasks, nor do they differ significantly by the chosen response.
-
 \kr{Do we have anything about total time spent on tasks across the 3 structures? We have time on each question for rounds 2-3. And can we pull in some of the certainty stuff?}
@@
 ### Differences across demographic groups
+
+Finally, we turn to investigating responses across demographic groups, a core benefit provided by the large and representative nature of our survey samples.
+\autoref{fig-demographics} displays response patterns across our three structural variations by gender, age (4 groups), educational attainment (5 groups), and income level (4 groups).
+Our pattern of higher accuracy in selecting the correct response on the aligned task relative to the unaligned task holds across all demographic groups and structural variations.
+However, we do observe different response patterns, particularly on the unaligned task, across demographic groups.
 
@@
 $$
+The resulting model estimates and confidence intervals are shown in \autoref{tbl-demographics}.
+When considering the easier (aligned) task, response patterns do not differ significantly by age, gender, or education level.
+We do observe, however, a significantly higher log odds of selecting the 'correct' or 'they are the same' response (or significantly lower log odds of selecting 'incorrect') among those with an income between \$60,000 and \$100,000.
+When we consider the more difficult (unaligned) task, we see significant separation in response patterns across gender, educational attainment, and income groups.
+We do not see significant differences in responses by age for either task.
+Interestingly, we see lower log odds of selecting the 'correct' response among those with higher educational attainment (those with a bachelor's degree or post graduate study) *and* lower log odds of selecting the 'incorrect' response; as income and educational attainment increase, respondents are more likely to select the 'they are the same' option during the difficult task.
+This speaks to the difficulty of the task, and respondents' interpretation of the task when it is more difficult to perceive small differences.
 
@@
 
--   Differences in accuracy between aligned and unaligned bar
-
--   Differences across structures
-
--   Differences in timing and zooming behavior across all designs
-
--   Demographic differences and their interaction with alignment
-
-# Conclusion
-
-Key findings to discuss:
-
--   Surveys can be used to ask these types of questions -- in the midst of other topics and with a limited number of questions, we are able to ask perception questions and produce results consistent with prior studies
-
--   Reproduced (some) prior convenience sample results using a large, nationally-representative survey population:
-
-    -   Cleveland: position (aligned) versus unaligned
-    -   unaligned tiles within stacked bar are a hybrid between framed bar and floats
-    -   Talbot: pie charts - individual wedges vs piecharts
-
--   Design choices impact viewer accuracy:
-
-    -   new finding: pie charts accuracy in work better when framed (wedges)
+There are differences by alignment
+There are not huge differences across structures
+However, we see a difference in how people react across different structures (zooming behavior, certainty? time?)
+
+- We can reproduce prior findings with nat rep sample 
+- We can rank 
+- We can rank better because we have a bigger sample size 
+- We can find changes in viewer behavior 
+- Demographics matter to accuracy and choices
+
+# Conclusions
+
+Through this work, we have tested the use of probability-based survey panels to ask perception questions; within a large survey covering a variety of topics, and with a limited number of questions for our specific tasks, we are able to measure viewer perception and produce results consistent with prior studies.
+Testing data visualization design structures and viewer behavior on a nationally-representative sample is of broad interest and applicability to the scientific communication community, and this study demonstrates a framework under which to complete further work in this area.
+
+With our survey framework, we have shown that we can rank structural variations on design and tasks by their relative accuracy; we identified significantly different rates of respondent accuracy under different designs presenting the same data.
+The size of our sample aids in identifying this signal.
+
+Further, we can also study respondent behavior and how viewers interact with the chart.
+Paradata on viewer interaction with the chart (e.g., zooming and time spent on each task) can be collected as well as asking directly asking respondents questions about their certainty.
+Not only can we collect this information, but we can also identify significant differences in this behavior across different designs.
+Understanding viewer interaction and engagement is a key component to designing effective data visualizations.
+If a particular design leads viewers to zoom in more to investigate it when determining a response, we might consider that design to be more frustrating to viewers 'in the wild' when viewing it outside of the context of our tests.
+\kr{I don't love the last sentence but I think we can clean it up}.
+
+Perhaps most salient among our results are the patterns in response behavior across demographic groups.
+The large size of our sample, paired with the probability-based survey panel approach, afford us the ability to dive into response patterns among demographics across tasks and identify significant differences among them.
+In particular, the differences across educational attainment groups and income groups underscore the importance of utilizing a representative population when testing perception; prior results in this area may be absorbing bias in responses due to the larger representation of individuals with higher education levels among those study respondents.
+
+While our results are limited to mapping in a stacked bar chart with a specific topic, they demonstrate among them a wealth of insights about how respondents perceive and interact with data visualizations.
+More expansive work on this topic should be completed to understand how the general public interacts with and understands charts, including expanding to a wider set of structural variations, more aesthetic design variations, and considering differences in respondent behavior across different data topic areas.
+
+<!-- Key findings to discuss: -->
+
+<!-- -   Surveys can be used to ask these types of questions -- in the midst of other topics and with a limited number of questions, we are able to ask perception questions and produce results consistent with prior studies -->
+
+<!-- -   Reproduced (some) prior convenience sample results using a large, nationally-representative survey population: -->
+
+<!--     -   Cleveland: position (aligned) versus unaligned -->
+<!--     -   unaligned tiles within stacked bar are a hybrid between framed bar and floats -->
+<!--     -   Talbot: pie charts - individual wedges vs piecharts -->
+
+<!-- -   Design choices impact viewer accuracy: -->
+
+<!--     -   new finding: pie charts accuracy in work better when framed (wedges) -->
 

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
oops - commenting out some notes

@@
 
-Results to discuss:
-
-There are differences by alignment
-There are not huge differences across structures
-However, we see a difference in how people react across different structures (zooming behavior, certainty? time?)
-
-- We can reproduce prior findings with nat rep sample 
-- We can rank 
-- We can rank better because we have a bigger sample size 
-- We can find changes in viewer behavior 
-- Demographics matter to accuracy and choices
+<!-- Results to discuss: -->
+
+<!-- There are differences by alignment -->
+<!-- There are not huge differences across structures -->
+<!-- However, we see a difference in how people react across different structures (zooming behavior, certainty? time?) -->
+
+<!-- - We can reproduce prior findings with nat rep sample  -->
+<!-- - We can rank  -->
+<!-- - We can rank better because we have a bigger sample size  -->
+<!-- - We can find changes in viewer behavior  -->
+<!-- - Demographics matter to accuracy and choices -->
 
@@
 
--   Other important measures beyond accuracy: time to completion (automatic on web), certainty in response (asked).
-
--   Studying how viewers interact with charts can be done in a survey, and more expansive work on this topic should be completed to understand how the general public interacts with and understands charts.
-
-    -   Design choices impact viewer behavior: zooming
-
--   (some) Demographics matter: age and gender does not seem to matter for perception economic background and education does matter for perception: we need to know, we do not want to create a hurdle in communication
+<!-- -   Other important measures beyond accuracy: time to completion (automatic on web), certainty in response (asked). -->
+
+<!-- -   Studying how viewers interact with charts can be done in a survey, and more expansive work on this topic should be completed to understand how the general public interacts with and understands charts.
+-->
+
+<!--     -   Design choices impact viewer behavior: zooming -->
+
+<!-- -   (some) Demographics matter: age and gender does not seem to matter for perception economic background and education does matter for perception: we need to know, we do not want to create a hurdle in communication -->
 

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
pdf rendered, small changes to address kr's questions

@@
 \kr{The mackinlay thing feels like maybe we are overemphasizing the ranking done by cleveland/mcgill and then mackinlay - I think we could emphasize less if we want}
-
-Our work -- as that of Cleveland and McGill and Heer and Bostock -- centers around studying data visualization design choices and their impact on viewer behavior and accuracy of viewers' responses.
+\hh{my point was to emphasize that Mackinlay is a foundational paper (rightfully so), but not based on any data beyond Cleveland}
+
+<!--Our work -- as that of Cleveland and McGill and Heer and Bostock -- -->
+
+Our work also centers on studying data visualization design choices and their impact on viewer behavior and accuracy of viewers' responses.
 We seek to answer whether it is possible to reproduce some of their findings in the context of a survey with a large, nationally-representative set of respondents, and within that context we focus on the following research questions:
@@
 \kr{The updated sketch is great here!! One note -- can we move the label for the "positions along the same, but unaligned scale", it's not aligned with the others}
+\hh{that was on purpose :(  that goes beyond Cleveland and McGill ... I'll see what I can do about that}
 
@@
 There were no significant differences in respondent accuracy or behavior across color schemes, and we combine them here into one set of stimulus responses to compare structural differences.
-\kr{Should we put the color variations and a test for significance in accuracy/a visual in the appendix/supplementary materials? }
+\kr{Should we put the color variations and a test for significance in accuracy/a visual in the appendix/supplementary materials?
+} \hh{we could do that, but I'm not sure that we need to mention color at all.}
 
@@
 The images utilized in the third task again represent the same data as in the first two tasks, and the overall image has dimensions which are identical to the images in the first two tasks, displaying at the same size during the survey.
-However, the aspect ratio of the plotting area is adjusted to increase the length in pixels of the stacked bar elements, and decrease the relative thickness \kr{wording?} of each bar. 
-\kr{Do we have a measurement of the pixels for this? I think a sentence on that would close out this section nicely.}
+\hh{The length of bars is increased to fit the new dimensions. The height of the tiles are adjusted accordingly (from 50 to 30 pixels) to keep the overall area the same as before}
+<!--However, the aspect ratio of the plotting area is adjusted to increase the length in pixels of the stacked bar elements, and  decrease the relative thickness \kr{wording?} of each bar. -->
+\kr{Do we have a measurement of the pixels for this? I think a sentence on that would close out this section nicely.}\hh{like that?}
 
@@
 
-\hh{XXX I think what you are trying to say is that we do not know, if participants have seen the stimuli multiple times, but if they did, there was at least a month in between?} \kr{Yes that's what I was trying to say here!!}
-
-Given the nature of pulling a sample from the panel for each Omnibus round, there is a possibility that some participants may have been included in multiple rounds; however, our collected data is not a longitudinal or panel study, so we do not have repeated responses from the same participants across each of our tests.
+<!--\hh{XXX I think what you are trying to say is that we do not know, if participants have seen the stimuli multiple times, but if they did, there was at least a month in between?} \kr{Yes that's what I was trying to say here!!}-->
+
+\hh{The Omnibus sample is not longitudinal in nature, i.e.\ we do not have any information about whether the same panelists were included in multiple rounds.
+While there is a (small) chance that panelists are included in multiple rounds of our data, there was at least a one month gap between viewing each visual stimulus, and for the purposes of the analysis we will consider data from these viewings as independent.
+}
+
+<!-- Given the nature of pulling a sample from the panel for each Omnibus round, there is a possibility that some participants may have been included in multiple rounds; however, our collected data is not a longitudinal or panel study, so we do not have repeated responses from the same participants across each of our tests.
 Each of the tests was run at least a month apart, so if respondents did participate in multiple rounds, there was at least a one month gap between viewing each visual stimulus.
+-->
 
@@
 
-When combining responses that were gathered during distinct rounds of the Omnibus survey, we make a weighting adjustment to ensure that weights in each sample are properly calibrated to the population total across all rounds in the resulting model.
+When combining responses from different rounds of the Omnibus survey, we calibrate weights in each round to the total population  across all rounds in the resulting model.
 
@@
 
-We *combine* (rather than cumulate) surveys $S_1$ and $S_2$, as described in @omuircheartaighCombiningSamplesVs2002, by multiplying weights in $S_1$ and $S_2$ by $\lambda$ and $1 - \lambda$ respectively.
+We *combine* (rather than cumulate) two surveys $S_1$ and $S_2$, as described in @omuircheartaighCombiningSamplesVs2002, by multiplying weights in $S_1$ and $S_2$ by $\lambda$ and $1 - \lambda$ respectively.
+$\lambda \in [0,1]$ is given as
 $$
@@
 Here, $d_1$ and $d_2$ are estimated as $$
-d_1 = 1 + CV(w_i \in S_1)^2 \ \ \ \text{ and } \ \ \ d_2 = 1 + CV(w_i \in S_2)^2
+d_1 = 1 + CV(w_i \in S_1)^2 \ \ \ \text{ and } \ \ \ d_2 = 1 + CV(w_i \in S_2)^2,
 $$ where $CV$ is the coefficient of variation of the weights within each sample, and is estimated as in @kishSurveySampling1965 :
@@
 
-Note that @omuircheartaighCombiningSamplesVs2002 estimate $\lambda$ separately for any combination of race/ethnicity by sex.
-We employ that strategy whenever we include demographic variables in the analysis, otherwise we use a single adjustment for the weights.
+<!--Note that @omuircheartaighCombiningSamplesVs2002 estimate $\lambda$ separately for any combination of race/ethnicity by sex.
+We employ that strategy whenever we include demographic variables in the analysis, otherwise we use a single adjustment for the weights. -->
 
@@
 I would suggest to move it up to the survey methods and add the lambdas} \kr{I think moving it here helps, I did also add another table below.
-Let me know what you think.}
+Let me know what you think.}\hh{I like it!}
 
@@
 | Round 3 | Jun 2022   | `r nrow(round3_adj)` | `r round(neff(round3_adj$weight),1)` | `r round(sum(round3_adj$weight),1)` | $\lambda_3$  = `r round(lambdas[3],3)` |
-| **Total**        |    ---        |    `r nrow(round_123)`   |  `r round(neff(round_123$weight),1)`   |    |  $\lambda_1w_1 + \lambda_2w_2 + \lambda_3w_3$ |
+| **Total**        |    ---        |    `r nrow(round_123)`   |  `r round(neff(round_123$weight),1)`   |    |   |
 
@@
 \kr{Is this table useful?
-I want to be clear about the difference between tasks and rounds and make sure it's clear what our resulting sizes are..}
+I want to be clear about the difference between tasks and rounds and make sure it's clear what our resulting sizes are..}\hh{yes!}
 
@@
 I feel like since we later split by task, maybe we should show the demos by task?
-Or we can just talk about how Round 3 was split randomly along demos for Tasks 2/3?}
+Or we can just talk about how Round 3 was split randomly along demos for Tasks 2/3?}\hh{changed the figure to show demos by task -- do we need confidence intervals?
+-- or a comparison against census data?}
 
@@
 
-round_123 <- round_123 %>% group_by(round) %>% 
+round_123 <- round_123 %>% group_by(chart) %>% 
   mutate(
@@
   xlab("") +
-  facet_grid(.~round)
+  facet_grid(.~chart)
 
@@
   xlab("") + ylim(c(0,35)) +
-  facet_grid(.~round)
+  facet_grid(.~chart)
 
@@
 
+\hh{we could include certainty - I'll look into it}
+
 ### Differences across demographic groups
@@
 
+\hh{XXX the formatting isn't right yet here ... latex problem with large floats ... will fix it.}
+
 ```{r}
 #| label: fig-demographics
-#| fig-cap: "Panelist's demographics matter, particularly, when the task difficulty increases.
+#| fig-cap: "Panelists' demographics matter, particularly, when the task difficulty increases.
 For aligned tiles, gender, age, and education are not significant factors.
-However, income levels do have - a small - effect.
+However, income levels do have a (small) effect.
 When income levels increase, the percentage of wrong answers (orange) decreases, while the percentage of correct answers (blue) increases slightly (significant at below 0.05).
-The more difficult task of comparing unaligned tiles, demographics are more significant.
-XXX With increasing levels in education, we see a small drop in correct answers, as well as a small drop in wrong answers, for a significant increase in 'they are the same'."
+For the  more difficult task of comparing unaligned tiles, demographics are more significant.
+Higher levels of education and higher levels of income are associated with a significant increase of panelists choosing a response of 'they are the same', resulting in significant decreases of both correct answers and wrong answers."
 #| fig-subcap: 
@@
 coefs_ab_correct_same <- broom::tidy(logit_demo_task_ab_correct_same)
-```
-
-```{=tex}
-\hh{@tbl-demographics doesn't survive the formatting to pdf very well - need to shorten the row names or put into multiple lines} \hh{table rotates in pdf}
 ```

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
closing out some hh/kr comment threads that are complete

@@
 <!--However, the aspect ratio of the plotting area is adjusted to increase the length in pixels of the stacked bar elements, and  decrease the relative thickness \kr{wording?} of each bar. -->
-\kr{Do we have a measurement of the pixels for this? I think a sentence on that would close out this section nicely.}\hh{like that?}
+\kr{Do we have a measurement of the pixels for this?
+I think a sentence on that would close out this section nicely.}\hh{like that?}\kr{I was thinking the length pixels, since we talk about how many they are in Task 1, so we could talk about how many JND the difference is.
+I like the height thing though too, so maybe both?
+}
 
@@
 
-\hh{The Omnibus sample is not longitudinal in nature, i.e.\ we do not have any information about whether the same panelists were included in multiple rounds.
+The Omnibus sample is not longitudinal in nature, i.e.\ we do not have any information about whether the same panelists were included in multiple rounds.
 While there is a (small) chance that panelists are included in multiple rounds of our data, there was at least a one month gap between viewing each visual stimulus, and for the purposes of the analysis we will consider data from these viewings as independent.
-}
 
@@
 
-\kr{We need to be careful about not making it seem like the rounds are the same as the tasks, since we have 3 rounds and 3 tasks.
-We describe it, but I feel like this table kind of muddies the waters.
-Could we instead do a table of respondents by task?} \hh{does this comment still stand after the re-organization?
-...
-I would suggest to move it up to the survey methods and add the lambdas} \kr{I think moving it here helps, I did also add another table below.
-Let me know what you think.}\hh{I like it!}
-
 | Name    | Date       |      \# Participants |                effective sample size |         Sum of weights $\sum_i w_i$ | $\lambda_i$ |
@@
 The resulting sample sizes corresponding to each task are shown in \autoref{tbl-tasks}.
-\kr{Is this table useful?
-I want to be clear about the difference between tasks and rounds and make sure it's clear what our resulting sizes are..}\hh{yes!}
 

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
two typos

--- 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
@@
 
--   They are the same"
+-   They are the same
 
@@
 Because each participant was shown both the aligned and unaligned versions of the chart, we can use a paired $t$-test to compare mean accuracy between the two charts.
-The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: $<< 0.0001$).
+The resulting $t$-statistic is highly significant ($t$ statistic: `r round(paired$statistic,1)`, df: `r paired$parameter`, $p$-value: $\ll 0.0001$).
 

### 2023-07-20 SDSS-2023/testing-charts-accuracy.qmd
remove the lagging AND

@@
 short-title: Testing charts in surveys # running title
-author:
-  - first: Kiegan                 # first name
-    first-init: K.              # initial of first name
-    last: Rice                  # last name
-    marker: 1               # identifier for the affiliation
-    footnote-id: 1              # optional identifier for the footnote
-    email: foo@bar.com          # optional email address
-  - first: Heike
-    first-init: H.
-    last: Hofmann
-    marker: 2
-  - first: Nola
-    first-init: N.
-    last: duToit
-    marker: 1
-  - first: Edward
-    first-init: E.
-    last: Mulrow
-    marker: 1
-footnote:
-  - content: Corresponding author
-    type: corresp               # set type for the corresponding author
-    id: 1
-affiliation:
-  - institution: NORC at the University of Chicago
-    country: United States
-    marker: 1
-  - institution: Iowa State University
-    department: Statistics
-    country: United States
-    marker: 2
 bibliography: references.bib
@@
 ```{=latex}
-  \newcommand{\hh}[1]{{\textcolor{orange}{#1}}} % take out 
-  \newcommand{\kr}[1]{{\textcolor{teal}{#1}}}
+  \newcommand{\hh}[1]{{\textcolor{orange}{#1}}} 
+  \newcommand{\kr}[1]{{\textcolor{teal}{#1}}}.
+
+% take out everything above later
+% keep everything below
 
@@
                       fig.env="figure*", 
+                      fig.pos = "hbt", # preference of figure placement: here, bottom, top 
                  #     fig.align = "center",
@@
 #| fig-cap: "Demographics of respondents"
-#| fig-height: 7
-#| fig-width: 10
+#| fig-height: 6
+#| fig-width: 12
 
@@
 
-\hh{XXX the formatting isn't right yet here ... latex problem with large floats ... will fix it.}
+<!--\hh{XXX the formatting isn't right yet here ... latex problem with large floats ... will fix it.}-->
 
@@
 #| fig-height: 4
-#| fig-width: 7
+#| fig-width: 10
 #| layout-nrow: 2

### 2023-07-20 analysis/analysis.qmd
floating bars

@@
   ids = ~CaseId, weights = ~weight)
+```
+
+```{r data-rnd5}
+# other github repo. 
+round5 <- readRDS(file.path(vs_repo_local, "round5.rds"))
+
+# floating bars (unaligned) are shown to all participants of round 5
+round5_adj <- round5 %>% mutate(
+  ab = ab1,
+  ab_zoom = ab1_zoom,
+  ab_certain = ab1_certain,
+  weight = WEIGHT_NOLA
+) %>% filter(
+  ab != "SKIPPED ON WEB"
+)
 ```
@@
 
+## Model 4: Unaligned Elements in horizontal wide barcharts compared to floating bars
+
+```{r}
+#| label: fig-context-bars
+#| fig-cap: Floating bars show and horizontal wide barcharts show the same areas  with the same distances
+#| fig-subcap: 
+#|   - floating bars
+#|   - horizontal wide
+#| layout-nrow: 1
+#| out-width: 100%
+knitr::include_graphics("images/floating - all.PNG")
+knitr::include_graphics("images/round3/hwide-ab.PNG")
+```
+
+```{r combine-r35}
+# wide horizontal
+wide_horiz <- round3_adj %>% filter(design=="Horizontal wide")
+
+# floating bars are in round5
+lambda <- calculate_lambda(wide_horiz$weight, round5_adj$weight)
+
+round_35 <- rbind(
+  wide_horiz %>% mutate(
+    weight = lambda*weight/sum(weight),
+    task = "horizontal wide"
+  ) %>% dplyr::select(CaseId, weight, ab, ab_zoom, ab_certain, task),
+  round5_adj %>% mutate(
+    weight = (1-lambda)*weight/sum(weight),
+    task = "floating"
+  ) %>% dplyr::select(CaseId, weight, ab, ab_zoom, ab_certain, task)
+) %>% mutate(
+  correct = ab=="B is bigger",
+  not_wrong = ab != "A is bigger",
+  task = factor(task)
+)
+
+survey_35 <- svydesign(ids=~CaseId, weights=~weight, data = round_35)
+```
+
+```{r model-r35}
+logit_35_correct <- svyglm(correct~task-1, design = survey_35, family=quasibinomial())
+logit_35_not_wrong <- svyglm(not_wrong~task-1, design = survey_35, family=quasibinomial())
+
+coefs_35_correct <- broom::tidy(logit_35_correct) %>% 
+  mutate(task = factor(gsub("task", "", term), levels=levels(round_35$task)))
+coefs_35_notwrong <- broom::tidy(logit_35_not_wrong) %>% 
+  mutate(task = factor(gsub("task", "", term), levels=levels(round_35$task)))
+```
+
+```{r}
+#| label: fig-round-floats
+#| fig-cap: "Accuracy of responses by design.
+Free floating bars are comparisons of lengths.
+The additional context of stacked barcharts anchors the bars for a significant  increase in correct responses.
+However, the percentage of wrong responses also significantly increases.
+"
+#| out-width: 70%
+
+round_35 %>% 
+  mutate(
+    response3 = c("Wrong", "They are the same", "Correct")[correct+not_wrong+1],
+    response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong")))
+  ) %>% 
+  ggplot(aes(x = task)) + 
+  geom_bar(aes(fill = response3, weight = weight), position="fill") +
+#  facet_grid(.~task) + 
+  theme(legend.position="bottom") + 
+  scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
+  xlab("") +
+  scale_y_continuous("Percent responses", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25)) + 
+  geom_rect(
+    aes(
+      xmin = as.numeric(task) - 0.45,
+      ymin = inv_logit(estimate - 1.96* std.error),
+      xmax = as.numeric(task) + 0.45,
+      ymax = inv_logit(estimate + 1.96* std.error),
+    ), 
+    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+    data = coefs_35_correct
+  ) +
+  geom_segment(
+    aes(
+      x = as.numeric(task) - 0.45,
+      y = inv_logit(estimate ),
+      xend = as.numeric(task) + 0.45,
+      yend = inv_logit(estimate )
+    ), 
+    alpha = 1, colour = "black", linewidth = 0.25,
+    data = coefs_35_correct
+  ) +
+  geom_rect(
+    aes(
+      xmin = as.numeric(task) - 0.45,
+      ymin = inv_logit(estimate - 1.96* std.error),
+      xmax = as.numeric(task) + 0.45,
+      ymax = inv_logit(estimate + 1.96* std.error),
+    ), 
+    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+    data = coefs_35_notwrong
+  ) +
+  geom_segment(
+    aes(
+      x = as.numeric(task) - 0.45,
+      y = inv_logit(estimate ),
+      xend = as.numeric(task) + 0.45,
+      yend = inv_logit(estimate )
+    ), 
+    alpha = 1, colour = "black", linewidth = 0.25,
+    data = coefs_35_notwrong
+  )
+#+ 
+#  geom_text(aes(label=letters), y = -0.05, data = ci_confint, colour = "steelblue") +
+#  geom_text(aes(label=letters), y = 1.05, data = ci_confint3, colour = "darkorange")
+
+```
+
 ## Model X: Comparing Aligned and Unaligned Elements Facets {#sec-modelX}
@@
 
+How do our findings relate to @cejaTruthSquareAspect2021 ?
+
+Science communication review of visuals: @desnoyersTaxonomyVisualsScience2023
+
 ## References{-}

### 2023-07-21 analysis/analysis.qmd
comparison to floating bars
floating bars

@@
     data = coefs_35_notwrong
-  )
+  ) +
+  ggtitle("Comparison of unaligned tiles")
 #+ 
@@
 #  geom_text(aes(label=letters), y = 1.05, data = ci_confint3, colour = "darkorange")
+
+```
+
+```{r m35-summary}
 

### 2023-07-21 SDSS-2023/testing-charts-accuracy.qmd
analysis

@@
 #| fig-cap: "Demographics of respondents"
-#| fig-height: 6
+#| fig-height: 7
 #| fig-width: 12
@@
 The overlaid rectangles represent 95% confidence intervals.
-The letters in blue and orange encode significances between pairwise proportions: two bars have a significantly different proportion (at a 5% significance level) if they do not share a letter.
+The letters in blue and orange encode significances between pairwise proportions [@piephoAlgorithmLetterBasedRepresentation2004]: two bars have a significantly different proportion (at a 5% significance level) if they do not share a letter.
 There is no significant difference between the three designs for wrong responses.
@@
 
+### How certain are participants?
+
+<!--
 \kr{Do we have anything about total time spent on tasks across the 3 structures? We have time on each question for rounds 2-3. And can we pull in some of the certainty stuff?}
@@
 \hh{we could include certainty - I'll look into it}
+-->
+
+```{r certainty-data}
+round_123_certainty_long <- round_123 %>% 
+  pivot_longer(ab_certain:cd_certain, names_to = "Question", values_to = "Response") %>%
+  filter(Response != "SKIPPED ON WEB") %>%  # 20 participants
+  mutate(
+    Correct = ifelse(Question == "ab_certain", CorrectAB, CorrectCD),
+    Question = ifelse(Question=="ab_certain", "Unaligned", "Aligned"),
+# ordering of levels in Response is from extremely=6 to not at all=2 - not very intuitive
+    Response = factor(Response, levels = c("not at all", "slightly", "moderately", "very", "extremely")) # gets rid of the 'SKIPPED ON WEB' and makes the ordering go from 1 to 5
+  )
+
+round_123_certainty_long <- round_123_certainty_long %>% mutate(
+  correct_question_chart = factor(paste(Correct, Question, chart, sep="-"))
+)
+survey_123_crt <- 
+  survey::svydesign(
+  data = round_123_certainty_long, 
+  ids = ~CaseId, weights = ~weight)
+```
+
+Using linear scores for the response of `Certainty`, with `not certain at all` assigned a score of 1 and `extremely certain` assigned a score of 5, we can estimate the effects of task, chart design, and correctness on certainty  by using  a cell-means model of the form:
+$$
+Y_{k} = \mu_{ij\ell(k)} + \epsilon_{k},
+$$
+where $k = 1, ..., N$, $\mu_{ij\ell(k)}$ is average certainty (measured on a scale from 1 to 5) of the four combinations of task and correctness by each of the three designs, where $i = 1, 2$ encodes unaligned/aligned, and $j=1,2$ encodes wrong, correct, and $\ell = 1, 2, 3$ encodes horizontal, wide horizontal, and vertical stacked bar charts, respectively.
+We also assume that errors are normally distributed, i.e.
+$\epsilon_k \stackrel{i.i.d}{\sim} N(0, \sigma^2)$ for all $k = 1, ..., N$.
+The results are shown in @fig-certainty. 
+What we find, is that the highest scores for certainty are associated with comparing aligned tiles.
+Certainty scores are not significantly different between correct and wrong responses.
+Scores are (mostly) significantly lower for the unaligned task.
+Interestingly, the lowest scores of certainty are associated with correct responses, here.
+
+```{r model-certainty}
+model_certain <- svyglm(as.numeric(Response)~ correct_question_chart-1, design = survey_123_crt)
+
+coefs_certainty <- broom::tidy(model_certain)
+coefs_certainty <- coefs_certainty %>% 
+  mutate(
+    term = gsub("correct_question_chart","", term)
+  ) %>%
+  separate(term, into=c("Correct", "Question", "chart"), sep="-", remove = FALSE)
+
+coefs_certainty <- coefs_certainty %>% 
+  mutate(
+    Question = factor(Question, 
+                      levels = c("Aligned", "Unaligned")),
+    Correct = factor(Correct, levels=c("FALSE", "TRUE"), 
+                     labels = c("Wrong", "Correct")),
+    chart = factor(chart, levels=c("Horizontal wide", "Horizontal", "Vertical"))
+  )
+
+round_123_certainty_long <- round_123_certainty_long %>%
+  mutate(
+    Question = factor(Question, 
+                      levels =c("Aligned", "Unaligned")),
+    Correct = factor(Correct, levels=c("FALSE", "TRUE"), 
+                     labels = c("Wrong", "Correct")),
+    chart = factor(chart, levels=c("Horizontal wide", "Horizontal", "Vertical"))
+  )
+
+```
+
+```{r model-letters, warning=FALSE}
+# create all pairwise comparisons for tasks and questions.
+cross_compare_certainty <- glht(model_certain, mcp(correct_question_chart="Tukey"))
+
+foo <- cld(cross_compare_certainty)
+
+# get letters for each confidence 
+coefs_certainty$letters <- cld(cross_compare_certainty)$mcletters$Letters
+```
+
+```{r}
+#| label: fig-certainty
+#| out-width: 100%
+#| fig-cap: "Certainty by task, chart design, and correctness.
+Correct answers in the unaligned task have the lowest certainty scores associated with them.
+Incorrect responses on the unaligned task show a significant boost in certainty in the vertical stacked bar.
+For horizontal stacked bars the difference between aligned and unaligned charts matters significantly.
+Errorbars around the points show estimated scores (on a scale of 1 to 5) with corresponding 95% CI.
+The letters underneath the bars indicate significances.
+The scores for two bars are significantly different if they do not share a letter.
+"
+#| fig-width: 10
+#| fig-height: 4.5
+
+round_123_certainty_long %>%
+  ggplot(aes(x = chart)) +
+  facet_grid(.~Question+Correct) +
+  geom_bar(aes(fill=Response, weight=weight), 
+           position=position_fill(reverse=TRUE), 
+           colour="grey40", linewidth=0.25) +
+  theme(
+    axis.text.x = element_text(angle=30, hjust=1)
+  ) + 
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), 
+                     labels = seq(0,1, by=0.25)*100, limits = c(-0.1, 1)) +
+  xlab("") + 
+  scale_fill_brewer(palette = "PRGn", guide = guide_legend(reverse = TRUE)) + 
+  geom_point(
+    aes(y = estimate/5),
+    size = 3,
+    data = coefs_certainty
+  ) + 
+  geom_errorbar(
+    aes(ymin = (estimate-1.96*std.error)/5,
+        ymax = (estimate+1.96*std.error)/5),
+    width=0.3,
+    data = coefs_certainty
+  ) + 
+  geom_text(aes(label=letters), y = -0.05, data = coefs_certainty) 
+```
 

### 2023-07-21 SDSS-2023/testing-charts-accuracy.qmd
certainty of participants

@@
 
+scale_x <- function(value, from, to) {
+  (value-from[1])/diff(from)*diff(to)
+}
+
+coefs_certainty <- coefs_certainty %>% 
+  mutate(
+    estimate = scale_x(estimate, from=c(1,5), to=c(0,1)),
+    std.error = std.error/4
+  )
+
 round_123_certainty_long <- round_123_certainty_long %>%
@@
   scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), 
-                     labels = seq(0,1, by=0.25)*100, limits = c(-0.1, 1)) +
+                     labels = seq(0,1, by=0.25)*100, limits = c(-0.1, 1),
+                     sec.axis = sec_axis( trans=~(.*4)+1, name="Certainty scores")) +
   xlab("") + 
@@
   geom_point(
-    aes(y = estimate/5),
+    aes(y = estimate),
     size = 3,
@@
   geom_errorbar(
-    aes(ymin = (estimate-1.96*std.error)/5,
-        ymax = (estimate+1.96*std.error)/5),
+    aes(ymin = (estimate-1.96*std.error),
+        ymax = (estimate+1.96*std.error)),
     width=0.3,

### 2023-07-21 SDSS-2023/testing-charts-accuracy.qmd
Odds, not log odds

@@
 The overlaid rectangles represent 95% confidence intervals.
-The letters in blue and orange encode significances between pairwise proportions [@piephoAlgorithmLetterBasedRepresentation2004]: two bars have a significantly different proportion (at a 5% significance level) if they do not share a letter.
+The letters in blue and orange encode significances between pairwise proportions: two bars have a significantly different proportion (at a 5% significance level) if they do not share a letter.
 There is no significant difference between the three designs for wrong responses.
@@
 The estimated values and 95% confidence intervals are shown in @tbl-rounds-123.
+\hh{The lowercase letters indicate significances between pairs of estimates: two estimates (in the same column) are significantly different at 5% level, if they do not have any letter in common [@piephoAlgorithmLetterBasedRepresentation2004].
 
@@
   gt() %>%
-  tab_header(title = "Log odds of accuracy by task and chart type") %>%
+  tab_header(title = "Odds of accuracy by task and chart type") %>%
   fmt_number(
@@
 ```{r zoom-123}
-logit_zoom <- svyglm(I(zoom=="yes")~response3+task+chart, family=quasibinomial(), design=survey_123)
+logit_zoom <- 
+  svyglm(I(zoom=="yes")~response3+task+chart, family=quasibinomial(), design=survey_123)
 
@@
   xlab("") + 
-  ggtitle("Zooming behavior by task")
+  ggtitle("Zooming behavior\nby task") +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), 
+                     labels = seq(0,1, by=0.25)*100)
 
@@
   xlab("") + 
-  ggtitle("Zooming behavior by chart") +
-  theme(axis.text.x = element_text(angle=30, hjust = 1))
+  ggtitle("\nby chart") +
+  theme(axis.text.x = element_text(angle=30, hjust = 1)) +
+  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), 
+                     labels = seq(0,1, by=0.25)*100)
 
@@
 
+#    term = TeX(c("$\\mu$", "They are the same $\\rho_2$", "wrong $\\rho_3$", "Aligned $\\tau_2$", "Horizontal wide $\\delta_2$", "Vertical $\\delta_3$"))
+
+coefs_zoom$term <- c("$$\\hat{\\mu}$$", "$$\\text{They are the same } \\widehat{\\rho}_{2}$$", 
+                     "Wrong $$\\widehat{\\rho}_{3}$$", "Aligned $$\\widehat{\\tau}_{2}$$", 
+                     "Horizontal wide $$\\widehat{\\delta}_{2}$$", "Vertical $$\\widehat{\\delta}_{3}$$")
+
 coefs_zoom %>%
@@
     term = ifelse(term=="(Intercept)", "Intercept", term),
-    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value))
-  ) %>%
+    p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value)),
+    variable = c("", "Response", "Response", "Task", "Chart design", "Chart design")
+  ) %>%
+  group_by(variable) %>%
   gt() %>%
   tab_header("Estimates for logistic regression on zooming behavior") %>%
+  tab_options(
+    column_labels.font.weight = "bold",
+    row_group.font.weight = "bold"
+  ) %>%
   cols_label(
@@
  gt() %>%
-  tab_header(title = "Log odds of accuracy by task and demographics of respondents") %>%
+  tab_header(title = "Odds of accuracy by task and demographics of respondents") %>%
   tab_footnote("Signif. codes:  0 `***' 0.001 `**' 0.01 `*' 0.05 `.' 0.1 ` ' 1") %>%

### 2023-07-21 SDSS-2023/testing-charts-accuracy.qmd
massaging tables

@@
   cols_merge(
+    columns = contains("ci") & contains("same or wrong"),
+    pattern = "<< [{1}, {2}]>>"
+  ) %>%
+  tab_spanner(
+    label = "correct | same or wrong",
     columns =  c("estimate_correct | same or wrong", "ci_low_correct | same or wrong",
-                "ci_high_correct | same or wrong", "letters_correct | same or wrong"),
-    pattern = "<<{1}>><< [{2}, {3}]>><< {4}>>"
+                 "letters_correct | same or wrong")
   ) %>%
   cols_merge(
+    columns = contains("ci") & contains("correct or same"),
+    pattern = "<< [{1}, {2}]>>"
+  ) %>%
+  tab_spanner(
+    label = "correct or same | wrong",
     columns =  c("estimate_correct or same | wrong", "ci_low_correct or same | wrong",
-                "ci_high_correct or same | wrong", "letters_correct or same | wrong"),
-    pattern = "<<{1}>><< [{2}, {3}]>><< {4}>>"
+                "letters_correct or same | wrong")
   ) %>%
@@
       "estimate_correct | same or wrong" = "correct | same or wrong",
-      "estimate_correct or same | wrong" = "correct or same | wrong"
+      "estimate_correct or same | wrong" = "correct or same | wrong",
+      "estimate_correct or same | wrong" = "Est.",
+      "ci_low_correct or same | wrong" = "          [95% CI]",
+      "letters_correct or same | wrong" = " ",
+      "estimate_correct | same or wrong" = "Est.",
+      "ci_low_correct | same or wrong" = "          [95% CI]",
+      "letters_correct | same or wrong" = " "
     )
@@
     .list = c(
+      "term" = " ",
       "estimate_Aligned_correct" = "Est.",

### 2023-07-21 SDSS-2023/testing-charts-accuracy.qmd
size of stimulus

@@
 The images utilized in the third task again represent the same data as in the first two tasks, and the overall image has dimensions which are identical to the images in the first two tasks, displaying at the same size during the survey.
-\hh{The length of bars is increased to fit the new dimensions. The height of the tiles are adjusted accordingly (from 50 to 30 pixels) to keep the overall area the same as before}
+\hh{The length of bars is increased to fit the new dimensions.
+This increases the difference in the length of bars from previously 8 pixels to 13 pixels.
+The widths of the tiles are adjusted accordingly (from 50 to 30 pixels) to keep the overall area of the tiles approximately constant.
+}
 <!--However, the aspect ratio of the plotting area is adjusted to increase the length in pixels of the stacked bar elements, and  decrease the relative thickness \kr{wording?} of each bar. -->
-\kr{Do we have a measurement of the pixels for this?
+<!--kr{Do we have a measurement of the pixels for this?
 I think a sentence on that would close out this section nicely.}\hh{like that?}\kr{I was thinking the length pixels, since we talk about how many they are in Task 1, so we could talk about how many JND the difference is.
 I like the height thing though too, so maybe both?
-}
+}-->
 

### 2023-07-21 SDSS-2023/testing-charts-accuracy.qmd
combination of multiple surveys

@@
 
-We *combine* (rather than cumulate) two surveys $S_1$ and $S_2$, as described in @omuircheartaighCombiningSamplesVs2002, by multiplying weights in $S_1$ and $S_2$ by $\lambda$ and $1 - \lambda$ respectively.
-$\lambda \in [0,1]$ is given as
+We *combine* (rather than cumulate) $\ell \ge 2$ surveys $S_1$, $S_2$, ...
+$S_\ell$ as described in @omuircheartaighCombiningSamplesVs2002, by multiplying weights in $S_i$ by $\lambda_i$, for $1 \le i \le \ell$.
+$\lambda_i \in [0,1]$ with $\sum_i \lambda_i = 1$ is given as
 $$
-\lambda = \frac{n_1/d_1}{n_1/d_1 + n_2/d_2},
-$$ where $n_1$ and $n_2$ are the nominal sample sizes and $d_1$ and $d_2$ are the design effects for the estimators.
-Here, $d_1$ and $d_2$ are estimated as $$
-d_1 = 1 + CV(w_i \in S_1)^2 \ \ \ \text{ and } \ \ \ d_2 = 1 + CV(w_i \in S_2)^2,
-$$ where $CV$ is the coefficient of variation of the weights within each sample, and is estimated as in @kishSurveySampling1965 :
+\lambda_i = \frac{n_i/d_i}{\sum_{j=1}^{\ell}n_j/d_j},
+$$ where $n_i$ is the nominal sample of survey $S_i$ and $d_i$ are the design effects for the estimators.
+Here, $d_i$ are estimated as 
+$$
+d_i = 1 + CV(w \in S_i)^2
+$$ 
+where $CV$ is the coefficient of variation of the weights $w$ within each sample, and is estimated as in @kishSurveySampling1965 :
 

### 2023-07-21 analysis/analysis.qmd
model for floating bars vs wide bars

@@
 
+coefs_zoom$term <- c("$$\\hat{\\mu}$$", "They are the same $$\\widehat{\\rho}_{2}$$", 
+                     "Wrong $$\\widehat{\\rho}_{3}$$", "Aligned $$\\widehat{\\tau}_{2}$$", 
+                     "Horizontal wide $$\\widehat{\\delta}_{2}$$", "Vertical $$\\widehat{\\delta}_{3}$$")
+
 coefs_zoom %>%
   mutate(
-    term = ifelse(term=="(Intercept)", "Intercept", term),
+#    term = ifelse(term=="(Intercept)", "Intercept", term),
     p.value = ifelse(p.value<0.0001, "< 0.0001", sprintf("%.4f", p.value))
@@
 ```{r m35-summary}
-
+#| label: tbl-floating
+#| tbl-cap: "Percentage of accuracy of responses by design."
+coefs_35 <- coefs_35_notwrong %>% mutate(
+  level = c("correct or same | wrong")
+) %>% rbind(
+coefs_35_correct %>% mutate(
+  level = c("correct | same or wrong")
+))
+
+coefs_35 %>% 
+  mutate(
+    ci_low = estimate - 1.96*std.error,
+    ci_high = estimate + 1.96*std.error,
+    estimate = inv_logit(estimate),
+    ci_low = inv_logit(ci_low),
+    ci_high = inv_logit(ci_high),
+    perc_ci = sprintf(" %.2f [%.2f, %.2f]", estimate, ci_low, ci_high)
+    # perc_ci = sprintf(" %.2f [%.2f, %.2f] %s", estimate, ci_low, ci_high,
+    #                    ifelse(p.value < 0.001, "***",
+    #                      ifelse(p.value < 0.01, "**",
+    #                        ifelse(p.value < 0.05, "*",
+    #                          ifelse(p.value < 0.1, ".", " ")))))
+  ) %>% 
+  dplyr::select(level, task, perc_ci) %>%
+  pivot_wider(names_from="task", values_from="perc_ci") %>%
+  gt() %>%
+  tab_header("Estimated percentage of responses [95% CI]") %>%
+  cols_label(
+    "level" = " "
+  )
 ```

### 2023-07-21 SDSS-2023/testing-charts-accuracy.qmd
including floating bars

@@
 
+round5 <- readRDS(file.path(vs_repo_local, "round5.rds"))
+
+# floating bars (unaligned) are shown to all participants of round 5
+round5_adj <- round5 %>% mutate(
+  ab = ab1,
+  ab_zoom = ab1_zoom,
+  ab_certain = ab1_certain,
+  weight = WEIGHT_NOLA
+) %>% filter(
+  ab != "SKIPPED ON WEB"
+)
+
 round6 <- readRDS(file.path(vs_repo_local, "round6.rds"))
@@
   weight = WEIGHT_NOLA
-)
-```
-
-```{r data-rnd3}
-round3 <- readRDS(file.path(vs_repo_local, "round3.rds"))
-
-# filter out all of the 'skipped on web':
-round3_adj <- round3 %>% mutate(
-  weight = WEIGHT_NOLA
-) %>% filter(
-  ab != "SKIPPED ON WEB",
-  cd != "SKIPPED ON WEB"
 )
@@
 
+lambda_1235 <- calculate_lambda(round_123$weight, round5_adj$weight)
+round_123b <- round_123 %>% 
+              mutate(weight = lambda_1235*weight/sum(weight)) %>% 
+  rbind(
+    round5_adj %>% 
+      mutate(
+        weight = (1-lambda_1235)*weight/sum(weight),
+        round = "Round 5",
+        cd = NA,
+        cd_certain=NA,
+        cd_zoom=NA,
+        CorrectAB = ab == "B is bigger",
+        CorrectCD = NA,
+        CorrectABCD = NA,
+        chart = "Floating bars"
+       ) %>% 
+        dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
+                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, chart, weight,
+                   age4, gender, income4, racethnicity, educ5) 
+  )
+
 levels(round_123$educ5) <- c("Less than HS", "HS or equivalent", "Some college", "Bachelor", "Post graduate")
@@
   ids = ~CaseId, weights = ~weight)
+
+round_123b_long <- round_123b %>% 
+  pivot_longer(ab:cd, names_to = "task", values_to = "Response") %>%
+  mutate(
+    correct = Response %in% c("B is bigger", "D is bigger"),
+    correct3 = correct | (Response == "They are the same"),
+    task_chart = factor(paste(task, chart, sep="-")),
+    zoom = ifelse(task=="ab", ab_zoom, cd_zoom),
+    response3 = ifelse(correct, "Correct", ifelse(correct3, "They are the same", "Wrong"))
+  ) %>% filter(
+    !is.na(Response)
+  )
+
+survey_123b <- 
+  survey::svydesign(
+  data = round_123b_long, 
+  ids = ~CaseId, weights = ~weight)
+
 ```
@@
 
-logit_123_correct <- svyglm(correct~task_chart-1, family=quasibinomial(), design=survey_123)
-logit_123_correct3 <- svyglm(correct3~task_chart-1, family=quasibinomial(), design=survey_123)
-
-ci_confint <- svyby(~correct, by=~task_chart, design=survey_123, svyciprop)
-ci_confint3 <- svyby(~correct3, by=~task_chart, design=survey_123, svyciprop)
+logit_123_correct <- svyglm(correct~task_chart-1, family=quasibinomial(), design=survey_123b)
+logit_123_correct3 <- svyglm(correct3~task_chart-1, family=quasibinomial(), design=survey_123b)
+
+ci_confint <- svyby(~correct, by=~task_chart, design=survey_123b, svyciprop)
+ci_confint3 <- svyby(~correct3, by=~task_chart, design=survey_123b, svyciprop)
 
@@
   task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
-  chart = factor(chart, levels =c("Vertical", "Horizontal", "Horizontal wide"))
+  chart = factor(chart, levels =c("Vertical", "Horizontal", "Horizontal wide", "Floating bars"))
 )
@@
   task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
-  chart = factor(chart, levels =c("Vertical", "Horizontal", "Horizontal wide"))
+  chart = factor(chart, levels =c("Vertical", "Horizontal", "Horizontal wide", "Floating bars"))
 ) 
@@
 
-round_123_long %>% 
+round_123b_long %>% 
   mutate(
@@
     task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
-    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide", "Floating bars"))
   ) %>% 
@@
   geom_bar(aes(fill = response3, weight = weight), position="fill") +
-  facet_grid(.~task) + 
+  facet_grid(.~task, scales = "free_x", space="free_x") + 
   theme(legend.position="bottom") + 
@@
     column_labels.font.weight = "bold",
-    row_group.font.weight = "bold"
+    row_group.font.weight = "bold",
+    footnotes.font.size = 8
   ) %>%
@@
     )
-  ) 
+  ) %>% tab_footnote(
+    footnote = "We are including here, responses on an A vs B comparison for floating bars, corresponding to a length assessment, see bottom right of Figure 1.",
+    locations = cells_body(
+      rows = 1,
+      columns = "chart"
+    )
+  )
+
 ```

### 2023-07-23 SDSS-2023/testing-charts-accuracy.qmd
census estimates added to demographics

@@
 
-```{r}
-#| label: fig-tasks
+<!--
+```{r eval=FALSE}
+#| label: fig-tasks-old
 #| fig-cap: "Each of the visual stimuli presented to viewers. In each bar chart, two pieces are marked. The larger piece in each chart are pieces B and D. "
 #| fig-subcap:
-#| - "Vertical stacked bar, unaligned"
-#| - "Horizontal stacked bar, unaligned"
-#| - "Wide horizontal stacked bar, unaligned"
-#| - "Vertical stacked bar, aligned"
-#| - "Horizontal stacked bar, aligned"
-#| - "Wide horizontal stacked bar, aligned"
-#| layout-nrow: 2
+#| - "Vertical, unaligned"
+#| - "Horizontal, unaligned"
+#| - "Wide horizontal, unaligned"
+#| - "Vertical, aligned"
+#| - "Horizontal, aligned"
+#| - "Wide horizontal, aligned"
+#| layout-nrow: 3
 #| out-width: 90%
+cat("Task 1")
+cat("Task 2")
+cat("Task 3")
 knitr::include_graphics("images/vertical-ab.PNG")
@@
 ```
+-->
+
+```{r}
+#| label: fig-tasks
+#| fig-cap: "Each of the visual stimuli presented to viewers. In each bar chart, two pieces are marked. The larger piece in each chart are pieces B and D. "
+#| out-width: 90%
+knitr::include_graphics("images/stimuli-overview.png")
+```
 
@@
 
-Task 1 was presented to viewers in one of two rounds (Rounds 1-2), with four color scheme variations each presented to 50% of respondents in each round. 
-There were no significant differences in respondent accuracy or behavior across color schemes, and we combine them here into one set of stimulus responses to compare structural differences.
+Task 1 was presented to viewers in two rounds (Rounds 1-2), with four color scheme variations each presented to 50% of respondents in each round. 
+There were no significant differences in respondent accuracy or behavior across color schemes, and we combine them here into one set of stimulus responses to focus on the comparison of structural differences.
 \kr{Should we put the color variations and a test for significance in accuracy/a visual in the appendix/supplementary materials?
@@
 
-The visual mappings in the second task, shown in \autoref{fig-tasks}(b, e), are identical to the first task, but the axes of the chart are rotated so that the stacked bar is represented in a horizontal format.
+The visual mappings in the second task, shown in \autoref{fig-tasks}(b, e), are identical to the first task, but the axes of the chart are rotated so that the stacked bars are represented in a horizontal format.
 This represents a structural change in how the data are presented to the viewer while preserving the pixel size of the elements viewers must compare.
@@
 households.
-The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households @dennis2019technical.
+The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households [@dennis2019technical].
 \hh{I don't think I've got the right citation} Each test was conducted using the AmeriSpeak Omnibus survey, which runs biweekly and samples around 1,000 U.S.
@@
 
-When combining responses from different rounds of the Omnibus survey, we calibrate weights in each round to the total population  across all rounds in the resulting model.
-
-\kr{Do we need to update this description to just say how we did it for combining across all the rounds represented?}
-
-We *combine* (rather than cumulate) $\ell \ge 2$ surveys $S_1$, $S_2$, ...
-$S_\ell$ as described in @omuircheartaighCombiningSamplesVs2002, by multiplying weights in $S_i$ by $\lambda_i$, for $1 \le i \le \ell$.
+When combining responses from different surveys, weights need to be calibrated to the total population, so that their order after combining still reflects their relative importance.
+
+<!--
+\kr{Do we need to update this description to just say how we did it for combining across all the rounds represented?}\hh{done} -->
+
+We *combine* (rather than cumulate) a set of $\ell$ surveys $S_1$, $S_2$, ...
+$S_\ell$ (with $ \ell \ge 2$ ) as described in @omuircheartaighCombiningSamplesVs2002, by multiplying weights in $S_i$ by $\lambda_i$, for $1 \le i \le \ell$.
 $\lambda_i \in [0,1]$ with $\sum_i \lambda_i = 1$ is given as
@@
 where $CV$ is the coefficient of variation of the weights $w$ within each sample, and is estimated as in @kishSurveySampling1965 :
-
 $$ 
@@
 
-The data for this paper were collected in several rounds as part of the NORC Omnibus, and the resulting number of participants, effective sample sizes, and $\lambda$ values are shown in \autoref{tbl-rounds}.
+The data for this paper were collected in several rounds as part of the NORC Omnibus.
+The resulting number of participants, effective sample sizes, and $\lambda$ values are shown in \autoref{tbl-rounds}.
 
@@
 | Round 2 | May 2022   | `r nrow(round2_adj)` | `r round(neff(round2_adj$weight),1)` | `r round(sum(round2_adj$weight),1)` | $\lambda_2$  = `r sprintf("%.3f",lambdas[2])` |
-| Round 3 | Jun 2022   | `r nrow(round3_adj)` | `r round(neff(round3_adj$weight),1)` | `r round(sum(round3_adj$weight),1)` | $\lambda_3$  = `r round(lambdas[3],3)` |
+| Round 3 | June 2022   | `r nrow(round3_adj)` | `r round(neff(round3_adj$weight),1)` | `r round(sum(round3_adj$weight),1)` | $\lambda_3$  = `r round(lambdas[3],3)` |
 | **Total**        |    ---        |    `r nrow(round_123)`   |  `r round(neff(round_123$weight),1)`   |    |   |
@@
 #| label: fig-rounds-demographics
-#| fig-cap: "Demographics of respondents"
+#| fig-cap: "Demographics of respondents in each of the tasks. The dots show percentages of the national demographics as estimated by the American Community Survery (ACS) 2021."
 #| fig-height: 7
@@
 
+# from ACS 2001
+# https://data.census.gov/table?q=sex+of+adults+age+greater+than+18+across+the+US&tid=ACSDP1Y2021.DP05	
+acs_gender <- data.frame(gender = c("Male", "Female"), 
+                         estimate = c(126690177, 131728290), moe=c(22354, 18838),
+                         percent = c(49, 51), moe_percent = 0.1)
+
 gender <- round_123 %>% 
-  ggplot(aes(x = gender, weight = weight_pct_round)) + geom_bar() + 
+  ggplot(aes(x = gender)) + 
+  geom_bar(aes(weight = weight_pct_round), fill="grey", colour = "grey30", linewidth=0.5) + 
   ylab("Percent") + 
@@
   xlab("") +
-  facet_grid(.~chart)
+  facet_grid(.~chart) + 
+  geom_point(aes(x = gender, y = percent), data = acs_gender)
+
+acs_age <- data.frame(age4 = c("18-29", "30-44", "45-59", "60+"), 
+                      estimate = c(52325456, 66712246, 61814869, 77565896))
+acs_age <- acs_age %>% mutate(percent = estimate/sum(estimate)*100)
 
 age <- round_123 %>% 
-  ggplot(aes(x = age4, weight = weight_pct_round)) + geom_bar() + 
+  ggplot(aes(x = age4)) + 
+  geom_bar(aes(weight = weight_pct_round), fill="grey", colour = "grey30", linewidth=0.5) + 
   ylab("Percent") + 
@@
   xlab("") + ylim(c(0,35)) +
-  facet_grid(.~chart)
+  facet_grid(.~chart) +
+  geom_point(aes(x = age4, y = percent), data = acs_age)
+
+acs_educ5 <- data.frame(educ5 = c("Less than HS", "HS or equivalent", "Some college",  "Bachelor", "Post graduate"), 
+                      estimate = c(28687047, 69610758, 77707408, 50031690, 29454503))
+acs_educ5 <- acs_educ5 %>% mutate(percent = estimate/sum(estimate)*100)
 
 education <- round_123 %>% 
-  ggplot(aes(x = educ5, weight = weight_pct_round)) + geom_bar() + 
+  ggplot(aes(x = educ5)) + 
+  geom_bar(aes(weight = weight_pct_round), fill="grey", colour = "grey30", linewidth=0.5) + 
   ylab("Percent") + 
@@
   xlab("") + ylim(c(0,35)) +
-  facet_grid(.~chart)
+  facet_grid(.~chart) +
+  geom_point(aes(y = percent), data = acs_educ5)
 

### 2023-07-23 SDSS-2023/testing-charts-accuracy.qmd
margin of error

@@
 
-Distributions of demographic characteristics across each survey round are shown in \autoref{fig-rounds-demographics}. 
-\kr{Split by rounds or by tasks?
+Distributions of demographic characteristics across each task are shown in \autoref{fig-rounds-demographics}, the dark points show percentages for US adults based on the American Community Survey 2021.
+The distribution of demographics are all quite similar across tasks and their 90% margin of errors (hashed lines) all cover the Census estimates.
+<!--\kr{Split by rounds or by tasks?
 I feel like since we later split by task, maybe we should show the demos by task?
 Or we can just talk about how Round 3 was split randomly along demos for Tasks 2/3?}\hh{changed the figure to show demos by task -- do we need confidence intervals?
--- or a comparison against census data?}
+-- or a comparison against census data?}-->
 
@@
 
+omni_gender <- data.frame(svyby(~gender, by=~chart, design=survey_123, svymean))
+omni_gender <- rbind(
+  omni_gender %>% dplyr::select(chart, ends_with("Male")) %>% mutate(sex="Male") %>%
+    rename(mean = "genderMale", SE = "se.genderMale") %>% 
+    dplyr::select(chart, mean, SE, sex),
+  omni_gender %>% dplyr::select(chart, ends_with("Female")) %>% mutate(sex="Female") %>%
+    rename(mean = "genderFemale", SE = "se.genderFemale")
+  ) %>%
+  mutate(
+    sex = factor(sex, levels=c("Male", "Female")),
+    mean = 100*mean,
+    SE = 100*SE
+  )
+omni_gender <- omni_gender %>% mutate(
+  low = mean - 1.6445*SE, # for 90% MoE 
+  high = mean + 1.6445*SE
+)
+omni_gender <- omni_gender %>%
+  dplyr::select(-SE) %>%
+  pivot_longer(cols =c(mean, low, high), names_to="group", values_to="value")
+
 gender <- round_123 %>% 
-  ggplot(aes(x = gender)) + 
-  geom_bar(aes(weight = weight_pct_round), fill="grey", colour = "grey30", linewidth=0.5) + 
+  ggplot() + 
+  geom_bar(aes(x = gender, weight = weight_pct_round), fill="grey", colour = "grey30", linewidth=0.5) + 
+  # geom_rect(
+  #   aes(
+  #     xmin = as.numeric(sex) - 0.45,
+  #     ymin = mean - 1.96* SE,
+  #     xmax = as.numeric(sex) + 0.45,
+  #     ymax = mean + 1.96* SE,
+  #   ), 
+  #   fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+  #   data = omni_gender
+  # ) +
+  geom_segment(
+    aes(
+      x = as.numeric(sex) - 0.45,
+      y = value,
+      xend = as.numeric(sex) + 0.45,
+      yend = value,
+      linetype=group
+    ), 
+    alpha = 1, colour = "black", linewidth = 0.25,
+    data = omni_gender
+  )  +
   ylab("Percent") + 
@@
   facet_grid(.~chart) + 
-  geom_point(aes(x = gender, y = percent), data = acs_gender)
+  geom_point(aes(x = gender, y = percent), data = acs_gender) +
+  theme(legend.position ="none") +
+  scale_linetype_manual(values=c(2,2,1))
 

### 2023-07-23 SDSS-2023/testing-charts-accuracy.qmd
5 year estimates

@@
 acs_gender <- data.frame(gender = c("Male", "Female"), 
-                         estimate = c(126690177, 131728290), moe=c(22354, 18838),
+                         estimate = c(125215171, 130276235), moe=c(4770, 4818),
                          percent = c(49, 51), moe_percent = 0.1)
@@
 acs_age <- data.frame(age4 = c("18-29", "30-44", "45-59", "60+"), 
-                      estimate = c(52325456, 66712246, 61814869, 77565896))
+                      estimate = c(22854328 + 30339089, 
+                                   22506614 + 21920290 + 20521593, 
+                                   20582187 + 21049271 + 21928936, 
+                                   73789098))
 acs_age <- acs_age %>% mutate(percent = estimate/sum(estimate)*100)
+
+omni_age <- data.frame(svyby(~age4, by=~chart, design=survey_123, svymean))
+omni_age_long <- omni_age %>% 
+  pivot_longer(starts_with("age"), names_to="age", values_to = "mean") %>% 
+  dplyr::select(chart, age, mean) 
+omni_age_long$SE <- omni_age %>% pivot_longer(starts_with("se."), names_to="age", values_to="SE") %>% pull(SE)
+omni_age <- omni_age_long %>% mutate(
+  age = gsub("age4", "", age),
+  age = gsub(".", "-", age, fixed=TRUE),
+  age = ifelse(age=="60-", "60+", age)
+)  
+omni_age <- omni_age %>% mutate(
+  age = factor(age, levels = c("18-29", "30-44", "45-59", "60+"))
+)
+
+omni_age <- omni_age %>% mutate(
+  low = mean - 1.6445*SE, # for 90% MoE 
+  high = mean + 1.6445*SE
+)
+omni_age <- omni_age %>%
+  dplyr::select(-SE) %>%
+  pivot_longer(cols =c(mean, low, high), names_to="group", values_to="value") %>%
+  mutate(
+    value = 100*value
+  )
 
@@
   theme(axis.text.x = element_text(angle=30, hjust = 1)) +
-  xlab("") + ylim(c(0,35)) +
+  xlab("") + ylim(c(0,40)) +
   facet_grid(.~chart) +
-  geom_point(aes(x = age4, y = percent), data = acs_age)
+  geom_point(aes(x = age4, y = percent), data = acs_age) +
+  geom_segment(
+    aes(
+      x = as.numeric(age) - 0.45,
+      y = value,
+      xend = as.numeric(age) + 0.45,
+      yend = value,
+      linetype=group
+    ), 
+    alpha = 1, colour = "black", linewidth = 0.25,
+    data = omni_age
+  )  +
+  scale_linetype_manual(values=c(2,2,1)) +
+  theme(legend.position = "none")
 

### 2023-07-23 SDSS-2023/testing-charts-accuracy.qmd
5 year ACS estimates

@@
 
-Distributions of demographic characteristics across each task are shown in \autoref{fig-rounds-demographics}, the dark points show percentages for US adults based on the American Community Survey 2021.
-The distribution of demographics are all quite similar across tasks and their 90% margin of errors (hashed lines) all cover the Census estimates.
+Distributions of demographic characteristics across each task are shown in \autoref{fig-rounds-demographics}, the dark points show percentages for US adults based on the 5 year estimates of the American Community Survey 2021.
+The distribution of demographics are all quite similar across tasks and their 90% margin of errors (hashed lines) mostly cover the Census estimates.
 <!--\kr{Split by rounds or by tasks?
@@
 
+omni_educ <- data.frame(svyby(~educ5, by=~chart, design=survey_123, svymean))
+omni_educ_long <- omni_educ %>% 
+  pivot_longer(starts_with("educ"), names_to="educ", values_to = "mean") %>% 
+  dplyr::select(chart, educ, mean) 
+omni_educ_long$SE <- omni_educ %>% pivot_longer(starts_with("se."), names_to="educ", values_to="SE") %>% pull(SE)
+omni_educ <- omni_educ_long %>% mutate(
+  educ = gsub("educ5", "", educ),
+  educ = gsub(".", " ", educ, fixed=TRUE)
+)  
+omni_educ <- omni_educ %>% mutate(
+  educ = factor(educ, levels = c("Less than HS", "HS or equivalent", "Some college", "Bachelor", "Post graduate"))
+)
+
+omni_educ <- omni_educ %>% mutate(
+  low = mean - 1.6445*SE, # for 90% MoE 
+  high = mean + 1.6445*SE
+)
+omni_educ <- omni_educ %>%
+  dplyr::select(-SE) %>%
+  pivot_longer(cols =c(mean, low, high), names_to="group", values_to="value") %>%
+  mutate(
+    value = 100*value
+  )
+
 education <- round_123 %>% 
@@
   theme(axis.text.x = element_text(angle=30, hjust = 1)) +
-  xlab("") + ylim(c(0,35)) +
+  xlab("") + ylim(c(0,40)) +
   facet_grid(.~chart) +
-  geom_point(aes(y = percent), data = acs_educ5)
+  geom_point(aes(y = percent), data = acs_educ5) +
+  geom_segment(
+    aes(
+      x = as.numeric(educ) - 0.45,
+      y = value,
+      xend = as.numeric(educ) + 0.45,
+      yend = value,
+      linetype=group
+    ), 
+    alpha = 1, colour = "black", linewidth = 0.25,
+    data = omni_educ
+  )  +
+  scale_linetype_manual(values=c(2,2,1)) +
+  theme(legend.position = "none")
+
+omni_income <- data.frame(svyby(~income4, by=~chart, design=survey_123, svymean))
+omni_income_long <- omni_income %>% 
+  pivot_longer(starts_with("income"), names_to="income", values_to = "mean") %>% 
+  dplyr::select(chart, income, mean) 
+omni_income_long$SE <- omni_income %>% pivot_longer(starts_with("se."), names_to="income", values_to="SE") %>% pull(SE)
+omni_income <- omni_income_long %>% mutate(
+  income = gsub("income4", "", income),
+  income = gsub(".", " ", income, fixed=TRUE)
+)  
+omni_income <- omni_income %>% mutate(
+  income = factor(income, levels = c("   30k", " 30k    60k", " 60k    100k" , "more than  100k"))
+)
+
+omni_income <- omni_income %>% mutate(
+  low = mean - 1.6445*SE, # for 90% MoE 
+  high = mean + 1.6445*SE
+)
+omni_income <- omni_income %>%
+  dplyr::select(-SE) %>%
+  pivot_longer(cols =c(mean, low, high), names_to="group", values_to="value") %>%
+  mutate(
+    value = 100*value
+  )
+levels(omni_income$income) <- levels(round_123$income4)
 
 income <- round_123 %>% 
-  ggplot(aes(x = income4, weight = weight_pct_round)) + geom_bar() + 
+  ggplot(aes(x = income4)) + 
+  geom_bar(aes(weight = weight_pct_round), fill="grey", colour = "grey30", linewidth=0.5) + 
   ylab("Percent") + 
@@
   theme(axis.text.x = element_text(angle=30, hjust = 1)) +
-  xlab("") + ylim(c(0,35)) +
-  facet_grid(.~chart)
+  xlab("") + ylim(c(0,40)) +
+  facet_grid(.~chart) +
+  geom_segment(
+    aes(
+      x = as.numeric(income) - 0.45,
+      y = value,
+      xend = as.numeric(income) + 0.45,
+      yend = value,
+      linetype=group
+    ), 
+    alpha = 1, colour = "black", linewidth = 0.25,
+    data = omni_income
+  )  +
+  scale_linetype_manual(values=c(2,2,1)) +
+  theme(legend.position = "none")
 

### 2023-07-24 SDSS-2023/testing-charts-accuracy.qmd
household income

@@
 
+acs_income <- data.frame(
+  income4 = c("< $30k", "$30k - $60k", 
+              "$60k - $100k", "more than $100k"), 
+  percent = c(6+3.9+7.5+ .5*7.8, 
+              .5*7.8 + 11.3 + .4*16.8, 
+              .6*16.8+12.8,
+              16.3+7.9+9.8))
+
+# Household income
+#     Less than $10,000	6.0
+#     $10,000 to $14,999	3.9
+#     $15,000 to $24,999	7.5
+#     $25,000 to $34,999	7.8
+#     $35,000 to $49,999	11.3
+#     $50,000 to $74,999	16.8
+#     $75,000 to $99,999	12.8
+#     $100,000 to $149,999	16.3
+#     $150,000 to $199,999	7.9
+#     $200,000 or more	9.8
+
+# 
+# acs_income <- data.frame(
+#   income4 = c("< $30k", "$30k - $60k", 
+#               "$60k - $100k", "more than $100k"), 
+#   percent = c(1807928 + 2669815 + 10192511 + .5*16440260, 
+#               .5*16440260+ 21573641+ 2/3*17579780, 
+#               1/3*17579780 + 8347923+13881892, 
+#               21161141)/113654891*100 )
+# Population age 18 and over
+# FULL-TIME, YEAR-ROUND WORKERS WITH EARNINGS	113654891	79395	
+#     $1 to $9,999 or loss	1807928	11027	1.6	0.1
+#     $10,000 to $14,999	2669815	16330	2.3	0.1
+#     $15,000 to $24,999	10192511	50628	9.0	0.1
+#     $25,000 to $34,999	16440260	63750	14.5	0.1
+#     $35,000 to $49,999	21573641	44746	19.0	0.1
+#     $50,000 to $64,999	17579780	35183	15.5	0.1
+#     $65,000 to $74,999	8347923	32204	7.3	0.1
+#     $75,000 to $99,999	13881892	50191	12.2	0.1
+#     $100,000 or more	21161141	100258	18.6	0.1
+
 omni_income <- data.frame(svyby(~income4, by=~chart, design=survey_123, svymean))
@@
   ylab("Percent") + 
-  ggtitle("Income levels") +
+  ggtitle("Income levels (Family)") +
   theme(axis.text.x = element_text(angle=30, hjust = 1)) +
@@
   facet_grid(.~chart) +
+  geom_point(aes(y = percent), data = acs_income) +
   geom_segment(

### 2023-07-24 SDSS-2023/testing-charts-accuracy.qmd
just a comment

@@
 The resulting model estimates and confidence intervals are displayed in \autoref{tbl-zoom-123}.
-We can observe that zooming behavior *only* differs significantly by structural design; rates do not differ significantly between the aligned and unaligned tasks, nor do they differ significantly by the chosen response.
+We observe that zooming behavior *only* differs significantly by structural design; rates do not differ significantly between the aligned and unaligned tasks, nor do they differ significantly by the chosen response.
 This higher rate of zooming does not necessarily lead to higher accuracy; although rates of accuracy are significantly higher for the vertical orientation relative to the horizontal orientation, they are not significantly higher -- and in fact, are significantly lower on the aligned task -- than the accuracy for the horizontal wide orientation.
 \hh{I'll pretty up @tbl-zoom-123 if we are going to keep it.} \kr{I think we should!
-it's nice to have a model for this}
+it's nice to have a model for this}\hh{gt and rmarkdown are frustrating!
+I'll re-set the table in its final form directly in tex}
 

### 2023-07-24 SDSS-2023/testing-charts-accuracy.qmd
ed's comments

@@
   \newcommand{\kr}[1]{{\textcolor{teal}{#1}}}.  
+  \newcommand{\ejm}[1]{{\textcolor{ForestGreen}{#1}}}.
 
@@
 The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households [@dennis2019technical].
-\hh{I don't think I've got the right citation} Each test was conducted using the AmeriSpeak Omnibus survey, which runs biweekly and samples around 1,000 U.S.
+\ejm{46,565 panelists are CAWI; they are 93.3% of the panel and 93.6% of the total panel weights.}
+Each test was conducted using the AmeriSpeak Omnibus survey, which runs biweekly and samples around 1,000 U.S.
 adults to answer questions on a variety of topics.
@@
 
-When combining responses from different surveys, weights need to be calibrated to the total population, so that their order after combining still reflects their relative importance.
+When combining responses from different surveys, weights \ejm{are rescaled} with respect to the total population, so that their order after combining still reflects their relative importance.
 
@@
 The resulting model estimates and confidence intervals are shown in \autoref{tbl-demographics}.
-When considering the easier (aligned) task, response patterns do not differ significantly by age, gender, or education level.
+\ejm{add legend for colors} When considering the easier (aligned) task, response patterns do not differ significantly by age, gender, or education level.
 We do observe, however, a significantly higher log odds of selecting the 'correct' or 'they are the same' response (or significantly lower log odds of selecting 'incorrect') among those with an income between \$60,000 and \$100,000.
@@
 For the  more difficult task of comparing unaligned tiles, demographics are more significant.
-Higher levels of education and higher levels of income are associated with a significant increase of panelists choosing a response of 'they are the same', resulting in significant decreases of both correct answers and wrong answers."
+Higher levels of education and higher levels of income are associated with a significant increase of panelists choosing a response of 'they are the same', resulting in significant decreases of both correct answers and wrong answers.
+"
 #| fig-subcap: 

### 2023-07-24 SDSS-2023/testing-charts-accuracy.qmd
some of these changes are just github's inability to deal with individual sentences

@@
 <!-- What is our main objective? -->
-
+<!--
 \kr{Should the abstract match -- or be close to -- our SDSS short abstract?}
 \hh{I don't think it has to}
+-->
 
 What do viewers see when we show them a data chart?
-A data chart -- at its core -- maps quantitative values to graphical elements representing their relative values.
+A data chart -- at its core -- maps values to graphical elements: quantitative elements are represented as measurable features, such as position, size, or shade.
+\hh{do we need to also go into qualitative variables?}
 Modern data visualizations are much more than a simple, objective mapping of values to a plane; they contain contextual and design elements, and are often structured to support the viewer in understanding a particular view of a set of data or specific pattern underlying the values.
-The design of a data visualization impacts a viewer's ability to achieve that understanding; a poorly designed data visualization may leave viewers struggling to understand the content or context, or make it difficult to complete accurate and useful comparison of values across groups or time points.
+The design of a data visualization impacts a viewer's ability to achieve that understanding; a poorly designed data visualization may leave viewers struggling to understand the content or context, or make it difficult to complete accurate and useful comparisons of values across groups or time points.
 More broadly, the design of a data visualization can change how viewers interact with the chart.
@@
 Cleveland and McGill's work -- while a foundational user study in graphical perception -- utilized a small convenience sample, consisting of only a few individuals recruited from among the authors' coworkers and their spouses.
-@heerCrowdsourcingGraphicalPerception2010 reproduced Cleveland and McGill's rankings using a larger sample from a crowd sourcing platform, employing a total of XXX Amazon Mechanical Turk (mTurk) workers for the study.
+@heerCrowdsourcingGraphicalPerception2010 reproduced Cleveland and McGill's rankings using a larger sample from a crowd sourcing platform, employing a total of 82 Amazon Mechanical Turk (mTurk) workers for the study.
 Crowd sourced samples were shown by @borgoCrowdsourcingInformationVisualization2017 to be biased towards more male, younger, and relatively higher education relative to the adult U.S.
@@
 
+<!--
 \kr{The mackinlay thing feels like maybe we are overemphasizing the ranking done by cleveland/mcgill and then mackinlay - I think we could emphasize less if we want}
 \hh{my point was to emphasize that Mackinlay is a foundational paper (rightfully so), but not based on any data beyond Cleveland}
+-->
 
@@
 
-Our work also centers on studying data visualization design choices and their impact on viewer behavior and accuracy of viewers' responses.
-We seek to answer whether it is possible to reproduce some of their findings in the context of a survey with a large, nationally-representative set of respondents, and within that context we focus on the following research questions:
+<!--Our work centers on studying data visualization design choices and their impact on viewer behavior and accuracy of viewers' responses.-->
+Here, we seek to answer whether it is possible to reproduce some of the previous findings in the context of a survey with a large, nationally-representative set of respondents, and within that context we focus on the following research questions:
 
@@
 
-We present viewers with structural variations of bar charts and ask them to answer questions comparing the size of elements within those charts.
+Specifically, we present viewers with structural variations of bar charts and ask them to answer questions comparing the size of elements within those charts.
 In this work, we present the series of tests we completed and the resulting findings.
-The remainder of the paper is organized as follows: first, we describe the design of the visual stimuli used in our perception tests.
+The remainder of the paper is organized as follows: first, we describe the design of the visual stimuli used in the perception tests.
 We then describe the population of study respondents and obtained survey sample.
-Subsequently, we share our analyses of the resulting survey responses across each of our tests, including analyses on accuracy of responses and response behavior.
+Subsequently, we share a summary of the resulting survey responses across each of the tests, including analyses on accuracy of responses and response behavior.
 Finally, we discuss implications of this work and next steps.
@@
 
-The comparison between sets of marked pairs is designed to be a difficult task, with the difference between the values represented in the two marked pieces being a just-noticeable difference.
+The comparison between sets of marked pairs is intentionally designed to be a difficult task, with the difference between the values represented in the two marked pieces being chosen close to their just-noticeable difference.
 The **Just-Noticeable Difference** (JND) is defined as the smallest difference that will be detected 50% of the time.
-Prior results from studies on bar charts and pie charts [@luModelingJustNoticeable2022] inform the differences in charts shown to our survey panelists.
+Prior results from studies on bar charts and pie charts [@luModelingJustNoticeable2022] inform the differences in charts shown to  survey panelists.
 
@@
 ```
-What we call 'aligned' and 'unaligned', here, is similar to Cleveland and McGill's set of rankings, but with some modifications: both 'aligned' and 'unaligned' bars share the same axis.
-Aligned tiles are additionally anchored in the same position in one dimension, i.e. the difference between their sizes can be reduced to a positional assessment.
-Unaligned tiles do not share this anchor, however, the context of the other tiles in the chart provide a frame, which *should* help with an assessment of the tiles' sizes beyond a comparison of (arc) lengths or areas.
+What we call 'aligned' and 'unaligned', here, is similar to Cleveland and McGill's positions along aligned and unaligned scales, but with some modifications: both 'aligned' and 'unaligned' bars share the same axis.
+Aligned tiles are additionally anchored in the same position in one dimension, i.e.
+the difference between their sizes can be reduced to a positional assessment, which in the absence of three-dimensional cues seems to be the general strategy for evaluating barcharts [@zacksReadingBarGraphs1998a].
+Unaligned tiles do not share this anchor, however, the context of the other tiles in the chart provide a frame, which *should* help with an assessment of the tiles' sizes.
 
@@
 \kr{The updated sketch is great here!! One note -- can we move the label for the "positions along the same, but unaligned scale", it's not aligned with the others}
-\hh{that was on purpose :(  that goes beyond Cleveland and McGill ... I'll see what I can do about that}
+\hh{What about the sketch now?}
+
+<!--
+@cejaTruthSquareAspect2021 suggest that the aspect ratio of areas causes biases in the viewers' assessments -- given that all of the stimuli are close in size (and therefore close in aspect ratio), we do not expect this to be an issue.
+-->
 

### 2023-07-24 analysis/analysis.qmd
some of these changes are just github's inability to deal with individual sentences

@@
 The data this section is based on comes from (a part) of Round 6 and all responses from Round 7.
+
+From @zacksReadingBarGraphs1998a a good explanation why the results from the facetted barcharts are so bad: we are running into a context-induced illusion.
+Lines are perceived to be bigger, if there is a bigger line next to them.
+-- Here, we see the opposite effect.
+
+> The graphical context in which an element occurs can produce distortions in judgments of color, angle, size, and orientation (see, e.g., Goldstein, 1989, especially chap.
+7; Howard, 1982).
+One particularly relevant example of such a distortion is given in Figure 1.
+This figure demonstrates the parallel lines illusion: When two parallel lines are viewed, the viewer tends to perceive assimilation (the lengths of the lines seem closer than they are) or contrast (the lengths of the lines seem more different than they are), depending on the ratio of the line lengths and the distance between them (Jordan & Schiano, 1986; Schiano, 1986).
 

### 2023-07-24 SDSS-2023/testing-charts-accuracy.qmd
legend added

@@
 lambda_1235 <- calculate_lambda(round_123$weight, round5_adj$weight)
+
 round_123b <- round_123 %>% 
@@
   rbind(
-    round5_adj %>% 
+    round5_adj
+     %>% 
       mutate(
@@
         chart = "Floating bars"
-       ) %>% 
-        dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
+       ) %>% dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
                    ab_certain, cd_certain, ab_zoom, cd_zoom, round, chart, weight,
-                   age4, gender, income4, racethnicity, educ5) 
-  )
+                   age4, gender, income4, racethnicity, educ5))
 
@@
 
-\hh{we might be able to include the lambda values for combining the tables here as well}
-
 <!-- | Name    | Date       |                \# Participants |                effective sample size |         Sum of weights $\sum_i w_i$ | -->
@@
 Distributions of demographic characteristics across each task are shown in \autoref{fig-rounds-demographics}, the dark points show percentages for US adults based on the 5 year estimates of the American Community Survey 2021.
-The distribution of demographics are all quite similar across tasks and their 90% margin of errors (hashed lines) mostly cover the Census estimates.
+The distribution of demographics are all quite similar across tasks and their 90% margin of errors (hashed lines) for the most part cover the Census estimates.
+
 <!--\kr{Split by rounds or by tasks?
@@
 The estimated values and 95% confidence intervals are shown in @tbl-rounds-123.
-\hh{The lowercase letters indicate significances between pairs of estimates: two estimates (in the same column) are significantly different at 5% level, if they do not have any letter in common [@piephoAlgorithmLetterBasedRepresentation2004].
+The lowercase letters indicate significances between pairs of estimates: two estimates (in the same column) are significantly different at 5% level, if they do not have any letter in common [@piephoAlgorithmLetterBasedRepresentation2004].
 
@@
 The resulting model estimates and confidence intervals are shown in \autoref{tbl-demographics}.
-\ejm{add legend for colors} When considering the easier (aligned) task, response patterns do not differ significantly by age, gender, or education level.
+When considering the easier (aligned) task, response patterns do not differ significantly by age, gender, or education level.
 We do observe, however, a significantly higher log odds of selecting the 'correct' or 'they are the same' response (or significantly lower log odds of selecting 'incorrect') among those with an income between \$60,000 and \$100,000.
@@
   geom_bar(aes(fill = response3, weight = weight), position="fill") +
-  theme(legend.position="none",
+  theme(legend.position="right",
         axis.text.x = element_text(angle=30, hjust=1)) + 
@@
   scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
-gender_cd + age_cd  + educ_cd +income_cd + plot_layout(nrow=1, guides = "collect")
-
-gender_ab + age_ab  + educ_ab +income_ab + plot_layout(nrow=1, guides = "collect")
+gender_cd + age_cd  + educ_cd +income_cd + 
+    plot_layout(nrow=1, guides = "collect", widths = c(2,4,5,4))
+
+gender_ab + age_ab  + educ_ab +income_ab + 
+  plot_layout(nrow=1, guides = "collect", widths = c(2,4,5,4))
 ```

### 2023-07-24 SDSS-2023/testing-charts-accuracy.qmd
results for 3b

@@
 ```
-What we call 'aligned' and 'unaligned', here, is similar to Cleveland and McGill's positions along aligned and unaligned scales, but with some modifications: both 'aligned' and 'unaligned' bars share the same axis.
-Aligned tiles are additionally anchored in the same position in one dimension, i.e.
-the difference between their sizes can be reduced to a positional assessment, which in the absence of three-dimensional cues seems to be the general strategy for evaluating barcharts [@zacksReadingBarGraphs1998a].
-Unaligned tiles do not share this anchor, however, the context of the other tiles in the chart provide a frame, which *should* help with an assessment of the tiles' sizes.
-
-We would expect that comparing unaligned tiles is a harder task (with correspondingly lower levels of accuracy) than a comparison of aligned tiles, with the framing given by the context of the other tiles in the same column mitigating some of this difficulty.
+What we call 'aligned' and 'unaligned', here, is similar to Cleveland and McGill's positions along aligned and unaligned scales, but with some modifications.
 @fig-tasks-explain gives an overview of the comparisons of tasks 1 through 3 and the closest corresponding tasks in Cleveland and McGill.
-\kr{The updated sketch is great here!! One note -- can we move the label for the "positions along the same, but unaligned scale", it's not aligned with the others}
-\hh{What about the sketch now?}
+
+In this paper, both 'aligned' and 'unaligned' bars share the same axis.
+
+Aligned tiles are additionally anchored in the same position along the axis, i.e.
+the difference between their sizes can be reduced to a positional assessment, which in the absence of three-dimensional cues seems to be the general strategy viewers use when evaluating barcharts [@zacksReadingBarGraphs1998a].
+Unaligned tiles do not share this anchor of position, i.e. the difference between their sizes needs at least two comparisons of positions.
+However, unaligned bars in a stacked barchart are not just assessments of the lengths of bars -- the other tiles in the chart provide a reference frame of the shared axis, which *should* help with an assessment of the tiles' sizes.
+As shown in the sketch of @fig-tasks-explain, this framing leads to a stimulus that is between Cleveland and McGill's task of 'position along the same but unaligned axis' and an assessment of length.
+
+We would expect that comparing unaligned tiles is a harder task (with correspondingly lower levels of accuracy) than a comparison of aligned tiles, with the framing given by the other tiles in the same column mitigating some of this difficulty.
+\hh{In particular, the additional context of the other tiles should *help* rather than *hurt* the comparison, unlike other situations, where additional information besides the pieces involved in the comparison were found to be distractors} [@talbotFourExperimentsPerception2014].
+<!--\kr{The updated sketch is great here!! One note -- can we move the label for the "positions along the same, but unaligned scale", it's not aligned with the others}-->
+<!--\hh{What about the sketch now?}-->
 
@@
 -->
-
-```{=tex}
-\hh{I think we should include the floating bars here to demonstrate that they are worse than the wide horizontal bars. }
-```
 
@@
 ```
-
-\hh{expectations - end}
 
@@
 Respondents were also asked to rate their certainty in their response to each question on a five-point scale.
-
-<!--
-```{r eval=FALSE}
-#| label: fig-tasks-old
-#| fig-cap: "Each of the visual stimuli presented to viewers. In each bar chart, two pieces are marked. The larger piece in each chart are pieces B and D. "
-#| fig-subcap:
-#| - "Vertical, unaligned"
-#| - "Horizontal, unaligned"
-#| - "Wide horizontal, unaligned"
-#| - "Vertical, aligned"
-#| - "Horizontal, aligned"
-#| - "Wide horizontal, aligned"
-#| layout-nrow: 3
-#| out-width: 90%
-cat("Task 1")
-cat("Task 2")
-cat("Task 3")
-knitr::include_graphics("images/vertical-ab.PNG")
-knitr::include_graphics("images/round3/horizontal-ab.PNG")
-knitr::include_graphics("images/round3/hwide-ab.PNG")
-knitr::include_graphics("images/vertical-cd.PNG")
-knitr::include_graphics("images/round3/horizontal-cd.PNG")
-knitr::include_graphics("images/round3/hwide-cd.PNG")
-```
--->
 
@@
 }-->
+
+### Task 3b: Floating bars
+
+```{r}
+n5 <- nrow(round5_adj)
+n5_eff <- neff(round5_adj$weight)
+```
+
+\hh{
+In order to measure the impact of the chart's context, we asked a subset of participants to assess the size of the two tiles A and B (as shown in the bottom right of @fig-tasks-explain) to each other without giving any additional information.}
 
@@
 | Task 3 | Horizontal wide |`r nrow(round3_adj[round3_adj$P_EXP == 2,])` | `r round(neff(round3_adj[round3_adj$P_EXP == 2,]$weight),1)` |
+| Task 3b | Floating bars |`r n5` | `r sprintf("%.1f",n5_eff)` |
 
@@
 The re-scaled design of the wide horizontal bars reclaims some of the loss for unaligned bars and outperforms the vertical design by a similar margin in aligned bars, but does not out-perform the vertical design when comparing unaligned tiles.
+\hh{The beneficial effect of frames/context of the chart is apparent when comparing accuracy of responses for the Horizontal wide design and the floating bars: the rate of correct responses in floating bars is significantly lower than in the wide horizontal charts.
+}
 
@@
 
+\hh{The actionable insights are:}
+
+- make sure that the main comparison in a plot is lined up -- in an interactive setting it would be beneficial to have the possibility of re-ordering the categories to help with comparisons. 
+- offering zoom-ins help with certainty, even if it does not help with perceptual accuracy
+
 <!-- Key findings to discuss: -->

### 2023-07-25 SDSS-2023/testing-charts-accuracy.qmd
include Nola's comments

@@
   \newcommand{\ejm}[1]{{\textcolor{ForestGreen}{#1}}}.  
+  \newcommand{\ndt}[1]{{\textcolor{purple}{#1}}}.
 
@@
 theme_set(theme_bw())
+theme_update(axis.ticks = element_blank())
 
@@
 
+\ndt{break up first sentence}
+
 ## Introduction
@@
 
-A crucial step in the process of interacting with and understanding a chart is the viewer's employment of comparisons of the parts within.
+\ndt{define structural}
+
+\ndt{The viewer's employment of comparisons of the components is a crucial step in the process of interacting with and understanding of chart.}
 @clevelandGraphicalPerceptionTheory1984 observed as such, and in their seminal study defined the better visual among a pair as the one that allows viewers to make more accurate comparisons.
-Based on mappings of quantitative variables to different graphical elements, Cleveland and McGill's study resulted in a ranking of perceptual tasks from most accurate to least accurate, which was then extended by @mackinlayAutomatingDesignGraphical1986 to a theoretical framework ranking tasks' order along their ordinal and nominal scales.
+Based on mappings of quantitative variables to different graphical elements \ndt{example of element to explain it XXX}, Cleveland and McGill's study resulted in a ranking of perceptual tasks \ndt{give example of task} from most accurate to least accurate, which was then extended by  @mackinlayAutomatingDesignGraphical1986 to a theoretical framework ranking tasks' order along their ordinal and nominal scales.
 
@@
 
+\ndt{I think the next paragraph can be moved above the RQs. Use this paragraph to set up the concepts, define constructs, then explain the study, and then note the RQs.}
 Specifically, we present viewers with structural variations of bar charts and ask them to answer questions comparing the size of elements within those charts.
@@
 
-Each task is made up of two elements: a visual stimulus and a question about that stimulus that the viewer is asked to respond to.
+\ndt{
+I think this paragraph should go first in this section.
+1. the survey format
+2. the type of data
+3. The stimulus
+3. how we set up the stimulus and why
+4. the tasks and alighed v unaligned
+4. qtext
+5. fig 1}
+
+Each task is made up of two elements: a visual stimulus and a question about the stimulus.
 In our study, each visual stimulus is an image of a data visualization, while each question prompts viewers to identify which of two marked pieces in the data visualization is larger.
@@
 Rather than asking the same (or similar) type of question with varied signal strength dozens or hundreds of times, we are limited to only a few questions at a time.
+\ndt{note why we are limited to only a few questions at a time - that is not clear.}
 With a small set of tasks, we need to present tasks that are perceptually hard, and thus ask questions about stimuli that are close to our perceptual threshold.
@@
 
-We ask participants to determine which of two just-noticeably different marked pieces is larger within each data visualization image, and focus on three main sets of structural variation in the design of that data visualization image.
-First, we vary the alignment of the pieces in question.
-Viewers are presented with two marked pieces in a chart that do not share a common baseline, then two pieces that do share a common baseline.
+We ask participants to determine which of two just-noticeably different marked pieces \ndt{is pieces the same as tiles below?} is larger within each chart, and focus on three main sets of structural variation in the design of that chart.
+First, we vary the alignment of the pieces in question. \ndt{perhaps a bulleted list would work better here}
+Viewers are presented with two marked pieces in a chart that do not share a common baseline, then two pieces that do share a common baseline. \ndt{"we refer to this as aligned and unaligned"}
 Second, we vary the orientation of the chart -- we rotate the vertical stacked bar chart, and present a horizontally oriented version of the same chart, with identically sized marked pieces.
@@
 
+\ndt{"The survey format guides the design and structure of the questions asked...." Why is this so?}
 Our use of a survey format guides the format and design of the questions asked and how they are presented to respondents.
@@
 adult; this risks participants reacting to the subject matter within the chart rather than focusing on the presented task.
+\ndt{stimulus?}
 For this reason, we utilize data on living arrangements of older U.S. adults -- a topic which most U.S. adults will have some familiarity with, but is not inherently polarizing.
@@
 When viewing each chart, participants were asked to compare the relative sizes of marked elements within the chart:
+
+\ndt{This should go above fig 1 to explain the set up for fig 1}
 
@@
 }-->
+\ndt{agree to drop note on color. Can perhaps just say we used CBlind  friendly}
 
@@
 \hh{
-In order to measure the impact of the chart's context, we asked a subset of participants to assess the size of the two tiles A and B (as shown in the bottom right of @fig-tasks-explain) to each other without giving any additional information.}
+In order to measure the impact of the chart's context, we asked a subset of participants to assess the relative sizes of the two tiles A and B (as shown in the bottom right of @fig-tasks-explain) without giving any additional information.}
 
@@
 
-<!-- | Name    | Date       |                \# Participants |                effective sample size |         Sum of weights $\sum_i w_i$ | -->
+<!-- | Name    | Date       |                \# Participants |                Effective Sample Size |         Sum of weights $\sum_i w_i$ | -->
 
@@
 
-| Name    | Date       |      \# Participants |                effective sample size |         Sum of weights $\sum_i w_i$ | $\lambda_i$ |
+| Name    | Date       |      \# Participants |                Effective Sample Size |         Sum of weights $\sum_i w_i$ | $\lambda_i$ |
 |---------------|---------------|--------------:|--------------:|--------------:|--------------:|
@@
 
-: Survey rounds: dates, number of participants (nominal sample size), effective sample size, sum of weights, and factors for the convex combination as discussed above. {#tbl-rounds}
+: Survey rounds: dates, number of participants (nominal sample size), effective sample size, sum of weights, and factors for the combining surveys as discussed above. {#tbl-rounds}
 
@@
 
-| Name    | Description    |      \# Participants |                effective sample size |
+| Name    | Description    |      \# Participants |                Effective Sample Size |
 |---------------|---------------|--------------:|--------------:|
@@
 #| label: fig-alignment
-#| out-width: 80%
-#| fig-height: 5
-#| fig-width: 10
+#| out-width: 90%
+#| fig-height: 4
+#| fig-width: 8.5
 
@@
   xlab("") +
-  scale_fill_manual(values=c("grey", "forestgreen")) +
+  scale_fill_manual("", values=c("grey", "forestgreen")) +
   scale_y_continuous(
@@
   ) +
-  theme(legend.position = "bottom") +
+  theme(legend.position = "left") +
   ggtitle("(a) Accuracy of survey responses")  +
@@
 #| label: fig-response-123
-#| fig-cap: "Responses for accuracy in the three designs.
-Responses to the same task are shown side-by-side for the three designs.
+#| fig-cap: "Responses for accuracy by task across the designs.
 The overlaid rectangles represent 95% confidence intervals.
 The letters in blue and orange encode significances between pairwise proportions: two bars have a significantly different proportion (at a 5% significance level) if they do not share a letter.
-There is no significant difference between the three designs for wrong responses.
+There is no significant difference between the designs for wrong responses.
 When tiles are unaligned, the horizontal wide barchart is showing the highest accuracy.
-For aligned tile, the horizontal wide design and the vertical design do not show a significant difference in accuracy."
-#| fig-height: 4.25
-#| fig-width: 7
-#| out-width: 80%
+For aligned tile, the horizontal wide design and the vertical design do not show a significant difference in accuracy.
+Floating bars have a significantly lower rate of correct answers than all the other designs."
+#| fig-height: 4
+#| fig-width: 8
+#| out-width: 90%
 
@@
   facet_grid(.~task, scales = "free_x", space="free_x") + 
-  theme(legend.position="bottom") + 
+  theme(legend.position="right") + 
   scale_fill_manual("Response", values=c("orange", "grey", "steelblue")) +
@@
 #| label: tbl-rounds-123
-#| tbl-cap: Log-odds for the cell-means model, letters behind numbers indicate pairwise significances.
-Within the same **column** values are significantly different (at 5%) if they do not share the same letter.
+#| tbl-cap: OOdds for the cell-means model, letters behind numbers indicate pairwise significances.
+Within the same *column* values are significantly different (at 5%) if they do not share the same letter.
 
@@
 
-One pattern in accuracy holds across each of the three structural variations: the aligned task has a higher level of accuracy than its unaligned counterpart.
+One pattern in accuracy holds across each of the three structural variations: **the aligned task has a higher level of accuracy than its unaligned counterpart**.
 Interestingly, while we expect an improvement in accuracy when shifting from the horizontal to the horizontal wide design given the larger difference in pixel length between the two pieces, the resulting effects on the accuracy of the responses are not completely straightforward: the shift from a vertical to the (tall) horizontal design is detrimental to an accurate perception for both aligned and unaligned comparisons.
@@
 \hh{The beneficial effect of frames/context of the chart is apparent when comparing accuracy of responses for the Horizontal wide design and the floating bars: the rate of correct responses in floating bars is significantly lower than in the wide horizontal charts.
-}
+This shows that the additional context }
 

### 2023-07-25 SDSS-2023/testing-charts-accuracy.qmd
typo

@@
 #| out-width: 90%
-#| fig-height: 4
 #| fig-width: 8.5
@@
 #| label: tbl-rounds-123
-#| tbl-cap: OOdds for the cell-means model, letters behind numbers indicate pairwise significances.
+#| tbl-cap: Odds for the cell-means model, letters behind numbers indicate pairwise significances.
 Within the same *column* values are significantly different (at 5%) if they do not share the same letter.
@@
   xlab("") + 
-  scale_fill_brewer(palette = "PRGn", guide = guide_legend(reverse = TRUE)) + 
+  scale_fill_brewer("How certain\nare you?", palette = "PRGn", guide = guide_legend(reverse = TRUE)) + 
   geom_point(

### 2023-07-25 SDSS-2023/testing-charts-accuracy.qmd
flipping attempt

@@
 
+# # attempt at flipping
+# gender + coord_flip() + education + coord_flip()+ age + coord_flip() + income + coord_flip()+ plot_layout(ncol=2, nrow=2, heights=c(2, 5, 4,4), byrow=FALSE)
+
 (gender + education + plot_layout(nrow=1, widths=c(2,5))) / 
   (age + income + plot_layout(nrow=1, widths = c(4,4)))
-
 ```

### 2023-07-25 SDSS-2023/testing-charts-accuracy.tex
tex file

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{jds}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{lscape}
\author[1]{Kiegan Rice\thanks{Corresponding author. Email: rice-kiegan@norc.org}}
\author[2]{Heike Hofmann\footnote{Email: hofmann@iastate.edu}}
\author[1]{Nola du Toit\footnote{Email: dutoit-nola@norc.org}}
\author[1]{Edward Mulrow\footnote{Email: mulrow-edward@norc.org}}
\affil[1]{NORC at the University of Chicago}
\affil[2]{Department of Statistics, Iowa State University}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Testing Charts: viewers' perceptual accuracy in surveys},
  pdfkeywords={list, keywords, here, 3-6, are, fine},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Testing Charts: viewers' perceptual accuracy in surveys}
\author{}
\date{}

\begin{document}
\maketitle
\begin{abstract}
The use of visuals is a key component in scientific communication, and
decisions about the design of a data visualization should be informed by
what design elements best support the audience's ability to perceive and
understand the components of the data visualization. We build on the
foundations of Cleveland and McGill's work in graphical perception,
employing a large, nationally-representative, probability-based panel of
survey respondents to test perception in statistical charts. Our
findings provide actionable guidance for data visualization
practitioners to employ in their work.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, breakable, enhanced]}{\end{tcolorbox}}\fi

  \newcommand{\hh}[1]{{\textcolor{orange}{#1}}} 
  \newcommand{\kr}[1]{{\textcolor{teal}{#1}}}.  
  \newcommand{\ejm}[1]{{\textcolor{ForestGreen}{#1}}}.  
  \newcommand{\ndt}[1]{{\textcolor{purple}{#1}}}.

% take out everything above later
% keep everything below

  \newcommand{\blandscape}{\begin{landscape}} 
  \newcommand{\elandscape}{\end{landscape}}

  \setlength{\parindent}{0pt}
  \singlespacing

\ndt{break up first sentence}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

What do viewers see when we show them a data chart? A data chart -- at
its core -- maps values to graphical elements: quantitative elements are
represented as measurable features, such as position, size, or shade.
\hh{do we need to also go into qualitative variables?} Modern data
visualizations are much more than a simple, objective mapping of values
to a plane; they contain contextual and design elements, and are often
structured to support the viewer in understanding a particular view of a
set of data or specific pattern underlying the values. The design of a
data visualization impacts a viewer's ability to achieve that
understanding; a poorly designed data visualization may leave viewers
struggling to understand the content or context, or make it difficult to
complete accurate and useful comparisons of values across groups or time
points. More broadly, the design of a data visualization can change how
viewers interact with the chart.

\ndt{define structural}

\ndt{The viewer's employment of comparisons of the components is a crucial step in the process of interacting with and understanding of chart.}

Cleveland and McGill (1984) observed as such, and in their seminal study
defined the better visual among a pair as the one that allows viewers to
make more accurate comparisons. Based on mappings of quantitative
variables to different graphical elements
\ndt{example of element to explain it XXX}, Cleveland and McGill's study
resulted in a ranking of perceptual tasks \ndt{give example of task}
from most accurate to least accurate, which was then extended by
Mackinlay (1986) to a theoretical framework ranking tasks' order along
their ordinal and nominal scales.

Cleveland and McGill's work -- while a foundational user study in
graphical perception -- utilized a small convenience sample, consisting
of only a few individuals recruited from among the authors' coworkers
and their spouses. Heer and Bostock (2010) reproduced Cleveland and
McGill's rankings using a larger sample from a crowd sourcing platform,
employing a total of 82 Amazon Mechanical Turk (mTurk) workers for the
study. Crowd sourced samples were shown by Borgo et al. (2017) to be
biased towards more male, younger, and relatively higher education
relative to the adult U.S. population as a whole; generally, convenience
and opt-in samples are not representative of the general population, a
common target audience for data visualization and scientific
communication work. Further, the populations' emphasis on higher
education individuals also leads to results which hold for groups of
individuals who may be more likely to have prior exposure to data
visualization in the context of scientific communication, or more
exposure to data topics in higher education, but may not hold across
other groups within the population.

Here, we seek to answer whether it is possible to reproduce some of the
previous findings in the context of a survey with a large,
nationally-representative set of respondents, and within that context we
focus on the following research questions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  How do structural design choices in a data visualization impact
  viewers' ability to identify the larger of two elements?
\item
  How is viewer interaction with the task impacted by structural design
  choices in a data visualization?
\item
  Are there differences in perception and interaction with the tasks
  across demographic groups?
\end{enumerate}

We employ a probability-based survey panel and run a series of
perception tests with nationally-representative samples of respondents
from that panel. The advantage of using a probability-based approach is
two-fold. First, we have access to a large sample of survey participants
and thus have greater power in making inference about graphical
perceptional abilities. Second, the sample is representative of the
general adult public in the U.S., and this allows us to test whether
prior results from convenience samples hold with a nationally
representative sample and whether there are differences in those results
across demographic subgroups.

\ndt{I think the next paragraph can be moved above the RQs. Use this paragraph to set up the concepts, define constructs, then explain the study, and then note the RQs.}

Specifically, we present viewers with structural variations of bar
charts and ask them to answer questions comparing the size of elements
within those charts. In this work, we present the series of tests we
completed and the resulting findings. The remainder of the paper is
organized as follows: first, we describe the design of the visual
stimuli used in the perception tests. We then describe the population of
study respondents and obtained survey sample. Subsequently, we share a
summary of the resulting survey responses across each of the tests,
including analyses on accuracy of responses and response behavior.
Finally, we discuss implications of this work and next steps.

\hypertarget{study-design-stimulus}{%
\subsection{Study design -- stimulus}\label{study-design-stimulus}}

\ndt{
I think this paragraph should go first in this section.
1. the survey format
2. the type of data
3. The stimulus
3. how we set up the stimulus and why
4. the tasks and alighed v unaligned
4. qtext
5. fig 1}

Each task is made up of two elements: a visual stimulus and a question
about the stimulus. In our study, each visual stimulus is an image of a
data visualization, while each question prompts viewers to identify
which of two marked pieces in the data visualization is larger.

The comparison between sets of marked pairs is intentionally designed to
be a difficult task, with the difference between the values represented
in the two marked pieces being chosen close to their just-noticeable
difference. The \textbf{Just-Noticeable Difference} (JND) is defined as
the smallest difference that will be detected 50\% of the time. Prior
results from studies on bar charts and pie charts (Lu et al. 2022)
inform the differences in charts shown to survey panelists.

We employ comparisons at the JND in our tasks in order to maximize our
ability to identify the impact of design changes on viewer accuracy and
behavior. Asking perception tasks in a survey differs from the
controlled environment of a cognitive lab, where these kind of questions
may usually be assessed. Rather than asking the same (or similar) type
of question with varied signal strength dozens or hundreds of times, we
are limited to only a few questions at a time.
\ndt{note why we are limited to only a few questions at a time - that is not clear.}
With a small set of tasks, we need to present tasks that are
perceptually hard, and thus ask questions about stimuli that are close
to our perceptual threshold. Therefore, we focus on questions which vary
the structure of the presented image, but ask viewers to compare the
same underlying data values across those varied images.

We ask participants to determine which of two just-noticeably different
marked pieces \ndt{is pieces the same as tiles below?} is larger within
each chart, and focus on three main sets of structural variation in the
design of that chart. First, we vary the alignment of the pieces in
question. \ndt{perhaps a bulleted list would work better here} Viewers
are presented with two marked pieces in a chart that do not share a
common baseline, then two pieces that do share a common baseline.
\ndt{"we refer to this as aligned and unaligned"} Second, we vary the
orientation of the chart -- we rotate the vertical stacked bar chart,
and present a horizontally oriented version of the same chart, with
identically sized marked pieces. Finally, we change the aspect ratio of
the chart and present a wider version of the horizontally oriented chart
which has longer, but thinner, marked pieces.

\hh{Not sure where should put this, parking it here for right now:}
\hh{expectations - start}

What we call `aligned' and `unaligned', here, is similar to Cleveland
and McGill's positions along aligned and unaligned scales, but with some
modifications. Figure~\ref{fig-tasks-explain} gives an overview of the
comparisons of tasks 1 through 3 and the closest corresponding tasks in
Cleveland and McGill.

In this paper, both `aligned' and `unaligned' bars share the same axis.

Aligned tiles are additionally anchored in the same position along the
axis, i.e.~the difference between their sizes can be reduced to a
positional assessment, which in the absence of three-dimensional cues
seems to be the general strategy viewers use when evaluating barcharts
(Zacks et al. 1998).\\
Unaligned tiles do not share this anchor of position, i.e.~the
difference between their sizes needs at least two comparisons of
positions. However, unaligned bars in a stacked barchart are not just
assessments of the lengths of bars -- the other tiles in the chart
provide a reference frame of the shared axis, which \emph{should} help
with an assessment of the tiles' sizes. As shown in the sketch of
Figure~\ref{fig-tasks-explain}, this framing leads to a stimulus that is
between Cleveland and McGill's task of `position along the same but
unaligned axis' and an assessment of length.

We would expect that comparing unaligned tiles is a harder task (with
correspondingly lower levels of accuracy) than a comparison of aligned
tiles, with the framing given by the other tiles in the same column
mitigating some of this difficulty.
\hh{In particular, the additional context of the other tiles should *help* rather than *hurt* the comparison, unlike other situations, where additional information besides the pieces involved in the comparison were found to be distractors}
(Talbot, Setlur, and Anand 2014).

\begin{figure*}[hbt]

{\centering \includegraphics[width=0.6\textwidth,height=\textheight]{images/sketch-tasks.png}

}

\caption{\label{fig-tasks-explain}Comparisons made in charts within the
Cleveland and McGill ranking}

\end{figure*}

\ndt{"The survey format guides the design and structure of the questions asked...." Why is this so?}

Our use of a survey format guides the format and design of the questions
asked and how they are presented to respondents. First, participant
instructions must be delivered in a very short and easily understandable
format, because participants cannot ask clarifying questions about the
task as they might be able to in a cognitive lab setting. Second, we
want to utilize content within the data visualization image which is not
socially or politically charged for the average U.S. adult; this risks
participants reacting to the subject matter within the chart rather than
focusing on the presented task. \ndt{stimulus?} For this reason, we
utilize data on living arrangements of older U.S. adults -- a topic
which most U.S. adults will have some familiarity with, but is not
inherently polarizing. Finally, to prevent viewers from being exposed to
slight variations of the same stimulus in a row (and risk unforeseen
order effects or respondents using prior questions to inform their
responses), we either split a survey sample in two and show each
subsample a distinct structural version of the chart or test variations
of a chart across distinct rounds of the survey.

\hypertarget{question-text}{%
\paragraph{Question text}\label{question-text}}

When viewing each chart, participants were asked to compare the relative
sizes of marked elements within the chart:

\ndt{This should go above fig 1 to explain the set up for fig 1}

``There are many charts used in the news media to portray data visually.
Looking at the chart below, which of the marked dark blue pieces is
bigger, A or B? Just your best guess is fine.

\begin{itemize}
\item
  A is bigger
\item
  B is bigger
\item
  They are the same
\end{itemize}

When presented with the aligned version of each chart, pieces were
marked with a C and D and the question text is updated accordingly. In
all scenarios, the second option (B {[}D{]} is bigger) is the correct
response.

For a given task, viewers are first presented with the unaligned version
of the task, followed by the aligned version of the task. The time in
seconds that each respondent spent on each task was recorded, as well as
whether the participant zoomed in on each chart while answering the
question. Respondents were also asked to rate their certainty in their
response to each question on a five-point scale.

\begin{figure*}[hbt]

{\centering \includegraphics[width=0.9\textwidth,height=\textheight]{images/stimuli-overview.png}

}

\caption{\label{fig-tasks}Each of the visual stimuli presented to
viewers. In each bar chart, two pieces are marked. The larger piece in
each chart are pieces B and D.}

\end{figure*}

\hypertarget{task-1-vertical-stacked-bar}{%
\subsubsection{Task 1: Vertical stacked
bar}\label{task-1-vertical-stacked-bar}}

\autoref{fig-tasks}(a,d) shows the two stacked bar charts shown to
participants in the first task. The marked tiles in each plot are 155
pixels apart, which leads to a JND of 3.5 pixels based on Lu et al.
(2022)`s model. The heights of the bars are 205 (left) and 213 pixels
(right), corresponding to about twice the JND. This difference should
lead to a relatively high accuracy rate for participants and
simultaneously limit the amount of frustration resulting from a task
that is perceived as 'too hard'.

Both charts show the same data with slight modifications to the order of
the levels -- the first and second level in each of the bars are
reversed between the left and the right chart. Both charts are displayed
at the same size, i.e.~in both cases both the difference in size between
the bars and the horizontal distance between the bars is the exact same
amount. This leaves the vertical positioning of the bars as the only
difference between the charts. Any differences in observed responses can
therefore be attributed to this difference in presentation.

Task 1 was presented to viewers in two rounds (Rounds 1-2), with four
color scheme variations each presented to 50\% of respondents in each
round. There were no significant differences in respondent accuracy or
behavior across color schemes, and we combine them here into one set of
stimulus responses to focus on the comparison of structural differences.
\kr{Should we put the color variations and a test for significance in accuracy/a visual in the appendix/supplementary materials? }
\hh{we could do that, but I'm not sure that we need to mention color at all.}

\hypertarget{task-2-horizontal-stacked-bar}{%
\subsubsection{Task 2: Horizontal stacked
bar}\label{task-2-horizontal-stacked-bar}}

The visual mappings in the second task, shown in \autoref{fig-tasks}(b,
e), are identical to the first task, but the axes of the chart are
rotated so that the stacked bars are represented in a horizontal format.
This represents a structural change in how the data are presented to the
viewer while preserving the pixel size of the elements viewers must
compare. Tasks 2 and 3 were asked within the same survey round (Round 3)
and the full sample was randomly split among respondents, with 50\% of
participants seeing the Task 2 stimuli and 50\% of participants seeing
the Task 3 stimuli.

\hypertarget{task-3-horizontal-wide-stacked-bar}{%
\subsubsection{Task 3: Horizontal wide stacked
bar}\label{task-3-horizontal-wide-stacked-bar}}

The images utilized in the third task again represent the same data as
in the first two tasks, and the overall image has dimensions which are
identical to the images in the first two tasks, displaying at the same
size during the survey.
\hh{The length of bars is increased to fit the new dimensions. This increases the difference in the length of bars from previously 8 pixels to 13 pixels. The widths of the tiles are adjusted accordingly (from 50 to 30 pixels) to keep the overall area of the tiles approximately constant. }
\ndt{agree to drop note on color. Can perhaps just say we used CBlind  friendly}

\hypertarget{task-3b-floating-bars}{%
\subsubsection{Task 3b: Floating bars}\label{task-3b-floating-bars}}

\hh{
In order to measure the impact of the chart's context, we asked a subset of participants to assess the relative sizes of the two tiles A and B (as shown in the bottom right of @fig-tasks-explain) without giving any additional information.}

\hypertarget{study-design-participants}{%
\subsection{Study design --
participants}\label{study-design-participants}}

Participants were recruited as part of NORC's AmeriSpeak panel, which
utilizes a probability-based sampling methodology and samples U.S.
households from NORC's National Sample Frame that provides coverage of
over 97\% of U.S. households. The current panel size is 54,001 panel
members aged 13 and over residing in over 43,000 households (Dennis
2019, updated 2022). \textbackslash ejm\{46,565 panelists are CAWI; they
are 93.3\% of the panel and 93.6\% of the total panel weights.\} Each
test was conducted using the AmeriSpeak Omnibus survey, which runs
biweekly and samples around 1,000 U.S. adults to answer questions on a
variety of topics. Our tests require visual inspection of an image; for
this reason, our survey questions were only presented to web-based
panelists and not panelists who respond via phone interviews.

The Omnibus sample is not longitudinal in nature, i.e.~we do not have
any information about whether the same panelists were included in
multiple rounds. While there is a (small) chance that panelists are
included in multiple rounds of our data, there was at least a one month
gap between viewing each visual stimulus, and for the purposes of the
analysis we will consider data from these viewings as independent.

\hypertarget{study-design-survey-weighting}{%
\subsection{Study design -- survey
weighting}\label{study-design-survey-weighting}}

Paragraph on how survey weights are derived? -- \textbf{Question for Ed
-- is this something AmeriSpeak can provide boilerplate language for?}

All calculations in this paper are done in R (R Core Team 2022), and
weights are applied in analyses using the \texttt{survey} package
(Lumley 2004) version 4.0 (Lumley 2020) based on Lumley (2010).

When combining responses from different surveys, weights
\ejm{are rescaled} with respect to the total population, so that their
order after combining still reflects their relative importance.

We \emph{combine} (rather than cumulate) a set of \(\ell\) surveys
\(S_1\), \(S_2\), \ldots{} \(S_\ell\) (with \$ \ell \ge 2\$ ) as
described in O'Muircheartaigh and Pedlow (2002), by multiplying weights
in \(S_i\) by \(\lambda_i\), for \(1 \le i \le \ell\).
\(\lambda_i \in [0,1]\) with \(\sum_i \lambda_i = 1\) is given as \[
\lambda_i = \frac{n_i/d_i}{\sum_{j=1}^{\ell}n_j/d_j},
\] where \(n_i\) is the nominal sample of survey \(S_i\) and \(d_i\) are
the design effects for the estimators. Here, \(d_i\) are estimated as \[
d_i = 1 + CV(w \in S_i)^2
\] where \(CV\) is the coefficient of variation of the weights \(w\)
within each sample, and is estimated as in Kish (1965) : \[ 
CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
\]

The data for this paper were collected in several rounds as part of the
NORC Omnibus. The resulting number of participants, effective sample
sizes, and \(\lambda\) values are shown in \autoref{tbl-rounds}.

\hypertarget{tbl-rounds}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}@{}}
\caption{\label{tbl-rounds}Survey rounds: dates, number of participants
(nominal sample size), effective sample size, sum of weights, and
factors for the combining surveys as discussed above.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Date
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\# Participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Effective Sample Size
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum of weights \(\sum_i w_i\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\lambda_i\)
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Date
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\# Participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Effective Sample Size
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum of weights \(\sum_i w_i\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\lambda_i\)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Round 1 & April 2022 & 933 & 521.1 & 934.9 & \(\lambda_1\) = 0.343 \\
Round 2 & May 2022 & 953 & 485.7 & 953.4 & \(\lambda_2\) = 0.320 \\
Round 3 & June 2022 & 921 & 513.5 & 923.1 & \(\lambda_3\) = 0.338 \\
\textbf{Total} & --- & 2807 & 1520.8 & & \\
\end{longtable}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{respondents}{%
\subsubsection{Respondents}\label{respondents}}

A total of 2807 respondents participated across the three rounds. The
number of responses and corresponding effective sample sizes in each
round are shown in \autoref{tbl-rounds}. All responses were combined
into one set of survey responses, with adjusted combined sample weights
and indicators for which task each respondent was exposed to. The
resulting sample sizes corresponding to each task are shown in
\autoref{tbl-tasks}.

\hypertarget{tbl-tasks}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}@{}}
\caption{\label{tbl-tasks}Survey tasks: number of participants (nominal
sample size) and effective sample size for each task.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\# Participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Effective Sample Size
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\# Participants
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Effective Sample Size
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Task 1 & Vertical & 1886 & 1007.1 \\
Task 2 & Horizontal & 459 & 266.5 \\
Task 3 & Horizontal wide & 462 & 249.1 \\
Task 3b & Floating bars & 915 & 473.0 \\
\end{longtable}

Distributions of demographic characteristics across each task are shown
in \autoref{fig-rounds-demographics}, the dark points show percentages
for US adults based on the 5 year estimates of the American Community
Survey 2021. The distribution of demographics are all quite similar
across tasks and their 90\% margin of errors (hashed lines) for the most
part cover the Census estimates.

\begin{figure*}[hbt]

{\centering \includegraphics{./figures/fig-rounds-demographics-1.png}

}

\caption{\label{fig-rounds-demographics}Demographics of respondents in
each of the tasks. The dots show percentages of the national
demographics as estimated by the American Community Survery (ACS) 2021.}

\end{figure*}

\hypertarget{accuracy-of-responses}{%
\subsubsection{Accuracy of responses}\label{accuracy-of-responses}}

We first investigate respondent accuracy in selecting the correct
response. As participants were able to select the option `They are the
same', there are several ways to model accuracy and response. The
argument could be made that for the purposes of practical
interpretation, `they are the same' is a correct choice, as the options
are visually very similar and not substantially different values within
the context of the data shown in the chart. However, we are interested
in understanding whether viewers \emph{can} perceive the difference and
correctly identify which piece is larger and thus consider multiple ways
of modeling response to investigate participant response patterns. We
begin by defining a measure of binary `correctness', for which all
answers that are not the correct option (B {[}D{]} is bigger) are
`incorrect', including the selection of `they are the same'.

\autoref{fig-alignment}(a) displays all responses along the binary
correctness measure, separated by whether the stimulus was an aligned
task or unaligned task. We can see that levels of accuracy for all
responses are significantly higher for the `easier' (aligned) task, with
about twice as many respondents correctly selecting the larger of the
two marked elements.

Because each participant was shown both the aligned and unaligned
versions of the chart, we can use a paired \(t\)-test to compare mean
accuracy between the two charts. The resulting \(t\)-statistic is highly
significant (\(t\) statistic: 21.2, df: 2265, \(p\)-value:
\(\ll 0.0001\)).

\begin{figure*}[hbt]

{\centering \includegraphics[width=0.9\textwidth,height=\textheight]{./figures/fig-alignment-1.png}

}

\caption{\label{fig-alignment}On the left (a), a stacked barchart shows
the number of respondents with correct (green) and incorrect (grey)
responses to the two comparison questions. When tiles are aligned along
the same axis, more than twice the number of responses are correct. The
shaded area along the top of the green tiles corresponds to 95\%
confidence intervals around (marginal) correct responses. On the right
(b), a parallel coordinate plot shows all combinations of responses.
There's a huge asymmetry in the number of responses where participants
answered only one of the questions correctly. A lot more responses are
correct when comparing aligned tiles than unaligned tiles.}

\end{figure*}

Next, we consider an ordinal model to investigate response behavior
across all three options -- `A {[}C{]} is bigger', `B {[}D{]} is
bigger', and `They are the same', and consider these response patterns
across each of the stimuli.

\begin{figure*}[hbt]

{\centering \includegraphics[width=0.9\textwidth,height=\textheight]{./figures/fig-response-123-1.png}

}

\caption{\label{fig-response-123}Responses for accuracy by task across
the designs. The overlaid rectangles represent 95\% confidence
intervals. The letters in blue and orange encode significances between
pairwise proportions: two bars have a significantly different proportion
(at a 5\% significance level) if they do not share a letter. There is no
significant difference between the designs for wrong responses. When
tiles are unaligned, the horizontal wide barchart is showing the highest
accuracy. For aligned tile, the horizontal wide design and the vertical
design do not show a significant difference in accuracy. Floating bars
have a significantly lower rate of correct answers than all the other
designs.}

\end{figure*}

Figure~\ref{fig-response-123} shows the results of a cell-means model
with ordinal response \(Y_k\), where \(Y_k\) is the \(k\)th
participant's response, \(Y_k \in \{1, 2, 3\}\), where `correct' is
encoded as 1, `they are the same' is encoded as 2, and `wrong' is
encoded as 3:

\[
\text{logit }P(Y_k \le \ell) = \mu_{ij\ell(k)},
\]

where \(\ell \in \{1, 2\}\); \(i \in \{1, 2\}\) is the comparison type
(1 = Aligned, 2 = Unaligned), and \(j \in \{1, 2, 3\}\) is the chart
design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide. The
estimated values and 95\% confidence intervals are shown in
Table~\ref{tbl-rounds-123}. The lowercase letters indicate significances
between pairs of estimates: two estimates (in the same column) are
significantly different at 5\% level, if they do not have any letter in
common (Piepho 2004).

\hypertarget{tbl-rounds-123}{}
\setlength{\LTpost}{0mm}
\begin{longtable}{lrrlrrl}
\caption{\label{tbl-rounds-123}Odds for the cell-means model, letters behind numbers indicate pairwise
significances. Within the same \emph{column} values are significantly
different (at 5\%) if they do not share the same letter. }\tabularnewline

\caption*{
{\large Odds of accuracy by task and chart type}
} \\ 
\toprule
 & \multicolumn{3}{c}{correct | same or wrong} & \multicolumn{3}{c}{correct or same | wrong} \\ 
\cmidrule(lr){2-4} \cmidrule(lr){5-7}
 & Est. &           [95\% CI] &   & Est. &           [95\% CI] &   \\ 
\midrule
\multicolumn{7}{l}{Unaligned} \\ 
\midrule
Floating bars\textsuperscript{\textit{1}} & $0.08$ &  [$0.06$, $0.12$] & a & $12.35$ &  [$8.10$, $18.83$] & ab \\ 
Horizontal & $0.22$ &  [$0.15$, $0.32$] & b & $5.54$ &  [$4.04$, $7.59$] & c \\ 
Horizontal wide & $0.26$ &  [$0.19$, $0.37$] & bc & $6.20$ &  [$4.46$, $8.61$] & ac \\ 
Vertical & $0.41$ &  [$0.36$, $0.47$] & cd & $6.90$ &  [$5.75$, $8.28$] & ac \\ 
\midrule
\multicolumn{7}{l}{Aligned} \\ 
\midrule
Horizontal & $0.63$ &  [$0.49$, $0.81$] & d & $14.02$ &  [$9.23$, $21.30$] & b \\ 
Horizontal wide & $2.67$ &  [$2.04$, $3.48$] & e & $17.12$ &  [$10.49$, $27.94$] & b \\ 
Vertical & $1.51$ &  [$1.33$, $1.71$] & f & $11.33$ &  [$8.99$, $14.28$] & b \\ 
\bottomrule
\end{longtable}
\begin{minipage}{\linewidth}
\textsuperscript{\textit{1}}We are including here, responses on an A vs B comparison for floating bars, corresponding to a length assessment, see bottom right of Figure 1.\\
\end{minipage}

One pattern in accuracy holds across each of the three structural
variations: \textbf{the aligned task has a higher level of accuracy than
its unaligned counterpart}. Interestingly, while we expect an
improvement in accuracy when shifting from the horizontal to the
horizontal wide design given the larger difference in pixel length
between the two pieces, the resulting effects on the accuracy of the
responses are not completely straightforward: the shift from a vertical
to the (tall) horizontal design is detrimental to an accurate perception
for both aligned and unaligned comparisons. The re-scaled design of the
wide horizontal bars reclaims some of the loss for unaligned bars and
outperforms the vertical design by a similar margin in aligned bars, but
does not out-perform the vertical design when comparing unaligned tiles.
\hh{The beneficial effect of frames/context of the chart is apparent when comparing accuracy of responses for the Horizontal wide design and the floating bars: the rate of correct responses in floating bars is significantly lower than in the wide horizontal charts. This shows that the additional context }

Rates of selecting the incorrect response are low across all three
structural variations and the aligned and unaligned tasks; while viewers
select the incorrect response at significantly higher rates for all
three unaligned tasks relative to the aligned tasks, those rates do not
differ significantly across the three structural variations. Most of the
observed differences in response patterns across structural variations
are attributed to respondents' selection between the correct option or
the `they are the same' option.

To better comprehend these observed patterns, we investigate viewer
interaction with the tasks more broadly as well as differences in
response selection across demographic groups.

\hypertarget{respondent-behavior}{%
\subsubsection{Respondent behavior}\label{respondent-behavior}}

One contributing factor to the observed patterns in response accuracy
might be the way that participants interact with the different designs.
Across all tasks, about half of all participants make use of the option
to zoom into charts. We observe that while zooming does help with the
overall accuracy (which is in agreement with the findings by Lu et al.
(2022) about the physical size of stimuli), the increase is not
significant. However, different designs lead to different rates of
zooming: we observe in \autoref{fig-zoom-123} that when dealing with the
vertical design, the rate of zooming is significantly higher than for
the two horizontal designs.

\begin{figure*}[hbt]

{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{./figures/fig-zoom-123-1.png}

}

\caption{\label{fig-zoom-123}Zooming - not significant for accuracy or
task, but changes by the type of chart.}

\end{figure*}

To formalize this pattern in a model, let \(Y_{jk}\) describe the
zooming behavior of panelist \(k\) on task \(j\). We model zooming
behavior (no = 0, yes = 1) as a logistic regression by correctness of
response (\(\rho\)), task (\(\tau\)), and design (\(\delta\)) of the
chart:

\[
\text{logit } P(Y_k \le 1) = \mu + \rho_{i(k)} + \tau_{j(k)} + \delta_{\ell(k)},
\]

The resulting model estimates and confidence intervals are displayed in
\autoref{tbl-zoom-123}. We observe that zooming behavior \emph{only}
differs significantly by structural design; rates do not differ
significantly between the aligned and unaligned tasks, nor do they
differ significantly by the chosen response. This higher rate of zooming
does not necessarily lead to higher accuracy; although rates of accuracy
are significantly higher for the vertical orientation relative to the
horizontal orientation, they are not significantly higher -- and in
fact, are significantly lower on the aligned task -- than the accuracy
for the horizontal wide orientation.
\hh{I'll pretty up @tbl-zoom-123 if we are going to keep it.}
\kr{I think we should! it's nice to have a model for this}\hh{gt and rmarkdown are frustrating! I'll re-set the table in its final form directly in tex}

\hypertarget{tbl-zoom-123}{}
\begin{longtable}{lrrrr}
\caption{\label{tbl-zoom-123}Coefficients for logistic regression of zooming by task }\tabularnewline

\caption*{
{\large Estimates for logistic regression on zooming behavior}
} \\ 
\toprule
term & estimate & SE & t-statistic & p-value \\ 
\midrule
\multicolumn{5}{l}{} \\ 
\midrule
\$\$\textbackslash{}hat\{\textbackslash{}mu\}\$\$ & $-0.35$ & $0.15$ & $-2.4$ & 0.0155 \\ 
\midrule
\multicolumn{5}{l}{Response} \\ 
\midrule
\$\$\textbackslash{}text\{They are the same \} \textbackslash{}widehat\{\textbackslash{}rho\}\_\{2\}\$\$ & $-0.17$ & $0.09$ & $-1.9$ & 0.0578 \\ 
Wrong \$\$\textbackslash{}widehat\{\textbackslash{}rho\}\_\{3\}\$\$ & $-0.06$ & $0.13$ & $-0.5$ & 0.6466 \\ 
\midrule
\multicolumn{5}{l}{Task} \\ 
\midrule
Aligned \$\$\textbackslash{}widehat\{\textbackslash{}tau\}\_\{2\}\$\$ & $0.00$ & $0.06$ & $-0.1$ & 0.9435 \\ 
\midrule
\multicolumn{5}{l}{Chart design} \\ 
\midrule
Horizontal wide \$\$\textbackslash{}widehat\{\textbackslash{}delta\}\_\{2\}\$\$ & $0.27$ & $0.15$ & $1.8$ & 0.0752 \\ 
Vertical \$\$\textbackslash{}widehat\{\textbackslash{}delta\}\_\{3\}\$\$ & $0.98$ & $0.13$ & $7.7$ & < 0.0001 \\ 
\bottomrule
\end{longtable}

\hypertarget{how-certain-are-participants}{%
\subsubsection{How certain are
participants?}\label{how-certain-are-participants}}

Using linear scores for the response of \texttt{Certainty}, with
\texttt{not\ certain\ at\ all} assigned a score of 1 and
\texttt{extremely\ certain} assigned a score of 5, we can estimate the
effects of task, chart design, and correctness on certainty by using a
cell-means model of the form: \[
Y_{k} = \mu_{ij\ell(k)} + \epsilon_{k},
\] where \(k = 1, ..., N\), \(\mu_{ij\ell(k)}\) is average certainty
(measured on a scale from 1 to 5) of the four combinations of task and
correctness by each of the three designs, where \(i = 1, 2\) encodes
unaligned/aligned, and \(j=1,2\) encodes wrong, correct, and
\(\ell = 1, 2, 3\) encodes horizontal, wide horizontal, and vertical
stacked bar charts, respectively. We also assume that errors are
normally distributed,
i.e.~\(\epsilon_k \stackrel{i.i.d}{\sim} N(0, \sigma^2)\) for all
\(k = 1, ..., N\). The results are shown in Figure~\ref{fig-certainty}.
What we find, is that the highest scores for certainty are associated
with comparing aligned tiles. Certainty scores are not significantly
different between correct and wrong responses. Scores are (mostly)
significantly lower for the unaligned task. Interestingly, the lowest
scores of certainty are associated with correct responses, here.

\begin{figure*}[hbt]

{\centering \includegraphics[width=1\textwidth,height=\textheight]{./figures/fig-certainty-1.png}

}

\caption{\label{fig-certainty}Certainty by task, chart design, and
correctness. Correct answers in the unaligned task have the lowest
certainty scores associated with them. Incorrect responses on the
unaligned task show a significant boost in certainty in the vertical
stacked bar. For horizontal stacked bars the difference between aligned
and unaligned charts matters significantly. Errorbars around the points
show estimated scores (on a scale of 1 to 5) with corresponding 95\% CI.
The letters underneath the bars indicate significances. The scores for
two bars are significantly different if they do not share a letter.}

\end{figure*}

\hypertarget{differences-across-demographic-groups}{%
\subsubsection{Differences across demographic
groups}\label{differences-across-demographic-groups}}

Finally, we turn to investigating responses across demographic groups, a
core benefit provided by the large and representative nature of our
survey samples. \autoref{fig-demographics} displays response patterns
across our three structural variations by gender, age (4 groups),
educational attainment (5 groups), and income level (4 groups). Our
pattern of higher accuracy in selecting the correct response on the
aligned task relative to the unaligned task holds across all demographic
groups and structural variations. However, we do observe different
response patterns, particularly on the unaligned task, across
demographic groups.

Let \(Y_k\) be the response of participant \(k\), on a scale from 1 =
`wrong', 2 = `they are the same' to 3 = `correct'. We use a generalized
cumulative logistic regression, where \(\mu_i\) are intercepts
\(1 \le i < 3\), \(X_k\) are demographics of the \(k\)th participant (in
form of the model matrix), and \(\beta_{i}\) are the coefficients.

\[
\text{ logit } P(Y_k \le i \mid X_k) = \mu_i + X_{k}'\beta_{i}
\] The resulting model estimates and confidence intervals are shown in
\autoref{tbl-demographics}. When considering the easier (aligned) task,
response patterns do not differ significantly by age, gender, or
education level. We do observe, however, a significantly higher log odds
of selecting the `correct' or `they are the same' response (or
significantly lower log odds of selecting `incorrect') among those with
an income between \$60,000 and \$100,000. When we consider the more
difficult (unaligned) task, we see significant separation in response
patterns across gender, educational attainment, and income groups. We do
not see significant differences in responses by age for either task.
Interestingly, we see lower log odds of selecting the `correct' response
among those with higher educational attainment (those with a bachelor's
degree or post graduate study) \emph{and} lower log odds of selecting
the `incorrect' response; as income and educational attainment increase,
respondents are more likely to select the `they are the same' option
during the difficult task. This speaks to the difficulty of the task,
and respondents' interpretation of the task when it is more difficult to
perceive small differences.

\begin{figure}

\begin{minipage}[t]{\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{./figures/fig-demographics-1.png}

}

}

\subcaption{\label{fig-demographics-1}Comparisons of aligned tiles}
\end{minipage}%
\newline
\begin{minipage}[t]{\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{./figures/fig-demographics-2.png}

}

}

\subcaption{\label{fig-demographics-2}Comparisons of unaligned tiles}
\end{minipage}%

\caption{\label{fig-demographics}Panelists' demographics matter,
particularly, when the task difficulty increases. For aligned tiles,
gender, age, and education are not significant factors. However, income
levels do have a (small) effect. When income levels increase, the
percentage of wrong answers (orange) decreases, while the percentage of
correct answers (blue) increases slightly (significant at below 0.05).
For the more difficult task of comparing unaligned tiles, demographics
are more significant. Higher levels of education and higher levels of
income are associated with a significant increase of panelists choosing
a response of `they are the same', resulting in significant decreases of
both correct answers and wrong answers.}

\end{figure}

\blandscape

\hypertarget{tbl-demographics}{}
\setlength{\LTpost}{0mm}
\begin{longtable}{lrrlrrlrrlrrl}
\caption{\label{tbl-demographics}Demographics matter for perception, particularly when the tasks get
harder. }\tabularnewline

\caption*{
{\large Odds of accuracy by task and demographics of respondents}
} \\ 
\toprule
 & \multicolumn{6}{c}{Aligned tiles} & \multicolumn{6}{c}{Unaligned tiles} \\ 
\cmidrule(lr){2-7} \cmidrule(lr){8-13}
 & \multicolumn{3}{c}{correct | same or wrong} & \multicolumn{3}{c}{correct or same | wrong} & \multicolumn{3}{c}{correct | same or wrong } & \multicolumn{3}{c}{correct or same | wrong } \\ 
\cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10} \cmidrule(lr){11-13}
  & Est. &           [95\% CI] &   & Est. &           [95\% CI] &   & Est. &           [95\% CI] &   & Est. &           [95\% CI] &   \\ 
\midrule
\multicolumn{13}{l}{Intercept} \\ 
\midrule
 & $1.51$ &  [$0.91$, $2.50$] &  & $12.75$ &  [$5.59$, $29.10$] & *** & $0.60$ &  [$0.35$, $1.03$] & . & $3.24$ &  [$1.70$, $6.19$] & *** \\ 
\midrule
\multicolumn{13}{l}{Gender} \\ 
\midrule
Female & $0.82$ &  [$0.67$, $1.01$] & . & $1.19$ &  [$0.81$, $1.74$] &  & $1.03$ &  [$0.82$, $1.30$] &  & $1.42$ &  [$1.07$, $1.90$] & * \\ 
\midrule
\multicolumn{13}{l}{Age} \\ 
\midrule
30-44 & $1.08$ &  [$0.78$, $1.49$] &  & $0.87$ &  [$0.43$, $1.75$] &  & $0.79$ &  [$0.55$, $1.13$] &  & $0.94$ &  [$0.59$, $1.49$] &  \\ 
45-59 & $1.13$ &  [$0.80$, $1.60$] &  & $0.95$ &  [$0.46$, $1.95$] &  & $0.92$ &  [$0.62$, $1.35$] &  & $0.76$ &  [$0.48$, $1.22$] &  \\ 
60+ & $0.96$ &  [$0.69$, $1.34$] &  & $0.70$ &  [$0.35$, $1.40$] &  & $1.01$ &  [$0.70$, $1.45$] &  & $0.73$ &  [$0.46$, $1.15$] &  \\ 
\midrule
\multicolumn{13}{l}{Education} \\ 
\midrule
HS or equivalent & $0.77$ &  [$0.46$, $1.30$] &  & $0.50$ &  [$0.20$, $1.21$] &  & $0.81$ &  [$0.47$, $1.40$] &  & $1.22$ &  [$0.66$, $2.27$] &  \\ 
Some college & $0.91$ &  [$0.56$, $1.49$] &  & $0.74$ &  [$0.32$, $1.71$] &  & $0.63$ &  [$0.38$, $1.06$] & . & $1.17$ &  [$0.65$, $2.12$] &  \\ 
Bachelor & $0.79$ &  [$0.48$, $1.32$] &  & $1.12$ &  [$0.42$, $2.97$] &  & $0.53$ &  [$0.31$, $0.93$] & * & $1.82$ &  [$0.91$, $3.64$] & . \\ 
Post graduate & $0.84$ &  [$0.49$, $1.44$] &  & $1.16$ &  [$0.41$, $3.30$] &  & $0.51$ &  [$0.28$, $0.91$] & * & $2.18$ &  [$1.04$, $4.57$] & * \\ 
\midrule
\multicolumn{13}{l}{Income} \\ 
\midrule
\$30k - \$60k & $1.19$ &  [$0.87$, $1.63$] &  & $1.25$ &  [$0.77$, $2.05$] &  & $1.00$ &  [$0.71$, $1.41$] &  & $1.22$ &  [$0.82$, $1.82$] &  \\ 
\$60k - \$100k & $1.23$ &  [$0.90$, $1.68$] &  & $2.14$ &  [$1.18$, $3.86$] & * & $0.87$ &  [$0.60$, $1.27$] &  & $1.88$ &  [$1.24$, $2.87$] & ** \\ 
more than \$100k & $1.36$ &  [$0.99$, $1.88$] & . & $1.57$ &  [$0.86$, $2.86$] &  & $0.86$ &  [$0.59$, $1.25$] &  & $2.18$ &  [$1.34$, $3.56$] & ** \\ 
\bottomrule
\end{longtable}
\begin{minipage}{\linewidth}
Signif. codes:  0 `***' 0.001 `**' 0.01 `*' 0.05 `.' 0.1 ` ' 1\\
\end{minipage}

\elandscape

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

Through this work, we have tested the use of probability-based survey
panels to ask perception questions; within a large survey covering a
variety of topics, and with a limited number of questions for our
specific tasks, we are able to measure viewer perception and produce
results consistent with prior studies. Testing data visualization design
structures and viewer behavior on a nationally-representative sample is
of broad interest and applicability to the scientific communication
community, and this study demonstrates a framework under which to
complete further work in this area.

With our survey framework, we have shown that we can rank structural
variations on design and tasks by their relative accuracy; we identified
significantly different rates of respondent accuracy under different
designs presenting the same data. The size of our sample aids in
identifying this signal.

Further, we can also study respondent behavior and how viewers interact
with the chart. Paradata on viewer interaction with the chart (e.g.,
zooming and time spent on each task) can be collected as well as asking
directly asking respondents questions about their certainty. Not only
can we collect this information, but we can also identify significant
differences in this behavior across different designs. Understanding
viewer interaction and engagement is a key component to designing
effective data visualizations. If a particular design leads viewers to
zoom in more to investigate it when determining a response, we might
consider that design to be more frustrating to viewers `in the wild'
when viewing it outside of the context of our tests.
\kr{I don't love the last sentence but I think we can clean it up}.

Perhaps most salient among our results are the patterns in response
behavior across demographic groups. The large size of our sample, paired
with the probability-based survey panel approach, afford us the ability
to dive into response patterns among demographics across tasks and
identify significant differences among them. In particular, the
differences across educational attainment groups and income groups
underscore the importance of utilizing a representative population when
testing perception; prior results in this area may be absorbing bias in
responses due to the larger representation of individuals with higher
education levels among those study respondents.

While our results are limited to mapping in a stacked bar chart with a
specific topic, they demonstrate among them a wealth of insights about
how respondents perceive and interact with data visualizations. More
expansive work on this topic should be completed to understand how the
general public interacts with and understands charts, including
expanding to a wider set of structural variations, more aesthetic design
variations, and considering differences in respondent behavior across
different data topic areas.

\hh{The actionable insights are:}

\begin{itemize}
\tightlist
\item
  make sure that the main comparison in a plot is lined up -- in an
  interactive setting it would be beneficial to have the possibility of
  re-ordering the categories to help with comparisons.
\item
  offering zoom-ins help with certainty, even if it does not help with
  perceptual accuracy
\end{itemize}

\hypertarget{supplementary-material}{%
\section*{Supplementary Material}\label{supplementary-material}}
\addcontentsline{toc}{section}{Supplementary Material}

\begin{itemize}
\tightlist
\item
  \textbf{Participant Data (Linear):} Link to csv file with the data.
\item
  \textbf{Data Analysis Code:} Link to an html document with annotated
  code chunks.
\end{itemize}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-borgoCrowdsourcingInformationVisualization2017}{}}%
Borgo, Rita, Bongshin Lee, Benjamin Bach, Sara Fabrikant, Radu Jianu,
Andreas Kerren, Stephen Kobourov, et al. 2017. {``Crowdsourcing for
{Information Visualization}: {Promises} and {Pitfalls}.''} In
\emph{Evaluation in the {Crowd}. {Crowdsourcing} and {Human-Centered
Experiments}}, edited by Daniel Archambault, Helen Purchase, and Tobias
Hoßfeld, 96--138. Lecture {Notes} in {Computer Science}. {Springer
International Publishing}.

\leavevmode\vadjust pre{\hypertarget{ref-clevelandGraphicalPerceptionTheory1984}{}}%
Cleveland, William S., and Robert McGill. 1984. {``Graphical Perception:
{Theory}, Experimentation, and Application to the Development of
Graphical Methods.''} \emph{Journal of the American Statistical
Association} 79 (387): 531--54.
\url{https://doi.org/10.1080/01621459.1984.10478080}.

\leavevmode\vadjust pre{\hypertarget{ref-dennis2019technical}{}}%
Dennis, J Michael. 2019, updated 2022. {``Technical Overview of the
{AmeriSpeak} Panel {NORC}'s Probability-Based Household Panel.''}
\emph{NORC at the University of Chicago}, 2019, updated 2022.

\leavevmode\vadjust pre{\hypertarget{ref-heerCrowdsourcingGraphicalPerception2010}{}}%
Heer, J., and M. Bostock. 2010. {``Crowdsourcing Graphical Perception:
{Using} Mechanical {Turk} to Assess Visualization Design.''} In
\emph{Conference on {Human Factors} in {Computing Systems} -
{Proceedings}}, 1:203--12. {ACM}.
\url{https://doi.org/10.1145/1753326.1753357}.

\leavevmode\vadjust pre{\hypertarget{ref-kishSurveySampling1965}{}}%
Kish, Leslie. 1965. \emph{Survey {Sampling}}. {Wiley}.

\leavevmode\vadjust pre{\hypertarget{ref-luModelingJustNoticeable2022}{}}%
Lu, Min, Joel Lanir, Chufeng Wang, Yucong Yao, Wen Zhang, Oliver
Deussen, and Hui Huang. 2022. {``Modeling {Just Noticeable Differences}
in {Charts}.''} \emph{IEEE Transactions on Visualization and Computer
Graphics} 28 (1): 718--26.
\url{https://doi.org/10.1109/TVCG.2021.3114874}.

\leavevmode\vadjust pre{\hypertarget{ref-lumleyAnalysisComplexSurvey2004}{}}%
Lumley, Thomas. 2004. {``Analysis of Complex Survey Samples.''}
\emph{Journal of Statistical Software} 9 (1): 1--19.

\leavevmode\vadjust pre{\hypertarget{ref-lumleyComplexSurveysGuide2010}{}}%
---------. 2010. \emph{Complex Surveys: {A} Guide to Analysis Using {R}:
{A} Guide to Analysis Using {R}}. {John Wiley and Sons}.

\leavevmode\vadjust pre{\hypertarget{ref-survey}{}}%
---------. 2020. {``Survey: Analysis of Complex Survey Samples.''}

\leavevmode\vadjust pre{\hypertarget{ref-mackinlayAutomatingDesignGraphical1986}{}}%
Mackinlay, Jock. 1986. {``Automating the Design of Graphical
Presentations of Relational Information.''} \emph{ACM Transactions on
Graphics} 5 (2): 110--41. \url{https://doi.org/10.1145/22949.22950}.

\leavevmode\vadjust pre{\hypertarget{ref-omuircheartaighCombiningSamplesVs2002}{}}%
O'Muircheartaigh, Colm, and Steven Pedlow. 2002. {``Combining {Samples
Vs}. {Cumulating Cases}: {A Comparison} of {Two Weighting Strategies} in
{NLSY97}.''} In \emph{{ASA Proceedings} of the {Joint Statistical
Meetings}}, 2557--62.
\url{http://www.asasrms.org/Proceedings/y2002/Files/JSM2002-001082.pdf}.

\leavevmode\vadjust pre{\hypertarget{ref-piephoAlgorithmLetterBasedRepresentation2004}{}}%
Piepho, Hans-Peter. 2004. {``An {Algorithm} for a {Letter-Based
Representation} of {All-Pairwise Comparisons}.''} \emph{Journal of
Computational and Graphical Statistics} 13 (2): 456--66.
\url{https://doi.org/10.1198/1061860043515}.

\leavevmode\vadjust pre{\hypertarget{ref-RLanguage}{}}%
R Core Team. 2022. \emph{R: {A} Language and Environment for Statistical
Computing}. Manual. {Vienna, Austria}: {R Foundation for Statistical
Computing}. \url{https://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-talbotFourExperimentsPerception2014}{}}%
Talbot, Justin, Vidya Setlur, and Anushka Anand. 2014. {``Four
{Experiments} on the {Perception} of {Bar Charts}.''} \emph{IEEE
Transactions on Visualization and Computer Graphics} 20 (12): 2152--60.
\url{https://doi.org/10.1109/TVCG.2014.2346320}.

\leavevmode\vadjust pre{\hypertarget{ref-zacksReadingBarGraphs1998a}{}}%
Zacks, Jeff, Ellen Levy, Barbara Tversky, and Diane J. Schiano. 1998.
{``Reading Bar Graphs: {Effects} of Extraneous Depth Cues and Graphical
Context.''} \emph{Journal of Experimental Psychology: Applied} 4 (2):
119--38. \url{https://doi.org/10.1037/1076-898X.4.2.119}.

\end{CSLReferences}

\end{document}

### 2023-07-27 SDSS-2023/testing-charts-accuracy.qmd
addressing some of NDT comments

@@
 A data chart -- at its core -- maps values to graphical elements: quantitative elements are represented as measurable features, such as position, size, or shade. 
-\hh{do we need to also go into qualitative variables?}
 Modern data visualizations are much more than a simple, objective mapping of values to a plane; they contain contextual and design elements, and are often structured to support the viewer in understanding a particular view of a set of data or specific pattern underlying the values.
+Structural choices, such as the choice of a graphical element (e.g., bar or pie), orientation of the chart, and sizing and relative positioning of elements within the chart, are determined by the data visualization practitioner.
 The design of a data visualization impacts a viewer's ability to achieve that understanding; a poorly designed data visualization may leave viewers struggling to understand the content or context, or make it difficult to complete accurate and useful comparisons of values across groups or time points.
@@
 
-\ndt{define structural}
-
-\ndt{The viewer's employment of comparisons of the components is a crucial step in the process of interacting with and understanding of chart.}
+The viewer's employment of comparisons of the components is a crucial step in the process of interacting with and understanding of chart.
 @clevelandGraphicalPerceptionTheory1984 observed as such, and in their seminal study defined the better visual among a pair as the one that allows viewers to make more accurate comparisons.
-Based on mappings of quantitative variables to different graphical elements \ndt{example of element to explain it XXX}, Cleveland and McGill's study resulted in a ranking of perceptual tasks \ndt{give example of task} from most accurate to least accurate, which was then extended by  @mackinlayAutomatingDesignGraphical1986 to a theoretical framework ranking tasks' order along their ordinal and nominal scales.
+Based on mappings of quantitative variables to different graphical elements, Cleveland and McGill's study resulted in a ranking of perceptual tasks (such as ranking the larger of two lengths or areas) from most accurate to least accurate, which was then extended by  @mackinlayAutomatingDesignGraphical1986 to a theoretical framework ranking tasks' order along their ordinal and nominal scales.
 
@@
 <!--Our work centers on studying data visualization design choices and their impact on viewer behavior and accuracy of viewers' responses.-->
-Here, we seek to answer whether it is possible to reproduce some of the previous findings in the context of a survey with a large, nationally-representative set of respondents, and within that context we focus on the following research questions:
-
-1.  How do structural design choices in a data visualization impact viewers' ability to identify the larger of two elements?
-
-2.  How is viewer interaction with the task impacted by structural design choices in a data visualization?
-
-3.  Are there differences in perception and interaction with the tasks across demographic groups?
-
+Here, we seek to answer whether it is possible to reproduce some of the previous findings in the context of a survey with a large, nationally-representative set of respondents.
+Specifically, we present viewers with structural variations of bar charts and ask them to answer questions comparing the size of elements within those charts.
 We employ a probability-based survey panel and run a series of perception tests with nationally-representative samples of respondents from that panel.
@@
 Second, the sample is representative of the general adult public in the U.S., and this allows us to test whether prior results from convenience samples hold with a nationally representative sample and whether there are differences in those results across demographic subgroups.
-
-\ndt{I think the next paragraph can be moved above the RQs. Use this paragraph to set up the concepts, define constructs, then explain the study, and then note the RQs.}
-Specifically, we present viewers with structural variations of bar charts and ask them to answer questions comparing the size of elements within those charts.
+Within that context we focus on the following research questions:
+
+1.  How do structural design choices in a data visualization impact viewers' ability to identify the larger of two elements?
+
+2.  How is viewer interaction with the task impacted by structural design choices in a data visualization?
+
+3.  Are there differences in perception and interaction with the tasks across demographic groups?
+
 In this work, we present the series of tests we completed and the resulting findings.
@@
 
+\ndt{"The survey format guides the design and structure of the questions asked...." Why is this so?}
+Our use of a survey format guides the format and design of the questions asked and how they are presented to respondents.
+First, participant instructions must be delivered in a very short and easily understandable format, because participants cannot ask clarifying questions about the task as they might be able to in a cognitive lab setting.
+Our tasks are presented as a short series of questions within a larger survey on a variety of topics.
+Second, we want to utilize content within the stimulus which is not socially or politically charged for the average U.S.
+adult; this risks participants reacting to the subject matter within the chart rather than focusing on the task.
+For this reason, we utilize data on living arrangements of older U.S. adults -- a topic which most U.S. adults will have some familiarity with, but is not inherently polarizing.
+Finally, to prevent viewers from being exposed to slight variations of the same stimulus in a row (and risk unforeseen order effects or respondents using prior questions to inform their responses), we either split a survey sample in two and show each subsample a distinct structural version of the chart or test variations of a chart across distinct rounds of the survey.
+
 \ndt{
@@
 First, we vary the alignment of the pieces in question. \ndt{perhaps a bulleted list would work better here}
-Viewers are presented with two marked pieces in a chart that do not share a common baseline, then two pieces that do share a common baseline. \ndt{"we refer to this as aligned and unaligned"}
+Viewers are presented with two marked pieces in a chart that do not share a common baseline (aligned pieces), then two pieces that do share a common baseline (unaligned pieces). 
 Second, we vary the orientation of the chart -- we rotate the vertical stacked bar chart, and present a horizontally oriented version of the same chart, with identically sized marked pieces.
@@
 
-\ndt{"The survey format guides the design and structure of the questions asked...." Why is this so?}
-Our use of a survey format guides the format and design of the questions asked and how they are presented to respondents.
-First, participant instructions must be delivered in a very short and easily understandable format, because participants cannot ask clarifying questions about the task as they might be able to in a cognitive lab setting.
-Second, we want to utilize content within the data visualization image which is not socially or politically charged for the average U.S.
-adult; this risks participants reacting to the subject matter within the chart rather than focusing on the presented task.
-\ndt{stimulus?}
-For this reason, we utilize data on living arrangements of older U.S. adults -- a topic which most U.S. adults will have some familiarity with, but is not inherently polarizing.
-Finally, to prevent viewers from being exposed to slight variations of the same stimulus in a row (and risk unforeseen order effects or respondents using prior questions to inform their responses), we either split a survey sample in two and show each subsample a distinct structural version of the chart or test variations of a chart across distinct rounds of the survey.
-
-#### Question text  
-When viewing each chart, participants were asked to compare the relative sizes of marked elements within the chart:
-
-\ndt{This should go above fig 1 to explain the set up for fig 1}
-
-"There are many charts used in the news media to portray data visually. Looking at the chart below, which of the marked dark blue pieces is bigger, A or B? Just your best guess is fine.
-
--   A is bigger
-
--   B is bigger
-
--   They are the same
-
-When presented with the aligned version of each chart, pieces were marked with a C and D and the question text is updated accordingly.
-In all scenarios, the second option (B \[D\] is bigger) is the correct response.
-
-For a given task, viewers are first presented with the unaligned version of the task, followed by the aligned version of the task.
-The time in seconds that each respondent spent on each task was recorded, as well as whether the participant zoomed in on each chart while answering the question.
-Respondents were also asked to rate their certainty in their response to each question on a five-point scale.
-
 ```{r}
@@
 
-Task 1 was presented to viewers in two rounds (Rounds 1-2), with four color scheme variations each presented to 50% of respondents in each round. 
-There were no significant differences in respondent accuracy or behavior across color schemes, and we combine them here into one set of stimulus responses to focus on the comparison of structural differences.
-\kr{Should we put the color variations and a test for significance in accuracy/a visual in the appendix/supplementary materials?
-} \hh{we could do that, but I'm not sure that we need to mention color at all.}
+<!-- Task 1 was presented to viewers in two rounds (Rounds 1-2), with four color scheme variations each presented to 50% of respondents in each round.  -->
+<!-- There were no significant differences in respondent accuracy or behavior across color schemes, and we combine them here into one set of stimulus responses to focus on the comparison of structural differences.
+-->
+<!-- \kr{Should we put the color variations and a test for significance in accuracy/a visual in the appendix/supplementary materials?
+} \hh{we could do that, but I'm not sure that we need to mention color at all.} -->
 
@@
 
-The images utilized in the third task again represent the same data as in the first two tasks, and the overall image has dimensions which are identical to the images in the first two tasks, displaying at the same size during the survey.
-\hh{The length of bars is increased to fit the new dimensions.
+The images utilized in the third task, \autoref{fig-tasks}(c,f) again represent the same data as in the first two tasks, and the overall image has dimensions which are identical to the images in the first two tasks, displaying at the same size during the survey.
+The length of bars is increased to fit the new dimensions.
 This increases the difference in the length of bars from previously 8 pixels to 13 pixels.
 The widths of the tiles are adjusted accordingly (from 50 to 30 pixels) to keep the overall area of the tiles approximately constant.
-}
 <!--However, the aspect ratio of the plotting area is adjusted to increase the length in pixels of the stacked bar elements, and  decrease the relative thickness \kr{wording?} of each bar. -->
@@
 }-->
-\ndt{agree to drop note on color. Can perhaps just say we used CBlind  friendly}
-
-### Task 3b: Floating bars
-
-```{r}
-n5 <- nrow(round5_adj)
-n5_eff <- neff(round5_adj$weight)
-```
-
-\hh{
-In order to measure the impact of the chart's context, we asked a subset of participants to assess the relative sizes of the two tiles A and B (as shown in the bottom right of @fig-tasks-explain) without giving any additional information.}
+<!-- \ndt{agree to drop note on color. Can perhaps just say we used CBlind  friendly} -->
+
+<!-- ### Task 3b: Floating bars -->
+
+<!-- \hh{ -->
+<!-- In order to measure the impact of the chart's context, we asked a subset of participants to assess the relative sizes of the two tiles A and B (as shown in the bottom right of @fig-tasks-explain) without giving any additional information.}   -->
+
+#### Question text  
+When viewing each chart, participants were asked to compare the relative sizes of marked elements within the chart:
+
+\ndt{This should go above fig 1 to explain the set up for fig 1}
+
+"There are many charts used in the news media to portray data visually. Looking at the chart below, which of the marked dark blue pieces is bigger, A or B? Just your best guess is fine.
+
+-   A is bigger
+
+-   B is bigger
+
+-   They are the same
+
+When presented with the aligned version of each chart, pieces were marked with a C and D and the question text is updated accordingly.
+In all scenarios, the second option (B \[D\] is bigger) is the correct response.
+
+For a given task, viewers are first presented with the unaligned version of the task, followed by the aligned version of the task.
+The time in seconds that each respondent spent on each task was recorded, as well as whether the participant zoomed in on each chart while answering the question.
+Respondents were also asked to rate their certainty in their response to each question on a five-point scale.
 
@@
 The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households [@dennis2019technical].
-\ejm{46,565 panelists are CAWI; they are 93.3% of the panel and 93.6% of the total panel weights.}
 Each test was conducted using the AmeriSpeak Omnibus survey, which runs biweekly and samples around 1,000 U.S.
@@
 Our tests require visual inspection of an image; for this reason, our survey questions were only presented to web-based panelists and not panelists who respond via phone interviews.
+The panel has 45,565 web-based panelists, representing 93.3% of the total panel and 96.3% of the total panel weights.
 
@@
 
-Paragraph on how survey weights are derived?
--- **Question for Ed -- is this something AmeriSpeak can provide boilerplate language for?**
-
+The sample used for each Omnibus study is selected from the AmeriSpeak Panel using 48 sampling strata, split by age, race/Hispanic ethnicity, education, and gender, with the size of each sample per sampling stratum determined by the population distribution for each stratum.
+In addition, sample selection takes into account expected differential survey completion rates by demographic groups in order to achieve a representative sample of the target population.
+Collected data are weighted to the latest Current Population Survey (CPS) benchmarks from the U.S.
+Census Bureau.
+They are balanced by gender, age, education, race/ethnicity, and geographic region.
 All calculations in this paper are done in R [@RLanguage], and weights are applied in analyses using the `survey` package [@lumleyAnalysisComplexSurvey2004] version 4.0 [@survey] based on @lumleyComplexSurveysGuide2010.
 
-When combining responses from different surveys, weights \ejm{are rescaled} with respect to the total population, so that their order after combining still reflects their relative importance.
+When combining responses from different surveys, weights are rescaled with respect to the total population, so that their order after combining still reflects their relative importance.
 
@@
 | Task 3 | Horizontal wide |`r nrow(round3_adj[round3_adj$P_EXP == 2,])` | `r round(neff(round3_adj[round3_adj$P_EXP == 2,]$weight),1)` |
-| Task 3b | Floating bars |`r n5` | `r sprintf("%.1f",n5_eff)` |
 
@@
 
-logit_123_correct <- svyglm(correct~task_chart-1, family=quasibinomial(), design=survey_123b)
-logit_123_correct3 <- svyglm(correct3~task_chart-1, family=quasibinomial(), design=survey_123b)
-
-ci_confint <- svyby(~correct, by=~task_chart, design=survey_123b, svyciprop)
-ci_confint3 <- svyby(~correct3, by=~task_chart, design=survey_123b, svyciprop)
+logit_123_correct <- svyglm(correct~task_chart-1, family=quasibinomial(), design=survey_123)
+logit_123_correct3 <- svyglm(correct3~task_chart-1, family=quasibinomial(), design=survey_123)
+
+ci_confint <- svyby(~correct, by=~task_chart, design=survey_123, svyciprop)
+ci_confint3 <- svyby(~correct3, by=~task_chart, design=survey_123, svyciprop)
 
@@
   task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
-  chart = factor(chart, levels =c("Vertical", "Horizontal", "Horizontal wide", "Floating bars"))
+    chart = factor(chart, levels =c("Vertical", "Horizontal", "Horizontal wide"))
+
+  #chart = factor(chart, levels =c("Vertical", "Horizontal", "Horizontal wide", "Floating bars"))
 )
@@
   task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
-  chart = factor(chart, levels =c("Vertical", "Horizontal", "Horizontal wide", "Floating bars"))
+    chart = factor(chart, levels =c("Vertical", "Horizontal", "Horizontal wide"))
+  #chart = factor(chart, levels =c("Vertical", "Horizontal", "Horizontal wide", "Floating bars"))
 ) 
@@
 When tiles are unaligned, the horizontal wide barchart is showing the highest accuracy.
-For aligned tile, the horizontal wide design and the vertical design do not show a significant difference in accuracy.
-Floating bars have a significantly lower rate of correct answers than all the other designs."
+For aligned tile, the horizontal wide design and the vertical design do not show a significant difference in accuracy."
 #| fig-height: 4
@@
 
-round_123b_long %>% 
+#round_123b_long %>%
+round_123_long %>%
   mutate(
@@
     task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
-    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide", "Floating bars"))
+    chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide"))
+    #chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide", "Floating bars"))
   ) %>% 
@@
     )
-  ) %>% tab_footnote(
-    footnote = "We are including here, responses on an A vs B comparison for floating bars, corresponding to a length assessment, see bottom right of Figure 1.",
-    locations = cells_body(
-      rows = 1,
-      columns = "chart"
-    )
-  )
+  ) #%>% tab_footnote(
+  #   footnote = "We are including here, responses on an A vs B comparison for floating bars, corresponding to a length assessment, see bottom right of Figure 1.",
+  #   locations = cells_body(
+  #     rows = 1,
+  #     columns = "chart"
+  #   )
+  # )
 
@@
 
-Using linear scores for the response of `Certainty`, with `not certain at all` assigned a score of 1 and `extremely certain` assigned a score of 5, we can estimate the effects of task, chart design, and correctness on certainty  by using  a cell-means model of the form:
+Using linear scores for the response of Certainty, with 'not certain at all' assigned a score of 1 and 'extremely certain' assigned a score of 5, we can estimate the effects of task, chart design, and correctness on certainty  by using  a cell-means model of the form:
 $$
@@
 If a particular design leads viewers to zoom in more to investigate it when determining a response, we might consider that design to be more frustrating to viewers 'in the wild' when viewing it outside of the context of our tests.
-\kr{I don't love the last sentence but I think we can clean it up}.
 

### 2023-07-27 SDSS-2023/testing-charts-accuracy.qmd
a bunch of NDT comments addressed

@@
 bibliography: references.bib
-abstract: "The use of visuals is a key component in scientific communication, and decisions about the design of a data visualization should be informed by what design elements best support the audience's ability to perceive and understand the components of the data visualization.
+abstract: "The use of visuals is a key component in scientific communication.
+Decisions about the design of a data visualization should be informed by what design elements best support the audience's ability to perceive and understand the components of the data visualization.
 We build on the foundations of Cleveland and McGill's work in graphical perception, employing a large, nationally-representative, probability-based panel of survey respondents to test perception in statistical charts.
@@
 
-\ndt{break up first sentence}
-
 ## Introduction
@@
 <!-- What is our main objective? -->
+
+```{=html}
 <!--
@@
 -->
-
+```
 What do viewers see when we show them a data chart?
@@
 
-<!--
-\kr{The mackinlay thing feels like maybe we are overemphasizing the ranking done by cleveland/mcgill and then mackinlay - I think we could emphasize less if we want}
-\hh{my point was to emphasize that Mackinlay is a foundational paper (rightfully so), but not based on any data beyond Cleveland}
--->
-
 <!--Our work -- as that of Cleveland and McGill and Heer and Bostock -- -->
@@
 <!--Our work centers on studying data visualization design choices and their impact on viewer behavior and accuracy of viewers' responses.-->
+
 Here, we seek to answer whether it is possible to reproduce some of the previous findings in the context of a survey with a large, nationally-representative set of respondents.
@@
 <!-- # ```{r} -->
+
 <!-- # #| label: fig-perceptual-tasks -->
+
 <!-- # #| fig-cap: Ranking of perceptual tasks, as given by @mackinlayAutomatingDesignGraphical1986.
@@
 -->
+
 <!-- # knitr::include_graphics("images/mackinlay-tasks.png") -->
+
 <!-- # ``` -->
@@
 <!-- ```{r} -->
+
 <!-- #| label: fig-aligned-unaligned -->
+
 <!-- #| fig-cap: "The only difference between the two pairs of rectangles A, B and C, D is their alignment, i.e.
@@
 In contrast, more than half of the viewers respond with 'D is bigger' to the aligned pair of bars on the right." -->
+
 <!-- #| fig-subcap: -->
+
 <!-- #| - "**Unaligned bars**" -->
+
 <!-- #| - "**Aligned bars**" -->
+
 <!-- #| layout-nrow: 1 -->
+
 <!-- #| fig-align: "center" -->
+
 <!-- #| out-width: 50% -->
+
 <!-- knitr::include_graphics("images/round5-modified.PNG") -->
+
 <!-- knitr::include_graphics("images/round3-modified.PNG") -->
+
 <!-- ``` -->
@@
 
-\ndt{"The survey format guides the design and structure of the questions asked...." Why is this so?}
 Our use of a survey format guides the format and design of the questions asked and how they are presented to respondents.
@@
 
-\ndt{
-I think this paragraph should go first in this section.
-1. the survey format
-2. the type of data
-3. The stimulus
-3. how we set up the stimulus and why
-4. the tasks and alighed v unaligned
-4. qtext
-5. fig 1}
-
 Each task is made up of two elements: a visual stimulus and a question about the stimulus.
@@
 Rather than asking the same (or similar) type of question with varied signal strength dozens or hundreds of times, we are limited to only a few questions at a time.
-\ndt{note why we are limited to only a few questions at a time - that is not clear.}
 With a small set of tasks, we need to present tasks that are perceptually hard, and thus ask questions about stimuli that are close to our perceptual threshold.
@@
 
-We ask participants to determine which of two just-noticeably different marked pieces \ndt{is pieces the same as tiles below?} is larger within each chart, and focus on three main sets of structural variation in the design of that chart.
-First, we vary the alignment of the pieces in question. \ndt{perhaps a bulleted list would work better here}
+We ask participants to determine which of two just-noticeably different marked pieces (tiles) is larger within each chart, and focus on three main sets of structural variation in the design of that chart:
+
+-   First, we vary the alignment of the pieces in question.
     Viewers are presented with two marked pieces in a chart that do not share a common baseline (aligned pieces), then two pieces that do share a common baseline (unaligned pieces).
-Second, we vary the orientation of the chart -- we rotate the vertical stacked bar chart, and present a horizontally oriented version of the same chart, with identically sized marked pieces.
-Finally, we change the aspect ratio of the chart and present a wider version of the horizontally oriented chart which has longer, but thinner, marked pieces.
-
-```{=tex}
-\hh{Not sure where should put this, parking it here for right now:}
-\hh{expectations - start}
-```
+
+-   Second, we vary the orientation of the chart -- we rotate the vertical stacked bar chart, and present a horizontally oriented version of the same chart, with identically sized marked pieces.
+
+-   Finally, we change the aspect ratio of the chart and present a wider version of the horizontally oriented chart which has longer, but thinner, marked pieces.
+
+```{r}
+#| label: fig-tasks
+#| fig-cap: "Each of the visual stimuli presented to viewers. In each bar chart, two pieces are marked. The larger piece in each chart are pieces B and D. "
+#| out-width: 90%
+knitr::include_graphics("images/stimuli-overview.png")
+```
+
+### Task 1: Vertical stacked bar
+
+\autoref{fig-tasks}(a,d) shows the two stacked bar charts shown to participants in the first task.
+The marked tiles in each plot are 155 pixels apart, which leads to a JND of 3.5 pixels based on @luModelingJustNoticeable2022's model.
+The heights of the bars are 205 (left) and 213 pixels (right), corresponding to about twice the JND.
+This difference should lead to a relatively high accuracy rate for participants and simultaneously limit the amount of frustration resulting from a task that is perceived as 'too hard'.
+
+<!-- Lu et al calculation: log_{10}(JND) = -.4653 + .0065 distance -->
+
+<!-- describe stimulus -->
+
+Both charts show the same data with slight modifications to the order of the levels -- the first and second level in each of the bars are reversed between the left and the right chart.
+Both charts are displayed at the same size, i.e. in both cases both the difference in size between the bars and the horizontal distance between the bars is the exact same amount.
+This leaves the vertical positioning of the bars as the only difference between the charts.
+Any differences in observed responses can therefore be attributed to this difference in presentation.
+
+<!-- Task 1 was presented to viewers in two rounds (Rounds 1-2), with four color scheme variations each presented to 50% of respondents in each round.  -->
+
+<!-- There were no significant differences in respondent accuracy or behavior across color schemes, and we combine them here into one set of stimulus responses to focus on the comparison of structural differences.
+-->
+
+<!-- \kr{Should we put the color variations and a test for significance in accuracy/a visual in the appendix/supplementary materials?
+} \hh{we could do that, but I'm not sure that we need to mention color at all.} -->
+
+### Task 2: Horizontal stacked bar
+
+The visual mappings in the second task, shown in \autoref{fig-tasks}(b, e), are identical to the first task, but the axes of the chart are rotated so that the stacked bars are represented in a horizontal format.
+This represents a structural change in how the data are presented to the viewer while preserving the pixel size of the elements viewers must compare.
+Tasks 2 and 3 were asked within the same survey round (Round 3) and the full sample was randomly split among respondents, with 50% of participants seeing the Task 2 stimuli and 50% of participants seeing the Task 3 stimuli.
+
+### Task 3: Horizontal wide stacked bar
+
+The images utilized in the third task, \autoref{fig-tasks}(c,f), again represent the same data as in the first two tasks, and the overall image has dimensions which are identical to the images in the first two tasks, displaying at the same size during the survey.
+The length of bars is increased to fit the new dimensions.
+This increases the difference in the length of bars from previously 8 pixels to 13 pixels.
+The widths of the tiles are adjusted accordingly (from 50 to 30 pixels) to keep the overall area of the tiles approximately constant.
+<!--However, the aspect ratio of the plotting area is adjusted to increase the length in pixels of the stacked bar elements, and  decrease the relative thickness \kr{wording?} of each bar.
+--> <!--kr{Do we have a measurement of the pixels for this?
+I think a sentence on that would close out this section nicely.}\hh{like that?}\kr{I was thinking the length pixels, since we talk about how many they are in Task 1, so we could talk about how many JND the difference is.
+I like the height thing though too, so maybe both?
+}--> <!-- \ndt{agree to drop note on color.
+Can perhaps just say we used CBlind  friendly} -->
+
+<!-- ### Task 3b: Floating bars -->
+
+<!-- \hh{ -->
+
+<!-- In order to measure the impact of the chart's context, we asked a subset of participants to assess the relative sizes of the two tiles A and B (as shown in the bottom right of @fig-tasks-explain) without giving any additional information.}   -->
+
+#### Question text
+
+When viewing each chart, participants were asked to compare the relative sizes of marked elements within the chart:
+
+"There are many charts used in the news media to portray data visually. Looking at the chart below, which of the marked dark blue pieces is bigger, A or B? Just your best guess is fine.
+
+-   A is bigger
+
+-   B is bigger
+
+-   They are the same
+
+When presented with the aligned version of each chart, pieces were marked with a C and D and the question text is updated accordingly.
+In all scenarios, the second option (B \[D\] is bigger) is the correct response.
+
+For a given task, viewers are first presented with the unaligned version of the task, followed by the aligned version of the task.
+The time in seconds that each respondent spent on each task was recorded, as well as whether the participant zoomed in on each chart while answering the question.
+Respondents were also asked to rate their certainty in their response to each question on a five-point scale.
+
+```{r}
+#| label: fig-tasks-explain
+#| fig-cap: Comparisons made in charts within the Cleveland and McGill ranking (left) and their corresponding representations in our study tasks (right).
+#| out-width: 60%
+
+include_graphics("images/sketch-tasks.png")
+```
+
+#### Alignment
+
 What we call 'aligned' and 'unaligned', here, is similar to Cleveland and McGill's positions along aligned and unaligned scales, but with some modifications.
 @fig-tasks-explain gives an overview of the comparisons of tasks 1 through 3 and the closest corresponding tasks in Cleveland and McGill.
-
 In this paper, both 'aligned' and 'unaligned' bars share the same axis.
@@
 We would expect that comparing unaligned tiles is a harder task (with correspondingly lower levels of accuracy) than a comparison of aligned tiles, with the framing given by the other tiles in the same column mitigating some of this difficulty.
-\hh{In particular, the additional context of the other tiles should *help* rather than *hurt* the comparison, unlike other situations, where additional information besides the pieces involved in the comparison were found to be distractors} [@talbotFourExperimentsPerception2014].
-<!--\kr{The updated sketch is great here!! One note -- can we move the label for the "positions along the same, but unaligned scale", it's not aligned with the others}-->
-<!--\hh{What about the sketch now?}-->
-
+In particular, the additional context of the other tiles should *help* rather than *hurt* the comparison, unlike other situations, where additional information besides the pieces involved in the comparison were found to be distractors [@talbotFourExperimentsPerception2014].
+<!--\kr{The updated sketch is great here!!
+One note -- can we move the label for the "positions along the same, but unaligned scale", it's not aligned with the others}--> <!--\hh{What about the sketch now?}-->
+
+```{=html}
 <!--
@@
 -->
-
-```{r}
-#| label: fig-tasks-explain
-#| fig-cap: Comparisons made in charts within the Cleveland and McGill ranking 
-#| out-width: 60%
-
-include_graphics("images/sketch-tasks.png")
-```
-
-```{r}
-#| label: fig-tasks
-#| fig-cap: "Each of the visual stimuli presented to viewers. In each bar chart, two pieces are marked. The larger piece in each chart are pieces B and D. "
-#| out-width: 90%
-knitr::include_graphics("images/stimuli-overview.png")
-```
-
-### Task 1: Vertical stacked bar
-
-\autoref{fig-tasks}(a,d) shows the two stacked bar charts shown to participants in the first task.
-The marked tiles in each plot are 155 pixels apart, which leads to a JND of 3.5 pixels based on @luModelingJustNoticeable2022's model.
-The heights of the bars are 205 (left) and 213 pixels (right), corresponding to about twice the JND.
-This difference should lead to a relatively high accuracy rate for participants and simultaneously limit the amount of frustration resulting from a task that is perceived as 'too hard'.
-
-<!-- Lu et al calculation: log_{10}(JND) = -.4653 + .0065 distance -->
-
-<!-- describe stimulus -->
-
-Both charts show the same data with slight modifications to the order of the levels -- the first and second level in each of the bars are reversed between the left and the right chart.
-Both charts are displayed at the same size, i.e. in both cases both the difference in size between the bars and the horizontal distance between the bars is the exact same amount.
-This leaves the vertical positioning of the bars as the only difference between the charts.
-Any differences in observed responses can therefore be attributed to this difference in presentation.
-
-<!-- Task 1 was presented to viewers in two rounds (Rounds 1-2), with four color scheme variations each presented to 50% of respondents in each round.  -->
-<!-- There were no significant differences in respondent accuracy or behavior across color schemes, and we combine them here into one set of stimulus responses to focus on the comparison of structural differences.
--->
-<!-- \kr{Should we put the color variations and a test for significance in accuracy/a visual in the appendix/supplementary materials?
-} \hh{we could do that, but I'm not sure that we need to mention color at all.} -->
-
-### Task 2: Horizontal stacked bar
-
-The visual mappings in the second task, shown in \autoref{fig-tasks}(b, e), are identical to the first task, but the axes of the chart are rotated so that the stacked bars are represented in a horizontal format.
-This represents a structural change in how the data are presented to the viewer while preserving the pixel size of the elements viewers must compare.
-Tasks 2 and 3 were asked within the same survey round (Round 3) and the full sample was randomly split among respondents, with 50% of participants seeing the Task 2 stimuli and 50% of participants seeing the Task 3 stimuli.
-
-### Task 3: Horizontal wide stacked bar
-
-The images utilized in the third task, \autoref{fig-tasks}(c,f) again represent the same data as in the first two tasks, and the overall image has dimensions which are identical to the images in the first two tasks, displaying at the same size during the survey.
-The length of bars is increased to fit the new dimensions.
-This increases the difference in the length of bars from previously 8 pixels to 13 pixels.
-The widths of the tiles are adjusted accordingly (from 50 to 30 pixels) to keep the overall area of the tiles approximately constant.
-<!--However, the aspect ratio of the plotting area is adjusted to increase the length in pixels of the stacked bar elements, and  decrease the relative thickness \kr{wording?} of each bar. -->
-<!--kr{Do we have a measurement of the pixels for this?
-I think a sentence on that would close out this section nicely.}\hh{like that?}\kr{I was thinking the length pixels, since we talk about how many they are in Task 1, so we could talk about how many JND the difference is.
-I like the height thing though too, so maybe both?
-}-->
-<!-- \ndt{agree to drop note on color. Can perhaps just say we used CBlind  friendly} -->
-
-<!-- ### Task 3b: Floating bars -->
-
-<!-- \hh{ -->
-<!-- In order to measure the impact of the chart's context, we asked a subset of participants to assess the relative sizes of the two tiles A and B (as shown in the bottom right of @fig-tasks-explain) without giving any additional information.}   -->
-
-#### Question text  
-When viewing each chart, participants were asked to compare the relative sizes of marked elements within the chart:
-
-\ndt{This should go above fig 1 to explain the set up for fig 1}
-
-"There are many charts used in the news media to portray data visually. Looking at the chart below, which of the marked dark blue pieces is bigger, A or B? Just your best guess is fine.
-
--   A is bigger
-
--   B is bigger
-
--   They are the same
-
-When presented with the aligned version of each chart, pieces were marked with a C and D and the question text is updated accordingly.
-In all scenarios, the second option (B \[D\] is bigger) is the correct response.
-
-For a given task, viewers are first presented with the unaligned version of the task, followed by the aligned version of the task.
-The time in seconds that each respondent spent on each task was recorded, as well as whether the participant zoomed in on each chart while answering the question.
-Respondents were also asked to rate their certainty in their response to each question on a five-point scale.
+```
 
@@
 
-Participants were recruited as part of NORC's AmeriSpeak panel, which utilizes a probability-based sampling methodology and samples U.S.
+Participants were recruited as part of NORC's AmeriSpeak panel^[NORC is a non-partisan organization that collects data for government and non-government clients], which utilizes a probability-based sampling methodology and samples U.S.
 households from NORC's National Sample Frame that provides coverage of over 97% of U.S.
@@
 The current panel size is 54,001 panel members aged 13 and over residing in over 43,000 households [@dennis2019technical].
-Each test was conducted using the AmeriSpeak Omnibus survey, which runs biweekly and samples around 1,000 U.S.
-adults to answer questions on a variety of topics.
+Each test was conducted using the AmeriSpeak Omnibus survey, which runs biweekly and samples around 1,000 U.S. adults to answer questions on a variety of topics.
 Our tests require visual inspection of an image; for this reason, our survey questions were only presented to web-based panelists and not panelists who respond via phone interviews.
@@
 
-The Omnibus sample is not longitudinal in nature, i.e.\ we do not have any information about whether the same panelists were included in multiple rounds.
+The Omnibus sample is not longitudinal in nature, i.e. we do not have any information about whether the same panelists were included in multiple rounds.
 While there is a (small) chance that panelists are included in multiple rounds of our data, there was at least a one month gap between viewing each visual stimulus, and for the purposes of the analysis we will consider data from these viewings as independent.
-
-<!-- Given the nature of pulling a sample from the panel for each Omnibus round, there is a possibility that some participants may have been included in multiple rounds; however, our collected data is not a longitudinal or panel study, so we do not have repeated responses from the same participants across each of our tests.
-Each of the tests was run at least a month apart, so if respondents did participate in multiple rounds, there was at least a one month gap between viewing each visual stimulus.
--->
 
@@
 
-<!--
-\kr{Do we need to update this description to just say how we did it for combining across all the rounds represented?}\hh{done} -->
-
 We *combine* (rather than cumulate) a set of $\ell$ surveys $S_1$, $S_2$, ...
-$S_\ell$ (with $ \ell \ge 2$ ) as described in @omuircheartaighCombiningSamplesVs2002, by multiplying weights in $S_i$ by $\lambda_i$, for $1 \le i \le \ell$.
+$S_\ell$ (with \$ \ell \ge 2\$ ) as described in @omuircheartaighCombiningSamplesVs2002, by multiplying weights in $S_i$ by $\lambda_i$, for $1 \le i \le \ell$.
 $\lambda_i \in [0,1]$ with $\sum_i \lambda_i = 1$ is given as 
@@
 \lambda_i = \frac{n_i/d_i}{\sum_{j=1}^{\ell}n_j/d_j},
-$$ where $n_i$ is the nominal sample of survey $S_i$ and $d_i$ are the design effects for the estimators.
+$$ 
+where $n_i$ is the nominal sample of survey $S_i$ and $d_i$ are the design effects for the estimators.
 Here, $d_i$ are estimated as 
@@
 
+```{=html}
 <!--Note that @omuircheartaighCombiningSamplesVs2002 estimate $\lambda$ separately for any combination of race/ethnicity by sex.
 We employ that strategy whenever we include demographic variables in the analysis, otherwise we use a single adjustment for the weights. -->
-
+```
 The data for this paper were collected in several rounds as part of the NORC Omnibus.
@@
 | Name      | Date       |      \# Participants |                Effective Sample Size |         Sum of weights $\sum_i w_i$ |                                  $\lambda_i$ |
-|---------------|---------------|--------------:|--------------:|--------------:|--------------:|
+|------------|------------|-----------:|-----------:|-----------:|-----------:|
 | Round 1   | April 2022 | `r nrow(round1_adj)` | `r round(neff(round1_adj$weight),1)` | `r round(sum(round1_adj$weight),1)` |        $\lambda_1$ = `r round(lambdas[1],3)` |
@@
 | Name   | Description     |                              \# Participants |                                        Effective Sample Size |
-|---------------|---------------|--------------:|--------------:|
-| Task 1 | Vertical| `r nrow(round1_adj) + nrow(round2_adj)` | `r round(neff(round_12$weight),1)` |
-| Task 2 | Horizontal |`r nrow(round3_adj[round3_adj$P_EXP == 1,])` | `r round(neff(round3_adj[round3_adj$P_EXP == 1,]$weight),1)` |
-| Task 3 | Horizontal wide |`r nrow(round3_adj[round3_adj$P_EXP == 2,])` | `r round(neff(round3_adj[round3_adj$P_EXP == 2,]$weight),1)` |
+|------------------|------------------|-----------------:|-----------------:|
+| Task 1 | Vertical        |      `r nrow(round1_adj) + nrow(round2_adj)` |                           `r round(neff(round_12$weight),1)` |
+| Task 2 | Horizontal      | `r nrow(round3_adj[round3_adj$P_EXP == 1,])` | `r round(neff(round3_adj[round3_adj$P_EXP == 1,]$weight),1)` |
+| Task 3 | Horizontal wide | `r nrow(round3_adj[round3_adj$P_EXP == 2,])` | `r round(neff(round3_adj[round3_adj$P_EXP == 2,]$weight),1)` |
 
@@
   (age + income + plot_layout(nrow=1, widths = c(4,4)))
+
 ```
@@
 However, we are interested in understanding whether viewers *can* perceive the difference and correctly identify which piece is larger and thus consider multiple ways of modeling response to investigate participant response patterns.
-We begin by defining a measure of binary 'correctness', for which all answers that are not the correct option (B [D] is bigger) are 'incorrect', including the selection of 'they are the same'.
+We begin by defining a measure of binary 'correctness', for which all answers that are not the correct option (B \[D\] is bigger) are 'incorrect', including the selection of 'they are the same'.
 
@@
 
-Next, we consider an ordinal model to investigate response behavior across all three options -- 'A [C] is bigger', 'B [D] is bigger', and 'They are the same', and consider these response patterns across each of the stimuli.
+Next, we consider an ordinal model to investigate response behavior across all three options -- 'A \[C\] is bigger', 'B \[D\] is bigger', and 'They are the same', and consider these response patterns across each of the stimuli.
 
 ```{r model-123, warning = FALSE, message = FALSE}
-
 logit_123_correct <- svyglm(correct~task_chart-1, family=quasibinomial(), design=survey_123)
@@
 )
+
 ```
@@
 #| label: fig-response-123
-#| fig-cap: "Responses for accuracy by task across the designs.
+#| fig-cap: "Responses by task across the designs.
 The overlaid rectangles represent 95% confidence intervals.
 The letters in blue and orange encode significances between pairwise proportions: two bars have a significantly different proportion (at a 5% significance level) if they do not share a letter.
-There is no significant difference between the designs for wrong responses.
-When tiles are unaligned, the horizontal wide barchart is showing the highest accuracy.
-For aligned tile, the horizontal wide design and the vertical design do not show a significant difference in accuracy."
+There is no significant difference between the designs for wrong responses."
 #| fig-height: 4
@@
 The re-scaled design of the wide horizontal bars reclaims some of the loss for unaligned bars and outperforms the vertical design by a similar margin in aligned bars, but does not out-perform the vertical design when comparing unaligned tiles.
-\hh{The beneficial effect of frames/context of the chart is apparent when comparing accuracy of responses for the Horizontal wide design and the floating bars: the rate of correct responses in floating bars is significantly lower than in the wide horizontal charts.
-This shows that the additional context }
+<!-- \hh{The beneficial effect of frames/context of the chart is apparent when comparing accuracy of responses for the Horizontal wide design and the floating bars: the rate of correct responses in floating bars is significantly lower than in the wide horizontal charts.
+This shows that the additional context } -->
 
@@
 #| label: fig-zoom-123
-#| fig-cap: "Zooming - not significant for accuracy or task, but changes by the type of chart."
+#| fig-cap: "Rates of zooming by task, chart design, and correctness.
+There are significantly higher rates of zooming by respondents for the vertical task compared to both horizontal tasks; rates of zooming do not differ by correctness of response or alignment in the stimulus."
 #| fig-height: 4.25
@@
 
-To formalize this pattern in a model, let $Y_{jk}$ describe the zooming behavior of panelist $k$ on task $j$.
-We model zooming behavior (no = 0, yes = 1) as a logistic regression by correctness of response ($\rho$), task ($\tau$), and design ($\delta$) of the chart:
-
-$$
-\text{logit } P(Y_k \le 1) = \mu + \rho_{i(k)} + \tau_{j(k)} + \delta_{\ell(k)},
-$$
-
-The resulting model estimates and confidence intervals are displayed in \autoref{tbl-zoom-123}.
-We observe that zooming behavior *only* differs significantly by structural design; rates do not differ significantly between the aligned and unaligned tasks, nor do they differ significantly by the chosen response.
-This higher rate of zooming does not necessarily lead to higher accuracy; although rates of accuracy are significantly higher for the vertical orientation relative to the horizontal orientation, they are not significantly higher -- and in fact, are significantly lower on the aligned task -- than the accuracy for the horizontal wide orientation.
-\hh{I'll pretty up @tbl-zoom-123 if we are going to keep it.} \kr{I think we should!
-it's nice to have a model for this}\hh{gt and rmarkdown are frustrating!
-I'll re-set the table in its final form directly in tex}
-
 ```{r}
@@
 
-### How certain are participants?
-
-<!--
-\kr{Do we have anything about total time spent on tasks across the 3 structures? We have time on each question for rounds 2-3. And can we pull in some of the certainty stuff?}
-
-\hh{we could include certainty - I'll look into it}
--->
+To formalize this pattern in a model, let $Y_{jk}$ describe the zooming behavior of panelist $k$ on task $j$.
+We model zooming behavior (no = 0, yes = 1) as a logistic regression by correctness of response ($\rho$), task ($\tau$), and design ($\delta$) of the chart:
+
+$$
+\text{logit } P(Y_k \le 1) = \mu + \rho_{i(k)} + \tau_{j(k)} + \delta_{\ell(k)},
+$$
+
+The resulting model estimates and confidence intervals are displayed in \autoref{tbl-zoom-123}.
+We observe that zooming behavior *only* differs significantly by structural design; rates do not differ significantly between the aligned and unaligned tasks, nor do they differ significantly by the chosen response.
+This higher rate of zooming does not necessarily lead to higher accuracy; although rates of accuracy are significantly higher for the vertical orientation relative to the horizontal orientation, they are not significantly higher -- and in fact, are significantly lower on the aligned task -- than the accuracy for the horizontal wide orientation.
+\hh{gt and rmarkdown are frustrating! I'll re-set the table in its final form directly in tex}
 
@@
 where $k = 1, ..., N$, $\mu_{ij\ell(k)}$ is average certainty (measured on a scale from 1 to 5) of the four combinations of task and correctness by each of the three designs, where $i = 1, 2$ encodes unaligned/aligned, and $j=1,2$ encodes wrong, correct, and $\ell = 1, 2, 3$ encodes horizontal, wide horizontal, and vertical stacked bar charts, respectively.
-We also assume that errors are normally distributed, i.e.
-$\epsilon_k \stackrel{i.i.d}{\sim} N(0, \sigma^2)$ for all $k = 1, ..., N$.
+We also assume that errors are normally distributed, i.e. $\epsilon_k \stackrel{i.i.d}{\sim} N(0, \sigma^2)$ for all $k = 1, ..., N$.
 The results are shown in @fig-certainty.
@@
 $$
+
 The resulting model estimates and confidence intervals are shown in \autoref{tbl-demographics}.
@@
 However, income levels do have a (small) effect.
-When income levels increase, the percentage of wrong answers (orange) decreases, while the percentage of correct answers (blue) increases slightly (significant at below 0.05).
+When income levels increase, the percentage of wrong answers decreases, while the percentage of correct answers increases slightly (significant at below 0.05).
 For the  more difficult task of comparing unaligned tiles, demographics are more significant.
@@
 <!-- There are differences by alignment -->
+
 <!-- There are not huge differences across structures -->
+
 <!-- However, we see a difference in how people react across different structures (zooming behavior, certainty? time?) -->
@@
 <!-- - We can reproduce prior findings with nat rep sample  -->
+
 <!-- - We can rank  -->
+
 <!-- - We can rank better because we have a bigger sample size  -->
+
 <!-- - We can find changes in viewer behavior  -->
+
 <!-- - Demographics matter to accuracy and choices -->
@@
 
+Our study results provide actionable insights that data visualization practitioners can utilize in the design of data visualizations.
+Comparisons of interest should be lined up and share a common baseline within a chart; in an interactive setting, it may be beneficial to give viewers the possibility of re-ordering categories to support effective comparison across groupings.
+Allowing users to zoom in on images may not impact perceptual accuracy, but it leads to higher certainty in viewers' comparisons.
+Considering the target audience for visualizations and what key findings are most important to convey should be an important step in the process of designing a data visualization.
+
 While our results are limited to mapping in a stacked bar chart with a specific topic, they demonstrate among them a wealth of insights about how respondents perceive and interact with data visualizations.
@@
 
-\hh{The actionable insights are:}
-
-- make sure that the main comparison in a plot is lined up -- in an interactive setting it would be beneficial to have the possibility of re-ordering the categories to help with comparisons. 
-- offering zoom-ins help with certainty, even if it does not help with perceptual accuracy
-
 <!-- Key findings to discuss: -->
@@
 <!--     -   Cleveland: position (aligned) versus unaligned -->
+
 <!--     -   unaligned tiles within stacked bar are a hybrid between framed bar and floats -->
+
 <!--     -   Talbot: pie charts - individual wedges vs piecharts -->

### 2023-07-27 SDSS-2023/testing-charts-accuracy.tex
a bunch of NDT comments addressed

@@
 \usepackage{amsmath,amssymb}
+\usepackage{lmodern}
 \usepackage{iftex}
@@
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-\usepackage{lmodern}
-\ifPDFTeX\else  
-    % xetex/luatex font selection
 \fi
@@
 \makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
+\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
 \makeatother
@@
 \@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
-\makeatother
-\makeatletter
 \makeatother
@@
 \begin{abstract}
-The use of visuals is a key component in scientific communication, and
-decisions about the design of a data visualization should be informed by
+The use of visuals is a key component in scientific communication.
+Decisions about the design of a data visualization should be informed by
 what design elements best support the audience's ability to perceive and
@@
 \end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, breakable, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, breakable, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 
-\ndt{break up first sentence}
-
 \hypertarget{introduction}{%
@@
 represented as measurable features, such as position, size, or shade.
-\hh{do we need to also go into qualitative variables?} Modern data
-visualizations are much more than a simple, objective mapping of values
-to a plane; they contain contextual and design elements, and are often
-structured to support the viewer in understanding a particular view of a
-set of data or specific pattern underlying the values. The design of a
-data visualization impacts a viewer's ability to achieve that
-understanding; a poorly designed data visualization may leave viewers
-struggling to understand the content or context, or make it difficult to
-complete accurate and useful comparisons of values across groups or time
-points. More broadly, the design of a data visualization can change how
-viewers interact with the chart.
-
-\ndt{define structural}
-
-\ndt{The viewer's employment of comparisons of the components is a crucial step in the process of interacting with and understanding of chart.}
-
+Modern data visualizations are much more than a simple, objective
+mapping of values to a plane; they contain contextual and design
+elements, and are often structured to support the viewer in
+understanding a particular view of a set of data or specific pattern
+underlying the values. Structural choices, such as the choice of a
+graphical element (e.g., bar or pie), orientation of the chart, and
+sizing and relative positioning of elements within the chart, are
+determined by the data visualization practitioner. The design of a data
+visualization impacts a viewer's ability to achieve that understanding;
+a poorly designed data visualization may leave viewers struggling to
+understand the content or context, or make it difficult to complete
+accurate and useful comparisons of values across groups or time points.
+More broadly, the design of a data visualization can change how viewers
+interact with the chart.
+
+The viewer's employment of comparisons of the components is a crucial
+step in the process of interacting with and understanding of chart.
 Cleveland and McGill (1984) observed as such, and in their seminal study
@@
 make more accurate comparisons. Based on mappings of quantitative
-variables to different graphical elements
-\ndt{example of element to explain it XXX}, Cleveland and McGill's study
-resulted in a ranking of perceptual tasks \ndt{give example of task}
-from most accurate to least accurate, which was then extended by
-Mackinlay (1986) to a theoretical framework ranking tasks' order along
-their ordinal and nominal scales.
+variables to different graphical elements, Cleveland and McGill's study
+resulted in a ranking of perceptual tasks (such as ranking the larger of
+two lengths or areas) from most accurate to least accurate, which was
+then extended by Mackinlay (1986) to a theoretical framework ranking
+tasks' order along their ordinal and nominal scales.
 
@@
 previous findings in the context of a survey with a large,
-nationally-representative set of respondents, and within that context we
-focus on the following research questions:
+nationally-representative set of respondents. Specifically, we present
+viewers with structural variations of bar charts and ask them to answer
+questions comparing the size of elements within those charts. We employ
+a probability-based survey panel and run a series of perception tests
+with nationally-representative samples of respondents from that panel.
+The advantage of using a probability-based approach is two-fold. First,
+we have access to a large sample of survey participants and thus have
+greater power in making inference about graphical perceptional
+abilities. Second, the sample is representative of the general adult
+public in the U.S., and this allows us to test whether prior results
+from convenience samples hold with a nationally representative sample
+and whether there are differences in those results across demographic
+subgroups. Within that context we focus on the following research
+questions:
 
@@
 
-We employ a probability-based survey panel and run a series of
-perception tests with nationally-representative samples of respondents
-from that panel. The advantage of using a probability-based approach is
-two-fold. First, we have access to a large sample of survey participants
-and thus have greater power in making inference about graphical
-perceptional abilities. Second, the sample is representative of the
-general adult public in the U.S., and this allows us to test whether
-prior results from convenience samples hold with a nationally
-representative sample and whether there are differences in those results
-across demographic subgroups.
-
-\ndt{I think the next paragraph can be moved above the RQs. Use this paragraph to set up the concepts, define constructs, then explain the study, and then note the RQs.}
-
-Specifically, we present viewers with structural variations of bar
-charts and ask them to answer questions comparing the size of elements
-within those charts. In this work, we present the series of tests we
-completed and the resulting findings. The remainder of the paper is
-organized as follows: first, we describe the design of the visual
-stimuli used in the perception tests. We then describe the population of
-study respondents and obtained survey sample. Subsequently, we share a
-summary of the resulting survey responses across each of the tests,
-including analyses on accuracy of responses and response behavior.
-Finally, we discuss implications of this work and next steps.
+In this work, we present the series of tests we completed and the
+resulting findings. The remainder of the paper is organized as follows:
+first, we describe the design of the visual stimuli used in the
+perception tests. We then describe the population of study respondents
+and obtained survey sample. Subsequently, we share a summary of the
+resulting survey responses across each of the tests, including analyses
+on accuracy of responses and response behavior. Finally, we discuss
+implications of this work and next steps.
 
@@
 \subsection{Study design -- stimulus}\label{study-design-stimulus}}
-
-\ndt{
-I think this paragraph should go first in this section.
-1. the survey format
-2. the type of data
-3. The stimulus
-3. how we set up the stimulus and why
-4. the tasks and alighed v unaligned
-4. qtext
-5. fig 1}
-
-Each task is made up of two elements: a visual stimulus and a question
-about the stimulus. In our study, each visual stimulus is an image of a
-data visualization, while each question prompts viewers to identify
-which of two marked pieces in the data visualization is larger.
-
-The comparison between sets of marked pairs is intentionally designed to
-be a difficult task, with the difference between the values represented
-in the two marked pieces being chosen close to their just-noticeable
-difference. The \textbf{Just-Noticeable Difference} (JND) is defined as
-the smallest difference that will be detected 50\% of the time. Prior
-results from studies on bar charts and pie charts (Lu et al. 2022)
-inform the differences in charts shown to survey panelists.
-
-We employ comparisons at the JND in our tasks in order to maximize our
-ability to identify the impact of design changes on viewer accuracy and
-behavior. Asking perception tasks in a survey differs from the
-controlled environment of a cognitive lab, where these kind of questions
-may usually be assessed. Rather than asking the same (or similar) type
-of question with varied signal strength dozens or hundreds of times, we
-are limited to only a few questions at a time.
-\ndt{note why we are limited to only a few questions at a time - that is not clear.}
-With a small set of tasks, we need to present tasks that are
-perceptually hard, and thus ask questions about stimuli that are close
-to our perceptual threshold. Therefore, we focus on questions which vary
-the structure of the presented image, but ask viewers to compare the
-same underlying data values across those varied images.
-
-We ask participants to determine which of two just-noticeably different
-marked pieces \ndt{is pieces the same as tiles below?} is larger within
-each chart, and focus on three main sets of structural variation in the
-design of that chart. First, we vary the alignment of the pieces in
-question. \ndt{perhaps a bulleted list would work better here} Viewers
-are presented with two marked pieces in a chart that do not share a
-common baseline, then two pieces that do share a common baseline.
-\ndt{"we refer to this as aligned and unaligned"} Second, we vary the
-orientation of the chart -- we rotate the vertical stacked bar chart,
-and present a horizontally oriented version of the same chart, with
-identically sized marked pieces. Finally, we change the aspect ratio of
-the chart and present a wider version of the horizontally oriented chart
-which has longer, but thinner, marked pieces.
-
-\hh{Not sure where should put this, parking it here for right now:}
-\hh{expectations - start}
-
-What we call `aligned' and `unaligned', here, is similar to Cleveland
-and McGill's positions along aligned and unaligned scales, but with some
-modifications. Figure~\ref{fig-tasks-explain} gives an overview of the
-comparisons of tasks 1 through 3 and the closest corresponding tasks in
-Cleveland and McGill.
-
-In this paper, both `aligned' and `unaligned' bars share the same axis.
-
-Aligned tiles are additionally anchored in the same position along the
-axis, i.e.~the difference between their sizes can be reduced to a
-positional assessment, which in the absence of three-dimensional cues
-seems to be the general strategy viewers use when evaluating barcharts
-(Zacks et al. 1998).\\
-Unaligned tiles do not share this anchor of position, i.e.~the
-difference between their sizes needs at least two comparisons of
-positions. However, unaligned bars in a stacked barchart are not just
-assessments of the lengths of bars -- the other tiles in the chart
-provide a reference frame of the shared axis, which \emph{should} help
-with an assessment of the tiles' sizes. As shown in the sketch of
-Figure~\ref{fig-tasks-explain}, this framing leads to a stimulus that is
-between Cleveland and McGill's task of `position along the same but
-unaligned axis' and an assessment of length.
-
-We would expect that comparing unaligned tiles is a harder task (with
-correspondingly lower levels of accuracy) than a comparison of aligned
-tiles, with the framing given by the other tiles in the same column
-mitigating some of this difficulty.
-\hh{In particular, the additional context of the other tiles should *help* rather than *hurt* the comparison, unlike other situations, where additional information besides the pieces involved in the comparison were found to be distractors}
-(Talbot, Setlur, and Anand 2014).
-
-\begin{figure*}[hbt]
-
-{\centering \includegraphics[width=0.6\textwidth,height=\textheight]{images/sketch-tasks.png}
-
-}
-
-\caption{\label{fig-tasks-explain}Comparisons made in charts within the
-Cleveland and McGill ranking}
-
-\end{figure*}
-
-\ndt{"The survey format guides the design and structure of the questions asked...." Why is this so?}
 
@@
 format, because participants cannot ask clarifying questions about the
-task as they might be able to in a cognitive lab setting. Second, we
-want to utilize content within the data visualization image which is not
-socially or politically charged for the average U.S. adult; this risks
-participants reacting to the subject matter within the chart rather than
-focusing on the presented task. \ndt{stimulus?} For this reason, we
+task as they might be able to in a cognitive lab setting. Our tasks are
+presented as a short series of questions within a larger survey on a
+variety of topics. Second, we want to utilize content within the
+stimulus which is not socially or politically charged for the average
+U.S. adult; this risks participants reacting to the subject matter
+within the chart rather than focusing on the task. For this reason, we
 utilize data on living arrangements of older U.S. adults -- a topic
@@
 
-\hypertarget{question-text}{%
-\paragraph{Question text}\label{question-text}}
-
-When viewing each chart, participants were asked to compare the relative
-sizes of marked elements within the chart:
-
-\ndt{This should go above fig 1 to explain the set up for fig 1}
-
-``There are many charts used in the news media to portray data visually.
-Looking at the chart below, which of the marked dark blue pieces is
-bigger, A or B? Just your best guess is fine.
+Each task is made up of two elements: a visual stimulus and a question
+about the stimulus. In our study, each visual stimulus is an image of a
+data visualization, while each question prompts viewers to identify
+which of two marked pieces in the data visualization is larger.
+
+The comparison between sets of marked pairs is intentionally designed to
+be a difficult task, with the difference between the values represented
+in the two marked pieces being chosen close to their just-noticeable
+difference. The \textbf{Just-Noticeable Difference} (JND) is defined as
+the smallest difference that will be detected 50\% of the time. Prior
+results from studies on bar charts and pie charts (Lu et al. 2022)
+inform the differences in charts shown to survey panelists.
+
+We employ comparisons at the JND in our tasks in order to maximize our
+ability to identify the impact of design changes on viewer accuracy and
+behavior. Asking perception tasks in a survey differs from the
+controlled environment of a cognitive lab, where these kind of questions
+may usually be assessed. Rather than asking the same (or similar) type
+of question with varied signal strength dozens or hundreds of times, we
+are limited to only a few questions at a time. With a small set of
+tasks, we need to present tasks that are perceptually hard, and thus ask
+questions about stimuli that are close to our perceptual threshold.
+Therefore, we focus on questions which vary the structure of the
+presented image, but ask viewers to compare the same underlying data
+values across those varied images.
+
+We ask participants to determine which of two just-noticeably different
+marked pieces (tiles) is larger within each chart, and focus on three
+main sets of structural variation in the design of that chart:
 
@@
 \item
-  A is bigger
+  First, we vary the alignment of the pieces in question. Viewers are
+  presented with two marked pieces in a chart that do not share a common
+  baseline (aligned pieces), then two pieces that do share a common
+  baseline (unaligned pieces).
 \item
-  B is bigger
+  Second, we vary the orientation of the chart -- we rotate the vertical
+  stacked bar chart, and present a horizontally oriented version of the
+  same chart, with identically sized marked pieces.
 \item
-  They are the same
+  Finally, we change the aspect ratio of the chart and present a wider
+  version of the horizontally oriented chart which has longer, but
+  thinner, marked pieces.
 \end{itemize}
-
-When presented with the aligned version of each chart, pieces were
-marked with a C and D and the question text is updated accordingly. In
-all scenarios, the second option (B {[}D{]} is bigger) is the correct
-response.
-
-For a given task, viewers are first presented with the unaligned version
-of the task, followed by the aligned version of the task. The time in
-seconds that each respondent spent on each task was recorded, as well as
-whether the participant zoomed in on each chart while answering the
-question. Respondents were also asked to rate their certainty in their
-response to each question on a five-point scale.
 
@@
 therefore be attributed to this difference in presentation.
-
-Task 1 was presented to viewers in two rounds (Rounds 1-2), with four
-color scheme variations each presented to 50\% of respondents in each
-round. There were no significant differences in respondent accuracy or
-behavior across color schemes, and we combine them here into one set of
-stimulus responses to focus on the comparison of structural differences.
-\kr{Should we put the color variations and a test for significance in accuracy/a visual in the appendix/supplementary materials? }
-\hh{we could do that, but I'm not sure that we need to mention color at all.}
 
@@
 
-The images utilized in the third task again represent the same data as
-in the first two tasks, and the overall image has dimensions which are
-identical to the images in the first two tasks, displaying at the same
-size during the survey.
-\hh{The length of bars is increased to fit the new dimensions.
-This increases the difference in the length of bars from previously 8 pixels to 13 pixels.
-The widths of the tiles are adjusted accordingly (from 50 to 30 pixels) to keep the overall area of the tiles approximately constant.
+The images utilized in the third task, \autoref{fig-tasks}(c,f), again
+represent the same data as in the first two tasks, and the overall image
+has dimensions which are identical to the images in the first two tasks,
+displaying at the same size during the survey. The length of bars is
+increased to fit the new dimensions. This increases the difference in
+the length of bars from previously 8 pixels to 13 pixels. The widths of
+the tiles are adjusted accordingly (from 50 to 30 pixels) to keep the
+overall area of the tiles approximately constant.
+
+\hypertarget{question-text}{%
+\paragraph{Question text}\label{question-text}}
+
+When viewing each chart, participants were asked to compare the relative
+sizes of marked elements within the chart:
+
+``There are many charts used in the news media to portray data visually.
+Looking at the chart below, which of the marked dark blue pieces is
+bigger, A or B? Just your best guess is fine.
+
+\begin{itemize}
+\item
+  A is bigger
+\item
+  B is bigger
+\item
+  They are the same
+\end{itemize}
+
+When presented with the aligned version of each chart, pieces were
+marked with a C and D and the question text is updated accordingly. In
+all scenarios, the second option (B {[}D{]} is bigger) is the correct
+response.
+
+For a given task, viewers are first presented with the unaligned version
+of the task, followed by the aligned version of the task. The time in
+seconds that each respondent spent on each task was recorded, as well as
+whether the participant zoomed in on each chart while answering the
+question. Respondents were also asked to rate their certainty in their
+response to each question on a five-point scale.
+
+\begin{figure*}[hbt]
+
+{\centering \includegraphics[width=0.6\textwidth,height=\textheight]{images/sketch-tasks.png}
+
 }
-\ndt{agree to drop note on color. Can perhaps just say we used CBlind  friendly}
-
-\hypertarget{task-3b-floating-bars}{%
-\subsubsection{Task 3b: Floating bars}\label{task-3b-floating-bars}}
-
-\hh{
-In order to measure the impact of the chart's context, we asked a subset of participants to assess the relative sizes of the two tiles A and B (as shown in the bottom right of @fig-tasks-explain) without giving any additional information.}
+
+\caption{\label{fig-tasks-explain}Comparisons made in charts within the
+Cleveland and McGill ranking (left) and their corresponding
+representations in our study tasks (right).}
+
+\end{figure*}
+
+\hypertarget{alignment}{%
+\paragraph{Alignment}\label{alignment}}
+
+What we call `aligned' and `unaligned', here, is similar to Cleveland
+and McGill's positions along aligned and unaligned scales, but with some
+modifications. Figure~\ref{fig-tasks-explain} gives an overview of the
+comparisons of tasks 1 through 3 and the closest corresponding tasks in
+Cleveland and McGill. In this paper, both `aligned' and `unaligned' bars
+share the same axis.
+
+Aligned tiles are additionally anchored in the same position along the
+axis, i.e.~the difference between their sizes can be reduced to a
+positional assessment, which in the absence of three-dimensional cues
+seems to be the general strategy viewers use when evaluating barcharts
+(Zacks et al. 1998). Unaligned tiles do not share this anchor of
+position, i.e.~the difference between their sizes needs at least two
+comparisons of positions. However, unaligned bars in a stacked barchart
+are not just assessments of the lengths of bars -- the other tiles in
+the chart provide a reference frame of the shared axis, which
+\emph{should} help with an assessment of the tiles' sizes. As shown in
+the sketch of Figure~\ref{fig-tasks-explain}, this framing leads to a
+stimulus that is between Cleveland and McGill's task of `position along
+the same but unaligned axis' and an assessment of length.
+
+We would expect that comparing unaligned tiles is a harder task (with
+correspondingly lower levels of accuracy) than a comparison of aligned
+tiles, with the framing given by the other tiles in the same column
+mitigating some of this difficulty. In particular, the additional
+context of the other tiles should \emph{help} rather than \emph{hurt}
+the comparison, unlike other situations, where additional information
+besides the pieces involved in the comparison were found to be
+distractors (Talbot, Setlur, and Anand 2014).
 
@@
 
-Participants were recruited as part of NORC's AmeriSpeak panel, which
-utilizes a probability-based sampling methodology and samples U.S.
-households from NORC's National Sample Frame that provides coverage of
-over 97\% of U.S. households. The current panel size is 54,001 panel
-members aged 13 and over residing in over 43,000 households (Dennis
-2019, updated 2022). \textbackslash ejm\{46,565 panelists are CAWI; they
-are 93.3\% of the panel and 93.6\% of the total panel weights.\} Each
-test was conducted using the AmeriSpeak Omnibus survey, which runs
+Participants were recruited as part of NORC's AmeriSpeak
+panel\footnote{NORC is a non-partisan organization that collects data
+  for government and non-government clients}, which utilizes a
+probability-based sampling methodology and samples U.S. households from
+NORC's National Sample Frame that provides coverage of over 97\% of U.S.
+households. The current panel size is 54,001 panel members aged 13 and
+over residing in over 43,000 households (Dennis 2019, updated 2022).
+Each test was conducted using the AmeriSpeak Omnibus survey, which runs
 biweekly and samples around 1,000 U.S. adults to answer questions on a
@@
 this reason, our survey questions were only presented to web-based
-panelists and not panelists who respond via phone interviews.
+panelists and not panelists who respond via phone interviews. The panel
+has 45,565 web-based panelists, representing 93.3\% of the total panel
+and 96.3\% of the total panel weights.
 
@@
 
-Paragraph on how survey weights are derived? -- \textbf{Question for Ed
--- is this something AmeriSpeak can provide boilerplate language for?}
-
-All calculations in this paper are done in R (R Core Team 2022), and
-weights are applied in analyses using the \texttt{survey} package
-(Lumley 2004) version 4.0 (Lumley 2020) based on Lumley (2010).
-
-When combining responses from different surveys, weights
-\ejm{are rescaled} with respect to the total population, so that their
-order after combining still reflects their relative importance.
+The sample used for each Omnibus study is selected from the AmeriSpeak
+Panel using 48 sampling strata, split by age, race/Hispanic ethnicity,
+education, and gender, with the size of each sample per sampling stratum
+determined by the population distribution for each stratum. In addition,
+sample selection takes into account expected differential survey
+completion rates by demographic groups in order to achieve a
+representative sample of the target population. Collected data are
+weighted to the latest Current Population Survey (CPS) benchmarks from
+the U.S. Census Bureau. They are balanced by gender, age, education,
+race/ethnicity, and geographic region. All calculations in this paper
+are done in R (R Core Team 2022), and weights are applied in analyses
+using the \texttt{survey} package (Lumley 2004) version 4.0 (Lumley
+2020) based on Lumley (2010).
+
+When combining responses from different surveys, weights are rescaled
+with respect to the total population, so that their order after
+combining still reflects their relative importance.
 
@@
 in \(S_i\) by \(\lambda_i\), for \(1 \le i \le \ell\).
-\(\lambda_i \in [0,1]\) with \(\sum_i \lambda_i = 1\) is given as \[
+\(\lambda_i \in [0,1]\) with \(\sum_i \lambda_i = 1\) is given as
+
+\[
 \lambda_i = \frac{n_i/d_i}{\sum_{j=1}^{\ell}n_j/d_j},
-\] where \(n_i\) is the nominal sample of survey \(S_i\) and \(d_i\) are
-the design effects for the estimators. Here, \(d_i\) are estimated as \[
+\]
+
+where \(n_i\) is the nominal sample of survey \(S_i\) and \(d_i\) are
+the design effects for the estimators. Here, \(d_i\) are estimated as
+
+\[
 d_i = 1 + CV(w \in S_i)^2
-\] where \(CV\) is the coefficient of variation of the weights \(w\)
-within each sample, and is estimated as in Kish (1965) : \[ 
+\]
+
+where \(CV\) is the coefficient of variation of the weights \(w\) within
+each sample, and is estimated as in Kish (1965) :
+
+\[ 
 CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
@@
 factors for the combining surveys as discussed above.}\tabularnewline
-\toprule\noalign{}
+\toprule()
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule\noalign{}
+\midrule()
 \endfirsthead
-\toprule\noalign{}
+\toprule()
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule\noalign{}
+\midrule()
 \endhead
-\bottomrule\noalign{}
-\endlastfoot
 Round 1 & April 2022 & 933 & 521.1 & 934.9 & \(\lambda_1\) = 0.343 \\
@@
 \textbf{Total} & --- & 2807 & 1520.8 & & \\
+\bottomrule()
 \end{longtable}
@@
 sample size) and effective sample size for each task.}\tabularnewline
-\toprule\noalign{}
+\toprule()
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule\noalign{}
+\midrule()
 \endhead
-\bottomrule\noalign{}
-\endlastfoot
 Task 1 & Vertical & 1886 & 1007.1 \\
@@
 Task 3 & Horizontal wide & 462 & 249.1 \\
-Task 3b & Floating bars & 915 & 473.0 \\
+\bottomrule()
 \end{longtable}
@@
 
-\caption{\label{fig-response-123}Responses for accuracy by task across
-the designs. The overlaid rectangles represent 95\% confidence
-intervals. The letters in blue and orange encode significances between
-pairwise proportions: two bars have a significantly different proportion
-(at a 5\% significance level) if they do not share a letter. There is no
-significant difference between the designs for wrong responses. When
-tiles are unaligned, the horizontal wide barchart is showing the highest
-accuracy. For aligned tile, the horizontal wide design and the vertical
-design do not show a significant difference in accuracy. Floating bars
-have a significantly lower rate of correct answers than all the other
-designs.}
+\caption{\label{fig-response-123}Responses by task across the designs.
+The overlaid rectangles represent 95\% confidence intervals. The letters
+in blue and orange encode significances between pairwise proportions:
+two bars have a significantly different proportion (at a 5\%
+significance level) if they do not share a letter. There is no
+significant difference between the designs for wrong responses.}
 
@@
 \hypertarget{tbl-rounds-123}{}
-\setlength{\LTpost}{0mm}
 \begin{longtable}{lrrlrrl}
 \caption{\label{tbl-rounds-123}Odds for the cell-means model, letters behind numbers indicate pairwise
-significances. Within the same \emph{column} values are significantly
-different (at 5\%) if they do not share the same letter. }\tabularnewline
+significances. Within the same column values are significantly different
+(at 5\%) if they do not share the same letter. }\tabularnewline
 
@@
 \midrule
-Floating bars\textsuperscript{\textit{1}} & $0.08$ &  [$0.06$, $0.12$] & a & $12.35$ &  [$8.10$, $18.83$] & ab \\ 
-Horizontal & $0.22$ &  [$0.15$, $0.32$] & b & $5.54$ &  [$4.04$, $7.59$] & c \\ 
-Horizontal wide & $0.26$ &  [$0.19$, $0.37$] & bc & $6.20$ &  [$4.46$, $8.61$] & ac \\ 
-Vertical & $0.41$ &  [$0.36$, $0.47$] & cd & $6.90$ &  [$5.75$, $8.28$] & ac \\ 
+Horizontal & $0.22$ &  [$0.15$, $0.32$] & a & $5.54$ &  [$4.04$, $7.59$] & a \\ 
+Horizontal wide & $0.26$ &  [$0.19$, $0.37$] & ab & $6.20$ &  [$4.46$, $8.61$] & a \\ 
+Vertical & $0.41$ &  [$0.36$, $0.47$] & b & $6.90$ &  [$5.75$, $8.28$] & a \\ 
 \midrule
@@
 \midrule
-Horizontal & $0.63$ &  [$0.49$, $0.81$] & d & $14.02$ &  [$9.23$, $21.30$] & b \\ 
-Horizontal wide & $2.67$ &  [$2.04$, $3.48$] & e & $17.12$ &  [$10.49$, $27.94$] & b \\ 
-Vertical & $1.51$ &  [$1.33$, $1.71$] & f & $11.33$ &  [$8.99$, $14.28$] & b \\ 
+Horizontal & $0.63$ &  [$0.49$, $0.81$] & c & $14.02$ &  [$9.23$, $21.30$] & b \\ 
+Horizontal wide & $2.67$ &  [$2.04$, $3.48$] & d & $17.12$ &  [$10.49$, $27.95$] & b \\ 
+Vertical & $1.51$ &  [$1.33$, $1.71$] & e & $11.33$ &  [$8.99$, $14.28$] & b \\ 
 \bottomrule
 \end{longtable}
-\begin{minipage}{\linewidth}
-\textsuperscript{\textit{1}}We are including here, responses on an A vs B comparison for floating bars, corresponding to a length assessment, see bottom right of Figure 1.\\
-\end{minipage}
 
@@
 does not out-perform the vertical design when comparing unaligned tiles.
-\hh{The beneficial effect of frames/context of the chart is apparent when comparing accuracy of responses for the Horizontal wide design and the floating bars: the rate of correct responses in floating bars is significantly lower than in the wide horizontal charts.
-This shows that the additional context }
 
@@
 
-\caption{\label{fig-zoom-123}Zooming - not significant for accuracy or
-task, but changes by the type of chart.}
+\caption{\label{fig-zoom-123}Rates of zooming by task, chart design, and
+correctness. There are significantly higher rates of zooming by
+respondents for the vertical task compared to both horizontal tasks;
+rates of zooming do not differ by correctness of response or alignment
+in the stimulus.}
 
 \end{figure*}
+
+\hypertarget{tbl-zoom-123}{}
+\begin{longtable}{lrrrr}
+\caption{\label{tbl-zoom-123}Coefficients for logistic regression of zooming by task }\tabularnewline
+
+\caption*{
+{\large Estimates for logistic regression on zooming behavior}
+} \\ 
+\toprule
+term & estimate & SE & t-statistic & p-value \\ 
+\midrule
+\multicolumn{5}{l}{} \\ 
+\midrule
+\$\$\textbackslash{}hat\{\textbackslash{}mu\}\$\$ & $-0.35$ & $0.15$ & $-2.4$ & 0.0155 \\ 
+\midrule
+\multicolumn{5}{l}{Response} \\ 
+\midrule
+\$\$\textbackslash{}text\{They are the same \} \textbackslash{}widehat\{\textbackslash{}rho\}\_\{2\}\$\$ & $-0.17$ & $0.09$ & $-1.9$ & 0.0578 \\ 
+Wrong \$\$\textbackslash{}widehat\{\textbackslash{}rho\}\_\{3\}\$\$ & $-0.06$ & $0.13$ & $-0.5$ & 0.6466 \\ 
+\midrule
+\multicolumn{5}{l}{Task} \\ 
+\midrule
+Aligned \$\$\textbackslash{}widehat\{\textbackslash{}tau\}\_\{2\}\$\$ & $0.00$ & $0.06$ & $-0.1$ & 0.9435 \\ 
+\midrule
+\multicolumn{5}{l}{Chart design} \\ 
+\midrule
+Horizontal wide \$\$\textbackslash{}widehat\{\textbackslash{}delta\}\_\{2\}\$\$ & $0.27$ & $0.15$ & $1.8$ & 0.0752 \\ 
+Vertical \$\$\textbackslash{}widehat\{\textbackslash{}delta\}\_\{3\}\$\$ & $0.98$ & $0.13$ & $7.7$ & < 0.0001 \\ 
+\bottomrule
+\end{longtable}
 
@@
 for the horizontal wide orientation.
-\hh{I'll pretty up @tbl-zoom-123 if we are going to keep it.}
-\kr{I think we should! it's nice to have a model for this}\hh{gt and rmarkdown are frustrating! I'll re-set the table in its final form directly in tex}
-
-\hypertarget{tbl-zoom-123}{}
-\begin{longtable}{lrrrr}
-\caption{\label{tbl-zoom-123}Coefficients for logistic regression of zooming by task }\tabularnewline
-
-\caption*{
-{\large Estimates for logistic regression on zooming behavior}
-} \\ 
-\toprule
-term & estimate & SE & t-statistic & p-value \\ 
-\midrule
-\multicolumn{5}{l}{} \\ 
-\midrule
-\$\$\textbackslash{}hat\{\textbackslash{}mu\}\$\$ & $-0.35$ & $0.15$ & $-2.4$ & 0.0155 \\ 
-\midrule
-\multicolumn{5}{l}{Response} \\ 
-\midrule
-\$\$\textbackslash{}text\{They are the same \} \textbackslash{}widehat\{\textbackslash{}rho\}\_\{2\}\$\$ & $-0.17$ & $0.09$ & $-1.9$ & 0.0578 \\ 
-Wrong \$\$\textbackslash{}widehat\{\textbackslash{}rho\}\_\{3\}\$\$ & $-0.06$ & $0.13$ & $-0.5$ & 0.6466 \\ 
-\midrule
-\multicolumn{5}{l}{Task} \\ 
-\midrule
-Aligned \$\$\textbackslash{}widehat\{\textbackslash{}tau\}\_\{2\}\$\$ & $0.00$ & $0.06$ & $-0.1$ & 0.9435 \\ 
-\midrule
-\multicolumn{5}{l}{Chart design} \\ 
-\midrule
-Horizontal wide \$\$\textbackslash{}widehat\{\textbackslash{}delta\}\_\{2\}\$\$ & $0.27$ & $0.15$ & $1.8$ & 0.0752 \\ 
-Vertical \$\$\textbackslash{}widehat\{\textbackslash{}delta\}\_\{3\}\$\$ & $0.98$ & $0.13$ & $7.7$ & < 0.0001 \\ 
-\bottomrule
-\end{longtable}
-
-\hypertarget{how-certain-are-participants}{%
-\subsubsection{How certain are
-participants?}\label{how-certain-are-participants}}
-
-Using linear scores for the response of \texttt{Certainty}, with
-\texttt{not\ certain\ at\ all} assigned a score of 1 and
-\texttt{extremely\ certain} assigned a score of 5, we can estimate the
-effects of task, chart design, and correctness on certainty by using a
-cell-means model of the form: \[
+\hh{gt and rmarkdown are frustrating! I'll re-set the table in its final form directly in tex}
+
+Using linear scores for the response of Certainty, with `not certain at
+all' assigned a score of 1 and `extremely certain' assigned a score of
+5, we can estimate the effects of task, chart design, and correctness on
+certainty by using a cell-means model of the form:
+
+\[
 Y_{k} = \mu_{ij\ell(k)} + \epsilon_{k},
-\] where \(k = 1, ..., N\), \(\mu_{ij\ell(k)}\) is average certainty
+\]
+
+where \(k = 1, ..., N\), \(\mu_{ij\ell(k)}\) is average certainty
 (measured on a scale from 1 to 5) of the four combinations of task and
@@
 \text{ logit } P(Y_k \le i \mid X_k) = \mu_i + X_{k}'\beta_{i}
-\] The resulting model estimates and confidence intervals are shown in
+\]
+
+The resulting model estimates and confidence intervals are shown in
 \autoref{tbl-demographics}. When considering the easier (aligned) task,
@@
 levels do have a (small) effect. When income levels increase, the
-percentage of wrong answers (orange) decreases, while the percentage of
-correct answers (blue) increases slightly (significant at below 0.05).
-For the more difficult task of comparing unaligned tiles, demographics
-are more significant. Higher levels of education and higher levels of
-income are associated with a significant increase of panelists choosing
-a response of `they are the same', resulting in significant decreases of
-both correct answers and wrong answers.}
+percentage of wrong answers decreases, while the percentage of correct
+answers increases slightly (significant at below 0.05). For the more
+difficult task of comparing unaligned tiles, demographics are more
+significant. Higher levels of education and higher levels of income are
+associated with a significant increase of panelists choosing a response
+of `they are the same', resulting in significant decreases of both
+correct answers and wrong answers.}
 
@@
 when viewing it outside of the context of our tests.
-\kr{I don't love the last sentence but I think we can clean it up}.
 
@@
 
+Our study results provide actionable insights that data visualization
+practitioners can utilize in the design of data visualizations.
+Comparisons of interest should be lined up and share a common baseline
+within a chart; in an interactive setting, it may be beneficial to give
+viewers the possibility of re-ordering categories to support effective
+comparison across groupings. Allowing users to zoom in on images may not
+impact perceptual accuracy, but it leads to higher certainty in viewers'
+comparisons. Considering the target audience for visualizations and what
+key findings are most important to convey should be an important step in
+the process of designing a data visualization.
+
 While our results are limited to mapping in a stacked bar chart with a
@@
 different data topic areas.
-
-\hh{The actionable insights are:}
-
-\begin{itemize}
-\tightlist
-\item
-  make sure that the main comparison in a plot is lined up -- in an
-  interactive setting it would be beneficial to have the possibility of
-  re-ordering the categories to help with comparisons.
-\item
-  offering zoom-ins help with certainty, even if it does not help with
-  perceptual accuracy
-\end{itemize}
 
@@
 Vs}. {Cumulating Cases}: {A Comparison} of {Two Weighting Strategies} in
-{NLSY97}.''} In \emph{{ASA Proceedings} of the {Joint Statistical
+{Nlsy97}.''} In \emph{{ASA Proceedings} of the {Joint Statistical
 Meetings}}, 2557--62.

### 2023-07-29 SDSS-2023/testing-charts-accuracy.qmd
a few wording edits from KR full editing pass-through

@@
 Our findings provide actionable guidance for data visualization practitioners to employ in their work."
-keywords: 'list, keywords, here, 3-6, are, fine'
+keywords: 'graphical perception, data visualization, survey data'
 preamble: >
@@
 Structural choices, such as the choice of a graphical element (e.g., bar or pie), orientation of the chart, and sizing and relative positioning of elements within the chart, are determined by the data visualization practitioner.
-The design of a data visualization impacts a viewer's ability to achieve that understanding; a poorly designed data visualization may leave viewers struggling to understand the content or context, or make it difficult to complete accurate and useful comparisons of values across groups or time points.
+The design of a data visualization impacts a viewer's ability to achieve the intended understanding; a poorly designed data visualization may leave viewers struggling to understand the content or context, or make it difficult to complete accurate and useful comparisons of values across groups or time points.
 More broadly, the design of a data visualization can change how viewers interact with the chart.
 
-The viewer's employment of comparisons of the components is a crucial step in the process of interacting with and understanding of chart.
+The viewer's employment of comparisons of the components is a crucial step in the process of interacting with and understanding the chart.
 @clevelandGraphicalPerceptionTheory1984 observed as such, and in their seminal study defined the better visual among a pair as the one that allows viewers to make more accurate comparisons.
-Based on mappings of quantitative variables to different graphical elements, Cleveland and McGill's study resulted in a ranking of perceptual tasks (such as ranking the larger of two lengths or areas) from most accurate to least accurate, which was then extended by @mackinlayAutomatingDesignGraphical1986 to a theoretical framework ranking tasks' order along their ordinal and nominal scales.
+Based on mappings of quantitative variables to different graphical elements, Cleveland and McGill's study resulted in a ranking of perceptual tasks (such as identifying the larger of two lengths or areas) from most accurate to least accurate, which was then extended by @mackinlayAutomatingDesignGraphical1986 to a theoretical framework ranking tasks' order along their ordinal and nominal scales.
 
@@
 
-In this work, we present the series of tests we completed and the resulting findings.
+In this work, we present a series of completed tests and the resulting findings.
 The remainder of the paper is organized as follows: first, we describe the design of the visual stimuli used in the perception tests.
@@
 The heights of the bars are 205 (left) and 213 pixels (right), corresponding to about twice the JND.
-This difference should lead to a relatively high accuracy rate for participants and simultaneously limit the amount of frustration resulting from a task that is perceived as 'too hard'.
+This difference should lead to a relatively high accuracy rate for participants and simultaneously limit the amount of frustration resulting from a task that is perceived as `too hard'.
 
@@
 
-Both charts show the same data with slight modifications to the order of the levels -- the first and second level in each of the bars are reversed between the left and the right chart.
+Both charts show the same data with slight modifications to the order of the levels -- the first and second level in each of the bars are reversed between the chart in (a) and the chart in (d).
 Both charts are displayed at the same size, i.e. in both cases both the difference in size between the bars and the horizontal distance between the bars is the exact same amount.
@@
 
-The visual mappings in the second task, shown in \autoref{fig-tasks}(b, e), are identical to the first task, but the axes of the chart are rotated so that the stacked bars are represented in a horizontal format.
+The visual mappings in the second task, shown in \autoref{fig-tasks}(b,e), are identical to the first task, but the axes of the chart are rotated so that the stacked bars are represented in a horizontal format.
 This represents a structural change in how the data are presented to the viewer while preserving the pixel size of the elements viewers must compare.
@@
 The images utilized in the third task, \autoref{fig-tasks}(c,f), again represent the same data as in the first two tasks, and the overall image has dimensions which are identical to the images in the first two tasks, displaying at the same size during the survey.
+However, the aspect ratio is adjusted to make the plotting area match that of Task 1, with the mapping area being wider rather than taller.
 The length of bars is increased to fit the new dimensions.
-This increases the difference in the length of bars from previously 8 pixels to 13 pixels.
+This increases the difference in the length of bars to 13 pixels (previously 8 pixels).
 The widths of the tiles are adjusted accordingly (from 50 to 30 pixels) to keep the overall area of the tiles approximately constant.
@@
 
--   They are the same
+-   They are the same"
 
@@
 
-Participants were recruited as part of NORC's AmeriSpeak panel^[NORC is a non-partisan organization that collects data for government and non-government clients], which utilizes a probability-based sampling methodology and samples U.S.
+Participants were recruited as part of NORC's AmeriSpeak panel^[NORC is a non-partisan organization that collects data for government and non-government clients.], which utilizes a probability-based sampling methodology and samples U.S.
 households from NORC's National Sample Frame that provides coverage of over 97% of U.S.
@@
 When combining responses from different surveys, weights are rescaled with respect to the total population, so that their order after combining still reflects their relative importance.
-
 We *combine* (rather than cumulate) a set of $\ell$ surveys $S_1$, $S_2$, ...
-$S_\ell$ (with \$ \ell \ge 2\$ ) as described in @omuircheartaighCombiningSamplesVs2002, by multiplying weights in $S_i$ by $\lambda_i$, for $1 \le i \le \ell$.
+$S_\ell$ (with $ \ell \ge 2$ ) as described in @omuircheartaighCombiningSamplesVs2002, by multiplying weights in $S_i$ by $\lambda_i$, for $1 \le i \le \ell$.
 $\lambda_i \in [0,1]$ with $\sum_i \lambda_i = 1$ is given as 
@@
 ```
-The data for this paper were collected in several rounds as part of the NORC Omnibus.
+The data for this paper were collected in several rounds.
 The resulting number of participants, effective sample sizes, and $\lambda$ values are shown in \autoref{tbl-rounds}.
@@
 
-: Survey rounds: dates, number of participants (nominal sample size), effective sample size, sum of weights, and factors for the combining surveys as discussed above. {#tbl-rounds}
+: Survey rounds: dates, number of participants (nominal sample size), effective sample size, sum of weights, and factors for the combining of surveys as discussed above. {#tbl-rounds}
 
@@
 When tiles are aligned along the same axis, more than twice the number of responses are correct.
-The shaded area along the top of the green tiles corresponds to  95\\% confidence intervals around (marginal) correct responses.
+The shaded area along the top of the green bars corresponds to  95\\% confidence intervals around (marginal) correct responses.
 On the right (b), a parallel coordinate plot shows all combinations of responses.
-There's a huge asymmetry in the number of responses where participants answered only one of the questions correctly.
-A lot more responses are correct when comparing aligned tiles than unaligned tiles."
+There's a huge asymmetry in the number of responses where participants answered only one of the questions correctly."
 #| label: fig-alignment
@@
 The estimated values and 95% confidence intervals are shown in @tbl-rounds-123.
-The lowercase letters indicate significances between pairs of estimates: two estimates (in the same column) are significantly different at 5% level, if they do not have any letter in common [@piephoAlgorithmLetterBasedRepresentation2004].
+The lowercase letters indicate significances between pairs of estimates: two estimates within the same column are significantly different at 5% level if they do not have any letter in common [@piephoAlgorithmLetterBasedRepresentation2004].
 
@@
 #| label: tbl-rounds-123
-#| tbl-cap: Odds for the cell-means model, letters behind numbers indicate pairwise significances.
-Within the same *column* values are significantly different (at 5%) if they do not share the same letter.
+#| tbl-cap: Estimated odds from the cell-means model for response patterns.
+Letters behind numbers indicate pairwise significances.
+Within the same column, values are significantly different (at 5%) if they do not share the same letter.
 
@@
   gt() %>%
-  tab_header(title = "Odds of accuracy by task and chart type") %>%
+  tab_header(title = "Odds of selecting responses by task and chart type") %>%
   fmt_number(
@@
 
-Using linear scores for the response of Certainty, with 'not certain at all' assigned a score of 1 and 'extremely certain' assigned a score of 5, we can estimate the effects of task, chart design, and correctness on certainty by using a cell-means model of the form:
+Using linear scores for the response of certainty, with 'not certain at all' assigned a score of 1 and 'extremely certain' assigned a score of 5, we can estimate the effects of task, chart design, and correctness on certainty by using a cell-means model of the form:
 $$
@@
 The results are shown in @fig-certainty.
-What we find, is that the highest scores for certainty are associated with comparing aligned tiles.
-Certainty scores are not significantly different between correct and wrong responses.
+We find that the highest scores for certainty are associated with the aligned task, but certainty scores are not significantly different between correct and wrong responses.
 Scores are (mostly) significantly lower for the unaligned task.
-Interestingly, the lowest scores of certainty are associated with correct responses, here.
+Interestingly, the lowest certainty scores are associated with correct responses to the unaligned task; although respondents successfully complete the task, they are less certain in their selection.
 
@@
 For horizontal stacked bars the difference between aligned and unaligned charts matters significantly.
-Errorbars around the points show estimated scores (on a scale of 1 to 5) with corresponding 95% CI.
-The letters underneath the bars indicate significances.
+Error bars around the points show estimated scores (on a scale of 1 to 5) with corresponding 95% CI.
+The letters underneath the bars indicate significance at 5%.
 The scores for two bars are significantly different if they do not share a letter.
@@
 #| label: fig-demographics
-#| fig-cap: "Panelists' demographics matter, particularly, when the task difficulty increases.
+#| fig-cap: "Response patterns by alignment and demographic levels.
 For aligned tiles, gender, age, and education are not significant factors.
 However, income levels do have a (small) effect.
-When income levels increase, the percentage of wrong answers decreases, while the percentage of correct answers increases slightly (significant at below 0.05).
-For the  more difficult task of comparing unaligned tiles, demographics are more significant.
-Higher levels of education and higher levels of income are associated with a significant increase of panelists choosing a response of 'they are the same', resulting in significant decreases of both correct answers and wrong answers.
+When income levels increase, the percentage of wrong answers decreases, while the percentage of correct answers increases slightly (significant at 5%).
+For the  more difficult task of comparing unaligned tiles, higher levels of education and higher levels of income are associated with a significant increase of panelists choosing a response of 'they are the same', resulting in significant decreases of both correct answers and wrong answers.
 "

### 2023-07-29 SDSS-2023/testing-charts-accuracy.tex
a few wording edits from KR full editing pass-through

@@
   pdftitle={Testing Charts: viewers' perceptual accuracy in surveys},
-  pdfkeywords={list, keywords, here, 3-6, are, fine},
+  pdfkeywords={graphical perception, data visualization, survey data},
   colorlinks=true,
@@
 \end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden, breakable, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 determined by the data visualization practitioner. The design of a data
-visualization impacts a viewer's ability to achieve that understanding;
-a poorly designed data visualization may leave viewers struggling to
-understand the content or context, or make it difficult to complete
-accurate and useful comparisons of values across groups or time points.
-More broadly, the design of a data visualization can change how viewers
-interact with the chart.
+visualization impacts a viewer's ability to achieve the intended
+understanding; a poorly designed data visualization may leave viewers
+struggling to understand the content or context, or make it difficult to
+complete accurate and useful comparisons of values across groups or time
+points. More broadly, the design of a data visualization can change how
+viewers interact with the chart.
 
 The viewer's employment of comparisons of the components is a crucial
-step in the process of interacting with and understanding of chart.
+step in the process of interacting with and understanding the chart.
 Cleveland and McGill (1984) observed as such, and in their seminal study
@@
 variables to different graphical elements, Cleveland and McGill's study
-resulted in a ranking of perceptual tasks (such as ranking the larger of
-two lengths or areas) from most accurate to least accurate, which was
-then extended by Mackinlay (1986) to a theoretical framework ranking
-tasks' order along their ordinal and nominal scales.
+resulted in a ranking of perceptual tasks (such as identifying the
+larger of two lengths or areas) from most accurate to least accurate,
+which was then extended by Mackinlay (1986) to a theoretical framework
+ranking tasks' order along their ordinal and nominal scales.
 
@@
 
-In this work, we present the series of tests we completed and the
-resulting findings. The remainder of the paper is organized as follows:
-first, we describe the design of the visual stimuli used in the
-perception tests. We then describe the population of study respondents
-and obtained survey sample. Subsequently, we share a summary of the
-resulting survey responses across each of the tests, including analyses
-on accuracy of responses and response behavior. Finally, we discuss
-implications of this work and next steps.
+In this work, we present a series of completed tests and the resulting
+findings. The remainder of the paper is organized as follows: first, we
+describe the design of the visual stimuli used in the perception tests.
+We then describe the population of study respondents and obtained survey
+sample. Subsequently, we share a summary of the resulting survey
+responses across each of the tests, including analyses on accuracy of
+responses and response behavior. Finally, we discuss implications of
+this work and next steps.
 
@@
 simultaneously limit the amount of frustration resulting from a task
-that is perceived as 'too hard'.
+that is perceived as `too hard'.
 
@@
 the levels -- the first and second level in each of the bars are
-reversed between the left and the right chart. Both charts are displayed
-at the same size, i.e.~in both cases both the difference in size between
-the bars and the horizontal distance between the bars is the exact same
-amount. This leaves the vertical positioning of the bars as the only
-difference between the charts. Any differences in observed responses can
-therefore be attributed to this difference in presentation.
+reversed between the chart in (a) and the chart in (d). Both charts are
+displayed at the same size, i.e.~in both cases both the difference in
+size between the bars and the horizontal distance between the bars is
+the exact same amount. This leaves the vertical positioning of the bars
+as the only difference between the charts. Any differences in observed
+responses can therefore be attributed to this difference in
+presentation.
 
@@
 
-The visual mappings in the second task, shown in \autoref{fig-tasks}(b,
-e), are identical to the first task, but the axes of the chart are
-rotated so that the stacked bars are represented in a horizontal format.
-This represents a structural change in how the data are presented to the
-viewer while preserving the pixel size of the elements viewers must
-compare. Tasks 2 and 3 were asked within the same survey round (Round 3)
-and the full sample was randomly split among respondents, with 50\% of
-participants seeing the Task 2 stimuli and 50\% of participants seeing
-the Task 3 stimuli.
+The visual mappings in the second task, shown in
+\autoref{fig-tasks}(b,e), are identical to the first task, but the axes
+of the chart are rotated so that the stacked bars are represented in a
+horizontal format. This represents a structural change in how the data
+are presented to the viewer while preserving the pixel size of the
+elements viewers must compare. Tasks 2 and 3 were asked within the same
+survey round (Round 3) and the full sample was randomly split among
+respondents, with 50\% of participants seeing the Task 2 stimuli and
+50\% of participants seeing the Task 3 stimuli.
 
@@
 has dimensions which are identical to the images in the first two tasks,
-displaying at the same size during the survey. The length of bars is
+displaying at the same size during the survey. However, the aspect ratio
+is adjusted to make the plotting area match that of Task 1, with the
+mapping area being wider rather than taller. The length of bars is
 increased to fit the new dimensions. This increases the difference in
-the length of bars from previously 8 pixels to 13 pixels. The widths of
-the tiles are adjusted accordingly (from 50 to 30 pixels) to keep the
+the length of bars to 13 pixels (previously 8 pixels). The widths of the
+tiles are adjusted accordingly (from 50 to 30 pixels) to keep the
 overall area of the tiles approximately constant.
@@
 \item
-  They are the same
+  They are the same''
 \end{itemize}
@@
 panel\footnote{NORC is a non-partisan organization that collects data
-  for government and non-government clients}, which utilizes a
+  for government and non-government clients.}, which utilizes a
 probability-based sampling methodology and samples U.S. households from
@@
 with respect to the total population, so that their order after
-combining still reflects their relative importance.
-
-We \emph{combine} (rather than cumulate) a set of \(\ell\) surveys
-\(S_1\), \(S_2\), \ldots{} \(S_\ell\) (with \$ \ell \ge 2\$ ) as
-described in O'Muircheartaigh and Pedlow (2002), by multiplying weights
-in \(S_i\) by \(\lambda_i\), for \(1 \le i \le \ell\).
-\(\lambda_i \in [0,1]\) with \(\sum_i \lambda_i = 1\) is given as
+combining still reflects their relative importance. We \emph{combine}
+(rather than cumulate) a set of \(\ell\) surveys \(S_1\), \(S_2\),
+\ldots{} \(S_\ell\) (with \$ \ell \ge 2\$ ) as described in
+O'Muircheartaigh and Pedlow (2002), by multiplying weights in \(S_i\) by
+\(\lambda_i\), for \(1 \le i \le \ell\). \(\lambda_i \in [0,1]\) with
+\(\sum_i \lambda_i = 1\) is given as
 
@@
 
-The data for this paper were collected in several rounds as part of the
-NORC Omnibus. The resulting number of participants, effective sample
-sizes, and \(\lambda\) values are shown in \autoref{tbl-rounds}.
+The data for this paper were collected in several rounds. The resulting
+number of participants, effective sample sizes, and \(\lambda\) values
+are shown in \autoref{tbl-rounds}.
 
@@
 (nominal sample size), effective sample size, sum of weights, and
-factors for the combining surveys as discussed above.}\tabularnewline
+factors for the combining of surveys as discussed above.}\tabularnewline
 \toprule()
@@
 the same axis, more than twice the number of responses are correct. The
-shaded area along the top of the green tiles corresponds to 95\%
+shaded area along the top of the green bars corresponds to 95\%
 confidence intervals around (marginal) correct responses. On the right
@@
 There's a huge asymmetry in the number of responses where participants
-answered only one of the questions correctly. A lot more responses are
-correct when comparing aligned tiles than unaligned tiles.}
+answered only one of the questions correctly.}
 
@@
 Table~\ref{tbl-rounds-123}. The lowercase letters indicate significances
-between pairs of estimates: two estimates (in the same column) are
-significantly different at 5\% level, if they do not have any letter in
+between pairs of estimates: two estimates within the same column are
+significantly different at 5\% level if they do not have any letter in
 common (Piepho 2004).
@@
 \begin{longtable}{lrrlrrl}
-\caption{\label{tbl-rounds-123}Odds for the cell-means model, letters behind numbers indicate pairwise
-significances. Within the same column values are significantly different
-(at 5\%) if they do not share the same letter. }\tabularnewline
+\caption{\label{tbl-rounds-123}Estimated odds from the cell-means model for response patterns. Letters
+behind numbers indicate pairwise significances. Within the same column,
+values are significantly different (at 5\%) if they do not share the
+same letter. }\tabularnewline
 
 \caption*{
-{\large Odds of accuracy by task and chart type}
+{\large Odds of selecting responses by task and chart type}
 } \\ 
@@
 
-Using linear scores for the response of Certainty, with `not certain at
+Using linear scores for the response of certainty, with `not certain at
 all' assigned a score of 1 and `extremely certain' assigned a score of
@@
 \(k = 1, ..., N\). The results are shown in Figure~\ref{fig-certainty}.
-What we find, is that the highest scores for certainty are associated
-with comparing aligned tiles. Certainty scores are not significantly
-different between correct and wrong responses. Scores are (mostly)
-significantly lower for the unaligned task. Interestingly, the lowest
-scores of certainty are associated with correct responses, here.
+We find that the highest scores for certainty are associated with the
+aligned task, but certainty scores are not significantly different
+between correct and wrong responses. Scores are (mostly) significantly
+lower for the unaligned task. Interestingly, the lowest certainty scores
+are associated with correct responses to the unaligned task; although
+respondents successfully complete the task, they are less certain in
+their selection.
 
@@
 stacked bar. For horizontal stacked bars the difference between aligned
-and unaligned charts matters significantly. Errorbars around the points
+and unaligned charts matters significantly. Error bars around the points
 show estimated scores (on a scale of 1 to 5) with corresponding 95\% CI.
-The letters underneath the bars indicate significances. The scores for
-two bars are significantly different if they do not share a letter.}
+The letters underneath the bars indicate significance at 5\%. The scores
+for two bars are significantly different if they do not share a letter.}
 
@@
 
-\caption{\label{fig-demographics}Panelists' demographics matter,
-particularly, when the task difficulty increases. For aligned tiles,
-gender, age, and education are not significant factors. However, income
-levels do have a (small) effect. When income levels increase, the
-percentage of wrong answers decreases, while the percentage of correct
-answers increases slightly (significant at below 0.05). For the more
-difficult task of comparing unaligned tiles, demographics are more
-significant. Higher levels of education and higher levels of income are
+\caption{\label{fig-demographics}Response patterns by alignment and
+demographic levels. For aligned tiles, gender, age, and education are
+not significant factors. However, income levels do have a (small)
+effect. When income levels increase, the percentage of wrong answers
+decreases, while the percentage of correct answers increases slightly
+(significant at 5\%). For the more difficult task of comparing unaligned
+tiles, higher levels of education and higher levels of income are
 associated with a significant increase of panelists choosing a response

### 2023-07-30 SDSS-2023/testing-charts-accuracy.qmd
adding a bit to README; deleting unecessary files from repo; adding CIs to zooming figure

@@
 ```
-<!-- fig-width: 6.5 -->
-
-<!-- fig-height: 4.5 -->
-
-<!-- template: template-jdsart.cls.tex -->
 
@@
 
-round5 <- readRDS(file.path(vs_repo_local, "round5.rds"))
-
-# floating bars (unaligned) are shown to all participants of round 5
-round5_adj <- round5 %>% mutate(
-  ab = ab1,
-  ab_zoom = ab1_zoom,
-  ab_certain = ab1_certain,
-  weight = WEIGHT_NOLA
-) %>% filter(
-  ab != "SKIPPED ON WEB"
-)
-
-round6 <- readRDS(file.path(vs_repo_local, "round6.rds"))
-
-round6_adj <- round6 %>% mutate(
-  weight = WEIGHT_NOLA
-) %>% filter(RND_01 == 0) # same type of questions as round 7
-
-round7 <- readRDS(file.path(vs_repo_local, "round7.rds"))
-
-# filter out all of the 'skipped on web':
-round7_adj <- round7 %>% mutate(
-  weight = WEIGHT_NOLA
-)
 ```
@@
 
-lambda_1235 <- calculate_lambda(round_123$weight, round5_adj$weight)
-
-round_123b <- round_123 %>% 
-              mutate(weight = lambda_1235*weight/sum(weight)) %>% 
-  rbind(
-    round5_adj
-     %>% 
-      mutate(
-        weight = (1-lambda_1235)*weight/sum(weight),
-        round = "Round 5",
-        cd = NA,
-        cd_certain=NA,
-        cd_zoom=NA,
-        CorrectAB = ab == "B is bigger",
-        CorrectCD = NA,
-        CorrectABCD = NA,
-        chart = "Floating bars"
-       ) %>% dplyr::select(CaseId, ab, cd, CorrectAB, CorrectCD, CorrectABCD,
-                   ab_certain, cd_certain, ab_zoom, cd_zoom, round, chart, weight,
-                   age4, gender, income4, racethnicity, educ5))
-
 levels(round_123$educ5) <- c("Less than HS", "HS or equivalent", "Some college", "Bachelor", "Post graduate")
@@
 
-round_123b_long <- round_123b %>% 
-  pivot_longer(ab:cd, names_to = "task", values_to = "Response") %>%
-  mutate(
-    correct = Response %in% c("B is bigger", "D is bigger"),
-    correct3 = correct | (Response == "They are the same"),
-    task_chart = factor(paste(task, chart, sep="-")),
-    zoom = ifelse(task=="ab", ab_zoom, cd_zoom),
-    response3 = ifelse(correct, "Correct", ifelse(correct3, "They are the same", "Wrong"))
-  ) %>% filter(
-    !is.na(Response)
-  )
-
-survey_123b <- 
-  survey::svydesign(
-  data = round_123b_long, 
-  ids = ~CaseId, weights = ~weight)
+# round_123b_long <- round_123b %>% 
+#   pivot_longer(ab:cd, names_to = "task", values_to = "Response") %>%
+#   mutate(
+#     correct = Response %in% c("B is bigger", "D is bigger"),
+#     correct3 = correct | (Response == "They are the same"),
+#     task_chart = factor(paste(task, chart, sep="-")),
+#     zoom = ifelse(task=="ab", ab_zoom, cd_zoom),
+#     response3 = ifelse(correct, "Correct", ifelse(correct3, "They are the same", "Wrong"))
+#   ) %>% filter(
+#     !is.na(Response)
+#   )
+# 
+# survey_123b <- 
+#   survey::svydesign(
+#   data = round_123b_long, 
+#   ids = ~CaseId, weights = ~weight)
 
@@
 
-<!-- What is our main objective? -->
-
-```{=html}
-<!--
-\kr{Should the abstract match -- or be close to -- our SDSS short abstract?}
-\hh{I don't think it has to}
--->
-```
 What do viewers see when we show them a data chart?
@@
 Further, the populations' emphasis on higher education individuals also leads to results which hold for groups of individuals who may be more likely to have prior exposure to data visualization in the context of scientific communication, or more exposure to data topics in higher education, but may not hold across other groups within the population.
-<!-- Bias from crowd-sourcing? sample size?  -->
-
-<!--Our work -- as that of Cleveland and McGill and Heer and Bostock -- -->
-
-<!--Our work centers on studying data visualization design choices and their impact on viewer behavior and accuracy of viewers' responses.-->
 
@@
 
-<!-- # ```{r} -->
-
-<!-- # #| label: fig-perceptual-tasks -->
-
-<!-- # #| fig-cap: Ranking of perceptual tasks, as given by @mackinlayAutomatingDesignGraphical1986.
-The ranking of tasks on the quantitative scale are empirically verified by @clevelandGraphicalPerceptionTheory1984.
--->
-
-<!-- # knitr::include_graphics("images/mackinlay-tasks.png") -->
-
-<!-- # ``` -->
-
-<!-- \kr{The figure below is still a bit of an orphan, but I think we can revisit the intro ad study setup after we get more of the results in the paper} -->
-
-<!-- ```{r} -->
-
-<!-- #| label: fig-aligned-unaligned -->
-
-<!-- #| fig-cap: "The only difference between the two pairs of rectangles A, B and C, D is their alignment, i.e.
-A and C are of identical size, as are B and D.
-When participants are asked to compare the size of these tiles in barcharts, the predominant response for the unaligned pair on the left is 'they are of the same size'.
-In contrast, more than half of the viewers respond with 'D is bigger' to the aligned pair of bars on the right." -->
-
-<!-- #| fig-subcap: -->
-
-<!-- #| - "**Unaligned bars**" -->
-
-<!-- #| - "**Aligned bars**" -->
-
-<!-- #| layout-nrow: 1 -->
-
-<!-- #| fig-align: "center" -->
-
-<!-- #| out-width: 50% -->
-
-<!-- knitr::include_graphics("images/round5-modified.PNG") -->
-
-<!-- knitr::include_graphics("images/round3-modified.PNG") -->
-
-<!-- ``` -->
-
 ## Study design -- stimulus
@@
 
-<!-- describe stimulus -->
-
 Both charts show the same data with slight modifications to the order of the levels -- the first and second level in each of the bars are reversed between the chart in (a) and the chart in (d).
@@
 Any differences in observed responses can therefore be attributed to this difference in presentation.
-
-<!-- Task 1 was presented to viewers in two rounds (Rounds 1-2), with four color scheme variations each presented to 50% of respondents in each round.  -->
-
-<!-- There were no significant differences in respondent accuracy or behavior across color schemes, and we combine them here into one set of stimulus responses to focus on the comparison of structural differences.
--->
-
-<!-- \kr{Should we put the color variations and a test for significance in accuracy/a visual in the appendix/supplementary materials?
-} \hh{we could do that, but I'm not sure that we need to mention color at all.} -->
 
@@
 The widths of the tiles are adjusted accordingly (from 50 to 30 pixels) to keep the overall area of the tiles approximately constant.
-<!--However, the aspect ratio of the plotting area is adjusted to increase the length in pixels of the stacked bar elements, and  decrease the relative thickness \kr{wording?} of each bar.
---> <!--kr{Do we have a measurement of the pixels for this?
-I think a sentence on that would close out this section nicely.}\hh{like that?}\kr{I was thinking the length pixels, since we talk about how many they are in Task 1, so we could talk about how many JND the difference is.
-I like the height thing though too, so maybe both?
-}--> <!-- \ndt{agree to drop note on color.
-Can perhaps just say we used CBlind  friendly} -->
-
-<!-- ### Task 3b: Floating bars -->
-
-<!-- \hh{ -->
-
-<!-- In order to measure the impact of the chart's context, we asked a subset of participants to assess the relative sizes of the two tiles A and B (as shown in the bottom right of @fig-tasks-explain) without giving any additional information.}   -->
 
@@
 The panel has 45,565 web-based panelists, representing 93.3% of the total panel and 96.3% of the total panel weights.
-
-<!--\hh{XXX I think what you are trying to say is that we do not know, if participants have seen the stimuli multiple times, but if they did, there was at least a month in between?} \kr{Yes that's what I was trying to say here!!}-->
 
@@
 
-```{=html}
-<!--Note that @omuircheartaighCombiningSamplesVs2002 estimate $\lambda$ separately for any combination of race/ethnicity by sex.
-We employ that strategy whenever we include demographic variables in the analysis, otherwise we use a single adjustment for the weights. -->
-```
 The data for this paper were collected in several rounds.
 The resulting number of participants, effective sample sizes, and $\lambda$ values are shown in \autoref{tbl-rounds}.
-
-<!-- | Name    | Date       |                \# Participants |                Effective Sample Size |         Sum of weights $\sum_i w_i$ | -->
-
-<!-- |---------------|---------------|--------------:|--------------:|--------------:| -->
-
-<!-- | Round 1 | April 2022 |           `r nrow(round1_adj)` | `r round(neff(round1_adj$weight),1)` | `r round(sum(round1_adj$weight),1)` | -->
-
-<!-- | Round 2 | May 2022   |           `r nrow(round2_adj)` | `r round(neff(round2_adj$weight),1)` | `r round(sum(round2_adj$weight),1)` | -->
-
-<!-- | Round 3 | Jun 2022   |           `r nrow(round3_adj)` | `r round(neff(round3_adj$weight),1)` | `r round(sum(round3_adj$weight),1)` | -->
-
-<!-- | Round 6 | Sep 2022   | `r nrow(round6_adj)` \[Split\] | `r round(neff(round6_adj$weight),1)` | `r round(sum(round6_adj$weight),1)` | -->
-
-<!-- | Round 7 | Oct 2022   |           `r nrow(round7_adj)` | `r round(neff(round7_adj$weight),1)` | `r round(sum(round7_adj$weight),1)` | -->
-
-<!-- |         |            |                                |                                      |                                     | -->
-
-<!-- : Survey rounds: dates, number of participants (nominal sample size), effective sample size, and sum of weights. {#tbl-rounds} -->
 
@@
 The distribution of demographics are all quite similar across tasks and their 90% margin of errors (hashed lines) for the most part cover the Census estimates.
-
-<!--\kr{Split by rounds or by tasks?
-I feel like since we later split by task, maybe we should show the demos by task?
-Or we can just talk about how Round 3 was split randomly along demos for Tasks 2/3?}\hh{changed the figure to show demos by task -- do we need confidence intervals?
--- or a comparison against census data?}-->
 
@@
 ```
-
-<!--Rounds 1 and 2 are combined by adjusting the weights with $\lambda$ = `r round(lambdas[1],3)` for an effective sample size of `r round(neff(round_12$weight),1)`.-->
 
@@
 The re-scaled design of the wide horizontal bars reclaims some of the loss for unaligned bars and outperforms the vertical design by a similar margin in aligned bars, but does not out-perform the vertical design when comparing unaligned tiles.
-<!-- \hh{The beneficial effect of frames/context of the chart is apparent when comparing accuracy of responses for the Horizontal wide design and the floating bars: the rate of correct responses in floating bars is significantly lower than in the wide horizontal charts.
-This shows that the additional context } -->
 
@@
 
-<!-- The rate of zooming into charts is about twice as high for participants filling out the survey on their smart phones than on other, usually bigger devices such as tablets or desktops. -->
-
 ```{r zoom-123}
@@
 #| fig-cap: "Rates of zooming by task, chart design, and correctness.
+The overlaid rectangles represent 95% confidence intervals.
 There are significantly higher rates of zooming by respondents for the vertical task compared to both horizontal tasks; rates of zooming do not differ by correctness of response or alignment in the stimulus."
@@
 #| out-width: 80%
+
+confint_zoom1 <- svyby(formula = ~zoom, by = ~task+response3, design = survey_123, svyciprop) %>%
+  mutate(task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")), 
+         zoom2 = zoom, 
+        response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong")))) %>%
+  dplyr::select(-c("zoom"))
+
+confint_zoom2 <- svyby(formula = ~zoom, by = ~chart+response3, design = survey_123, svyciprop) %>%
+  mutate(chart = factor(chart, levels=c("Vertical", "Horizontal", "Horizontal wide")), 
+         zoom2 = zoom, 
+        response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong")))) %>%
+  dplyr::select(-c("zoom"))
 
@@
   scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), 
-                     labels = seq(0,1, by=0.25)*100)
+                     labels = seq(0,1, by=0.25)*100) + 
+  geom_rect(
+    aes(
+      xmin = as.numeric(response3) - 0.45,
+      ymin = zoom2 - 1.96* `se.as.numeric(zoom)`,
+      xmax = as.numeric(response3) + 0.45,
+      ymax = zoom2 + 1.96* `se.as.numeric(zoom)`,
+    ), 
+    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+    data = confint_zoom1
+  )  + 
+  geom_segment(
+    aes(
+      x = as.numeric(response3) - 0.45,
+      y = zoom2,
+      xend = as.numeric(response3) + 0.45,
+      yend = zoom2
+    ), 
+    alpha = 1, colour = "black", linewidth = 0.25,
+    data = confint_zoom1
+  )
 
@@
   scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), 
-                     labels = seq(0,1, by=0.25)*100)
+                     labels = seq(0,1, by=0.25)*100) + 
+  geom_rect(
+    aes(
+      xmin = as.numeric(response3) - 0.45,
+      ymin = zoom2 - 1.96* `se.as.numeric(zoom)`,
+      xmax = as.numeric(response3) + 0.45,
+      ymax = zoom2 + 1.96* `se.as.numeric(zoom)`,
+    ), 
+    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+    data = confint_zoom2
+  )  + 
+  geom_segment(
+    aes(
+      x = as.numeric(response3) - 0.45,
+      y = zoom2,
+      xend = as.numeric(response3) + 0.45,
+      yend = zoom2
+    ), 
+    alpha = 1, colour = "black", linewidth = 0.25,
+    data = confint_zoom2
+  )
 
@@
 
-<!-- Results to discuss: -->
-
-<!-- There are differences by alignment -->
-
-<!-- There are not huge differences across structures -->
-
-<!-- However, we see a difference in how people react across different structures (zooming behavior, certainty? time?) -->
-
-<!-- - We can reproduce prior findings with nat rep sample  -->
-
-<!-- - We can rank  -->
-
-<!-- - We can rank better because we have a bigger sample size  -->
-
-<!-- - We can find changes in viewer behavior  -->
-
-<!-- - Demographics matter to accuracy and choices -->
-
 # Conclusions
@@
 
-<!-- Key findings to discuss: -->
-
-<!-- -   Surveys can be used to ask these types of questions -- in the midst of other topics and with a limited number of questions, we are able to ask perception questions and produce results consistent with prior studies -->
-
-<!-- -   Reproduced (some) prior convenience sample results using a large, nationally-representative survey population: -->
-
-<!--     -   Cleveland: position (aligned) versus unaligned -->
-
-<!--     -   unaligned tiles within stacked bar are a hybrid between framed bar and floats -->
-
-<!--     -   Talbot: pie charts - individual wedges vs piecharts -->
-
-<!-- -   Design choices impact viewer accuracy: -->
-
-<!--     -   new finding: pie charts accuracy in work better when framed (wedges) -->
-
-<!-- -   Other important measures beyond accuracy: time to completion (automatic on web), certainty in response (asked). -->
-
-<!-- -   Studying how viewers interact with charts can be done in a survey, and more expansive work on this topic should be completed to understand how the general public interacts with and understands charts.
--->
-
-<!--     -   Design choices impact viewer behavior: zooming -->
-
-<!-- -   (some) Demographics matter: age and gender does not seem to matter for perception economic background and education does matter for perception: we need to know, we do not want to create a hurdle in communication -->
-
-# Supplementary Material {.unnumbered}
-
--   **Participant Data (Linear):** Link to csv file with the data.
--   **Data Analysis Code:** Link to an html document with annotated code chunks.
+<!-- # Supplementary Material {.unnumbered} -->
+
+<!-- -   **Participant Data (Linear):** Link to csv file with the data. -->
+<!-- -   **Data Analysis Code:** Link to an html document with annotated code chunks. -->
 

### 2023-07-30 SDSS-2023/testing-charts-accuracy.tex
adding a bit to README; deleting unecessary files from repo; adding CIs to zooming figure

@@
 \end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, frame hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, interior hidden, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 \caption{\label{fig-zoom-123}Rates of zooming by task, chart design, and
-correctness. There are significantly higher rates of zooming by
+correctness. The overlaid rectangles represent 95\% confidence
+intervals. There are significantly higher rates of zooming by
 respondents for the vertical task compared to both horizontal tasks;
@@
 
-\hypertarget{supplementary-material}{%
-\section*{Supplementary Material}\label{supplementary-material}}
-\addcontentsline{toc}{section}{Supplementary Material}
-
-\begin{itemize}
-\tightlist
-\item
-  \textbf{Participant Data (Linear):} Link to csv file with the data.
-\item
-  \textbf{Data Analysis Code:} Link to an html document with annotated
-  code chunks.
-\end{itemize}
-
 \hypertarget{references}{%

### 2023-07-31 SDSS-2023/testing-charts-accuracy.qmd
commenting out remaining comments

@@
 In particular, the additional context of the other tiles should *help* rather than *hurt* the comparison, unlike other situations, where additional information besides the pieces involved in the comparison were found to be distractors [@talbotFourExperimentsPerception2014].
-<!--\kr{The updated sketch is great here!!
-One note -- can we move the label for the "positions along the same, but unaligned scale", it's not aligned with the others}--> <!--\hh{What about the sketch now?}-->
 
@@
 This higher rate of zooming does not necessarily lead to higher accuracy; although rates of accuracy are significantly higher for the vertical orientation relative to the horizontal orientation, they are not significantly higher -- and in fact, are significantly lower on the aligned task -- than the accuracy for the horizontal wide orientation.
-\hh{gt and rmarkdown are frustrating! I'll re-set the table in its final form directly in tex}
+<!-- \hh{gt and rmarkdown are frustrating! I'll re-set the table in its final form directly in tex} -->
 

### 2023-07-31 SDSS-2023/testing-charts-accuracy.qmd
adding equation numbers

@@
 $\lambda_i \in [0,1]$ with $\sum_i \lambda_i = 1$ is given as 
-$$
+\begin{equation}
 \lambda_i = \frac{n_i/d_i}{\sum_{j=1}^{\ell}n_j/d_j},
-$$ 
+\end{equation}
 where $n_i$ is the nominal sample of survey $S_i$ and $d_i$ are the design effects for the estimators.
 Here, $d_i$ are estimated as 
-$$
+\begin{equation}
 d_i = 1 + CV(w \in S_i)^2
-$$ 
+\end{equation}
 where $CV$ is the coefficient of variation of the weights $w$ within each sample, and is estimated as in @kishSurveySampling1965 : 
-$$ 
+\begin{equation}
 CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
-$$
+\end{equation}
 
@@
 
-$$
+\begin{equation}
 \text{logit }P(Y_k \le \ell) = \mu_{ij\ell(k)},
-$$
+\end{equation}
 
@@
 
-$$
+\begin{equation}
 \text{logit } P(Y_k \le 1) = \mu + \rho_{i(k)} + \tau_{j(k)} + \delta_{\ell(k)},
-$$
+\end{equation}
 
@@
 Using linear scores for the response of certainty, with 'not certain at all' assigned a score of 1 and 'extremely certain' assigned a score of 5, we can estimate the effects of task, chart design, and correctness on certainty by using a cell-means model of the form:
-$$
+\begin{equation}
 Y_{k} = \mu_{ij\ell(k)} + \epsilon_{k},
-$$ 
+\end{equation}
 where $k = 1, ..., N$, $\mu_{ij\ell(k)}$ is average certainty (measured on a scale from 1 to 5) of the four combinations of task and correctness by each of the three designs, where $i = 1, 2$ encodes unaligned/aligned, and $j=1,2$ encodes wrong, correct, and $\ell = 1, 2, 3$ encodes horizontal, wide horizontal, and vertical stacked bar charts, respectively.
@@
 
-$$
+\begin{equation}
 \text{ logit } P(Y_k \le i \mid X_k) = \mu_i + X_{k}'\beta_{i}
-$$
+\end{equation}
 

### 2023-07-31 SDSS-2023/testing-charts-accuracy.tex
adding equation numbers

@@
 \end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, interior hidden, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 \(\lambda_i\), for \(1 \le i \le \ell\). \(\lambda_i \in [0,1]\) with
-\(\sum_i \lambda_i = 1\) is given as
-
-\[
+\(\sum_i \lambda_i = 1\) is given as \begin{equation}
 \lambda_i = \frac{n_i/d_i}{\sum_{j=1}^{\ell}n_j/d_j},
-\]
-
-where \(n_i\) is the nominal sample of survey \(S_i\) and \(d_i\) are
-the design effects for the estimators. Here, \(d_i\) are estimated as
-
-\[
+\end{equation} where \(n_i\) is the nominal sample of survey \(S_i\) and
+\(d_i\) are the design effects for the estimators. Here, \(d_i\) are
+estimated as \begin{equation}
 d_i = 1 + CV(w \in S_i)^2
-\]
-
-where \(CV\) is the coefficient of variation of the weights \(w\) within
-each sample, and is estimated as in Kish (1965) :
-
-\[ 
+\end{equation} where \(CV\) is the coefficient of variation of the
+weights \(w\) within each sample, and is estimated as in Kish (1965) :
+\begin{equation}
 CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
-\]
+\end{equation}
 
@@
 
-\[
+\begin{equation}
 \text{logit }P(Y_k \le \ell) = \mu_{ij\ell(k)},
-\]
+\end{equation}
 
@@
 
-\[
+\begin{equation}
 \text{logit } P(Y_k \le 1) = \mu + \rho_{i(k)} + \tau_{j(k)} + \delta_{\ell(k)},
-\]
+\end{equation}
 
@@
 for the horizontal wide orientation.
-\hh{gt and rmarkdown are frustrating! I'll re-set the table in its final form directly in tex}
 
@@
 5, we can estimate the effects of task, chart design, and correctness on
-certainty by using a cell-means model of the form:
-
-\[
+certainty by using a cell-means model of the form: \begin{equation}
 Y_{k} = \mu_{ij\ell(k)} + \epsilon_{k},
-\]
-
-where \(k = 1, ..., N\), \(\mu_{ij\ell(k)}\) is average certainty
-(measured on a scale from 1 to 5) of the four combinations of task and
-correctness by each of the three designs, where \(i = 1, 2\) encodes
-unaligned/aligned, and \(j=1,2\) encodes wrong, correct, and
+\end{equation} where \(k = 1, ..., N\), \(\mu_{ij\ell(k)}\) is average
+certainty (measured on a scale from 1 to 5) of the four combinations of
+task and correctness by each of the three designs, where \(i = 1, 2\)
+encodes unaligned/aligned, and \(j=1,2\) encodes wrong, correct, and
 \(\ell = 1, 2, 3\) encodes horizontal, wide horizontal, and vertical
@@
 
-\[
+\begin{equation}
 \text{ logit } P(Y_k \le i \mid X_k) = \mu_i + X_{k}'\beta_{i}
-\]
+\end{equation}
 

### 2023-07-31 SDSS-2023/testing-charts-accuracy.qmd
confidence intervals

@@
 
+```{r demographics-model}
+logit_demo_task_cd_correct <- svyglm(correct~gender+age4+educ5+income4, design=survey_123, family=quasibinomial(), subset = task=="cd")
+
+logit_demo_task_cd_correct_same <- svyglm(correct3~gender+age4+educ5+income4, design=survey_123, family=quasibinomial(), subset = task=="cd")
+
+logit_demo_task_ab_correct <- svyglm(correct~gender+age4+educ5+income4, design=survey_123, family=quasibinomial(), subset = task=="ab")
+logit_demo_task_ab_correct_same <- svyglm(correct3~gender+age4+educ5+income4, design=survey_123, family=quasibinomial(), subset = task=="ab")
+
+coefs_cd_correct <- broom::tidy(logit_demo_task_cd_correct)
+coefs_cd_correct_same <- broom::tidy(logit_demo_task_cd_correct_same)
+
+coefs_ab_correct <- broom::tidy(logit_demo_task_ab_correct)
+coefs_ab_correct_same <- broom::tidy(logit_demo_task_ab_correct_same)
+```
+
 ```{r}
@@
 #|   - Comparisons of unaligned tiles
-#| fig-height: 4
+#| fig-height: 4.5
 #| fig-width: 10
@@
 #| out-width: 100%
+
+educ_glm <- svyglm(correct~educ5-1, design=survey_123, family=quasibinomial(), subset = task=="cd")
+educ_ci <- broom::tidy(educ_glm) %>%
+  mutate(
+    educ5 = factor(levels(round_123$educ5), levels = levels(round_123$educ5)),
+    low = estimate-1.96*std.error,
+    high = estimate+1.96*std.error,
+    estimate = inv_logit(estimate),
+    low = inv_logit(low),
+    high = inv_logit(high))
+
+educ_glm3 <- svyglm(correct3~educ5-1, design=survey_123, family=quasibinomial(), subset = task=="cd")
+
+educ_ci3 <- broom::tidy(educ_glm3) %>%
+  mutate(
+    educ5 = factor(levels(round_123$educ5), levels = levels(round_123$educ5)),
+    low = estimate-1.96*std.error,
+    high = estimate+1.96*std.error,
+    estimate = inv_logit(estimate),
+    low = inv_logit(low),
+    high = inv_logit(high))
+
+# create all pairwise comparisons for tasks and questions.
+educ_compare <- glht(educ_glm, mcp(educ5="Tukey"))
+educ_compare3 <- glht(educ_glm3, mcp(educ5="Tukey"))
+
+# get letters for each confidence
+educ_ci$letters <- cld(educ_compare)$mcletters$Letters
+educ_ci3$letters <- cld(educ_compare3)$mcletters$Letters
 
@@
     ggtitle("\nby education") +
-  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+    geom_segment(
+    aes(
+      x = as.numeric(educ5) - 0.45,
+      y = estimate,
+      xend = as.numeric(educ5) + 0.45,
+      yend = estimate
+    ),
+    alpha = 1, colour = "black", linewidth = 0.25,
+    data = educ_ci
+  ) +
+  geom_segment(
+    aes(
+      x = as.numeric(educ5) - 0.45,
+      y = estimate,
+      xend = as.numeric(educ5) + 0.45,
+      yend = estimate
+    ),
+    alpha = 1, colour = "black", linewidth = 0.25,
+    data = educ_ci3
+  ) +
+  geom_rect(
+    aes(
+      xmin = as.numeric(educ5) - 0.45,
+      ymin = low,
+      xmax = as.numeric(educ5) + 0.45,
+      ymax = high,
+    ),
+    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+    data = educ_ci
+  ) +
+  geom_rect(
+    aes(
+      xmin = as.numeric(educ5) - 0.45,
+      ymin = low,
+      xmax = as.numeric(educ5) + 0.45,
+      ymax = high,
+    ),
+    fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+    data = educ_ci3
+  ) +
+  scale_y_continuous("Percentage", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25), limit=c(-.05, 1.05))  +
+  geom_text(aes(label=letters), y = -0.05, data = educ_ci, colour = "steelblue") +
+  geom_text(aes(label=letters), y = 1.05, data = educ_ci3, colour = "darkorange")
+
+educ_glm <- svyglm(correct~educ5-1, design=survey_123, family=quasibinomial(), subset = task=="ab")
+educ_ci <- broom::tidy(educ_glm) %>%
+  mutate(
+    educ5 = factor(levels(round_123$educ5), levels = levels(round_123$educ5)),
+    low = estimate-1.96*std.error,
+    high = estimate+1.96*std.error,
+    estimate = inv_logit(estimate),
+    low = inv_logit(low),
+    high = inv_logit(high))
+
+educ_glm3 <- svyglm(correct3~educ5-1, design=survey_123, family=quasibinomial(), subset = task=="ab")
+
+educ_ci3 <- broom::tidy(educ_glm3) %>%
+  mutate(
+    educ5 = factor(levels(round_123$educ5), levels = levels(round_123$educ5)),
+    low = estimate-1.96*std.error,
+    high = estimate+1.96*std.error,
+    estimate = inv_logit(estimate),
+    low = inv_logit(low),
+    high = inv_logit(high))
+
+# create all pairwise comparisons for tasks and questions.
+educ_compare <- glht(educ_glm, mcp(educ5="Tukey"))
+educ_compare3 <- glht(educ_glm3, mcp(educ5="Tukey"))
+
+# get letters for each confidence
+educ_ci$letters <- cld(educ_compare)$mcletters$Letters
+educ_ci3$letters <- cld(educ_compare3)$mcletters$Letters
 
@@
     ggtitle("\nby education") +
-  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+#  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100) 
+geom_segment(
+  aes(
+    x = as.numeric(educ5) - 0.45,
+    y = estimate,
+    xend = as.numeric(educ5) + 0.45,
+    yend = estimate
+  ),
+  alpha = 1, colour = "black", linewidth = 0.25,
+  data = educ_ci
+) +
+geom_segment(
+  aes(
+    x = as.numeric(educ5) - 0.45,
+    y = estimate,
+    xend = as.numeric(educ5) + 0.45,
+    yend = estimate
+  ),
+  alpha = 1, colour = "black", linewidth = 0.25,
+  data = educ_ci3
+) +
+geom_rect(
+  aes(
+    xmin = as.numeric(educ5) - 0.45,
+    ymin = low,
+    xmax = as.numeric(educ5) + 0.45,
+    ymax = high,
+  ),
+  fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+  data = educ_ci
+) +
+geom_rect(
+  aes(
+    xmin = as.numeric(educ5) - 0.45,
+    ymin = low,
+    xmax = as.numeric(educ5) + 0.45,
+    ymax = high,
+  ),
+  fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+  data = educ_ci3
+) +
+  scale_y_continuous("Percentage", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25), limit=c(-.05, 1.05))  +
+ geom_text(aes(label=letters), y = -0.05, data = educ_ci, colour = "steelblue") +
+ geom_text(aes(label=letters), y = 1.05, data = educ_ci3, colour = "darkorange")
+
+########
+# gender
+
+gender_glm <- svyglm(correct~gender-1, design=survey_123, family=quasibinomial(), subset = task=="cd")
+gender_ci <- broom::tidy(gender_glm) %>%
+  mutate(
+    gender = factor(levels(round_123$gender), levels = levels(round_123$gender)),
+    low = estimate-1.96*std.error,
+    high = estimate+1.96*std.error,
+    estimate = inv_logit(estimate),
+    low = inv_logit(low),
+    high = inv_logit(high))
+
+gender_glm3 <- svyglm(correct3~gender-1, design=survey_123, family=quasibinomial(), subset = task=="cd")
+
+gender_ci3 <- broom::tidy(gender_glm3) %>%
+  mutate(
+    gender = factor(levels(round_123$gender), levels = levels(round_123$gender)),
+    low = estimate-1.96*std.error,
+    high = estimate+1.96*std.error,
+    estimate = inv_logit(estimate),
+    low = inv_logit(low),
+    high = inv_logit(high))
+
+# create all pairwise comparisons for tasks and questions.
+gender_compare <- glht(gender_glm, mcp(gender="Tukey"))
+gender_compare3 <- glht(gender_glm3, mcp(gender="Tukey"))
+
+# get letters for each confidence
+gender_ci$letters <- cld(gender_compare)$mcletters$Letters
+gender_ci3$letters <- cld(gender_compare3)$mcletters$Letters
 
@@
   ggtitle("Accuracy\nby gender") +
-  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+ geom_segment(
+  aes(
+    x = as.numeric(gender) - 0.45,
+    y = estimate,
+    xend = as.numeric(gender) + 0.45,
+    yend = estimate
+  ),
+  alpha = 1, colour = "black", linewidth = 0.25,
+  data = gender_ci
+) +
+geom_segment(
+  aes(
+    x = as.numeric(gender) - 0.45,
+    y = estimate,
+    xend = as.numeric(gender) + 0.45,
+    yend = estimate
+  ),
+  alpha = 1, colour = "black", linewidth = 0.25,
+  data = gender_ci3
+) +
+geom_rect(
+  aes(
+    xmin = as.numeric(gender) - 0.45,
+    ymin = low,
+    xmax = as.numeric(gender) + 0.45,
+    ymax = high,
+  ),
+  fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+  data = gender_ci
+) +
+geom_rect(
+  aes(
+    xmin = as.numeric(gender) - 0.45,
+    ymin = low,
+    xmax = as.numeric(gender) + 0.45,
+    ymax = high,
+  ),
+  fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+  data = gender_ci3
+) +
+  scale_y_continuous("Percentage", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25), limit=c(-.05, 1.05))  +
+ geom_text(aes(label=letters), y = -0.05, data = gender_ci, colour = "steelblue") +
+ geom_text(aes(label=letters), y = 1.05, data = gender_ci3, colour = "darkorange")
+
+gender_glm <- svyglm(correct~gender-1, design=survey_123, family=quasibinomial(), subset = task=="ab")
+gender_ci <- broom::tidy(gender_glm) %>%
+  mutate(
+    gender = factor(levels(round_123$gender), levels = levels(round_123$gender)),
+    low = estimate-1.96*std.error,
+    high = estimate+1.96*std.error,
+    estimate = inv_logit(estimate),
+    low = inv_logit(low),
+    high = inv_logit(high))
+
+gender_glm3 <- svyglm(correct3~gender-1, design=survey_123, family=quasibinomial(), subset = task=="ab")
+
+gender_ci3 <- broom::tidy(gender_glm3) %>%
+  mutate(
+    gender = factor(levels(round_123$gender), levels = levels(round_123$gender)),
+    low = estimate-1.96*std.error,
+    high = estimate+1.96*std.error,
+    estimate = inv_logit(estimate),
+    low = inv_logit(low),
+    high = inv_logit(high))
+
+# create all pairwise comparisons for tasks and questions.
+gender_compare <- glht(gender_glm, mcp(gender="Tukey"))
+gender_compare3 <- glht(gender_glm3, mcp(gender="Tukey"))
+
+# get letters for each confidence
+gender_ci$letters <- cld(gender_compare)$mcletters$Letters
+gender_ci3$letters <- cld(gender_compare3)$mcletters$Letters
 
@@
   ggtitle("Accuracy\nby gender") +
-  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+   geom_segment(
+  aes(
+    x = as.numeric(gender) - 0.45,
+    y = estimate,
+    xend = as.numeric(gender) + 0.45,
+    yend = estimate
+  ),
+  alpha = 1, colour = "black", linewidth = 0.25,
+  data = gender_ci
+) +
+geom_segment(
+  aes(
+    x = as.numeric(gender) - 0.45,
+    y = estimate,
+    xend = as.numeric(gender) + 0.45,
+    yend = estimate
+  ),
+  alpha = 1, colour = "black", linewidth = 0.25,
+  data = gender_ci3
+) +
+geom_rect(
+  aes(
+    xmin = as.numeric(gender) - 0.45,
+    ymin = low,
+    xmax = as.numeric(gender) + 0.45,
+    ymax = high,
+  ),
+  fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+  data = gender_ci
+) +
+geom_rect(
+  aes(
+    xmin = as.numeric(gender) - 0.45,
+    ymin = low,
+    xmax = as.numeric(gender) + 0.45,
+    ymax = high,
+  ),
+  fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+  data = gender_ci3
+) +
+  scale_y_continuous("Percentage", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25), limit=c(-.05, 1.05))  +
+ geom_text(aes(label=letters), y = -0.05, data = gender_ci, colour = "steelblue") +
+ geom_text(aes(label=letters), y = 1.05, data = gender_ci3, colour = "darkorange")
+###########
+
+# income
+
+income_glm <- svyglm(correct~income4-1, design=survey_123, family=quasibinomial(), subset = task=="cd")
+income_ci <- broom::tidy(income_glm) %>%
+  mutate(
+    income4 = factor(levels(round_123$income4), levels = levels(round_123$income4)),
+    low = estimate-1.96*std.error,
+    high = estimate+1.96*std.error,
+    estimate = inv_logit(estimate),
+    low = inv_logit(low),
+    high = inv_logit(high))
+
+income_glm3 <- svyglm(correct3~income4-1, design=survey_123, family=quasibinomial(), subset = task=="cd")
+
+income_ci3 <- broom::tidy(income_glm3) %>%
+  mutate(
+    income4 = factor(levels(round_123$income4), levels = levels(round_123$income4)),
+    low = estimate-1.96*std.error,
+    high = estimate+1.96*std.error,
+    estimate = inv_logit(estimate),
+    low = inv_logit(low),
+    high = inv_logit(high))
+
+# create all pairwise comparisons for tasks and questions.
+income_compare <- glht(income_glm, mcp(income4="Tukey"))
+income_compare3 <- glht(income_glm3, mcp(income4="Tukey"))
+
+# get letters for each confidence
+income_ci$letters <- cld(income_compare)$mcletters$Letters
+income_ci3$letters <- cld(income_compare3)$mcletters$Letters
 
@@
     ggtitle("\nby income") +
-  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+   geom_segment(
+  aes(
+    x = as.numeric(income4) - 0.45,
+    y = estimate,
+    xend = as.numeric(income4) + 0.45,
+    yend = estimate
+  ),
+  alpha = 1, colour = "black", linewidth = 0.25,
+  data = income_ci
+) +
+geom_segment(
+  aes(
+    x = as.numeric(income4) - 0.45,
+    y = estimate,
+    xend = as.numeric(income4) + 0.45,
+    yend = estimate
+  ),
+  alpha = 1, colour = "black", linewidth = 0.25,
+  data = income_ci3
+) +
+geom_rect(
+  aes(
+    xmin = as.numeric(income4) - 0.45,
+    ymin = low,
+    xmax = as.numeric(income4) + 0.45,
+    ymax = high,
+  ),
+  fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+  data = income_ci
+) +
+geom_rect(
+  aes(
+    xmin = as.numeric(income4) - 0.45,
+    ymin = low,
+    xmax = as.numeric(income4) + 0.45,
+    ymax = high,
+  ),
+  fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+  data = income_ci3
+) +
+  scale_y_continuous("Percentage", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25), limit=c(-.05, 1.05))  +
+ geom_text(aes(label=letters), y = -0.05, data = income_ci, colour = "steelblue") +
+ geom_text(aes(label=letters), y = 1.05, data = income_ci3, colour = "darkorange")
+
+income_glm <- svyglm(correct~income4-1, design=survey_123, family=quasibinomial(), subset = task=="ab")
+income_ci <- broom::tidy(income_glm) %>%
+  mutate(
+    income4 = factor(levels(round_123$income4), levels = levels(round_123$income4)),
+    low = estimate-1.96*std.error,
+    high = estimate+1.96*std.error,
+    estimate = inv_logit(estimate),
+    low = inv_logit(low),
+    high = inv_logit(high))
+
+income_glm3 <- svyglm(correct3~income4-1, design=survey_123, family=quasibinomial(), subset = task=="ab")
+
+income_ci3 <- broom::tidy(income_glm3) %>%
+  mutate(
+    income4 = factor(levels(round_123$income4), levels = levels(round_123$income4)),
+    low = estimate-1.96*std.error,
+    high = estimate+1.96*std.error,
+    estimate = inv_logit(estimate),
+    low = inv_logit(low),
+    high = inv_logit(high))
+
+# create all pairwise comparisons for tasks and questions.
+income_compare <- glht(income_glm, mcp(income4="Tukey"))
+income_compare3 <- glht(income_glm3, mcp(income4="Tukey"))
+
+# get letters for each confidence
+income_ci$letters <- cld(income_compare)$mcletters$Letters
+income_ci3$letters <- cld(income_compare3)$mcletters$Letters
 
@@
     ggtitle("\nby income") +
-  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+   geom_segment(
+  aes(
+    x = as.numeric(income4) - 0.45,
+    y = estimate,
+    xend = as.numeric(income4) + 0.45,
+    yend = estimate
+  ),
+  alpha = 1, colour = "black", linewidth = 0.25,
+  data = income_ci
+) +
+geom_segment(
+  aes(
+    x = as.numeric(income4) - 0.45,
+    y = estimate,
+    xend = as.numeric(income4) + 0.45,
+    yend = estimate
+  ),
+  alpha = 1, colour = "black", linewidth = 0.25,
+  data = income_ci3
+) +
+geom_rect(
+  aes(
+    xmin = as.numeric(income4) - 0.45,
+    ymin = low,
+    xmax = as.numeric(income4) + 0.45,
+    ymax = high,
+  ),
+  fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+  data = income_ci
+) +
+geom_rect(
+  aes(
+    xmin = as.numeric(income4) - 0.45,
+    ymin = low,
+    xmax = as.numeric(income4) + 0.45,
+    ymax = high,
+  ),
+  fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+  data = income_ci3
+) +
+  scale_y_continuous("Percentage", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25), limit=c(-.05, 1.05))  +
+ geom_text(aes(label=letters), y = -0.05, data = income_ci, colour = "steelblue") +
+ geom_text(aes(label=letters), y = 1.05, data = income_ci3, colour = "darkorange")
 
@@
 
+age_glm <- svyglm(correct~age4-1, design=survey_123, family=quasibinomial(), subset = task=="cd")
+age_ci <- broom::tidy(age_glm) %>%
+  mutate(
+    age4 = factor(levels(round_123$age4), levels = levels(round_123$age4)),
+    low = estimate-1.96*std.error,
+    high = estimate+1.96*std.error,
+    estimate = inv_logit(estimate),
+    low = inv_logit(low),
+    high = inv_logit(high))
+
+age_glm3 <- svyglm(correct3~age4-1, design=survey_123, family=quasibinomial(), subset = task=="cd")
+
+age_ci3 <- broom::tidy(age_glm3) %>%
+  mutate(
+    age4 = factor(levels(round_123$age4), levels = levels(round_123$age4)),
+    low = estimate-1.96*std.error,
+    high = estimate+1.96*std.error,
+    estimate = inv_logit(estimate),
+    low = inv_logit(low),
+    high = inv_logit(high))
+
+# create all pairwise comparisons for tasks and questions.
+age_compare <- glht(age_glm, mcp(age4="Tukey"))
+age_compare3 <- glht(age_glm3, mcp(age4="Tukey"))
+
+# get letters for each confidence
+age_ci$letters <- cld(age_compare)$mcletters$Letters
+age_ci3$letters <- cld(age_compare3)$mcletters$Letters
+
 age_cd <- round_123_long %>% 
@@
     ggtitle("\nby age") +
-  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+   geom_segment(
+  aes(
+    x = as.numeric(age4) - 0.45,
+    y = estimate,
+    xend = as.numeric(age4) + 0.45,
+    yend = estimate
+  ),
+  alpha = 1, colour = "black", linewidth = 0.25,
+  data = age_ci
+) +
+geom_segment(
+  aes(
+    x = as.numeric(age4) - 0.45,
+    y = estimate,
+    xend = as.numeric(age4) + 0.45,
+    yend = estimate
+  ),
+  alpha = 1, colour = "black", linewidth = 0.25,
+  data = age_ci3
+) +
+geom_rect(
+  aes(
+    xmin = as.numeric(age4) - 0.45,
+    ymin = low,
+    xmax = as.numeric(age4) + 0.45,
+    ymax = high,
+  ),
+  fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+  data = age_ci
+) +
+geom_rect(
+  aes(
+    xmin = as.numeric(age4) - 0.45,
+    ymin = low,
+    xmax = as.numeric(age4) + 0.45,
+    ymax = high,
+  ),
+  fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+  data = age_ci3
+) +
+  scale_y_continuous("Percentage", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25), limit=c(-.05, 1.05))  +
+ geom_text(aes(label=letters), y = -0.05, data = age_ci, colour = "steelblue") +
+ geom_text(aes(label=letters), y = 1.05, data = age_ci3, colour = "darkorange")
+
+age_glm <- svyglm(correct~age4-1, design=survey_123, family=quasibinomial(), subset = task=="ab")
+age_ci <- broom::tidy(age_glm) %>%
+  mutate(
+    age4 = factor(levels(round_123$age4), levels = levels(round_123$age4)),
+    low = estimate-1.96*std.error,
+    high = estimate+1.96*std.error,
+    estimate = inv_logit(estimate),
+    low = inv_logit(low),
+    high = inv_logit(high))
+
+age_glm3 <- svyglm(correct3~age4-1, design=survey_123, family=quasibinomial(), subset = task=="ab")
+
+age_ci3 <- broom::tidy(age_glm3) %>%
+  mutate(
+    age4 = factor(levels(round_123$age4), levels = levels(round_123$age4)),
+    low = estimate-1.96*std.error,
+    high = estimate+1.96*std.error,
+    estimate = inv_logit(estimate),
+    low = inv_logit(low),
+    high = inv_logit(high))
+
+# create all pairwise comparisons for tasks and questions.
+age_compare <- glht(age_glm, mcp(age4="Tukey"))
+age_compare3 <- glht(age_glm3, mcp(age4="Tukey"))
+
+# get letters for each confidence
+age_ci$letters <- cld(age_compare)$mcletters$Letters
+age_ci3$letters <- cld(age_compare3)$mcletters$Letters
 
@@
     ggtitle("\nby age") +
-  scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), labels = seq(0,1, by=0.25)*100)
+   geom_segment(
+  aes(
+    x = as.numeric(age4) - 0.45,
+    y = estimate,
+    xend = as.numeric(age4) + 0.45,
+    yend = estimate
+  ),
+  alpha = 1, colour = "black", linewidth = 0.25,
+  data = age_ci
+) +
+geom_segment(
+  aes(
+    x = as.numeric(age4) - 0.45,
+    y = estimate,
+    xend = as.numeric(age4) + 0.45,
+    yend = estimate
+  ),
+  alpha = 1, colour = "black", linewidth = 0.25,
+  data = age_ci3
+) +
+geom_rect(
+  aes(
+    xmin = as.numeric(age4) - 0.45,
+    ymin = low,
+    xmax = as.numeric(age4) + 0.45,
+    ymax = high,
+  ),
+  fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+  data = age_ci
+) +
+geom_rect(
+  aes(
+    xmin = as.numeric(age4) - 0.45,
+    ymin = low,
+    xmax = as.numeric(age4) + 0.45,
+    ymax = high,
+  ),
+  fill = "black", alpha = 0.25, #colour = "white", linewidth = 0.25,
+  data = age_ci3
+) +
+  scale_y_continuous("Percentage", breaks=seq(0,1,by=0.25), labels = 100*seq(0,1,by=0.25), limit=c(-.05, 1.05))  +
+ geom_text(aes(label=letters), y = -0.05, data = age_ci, colour = "steelblue") +
+ geom_text(aes(label=letters), y = 1.05, data = age_ci3, colour = "darkorange")
+
 gender_cd + age_cd  + educ_cd +income_cd + 
@@
   plot_layout(nrow=1, guides = "collect", widths = c(2,4,5,4))
-```
-
-```{r demographics-model}
-logit_demo_task_cd_correct <- svyglm(correct~gender+age4+educ5+income4, design=survey_123, family=quasibinomial(), subset = task=="cd")
-
-logit_demo_task_cd_correct_same <- svyglm(correct3~gender+age4+educ5+income4, design=survey_123, family=quasibinomial(), subset = task=="cd")
-
-logit_demo_task_ab_correct <- svyglm(correct~gender+age4+educ5+income4, design=survey_123, family=quasibinomial(), subset = task=="ab")
-logit_demo_task_ab_correct_same <- svyglm(correct3~gender+age4+educ5+income4, design=survey_123, family=quasibinomial(), subset = task=="ab")
-
-coefs_cd_correct <- broom::tidy(logit_demo_task_cd_correct)
-coefs_cd_correct_same <- broom::tidy(logit_demo_task_cd_correct_same)
-
-coefs_ab_correct <- broom::tidy(logit_demo_task_ab_correct)
-coefs_ab_correct_same <- broom::tidy(logit_demo_task_ab_correct_same)
 ```

### 2023-07-31 SDSS-2023/testing-charts-accuracy.tex
confidence intervals

@@
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
 \fi
@@
 \makeatletter
-\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
 \makeatother
@@
 \@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
 \makeatother
@@
 \end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 factors for the combining of surveys as discussed above.}\tabularnewline
-\toprule()
+\toprule\noalign{}
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule()
+\midrule\noalign{}
 \endfirsthead
-\toprule()
+\toprule\noalign{}
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule()
+\midrule\noalign{}
 \endhead
+\bottomrule\noalign{}
+\endlastfoot
 Round 1 & April 2022 & 933 & 521.1 & 934.9 & \(\lambda_1\) = 0.343 \\
@@
 \textbf{Total} & --- & 2807 & 1520.8 & & \\
-\bottomrule()
 \end{longtable}
@@
 sample size) and effective sample size for each task.}\tabularnewline
-\toprule()
+\toprule\noalign{}
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} \\
-\midrule()
+\midrule\noalign{}
 \endhead
+\bottomrule\noalign{}
+\endlastfoot
 Task 1 & Vertical & 1886 & 1007.1 \\
@@
 Task 3 & Horizontal wide & 462 & 249.1 \\
-\bottomrule()
 \end{longtable}
@@
 Vs}. {Cumulating Cases}: {A Comparison} of {Two Weighting Strategies} in
-{Nlsy97}.''} In \emph{{ASA Proceedings} of the {Joint Statistical
+{NLSY97}.''} In \emph{{ASA Proceedings} of the {Joint Statistical
 Meetings}}, 2557--62.

### 2023-07-31 SDSS-2023/testing-charts-accuracy.tex
remove 'and'
table 4

@@
 \title{Testing Charts: viewers' perceptual accuracy in surveys}
-\author{}
 \date{}
@@
 \end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 \midrule
-\multicolumn{5}{l}{} \\ 
-\midrule
-\$\$\textbackslash{}hat\{\textbackslash{}mu\}\$\$ & $-0.35$ & $0.15$ & $-2.4$ & 0.0155 \\ 
+$\hat{\mu}$ & $-0.35$ & $0.15$ & $-2.4$ & 0.0155 \\ 
 \midrule
@@
 \midrule
-\$\$\textbackslash{}text\{They are the same \} \textbackslash{}widehat\{\textbackslash{}rho\}\_\{2\}\$\$ & $-0.17$ & $0.09$ & $-1.9$ & 0.0578 \\ 
-Wrong \$\$\textbackslash{}widehat\{\textbackslash{}rho\}\_\{3\}\$\$ & $-0.06$ & $0.13$ & $-0.5$ & 0.6466 \\ 
+They are the same, $\widehat{\rho_2}$ & $-0.17$ & $0.09$ & $-1.9$ & 0.0578 \\ 
+Wrong, $\widehat{\rho_3}$ & $-0.06$ & $0.13$ & $-0.5$ & 0.6466 \\ 
 \midrule
@@
 \midrule
-Aligned \$\$\textbackslash{}widehat\{\textbackslash{}tau\}\_\{2\}\$\$ & $0.00$ & $0.06$ & $-0.1$ & 0.9435 \\ 
+Aligned, $\widehat{\tau_2}$ & $0.00$ & $0.06$ & $-0.1$ & 0.9435 \\ 
 \midrule
@@
 \midrule
-Horizontal wide \$\$\textbackslash{}widehat\{\textbackslash{}delta\}\_\{2\}\$\$ & $0.27$ & $0.15$ & $1.8$ & 0.0752 \\ 
-Vertical \$\$\textbackslash{}widehat\{\textbackslash{}delta\}\_\{3\}\$\$ & $0.98$ & $0.13$ & $7.7$ & < 0.0001 \\ 
+Horizontal wide, $\widehat{\delta_2}$ & $0.27$ & $0.15$ & $1.8$ & 0.0752 \\ 
+Vertical, $\widehat{\delta_3}$ & $0.98$ & $0.13$ & $7.7$ & < 0.0001 \\ 
 \bottomrule
@@
 
+\begin{verbatim}
+Warning in RET$pfunction("adjusted", ...): Completion with error > abseps
+\end{verbatim}
+
 \begin{figure}

### 2023-07-31 SDSS-supplement/supplement.Rmd
adding supplement PDF

---
title: "Testing Charts: viewers' perceptual accuracy in surveys"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Supplementary Materials

Additional materials for this paper are shared in the following GitHub repository: [https://github.com/kiegan/testing-charts-jds](https://github.com/kiegan/testing-charts-jds). The repository contains all data used in analyses, the original abstract submission file for SDSS 2023, and the code used to generate all figures and tables in the submission document.

### 2023-10-06 revisions-round1/response-to-reviewers.qmd
adding a doc to respond to reviewer comments

---
title: "Round 1: Response to Reviewers"
format: pdf
editor: source
---

## Reviewer 1

-   The introduction doesn't provide a clear rationale for the questions' importance.

-   The entire article reads more like a project report rather than a scientific paper, with an excess of questions and bullet-pointed sentences.

-   There is an overabundance of figures without clear emphasis.

-   I have doubts about whether the paper aligns with the Journal's scope.

    -   *Per AE and Editor, we are not responding to this comment.*

## Reviewer 2

-   There are a couple of assumptions that assume certain factors (order of question, rounds) are negligible to the outcome. While I think this is reasonable, for a higher confidence in results it might be a good idea to test all visual stimulus within each round and block (if it is hard to be individually randomised) so that some participants get different question order.

-   Major comments

    -   The title, abstract and conclusion suggests a wider context but the content is actually more limited in scope. I suggest changing the title to emphasise the main points in the article: 1) nationally representative sample and 2) testing stacked barplots. I think statements like "Our study results provide actionable insights that data visualization practitioners can utilize in the design of data visualizations." extrapolates too much. It should be more specific in that the actionable insights is *for stacked barplots* and also write concretely what the recommendation is (i.e., vertical display, align groups that most important to compare, etc.).

    -   Do the authors have recommendations for (non-representative) visualisation perception surveys given their experience on nationally representative sample? E.g. given that the surveys conducted via Mturks etc are biased toward highly educated young males and authors found differences between the income group, what do these non-representative survey inform us?

-   Minor comments

    -   I think the paper can benefit in the intro to include a stronger emphasis why we want to achieve a representative sample. The authors do address this in part but it reads like an aside ("..., a common target audience for data visualisation"). Maybe a paragraph on about how data visualisation is used by general public (e.g. results on polls in the news -- this can also lead to why the authors chose to focus on barplots too as polls are often count data). Authors can also discuss how it is **imperative to be inclusive** as these data visualisation can be informative and the results need to be accessible by the *whole* population and not just the elite/highly educated population (which seem to be the bias towards at lot of these surveys have). I think also a stronger emphasis on the importance of the authors' works is needed, e.g., "this is the first study to ...". Also in conclusion, the authors state "The size of our sample aids in identifying this signal.", but the authors can instead state e.g. "our study is orders of magnitude larger than other studies with similar scope, demonstrating...".

    -   While I think the content is fine, I think the grammar in the paper can be better polished. (It is not that the current grammar is bad, but I just think it can be improved to make this a better paper). I have outlined some of the suggested corrections below.

        -   page 2, second paragraph: "The viewer's employment of comparisons of the components" -\> "The comparison of components"

        -   page 2, third paragraph: "to be biased towards more male, younger and relatively higher education" -\> "to be biased towards more male, younger and relatively well-educated individuals"

        -   page 2, third paragraph: "Further, the populations' emphasis..." -\> This sentence was confusing. I kind of get the gist what the authors are trying to say but I wonder if it can be paraphrased better? E.g. "Arguably, we could limit the population of interest to highly educated individuals, however, this group is also likely to have prior exposure to data analysis in higher education, which may not be the case for other groups in the wider population."

        -   page 2, fourth paragraph: I"m not sure if you should start the paragraph with "Here".

    -   FIgure 1, the text within the visualisation are too small to see. Perhaps lay it out as a 3 x 2 instead?

    -   page 8, first paragraph: "\$ 2\$" is a typo?

    -   Table 1, effective sample size adds up to 1520.3 but the total says 1520.8?

    -   Figure 3: I take it that the American Community Survey 2021 is considered the Census estimates? I think it makes more sense to swap the bar and point so that the bar is the Census estimate. This way you can use the errorbar plot. The margin of error around a barplot is reminiscent of a dynamite plot, which I don't think is well thought of in the community (at least not me). I think you can also color the bars to indicate whether your survey estimate covers the census estimate so it is easier to see which one covers it or not. In this plot, you've facetted by the stimulus type but I think it makes sense to facet by the category level (the comparison of distribution is done with the census estimate directly, and I think the interest is to compare the distribution across stimulus so you want the primary factor of interest to compare to be close in proximity). Within a category level, the census estimate is constant across stimulus so you could substitute is with a horizontal line (instead of a bar or point) as well.

    -   Figure 4: I think (b) alone is sufficient as you can gain the same information as (a) (although perhaps it is less familiar for wider audience). You could arguably put figure (a) information in a table instead.

    -   Rephrase needed on page 18 third paragraph "... asking directly asking ...".

    -   "Our study results provide actionable insights ... in the design of data visualisations." -\> This suggests wider scope than what the results suggest. I think the authors should say "design of stacked barplots" instead.

    -   In the source code of the paper, authors use 1.96 and 1.645 for 0.875 and 0.95 quantiles but I think the authors should use `qnorm(0.975)` and `qnorm(0.95)` (I realize it doesn't make much difference in the end but I think you should avoid rounding unnecessary until the presentation of the results).

    -   I do appreciate that the authors present the results using vertical stacked barplot, reflected from their results. A minor point is that I actually don't like the angled display of the x-axis text and I would have preferred to present this in a horizontal display so the text stays horizontal. I realize this sacrifices partially the accuracy of the results for something more aesthetically pleasing for peripheral information. But this opens up another question: is communicating information about always presenting the most accurate results? Communication won't happen if the recipient isn't taking in that information in the first place and while it may be superficial, aesthetically pleasing display may perhaps drive *better engagement*. In that sense, horizontal display where axis text can be better presented may be a better alternative? Authors could perhaps include a discussion along this line and I'd be interested in knowing what the authors think/recommend.

## Editors

-   The AE and I do believe this is within the scope for this special issue, contrary to the final comment from Referee #1, so no need to address that comment.

-   Please add an unnumbered section Supplementary Material to list its content (e.g., data, code, additional simulation results, technical proofs, etc.). See recently published papers for examples. Reproducibility is held high at the JDS; the reproducibility team will check for reproducibility when the paper is accepted.

-   When addressing the comments, please keep in mind that there is a 20-page limit (including everything) in the production style.

-   In general, both reviewers express some concerns regarding some of the language and description regarding overall impact of the work. I would encourage the authors to make it more clear that their primary contribution is in using a representative probability sample, and that their conclusions are focused on stacked barplot displays, as Reviewer 2 specifically remarks, and highlight why conclusions from a non-representative sample may be problematic within the data visualization world.

-   I believe the figures are fairly clear (per Reviewer 1's comment), but the authors might also be inclined to be more selective with what is shown in Figures 3, 6-8, and maybe move some of the others to an appendix.

-   Can the authors comment further on modeling assumptions, particularly in relation to presenting all participants with the unaligned plot first followed by the aligned plot second (as opposed to randomizing this)? For instance, perhaps some participants notice that the plots are the same, with just the bars stacked in a different order, and their response to the second task is biased to be equivalent to that of the first task. This seems like it would be more of a problem if the plots were presented in the reverse order, but I'm curious if the authors looked into this.

-   Is there a reason why the meaning/values of i from 1 to 3 are reversed between the two models specified in Equations 4 and 7?

-   It appears the cumulative logistic regression model with respect to participant demographic covariates just considers first-order effects, but interactions would also be pretty interesting here. Are there any notable interaction effects?

-   The letter notation used to denote significance is a bit confusing to me (e.g., Figure 5, Table 3), can the authors clarify this within the text or modify accordingly? What do a, b, c, etc. actually mean? If this is convention from Piepho (2004), it would be useful for this to be explained more or changed for a more general audience.

### 2023-10-16 revisions-round1/response-to-reviewers.qmd
one edit to revisions
addressing language-based comments from reviewers in round 1

@@
 
--   The introduction doesn't provide a clear rationale for the questions' importance.
+-   **The introduction doesn't provide a clear rationale for the questions' importance.**
 
 -   The entire article reads more like a project report rather than a scientific paper, with an excess of questions and bullet-pointed sentences.
+    -   *To address this, we have removed two of the instances of bulleted lists: one which listed out the three variations in testing, each described in a bullet, and a second which demonstrated the options participants have to select from, each option shown in a bullet.
+The third list, which we have kept intact, is a list of the three research questions we pose to answer through our study.
+The authors could not identify a reason to remove the list structure for our research questions, as it clearly separates each of the distinct questions for the reader.
+There are four questions in the text of the paper: one at the very beginning, posing the question 'What do viewers see when we show them a data chart?', and three associated with the research questions.
+We have not removed any of these, as a question mark at the end of a research question is appropriate, and posing an opening question in the introduction is also appropriate.*
 
--   There is an overabundance of figures without clear emphasis.
+-   **There is an overabundance of figures without clear emphasis.**
 
@@
 
--   There are a couple of assumptions that assume certain factors (order of question, rounds) are negligible to the outcome.
-While I think this is reasonable, for a higher confidence in results it might be a good idea to test all visual stimulus within each round and block (if it is hard to be individually randomised) so that some participants get different question order.
+-   **There are a couple of assumptions that assume certain factors (order of question, rounds) are negligible to the outcome.
+While I think this is reasonable, for a higher confidence in results it might be a good idea to test all visual stimulus within each round and block (if it is hard to be individually randomised) so that some participants get different question order.**
 
@@
 
-    -   The title, abstract and conclusion suggests a wider context but the content is actually more limited in scope.
+    -   **The title, abstract and conclusion suggests a wider context but the content is actually more limited in scope.
 I suggest changing the title to emphasise the main points in the article: 1) nationally representative sample and 2) testing stacked barplots.
 I think statements like "Our study results provide actionable insights that data visualization practitioners can utilize in the design of data visualizations." extrapolates too much.
-It should be more specific in that the actionable insights is *for stacked barplots* and also write concretely what the recommendation is (i.e., vertical display, align groups that most important to compare, etc.).
+It should be more specific in that the actionable insights is *for stacked barplots* and also write concretely what the recommendation is (i.e., vertical display, align groups that most important to compare, etc.).**
+        -   *We have updated the statement in the conclusions per a specific recommendation in the minor comments.
+In addition, we have updated the paper title and some of the language in the abstract to reflect that the scope is limited to testing stacked bar charts.* **We will also update the recommendation language to be more concrete**
 
-    -   Do the authors have recommendations for (non-representative) visualisation perception surveys given their experience on nationally representative sample?
+    -   **Do the authors have recommendations for (non-representative) visualisation perception surveys given their experience on nationally representative sample?
 E.g.
-given that the surveys conducted via Mturks etc are biased toward highly educated young males and authors found differences between the income group, what do these non-representative survey inform us?
+given that the surveys conducted via Mturks etc are biased toward highly educated young males and authors found differences between the income group, what do these non-representative survey inform us?**
 
@@
 
-    -   I think the paper can benefit in the intro to include a stronger emphasis why we want to achieve a representative sample.
+    -   **I think the paper can benefit in the intro to include a stronger emphasis why we want to achieve a representative sample.
 The authors do address this in part but it reads like an aside ("..., a common target audience for data visualisation").
@@
 results on polls in the news -- this can also lead to why the authors chose to focus on barplots too as polls are often count data).
-Authors can also discuss how it is **imperative to be inclusive** as these data visualisation can be informative and the results need to be accessible by the *whole* population and not just the elite/highly educated population (which seem to be the bias towards at lot of these surveys have).
+Authors can also discuss how it is *imperative to be inclusive* as these data visualisation can be informative and the results need to be accessible by the *whole* population and not just the elite/highly educated population (which seem to be the bias towards at lot of these surveys have).
 I think also a stronger emphasis on the importance of the authors' works is needed, e.g., "this is the first study to ...".
 Also in conclusion, the authors state "The size of our sample aids in identifying this signal.", but the authors can instead state e.g.
-"our study is orders of magnitude larger than other studies with similar scope, demonstrating...".
+"our study is orders of magnitude larger than other studies with similar scope, demonstrating...".**
+        -   *There are multiple points here; we will address each.*
+        -   **Stronger emphasis on why we want to achieve a representative sample**
+        -   **Stronger emphasis on the importance of the authors' work is needed**
 
-    -   While I think the content is fine, I think the grammar in the paper can be better polished.
+    -   **While I think the content is fine, I think the grammar in the paper can be better polished.
 (It is not that the current grammar is bad, but I just think it can be improved to make this a better paper).
-I have outlined some of the suggested corrections below.
+I have outlined some of the suggested corrections below.**
 
         -   page 2, second paragraph: "The viewer's employment of comparisons of the components" -\> "The comparison of components"
+            -   *This change has been made in the text.*
 
         -   page 2, third paragraph: "to be biased towards more male, younger and relatively higher education" -\> "to be biased towards more male, younger and relatively well-educated individuals"
+            -   *This change has been made in the text.*
 
-        -   page 2, third paragraph: "Further, the populations' emphasis..." -\> This sentence was confusing.
+        -   **page 2, third paragraph: "Further, the populations' emphasis..." -\> This sentence was confusing.
 I kind of get the gist what the authors are trying to say but I wonder if it can be paraphrased better?
 E.g.
-"Arguably, we could limit the population of interest to highly educated individuals, however, this group is also likely to have prior exposure to data analysis in higher education, which may not be the case for other groups in the wider population."
+"Arguably, we could limit the population of interest to highly educated individuals, however, this group is also likely to have prior exposure to data analysis in higher education, which may not be the case for other groups in the wider population."**
 
         -   page 2, fourth paragraph: I"m not sure if you should start the paragraph with "Here".
+            -   *We have replaced 'Here' with 'In this work'*
 
-    -   FIgure 1, the text within the visualisation are too small to see. Perhaps lay it out as a 3 x 2 instead?
+    -   **FIgure 1, the text within the visualisation are too small to see. Perhaps lay it out as a 3 x 2 instead?**
 
-    -   page 8, first paragraph: "\$ 2\$" is a typo?
+    -   **page 8, first paragraph: "\$ 2\$" is a typo?**
+        -   *This has been resolved in the source code.* **Heike, please confirm in the rendered paper.**
 
-    -   Table 1, effective sample size adds up to 1520.3 but the total says 1520.8?
+    -   **Table 1, effective sample size adds up to 1520.3 but the total says 1520.8?**
 
-    -   Figure 3: I take it that the American Community Survey 2021 is considered the Census estimates?
+    -   **Figure 3: I take it that the American Community Survey 2021 is considered the Census estimates?
 I think it makes more sense to swap the bar and point so that the bar is the Census estimate.
@@
 In this plot, you've facetted by the stimulus type but I think it makes sense to facet by the category level (the comparison of distribution is done with the census estimate directly, and I think the interest is to compare the distribution across stimulus so you want the primary factor of interest to compare to be close in proximity).
-Within a category level, the census estimate is constant across stimulus so you could substitute is with a horizontal line (instead of a bar or point) as well.
+Within a category level, the census estimate is constant across stimulus so you could substitute is with a horizontal line (instead of a bar or point) as well.**
 
-    -   Figure 4: I think (b) alone is sufficient as you can gain the same information as (a) (although perhaps it is less familiar for wider audience).
-You could arguably put figure (a) information in a table instead.
+    -   **Figure 4: I think (b) alone is sufficient as you can gain the same information as (a) (although perhaps it is less familiar for wider audience).
+You could arguably put figure (a) information in a table instead.**
 
     -   Rephrase needed on page 18 third paragraph "... asking directly asking ...".
+        -   *We have removed the extraneous 'asking'.*
 
@@
 I think the authors should say "design of stacked barplots" instead.
+        -   *We have updated the language here to say 'design of stacked bar charts'.*
 
     -   In the source code of the paper, authors use 1.96 and 1.645 for 0.875 and 0.95 quantiles but I think the authors should use `qnorm(0.975)` and `qnorm(0.95)` (I realize it doesn't make much difference in the end but I think you should avoid rounding unnecessary until the presentation of the results).
+        -   *This change has been made in the source code.*
 
-    -   I do appreciate that the authors present the results using vertical stacked barplot, reflected from their results.
+    -   **I do appreciate that the authors present the results using vertical stacked barplot, reflected from their results.
 A minor point is that I actually don't like the angled display of the x-axis text and I would have preferred to present this in a horizontal display so the text stays horizontal.
@@
 In that sense, horizontal display where axis text can be better presented may be a better alternative?
-Authors could perhaps include a discussion along this line and I'd be interested in knowing what the authors think/recommend.
+Authors could perhaps include a discussion along this line and I'd be interested in knowing what the authors think/recommend.**
 
@@
 
--   Please add an unnumbered section Supplementary Material to list its content (e.g., data, code, additional simulation results, technical proofs, etc.).
+-   **Please add an unnumbered section Supplementary Material to list its content (e.g., data, code, additional simulation results, technical proofs, etc.).
 See recently published papers for examples.
-Reproducibility is held high at the JDS; the reproducibility team will check for reproducibility when the paper is accepted.
+Reproducibility is held high at the JDS; the reproducibility team will check for reproducibility when the paper is accepted.**
 
@@
 
--   In general, both reviewers express some concerns regarding some of the language and description regarding overall impact of the work.
-I would encourage the authors to make it more clear that their primary contribution is in using a representative probability sample, and that their conclusions are focused on stacked barplot displays, as Reviewer 2 specifically remarks, and highlight why conclusions from a non-representative sample may be problematic within the data visualization world.
+-   **In general, both reviewers express some concerns regarding some of the language and description regarding overall impact of the work.
+I would encourage the authors to make it more clear that their primary contribution is in using a representative probability sample, and that their conclusions are focused on stacked barplot displays, as Reviewer 2 specifically remarks, and highlight why conclusions from a non-representative sample may be problematic within the data visualization world.**
 
--   I believe the figures are fairly clear (per Reviewer 1's comment), but the authors might also be inclined to be more selective with what is shown in Figures 3, 6-8, and maybe move some of the others to an appendix.
+-   **I believe the figures are fairly clear (per Reviewer 1's comment), but the authors might also be inclined to be more selective with what is shown in Figures 3, 6-8, and maybe move some of the others to an appendix.**
 
--   Can the authors comment further on modeling assumptions, particularly in relation to presenting all participants with the unaligned plot first followed by the aligned plot second (as opposed to randomizing this)?
+-   **Can the authors comment further on modeling assumptions, particularly in relation to presenting all participants with the unaligned plot first followed by the aligned plot second (as opposed to randomizing this)?
 For instance, perhaps some participants notice that the plots are the same, with just the bars stacked in a different order, and their response to the second task is biased to be equivalent to that of the first task.
-This seems like it would be more of a problem if the plots were presented in the reverse order, but I'm curious if the authors looked into this.
+This seems like it would be more of a problem if the plots were presented in the reverse order, but I'm curious if the authors looked into this.**
+    -   *We acknowledge that this is a limitation.
+There is a chance that someone recognizes they are completing the same task with the same bars when we present the aligned task before the unaligned task; we made the assumption that the chance of recognizing the tasks is lower when seeing unaligned prior to aligned.
+Participants are not able to return to the prior question once they move on to the second question, which does make it more difficult for participants to compare the tasks to one another.* **We will also add some text to the paper to reflect this limitation a little more clearly**
 
--   Is there a reason why the meaning/values of i from 1 to 3 are reversed between the two models specified in Equations 4 and 7?
+-   **Is there a reason why the meaning/values of i from 1 to 3 are reversed between the two models specified in Equations 4 and 7?**
 
--   It appears the cumulative logistic regression model with respect to participant demographic covariates just considers first-order effects, but interactions would also be pretty interesting here.
-Are there any notable interaction effects?
+-   **It appears the cumulative logistic regression model with respect to participant demographic covariates just considers first-order effects, but interactions would also be pretty interesting here.
+Are there any notable interaction effects?**
 
--   The letter notation used to denote significance is a bit confusing to me (e.g., Figure 5, Table 3), can the authors clarify this within the text or modify accordingly?
+-   **The letter notation used to denote significance is a bit confusing to me (e.g., Figure 5, Table 3), can the authors clarify this within the text or modify accordingly?
 What do a, b, c, etc.
 actually mean?
-If this is convention from Piepho (2004), it would be useful for this to be explained more or changed for a more general audience.
+If this is convention from Piepho (2004), it would be useful for this to be explained more or changed for a more general audience.**

### 2023-10-16 SDSS-2023/testing-charts-accuracy.qmd
addressing language-based comments from reviewers in round 1

@@
 ---
-title: "Testing Charts: viewers' perceptual accuracy in surveys"
+title: "Testing Stacked Bar Charts: viewers' perceptual accuracy in surveys"
 short-title: Testing charts in surveys # running title
@@
 Decisions about the design of a data visualization should be informed by what design elements best support the audience's ability to perceive and understand the components of the data visualization.
-We build on the foundations of Cleveland and McGill's work in graphical perception, employing a large, nationally-representative, probability-based panel of survey respondents to test perception in statistical charts.
+We build on the foundations of Cleveland and McGill's work in graphical perception, employing a large, nationally-representative, probability-based panel of survey respondents to test perception in stacked bar charts.
 Our findings provide actionable guidance for data visualization practitioners to employ in their work."
@@
 ```{=latex}
-  \newcommand{\hh}[1]{{\textcolor{orange}{#1}}} 
-  \newcommand{\kr}[1]{{\textcolor{teal}{#1}}}.  
-  \newcommand{\ejm}[1]{{\textcolor{ForestGreen}{#1}}}.  
-  \newcommand{\ndt}[1]{{\textcolor{purple}{#1}}}.
+  \newcommand{\rev1}[1]{{\textcolor{teal}{#1}}}
+  \newcommand{\rev2}[1]{{\textcolor{purple}{#1}}}
+  \newcommand{\editors}[1]{{\textcolor{ForestGreen}{#1}}}
 
@@
 
-The viewer's employment of comparisons of the components is a crucial step in the process of interacting with and understanding the chart.
+The comparisons of components is a crucial step in the process of interacting with and understanding the chart.
 @clevelandGraphicalPerceptionTheory1984 observed as such, and in their seminal study defined the better visual among a pair as the one that allows viewers to make more accurate comparisons.
@@
 @heerCrowdsourcingGraphicalPerception2010 reproduced Cleveland and McGill's rankings using a larger sample from a crowd sourcing platform, employing a total of 82 Amazon Mechanical Turk (mTurk) workers for the study.
-Crowd sourced samples were shown by @borgoCrowdsourcingInformationVisualization2017 to be biased towards more male, younger, and relatively higher education relative to the adult U.S.
+Crowd sourced samples were shown by @borgoCrowdsourcingInformationVisualization2017 to be biased towards more male, younger, and relatively well-educated individuals relative to the adult U.S.
 population as a whole; generally, convenience and opt-in samples are not representative of the general population, a common target audience for data visualization and scientific communication work.
@@
 
-Here, we seek to answer whether it is possible to reproduce some of the previous findings in the context of a survey with a large, nationally-representative set of respondents.
+In this work, we seek to answer whether it is possible to reproduce some of the previous findings in the context of a survey with a large, nationally-representative set of respondents.
 Specifically, we present viewers with structural variations of bar charts and ask them to answer questions comparing the size of elements within those charts.
@@
 
-We ask participants to determine which of two just-noticeably different marked pieces (tiles) is larger within each chart, and focus on three main sets of structural variation in the design of that chart:
-
--   First, we vary the alignment of the pieces in question.
+We ask participants to determine which of two just-noticeably different marked pieces (tiles) is larger within each chart, and focus on three main sets of structural variation in the design of that chart.
+First, we vary the alignment of the pieces in question.
 Viewers are presented with two marked pieces in a chart that do not share a common baseline (aligned pieces), then two pieces that do share a common baseline (unaligned pieces).
-
--   Second, we vary the orientation of the chart -- we rotate the vertical stacked bar chart, and present a horizontally oriented version of the same chart, with identically sized marked pieces.
-
--   Finally, we change the aspect ratio of the chart and present a wider version of the horizontally oriented chart which has longer, but thinner, marked pieces.
+Second, we vary the orientation of the chart -- we rotate the vertical stacked bar chart, and present a horizontally oriented version of the same chart, with identically sized marked pieces.
+Finally, we change the aspect ratio of the chart and present a wider version of the horizontally oriented chart which has longer, but thinner, marked pieces.
 
@@
 
-"There are many charts used in the news media to portray data visually. Looking at the chart below, which of the marked dark blue pieces is bigger, A or B? Just your best guess is fine.
-
--   A is bigger
-
--   B is bigger
-
--   They are the same"
+"There are many charts used in the news media to portray data visually.
+Looking at the chart below, which of the marked dark blue pieces is bigger, A or B?
+Just your best guess is fine.
+[] A is bigger; [] B is bigger; [] They are the same."
 
@@
 We *combine* (rather than cumulate) a set of $\ell$ surveys $S_1$, $S_2$, ...
-$S_\ell$ (with $ \ell \ge 2$ ) as described in @omuircheartaighCombiningSamplesVs2002, by multiplying weights in $S_i$ by $\lambda_i$, for $1 \le i \le \ell$.
+$S_\ell$ (with $\ell \ge 2$) as described in @omuircheartaighCombiningSamplesVs2002, by multiplying weights in $S_i$ by $\lambda_i$, for $1 \le i \le \ell$.
 $\lambda_i \in [0,1]$ with $\sum_i \lambda_i = 1$ is given as 
@@
 omni_gender <- omni_gender %>% mutate(
-  low = mean - 1.6445*SE, # for 90% MoE 
-  high = mean + 1.6445*SE
+  low = mean - qnorm(0.95)*SE, # for 90% MoE 
+  high = mean + qnorm(0.95)*SE
 )
@@
 omni_age <- omni_age %>% mutate(
-  low = mean - 1.6445*SE, # for 90% MoE 
-  high = mean + 1.6445*SE
+  low = mean - qnorm(0.95)*SE, # for 90% MoE 
+  high = mean + qnorm(0.95)*SE
 )
@@
 omni_educ <- omni_educ %>% mutate(
-  low = mean - 1.6445*SE, # for 90% MoE 
-  high = mean + 1.6445*SE
+  low = mean - qnorm(0.95)*SE, # for 90% MoE 
+  high = mean + qnorm(0.95)*SE
 )
@@
 omni_income <- omni_income %>% mutate(
-  low = mean - 1.6445*SE, # for 90% MoE 
-  high = mean + 1.6445*SE
+  low = mean - qnorm(0.95)*SE, # for 90% MoE 
+  high = mean + qnorm(0.95)*SE
 )
@@
       xmin = 2:1 - 0.45,
-      ymin = correctTRUE - 1.96* se.correctTRUE,
+      ymin = correctTRUE - qnorm(0.975)* se.correctTRUE,
       xmax = 2:1 + 0.45,
-      ymax = correctTRUE + 1.96* se.correctTRUE,
+      ymax = correctTRUE + qnorm(0.975)* se.correctTRUE,
       weight = 1),
@@
       xmin = as.numeric(chart) - 0.45,
-      ymin = correct - 1.96* `se.as.numeric(correct)`,
+      ymin = correct - qnorm(0.975)* `se.as.numeric(correct)`,
       xmax = as.numeric(chart) + 0.45,
-      ymax = correct + 1.96* `se.as.numeric(correct)`,
+      ymax = correct + qnorm(0.975)* `se.as.numeric(correct)`,
     ), 
@@
       xmin = as.numeric(chart) - 0.45,
-      ymin = correct3 - 1.96* `se.as.numeric(correct3)`,
+      ymin = correct3 - qnorm(0.975)* `se.as.numeric(correct3)`,
       xmax = as.numeric(chart) + 0.45,
-      ymax = correct3 + 1.96* `se.as.numeric(correct3)`,
+      ymax = correct3 + qnorm(0.975)* `se.as.numeric(correct3)`,
     ), 
@@
   mutate(
-    ci_low = exp(estimate-1.96*std.error),
-    ci_high = exp(estimate+1.96*std.error),
+    ci_low = exp(estimate-qnorm(0.975)*std.error),
+    ci_high = exp(estimate+qnorm(0.975)*std.error),
     estimate = exp(estimate)
@@
       xmin = as.numeric(response3) - 0.45,
-      ymin = zoom2 - 1.96* `se.as.numeric(zoom)`,
+      ymin = zoom2 - qnorm(0.975)* `se.as.numeric(zoom)`,
       xmax = as.numeric(response3) + 0.45,
-      ymax = zoom2 + 1.96* `se.as.numeric(zoom)`,
+      ymax = zoom2 + qnorm(0.975)* `se.as.numeric(zoom)`,
     ), 
@@
   geom_errorbar(
-    aes(ymin = (estimate-1.96*std.error),
-        ymax = (estimate+1.96*std.error)),
+    aes(ymin = (estimate-qnorm(0.975)*std.error),
+        ymax = (estimate+qnorm(0.975)*std.error)),
     width=0.3,
@@
     educ5 = factor(levels(round_123$educ5), levels = levels(round_123$educ5)),
-    low = estimate-1.96*std.error,
-    high = estimate+1.96*std.error,
+    low = estimate-qnorm(0.975)*std.error,
+    high = estimate+qnorm(0.975)*std.error,
     estimate = inv_logit(estimate),
@@
     gender = factor(levels(round_123$gender), levels = levels(round_123$gender)),
-    low = estimate-1.96*std.error,
-    high = estimate+1.96*std.error,
+    low = estimate-qnorm(0.975)*std.error,
+    high = estimate+qnorm(0.975)*std.error,
     estimate = inv_logit(estimate),
@@
     income4 = factor(levels(round_123$income4), levels = levels(round_123$income4)),
-    low = estimate-1.96*std.error,
-    high = estimate+1.96*std.error,
+    low = estimate-qnorm(0.975)*std.error,
+    high = estimate+qnorm(0.975)*std.error,
     estimate = inv_logit(estimate),
@@
     age4 = factor(levels(round_123$age4), levels = levels(round_123$age4)),
-    low = estimate-1.96*std.error,
-    high = estimate+1.96*std.error,
+    low = estimate-qnorm(0.975)*std.error,
+    high = estimate+qnorm(0.975)*std.error,
     estimate = inv_logit(estimate),
@@
     term = gsub("income4", "", term),
-    ci_low = exp(estimate-1.96*std.error),
-    ci_high = exp(estimate+1.96*std.error),
+    ci_low = exp(estimate-qnorm(0.975)*std.error),
+    ci_high = exp(estimate+qnorm(0.975)*std.error),
     estimate = exp(estimate)
@@
 Further, we can also study respondent behavior and how viewers interact with the chart.
-Paradata on viewer interaction with the chart (e.g., zooming and time spent on each task) can be collected as well as asking directly asking respondents questions about their certainty.
+Paradata on viewer interaction with the chart (e.g., zooming and time spent on each task) can be collected as well as directly asking respondents questions about their certainty.
 Not only can we collect this information, but we can also identify significant differences in this behavior across different designs.
@@
 
-Our study results provide actionable insights that data visualization practitioners can utilize in the design of data visualizations.
+Our study results provide actionable insights that data visualization practitioners can utilize in the design of stacked bar charts.
 Comparisons of interest should be lined up and share a common baseline within a chart; in an interactive setting, it may be beneficial to give viewers the possibility of re-ordering categories to support effective comparison across groupings.

### 2023-10-16 revisions-round1/response-to-reviewers.qmd
language addition in conclusion per reviewer 2

@@
 
-    -   **Do the authors have recommendations for (non-representative) visualisation perception surveys given their experience on nationally representative sample?
+    -   Do the authors have recommendations for (non-representative) visualisation perception surveys given their experience on nationally representative sample?
 E.g.
-given that the surveys conducted via Mturks etc are biased toward highly educated young males and authors found differences between the income group, what do these non-representative survey inform us?**
+given that the surveys conducted via Mturks etc are biased toward highly educated young males and authors found differences between the income group, what do these non-representative survey inform us?
+        -   *We have added the following text to our conclusions to address this point: "If studies using non-representative or crowd-sourced samples find non-significant results when studying elements of design, those results may not actually be non-significant for other populations, particularly lower education or lower income populations.
+Researchers should not disregard non-significant findings in those survey studies but rather ask whether the lack of significance would hold across population subgroups."*
 

### 2023-10-16 SDSS-2023/testing-charts-accuracy.qmd
language addition in conclusion per reviewer 2

@@
 In particular, the differences across educational attainment groups and income groups underscore the importance of utilizing a representative population when testing perception; prior results in this area may be absorbing bias in responses due to the larger representation of individuals with higher education levels among those study respondents.
+If studies using non-representative or crowd-sourced samples find non-significant results when studying elements of design, those results may not actually be non-significant for other populations, particularly lower education or lower income populations.
+Researchers should not disregard non-significant findings in those survey studies but rather ask whether the lack of significance would hold across population subgroups.
 

### 2023-10-23 revisions-round1/response-to-reviewers.qmd
first couple of bullet points

@@
 
-    -   **FIgure 1, the text within the visualisation are too small to see. Perhaps lay it out as a 3 x 2 instead?**
+    -   **Figure 1, the text within the visualisation are too small to see. Perhaps lay it out as a 3 x 2 instead?**
+            - HH: *Figure 1 was not meant to have readable text within the stimuli visualizations but only give an overview of the different stimuli and the testing situations.
+We have increased the font size for the overview, expanded the discussion of the figure to reflect its purpose and added image (a) to the supplement in a larger format.* **create supplement and add stimulus**
 
     -   **page 8, first paragraph: "\$ 2\$" is a typo?**
-        -   *This has been resolved in the source code.* **Heike, please confirm in the rendered paper.**
+        -   *This has been resolved in the source code.* **Heike, please confirm in the rendered paper.** HH: yes, it's a typo, it was supposed to be the expression l <= 2 and the latex expression $\ell \ge 2$ did not render properly
 
-    -   **Table 1, effective sample size adds up to 1520.3 but the total says 1520.8?**
+    -   **Table 1, effective sample size adds up to 1520.3 but the total says 1520.8?** *HH: What is shown is the effective sample size of the combined survey rounds, not the total of each round's effective sample size.
+    The fact that these numbers are close is due to similar weighting strategies being used in each of the individual rounds.
+We have now renamed the line "Combined Rounds".
+* **and added a clarifying footnote in the table**
 
@@
     -   **Figure 4: I think (b) alone is sufficient as you can gain the same information as (a) (although perhaps it is less familiar for wider audience).
-You could arguably put figure (a) information in a table instead.**
+You could arguably put figure (a) information in a table instead.**.
+
+    XXX HH: What should we do here? I'd be OK with putting all of this into a table.
 

### 2023-10-23 SDSS-2023/testing-charts-accuracy.qmd
fig 6. horizontal labels

@@
 ---
-title: "Testing Stacked Bar Charts: viewers' perceptual accuracy in surveys"
-short-title: Testing charts in surveys # running title
+title: "What do they see? Testing perceptual accuracy  in surveys of the of the general US population using stacked barcharts"
+short-title: Testing stacked barcharts in surveys # running title
 bibliography: references.bib
@@
 ```
+
+# To-do
+
+Into appendix: Stimulus question(s)
 
@@
 #| label: fig-tasks
-#| fig-cap: "Each of the visual stimuli presented to viewers. In each bar chart, two pieces are marked. The larger piece in each chart are pieces B and D. "
-#| out-width: 90%
-knitr::include_graphics("images/stimuli-overview.png")
+#| fig-cap: "Overview of all of the visual stimuli presented to viewers. In each bar chart, two pieces are marked. The larger piece in each chart are pieces B and D. "
+#| out-width: 100%
+knitr::include_graphics("images/stimuli-overview-1.png")
 ```
@@
 | Round 3   | June 2022  | `r nrow(round3_adj)` | `r round(neff(round3_adj$weight),1)` | `r round(sum(round3_adj$weight),1)` |        $\lambda_3$ = `r round(lambdas[3],3)` |
-| **Total** | ---        |  `r nrow(round_123)` |  `r round(neff(round_123$weight),1)` |                                     |                                              |
+| **Combined Rounds** | ---        |  `r nrow(round_123)` |  `r round(neff(round_123$weight),1)` |                                     |                                              |
 
@@
          zoom2 = zoom, 
-        response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong")))) %>%
+        response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+        response3 = fct_recode(response3, "They are\nthe same" = "They are the same")) %>%
   dplyr::select(-c("zoom"))
@@
          zoom2 = zoom, 
-        response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong")))) %>%
+        response3 = factor(response3, 
+                           levels=rev(c("Correct", "They are the same", "Wrong"))),
+        response3 = fct_recode(response3, "They are\nthe same" = "They are the same")) %>%
   dplyr::select(-c("zoom"))
@@
     response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    response3 = fct_recode(response3, "They are\nthe same" = "They are the same"),
     task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
@@
   facet_grid(.~task) + 
-  theme(axis.text.x = element_text(angle=30, hjust = 1)) + 
+#  theme(axis.text.x = element_text(angle=30, hjust = 1)) + 
   scale_fill_manual("Zoom", values=c("grey", "#1274F3")) +
@@
     response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
+    response3 = fct_recode(response3, "They are\nthe same"="They are the same"),
     task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
@@
   ggtitle("\nby chart") +
-  theme(axis.text.x = element_text(angle=30, hjust = 1)) +
+#  theme(axis.text.x = element_text(angle=30, hjust = 1)) +
   scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), 

### 2023-10-23 SDSS-2023/testing-charts-accuracy.qmd
figure 6 labels

@@
 #| label: fig-tasks
-#| fig-cap: "Overview of all of the visual stimuli presented to viewers. In each bar chart, two pieces are marked. The larger piece in each chart are pieces B and D. "
+#| fig-cap: "Overview of all of the visual stimuli presented to viewers.
+In each bar chart, two pieces are marked.
+The larger piece in each chart are pieces B and D.
+Please refer to the appendix for full size images of the stimuli."
 #| out-width: 100%
@@
 | Round 3   | June 2022  | `r nrow(round3_adj)` | `r round(neff(round3_adj$weight),1)` | `r round(sum(round3_adj$weight),1)` |        $\lambda_3$ = `r round(lambdas[3],3)` |
-| **Combined Rounds** | ---        |  `r nrow(round_123)` |  `r round(neff(round_123$weight),1)` |                                     |                                              |
-
-: Survey rounds: dates, number of participants (nominal sample size), effective sample size, sum of weights, and factors for the combining of surveys as discussed above. {#tbl-rounds}
+| **Combined** | ---        |  `r nrow(round_123)` |  `r round(neff(round_123$weight),1)`* |                                     |                                              |
+
+: Survey rounds: dates, number of participants (nominal sample size), effective sample size, sum of weights, and factors for the combining of surveys as discussed above.
+{#tbl-rounds}
 
@@
 There are significantly higher rates of zooming by respondents for the vertical task compared to both horizontal tasks; rates of zooming do not differ by correctness of response or alignment in the stimulus."
-#| fig-height: 4.25
-#| fig-width: 7
-#| out-width: 80%
+#| fig-height: 3.5
+#| fig-width: 10
+#| out-width: 90%
 
@@
   facet_grid(.~task) + 
-#  theme(axis.text.x = element_text(angle=30, hjust = 1)) + 
+  theme(axis.text.x = element_text(angle=0, vjust = 0.5)) + 
   scale_fill_manual("Zoom", values=c("grey", "#1274F3")) +
@@
   ggtitle("\nby chart") +
+  theme(axis.text.x = element_text(angle=0, vjust = 0.5)) + 
 #  theme(axis.text.x = element_text(angle=30, hjust = 1)) +
@@
 
-zt + zd + plot_layout(widths=c(5, 7.5), guides='collect')
+zt + zd + plot_layout(widths=c(6, 9), guides='collect')
 

### 2023-10-23 SDSS-2023/testing-charts-accuracy.qmd
connected letters

@@
 The overlaid rectangles represent 95% confidence intervals.
-The letters in blue and orange encode significances between pairwise proportions: two bars have a significantly different proportion (at a 5% significance level) if they do not share a letter.
+The letters in blue and orange encode significances between pairwise proportions [@piephoAlgorithmLetterBasedRepresentation2004]: two bars have a significantly different proportion (at a 5% significance level) if they do not have a letter in common.
 There is no significant difference between the designs for wrong responses."
@@
 where $\ell \in \{1, 2\}$; $i \in \{1, 2\}$ is the comparison type (1 = Aligned, 2 = Unaligned), and $j \in \{1, 2, 3\}$ is the chart design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide.
+The letters allow us to assess whether differences between two means are significant:  two estimates within the same column are significantly different (at the 5% level) if they do not have any letter in common.
+This method is called connected letters and has been introduced by @piephoAlgorithmLetterBasedRepresentation2004.
+In @fig-response-123, we see, that the proportions of wrong responses are significantly different between tasks, but designs within each task are not significantly different.
+For correct responses, all proportions are significantly different except for the proportion of correctly assessing the difference between unaligned rectangles in the wide horizontal design: their label of `ab` shows us that assessing the differences between unaligned rectangles in the wide horizontal design is not significantly better than in the horizontal design (labelled `a`), but also not significantly worse than in the vertical design (labelled `b`).
+
 The estimated values and 95% confidence intervals are shown in @tbl-rounds-123.
-The lowercase letters indicate significances between pairs of estimates: two estimates within the same column are significantly different at 5% level if they do not have any letter in common [@piephoAlgorithmLetterBasedRepresentation2004].
 
@@
 #| fig-height: 3.5
-#| fig-width: 10
+#| fig-width: 11
 #| out-width: 90%

### 2023-10-23 SDSS-2023/testing-charts-accuracy.qmd
figure 7 - horizontal

@@
     estimate = scale_x(estimate, from=c(1,5), to=c(0,1)),
-    std.error = std.error/4
+    std.error = std.error/4,
+    chart = factor(chart, levels=c("Horizontal wide", "Horizontal", "Vertical"))
   )
@@
 
-round_123_certainty_long %>%
+round_123_certainty_long <- round_123_certainty_long %>% 
+  mutate(
+    chart = fct_recode(chart, "Horizontal\nwide" = "Horizontal wide")
+  )
+coefs_certainty <- coefs_certainty %>% 
+  mutate(
+    chart = fct_recode(chart, "Horizontal\nwide" = "Horizontal wide")
+  )
+
+gg <- round_123_certainty_long %>%
   ggplot(aes(x = chart)) +
@@
            colour="grey40", linewidth=0.25) +
-  theme(
-    axis.text.x = element_text(angle=30, hjust=1)
-  ) + 
+  theme(axis.text.x = element_text(angle=0, vjust = 0.5)) + 
+  #theme(
+  #  axis.text.x = element_text(angle=30, hjust=1)
+  #) + 
   scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), 
@@
   ) + 
-  geom_text(aes(label=letters), y = -0.05, data = coefs_certainty) 
+  geom_text(aes(label=letters), y = -0.05, data = coefs_certainty) #+
+#  scale_x_discrete(breaks=c("Horizontal", "Horizontal\nwide", "Vertical"))
+
+#ggb <- ggplot2::ggplot_build(gg)
+
+gg
 ```

### 2023-10-23 SDSS-2023/testing-charts-accuracy.qmd
using lineranges for all estimates

@@
 
-# from ACS 2001
+# from ACS 2021
 # https://data.census.gov/table?q=sex+of+adults+age+greater+than+18+across+the+US&tid=ACSDP1Y2021.DP05	
@@
 
-gender <- round_123 %>% 
+gender_bak <- round_123 %>% 
   ggplot() + 
@@
   scale_linetype_manual(values=c(2,2,1))
+
+gender <- acs_gender %>%
+  ggplot(aes(x = gender)) +
+  geom_errorbar(aes(ymin = percent-moe_percent, ymax=percent+moe_percent)) +
+  geom_point(aes(y = percent)) +
+  geom_point(
+    aes(x = sex, y = value, colour=chart),
+    position = position_dodge(width = 0.5),
+    data = omni_gender %>% filter(group=="mean")
+  ) +
+  geom_linerange(
+    aes(
+      x = sex,
+      ymin = low,
+      ymax=high,
+      colour=chart
+    ),
+    position = position_dodge(width = 0.5),
+    data = omni_gender %>% pivot_wider(names_from=group, values_from=value)
+  )
 
@@
 There are significantly higher rates of zooming by respondents for the vertical task compared to both horizontal tasks; rates of zooming do not differ by correctness of response or alignment in the stimulus."
-#| fig-height: 3.5
+#| fig-height: 4
 #| fig-width: 11
@@
 #| fig-width: 10
-#| fig-height: 4.5
+#| fig-height: 4.25
 

### 2023-10-25 SDSS-2023/testing-charts-accuracy.qmd
title change

--- 2023-10-23 SDSS-2023/testing-charts-accuracy.qmd
@@
 ---
-title: "What do they see? Testing perceptual accuracy  in surveys of the of the general US population using stacked barcharts"
-short-title: Testing stacked barcharts in surveys # running title
+title: "What do they see? Testing perceptual accuracy in a U.S. general population survey using stacked bar charts"
+short-title: Testing charts in surveys # running title
 bibliography: references.bib
@@
 ```
-
-# To-do
-
-Into appendix: Stimulus question(s)
 
@@
 #| label: fig-tasks
-#| fig-cap: "Overview of all of the visual stimuli presented to viewers.
-In each bar chart, two pieces are marked.
-The larger piece in each chart are pieces B and D.
-Please refer to the appendix for full size images of the stimuli."
-#| out-width: 100%
-knitr::include_graphics("images/stimuli-overview-1.png")
+#| fig-cap: "Each of the visual stimuli presented to viewers. In each bar chart, two pieces are marked. The larger piece in each chart are pieces B and D. "
+#| out-width: 90%
+knitr::include_graphics("images/stimuli-overview.png")
 ```
@@
 | Round 3   | June 2022  | `r nrow(round3_adj)` | `r round(neff(round3_adj$weight),1)` | `r round(sum(round3_adj$weight),1)` |        $\lambda_3$ = `r round(lambdas[3],3)` |
-| **Combined** | ---        |  `r nrow(round_123)` |  `r round(neff(round_123$weight),1)`* |                                     |                                              |
-
-: Survey rounds: dates, number of participants (nominal sample size), effective sample size, sum of weights, and factors for the combining of surveys as discussed above.
-{#tbl-rounds}
+| **Total** | ---        |  `r nrow(round_123)` |  `r round(neff(round_123$weight),1)` |                                     |                                              |
+
+: Survey rounds: dates, number of participants (nominal sample size), effective sample size, sum of weights, and factors for the combining of surveys as discussed above. {#tbl-rounds}
 
@@
 The overlaid rectangles represent 95% confidence intervals.
-The letters in blue and orange encode significances between pairwise proportions [@piephoAlgorithmLetterBasedRepresentation2004]: two bars have a significantly different proportion (at a 5% significance level) if they do not have a letter in common.
+The letters in blue and orange encode significances between pairwise proportions: two bars have a significantly different proportion (at a 5% significance level) if they do not share a letter.
 There is no significant difference between the designs for wrong responses."
@@
 where $\ell \in \{1, 2\}$; $i \in \{1, 2\}$ is the comparison type (1 = Aligned, 2 = Unaligned), and $j \in \{1, 2, 3\}$ is the chart design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide.
-The letters allow us to assess whether differences between two means are significant:  two estimates within the same column are significantly different (at the 5% level) if they do not have any letter in common.
-This method is called connected letters and has been introduced by @piephoAlgorithmLetterBasedRepresentation2004.
-In @fig-response-123, we see, that the proportions of wrong responses are significantly different between tasks, but designs within each task are not significantly different.
-For correct responses, all proportions are significantly different except for the proportion of correctly assessing the difference between unaligned rectangles in the wide horizontal design: their label of `ab` shows us that assessing the differences between unaligned rectangles in the wide horizontal design is not significantly better than in the horizontal design (labelled `a`), but also not significantly worse than in the vertical design (labelled `b`).
-
 The estimated values and 95% confidence intervals are shown in @tbl-rounds-123.
+The lowercase letters indicate significances between pairs of estimates: two estimates within the same column are significantly different at 5% level if they do not have any letter in common [@piephoAlgorithmLetterBasedRepresentation2004].
 
@@
 There are significantly higher rates of zooming by respondents for the vertical task compared to both horizontal tasks; rates of zooming do not differ by correctness of response or alignment in the stimulus."
-#| fig-height: 3.5
-#| fig-width: 11
-#| out-width: 90%
+#| fig-height: 4.25
+#| fig-width: 7
+#| out-width: 80%
 
@@
          zoom2 = zoom, 
-        response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
-        response3 = fct_recode(response3, "They are\nthe same" = "They are the same")) %>%
+        response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong")))) %>%
   dplyr::select(-c("zoom"))
@@
          zoom2 = zoom, 
-        response3 = factor(response3, 
-                           levels=rev(c("Correct", "They are the same", "Wrong"))),
-        response3 = fct_recode(response3, "They are\nthe same" = "They are the same")) %>%
+        response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong")))) %>%
   dplyr::select(-c("zoom"))
@@
     response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
-    response3 = fct_recode(response3, "They are\nthe same" = "They are the same"),
     task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
@@
   facet_grid(.~task) + 
-  theme(axis.text.x = element_text(angle=0, vjust = 0.5)) + 
+  theme(axis.text.x = element_text(angle=30, hjust = 1)) + 
   scale_fill_manual("Zoom", values=c("grey", "#1274F3")) +
@@
     response3 = factor(response3, levels=rev(c("Correct", "They are the same", "Wrong"))),
-    response3 = fct_recode(response3, "They are\nthe same"="They are the same"),
     task = factor(task, levels=c("cd", "ab"), labels=c("Aligned", "Unaligned")),
@@
   ggtitle("\nby chart") +
-  theme(axis.text.x = element_text(angle=0, vjust = 0.5)) + 
-#  theme(axis.text.x = element_text(angle=30, hjust = 1)) +
+  theme(axis.text.x = element_text(angle=30, hjust = 1)) +
   scale_y_continuous("Percentage", breaks=seq(0,1, by=0.25), 
@@
 
-zt + zd + plot_layout(widths=c(6, 9), guides='collect')
+zt + zd + plot_layout(widths=c(5, 7.5), guides='collect')
 

### 2023-10-25 revisions-round1/response-to-reviewers.qmd

@@
 Authors could perhaps include a discussion along this line and I'd be interested in knowing what the authors think/recommend.**
+    HH: figure 6: changed labels to horizontal - a vertical display of the results allows a comparison across all tasks and designs.
+A wide layout of the chart allows us to keep the number of pages down.
+    XXX: Kiegan, could you make sure that the workaround actually works in the pdf? I think that my pdf looks different from the one you get.
+
+    I'm not sure how much we can really go into liking something and engaging with charts. 
+    Signals from aesthetic changes are usually quite small and people like different things. Finding aesthetic changes that suit everybody is rare.
 
@@
 
+HH *sample size*
+
 -   **The letter notation used to denote significance is a bit confusing to me (e.g., Figure 5, Table 3), can the authors clarify this within the text or modify accordingly?
@@
 If this is convention from Piepho (2004), it would be useful for this to be explained more or changed for a more general audience.**
+
+HH *we have expanded the discussion on the interpretation of connected letters, and included an example for their first use.*

### 2023-10-25 SDSS-2023/testing-charts-accuracy.qmd

--- 2023-10-23 SDSS-2023/testing-charts-accuracy.qmd
@@
 ---
-title: "What do they see? Testing perceptual accuracy  in surveys of the of the general US population using stacked barcharts"
-short-title: Testing stacked barcharts in surveys # running title
+title: "What do they see? Testing perceptual accuracy in a U.S. general population survey using stacked bar charts"
+short-title: Testing stacked bar charts in surveys # running title
 bibliography: references.bib

### 2023-10-25 revisions-round1/response-to-reviewers.qmd
a few additional wording updates, fixing rendering issue, cleaning up responses.

--- 2023-10-23 revisions-round1/response-to-reviewers.qmd
@@
 ## Reviewer 2
-
--   **There are a couple of assumptions that assume certain factors (order of question, rounds) are negligible to the outcome.
-While I think this is reasonable, for a higher confidence in results it might be a good idea to test all visual stimulus within each round and block (if it is hard to be individually randomised) so that some participants get different question order.**
 
@@
 
-        -   **page 2, third paragraph: "Further, the populations' emphasis..." -\> This sentence was confusing.
+        -   page 2, third paragraph: "Further, the populations' emphasis..." -\> This sentence was confusing.
 I kind of get the gist what the authors are trying to say but I wonder if it can be paraphrased better?
 E.g.
-"Arguably, we could limit the population of interest to highly educated individuals, however, this group is also likely to have prior exposure to data analysis in higher education, which may not be the case for other groups in the wider population."**
+"Arguably, we could limit the population of interest to highly educated individuals, however, this group is also likely to have prior exposure to data analysis in higher education, which may not be the case for other groups in the wider population."
+            -   *We have rephrased this section to make the point more clearly.*
 
@@
 
-    -   **page 8, first paragraph: "\$ 2\$" is a typo?**
-        -   *This has been resolved in the source code.* **Heike, please confirm in the rendered paper.** HH: yes, it's a typo, it was supposed to be the expression l <= 2 and the latex expression $\ell \ge 2$ did not render properly
+    -   page 8, first paragraph: "\$ 2\$" is a typo?
+        -   *This was a typesetting error and has been resolved in the source code.*
 
-    -   **Table 1, effective sample size adds up to 1520.3 but the total says 1520.8?** *HH: What is shown is the effective sample size of the combined survey rounds, not the total of each round's effective sample size.
+    -   Table 1, effective sample size adds up to 1520.3 but the total says 1520.8? 
+        -   *What is shown is the effective sample size of the combined survey rounds, not the total of each round's effective sample size.
 The fact that these numbers are close is due to similar weighting strategies being used in each of the individual rounds.
-We have now renamed the line "Combined Rounds".
-* **and added a clarifying footnote in the table**
+We have now renamed the line "Combined Rounds" to reduce confusion.*
 
@@
 
-    XXX HH: What should we do here? I'd be OK with putting all of this into a table.
+    XXX HH: What should we do here?
+I'd be OK with putting all of this into a table.
+KR: I'm ok with moving to a table, though I like having at least one visual to really highlight the difference between aligned/unaligned.
+This would help us be responsive to the other comments about overuse of visuals, though.
 
@@
 
-    -   **I do appreciate that the authors present the results using vertical stacked barplot, reflected from their results.
+    -   I do appreciate that the authors present the results using vertical stacked barplot, reflected from their results.
 A minor point is that I actually don't like the angled display of the x-axis text and I would have preferred to present this in a horizontal display so the text stays horizontal.
@@
 In that sense, horizontal display where axis text can be better presented may be a better alternative?
-Authors could perhaps include a discussion along this line and I'd be interested in knowing what the authors think/recommend.**
+Authors could perhaps include a discussion along this line and I'd be interested in knowing what the authors think/recommend.
+        -   *We have changed the labels to horizontal for easier reading in Figures 6 and 7, but retained the angled labels in Figure 8.
+A vertical display of the results in Figure 6 allows an easier comparison across all tasks and designs.
+We note that the reviewer makes a good point about their preferences; much of data visualization design guidance in the past has been driven by aesthetic preferences, and different people find different layouts more appealing.
+This is one of the reasons it is important to carefully study the impact of those types of preferences with a representative sample.
+Finding an aesthetic or design that suits everyone is rare.
+While we have not yet covered our upcoming work on this paper, this is one of the reasons we are also using our survey framework to study participant engagement and whether design changes impact participants' willingness to engage with open-ended and interpretation questions.
+Because our experimental design does not support answering that question here, we did not add a discussion on that in the body of our paper, but do intend to discuss it more in future work more focused on engagement.*
 
@@
 This seems like it would be more of a problem if the plots were presented in the reverse order, but I'm curious if the authors looked into this.**
+
     -   *We acknowledge that this is a limitation.
@@
 
+    -   KR: **Heike, please confirm** which order they should be in.
+
 -   **It appears the cumulative logistic regression model with respect to participant demographic covariates just considers first-order effects, but interactions would also be pretty interesting here.
@@
 
--   **The letter notation used to denote significance is a bit confusing to me (e.g., Figure 5, Table 3), can the authors clarify this within the text or modify accordingly?
+    -   HH *sample size*
+
+-   The letter notation used to denote significance is a bit confusing to me (e.g., Figure 5, Table 3), can the authors clarify this within the text or modify accordingly?
 What do a, b, c, etc.
 actually mean?
-If this is convention from Piepho (2004), it would be useful for this to be explained more or changed for a more general audience.**
+If this is convention from Piepho (2004), it would be useful for this to be explained more or changed for a more general audience.
+    -   *We have expanded the discussion on the interpretation of connected letters, and included an example for their first use.*

### 2023-10-25 SDSS-2023/testing-charts-accuracy.qmd
a few additional wording updates, fixing rendering issue, cleaning up responses.

@@
 ```{=latex}
-  \newcommand{\rev1}[1]{{\textcolor{teal}{#1}}}
-  \newcommand{\rev2}[1]{{\textcolor{purple}{#1}}}
-  \newcommand{\editors}[1]{{\textcolor{ForestGreen}{#1}}}
-
 % take out everything above later
@@
 population as a whole; generally, convenience and opt-in samples are not representative of the general population, a common target audience for data visualization and scientific communication work.
-Further, the populations' emphasis on higher education individuals also leads to results which hold for groups of individuals who may be more likely to have prior exposure to data visualization in the context of scientific communication, or more exposure to data topics in higher education, but may not hold across other groups within the population.
+Further, higher education individuals are likely to have prior exposure to data visualization or data topics in higher education, which may not be the case for other groups within the population; existing differences in exposure to data may drive differences in results across groups.
 
@@
 Aligned tiles are additionally anchored in the same position along the axis, i.e.
-the difference between their sizes can be reduced to a positional assessment, which in the absence of three-dimensional cues seems to be the general strategy viewers use when evaluating barcharts [@zacksReadingBarGraphs1998a].
+the difference between their sizes can be reduced to a positional assessment, which in the absence of three-dimensional cues seems to be the general strategy viewers use when evaluating bar charts [@zacksReadingBarGraphs1998a].
 Unaligned tiles do not share this anchor of position, i.e. the difference between their sizes needs at least two comparisons of positions.
-However, unaligned bars in a stacked barchart are not just assessments of the lengths of bars -- the other tiles in the chart provide a reference frame of the shared axis, which *should* help with an assessment of the tiles' sizes.
+However, unaligned bars in a stacked bar chart are not just assessments of the lengths of bars -- the other tiles in the chart provide a reference frame of the shared axis, which *should* help with an assessment of the tiles' sizes.
 As shown in the sketch of @fig-tasks-explain, this framing leads to a stimulus that is between Cleveland and McGill's task of 'position along the same but unaligned axis' and an assessment of length.
@@
 | Round 3   | June 2022  | `r nrow(round3_adj)` | `r round(neff(round3_adj$weight),1)` | `r round(sum(round3_adj$weight),1)` |        $\lambda_3$ = `r round(lambdas[3],3)` |
-| **Combined** | ---        |  `r nrow(round_123)` |  `r round(neff(round_123$weight),1)`* |                                     |                                              |
+| **Combined** | ---        |  `r nrow(round_123)` |  `r round(neff(round_123$weight),1)` |                                     |                                              |
 

### 2023-10-25 SDSS-2023/testing-charts-accuracy.qmd
comparison to ACS

--- 2023-10-25 SDSS-2023/testing-charts-accuracy.qmd
@@
 ---
-title: "What do they see? Testing perceptual accuracy in a U.S. general population survey using stacked bar charts"
-short-title: Testing stacked bar charts in surveys # running title
+title: "What do they see? Testing perceptual accuracy  in surveys of the of the general US population using stacked barcharts"
+short-title: Testing stacked barcharts in surveys # running title
 bibliography: references.bib
@@
 #| label: fig-rounds-demographics
-#| fig-cap: "Demographics of respondents in each of the tasks. The dots show percentages of the national demographics as estimated by the American Community Survery (ACS) 2021."
+#| fig-cap: "Overview of participants' demographics in each of the tasks: the colored dots show marginal percentages for demographics across the three tasks.
+In comparison, the dark lines show upper and lower 95% confidence of these demographics as estimated by the American Community Survery (ACS) 2021.
+Note that for a lot of these intervals the estimates are so close together that they appear as a single line.
+"
 #| fig-height: 7
@@
 
+rounds_col <- RColorBrewer::brewer.pal(name="Accent", n=8)[c(1,3,5)]
+#[1] "#7FC97F" "#BEAED4" "#FDC086" "#FFFF99" "#386CB0" "#F0027F" "#BF5B17" "#666666"
 gender <- acs_gender %>%
   ggplot(aes(x = gender)) +
-  geom_errorbar(aes(ymin = percent-moe_percent, ymax=percent+moe_percent)) +
-  geom_point(aes(y = percent)) +
+  geom_errorbar(aes(ymin = percent-moe_percent, ymax=percent+moe_percent, colour="ACS estimate")) +
+#  geom_point(aes(y = percent)) +
   geom_point(
+    size=3,
     aes(x = sex, y = value, colour=chart),
@@
   ) +
+  ylim(c(0, NA)) +
   geom_linerange(
@@
     ),
+    linewidth = 1, 
     position = position_dodge(width = 0.5),
-    data = omni_gender %>% pivot_wider(names_from=group, values_from=value)
-  )
+    data = omni_gender %>% pivot_wider(names_from=group, values_from=value),
+    show.legend = FALSE
+  ) +
+  ylab("Percent") + 
+  ggtitle("Gender") +
+#  theme(axis.text.x = element_text(angle=0, vjust = 0.5)) +
+  xlab("") +
+#  theme(legend.position ="none") +
+  scale_linetype_manual(values=c(2,2,1)) +
+  scale_color_manual(values=c("grey20", rounds_col))
 
@@
                                    20582187 + 21049271 + 21928936, 
-                                   73789098))
-acs_age <- acs_age %>% mutate(percent = estimate/sum(estimate)*100)
+                                   73789098),
+                      percent = c(6.9 + 9.2, 
+                                  6.8 + 6.6 + 6.2, 
+                                  6.2 + 6.4 + 6.7, 
+                                  22.4
+                                  ),
+                      moe_percent = c(sqrt(0.1^2+0.1^2),
+                              sqrt(0.1^2+0.1^2 + 0.1^2),
+                              sqrt(0.1^2+0.1^2 + 0.1^2),
+                              0.1) )
+acs_age <- acs_age %>% mutate(percent2 = estimate/sum(estimate)*100)
 
@@
 
-age <- round_123 %>% 
+age_bak <- round_123 %>% 
   ggplot(aes(x = age4)) + 
@@
 
+age <- acs_age %>%
+  ggplot(aes(x = age4)) +
+  geom_errorbar(aes(ymin = percent2-moe_percent, ymax=percent2+moe_percent, colour="ACS estimate")) +
+#  geom_point(aes(y = percent)) +
+  geom_point(
+    size=3,
+    aes(x = age, y = value, colour=chart),
+    position = position_dodge(width = 0.5),
+    data = omni_age %>% filter(group=="mean")
+  ) +
+  ylim(c(0, NA)) +
+  geom_linerange(
+    aes(
+      x = age,
+      ymin = low,
+      ymax=high,
+      colour=chart
+    ),
+    linewidth = 1,
+    position = position_dodge(width = 0.5),
+    data = omni_age %>% pivot_wider(names_from=group, values_from=value)
+  ) +
+  ylab("Percent") + 
+  ggtitle("Age groups") +
+  theme(axis.text.x = element_text(angle=0, vjust = 0.5)) +
+  xlab("") +
+  theme(legend.position ="none") +
+  scale_linetype_manual(values=c(2,2,1)) +
+  scale_color_manual(values=c("grey20", rounds_col))
+
 acs_educ5 <- data.frame(educ5 = c("Less than HS", "HS or equivalent", "Some college",  "Bachelor", "Post graduate"), 
-                      estimate = c(28687047, 69610758, 77707408, 50031690, 29454503))
-acs_educ5 <- acs_educ5 %>% mutate(percent = estimate/sum(estimate)*100)
+                      estimate = c(
+                      3636691 + 10793507 + 14256849, 
+                      9974372 + 59636386,
+                      13050667 + 45042031 + 19614710,
+                      3677359 + 46354331,
+                      29454503),
+                      moe_percent = c(sqrt(0.1^2+0.1^2+0.1^2),
+                              sqrt(0.1^2+0.1^2 ),
+                              sqrt(0.1^2+0.1^2 + 0.1^2),
+                              sqrt(0.1^2+0.2^2 ),
+                              0.1) 
+                      )
+acs_educ5 <- acs_educ5 %>% mutate(percent = estimate/(30339089+225152317)*100)
 
@@
 
-education <- round_123 %>% 
+education_bak <- round_123 %>% 
   ggplot(aes(x = educ5)) + 
@@
 
+education <- acs_educ5 %>%
+  mutate(educ5 = factor(educ5, levels=levels(omni_educ$educ))) %>%
+  ggplot(aes(x = educ5)) +
+  geom_errorbar(aes(ymin = percent-moe_percent, ymax=percent+moe_percent, colour="ACS estimate")) +
+#  geom_point(aes(y = percent)) +
+  geom_point(
+    size=3,
+    aes(x = educ, y = value, colour=chart),
+    position = position_dodge(width = 0.5),
+    data = omni_educ %>% filter(group=="mean")
+  ) +
+  ylim(c(0, NA)) +
+  geom_linerange(
+    aes(
+      x = educ,
+      ymin = low,
+      ymax=high,
+      colour=chart
+    ),
+    linewidth = 1, 
+    position = position_dodge(width = 0.5),
+    data = omni_educ %>% pivot_wider(names_from=group, values_from=value),
+    show.legend = FALSE
+  ) +
+  ylab("Percent") + 
+  ggtitle("Education") +
+#  theme(axis.text.x = element_text(angle=0, vjust = 0.5)) +
+  xlab("") +
+  theme(legend.position ="none") +
+  scale_linetype_manual(values=c(2,2,1)) +
+  scale_color_manual(values=c("grey20", rounds_col))
+
 acs_income <- data.frame(
@@
               .6*16.8+12.8,
-              16.3+7.9+9.8))
+              16.3+7.9+9.8),
+  moe_percent = c(sqrt(0.1^2+0.1^2+0.1^2+0.1^2+0.1^2),
+                  sqrt(0.1^2+0.1^2+0.1^2),
+                  sqrt(0.1^2 + 0.1^2),
+                  sqrt(0.1^2+0.1^2+0.1^2))
+  )
 
@@
 
-income <- round_123 %>% 
+income_bak <- round_123 %>% 
   ggplot(aes(x = income4)) + 
@@
   theme(legend.position = "none")
+
+income <- acs_income %>%
+  mutate(income4 = factor(income4, levels=levels(omni_income$income))) %>%
+  ggplot(aes(x = income4)) +
+  geom_errorbar(aes(ymin = percent-moe_percent, ymax=percent+moe_percent, colour="ACS estimate")) +
+#  geom_point(aes(y = percent)) +
+  geom_point(
+    size=3,
+    aes(x = income, y = value, colour=chart),
+    position = position_dodge(width = 0.5),
+    data = omni_income %>% filter(group=="mean")
+  ) +
+  ylim(c(0, NA)) +
+  geom_linerange(
+    aes(
+      x = income,
+      ymin = low,
+      ymax=high,
+      colour=chart
+    ),
+    linewidth = 1, 
+    position = position_dodge(width = 0.5),
+    data = omni_income %>% pivot_wider(names_from=group, values_from=value),
+    show.legend = FALSE
+  ) +
+  ylab("Percent") + 
+  ggtitle("Income") +
+#  theme(axis.text.x = element_text(angle=0, vjust = 0.5)) +
+  xlab("") +
+  theme(legend.position ="none") +
+  scale_linetype_manual(values=c(2,2,1)) +
+  scale_color_manual(values=c("grey20", rounds_col))
 

### 2023-10-25 SDSS-2023/testing-charts-accuracy.qmd
updated income table

@@
 
-acs_income <- data.frame(
-  income4 = c("< $30k", "$30k - $60k", 
-              "$60k - $100k", "more than $100k"), 
-  percent = c(6+3.9+7.5+ .5*7.8, 
-              .5*7.8 + 11.3 + .4*16.8, 
-              .6*16.8+12.8,
-              16.3+7.9+9.8),
-  moe_percent = c(sqrt(0.1^2+0.1^2+0.1^2+0.1^2+0.1^2),
-                  sqrt(0.1^2+0.1^2+0.1^2),
-                  sqrt(0.1^2 + 0.1^2),
-                  sqrt(0.1^2+0.1^2+0.1^2))
-  )
-
-# Household income
-# Less than $10,000	6.0
-# $10,000 to $14,999	3.9
-# $15,000 to $24,999	7.5
-# $25,000 to $34,999	7.8
-# $35,000 to $49,999	11.3
-# $50,000 to $74,999	16.8
-# $75,000 to $99,999	12.8
-# $100,000 to $149,999	16.3
-# $150,000 to $199,999	7.9
-# $200,000 or more	9.8
-
-# 
-# acs_income <- data.frame(
-#   income4 = c("< $30k", "$30k - $60k", 
-#               "$60k - $100k", "more than $100k"), 
-#   percent = c(1807928 + 2669815 + 10192511 + .5*16440260, 
-#               .5*16440260+ 21573641+ 2/3*17579780, 
-#               1/3*17579780 + 8347923+13881892, 
-#               21161141)/113654891*100 )
-# Population age 18 and over
-# FULL-TIME, YEAR-ROUND WORKERS WITH EARNINGS	113654891	79395	
-# $1 to $9,999 or loss	1807928	11027	1.6	0.1
-# $10,000 to $14,999	2669815	16330	2.3	0.1
-# $15,000 to $24,999	10192511	50628	9.0	0.1
-# $25,000 to $34,999	16440260	63750	14.5	0.1
-# $35,000 to $49,999	21573641	44746	19.0	0.1
-# $50,000 to $64,999	17579780	35183	15.5	0.1
-# $65,000 to $74,999	8347923	32204	7.3	0.1
-# $75,000 to $99,999	13881892	50191	12.2	0.1
-# $100,000 or more	21161141	100258	18.6	0.1
+acs_income_raw <- read_xlsx("../data/ACSDT1Y2021.B19001-income.xlsx", sheet="Data", skip = 1) %>% mutate(
+  estimate = parse_number(Estimate), 
+  moe = parse_number(`Margin of Error`)
+)
+acs_income_raw <- acs_income_raw %>% mutate(
+  income4 = c("Total", rep("< $30k", 5), rep("$30k - $60k", 5), rep("$60k - $100k", 2), rep("more than $100k", 4))
+)
+
+acs_income2 <- acs_income_raw %>% group_by(income4) %>% 
+  summarize(
+    moe = tidycensus::moe_sum(moe, estimate),
+    estimate = sum(estimate)
+  )
+acs_income2 <- acs_income2 %>% mutate(
+  percent = estimate/estimate[income4=="Total"]*100,
+  moe_percent = moe_prop(estimate, rep(estimate[income4=="Total"], n()), 
+                      100*moe, 100*rep(moe[income4=="Total"], n()))
+)
+acs_income <- acs_income2 %>%
+  filter(income4 != "Total") %>%
+  mutate(
+  income4 = factor(income4, levels = c("< $30k", "$30k - $60k", 
+              "$60k - $100k", "more than $100k"))
+)
 

### 2023-10-26 SDSS-2023/testing-charts-accuracy.qmd
figure 8 - horizontal labels

@@
 library(tidyverse)
+library(tidycensus)
 library(knitr)
@@
 levels(round_123$educ5) <- c("Less than HS", "HS or equivalent", "Some college", "Bachelor", "Post graduate")
+
+#levels(round_123$educ5) <- c("< HS", "HS", "Some\ncollege", "Bachelor", "Post\ngraduate")
 
@@
 
-acs_educ5 <- data.frame(educ5 = c("Less than HS", "HS or equivalent", "Some college",  "Bachelor", "Post graduate"), 
+acs_educ5 <- data.frame(educ5 = levels(round_123$educ5), 
                       estimate = c(
@@
 omni_educ <- omni_educ %>% mutate(
-  educ = factor(educ, levels = c("Less than HS", "HS or equivalent", "Some college", "Bachelor", "Post graduate"))
+  educ = factor(educ, levels = levels(round_123$educ5))
 )
@@
 )
+
+levels(acs_income$income4) <- levels(round_123$income4)
 
@@
 We observe that while zooming does help with the overall accuracy (which is in agreement with the findings by @luModelingJustNoticeable2022 about the physical size of stimuli), the increase is not significant.
-However, different designs lead to different rates of zooming: we observe in \autoref{fig-zoom-123} that when dealing with the vertical design, the rate of zooming is significantly higher than for the two horizontal designs.
+However, different designs lead to different rates of zooming: we observe in \autoref{fig-zoom-123} that when a significantly lower rate of respondents chose to make use of zooming when answering the horizontal designs compared to the vertical design.
 
@@
 
-```{r}
+```{r results="asis"}
 #| label: tbl-zoom-123
@@
 
-coefs_zoom$term <- c("$$\\hat{\\mu}$$", "$$\\text{They are the same } \\widehat{\\rho}_{2}$$", 
-                     "Wrong $$\\widehat{\\rho}_{3}$$", "Aligned $$\\widehat{\\tau}_{2}$$", 
-                     "Horizontal wide $$\\widehat{\\delta}_{2}$$", "Vertical $$\\widehat{\\delta}_{3}$$")
+# coefs_zoom$term <- c("$$\\hat{\\mu}$$", "$$\\text{They are the same } \\widehat{\\rho}_{2}$$", 
+#                      "Wrong $$\\widehat{\\rho}_{3}$$", "Aligned $$\\widehat{\\tau}_{2}$$", 
+#                      "Horizontal wide $$\\widehat{\\delta}_{2}$$", "Vertical $$\\widehat{\\delta}_{3}$$")
+
+coefs_zoom$term <- c("\\hat{\\mu}", 
+                         "\\text{They are the same } \\widehat{\\rho}_{2}", 
+                         "Wrong $$\\widehat{\\rho}_{3}$$", 
+                         "Aligned $$\\widehat{\\tau}_{2}$$", 
+                         "Horizontal wide $$\\widehat{\\delta}_{2}$$", 
+                         "Vertical $$\\widehat{\\delta}_{3}$$")
 
@@
 
-The resulting model estimates and confidence intervals are shown in \autoref{tbl-demographics}.
-When considering the easier (aligned) task, response patterns do not differ significantly by age, gender, or education level.
-We do observe, however, a significantly higher log odds of selecting the 'correct' or 'they are the same' response (or significantly lower log odds of selecting 'incorrect') among those with an income between \$60,000 and \$100,000.
-When we consider the more difficult (unaligned) task, we see significant separation in response patterns across gender, educational attainment, and income groups.
-We do not see significant differences in responses by age for either task.
-Interestingly, we see lower log odds of selecting the 'correct' response among those with higher educational attainment (those with a bachelor's degree or post graduate study) *and* lower log odds of selecting the 'incorrect' response; as income and educational attainment increase, respondents are more likely to select the 'they are the same' option during the difficult task.
-This speaks to the difficulty of the task, and respondents' interpretation of the task when it is more difficult to perceive small differences.
-
-<!--\hh{XXX the formatting isn't right yet here ... latex problem with large floats ... will fix it.}-->
-
 ```{r demographics-model}
@@
 
+thm <- 
+  theme(axis.text.x = element_text(angle=0, hjust = 0.5))
+
+thm2 <- 
+  theme(axis.text.x = element_text(angle=0, hjust = 0.5),
+        axis.title.y = element_blank(), 
+        axis.text.y = element_blank())
+
 educ_glm <- svyglm(correct~educ5-1, design=survey_123, family=quasibinomial(), subset = task=="cd")
@@
 
-gender_cd + age_cd  + educ_cd +income_cd + 
-    plot_layout(nrow=1, guides = "collect", widths = c(2,4,5,4))
-
-gender_ab + age_ab  + educ_ab +income_ab + 
-  plot_layout(nrow=1, guides = "collect", widths = c(2,4,5,4))
+educ_cd <- educ_cd + 
+  scale_x_discrete(labels=c("< HS", "HS", "Some\ncollege", "Bachelor", "Post\ngraduate"))
+educ_ab <- educ_ab + 
+  scale_x_discrete(labels=c("< HS", "HS", "Some\ncollege", "Bachelor", "Post\ngraduate"))
+
+income_cd <- income_cd +
+  scale_x_discrete(labels=c("< $30k", "$30-60k", "$60-100k", "≥ $100k"))
+income_ab <- income_ab +
+  scale_x_discrete(labels=c("< $30k", "$30-60k", "$60-100k", "≥ $100k"))
+
+gender_cd + thm + 
+ age_cd  + thm2 + 
+  educ_cd  + thm2 + 
+  income_cd + thm2 + 
+    plot_layout(nrow=1, guides = "collect", widths = c(2,4,5,4)) &
+  theme(legend.position = "bottom",
+        legend.box.margin = margin(t = -15, r = 0, b = 0, l = 0, unit = "pt"))
+
+gender_ab + thm + 
+  age_ab  + thm2 + 
+  educ_ab + thm2 +
+  income_ab + thm2 +
+  plot_layout(nrow=1, guides = "collect", widths = c(2,4,5,4)) &
+  theme(legend.position = "bottom", 
+        legend.box.margin = margin(t = -15, r = 0, b = 0, l = 0, unit = "pt"))
 ```
+
+The resulting model estimates and confidence intervals are shown in \autoref{tbl-demographics}.
+When considering the easier (aligned) task, response patterns do not differ significantly by age, gender, or education level.
+We do observe, however, a significantly higher log odds of selecting the 'correct' or 'they are the same' response (or significantly lower log odds of selecting 'incorrect') among those with an income between \$60,000 and \$100,000.
+When we consider the more difficult (unaligned) task, we see significant separation in response patterns across gender, educational attainment, and income groups.
+We do not see significant differences in responses by age for either task.
+Interestingly, we see lower log odds of selecting the 'correct' response among those with higher educational attainment (those with a bachelor's degree or post graduate study) *and* lower log odds of selecting the 'incorrect' response; as income and educational attainment increase, respondents are more likely to select the 'they are the same' option during the difficult task.
+This speaks to the difficulty of the task, and respondents' interpretation of the task when it is more difficult to perceive small differences.
+
+<!--\hh{XXX the formatting isn't right yet here ... latex problem with large floats ... will fix it.}-->
 

### 2023-10-26 revisions-round1/response-to-reviewers.qmd
additional language updates to respond to R1, R2, and editor comments.

@@
 
--   **The introduction doesn't provide a clear rationale for the questions' importance.**
+-   The introduction doesn't provide a clear rationale for the questions' importance.
+
+    -   *In response to specific comments from Reviewer 2, we have strengthened some of the wording in the introduction to better emphasize the importance of our work and the survey framework.
+We hope this addresses this concern.*
 
 -   The entire article reads more like a project report rather than a scientific paper, with an excess of questions and bullet-pointed sentences.
+
     -   *To address this, we have removed two of the instances of bulleted lists: one which listed out the three variations in testing, each described in a bullet, and a second which demonstrated the options participants have to select from, each option shown in a bullet.
@@
 
-    -   **The title, abstract and conclusion suggests a wider context but the content is actually more limited in scope.
+    -   The title, abstract and conclusion suggests a wider context but the content is actually more limited in scope.
 I suggest changing the title to emphasise the main points in the article: 1) nationally representative sample and 2) testing stacked barplots.
 I think statements like "Our study results provide actionable insights that data visualization practitioners can utilize in the design of data visualizations." extrapolates too much.
-It should be more specific in that the actionable insights is *for stacked barplots* and also write concretely what the recommendation is (i.e., vertical display, align groups that most important to compare, etc.).**
+It should be more specific in that the actionable insights is *for stacked barplots* and also write concretely what the recommendation is (i.e., vertical display, align groups that most important to compare, etc.).
         -   *We have updated the statement in the conclusions per a specific recommendation in the minor comments.
-In addition, we have updated the paper title and some of the language in the abstract to reflect that the scope is limited to testing stacked bar charts.* **We will also update the recommendation language to be more concrete**
+In addition, we have updated the paper title and some of the language in the abstract to reflect that the scope is limited to testing stacked bar charts.
+We have also updated the recommendation language in the conclusions to be more concretely focused on stacked bar charts.*
 
@@
 
-    -   **I think the paper can benefit in the intro to include a stronger emphasis why we want to achieve a representative sample.
+    -   I think the paper can benefit in the intro to include a stronger emphasis why we want to achieve a representative sample.
 The authors do address this in part but it reads like an aside ("..., a common target audience for data visualisation").
@@
 Also in conclusion, the authors state "The size of our sample aids in identifying this signal.", but the authors can instead state e.g.
-"our study is orders of magnitude larger than other studies with similar scope, demonstrating...".**
+"our study is orders of magnitude larger than other studies with similar scope, demonstrating...".
         -   *There are multiple points here; we will address each.*
-        -   **Stronger emphasis on why we want to achieve a representative sample**
-        -   **Stronger emphasis on the importance of the authors' work is needed**
+        -   Stronger emphasis on why we want to achieve a representative sample
+            -   *We have added additional language in the introduction emphasizing the point of the general public as a target audience and the importance of using a representative survey population.
+We added the following language: "The general public is a common target audience for data visualization and scientific communication work; results of polls, surveys, and scientific work are presented often in the news, shared on social media, or posted online or in print materials.
+It is imperative that data visualizations used to communicate these topics are designed with the goal of reaching the entire population.
+To that end, an inclusive, representative approach to studying how viewers perceive information presented in charts is critical to understand how best to communicate with the general public".*
+        -   Stronger emphasis on the importance of the authors' work is needed
+            -   *Some of this has been addressed with the changes in the prior comment; however, we also updated the language in the conclusions to mention the large sample size relative to prior work.
+Further, in response to another comment, we expanded on the impact our findings have for interpreting non-representative studies, which we hope further underscores the importance of our work.*
 
-    -   **While I think the content is fine, I think the grammar in the paper can be better polished.
+    -   While I think the content is fine, I think the grammar in the paper can be better polished.
 (It is not that the current grammar is bad, but I just think it can be improved to make this a better paper).
-I have outlined some of the suggested corrections below.**
+I have outlined some of the suggested corrections below.
 
@@
 
--   **In general, both reviewers express some concerns regarding some of the language and description regarding overall impact of the work.
-I would encourage the authors to make it more clear that their primary contribution is in using a representative probability sample, and that their conclusions are focused on stacked barplot displays, as Reviewer 2 specifically remarks, and highlight why conclusions from a non-representative sample may be problematic within the data visualization world.**
+-   In general, both reviewers express some concerns regarding some of the language and description regarding overall impact of the work.
+I would encourage the authors to make it more clear that their primary contribution is in using a representative probability sample, and that their conclusions are focused on stacked barplot displays, as Reviewer 2 specifically remarks, and highlight why conclusions from a non-representative sample may be problematic within the data visualization world.
+
+    -   *We have updated the language in the title, abstract, introduction, and conclusion to be responsive to these concerns.
+See our specific responses to Reviewer 2's specific comments, but in general we added more emphasis on the importance of communicating a general audience, why non-representative survey findings don't necessarily hold for the entire population, and clarified that our scope was specifically on stacked bar chart displays.*
 
@@
 
--   **Can the authors comment further on modeling assumptions, particularly in relation to presenting all participants with the unaligned plot first followed by the aligned plot second (as opposed to randomizing this)?
+-   Can the authors comment further on modeling assumptions, particularly in relation to presenting all participants with the unaligned plot first followed by the aligned plot second (as opposed to randomizing this)?
 For instance, perhaps some participants notice that the plots are the same, with just the bars stacked in a different order, and their response to the second task is biased to be equivalent to that of the first task.
-This seems like it would be more of a problem if the plots were presented in the reverse order, but I'm curious if the authors looked into this.**
+This seems like it would be more of a problem if the plots were presented in the reverse order, but I'm curious if the authors looked into this.
 
@@
 There is a chance that someone recognizes they are completing the same task with the same bars when we present the aligned task before the unaligned task; we made the assumption that the chance of recognizing the tasks is lower when seeing unaligned prior to aligned.
-Participants are not able to return to the prior question once they move on to the second question, which does make it more difficult for participants to compare the tasks to one another.* **We will also add some text to the paper to reflect this limitation a little more clearly**
+Participants are not able to return to the prior question once they move on to the second question, which does make it more difficult for participants to compare the tasks to one another.
+We also added the following text in the section describing alignment: "Participants first saw the chart with the unaligned comparison, followed by the chart with the aligned comparison, but could not return to the first comparison after completing it.
+While there is some risk of participants recognizing the subsequent tasks to be the same value comparisons, the unaligned (harder) task was shown first to reduce the likelihood of participants carrying over an answer from the aligned (easier) task."*
 
@@
 
--   **It appears the cumulative logistic regression model with respect to participant demographic covariates just considers first-order effects, but interactions would also be pretty interesting here.
-Are there any notable interaction effects?**
+-   It appears the cumulative logistic regression model with respect to participant demographic covariates just considers first-order effects, but interactions would also be pretty interesting here.
+Are there any notable interaction effects?
 
-    -   HH *sample size*
+    -   *While we agree interactions would be interesting here, we lose sample size pretty quickly when interacting across demographic groups, particularly for demographics with several categories (e.g., education level and race/ethnicity).
+We are not comfortable with completing inference with those sample sizes, particularly given how important properly representing the relationship of demographic characteristics to outcomes is to our work.
+However, this is precisely the reason we hope to complete surveys with larger sample sizes, and possibly oversamples for underrepresented groups, in the future; studying those interactions could provide valuable insight into how subgroups in the general population interact with charts.*
 
@@
 If this is convention from Piepho (2004), it would be useful for this to be explained more or changed for a more general audience.
+
     -   *We have expanded the discussion on the interpretation of connected letters, and included an example for their first use.*

### 2023-10-26 SDSS-2023/testing-charts-accuracy.qmd
additional language updates to respond to R1, R2, and editor comments.

--- 2023-10-25 SDSS-2023/testing-charts-accuracy.qmd
@@
 Crowd sourced samples were shown by @borgoCrowdsourcingInformationVisualization2017 to be biased towards more male, younger, and relatively well-educated individuals relative to the adult U.S.
-population as a whole; generally, convenience and opt-in samples are not representative of the general population, a common target audience for data visualization and scientific communication work.
+population as a whole; generally, convenience and opt-in samples are not representative of the general population.
 Further, higher education individuals are likely to have prior exposure to data visualization or data topics in higher education, which may not be the case for other groups within the population; existing differences in exposure to data may drive differences in results across groups.
+The general public is a common target audience for data visualization and scientific communication work; results of polls, surveys, and scientific work are presented often in the news, shared on social media, or posted online or in print materials.
+It is imperative that data visualizations used to communicate these topics are designed with the goal of reaching the entire population.
+To that end, an inclusive, representative approach to studying how viewers perceive information presented in charts is critical to understand how best to communicate with the general public.
 
@@
 In particular, the additional context of the other tiles should *help* rather than *hurt* the comparison, unlike other situations, where additional information besides the pieces involved in the comparison were found to be distractors [@talbotFourExperimentsPerception2014].
+Participants first saw the chart with the unaligned comparison, followed by the chart with the aligned comparison, but could not return to the first comparison after completing it.
+While there is some risk of participants recognizing the subsequent tasks to be the same value comparisons, the unaligned (harder) task was shown first to reduce the likelihood of participants carrying over an answer from the aligned (easier) task.
 
@@
 #| label: fig-rounds-demographics
-#| fig-cap: "Overview of participants' demographics in each of the tasks: the colored dots show marginal percentages for demographics across the three tasks.
-In comparison, the dark lines show upper and lower 95% confidence of these demographics as estimated by the American Community Survery (ACS) 2021.
-Note that for a lot of these intervals the estimates are so close together that they appear as a single line.
-"
+#| fig-cap: "Demographics of respondents in each of the tasks. The dots show percentages of the national demographics as estimated by the American Community Survery (ACS) 2021."
 #| fig-height: 7
@@
 
-rounds_col <- RColorBrewer::brewer.pal(name="Accent", n=8)[c(1,3,5)]
-#[1] "#7FC97F" "#BEAED4" "#FDC086" "#FFFF99" "#386CB0" "#F0027F" "#BF5B17" "#666666"
 gender <- acs_gender %>%
   ggplot(aes(x = gender)) +
-  geom_errorbar(aes(ymin = percent-moe_percent, ymax=percent+moe_percent, colour="ACS estimate")) +
-#  geom_point(aes(y = percent)) +
+  geom_errorbar(aes(ymin = percent-moe_percent, ymax=percent+moe_percent)) +
+  geom_point(aes(y = percent)) +
   geom_point(
-    size=3,
     aes(x = sex, y = value, colour=chart),
@@
   ) +
-  ylim(c(0, NA)) +
   geom_linerange(
@@
     ),
-    linewidth = 1, 
     position = position_dodge(width = 0.5),
-    data = omni_gender %>% pivot_wider(names_from=group, values_from=value),
-    show.legend = FALSE
-  ) +
-  ylab("Percent") + 
-  ggtitle("Gender") +
-#  theme(axis.text.x = element_text(angle=0, vjust = 0.5)) +
-  xlab("") +
-#  theme(legend.position ="none") +
-  scale_linetype_manual(values=c(2,2,1)) +
-  scale_color_manual(values=c("grey20", rounds_col))
+    data = omni_gender %>% pivot_wider(names_from=group, values_from=value)
+  )
 
@@
                                    20582187 + 21049271 + 21928936, 
-                                   73789098),
-                      percent = c(6.9 + 9.2, 
-                                  6.8 + 6.6 + 6.2, 
-                                  6.2 + 6.4 + 6.7, 
-                                  22.4
-                                  ),
-                      moe_percent = c(sqrt(0.1^2+0.1^2),
-                              sqrt(0.1^2+0.1^2 + 0.1^2),
-                              sqrt(0.1^2+0.1^2 + 0.1^2),
-                              0.1) )
-acs_age <- acs_age %>% mutate(percent2 = estimate/sum(estimate)*100)
+                                   73789098))
+acs_age <- acs_age %>% mutate(percent = estimate/sum(estimate)*100)
 
@@
 
-age_bak <- round_123 %>% 
+age <- round_123 %>% 
   ggplot(aes(x = age4)) + 
@@
 
-age <- acs_age %>%
-  ggplot(aes(x = age4)) +
-  geom_errorbar(aes(ymin = percent2-moe_percent, ymax=percent2+moe_percent, colour="ACS estimate")) +
-#  geom_point(aes(y = percent)) +
-  geom_point(
-    size=3,
-    aes(x = age, y = value, colour=chart),
-    position = position_dodge(width = 0.5),
-    data = omni_age %>% filter(group=="mean")
-  ) +
-  ylim(c(0, NA)) +
-  geom_linerange(
-    aes(
-      x = age,
-      ymin = low,
-      ymax=high,
-      colour=chart
-    ),
-    linewidth = 1,
-    position = position_dodge(width = 0.5),
-    data = omni_age %>% pivot_wider(names_from=group, values_from=value)
-  ) +
-  ylab("Percent") + 
-  ggtitle("Age groups") +
-  theme(axis.text.x = element_text(angle=0, vjust = 0.5)) +
-  xlab("") +
-  theme(legend.position ="none") +
-  scale_linetype_manual(values=c(2,2,1)) +
-  scale_color_manual(values=c("grey20", rounds_col))
-
 acs_educ5 <- data.frame(educ5 = c("Less than HS", "HS or equivalent", "Some college",  "Bachelor", "Post graduate"), 
-                      estimate = c(
-                      3636691 + 10793507 + 14256849, 
-                      9974372 + 59636386,
-                      13050667 + 45042031 + 19614710,
-                      3677359 + 46354331,
-                      29454503),
-                      moe_percent = c(sqrt(0.1^2+0.1^2+0.1^2),
-                              sqrt(0.1^2+0.1^2 ),
-                              sqrt(0.1^2+0.1^2 + 0.1^2),
-                              sqrt(0.1^2+0.2^2 ),
-                              0.1) 
-                      )
-acs_educ5 <- acs_educ5 %>% mutate(percent = estimate/(30339089+225152317)*100)
+                      estimate = c(28687047, 69610758, 77707408, 50031690, 29454503))
+acs_educ5 <- acs_educ5 %>% mutate(percent = estimate/sum(estimate)*100)
 
@@
 
-education_bak <- round_123 %>% 
+education <- round_123 %>% 
   ggplot(aes(x = educ5)) + 
@@
 
-education <- acs_educ5 %>%
-  mutate(educ5 = factor(educ5, levels=levels(omni_educ$educ))) %>%
-  ggplot(aes(x = educ5)) +
-  geom_errorbar(aes(ymin = percent-moe_percent, ymax=percent+moe_percent, colour="ACS estimate")) +
-#  geom_point(aes(y = percent)) +
-  geom_point(
-    size=3,
-    aes(x = educ, y = value, colour=chart),
-    position = position_dodge(width = 0.5),
-    data = omni_educ %>% filter(group=="mean")
-  ) +
-  ylim(c(0, NA)) +
-  geom_linerange(
-    aes(
-      x = educ,
-      ymin = low,
-      ymax=high,
-      colour=chart
-    ),
-    linewidth = 1, 
-    position = position_dodge(width = 0.5),
-    data = omni_educ %>% pivot_wider(names_from=group, values_from=value),
-    show.legend = FALSE
-  ) +
-  ylab("Percent") + 
-  ggtitle("Education") +
-#  theme(axis.text.x = element_text(angle=0, vjust = 0.5)) +
-  xlab("") +
-  theme(legend.position ="none") +
-  scale_linetype_manual(values=c(2,2,1)) +
-  scale_color_manual(values=c("grey20", rounds_col))
-
 acs_income <- data.frame(
@@
               .6*16.8+12.8,
-              16.3+7.9+9.8),
-  moe_percent = c(sqrt(0.1^2+0.1^2+0.1^2+0.1^2+0.1^2),
-                  sqrt(0.1^2+0.1^2+0.1^2),
-                  sqrt(0.1^2 + 0.1^2),
-                  sqrt(0.1^2+0.1^2+0.1^2))
-  )
+              16.3+7.9+9.8))
 
@@
 
-income_bak <- round_123 %>% 
+income <- round_123 %>% 
   ggplot(aes(x = income4)) + 
@@
   theme(legend.position = "none")
-
-income <- acs_income %>%
-  mutate(income4 = factor(income4, levels=levels(omni_income$income))) %>%
-  ggplot(aes(x = income4)) +
-  geom_errorbar(aes(ymin = percent-moe_percent, ymax=percent+moe_percent, colour="ACS estimate")) +
-#  geom_point(aes(y = percent)) +
-  geom_point(
-    size=3,
-    aes(x = income, y = value, colour=chart),
-    position = position_dodge(width = 0.5),
-    data = omni_income %>% filter(group=="mean")
-  ) +
-  ylim(c(0, NA)) +
-  geom_linerange(
-    aes(
-      x = income,
-      ymin = low,
-      ymax=high,
-      colour=chart
-    ),
-    linewidth = 1, 
-    position = position_dodge(width = 0.5),
-    data = omni_income %>% pivot_wider(names_from=group, values_from=value),
-    show.legend = FALSE
-  ) +
-  ylab("Percent") + 
-  ggtitle("Income") +
-#  theme(axis.text.x = element_text(angle=0, vjust = 0.5)) +
-  xlab("") +
-  theme(legend.position ="none") +
-  scale_linetype_manual(values=c(2,2,1)) +
-  scale_color_manual(values=c("grey20", rounds_col))
 
@@
 With our survey framework, we have shown that we can rank structural variations on design and tasks by their relative accuracy; we identified significantly different rates of respondent accuracy under different designs presenting the same data.
-The size of our sample aids in identifying this signal.
+Our study sample is larger than prior studies with a similar scope. That large sample size improves our power to identify this signal.
 

### 2023-10-26 revisions-round1/response-to-reviewers.qmd
eqns 4 and 7
closing out response on eqns 4 and 7 items

@@
 
--   **Is there a reason why the meaning/values of i from 1 to 3 are reversed between the two models specified in Equations 4 and 7?**
+-   Is there a reason why the meaning/values of i from 1 to 3 are reversed between the two models specified in Equations 4 and 7?
 
-    -   KR: **Heike, please confirm** which order they should be in.
+    -   *The order in 4 is the correct one, and was actually the parameterization used for both models and shown in both corresponding tables.
+We have fixed the description for Equation 7 to reflect this.*
 

### 2023-10-27 revisions-round1/response-to-reviewers.qmd
finalizing response on Fig 3 comment

@@
 
-    -   **Figure 3: I take it that the American Community Survey 2021 is considered the Census estimates?
+    -   Figure 3: I take it that the American Community Survey 2021 is considered the Census estimates?
 I think it makes more sense to swap the bar and point so that the bar is the Census estimate.
@@
 In this plot, you've facetted by the stimulus type but I think it makes sense to facet by the category level (the comparison of distribution is done with the census estimate directly, and I think the interest is to compare the distribution across stimulus so you want the primary factor of interest to compare to be close in proximity).
-Within a category level, the census estimate is constant across stimulus so you could substitute is with a horizontal line (instead of a bar or point) as well.**
+Within a category level, the census estimate is constant across stimulus so you could substitute is with a horizontal line (instead of a bar or point) as well.
+        -   *We have changed the format of this figure to address these concerns, using a dot plot with error bars to represent the uncertainty.*
 

### 2023-10-27 SDSS-2023/testing-charts-accuracy.qmd
eqns 4 and 7
demographics

@@
   xlab("") +
-#  theme(legend.position ="none") +
+  theme(legend.position ="none") +
   scale_linetype_manual(values=c(2,2,1)) +
@@
 
+gender <- 
+  gender +
+  annotate("text", x = 0.83, y = 5, label="ACS estimate", hjust=0, 
+           size = 5, colour = "grey20") +
+  annotate("segment", x = 0.5, xend=0.7, y = 5,yend=5, 
+           linewidth = 1, colour="grey20") +
+  annotate("text", x = 0.83, y = 15, label="Horizontal design", hjust=0,
+           size = 5, colour = "grey20") +
+  annotate("segment", x = 0.8333, xend = 0.8333, y = 17, yend=40, 
+           linewidth = 0.75, colour=rounds_col[1], linetype=2) +
+  annotate("text", x = 1, y = 20, label="Horizontal wide", hjust=0,
+           size = 5, colour = "grey20") +
+  annotate("segment", x = 1, xend = 1, y = 22, yend=40, 
+           linewidth = 0.75, colour=rounds_col[2], linetype=2) +
+  annotate("text", x = 1.17, y = 25, label="Vertical design", hjust=0,
+           size = 5, colour = "grey20") +
+  annotate("segment", x = 1.166, xend = 1.166, y = 27, yend=40, 
+           linewidth = 0.75, colour=rounds_col[3], linetype=2)
+
 (gender + education + plot_layout(nrow=1, widths=c(2,5))) / 
   (age + income + plot_layout(nrow=1, widths = c(4,4)))
+
+```
+
+```{r eval=FALSE, include=FALSE}
+omni <- rbind(
+omni_gender %>% rename(levels = sex) %>% 
+  mutate(variable = "Gender"),
+omni_age %>% rename(levels = age) %>% 
+  mutate(variable = "Age groups"),
+omni_income %>% rename(levels = income) %>% 
+  mutate(variable = "Household income"),
+omni_educ %>% rename(levels = educ) %>% 
+  mutate(variable = "Education"))
+
+acs <- rbind(
+acs_age %>% rename(levels = age4) %>% mutate(variable = "Age groups") %>% 
+  dplyr::select(variable, levels, percent2, moe_percent) %>% rename(percent = percent2),
+acs_educ5 %>% rename(levels = educ5) %>% mutate(variable = "Education") %>% 
+  dplyr::select(variable, levels, percent, moe_percent), 
+acs_gender %>% rename(levels = gender) %>% mutate(variable = "Gender") %>% 
+  dplyr::select(variable, levels, percent, moe_percent), 
+acs_income %>% rename(levels = income4) %>% mutate(variable = "Household income") %>%
+  dplyr::select(variable, levels, percent, moe_percent)
+)
+
+acs %>% 
+  ggplot(aes(y = levels)) + 
+  geom_errorbar(aes(xmin = percent-moe_percent, xmax = percent + moe_percent, colour = "ACS Est.")) +
+  facet_grid(rows = vars(variable), space="free", scales = "free") +
+  xlim(c(0,NA)) + ylab("") + xlab("Percent") +
+  scale_color_manual(values=c("grey20", rounds_col)) +
+  geom_linerange(
+    aes(
+      y = levels,
+      xmin = low,
+      xmax=high,
+      colour=chart
+    ),
+    linewidth = 1, 
+    position = position_dodge(width = 0.5),
+    data = omni %>% pivot_wider(names_from=group, values_from=value),
+    show.legend = FALSE
+  )
 
@@
 #| label: fig-certainty
-#| out-width: 100%
+#| out-width: 95%
 #| fig-cap: "Certainty by task, chart design, and correctness.
@@
 
-Let $Y_k$ be the response of participant $k$, on a scale from 1 = 'wrong', 2 = 'they are the same' to 3 = 'correct'.
+Let $Y_k$ be the response of participant $k$, on a scale from 1 = 'correct', 2 = 'they are the same' to 3 = 'wrong'.
+
 We use a generalized cumulative logistic regression, where $\mu_i$ are intercepts $1 \le i < 3$, $X_k$ are demographics of the $k$th participant (in form of the model matrix), and $\beta_{i}$ are the coefficients.
@@
 #| layout-nrow: 2
-#| out-width: 100%
+#| out-width: 95%
 

### 2023-10-27 SDSS-2023/testing-charts-accuracy.qmd
cleaned up visual
supplementary materials section and finalizing response to reviewers
tex file overhauled and rendered as pdf

@@
 title: "What do they see? Testing perceptual accuracy in a U.S. general population survey using stacked bar charts"
-short-title: Testing stacked bar charts in surveys # running title
+runtitle: Testing stacked bar charts in surveys # running title
 bibliography: references.bib
@@
 
-# To-do
-
-Into appendix: Stimulus question(s)
-
 ```{r run-to-update-bibfile, eval=FALSE, include=FALSE}
@@
 #| fig-width: 11
-#| out-width: 90%
+#| out-width: 95%
 
@@
 
-# coefs_zoom$term <- c("$$\\hat{\\mu}$$", "$$\\text{They are the same } \\widehat{\\rho}_{2}$$", 
-#                      "Wrong $$\\widehat{\\rho}_{3}$$", "Aligned $$\\widehat{\\tau}_{2}$$", 
-#                      "Horizontal wide $$\\widehat{\\delta}_{2}$$", "Vertical $$\\widehat{\\delta}_{3}$$")
-
-coefs_zoom$term <- c("\\hat{\\mu}", 
-                         "\\text{They are the same } \\widehat{\\rho}_{2}", 
-                         "Wrong $$\\widehat{\\rho}_{3}$$", 
-                         "Aligned $$\\widehat{\\tau}_{2}$$", 
-                         "Horizontal wide $$\\widehat{\\delta}_{2}$$", 
-                         "Vertical $$\\widehat{\\delta}_{3}$$")
+# coefs_zoom$term <- c("\\hat{\\mu}", 
+#                          "\\text{They are the same } \\widehat{\\rho}_{2}", 
+#                          "Wrong $$\\widehat{\\rho}_{3}$$", 
+#                          "Aligned $$\\widehat{\\tau}_{2}$$", 
+#                          "Horizontal wide $$\\widehat{\\delta}_{2}$$", 
+#                          "Vertical $$\\widehat{\\delta}_{3}$$")
 

### 2023-10-27 SDSS-2023/testing-charts-accuracy.tex
keywords
removing warning message and italicizing keywords
tex file overhauled and rendered as pdf

@@
 \hypersetup{
-  pdftitle={Testing Charts: viewers' perceptual accuracy in surveys},
+  pdftitle={What do they see? Testing perceptual accuracy in a U.S. general population survey using stacked bar charts},
   pdfkeywords={graphical perception, data visualization, survey data},
@@
 
-\title{Testing Charts: viewers' perceptual accuracy in surveys}
+\title[Testing stacked bar charts in surveys]{What do they see? Testing perceptual accuracy in a U.S. general population survey using stacked bar charts}
 \date{}
@@
 employing a large, nationally-representative, probability-based panel of
-survey respondents to test perception in statistical charts. Our
+survey respondents to test perception in stacked bar charts. Our
 findings provide actionable guidance for data visualization
@@
 \end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi
-
-  \newcommand{\hh}[1]{{\textcolor{orange}{#1}}} 
-  \newcommand{\kr}[1]{{\textcolor{teal}{#1}}}.  
-  \newcommand{\ejm}[1]{{\textcolor{ForestGreen}{#1}}}.  
-  \newcommand{\ndt}[1]{{\textcolor{purple}{#1}}}.
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, boxrule=0pt, sharp corners, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
 
@@
 
-The viewer's employment of comparisons of the components is a crucial
-step in the process of interacting with and understanding the chart.
-Cleveland and McGill (1984) observed as such, and in their seminal study
-defined the better visual among a pair as the one that allows viewers to
-make more accurate comparisons. Based on mappings of quantitative
-variables to different graphical elements, Cleveland and McGill's study
-resulted in a ranking of perceptual tasks (such as identifying the
-larger of two lengths or areas) from most accurate to least accurate,
-which was then extended by Mackinlay (1986) to a theoretical framework
-ranking tasks' order along their ordinal and nominal scales.
+The comparisons of components is a crucial step in the process of
+interacting with and understanding the chart. Cleveland and McGill
+(1984) observed as such, and in their seminal study defined the better
+visual among a pair as the one that allows viewers to make more accurate
+comparisons. Based on mappings of quantitative variables to different
+graphical elements, Cleveland and McGill's study resulted in a ranking
+of perceptual tasks (such as identifying the larger of two lengths or
+areas) from most accurate to least accurate, which was then extended by
+Mackinlay (1986) to a theoretical framework ranking tasks' order along
+their ordinal and nominal scales.
 
@@
 study. Crowd sourced samples were shown by Borgo et al. (2017) to be
-biased towards more male, younger, and relatively higher education
-relative to the adult U.S. population as a whole; generally, convenience
-and opt-in samples are not representative of the general population, a
-common target audience for data visualization and scientific
-communication work. Further, the populations' emphasis on higher
-education individuals also leads to results which hold for groups of
-individuals who may be more likely to have prior exposure to data
-visualization in the context of scientific communication, or more
-exposure to data topics in higher education, but may not hold across
-other groups within the population.
-
-Here, we seek to answer whether it is possible to reproduce some of the
-previous findings in the context of a survey with a large,
+biased towards more male, younger, and relatively well-educated
+individuals relative to the adult U.S. population as a whole; generally,
+convenience and opt-in samples are not representative of the general
+population. Further, higher education individuals are likely to have
+prior exposure to data visualization or data topics in higher education,
+which may not be the case for other groups within the population;
+existing differences in exposure to data may drive differences in
+results across groups. The general public is a common target audience
+for data visualization and scientific communication work; results of
+polls, surveys, and scientific work are presented often in the news,
+shared on social media, or posted online or in print materials. It is
+imperative that data visualizations used to communicate these topics are
+designed with the goal of reaching the entire population. To that end,
+an inclusive, representative approach to studying how viewers perceive
+information presented in charts is critical to understand how best to
+communicate with the general public.
+
+In this work, we seek to answer whether it is possible to reproduce some
+of the previous findings in the context of a survey with a large,
 nationally-representative set of respondents. Specifically, we present
@@
 marked pieces (tiles) is larger within each chart, and focus on three
-main sets of structural variation in the design of that chart:
-
-\begin{itemize}
-\item
-  First, we vary the alignment of the pieces in question. Viewers are
-  presented with two marked pieces in a chart that do not share a common
-  baseline (aligned pieces), then two pieces that do share a common
-  baseline (unaligned pieces).
-\item
-  Second, we vary the orientation of the chart -- we rotate the vertical
-  stacked bar chart, and present a horizontally oriented version of the
-  same chart, with identically sized marked pieces.
-\item
-  Finally, we change the aspect ratio of the chart and present a wider
-  version of the horizontally oriented chart which has longer, but
+main sets of structural variation in the design of that chart. First, we
+vary the alignment of the pieces in question. Viewers are presented with
+two marked pieces in a chart that do not share a common baseline
+(aligned pieces), then two pieces that do share a common baseline
+(unaligned pieces). Second, we vary the orientation of the chart -- we
+rotate the vertical stacked bar chart, and present a horizontally
+oriented version of the same chart, with identically sized marked
+pieces. Finally, we change the aspect ratio of the chart and present a
+wider version of the horizontally oriented chart which has longer, but
 thinner, marked pieces.
-\end{itemize}
 
@@
 
-{\centering \includegraphics[width=0.9\textwidth,height=\textheight]{images/stimuli-overview.png}
+{\centering \includegraphics[width=1\textwidth,height=\textheight]{images/stimuli-overview-1.png}
 
@@
 
-\caption{\label{fig-tasks}Each of the visual stimuli presented to
-viewers. In each bar chart, two pieces are marked. The larger piece in
-each chart are pieces B and D.}
+\caption{\label{fig-tasks}Overview of all of the visual stimuli
+presented to viewers. In each bar chart, two pieces are marked. The
+larger piece in each chart are pieces B and D. Please refer to the
+appendix for full size images of the stimuli.}
 
@@
 Looking at the chart below, which of the marked dark blue pieces is
-bigger, A or B? Just your best guess is fine.
-
-\begin{itemize}
-\item
-  A is bigger
-\item
-  B is bigger
-\item
-  They are the same''
-\end{itemize}
+bigger, A or B? Just your best guess is fine. {[}{]} A is bigger; {[}{]}
+B is bigger; {[}{]} They are the same.''
 
@@
 positional assessment, which in the absence of three-dimensional cues
-seems to be the general strategy viewers use when evaluating barcharts
+seems to be the general strategy viewers use when evaluating bar charts
 (Zacks et al. 1998). Unaligned tiles do not share this anchor of
 position, i.e.~the difference between their sizes needs at least two
-comparisons of positions. However, unaligned bars in a stacked barchart
+comparisons of positions. However, unaligned bars in a stacked bar chart
 are not just assessments of the lengths of bars -- the other tiles in
@@
 besides the pieces involved in the comparison were found to be
-distractors (Talbot, Setlur, and Anand 2014).
+distractors (Talbot, Setlur, and Anand 2014). Participants first saw the
+chart with the unaligned comparison, followed by the chart with the
+aligned comparison, but could not return to the first comparison after
+completing it. While there is some risk of participants recognizing the
+subsequent tasks to be the same value comparisons, the unaligned
+(harder) task was shown first to reduce the likelihood of participants
+carrying over an answer from the aligned (easier) task.
 
@@
 (rather than cumulate) a set of \(\ell\) surveys \(S_1\), \(S_2\),
-\ldots{} \(S_\ell\) (with \$ \ell \ge 2\$ ) as described in
+\ldots{} \(S_\ell\) (with \(\ell \ge 2\)) as described in
 O'Muircheartaigh and Pedlow (2002), by multiplying weights in \(S_i\) by
@@
 (nominal sample size), effective sample size, sum of weights, and
-factors for the combining of surveys as discussed above.}\tabularnewline
+factors for the combining of surveys as discussed above.
+}\tabularnewline
 \toprule\noalign{}
@@
 Round 3 & June 2022 & 921 & 513.5 & 923.1 & \(\lambda_3\) = 0.338 \\
-\textbf{Total} & --- & 2807 & 1520.8 & & \\
+\textbf{Combined} & --- & 2807 & 1520.8 & & \\
 \end{longtable}
@@
 
-\caption{\label{fig-rounds-demographics}Demographics of respondents in
-each of the tasks. The dots show percentages of the national
-demographics as estimated by the American Community Survery (ACS) 2021.}
+\caption{\label{fig-rounds-demographics}Overview of participants'
+demographics in each of the tasks: the colored dots show marginal
+percentages for demographics across the three tasks. In comparison, the
+dark lines show upper and lower 95\% confidence of these demographics as
+estimated by the American Community Survery (ACS) 2021. Note that for a
+lot of these intervals the estimates are so close together that they
+appear as a single line.}
 
@@
 The overlaid rectangles represent 95\% confidence intervals. The letters
-in blue and orange encode significances between pairwise proportions:
-two bars have a significantly different proportion (at a 5\%
-significance level) if they do not share a letter. There is no
-significant difference between the designs for wrong responses.}
+in blue and orange encode significances between pairwise proportions
+(Piepho 2004): two bars have a significantly different proportion (at a
+5\% significance level) if they do not have a letter in common. There is
+no significant difference between the designs for wrong responses.}
 
@@
 design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide. The
-estimated values and 95\% confidence intervals are shown in
-Table~\ref{tbl-rounds-123}. The lowercase letters indicate significances
-between pairs of estimates: two estimates within the same column are
-significantly different at 5\% level if they do not have any letter in
-common (Piepho 2004).
+letters allow us to assess whether differences between two means are
+significant: two estimates within the same column are significantly
+different (at the 5\% level) if they do not have any letter in common.
+This method is called connected letters and has been introduced by
+Piepho (2004). In Figure~\ref{fig-response-123}, we see, that the
+proportions of wrong responses are significantly different between
+tasks, but designs within each task are not significantly different. For
+correct responses, all proportions are significantly different except
+for the proportion of correctly assessing the difference between
+unaligned rectangles in the wide horizontal design: their label of
+\texttt{ab} shows us that assessing the differences between unaligned
+rectangles in the wide horizontal design is not significantly better
+than in the horizontal design (labelled \texttt{a}), but also not
+significantly worse than in the vertical design (labelled \texttt{b}).
+
+The estimated values and 95\% confidence intervals are shown in
+Table~\ref{tbl-rounds-123}.
 
@@
 significant. However, different designs lead to different rates of
-zooming: we observe in \autoref{fig-zoom-123} that when dealing with the
-vertical design, the rate of zooming is significantly higher than for
-the two horizontal designs.
+zooming: we observe in \autoref{fig-zoom-123} that when a significantly
+lower rate of respondents chose to make use of zooming when answering
+the horizontal designs compared to the vertical design.
 
@@
 
-{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{./figures/fig-zoom-123-1.png}
+{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{./figures/fig-zoom-123-1.png}
 
@@
 \midrule
-$\hat{\mu}$ & $-0.35$ & $0.15$ & $-2.4$ & 0.0155 \\ 
+\multicolumn{5}{l}{} \\ 
+\midrule
+$\widehat{\mu}$ & $-0.35$ & $0.15$ & $-2.4$ & 0.0155 \\ 
 \midrule
@@
 \midrule
-They are the same, $\widehat{\rho_2}$ & $-0.17$ & $0.09$ & $-1.9$ & 0.0578 \\ 
-Wrong, $\widehat{\rho_3}$ & $-0.06$ & $0.13$ & $-0.5$ & 0.6466 \\ 
+They are the same $\widehat{\rho}_2$ & $-0.17$ & $0.09$ & $-1.9$ & 0.0578 \\ 
+Wrong $\widehat{\rho}_3$ & $-0.06$ & $0.13$ & $-0.5$ & 0.6466 \\ 
 \midrule
@@
 \midrule
-Aligned, $\widehat{\tau_2}$ & $0.00$ & $0.06$ & $-0.1$ & 0.9435 \\ 
+Aligned $\widehat{\tau}_2$ & $0.00$ & $0.06$ & $-0.1$ & 0.9435 \\ 
 \midrule
@@
 \midrule
-Horizontal wide, $\widehat{\delta_2}$ & $0.27$ & $0.15$ & $1.8$ & 0.0752 \\ 
-Vertical, $\widehat{\delta_3}$ & $0.98$ & $0.13$ & $7.7$ & < 0.0001 \\ 
+Horizontal wide $\widehat{\delta}_2$ & $0.27$ & $0.15$ & $1.8$ & 0.0752 \\ 
+Vertical $\widehat{\rho}_3$ & $0.98$ & $0.13$ & $7.7$ & < 0.0001 \\ 
 \bottomrule
@@
 
-{\centering \includegraphics[width=1\textwidth,height=\textheight]{./figures/fig-certainty-1.png}
+{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{./figures/fig-certainty-1.png}
 
@@
 Let \(Y_k\) be the response of participant \(k\), on a scale from 1 =
-`wrong', 2 = `they are the same' to 3 = `correct'. We use a generalized
-cumulative logistic regression, where \(\mu_i\) are intercepts
-\(1 \le i < 3\), \(X_k\) are demographics of the \(k\)th participant (in
-form of the model matrix), and \(\beta_{i}\) are the coefficients.
+`correct', 2 = `they are the same' to 3 = `wrong'.
+
+We use a generalized cumulative logistic regression, where \(\mu_i\) are
+intercepts \(1 \le i < 3\), \(X_k\) are demographics of the \(k\)th
+participant (in form of the model matrix), and \(\beta_{i}\) are the
+coefficients.
 
@@
 \end{equation}
+
+\begin{figure}
+
+\begin{minipage}[t]{\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{./figures/fig-demographics-1.png}
+
+}
+
+}
+
+\subcaption{\label{fig-demographics-1}Comparisons of aligned tiles}
+\end{minipage}%
+\newline
+\begin{minipage}[t]{\linewidth}
+
+{\centering
+
+\raisebox{-\height}{
+
+\includegraphics{./figures/fig-demographics-2.png}
+
+}
+
+}
+
+\subcaption{\label{fig-demographics-2}Comparisons of unaligned tiles}
+\end{minipage}%
+
+\caption{\label{fig-demographics}Response patterns by alignment and
+demographic levels. For aligned tiles, gender, age, and education are
+not significant factors. However, income levels do have a (small)
+effect. When income levels increase, the percentage of wrong answers
+decreases, while the percentage of correct answers increases slightly
+(significant at 5\%). For the more difficult task of comparing unaligned
+tiles, higher levels of education and higher levels of income are
+associated with a significant increase of panelists choosing a response
+of `they are the same', resulting in significant decreases of both
+correct answers and wrong answers.}
+
+\end{figure}
 
@@
 
-\begin{verbatim}
-Warning in RET$pfunction("adjusted", ...): Completion with error > abseps
-\end{verbatim}
-
-\begin{figure}
-
-\begin{minipage}[t]{\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{./figures/fig-demographics-1.png}
-
-}
-
-}
-
-\subcaption{\label{fig-demographics-1}Comparisons of aligned tiles}
-\end{minipage}%
-\newline
-\begin{minipage}[t]{\linewidth}
-
-{\centering
-
-\raisebox{-\height}{
-
-\includegraphics{./figures/fig-demographics-2.png}
-
-}
-
-}
-
-\subcaption{\label{fig-demographics-2}Comparisons of unaligned tiles}
-\end{minipage}%
-
-\caption{\label{fig-demographics}Response patterns by alignment and
-demographic levels. For aligned tiles, gender, age, and education are
-not significant factors. However, income levels do have a (small)
-effect. When income levels increase, the percentage of wrong answers
-decreases, while the percentage of correct answers increases slightly
-(significant at 5\%). For the more difficult task of comparing unaligned
-tiles, higher levels of education and higher levels of income are
-associated with a significant increase of panelists choosing a response
-of `they are the same', resulting in significant decreases of both
-correct answers and wrong answers.}
-
-\end{figure}
-
 \blandscape
@@
 \midrule
-HS or equivalent & $0.77$ &  [$0.46$, $1.30$] &  & $0.50$ &  [$0.20$, $1.21$] &  & $0.81$ &  [$0.47$, $1.40$] &  & $1.22$ &  [$0.66$, $2.27$] &  \\ 
+HS or equivalent & $0.77$ &  [$0.46$, $1.30$] &  & $0.50$ &  [$0.20$, $1.21$] &  & $0.81$ &  [$0.47$, $1.40$] &  & $1.22$ &  [$0.66$, $2.26$] &  \\ 
 Some college & $0.91$ &  [$0.56$, $1.49$] &  & $0.74$ &  [$0.32$, $1.71$] &  & $0.63$ &  [$0.38$, $1.06$] & . & $1.17$ &  [$0.65$, $2.12$] &  \\ 
@@
 significantly different rates of respondent accuracy under different
-designs presenting the same data. The size of our sample aids in
-identifying this signal.
+designs presenting the same data. Our study sample is larger than prior
+studies with a similar scope. That large sample size improves our power
+to identify this signal.
 
@@
 with the chart. Paradata on viewer interaction with the chart (e.g.,
-zooming and time spent on each task) can be collected as well as asking
+zooming and time spent on each task) can be collected as well as
 directly asking respondents questions about their certainty. Not only
@@
 responses due to the larger representation of individuals with higher
-education levels among those study respondents.
+education levels among those study respondents. If studies using
+non-representative or crowd-sourced samples find non-significant results
+when studying elements of design, those results may not actually be
+non-significant for other populations, particularly lower education or
+lower income populations. Researchers should not disregard
+non-significant findings in those survey studies but rather ask whether
+the lack of significance would hold across population subgroups.
 
 Our study results provide actionable insights that data visualization
-practitioners can utilize in the design of data visualizations.
+practitioners can utilize in the design of stacked bar charts.
 Comparisons of interest should be lined up and share a common baseline

### 2023-10-27 revisions-round1/response-to-reviewers.qmd

@@
 
--   **There is an overabundance of figures without clear emphasis.**
+-   There is an overabundance of figures without clear emphasis.
+
+    -   *We have removed a redundant sub-figure from Figure 4.
+Displaying the response patterns for each research question visually is a useful method of communicating our findings in addition to formal modeling results.
+The inclusion of our results figures highlights the nuance in our findings, particularly when it comes to comparisons across tasks and subgroups, and we feel they are important to convey our findings.
+For that reason, we have retained most of them.*
 
@@
 
-    -   **Figure 1, the text within the visualisation are too small to see. Perhaps lay it out as a 3 x 2 instead?**
-            - HH: *Figure 1 was not meant to have readable text within the stimuli visualizations but only give an overview of the different stimuli and the testing situations.
-We have increased the font size for the overview, expanded the discussion of the figure to reflect its purpose and added image (a) to the supplement in a larger format.* **create supplement and add stimulus**
+    -   Figure 1, the text within the visualisation are too small to see. Perhaps lay it out as a 3 x 2 instead?
+            - *Figure 1 was not meant to have readable text within the stimuli visualizations but only give an overview of the different stimuli and the testing situations.
+We have increased the font size for the overview, expanded the discussion of the figure to reflect its purpose and added image (a) to the supplement in a larger format.*
 
@@
 
-    -   **Figure 4: I think (b) alone is sufficient as you can gain the same information as (a) (although perhaps it is less familiar for wider audience).
-You could arguably put figure (a) information in a table instead.**.
-
-    XXX HH: What should we do here?
-I'd be OK with putting all of this into a table.
-KR: I'm ok with moving to a table, though I like having at least one visual to really highlight the difference between aligned/unaligned.
-This would help us be responsive to the other comments about overuse of visuals, though.
+    -   Figure 4: I think (b) alone is sufficient as you can gain the same information as (a) (although perhaps it is less familiar for wider audience).
+You could arguably put figure (a) information in a table instead.
+        -   *We have removed Figure 4(a) and just kept what was originally 4(b) to avoid this redundancy.*
 
@@
 
--   **Please add an unnumbered section Supplementary Material to list its content (e.g., data, code, additional simulation results, technical proofs, etc.).
+-   Please add an unnumbered section Supplementary Material to list its content (e.g., data, code, additional simulation results, technical proofs, etc.).
 See recently published papers for examples.
-Reproducibility is held high at the JDS; the reproducibility team will check for reproducibility when the paper is accepted.**
+Reproducibility is held high at the JDS; the reproducibility team will check for reproducibility when the paper is accepted.
+
+    -   *We have added this section to the end of the paper, and note that we also submitted a supplement document.*
 
@@
 
--   **I believe the figures are fairly clear (per Reviewer 1's comment), but the authors might also be inclined to be more selective with what is shown in Figures 3, 6-8, and maybe move some of the others to an appendix.**
+-   I believe the figures are fairly clear (per Reviewer 1's comment), but the authors might also be inclined to be more selective with what is shown in Figures 3, 6-8, and maybe move some of the others to an appendix.
+
+    -   *In response to a specific comment from Reviewer 2, we have removed a redundant sub-figure from Figure 4.
+We have retained the other figures given that we are still within the page limit and displaying the response patterns visually is a useful method of communicating our findings in addition to formal modeling results.
+The inclusion of our results figures highlights the nuance in our findings, particularly when it comes to comparisons across tasks and subgroups.*
 

### 2023-10-27 SDSS-2023/testing-charts-accuracy.qmd
figure 4

@@
 @heerCrowdsourcingGraphicalPerception2010 reproduced Cleveland and McGill's rankings using a larger sample from a crowd sourcing platform, employing a total of 82 Amazon Mechanical Turk (mTurk) workers for the study.
-Crowd sourced samples were shown by @borgoCrowdsourcingInformationVisualization2017 to be biased towards more male, younger, and relatively well-educated individuals relative to the adult U.S.
-population as a whole; generally, convenience and opt-in samples are not representative of the general population.
-Further, higher education individuals are likely to have prior exposure to data visualization or data topics in higher education, which may not be the case for other groups within the population; existing differences in exposure to data may drive differences in results across groups.
-The general public is a common target audience for data visualization and scientific communication work; results of polls, surveys, and scientific work are presented often in the news, shared on social media, or posted online or in print materials.
-It is imperative that data visualizations used to communicate these topics are designed with the goal of reaching the entire population.
+Crowd sourced samples were shown by @borgoCrowdsourcingInformationVisualization2017 to be biased towards more male, younger, and relatively well-educated individuals compared to the general adult U.S.
+population.
+Presumably, individuals with a higher education are more likely to have prior exposure to data visualization and might be more familiar with quantitative statements than less well-educated groups within the population; other pre-existing differences between population groups may also drive differences in understanding or interpreting charts.
+This raises the question of whether conclusions drawn from these studies hold for the general population -- 
+<!--generally, convenience and opt-in samples are not representative of the general population.-->
+ a common target audience for data visualization and scientific communication work.
+Results of polls, surveys, and scientific work are presented visually in the news, shared on social media, posted online or in print materials.
+It is imperative that data visualizations used to communicate quantitative information are designed with the goal of reaching the entire population.
 To that end, an inclusive, representative approach to studying how viewers perceive information presented in charts is critical to understand how best to communicate with the general public.
@@
 We employ a probability-based survey panel and run a series of perception tests with nationally-representative samples of respondents from that panel.
-The advantage of using a probability-based approach is two-fold.
-First, we have access to a large sample of survey participants and thus have greater power in making inference about graphical perceptional abilities.
-Second, the sample is representative of the general adult public in the U.S., and this allows us to test whether prior results from convenience samples hold with a nationally representative sample and whether there are differences in those results across demographic subgroups.
+The advantage of using a probability-based approach is two-fold:
+First, the relatively larger sample of survey participants provides greater statistical power in drawing inference about graphical perceptional abilities.
+Second, the sample is representative of the general adult public in the U.S., permitting an assessment inhowfar prior results from convenience samples hold with a nationally representative sample and whether there are differences in those results across demographic subgroups.
 Within that context we focus on the following research questions:
@@
   ggplot(aes(x = age4)) +
-  geom_errorbar(aes(ymin = percent2-moe_percent, ymax=percent2+moe_percent, colour="ACS estimate")) +
+  geom_errorbar(aes(ymin = percent2-moe_percent, ymax=percent2+moe_percent, colour=" ACS estimate")) +
 #  geom_point(aes(y = percent)) +
@@
   gender +
-  annotate("text", x = 0.83, y = 5, label="ACS estimate", hjust=0, 
+  annotate("text", x = 0.83, y = 5, label="ACS estimate & 95% C.I.", hjust=0, 
            size = 5, colour = "grey20") +
   annotate("segment", x = 0.5, xend=0.7, y = 5,yend=5, 
+           linewidth = 1, colour="grey20") +
+  annotate("segment", x = 0.5, xend=0.7, y = 5.1,yend=5.1, 
+           linewidth = 1, colour="grey20") +
+  annotate("segment", x = 0.5, xend=0.7, y = 4.9,yend=4.9, 
            linewidth = 1, colour="grey20") +
@@
 
-\autoref{fig-alignment}(a) displays all responses along the binary correctness measure, separated by whether the stimulus was an aligned task or unaligned task.
+\autoref{fig-alignment} displays all responses along the binary correctness measure, separated by whether the stimulus was an aligned task or unaligned task.
 We can see that levels of accuracy for all responses are significantly higher for the 'easier' (aligned) task, with about twice as many respondents correctly selecting the larger of the two marked elements.
@@
 ```{r accuracy,  warning=FALSE}
-#| fig-cap: "On the left (a), a stacked barchart shows the number of respondents with correct (green) and incorrect (grey) responses to the two comparison questions.
-When tiles are aligned along the same axis, more than twice the number of responses are correct.
-The shaded area along the top of the green bars corresponds to  95\\% confidence intervals around (marginal) correct responses.
-On the right (b), a parallel coordinate plot shows all combinations of responses.
+#| fig-cap: "Participants' response pattern to both questions.
+The coloring reflects the number of correct responses (two in full green, zero in grey, and one correct answer in grey-green).
+When tiles are aligned along the same axis, more than twice the number of respondents correctly identifies the larger tile.
 There's a huge asymmetry in the number of responses where participants answered only one of the questions correctly."
 #| label: fig-alignment
-#| out-width: 90%
-#| fig-width: 8.5
+#| out-width: 40%
+#| fig-width: 4
 
@@
 
-gg1 + gg2
+gg2
 ```
@@
 
+# Supplementary Material
+
+The source code and data used in this paper are available on a GitHub repository at [https://github.com/kiegan/testing-charts-jds](https://github.com/kiegan/testing-charts-jds).
+In addition, we provide an example stimulus image in that repository.
+
 <!-- # Supplementary Material {.unnumbered} -->

### 2023-10-27 SDSS-2023/testing-charts-accuracy.tex
modified tex file, rendered to pdf

@@
 \end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, boxrule=0pt, sharp corners, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 biased towards more male, younger, and relatively well-educated
-individuals relative to the adult U.S. population as a whole; generally,
-convenience and opt-in samples are not representative of the general
-population. Further, higher education individuals are likely to have
-prior exposure to data visualization or data topics in higher education,
-which may not be the case for other groups within the population;
-existing differences in exposure to data may drive differences in
-results across groups. The general public is a common target audience
-for data visualization and scientific communication work; results of
-polls, surveys, and scientific work are presented often in the news,
-shared on social media, or posted online or in print materials. It is
-imperative that data visualizations used to communicate these topics are
-designed with the goal of reaching the entire population. To that end,
-an inclusive, representative approach to studying how viewers perceive
-information presented in charts is critical to understand how best to
-communicate with the general public.
+individuals compared to the general adult U.S. population. Presumably,
+individuals with a higher education are more likely to have prior
+exposure to data visualization and might be more familiar with
+quantitative statements than less well-educated groups within the
+population; other pre-existing differences between population groups may
+also drive differences in understanding or interpreting charts. This
+raises the question of whether conclusions drawn from these studies hold
+for the general population -- a common target audience for data
+visualization and scientific communication work. Results of polls,
+surveys, and scientific work are presented visually in the news, shared
+on social media, posted online or in print materials. It is imperative
+that data visualizations used to communicate quantitative information
+are designed with the goal of reaching the entire population. To that
+end, an inclusive, representative approach to studying how viewers
+perceive information presented in charts is critical to understand how
+best to communicate with the general public.
 
@@
 with nationally-representative samples of respondents from that panel.
-The advantage of using a probability-based approach is two-fold. First,
-we have access to a large sample of survey participants and thus have
-greater power in making inference about graphical perceptional
+The advantage of using a probability-based approach is two-fold: First,
+the relatively larger sample of survey participants provides greater
+statistical power in drawing inference about graphical perceptional
 abilities. Second, the sample is representative of the general adult
-public in the U.S., and this allows us to test whether prior results
-from convenience samples hold with a nationally representative sample
-and whether there are differences in those results across demographic
+public in the U.S., permitting an assessment inhowfar prior results from
+convenience samples hold with a nationally representative sample and
+whether there are differences in those results across demographic
 subgroups. Within that context we focus on the following research
@@
 
-\autoref{fig-alignment}(a) displays all responses along the binary
+\autoref{fig-alignment} displays all responses along the binary
 correctness measure, separated by whether the stimulus was an aligned
@@
 
-{\centering \includegraphics[width=0.9\textwidth,height=\textheight]{./figures/fig-alignment-1.png}
+{\centering \includegraphics[width=0.4\textwidth,height=\textheight]{./figures/fig-alignment-1.png}
 
@@
 
-\caption{\label{fig-alignment}On the left (a), a stacked barchart shows
-the number of respondents with correct (green) and incorrect (grey)
-responses to the two comparison questions. When tiles are aligned along
-the same axis, more than twice the number of responses are correct. The
-shaded area along the top of the green bars corresponds to 95\%
-confidence intervals around (marginal) correct responses. On the right
-(b), a parallel coordinate plot shows all combinations of responses.
-There's a huge asymmetry in the number of responses where participants
-answered only one of the questions correctly.}
+\caption{\label{fig-alignment}Participants' response pattern to both
+questions. The coloring reflects the number of correct responses (two in
+full green, zero in grey, and one correct answer in grey-green). When
+tiles are aligned along the same axis, more than twice the number of
+respondents correctly identifies the larger tile. There's a huge
+asymmetry in the number of responses where participants answered only
+one of the questions correctly.}
 
@@
 
-\hypertarget{tbl-zoom-123}{}
+%\hypertarget{tbl-zoom-123}{}
 \begin{longtable}{lrrrr}
@@
 
+\hypertarget{supplementary-material}{%
+\section{Supplementary Material}\label{supplementary-material}}
+
+The source code and data used in this paper are available on a GitHub
+repository at \url{https://github.com/kiegan/testing-charts-jds}. In
+addition, we provide an example stimulus image in that repository.
+
 \hypertarget{references}{%

### 2023-12-04 revisions-round2/response-to-reviewers.qmd
prepping for revision 2

---
title: "Round 2: Response to the AE"
format: pdf
editor: source
---

Amanda,

thank you for your careful read-through and the thoughtful comments.

Please find belaow our point-by-point response,

all the best,

Point-by-point response:

Major comment:

I find the equations for the models inconsistent and confusing. Making these more consistent would greatly improve readability and understanding. Specific examples below:
+ Eq (3): Var hat and w bar are undefined.
+ Eq (5): the terms and the subscripts need to be defined; the parentheses in the subscript need to be explained (perhaps nested effects). Also, you say “let Y_{jk} describe the zooming behavior”, but only use Y_k in Eq (5). I also don’t understand what i, j, and l are referring to in this equation.
+ The indexing changes between Eq (4) and (6) (design is j in (4) and l in (6)).
+ k=1,...,N isn’t defined until Eq (6), even though it is used earlier. It isn’t clear what N is.

Minor comments:

The title is much improved, great change. I’d recommend removing the question from the title, as it seems repetitive with the first sentence of the paper.

Last paragraph of page 2 and first paragraph of page 3 (after numbered list) both start with “In this work”; one of these should be changed (could change the second to “To address these questions”).

Pg 3: “...we either split a survey sample in two and show each subsample a distinct structural version of the chart or test variations of a chart across distinct rounds of the survey.” Please add how much time passes between rounds.

Pg 4, 2nd paragraph: I think “aligned pieces” and “unaligned pieces” in the parentheses should be switched.

Figure 1 caption: “Please refer to the appendix for full size images of the stimuli.” -> “Please refer to the appendix for a full size image of stimuli (a).”

Pg 5: I really like the last sentence of the Task 1 section: “Any differences in observed responses can therefore be attributed to this difference in presentation.” A sentence like this after the other 2 tasks would be a helpful addition.

Pg 7, 1st sentence: “While there is some risk…” A sentence like this should appear earlier, like on page 3 in the study design section.

Pg 8, table 1 caption: Remove “as discussed above”

Pg 9, 1st paragraph: “the dark points” → “the black points and error bars”; “hashed lines” → “vertical lines”

Pg 9, fig 3: increase the axis text size

Pg 10, fig 4: I think the figure would be easier/more natural to read if you switch the ordering of the x axis (unaligned first). This better represents the order in which the questions are shown and would highlight the improvement of moving from unaligned to aligned. Also, this figure should be bigger. Finally, the sentence “There’s a huge asymmetry in the number of responses where participants answered only one of the questions correctly.” is very confusing to me, could you clarify this?

Pg 11: Thanks for adding all the detail about the letter comparisons in the figures and tables, very helpful. I’d start a new paragraph for this description, starting with “The letters allow us to assess…”. I’d also specify: “The letters in the figures and tables allow us to assess…”. Finally, the one sentence paragraph on this page should be combined with another paragraph.

Pg 11: “two estimates within the same column are significantly different (at the 5% level) if they do not have any letter in common”: this works for the table, but is confusing for the figures (and the reader is mostly looking at fig 5 when reading this description), be more specific here

Pg 12, Table 3 caption: “at 5%” → “at a 5% level of significance”

Pg 14, fig 7: Could you rephrase this, I find it hard to understand: “Incorrect responses on the unaligned task show a significant boost in certainty in the vertical stacked bar.” Also, in the next sentence “the difference…matters significantly” does not make sense.

Pg 15, 1st paragraph: “Our pattern of higher accuracy in selecting the correct response on the aligned task relative to the unaligned task holds across all demographic groups and structural variations.” To make this comparison, shouldn’t the aligned blue letters and the unaligned blue letters be different?

Pg 16, figure 8 caption: Some of the conclusions made seem to conflict with how I interpret the plot. Clarification is needed here. Specifically:
+ “For aligned tiles, gender, age, and education are not significant factors.” It looks like gender is significant? It has a different blue letter.
+ “When income levels increase, the percentage of wrong answers decreases, while the percentage of correct answers increases slightly (significant at 5%).” Looking at the pattern of the orange letters in the “by income, aligned” plot (a,ab,b,ab), it seems like the only significant difference is between the <$30k group and the $60k-$100k group, so it seems incorrect to say “the percentage of wrong answers decreases … (significant at 5%)”. Similarly, the blue letters in that plot are all a, so I don’t think it is correct to say “percentage of correct answers increases slightly (significant at 5%)”.
+ “higher levels of education and higher levels of income are associated with a significant increase of panelists choosing a response of ‘they are the same’, resulting in significant decreases of both correct answers and wrong answers.” Again, I don’t see evidence in the plot to say these differences are significant. Could you please clarify this? For example, in the “by income, unaligned” plot, the blue letters are all a, so there is no significant decrease of correct responses there.

Overall:
+ The paper suits better the Data Science in Action Section better than the Statistical Data Science Section, since this is a case study.
+ Number the sections as the JDS style does.

To submit your revision, please log in to EJMS and submit it as a revised file to the original submission. Please also include a point-by-point reply to each comment raised by the reviewers. In particular, most of the comments are relevant not just to the reviewers but also to general readers, so in each response, please state what were done to the manuscript.

Please submit your revision until 2023-12-22.

### 2023-12-04 SDSS-2023/testing-charts-accuracy.tex
moved tables to avoid hanging captions
prepping for revision 2

@@
 
-\title[Testing stacked bar charts in surveys]{What do they see? Testing perceptual accuracy in a U.S. general population survey using stacked bar charts}
+\title{What do they see? Testing perceptual accuracy in a U.S. general
+population survey using stacked bar charts}
+\author{}
 \date{}
@@
 \end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, frame hidden]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable]}{\end{tcolorbox}}\fi
 
@@
 
-%\hypertarget{tbl-zoom-123}{}
+\hypertarget{tbl-zoom-123}{}
 \begin{longtable}{lrrrr}
@@
 \midrule
-$\widehat{\mu}$ & $-0.35$ & $0.15$ & $-2.4$ & 0.0155 \\ 
+Intercept & $-0.35$ & $0.15$ & $-2.4$ & 0.0155 \\ 
 \midrule
@@
 \midrule
-They are the same $\widehat{\rho}_2$ & $-0.17$ & $0.09$ & $-1.9$ & 0.0578 \\ 
-Wrong $\widehat{\rho}_3$ & $-0.06$ & $0.13$ & $-0.5$ & 0.6466 \\ 
+response3They are the same & $-0.17$ & $0.09$ & $-1.9$ & 0.0578 \\ 
+response3Wrong & $-0.06$ & $0.13$ & $-0.5$ & 0.6466 \\ 
 \midrule
@@
 \midrule
-Aligned $\widehat{\tau}_2$ & $0.00$ & $0.06$ & $-0.1$ & 0.9435 \\ 
+taskcd & $0.00$ & $0.06$ & $-0.1$ & 0.9435 \\ 
 \midrule
@@
 \midrule
-Horizontal wide $\widehat{\delta}_2$ & $0.27$ & $0.15$ & $1.8$ & 0.0752 \\ 
-Vertical $\widehat{\rho}_3$ & $0.98$ & $0.13$ & $7.7$ & < 0.0001 \\ 
+chartHorizontal wide & $0.27$ & $0.15$ & $1.8$ & 0.0752 \\ 
+chartVertical & $0.98$ & $0.13$ & $7.7$ & < 0.0001 \\ 
 \bottomrule

### 2023-12-14 SDSS-2023/testing-charts-accuracy.qmd
prepping for revision 2
start of Round 2 Revisions work

@@
 ---
-title: "What do they see? Testing perceptual accuracy in a U.S. general population survey using stacked bar charts"
+title: "Testing perceptual accuracy in a U.S. general population survey using stacked bar charts"
 runtitle: Testing stacked bar charts in surveys # running title
@@
 
-In this work, we present a series of completed tests and the resulting findings.
+To address these questions, we present a series of completed tests and the resulting findings.
 The remainder of the paper is organized as follows: first, we describe the design of the visual stimuli used in the perception tests.
@@
 First, we vary the alignment of the pieces in question.
-Viewers are presented with two marked pieces in a chart that do not share a common baseline (aligned pieces), then two pieces that do share a common baseline (unaligned pieces).
+Viewers are presented with two marked pieces in a chart that do not share a common baseline (unaligned pieces), then two pieces that do share a common baseline (aligned pieces).
 Second, we vary the orientation of the chart -- we rotate the vertical stacked bar chart, and present a horizontally oriented version of the same chart, with identically sized marked pieces.
@@
 The larger piece in each chart are pieces B and D.
-Please refer to the appendix for full size images of the stimuli."
+Please refer to the appendix for afull size image  of stimulus (a)."
 #| out-width: 100%
@@
 
-: Survey rounds: dates, number of participants (nominal sample size), effective sample size, sum of weights, and factors for the combining of surveys as discussed above.
+: Survey rounds: dates, number of participants (nominal sample size), effective sample size, sum of weights, and factors for the combining of surveys.
 {#tbl-rounds}
@@
 
-Distributions of demographic characteristics across each task are shown in \autoref{fig-rounds-demographics}, the dark points show percentages for US adults based on the 5 year estimates of the American Community Survey 2021.
-The distribution of demographics are all quite similar across tasks and their 90% margin of errors (hashed lines) for the most part cover the Census estimates.
+Distributions of demographic characteristics across each task are shown in \autoref{fig-rounds-demographics}, the black points and error bars show percentages for US adults based on the 5 year estimates of the American Community Survey 2021.
+The distribution of demographics are all quite similar across tasks and their 90% margin of errors (vertical lines) for the most part cover the Census estimates.
 
@@
 Letters behind numbers indicate pairwise significances.
-Within the same column, values are significantly different (at 5%) if they do not share the same letter.
+Within the same column, values are significantly different (at a 5% level of significance) if they do not share the same letter.
 

### 2023-12-14 SDSS-2023/testing-charts-accuracy.qmd
updates to figures and figure captions per round 2 revision requests

@@
 Finally, to prevent viewers from being exposed to slight variations of the same stimulus in a row (and risk unforeseen order effects or respondents using prior questions to inform their responses), we either split a survey sample in two and show each subsample a distinct structural version of the chart or test variations of a chart across distinct rounds of the survey.
+Survey rounds are fielded once a month, so distinct rounds are separated by approximately one month.
 
@@
   xlab("") +
-  theme(legend.position ="none") +
+  theme(legend.position ="none", axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 12)) +
   scale_linetype_manual(values=c(2,2,1)) +
@@
   ggtitle("Age groups") +
-  theme(axis.text.x = element_text(angle=0, vjust = 0.5)) +
+  theme(axis.text.x = element_text(angle=0, vjust = 0.5, size = 12), axis.text.y = element_text(size = 12)) +
   xlab("") +
@@
 
-(gender + education + plot_layout(nrow=1, widths=c(2,5))) / 
+(gender + education + plot_spacer() + plot_layout(nrow=1, widths=c(2,5,1))) / 
   (age + income + plot_layout(nrow=1, widths = c(4,4)))
@@
 When tiles are aligned along the same axis, more than twice the number of respondents correctly identifies the larger tile.
-There's a huge asymmetry in the number of responses where participants answered only one of the questions correctly."
+The pattern in responses for participants who only answered one of the questions correctly is asymmetric; that is, a much higher percentage of participants answered the unaligned task incorrectly and the aligned task correctly than answered the unaligned task correctly and the aligned task incorrectly."
 #| label: fig-alignment
@@
     ) %>%
-  pcp_select(`Aligned: C or D?`, `Unaligned: A or B?`) %>%
+  pcp_select(`Unaligned: A or B?`, `Aligned: C or D?`) %>%
   pcp_scale() %>%
@@
 where $\ell \in \{1, 2\}$; $i \in \{1, 2\}$ is the comparison type (1 = Aligned, 2 = Unaligned), and $j \in \{1, 2, 3\}$ is the chart design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide.
-The letters allow us to assess whether differences between two means are significant:  two estimates within the same column are significantly different (at the 5% level) if they do not have any letter in common.
+
+The letters in the figures and tables allow us to assess whether differences between two means are significant:  two estimates within the same column are significantly different (at the 5% level) if they do not have any letter in common.
 This method is called connected letters and has been introduced by @piephoAlgorithmLetterBasedRepresentation2004.
@@
 For correct responses, all proportions are significantly different except for the proportion of correctly assessing the difference between unaligned rectangles in the wide horizontal design: their label of `ab` shows us that assessing the differences between unaligned rectangles in the wide horizontal design is not significantly better than in the horizontal design (labelled `a`), but also not significantly worse than in the vertical design (labelled `b`).
-
 The estimated values and 95% confidence intervals are shown in @tbl-rounds-123.
@@
 Correct answers in the unaligned task have the lowest certainty scores associated with them.
-Incorrect responses on the unaligned task show a significant boost in certainty in the vertical stacked bar.
-For horizontal stacked bars the difference between aligned and unaligned charts matters significantly.
+Incorrect responses to the unaligned task when viewing the vertical stacked bar are associated with a significantly higher certainty score.
 Error bars around the points show estimated scores (on a scale of 1 to 5) with corresponding 95% CI.
@@
 Let $Y_k$ be the response of participant $k$, on a scale from 1 = 'correct', 2 = 'they are the same' to 3 = 'wrong'.
-
 We use a generalized cumulative logistic regression, where $\mu_i$ are intercepts $1 \le i < 3$, $X_k$ are demographics of the $k$th participant (in form of the model matrix), and $\beta_{i}$ are the coefficients.
@@
 #| fig-cap: "Response patterns by alignment and demographic levels.
-For aligned tiles, gender, age, and education are not significant factors.
-However, income levels do have a (small) effect.
-When income levels increase, the percentage of wrong answers decreases, while the percentage of correct answers increases slightly (significant at 5%).
-For the  more difficult task of comparing unaligned tiles, higher levels of education and higher levels of income are associated with a significant increase of panelists choosing a response of 'they are the same', resulting in significant decreases of both correct answers and wrong answers.
+For aligned tiles, age and education are not significant factors.
+However, gender and income levels do have a (small) effect.
+For the  more difficult task of comparing unaligned tiles, higher levels of education and higher levels of income are associated with a significant increase of panelists choosing a response of 'they are the same', resulting in significant decreases of wrong answers for higher education attainment and higher income groups, and significant decreases in correct answers for higher levels of education.
 "

### 2023-12-14 SDSS-2023/testing-charts-accuracy.tex
updates to figures and figure captions per round 2 revision requests

@@
 \hypersetup{
-  pdftitle={What do they see? Testing perceptual accuracy in a U.S. general population survey using stacked bar charts},
+  pdftitle={Testing perceptual accuracy in a U.S. general population survey using stacked bar charts},
   pdfkeywords={graphical perception, data visualization, survey data},
@@
 
-\title{What do they see? Testing perceptual accuracy in a U.S. general
-population survey using stacked bar charts}
+\title{Testing perceptual accuracy in a U.S. general population survey
+using stacked bar charts}
 \author{}
@@
 \end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, breakable, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi
 
@@
 
-In this work, we present a series of completed tests and the resulting
-findings. The remainder of the paper is organized as follows: first, we
-describe the design of the visual stimuli used in the perception tests.
-We then describe the population of study respondents and obtained survey
-sample. Subsequently, we share a summary of the resulting survey
-responses across each of the tests, including analyses on accuracy of
-responses and response behavior. Finally, we discuss implications of
-this work and next steps.
+To address these questions, we present a series of completed tests and
+the resulting findings. The remainder of the paper is organized as
+follows: first, we describe the design of the visual stimuli used in the
+perception tests. We then describe the population of study respondents
+and obtained survey sample. Subsequently, we share a summary of the
+resulting survey responses across each of the tests, including analyses
+on accuracy of responses and response behavior. Finally, we discuss
+implications of this work and next steps.
 
@@
 subsample a distinct structural version of the chart or test variations
-of a chart across distinct rounds of the survey.
+of a chart across distinct rounds of the survey. Survey rounds are
+fielded once a month, so distinct rounds are separated by approximately
+one month.
 
@@
 two marked pieces in a chart that do not share a common baseline
-(aligned pieces), then two pieces that do share a common baseline
-(unaligned pieces). Second, we vary the orientation of the chart -- we
+(unaligned pieces), then two pieces that do share a common baseline
+(aligned pieces). Second, we vary the orientation of the chart -- we
 rotate the vertical stacked bar chart, and present a horizontally
@@
 larger piece in each chart are pieces B and D. Please refer to the
-appendix for full size images of the stimuli.}
+appendix for afull size image of stimulus (a).}
 
@@
 (nominal sample size), effective sample size, sum of weights, and
-factors for the combining of surveys as discussed above.
-}\tabularnewline
-\toprule\noalign{}
+factors for the combining of surveys. }\tabularnewline
+\toprule()
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \textbf{Combined} & --- & 2807 & 1520.8 & & \\
+\bottomrule()
 \end{longtable}
@@
 Task 3 & Horizontal wide & 462 & 249.1 \\
+\bottomrule()
 \end{longtable}
@@
 Distributions of demographic characteristics across each task are shown
-in \autoref{fig-rounds-demographics}, the dark points show percentages
-for US adults based on the 5 year estimates of the American Community
-Survey 2021. The distribution of demographics are all quite similar
-across tasks and their 90\% margin of errors (hashed lines) for the most
-part cover the Census estimates.
+in \autoref{fig-rounds-demographics}, the black points and error bars
+show percentages for US adults based on the 5 year estimates of the
+American Community Survey 2021. The distribution of demographics are all
+quite similar across tasks and their 90\% margin of errors (vertical
+lines) for the most part cover the Census estimates.
 
@@
 tiles are aligned along the same axis, more than twice the number of
-respondents correctly identifies the larger tile. There's a huge
-asymmetry in the number of responses where participants answered only
-one of the questions correctly.}
+respondents correctly identifies the larger tile. The pattern in
+responses for participants who only answered one of the questions
+correctly is asymmetric; that is, a much higher percentage of
+participants answered the unaligned task incorrectly and the aligned
+task correctly than answered the unaligned task correctly and the
+aligned task incorrectly.}
 
@@
 (1 = Aligned, 2 = Unaligned), and \(j \in \{1, 2, 3\}\) is the chart
-design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide. The
-letters allow us to assess whether differences between two means are
-significant: two estimates within the same column are significantly
-different (at the 5\% level) if they do not have any letter in common.
-This method is called connected letters and has been introduced by
-Piepho (2004). In Figure~\ref{fig-response-123}, we see, that the
-proportions of wrong responses are significantly different between
-tasks, but designs within each task are not significantly different. For
-correct responses, all proportions are significantly different except
-for the proportion of correctly assessing the difference between
-unaligned rectangles in the wide horizontal design: their label of
-\texttt{ab} shows us that assessing the differences between unaligned
-rectangles in the wide horizontal design is not significantly better
-than in the horizontal design (labelled \texttt{a}), but also not
-significantly worse than in the vertical design (labelled \texttt{b}).
-
-The estimated values and 95\% confidence intervals are shown in
-Table~\ref{tbl-rounds-123}.
+design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide.
+
+The letters in the figures and tables allow us to assess whether
+differences between two means are significant: two estimates within the
+same column are significantly different (at the 5\% level) if they do
+not have any letter in common. This method is called connected letters
+and has been introduced by Piepho (2004). In
+Figure~\ref{fig-response-123}, we see, that the proportions of wrong
+responses are significantly different between tasks, but designs within
+each task are not significantly different. For correct responses, all
+proportions are significantly different except for the proportion of
+correctly assessing the difference between unaligned rectangles in the
+wide horizontal design: their label of \texttt{ab} shows us that
+assessing the differences between unaligned rectangles in the wide
+horizontal design is not significantly better than in the horizontal
+design (labelled \texttt{a}), but also not significantly worse than in
+the vertical design (labelled \texttt{b}). The estimated values and 95\%
+confidence intervals are shown in Table~\ref{tbl-rounds-123}.
 
@@
 behind numbers indicate pairwise significances. Within the same column,
-values are significantly different (at 5\%) if they do not share the
-same letter. }\tabularnewline
+values are significantly different (at a 5\% level of significance) if
+they do not share the same letter. }\tabularnewline
 
@@
 correctness. Correct answers in the unaligned task have the lowest
-certainty scores associated with them. Incorrect responses on the
-unaligned task show a significant boost in certainty in the vertical
-stacked bar. For horizontal stacked bars the difference between aligned
-and unaligned charts matters significantly. Error bars around the points
+certainty scores associated with them. Incorrect responses to the
+unaligned task when viewing the vertical stacked bar are associated with
+a significantly higher certainty score. Error bars around the points
 show estimated scores (on a scale of 1 to 5) with corresponding 95\% CI.
@@
 Let \(Y_k\) be the response of participant \(k\), on a scale from 1 =
-`correct', 2 = `they are the same' to 3 = `wrong'.
-
-We use a generalized cumulative logistic regression, where \(\mu_i\) are
-intercepts \(1 \le i < 3\), \(X_k\) are demographics of the \(k\)th
-participant (in form of the model matrix), and \(\beta_{i}\) are the
-coefficients.
+`correct', 2 = `they are the same' to 3 = `wrong'. We use a generalized
+cumulative logistic regression, where \(\mu_i\) are intercepts
+\(1 \le i < 3\), \(X_k\) are demographics of the \(k\)th participant (in
+form of the model matrix), and \(\beta_{i}\) are the coefficients.
 
@@
 \end{equation}
+
+\begin{verbatim}
+Warning in RET$pfunction("adjusted", ...): Completion with error > abseps
+\end{verbatim}
 
@@
 \caption{\label{fig-demographics}Response patterns by alignment and
-demographic levels. For aligned tiles, gender, age, and education are
-not significant factors. However, income levels do have a (small)
-effect. When income levels increase, the percentage of wrong answers
-decreases, while the percentage of correct answers increases slightly
-(significant at 5\%). For the more difficult task of comparing unaligned
-tiles, higher levels of education and higher levels of income are
-associated with a significant increase of panelists choosing a response
-of `they are the same', resulting in significant decreases of both
-correct answers and wrong answers.}
+demographic levels. For aligned tiles, age and education are not
+significant factors. However, gender and income levels do have a (small)
+effect. For the more difficult task of comparing unaligned tiles, higher
+levels of education and higher levels of income are associated with a
+significant increase of panelists choosing a response of `they are the
+same', resulting in significant decreases of wrong answers for higher
+education attainment and higher income groups, and significant decreases
+in correct answers for higher levels of education.}
 

### 2023-12-14 SDSS-2023/testing-charts-accuracy.qmd
additional language changes in response to round 2 review comments

@@
 Viewers are presented with two marked pieces in a chart that do not share a common baseline (unaligned pieces), then two pieces that do share a common baseline (aligned pieces).
+We present the unaligned comparison first to reduce the likelihood that participants recognize that the tasks are identical excepting the alignment and carry over an answer from the aligned task to the unaligned task.
 Second, we vary the orientation of the chart -- we rotate the vertical stacked bar chart, and present a horizontally oriented version of the same chart, with identically sized marked pieces.
@@
 This represents a structural change in how the data are presented to the viewer while preserving the pixel size of the elements viewers must compare.
+Differences in participant accuracy across Tasks 1 and 2 can thus be assumed to be due to the orientation or layout of the chart rather than the comparison task itself.
 Tasks 2 and 3 were asked within the same survey round (Round 3) and the full sample was randomly split among respondents, with 50% of participants seeing the Task 2 stimuli and 50% of participants seeing the Task 3 stimuli.
@@
 This increases the difference in the length of bars to 13 pixels (previously 8 pixels).
+This increase in width allows us to test whether participant accuracy is higher when the elements in the comparison task have a larger visual difference in length relative to Tasks 1 and 2.
 The widths of the tiles are adjusted accordingly (from 50 to 30 pixels) to keep the overall area of the tiles approximately constant.
@@
 
-The letters in the figures and tables allow us to assess whether differences between two means are significant:  two estimates within the same column are significantly different (at the 5% level) if they do not have any letter in common.
+The letters in the figures and tables allow us to assess whether differences between two means are significant: in the tables, two estimates within the same column are significantly different (at the 5% level) if they do not have any letter in common.
+The same is true for letters along the top and bottom of the figures.
 This method is called connected letters and has been introduced by @piephoAlgorithmLetterBasedRepresentation2004.

### 2023-12-21 SDSS-2023/testing-charts-accuracy.qmd
quarto changes

@@
 \end{equation}
-where $CV$ is the coefficient of variation of the weights $w$ within each sample, and is estimated as in @kishSurveySampling1965 : 
+where $CV$ is the coefficient of variation of the weights $w$ within each survey sample $S_i$, and is estimated as in @kishSurveySampling1965 : 
 \begin{equation}
-CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
+CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2},
 \end{equation}
+and where $\bar{w}$ and $\widehat{Var(w)}$ are sample mean and sample variance of the weights $w$ of survey sample $S$, respectively.
 
 The data for this paper were collected in several rounds.
-The resulting number of participants, effective sample sizes, and $\lambda$ values are shown in \autoref{tbl-rounds}.
-
-| Name      | Date       |      \# Participants |                Effective Sample Size |         Sum of weights $\sum_i w_i$ |                                  $\lambda_i$ |
+The resulting number of panelists, effective sample sizes, and $\lambda$ values are shown in \autoref{tbl-rounds}.
+
+| Name      | Date       |      \# Panelists |                Effective Sample Size |         Sum of weights $\sum_i w_i$ |                                  $\lambda_i$ |
 |------------|------------|-----------:|-----------:|-----------:|-----------:|
@@
 
-: Survey rounds: dates, number of participants (nominal sample size), effective sample size, sum of weights, and factors for the combining of surveys.
+: Survey rounds: dates, number of panelists (nominal sample size), effective sample size, sum of weights, and factors for the combining of surveys.
 {#tbl-rounds}
@@
 
-A total of `r sum(c(nrow(round1_adj), nrow(round2_adj), nrow(round3_adj)))` respondents participated across the three rounds.
+A total of `r sum(c(nrow(round1_adj), nrow(round2_adj), nrow(round3_adj)))` panelists participated across the three rounds.
 The number of responses and corresponding effective sample sizes in each round are shown in \autoref{tbl-rounds}.
@@
   annotate("text", x = 0.83, y = 5, label="ACS estimate & 95% C.I.", hjust=0, 
-           size = 5, colour = "grey20") +
+           size = 4.5, colour = "grey20") +
   annotate("segment", x = 0.5, xend=0.7, y = 5,yend=5, 
@@
   annotate("text", x = 0.83, y = 15, label="Horizontal design", hjust=0,
-           size = 5, colour = "grey20") +
+           size = 4.5, colour = "grey20") +
   annotate("segment", x = 0.8333, xend = 0.8333, y = 17, yend=40, 
@@
   annotate("text", x = 1, y = 20, label="Horizontal wide", hjust=0,
-           size = 5, colour = "grey20") +
+           size = 4.5, colour = "grey20") +
   annotate("segment", x = 1, xend = 1, y = 22, yend=40, 
@@
   annotate("text", x = 1.17, y = 25, label="Vertical design", hjust=0,
-           size = 5, colour = "grey20") +
+           size = 4.5, colour = "grey20") +
   annotate("segment", x = 1.166, xend = 1.166, y = 27, yend=40, 
@@
 We first investigate respondent accuracy in selecting the correct response.
-As participants were able to select the option 'They are the same', there are several ways to model accuracy and response.
-The argument could be made that for the purposes of practical interpretation, 'they are the same' is a correct choice, as the options are visually very similar and not substantially different values within the context of the data shown in the chart.
+Participants were able to select the option 'They are the same', there are several ways to model accuracy and response.
+For the purposes of practical interpretation, the argument could be made that 'they are the same' is a correct choice, as the options are visually very similar and not substantially different values within the context of the data shown in the chart.
 However, we are interested in understanding whether viewers *can* perceive the difference and correctly identify which piece is larger and thus consider multiple ways of modeling response to investigate participant response patterns.
@@
   scale_colour_manual(values = c("grey", "#84AC84", "#70A570", "forestgreen")) +
-  ggtitle("(b) Combinations of responses") +
+  ggtitle("Combinations of responses") +
   theme(legend.position = "none")
@@
 
-@fig-response-123 shows the results of a cell-means model with ordinal response $Y_k$, where $Y_k$ is the $k$th participant's response, $Y_k \in \{1, 2, 3\}$, where 'correct' is encoded as 1, 'they are the same' is encoded as 2, and 'wrong' is encoded as 3:
+@fig-response-123 shows the results of a cell-means model with ordinal response $Y_{ik}$, where $Y_{ik}$ is the $k$th response, $1 \le k \le N$ for $N$ number of (nominal) responses. 
+We use an encoding of $Y_k \in \{1, 2, 3\}$ for 'correct'=1, 'they are the same'= 2, and 'wrong' = 3:
 
@@
 
-where $\ell \in \{1, 2\}$; $i \in \{1, 2\}$ is the comparison type (1 = Aligned, 2 = Unaligned), and $j \in \{1, 2, 3\}$ is the chart design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide.
+where $1 \le \ell < 3$; $i \in \{1, 2\}$ is the comparison type (1 = Aligned, 2 = Unaligned), and $j \in \{1, 2, 3\}$ is the chart design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide.
+Note that we are using the notation $ij\ell(k)$ to emphasize the functional relationship between the $k$th response and the model parameters while avoiding double-subscripts, i.e.
+$ij\ell(k)$ signifies the index combination of $i, j,$ and $\ell$ of the $k$th response, i.e.
+$ij\ell(k)$ is 123, if the $k$th response was a wrong response ($\ell$=3) for an aligned comparison ($i$=1) in a horizontal ($j$=2) barchart.
 
@@
 
-To formalize this pattern in a model, let $Y_{jk}$ describe the zooming behavior of panelist $k$ on task $j$.
+To formalize this pattern in a model, let $Y_k$ describe the zooming behavior on the $k$th response.
 We model zooming behavior (no = 0, yes = 1) as a logistic regression by correctness of response ($\rho$), task ($\tau$), and design ($\delta$) of the chart:
@@
 \end{equation}
-where $k = 1, ..., N$, $\mu_{ij\ell(k)}$ is average certainty (measured on a scale from 1 to 5) of the four combinations of task and correctness by each of the three designs, where $i = 1, 2$ encodes unaligned/aligned, and $j=1,2$ encodes wrong, correct, and $\ell = 1, 2, 3$ encodes horizontal, wide horizontal, and vertical stacked bar charts, respectively.
-We also assume that errors are normally distributed, i.e. $\epsilon_k \stackrel{i.i.d}{\sim} N(0, \sigma^2)$ for all $k = 1, ..., N$.
+where $1 \le k \le N$ for $N$ responses, $\mu_{ij\ell(k)}$ is average certainty (measured on a scale from 1 to 5) of the four combinations of task and correctness by each of the three designs, for task $i = 1, 2$ (unaligned/aligned), design $j=1, 2, 3$ encoded as horizontal, wide horizontal, and vertical stacked bar charts, and  participant's decision $\ell = 1, 2$, encoded as correct or wrong, respectively.
+We also assume that errors are normally distributed, i.e. $\epsilon_k \stackrel{i.i.d}{\sim} N(0, \sigma^2)$ for all $1 \le k \le N$.
 The results are shown in @fig-certainty.
@@
 
-Let $Y_k$ be the response of participant $k$, on a scale from 1 = 'correct', 2 = 'they are the same' to 3 = 'wrong'.
-We use a generalized cumulative logistic regression, where $\mu_i$ are intercepts $1 \le i < 3$, $X_k$ are demographics of the $k$th participant (in form of the model matrix), and $\beta_{i}$ are the coefficients.
+Let $Y_k$ be the response of participant $k$, on a scale from 1 = 'wrong', 2 = 'they are the same' to 3 = 'correct'.
+We use a generalized cumulative logistic regression, where $\mu_\ell$ are intercepts $1 \le \ell < 3$, $X_k$ are demographics of the $k$th participant (the $k$th row of the corresponding model matrix), and $\beta_{\ell}$ are the coefficients.
 
 \begin{equation}
-\text{ logit } P(Y_k \le i \mid X_k) = \mu_i + X_{k}'\beta_{i}
+\text{ logit } P(Y_k \le \ell \mid X_k) = \mu_\ell + X_{k}\beta_{\ell}
 \end{equation}

### 2023-12-21 SDSS-2023/testing-charts-accuracy.tex
corrected tex file, pdf and images

@@
 
-\title{Testing perceptual accuracy in a U.S. general population survey
+\title[Testing stacked bar charts in surveys]{Testing perceptual accuracy in a U.S. general population survey
 using stacked bar charts}
-\author{}
 \date{}
@@
 \end{abstract}
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, breakable, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, sharp corners, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi
 
@@
 (unaligned pieces), then two pieces that do share a common baseline
-(aligned pieces). Second, we vary the orientation of the chart -- we
-rotate the vertical stacked bar chart, and present a horizontally
+(aligned pieces). We present the unaligned comparison first to reduce
+the likelihood that participants recognize that the tasks are identical
+excepting the alignment and carry over an answer from the aligned task
+to the unaligned task. Second, we vary the orientation of the chart --
+we rotate the vertical stacked bar chart, and present a horizontally
 oriented version of the same chart, with identically sized marked
@@
 are presented to the viewer while preserving the pixel size of the
-elements viewers must compare. Tasks 2 and 3 were asked within the same
-survey round (Round 3) and the full sample was randomly split among
-respondents, with 50\% of participants seeing the Task 2 stimuli and
-50\% of participants seeing the Task 3 stimuli.
+elements viewers must compare. Differences in participant accuracy
+across Tasks 1 and 2 can thus be assumed to be due to the orientation or
+layout of the chart rather than the comparison task itself. Tasks 2 and
+3 were asked within the same survey round (Round 3) and the full sample
+was randomly split among respondents, with 50\% of participants seeing
+the Task 2 stimuli and 50\% of participants seeing the Task 3 stimuli.
 
@@
 increased to fit the new dimensions. This increases the difference in
-the length of bars to 13 pixels (previously 8 pixels). The widths of the
-tiles are adjusted accordingly (from 50 to 30 pixels) to keep the
-overall area of the tiles approximately constant.
+the length of bars to 13 pixels (previously 8 pixels). This increase in
+width allows us to test whether participant accuracy is higher when the
+elements in the comparison task have a larger visual difference in
+length relative to Tasks 1 and 2. The widths of the tiles are adjusted
+accordingly (from 50 to 30 pixels) to keep the overall area of the tiles
+approximately constant.
 
@@
 \end{equation} where \(CV\) is the coefficient of variation of the
-weights \(w\) within each sample, and is estimated as in Kish (1965) :
-\begin{equation}
-CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2}.
-\end{equation}
+weights \(w\) within each survey sample \(S_i\), and is estimated as in
+Kish (1965) : \begin{equation}
+CV(w \in S) = \frac{\widehat{Var(w)}}{\bar{w}^2},
+\end{equation} and where \(\bar{w}\) and \(\widehat{Var(w)}\) are sample
+mean and sample variance of the weights \(w\) of survey sample \(S\),
+respectively.
 
 The data for this paper were collected in several rounds. The resulting
-number of participants, effective sample sizes, and \(\lambda\) values
-are shown in \autoref{tbl-rounds}.
+number of panelists, effective sample sizes, and \(\lambda\) values are
+shown in \autoref{tbl-rounds}.
 
@@
   >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}@{}}
-\caption{\label{tbl-rounds}Survey rounds: dates, number of participants
+\caption{\label{tbl-rounds}Survey rounds: dates, number of panelists
 (nominal sample size), effective sample size, sum of weights, and
 factors for the combining of surveys. }\tabularnewline
-\toprule()
+\toprule\noalign{}
 \begin{minipage}[b]{\linewidth}\raggedright
@@
 \end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
-\# Participants
+\# Panelists
 \end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
@@
 \textbf{Combined} & --- & 2807 & 1520.8 & & \\
-\bottomrule()
 \end{longtable}
@@
 
-A total of 2807 respondents participated across the three rounds. The
+A total of 2807 panelists participated across the three rounds. The
 number of responses and corresponding effective sample sizes in each
@@
 We first investigate respondent accuracy in selecting the correct
-response. As participants were able to select the option `They are the
-same', there are several ways to model accuracy and response. The
-argument could be made that for the purposes of practical
-interpretation, `they are the same' is a correct choice, as the options
-are visually very similar and not substantially different values within
-the context of the data shown in the chart. However, we are interested
-in understanding whether viewers \emph{can} perceive the difference and
+response. Participants were able to select the option `They are the
+same', there are several ways to model accuracy and response. For the
+purposes of practical interpretation, the argument could be made that
+`they are the same' is a correct choice, as the options are visually
+very similar and not substantially different values within the context
+of the data shown in the chart. However, we are interested in
+understanding whether viewers \emph{can} perceive the difference and
 correctly identify which piece is larger and thus consider multiple ways
@@
 Figure~\ref{fig-response-123} shows the results of a cell-means model
-with ordinal response \(Y_k\), where \(Y_k\) is the \(k\)th
-participant's response, \(Y_k \in \{1, 2, 3\}\), where `correct' is
-encoded as 1, `they are the same' is encoded as 2, and `wrong' is
-encoded as 3:
+with ordinal response \(Y_{ik}\), where \(Y_{ik}\) is the \(k\)th
+response, \(1 \le k \le N\) for \(N\) number of (nominal) responses. We
+use an encoding of \(Y_k \in \{1, 2, 3\}\) for `correct'=1, `they are
+the same'= 2, and `wrong' = 3:
 
@@
 
-where \(\ell \in \{1, 2\}\); \(i \in \{1, 2\}\) is the comparison type
-(1 = Aligned, 2 = Unaligned), and \(j \in \{1, 2, 3\}\) is the chart
-design, with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide.
+where \(1 \le \ell < 3\); \(i \in \{1, 2\}\) is the comparison type (1 =
+Aligned, 2 = Unaligned), and \(j \in \{1, 2, 3\}\) is the chart design,
+with 1 = Vertical, 2 = Horizontal, and 3 = Horizontal wide. Note that we
+are using the notation \(ij\ell(k)\) to emphasize the functional
+relationship between the \(k\)th response and the model parameters while
+avoiding double-subscripts, i.e.~\(ij\ell(k)\) signifies the index
+combination of \(i, j,\) and \(\ell\) of the \(k\)th response,
+i.e.~\(ij\ell(k)\) is 123, if the \(k\)th response was a wrong response
+(\(\ell\)=3) for an aligned comparison (\(i\)=1) in a horizontal
+(\(j\)=2) barchart.
 
 The letters in the figures and tables allow us to assess whether
-differences between two means are significant: two estimates within the
-same column are significantly different (at the 5\% level) if they do
-not have any letter in common. This method is called connected letters
-and has been introduced by Piepho (2004). In
+differences between two means are significant: in the tables, two
+estimates within the same column are significantly different (at the 5\%
+level) if they do not have any letter in common. The same is true for
+letters along the top and bottom of the figures. This method is called
+connected letters and has been introduced by Piepho (2004). In
 Figure~\ref{fig-response-123}, we see, that the proportions of wrong
@@
 \midrule
-Intercept & $-0.35$ & $0.15$ & $-2.4$ & 0.0155 \\ 
+$\widehat{\mu}$ & $-0.35$ & $0.15$ & $-2.4$ & 0.0155 \\ 
 \midrule
@@
 \midrule
-response3They are the same & $-0.17$ & $0.09$ & $-1.9$ & 0.0578 \\ 
-response3Wrong & $-0.06$ & $0.13$ & $-0.5$ & 0.6466 \\ 
+They are the same $\widehat{\rho}_2$ & $-0.17$ & $0.09$ & $-1.9$ & 0.0578 \\ 
+Wrong $\widehat{\rho}_3$ & $-0.06$ & $0.13$ & $-0.5$ & 0.6466 \\ 
 \midrule
@@
 \midrule
-taskcd & $0.00$ & $0.06$ & $-0.1$ & 0.9435 \\ 
+Aligned $\widehat{\tau}_2$ & $0.00$ & $0.06$ & $-0.1$ & 0.9435 \\ 
 \midrule
@@
 \midrule
-chartHorizontal wide & $0.27$ & $0.15$ & $1.8$ & 0.0752 \\ 
-chartVertical & $0.98$ & $0.13$ & $7.7$ & < 0.0001 \\ 
+Horizontal wide $\widehat{\delta}_2$ & $0.27$ & $0.15$ & $1.8$ & 0.0752 \\ 
+Vertical $\widehat{\delta}_3$ & $0.98$ & $0.13$ & $7.7$ & < 0.0001 \\ 
 \bottomrule
@@
 
-To formalize this pattern in a model, let \(Y_{jk}\) describe the
-zooming behavior of panelist \(k\) on task \(j\). We model zooming
-behavior (no = 0, yes = 1) as a logistic regression by correctness of
-response (\(\rho\)), task (\(\tau\)), and design (\(\delta\)) of the
-chart:
+To formalize this pattern in a model, let \(Y_k\) describe the zooming
+behavior on the \(k\)th response. We model zooming behavior (no = 0, yes
+= 1) as a logistic regression by correctness of response (\(\rho\)),
+task (\(\tau\)), and design (\(\delta\)) of the chart:
 
 \begin{equation}
-\text{logit } P(Y_k \le 1) = \mu + \rho_{i(k)} + \tau_{j(k)} + \delta_{\ell(k)},
+\text{logit } P(Y_k \le 1) = \mu + \rho_{\ell(k)} + \tau_{i(k)} + \delta_{j(k)},
 \end{equation}
@@
 Y_{k} = \mu_{ij\ell(k)} + \epsilon_{k},
-\end{equation} where \(k = 1, ..., N\), \(\mu_{ij\ell(k)}\) is average
-certainty (measured on a scale from 1 to 5) of the four combinations of
-task and correctness by each of the three designs, where \(i = 1, 2\)
-encodes unaligned/aligned, and \(j=1,2\) encodes wrong, correct, and
-\(\ell = 1, 2, 3\) encodes horizontal, wide horizontal, and vertical
-stacked bar charts, respectively. We also assume that errors are
-normally distributed,
+\end{equation} where \(1 \le k \le N\) for \(N\) responses,
+\(\mu_{ij\ell(k)}\) is average certainty (measured on a scale from 1 to
+5) of the four combinations of task and correctness by each of the three
+designs, for task \(i = 1, 2\) (unaligned/aligned), design \(j=1, 2, 3\)
+encoded as horizontal, wide horizontal, and vertical stacked bar charts,
+and participant's decision \(\ell = 1, 2\), encoded as correct or wrong,
+respectively. We also assume that errors are normally distributed,
 i.e.~\(\epsilon_k \stackrel{i.i.d}{\sim} N(0, \sigma^2)\) for all
-\(k = 1, ..., N\). The results are shown in Figure~\ref{fig-certainty}.
+\(1 \le k \le N\). The results are shown in Figure~\ref{fig-certainty}.
 We find that the highest scores for certainty are associated with the
@@
 Let \(Y_k\) be the response of participant \(k\), on a scale from 1 =
-`correct', 2 = `they are the same' to 3 = `wrong'. We use a generalized
-cumulative logistic regression, where \(\mu_i\) are intercepts
-\(1 \le i < 3\), \(X_k\) are demographics of the \(k\)th participant (in
-form of the model matrix), and \(\beta_{i}\) are the coefficients.
+`wrong', 2 = `they are the same' to 3 = `correct'. We use a generalized
+cumulative logistic regression, where \(\mu_\ell\) are intercepts
+\(1 \le \ell < 3\), \(X_k\) are demographics of the \(k\)th participant
+(the \(k\)th row of the corresponding model matrix), and
+\(\beta_{\ell}\) are the coefficients.
 
 \begin{equation}
-\text{ logit } P(Y_k \le i \mid X_k) = \mu_i + X_{k}'\beta_{i}
+\text{ logit } P(Y_k \le \ell \mid X_k) = \mu_\ell + X_{k}\beta_{\ell}
 \end{equation}
-
-\begin{verbatim}
-Warning in RET$pfunction("adjusted", ...): Completion with error > abseps
-\end{verbatim}
 

### 2023-12-21 SDSS-2023/testing-charts-accuracy.tex
adding section numbers

@@
 \hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}}
 
@@
 
+\hypertarget{study-design}{%
+\section{Study design}\label{study-design}}
+
 \hypertarget{study-design-stimulus}{%
-\subsection{Study design -- stimulus}\label{study-design-stimulus}}
+\subsection{Stimulus}\label{study-design-stimulus}}
 
@@
 \hypertarget{question-text}{%
-\paragraph{Question text}\label{question-text}}
+\subsubsection{Question text}\label{question-text}}
 
@@
 \hypertarget{alignment}{%
-\paragraph{Alignment}\label{alignment}}
+\subsubsection{Alignment}\label{alignment}}
 
@@
 \hypertarget{study-design-participants}{%
-\subsection{Study design --
-participants}\label{study-design-participants}}
+\subsection{Participants}\label{study-design-participants}}
 
@@
 \hypertarget{study-design-survey-weighting}{%
-\subsection{Study design -- survey
+\subsection{Survey
 weighting}\label{study-design-survey-weighting}}
@@
 \hypertarget{respondents}{%
-\subsubsection{Respondents}\label{respondents}}
+\subsection{Respondents}\label{respondents}}
 
@@
 \hypertarget{accuracy-of-responses}{%
-\subsubsection{Accuracy of responses}\label{accuracy-of-responses}}
+\subsection{Accuracy of responses}\label{accuracy-of-responses}}
 
@@
 \hypertarget{respondent-behavior}{%
-\subsubsection{Respondent behavior}\label{respondent-behavior}}
+\subsection{Respondent behavior}\label{respondent-behavior}}
 
@@
 \hypertarget{differences-across-demographic-groups}{%
-\subsubsection{Differences across demographic
+\subsection{Differences across demographic
 groups}\label{differences-across-demographic-groups}}
@@
 
+\hypertarget{conclusions}{%
+\section{Conclusions}\label{conclusions}}
+
+Through this work, we have tested the use of probability-based survey
+panels to ask perception questions; within a large survey covering a
+variety of topics, and with a limited number of questions for our
+specific tasks, we are able to measure viewer perception and produce
+results consistent with prior studies. Testing data visualization design
+structures and viewer behavior on a nationally-representative sample is
+of broad interest and applicability to the scientific communication
+community, and this study demonstrates a framework under which to
+complete further work in this area.
+
+With our survey framework, we have shown that we can rank structural
+variations on design and tasks by their relative accuracy; we identified
+significantly different rates of respondent accuracy under different
+designs presenting the same data. Our study sample is larger than prior
+studies with a similar scope. That large sample size improves our power
+to identify this signal.
+
 \blandscape
@@
 \elandscape
-
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
-
-Through this work, we have tested the use of probability-based survey
-panels to ask perception questions; within a large survey covering a
-variety of topics, and with a limited number of questions for our
-specific tasks, we are able to measure viewer perception and produce
-results consistent with prior studies. Testing data visualization design
-structures and viewer behavior on a nationally-representative sample is
-of broad interest and applicability to the scientific communication
-community, and this study demonstrates a framework under which to
-complete further work in this area.
-
-With our survey framework, we have shown that we can rank structural
-variations on design and tasks by their relative accuracy; we identified
-significantly different rates of respondent accuracy under different
-designs presenting the same data. Our study sample is larger than prior
-studies with a similar scope. That large sample size improves our power
-to identify this signal.
 

### 2023-12-21 SDSS-2023/testing-charts-accuracy.tex
updating supplementary materials header for page limit

@@
 \hypertarget{supplementary-material}{%
-\section{Supplementary Material}\label{supplementary-material}}
-
-The source code and data used in this paper are available on a GitHub
-repository at \url{https://github.com/kiegan/testing-charts-jds}. In
-addition, we provide an example stimulus image in that repository.
+\section*{Supplementary Material}\label{supplementary-material}}
+\vspace{-4mm}
+
+The source code and data used in this paper and an example stimulus image are available on a GitHub
+repository at \url{https://github.com/kiegan/testing-charts-jds}.
 

### 2024-02-06 SDSS-2023/testing-charts-accuracy.tex
final round?

@@
 larger piece in each chart are pieces B and D. Please refer to the
-appendix for afull size image of stimulus (a).}
+appendix for a full size image of stimulus (a).}
 
@@
 pixels apart, which leads to a JND of 3.5 pixels based on Lu et al.
-(2022)`s model. The heights of the bars are 205 (left) and 213 pixels
+(2022)'s model. The heights of the bars are 205 (left) and 213 pixels
 (right), corresponding to about twice the JND. This difference should
@@
 demographics in each of the tasks: the colored dots show marginal
-percentages for demographics across the three tasks. In comparison, the
+percentages (with 90\% margins of error as vertical lines in the same color) for demographics across the three tasks.  
+In comparison, the
 dark lines show upper and lower 95\% confidence of these demographics as
@@
 We first investigate respondent accuracy in selecting the correct
-response. Participants were able to select the option `They are the
-same', there are several ways to model accuracy and response. For the
+response. Remember, that participants were able to select the option `They are the
+same' as response. This leads us to several ways to model accuracy and response. For the
 purposes of practical interpretation, the argument could be made that
@@
 Figure~\ref{fig-response-123} shows the results of a cell-means model
-with ordinal response \(Y_{ik}\), where \(Y_{ik}\) is the \(k\)th
+with ordinal response \(Y_{k}\), where \(Y_{k}\) is the \(k\)th
 response, \(1 \le k \le N\) for \(N\) number of (nominal) responses. We
-use an encoding of \(Y_k \in \{1, 2, 3\}\) for `correct'=1, `they are
-the same'= 2, and `wrong' = 3:
+use an encoding of \(Y_k \in \{1, 2, 3\}\) for `correct' ($\ell$=1), `they are
+the same' ($\ell$ = 2), and `wrong' ($\ell$ = 3):
 
@@
 
-where \(1 \le \ell < 3\); \(i \in \{1, 2\}\) is the comparison type (1 =
+where \(i \in \{1, 2\}\) is the comparison type (1 =
 Aligned, 2 = Unaligned), and \(j \in \{1, 2, 3\}\) is the chart design,
@@
 \toprule
- & \multicolumn{3}{c}{correct | same or wrong} & \multicolumn{3}{c}{correct or same | wrong} \\ 
+ & \multicolumn{3}{c}{correct vs. same or wrong} & \multicolumn{3}{c}{correct or same vs. wrong} \\ 
 \cmidrule(lr){2-4} \cmidrule(lr){5-7}
@@
 
-Rates of selecting the incorrect response are low across all three
+Rates of selecting the response `wrong' are low across all three
 structural variations and the aligned and unaligned tasks; while viewers
-select the incorrect response at significantly higher rates for all
+select the  response `wrong' at significantly higher rates for all
 three unaligned tasks relative to the aligned tasks, those rates do not
@@
 significant. However, different designs lead to different rates of
-zooming: we observe in \autoref{fig-zoom-123} that when a significantly
+zooming: we observe in \autoref{fig-zoom-123} that a significantly
 lower rate of respondents chose to make use of zooming when answering
@@
 for the horizontal wide orientation.
-
+%
 Using linear scores for the response of certainty, with `not certain at
@@
 (the \(k\)th row of the corresponding model matrix), and
-\(\beta_{\ell}\) are the coefficients.
+\(\beta_{\ell}\) is a vector of coefficients:
 
@@
 
+Note, that $\beta_\ell$ encompasses all estimable parameters of the demographic variables gender, age category, education category and income category for a total of eleven parameters.
 The resulting model estimates and confidence intervals are shown in
-\autoref{tbl-demographics}. When considering the easier (aligned) task,
+\autoref{tbl-demographics}.
+
+When considering the easier (aligned) task,
 response patterns do not differ significantly by age, gender, or
@@
 \cmidrule(lr){2-7} \cmidrule(lr){8-13}
- & \multicolumn{3}{c}{correct | same or wrong} & \multicolumn{3}{c}{correct or same | wrong} & \multicolumn{3}{c}{correct | same or wrong } & \multicolumn{3}{c}{correct or same | wrong } \\ 
+ & \multicolumn{3}{c}{correct vs. same or wrong} & \multicolumn{3}{c}{correct or same vs. wrong} & \multicolumn{3}{c}{correct vs. same or wrong } & \multicolumn{3}{c}{correct or same vs. wrong } \\ 
 \cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10} \cmidrule(lr){11-13}
